### 2023-01-12 docs/causal_roadmap.Rmd
Initial commit

---
title: "Causal Roadmap"
author: Lorenzo Fabbri
output:
  bookdown::pdf_document2:
    number_sections: false
  html_document: default
bibliography: references.bib
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
```

```{r, echo=FALSE}
# Type of adjustment set from DAG
type.mas <- "minimal"
# Type of effect to be estimated from DAG
type.effect <- "direct"
```

\textcolor{red}{Most of the following was **copy-pasted** from published papers. Some of the Sections must be filled depending on the chosen causal framework.}

<!-- ####################################################################### -->
# Step 0: Formulate the research question(s)
<!-- ####################################################################### -->

The **aim** of the present study was to research the...

The study **population** was based on..., consisting of $\text{N}=$... Inclusion criteria were:...

Measured **exposures** consisted of...

The **primary outcome** was... **Secondary outcomes** included...

<!-- ####################################################################### -->
## Descriptive analyses
<!-- ####################################################################### -->

<!-- ####################################################################### -->
# Step 1: Define a realistic statistical model
<!-- ####################################################################### -->

For time index $t$, let $W_t$ denote the set of potential confounders, $A_t$ the observed exposures, 
and $Y_t$ the clinical outcome of interest. For each subject $i$ and each time index $t$, 
we assumed its observed data $O_{ti} = (W_{ti}, A_{ti}, Y_{ti})$ were generated by sampling from a distribution 
$\mathbb{P}_{0,t}$ compatible with the causal model specified in \@ref(eq:npsem).

<!-- ####################################################################### -->
# Step 2: Specify a causal model and causal quantity of interest
<!-- ####################################################################### -->

We specified the following nonparametric structural equation model (NPSEM) to present the data generating process, 
including its measured confounders $W_t$, exposures $A_t$, and outcome $Y_t$, for each time index $t$:

\begin{equation}
\begin{aligned}
  W_t &= f_{W{_t}} (U_{W_{t}}) \\
  A_t &= f_{A_{t}} (W_t, U_{A_{t}}) \\
  Y_t &= f_{Y_{t}} (W_t, A_t, U_{Y_{t}}), 
\end{aligned}
(\#eq:npsem)
\end{equation}

where $(f_{W{_t}}, f_{A_{t}}, f_{Y_{t}})$ were the nonparametric structural equations, 
and $(U_{W_{t}}, U_{A_{t}}, U_{Y_{t}})$ were the unmeasured factors contributing the confounders, exposures, and outcome, respectively. 
The corresponding causal graph (without indexes for ease of notation) is given in Figure \@ref(fig:causal-graph).

```{r causal-graph, echo=FALSE, fig.cap="Causal graph corresponding to structural equations.", fig.height=3, fig.width=3}
causal.graph <- ggdag::dagify(A ~ W + U_A, Y ~ W + A + U_Y, W ~ U_W, 
                              exposure = "A", outcome = "Y") %>%
  ggdag::tidy_dagitty()
ggdag::ggdag(causal.graph) + ggdag::theme_dag()
```

For each time index, we specified our causal parameter $\phi$ as...

As **primary analysis**, we estimated....

As **secondary analyses**, we estimated...

In the following sections, we present the simplified DAG specific to each research question.

## Research question 1:...

```{r echo=FALSE, fig.cap="Simplified DAG for research question 1."}
sdag1 <- ggdag::dagify(X_ti ~ W_t, 
                       Y ~ W_t + X_ti, 
                       exposure = "X_ti", outcome = "Y") %>%
  ggdag::tidy_dagitty() %>%
  ggdag::ggdag(text_size = 2, node_size = 10) + ggdag::theme_dag() +
  ggplot2::ggtitle("RQ1")
```

```{r, echo=FALSE}
ggpubr::ggarrange(sdag1, ..., 
                  ncol = x, nrow = y)
```

<!-- ####################################################################### -->
# Step 3: Identification and the statistical estimand
<!-- ####################################################################### -->

For the causal parameter $\phi$, which involves a summary measure of the distribution of counterfactuals, 
to be identified in terms of the observed data distribution, several assumptions would be required:

* **No unmeasured confounding**: There are no unmeasured common causes of the exposure and the subsequent outcome. In our context, this assumption would be violated if, for example, an unmeasured variable influenced both the exposure levels and the outcome. We cannot guarantee that this assumption holds, and therefore limit our interpretations to statistical associations rather than causal effects.
* **Positivity**:....
* **Independence** of subjects. This assumption also implies **no interference**: the exposure level $a$ for a given subject does not affect the outcomes of the other subjects.
* **Consistency**: If $A = a$ for any subject, then $Y(a) = Y$. This means that the counterfactual outcome for a subject with its observed exposure level is the observed outcome.
* **Time-ordering**: The confounders $W$ precede the exposure $A$, which also precedes the outcome $Y$.

If these identifiability assumptions held, we could specify and focus our estimation efforts on a statistical estimand that equals the wished-for causal effect. In the (likely) case that they are not satisfied, we could still specify and focus our estimation efforts on a statistical estimand that is as close as possible to the causal parameter. Factoring the joint distribution of the observed data $\mathbb{P}_0$ into $\mathbb{P}_0 (O) = \mathbb{P}_0(Y|A,W) \mathbb{P}_0(A|W) \mathbb{P}_0(W)$, it can be shown that the statistical estimand corresponding to expected counterfactual outcome..., is given by

$$
..., 
$$

Therefore, our statistical estimand of interest was

$$
....
$$

<!-- ####################################################################### -->
# Step 4: Estimation from data and statistical inference
<!-- ####################################################################### -->

<!-- ####################################################################### -->
# Step 5: Interpretation and sensitivity analyses to inform a substantive conclusion
<!-- ####################################################################### -->

Highlight possible **limitations**:

* When considering only one exposure, we do not consider the effect of the others on that exposure and the outcome;
* When we consider only one outcome, we do not consider the effect of the other outcomes on that one;
* Although the interpretation of $\psi^{\text{causal}}$ is clear, how closely $\psi^{\text{stat}}$ matches it merits discussion. One considers the plausibility of each of the identifying assumptions in turn.

We performed **sensitivity analysis** to measure robustness to unmeasured confounders, based on the following considerations:

* Broadly, an open question is whether the estimated treatment effect is biased due to confounding by unmeasured covariates. We can examine how the substantive conclusion would be impacted under a range of presumed causal bias, $\delta = \psi^{\text{causal}} - \psi^{\text{stat}}$. That is, how *strong* would a particular confounder (or group of confounders) have to be in order to change the conclusions of this study? In a worst case scenario, how vulnerable are the study's results to many or all unobserved confounders acting **together**, possibly non-linearly? The exercise illustrates how the effect estimates, and confidence interval bounds change, depending on the magnitude and direction of the hypothesized gap;
* We implemented the following methods to conduct sensitivity analysis:
  * Selection bias and stratification...
  * Placebo treatment (replacement of the actual treatment with a random variable);
  * We performed leave-1-out analyses to assess the influence of single subjects on the effect of interest.

We took into consideration the following points to *validate* our results:

* We evaluated whether the obtained results are of any relevance from a public health point of view.

<!-- ####################################################################### -->
# Appendix
<!-- ####################################################################### -->

## Checklist A: Reproducibility

Table with random seed, names, description, and version numbers of all software packages.

| **Name** | **Version** | **Description** |
|:---------|:------------|:----------------|
| Random seed | NA | Will be set to $X$ |
| `R` | `r I(paste0(R.version$major, ".", R.version$minor))` | Statistical programming environment |
| ... | ... | ... |

## Checklist B1: `xxx` package specifications

Table providing values for all non-data arguments, and brief rationale when departing from the default specification.

| **Argument** | **Setting** | **Default** (Y/N) | **Comment** |
|:-------------|:------------|:------------------|:------------|
| ... | ... | ... | ... |

## DAGs

```{r, echo=FALSE}
source("../DAGs/dag_v1.R")
source("../code/dags.R")

dags <- dag.v1()
```

### Research question 1:...

```{r, echo=FALSE, fig.cap="DAG for Research question 1.", fig.height=10, fig.width=18}
dag.rq1 <- dags$rq1 %>%
  dagitty::dagitty()
dag.rq1 %>% ggdag::tidy_dagitty() %>%
  ggdag::ggdag_status(text_size = 10, text_col = "black") +
  ggdag::theme_dag() + ggdag::scale_adjusted() +
  ggplot2::theme(legend.text = ggplot2::element_text(size = 20))
```

The `r I(type.mas)` sufficient adjustment sets for estimating the `r I(type.effect)` effect of X on Y are:

```{r, echo=FALSE}
dagitty::adjustmentSets(x = dag.rq1, 
                        type = type.mas, effect = type.effect)
```

\newpage

<!-- ####################################################################### -->
# Bibliography
<!-- ####################################################################### -->

### 2023-01-12 code/dictionaries.R

@@ -1,0 +1,1 @@
+# Script containing dictionaries (e.g., exposures)

### 2023-01-12 housekeeping/list_files.sh

@@ -1,0 +1,16 @@
+## Update README file of given directory with list of files
+
+# Date and time
+dt=$(date '+%d/%m/%Y %H:%M:%S');
+
+# Loop over files in given directory
+for entry in "$1"/*
+do
+  if [[ -f $entry ]]
+  then
+    echo "${entry:0} $dt" >> "$1/README.md";
+  fi
+done
+
+# Sort README file
+sort -o "$1/README.md"{,}

### 2023-01-12 docs/causal_roadmap.qmd
Converted to qmd

--- 2023-01-12 docs/causal_roadmap.Rmd
@@
 
-```{r setup, include=FALSE}
+```{r}
+#| label: setup
+#| include: false
 library(tidyverse)
@@
 
-```{r, echo=FALSE}
+```{r}
+#| echo: false
 # Type of adjustment set from DAG
@@
 
-```{r causal-graph, echo=FALSE, fig.cap="Causal graph corresponding to structural equations.", fig.height=3, fig.width=3}
+```{r}
+#| label: causal-graph
+#| echo: false
+#| fig.cap: Causal graph corresponding to structural equations.
+#| fig.height: 3.0
+#| fig.width: 3.0
 causal.graph <- ggdag::dagify(A ~ W + U_A, Y ~ W + A + U_Y, W ~ U_W, 
@@
 
-```{r echo=FALSE, fig.cap="Simplified DAG for research question 1."}
+```{r}
+#| echo: false
+#| fig.cap: Simplified DAG for research question 1.
 sdag1 <- ggdag::dagify(X_ti ~ W_t, 
@@
 
-```{r, echo=FALSE}
+```{r}
+#| echo: false
 ggpubr::ggarrange(sdag1, ..., 
@@
 
-```{r, echo=FALSE}
+```{r}
+#| echo: false
 source("../DAGs/dag_v1.R")
@@
 
-```{r, echo=FALSE, fig.cap="DAG for Research question 1.", fig.height=10, fig.width=18}
+```{r}
+#| echo: false
+#| fig.cap: DAG for Research question 1.
+#| fig.height: 10.0
+#| fig.width: 18.0
 dag.rq1 <- dags$rq1 %>%
@@
 
-```{r, echo=FALSE}
+```{r}
+#| echo: false
 dagitty::adjustmentSets(x = dag.rq1, 

### 2023-01-12 output/paper/figures.qmd
Converted to qmd

---
title: "Figures"
output:
  bookdown::pdf_document2
---

```{r}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, message = FALSE, 
                      cache = TRUE)
```

<!-- ####################################################################### -->
<!-- Workflow -->
```{r}
```
<!-- ####################################################################### -->

### 2023-01-12 output/paper/manuscript.qmd
Converted to qmd

---
title: "Paper X"
author:
  - Lorenzo Fabbri:
      correspondence: "yes"
      email: "lorenzo.fabbri@isglobal.org"
      institute: [isg, upf]
institute:
  - isg: ISGlobal
  - upf: UPF
output:
  bookdown::pdf_document2:
    toc: no
    pandoc_args:
      - --filter=pandoc-crossref
      - --lua-filter=scholarly-metadata.lua
      - --lua-filter=author-info-blocks.lua
  bookdown::word_document: default
  bookdown::word_document2:
    pandoc_args:
      - --filter=pandoc-crossref
      - --lua-filter=scholarly-metadata.lua
      - --lua-filter=author-info-blocks.lua
      - --reference-doc=reference_document.docx
bibliography:
  - bibliography.bib
#csl: journal.csl
link-citations: yes
linkcolor: blue
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage[switch, pagewise, running]{lineno}
  - \linenumbers
  - \renewcommand\linenumberfont{\normalfont\small}
  - \usepackage{rotating}
  - \usepackage{float}
abstract: |-
 \singlespacing
 **Background**

 **Objectives**

 **Methods**

 **Results**

 **Discussion**

  **Keywords:** Causal Inference
---

```{=latex}
\setcounter{tocdepth}{4}
\tableofcontents
```

\twocolumn
\doublespacing
\pagewiselinenumbers

<!-- ####################################################################### -->
<!-- INTRODUCTION -->
# Introduction {#intro}
```{r}
#| child: sections/introduction.qmd
```
<!-- ####################################################################### -->

<!-- ####################################################################### -->
<!-- METHODS -->
# Methods {#methods}
```{r}
#| child: sections/methods.qmd
```
<!-- ####################################################################### -->

<!-- ####################################################################### -->
<!-- RESULTS -->
# Results {#res}
```{r}
#| child: sections/results.qmd
```
<!-- ####################################################################### -->

<!-- ####################################################################### -->
<!-- DISCUSSION -->
# Discussion {#discussion}
```{r}
#| child: sections/discussion.qmd
```
<!-- ####################################################################### -->

\newpage
\onecolumn
\singlespacing

# References

### 2023-01-12 output/paper/tables.qmd
Converted to qmd

---
title: "Tables"
output:
  bookdown::pdf_document2
---

```{r}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, message = FALSE, 
                      cache = TRUE)
```

<!-- ####################################################################### -->
<!-- Study populations -->
```{r}

```
<!-- ####################################################################### -->

<!-- ####################################################################### -->
<!-- Exposures description -->
```{r}

```
<!-- ####################################################################### -->

### 2023-01-13 output/paper/manuscript.qmd
Trying to fix qmd of manuscript without using lua filters

@@
 title: "Paper X"
+keywords:
+  - Causal Inference
+  - Epidemiology
 author:
-  - Lorenzo Fabbri:
-      correspondence: "yes"
-      email: "lorenzo.fabbri@isglobal.org"
-      institute: [isg, upf]
-institute:
-  - isg: ISGlobal
-  - upf: UPF
-output:
-  bookdown::pdf_document2:
-    toc: no
-    pandoc_args:
-      - --filter=pandoc-crossref
-      - --lua-filter=scholarly-metadata.lua
-      - --lua-filter=author-info-blocks.lua
-  bookdown::word_document: default
-  bookdown::word_document2:
-    pandoc_args:
-      - --filter=pandoc-crossref
-      - --lua-filter=scholarly-metadata.lua
-      - --lua-filter=author-info-blocks.lua
-      - --reference-doc=reference_document.docx
-bibliography:
-  - bibliography.bib
+  - name: "Lorenzo Fabbri"
+    number: 1
+    email: lorenzo.fabbri@isglobal.org
+    orcid: 0000-0003-3031-322X
+    attributes:
+      corresponding: true
+    affiliations:
+      - ref: isg
+      - ref: upf
+affiliations:
+  - id: isg
+    name: ISGlobal
+    city: Barcelona
+    state: Spain
+  - id: upf
+    name: UPF
+    city: Barcelona
+    state: Spain
+format:
+  pdf:
+    pdf-engine: pdflatex
+    toc: false
+  # docx:
 #csl: journal.csl
-link-citations: yes
+bibliography: bibliography.bib
+link-citations: true
 linkcolor: blue
@@
   - \usepackage{float}
-abstract: |-
+abstract: |
  \singlespacing
@@
  **Discussion**
-
-  **Keywords:** Causal Inference
 ---

### 2023-01-16 docs/causal_roadmap.qmd
Set up targets. Better bibtex to auto-update bibliography from sub-collection for paper 3 from Zotero. Targets: one pipeline for each research question of causal roadmap, with data saved outside of git repo

@@
 title: "Causal Roadmap"
-author: Lorenzo Fabbri
-output:
-  bookdown::pdf_document2:
-    number_sections: false
-  html_document: default
-bibliography: references.bib
-editor_options: 
-  chunk_output_type: console
+keywords:
+  - Causal Inference
+  - Epidemiology
+author:
+  - name: "Lorenzo Fabbri"
+    number: 1
+    email: lorenzo.fabbri@isglobal.org
+    orcid: 0000-0003-3031-322X
+    attributes:
+      corresponding: true
+    affiliations:
+      - ref: isg
+      - ref: upf
+affiliations:
+  - id: isg
+    name: ISGlobal
+    city: Barcelona
+    state: Spain
+  - id: upf
+    name: UPF
+    city: Barcelona
+    state: Spain
+format:
+  pdf:
+    pdf-engine: pdflatex
+    toc: false
+bibliography: ../output/paper/bibliography.bib
+link-citations: true
+linkcolor: blue
+header-includes:
+  - \usepackage{setspace}\doublespacing
+  - \usepackage[switch, pagewise, running]{lineno}
+  - \linenumbers
+  - \renewcommand\linenumberfont{\normalfont\small}
+  - \usepackage{rotating}
+  - \usepackage{float}
+execute:
+  echo: false
 ---
@@
 #| label: setup
-#| include: false
 library(tidyverse)
@@
 ```{r}
-#| echo: false
 # Type of adjustment set from DAG
@@
 #| label: causal-graph
-#| echo: false
 #| fig.cap: Causal graph corresponding to structural equations.
@@
 ```{r}
-#| echo: false
 #| fig.cap: Simplified DAG for research question 1.
@@
 ```{r}
-#| echo: false
-ggpubr::ggarrange(sdag1, ..., 
-                  ncol = x, nrow = y)
+ggpubr::ggarrange(sdag1, 
+                  ncol = 1, nrow = 1)
 ```
@@
 ```{r}
-#| echo: false
 source("../DAGs/dag_v1.R")
@@
 ```{r}
-#| echo: false
 #| fig.cap: DAG for Research question 1.
@@
 ```{r}
-#| echo: false
 dagitty::adjustmentSets(x = dag.rq1, 

### 2023-01-16 _targets_rq1.R

@@ -1,0 +1,25 @@
+# Load packages required to define the pipeline
+library(targets)
+library(tarchetypes)
+source("code/")
+
+# Set target options
+targets::tar_option_set(
+  packages = c(""), 
+  format = "qs"
+)
+options(clustermq.scheduler = "multicore")
+
+# Run the R scripts in the code/ folder with custom functions
+targets::tar_source()
+
+list(
+  targets::tar_target(
+    name = data, 
+    command = ""
+  ), 
+  targets::tar_target(
+    name = model, 
+    command = ""
+  )
+)

### 2023-03-01 docs/causal_roadmap.qmd
Started working on DAGs: all nodes, DAG for chemical to intelligence
Cleaned the causal roadmap.

--- 2023-01-12 docs/causal_roadmap.qmd
@@
 ---
-title: "Causal Roadmap"
-author: Lorenzo Fabbri
-output:
-  bookdown::pdf_document2:
-    number_sections: false
-  html_document: default
-bibliography: references.bib
+title: "Paper 3 - Causal Roadmap"
+keywords:
+  - Causal Inference
+  - Epidemiology
+  - Triangulation
+  - Modified Treatment Policy
+  - Mendelian Randomization
+  - Metabolomics
+  - Endocrine Disrupting Chemicals
+  - Neurodevelopment
+  - Neurobehavior
+
+author:
+  - name: "Lorenzo Fabbri"
+    number: 1
+    email: lorenzo.fabbri@isglobal.org
+    orcid: 0000-0003-3031-322X
+    attributes:
+      corresponding: true
+    affiliations:
+      - ref: isg
+      - ref: upf
+affiliations:
+  - id: isg
+    name: Barcelona Institute for Global Health (ISGlobal)
+    city: Barcelona
+    state: Spain
+  - id: upf
+    name: Pompeu Fabra University
+    city: Barcelona
+    state: Spain
+
+format:
+  pdf:
+    documentclass: report
+    pdf-engine: pdflatex
+    toc: true
+    margin-left: 15mm
+    margin-right: 15mm
+    margin-bottom: 15mm
+    margin-top: 15mm
+
+bibliography: ../output/paper/bibliography.bib
+link-citations: true
+linkcolor: blue
+
+header-includes:
+  - \usepackage{rotating}
+  - \usepackage{float}
+
+execute:
+  echo: false
+
 editor_options: 
@@
 #| label: setup
-#| include: false
+#| message: false
 library(tidyverse)
@@
 ```{r}
-#| echo: false
 # Type of adjustment set from DAG
-type.mas <- "minimal"
+type_mas <- "minimal"
 # Type of effect to be estimated from DAG
-type.effect <- "direct"
+type_effect <- "direct"
 ```
 
-\textcolor{red}{Most of the following was **copy-pasted** from published papers. Some of the Sections must be filled depending on the chosen causal framework.}
+\textcolor{red}{Most of the following was **copy-pasted** from published papers.}
 
@@
 we assumed its observed data $O_{ti} = (W_{ti}, A_{ti}, Y_{ti})$ were generated by sampling from a distribution 
-$\mathbb{P}_{0,t}$ compatible with the causal model specified in \@ref(eq:npsem).
+$\mathbb{P}_{0,t}$ compatible with the causal model specified in @eq-npsem.
 
@@
 
-\begin{equation}
+$$
 \begin{aligned}
@@
 \end{aligned}
-(\#eq:npsem)
-\end{equation}
+$$ {#eq-npsem}
 
@@
 and $(U_{W_{t}}, U_{A_{t}}, U_{Y_{t}})$ were the unmeasured factors contributing the confounders, exposures, and outcome, respectively. 
-The corresponding causal graph (without indexes for ease of notation) is given in Figure \@ref(fig:causal-graph).
+The corresponding causal graph (without indexes for ease of notation) is given in @fig-causal-graph.
 
 ```{r}
-#| label: causal-graph
-#| echo: false
+#| label: fig-causal-graph
 #| fig.cap: Causal graph corresponding to structural equations.
@@
 #| fig.width: 3.0
-causal.graph <- ggdag::dagify(A ~ W + U_A, Y ~ W + A + U_Y, W ~ U_W, 
+causal_graph <- ggdag::dagify(A ~ W + U_A, Y ~ W + A + U_Y, W ~ U_W, 
                               exposure = "A", outcome = "Y") %>%
   ggdag::tidy_dagitty()
-ggdag::ggdag(causal.graph) + ggdag::theme_dag()
+ggdag::ggdag(causal_graph) + ggdag::theme_dag()
 ```
@@
 
-As **secondary analyses**, we estimated...
-
 In the following sections, we present the simplified DAG specific to each research question.
@@
 
+## Research question 2:...
+
 ```{r}
-#| echo: false
-ggpubr::ggarrange(sdag1, ..., 
-                  ncol = x, nrow = y)
+#| fig.cap: Simplified DAG for research question 2.
+sdag2 <- ggdag::dagify(X_ti ~ W_t, 
+                       Y ~ W_t + X_ti, 
+                       exposure = "X_ti", outcome = "Y") %>%
+  ggdag::tidy_dagitty() %>%
+  ggdag::ggdag(text_size = 2, node_size = 10) + ggdag::theme_dag() +
+  ggplot2::ggtitle("RQ2")
+```
+
+## Research question 3:...
+
+```{r}
+#| fig.cap: Simplified DAG for research question 3.
+sdag3 <- ggdag::dagify(X_ti ~ W_t, 
+                       Y ~ W_t + X_ti, 
+                       exposure = "X_ti", outcome = "Y") %>%
+  ggdag::tidy_dagitty() %>%
+  ggdag::ggdag(text_size = 2, node_size = 10) + ggdag::theme_dag() +
+  ggplot2::ggtitle("RQ3")
+```
+
+```{r}
+ggpubr::ggarrange(sdag1, sdag2, sdag3, 
+                  ncol = 3, nrow = 1)
 ```
@@
 ```{r}
-#| echo: false
+#| message: false
 source("../DAGs/dag_v1.R")
@@
 
-dags <- dag.v1()
+dags <- dag_v1()
 ```
@@
 #| fig.width: 18.0
-dag.rq1 <- dags$rq1 %>%
+dag_rq1 <- dags$chem_to_out %>%
   dagitty::dagitty()
-dag.rq1 %>% ggdag::tidy_dagitty() %>%
+dag_rq1 %>% ggdag::tidy_dagitty() %>%
   ggdag::ggdag_status(text_size = 10, text_col = "black") +
@@
 
-The `r I(type.mas)` sufficient adjustment sets for estimating the `r I(type.effect)` effect of X on Y are:
+The `r I(type_mas)` sufficient adjustment sets for estimating the `r I(type_effect)` effect of X on Y are:
 
 ```{r}
-#| echo: false
-dagitty::adjustmentSets(x = dag.rq1, 
-                        type = type.mas, effect = type.effect)
+dagitty::adjustmentSets(x = dag_rq1, 
+                        type = type_mas, effect = type_effect)
 ```

### 2023-03-06 docs/causal_roadmap.qmd
Kept working on causal triangulation roadmap.

@@
   - Neurobehavior
-
 author:
@@
     state: Spain
-
 format:
@@
     margin-top: 15mm
-
-bibliography: ../output/paper/bibliography.bib
+    include-in-header:
+      - \usepackage{rotating}
+      - \usepackage{float}
+    pandoc_args:
+      - !expr acronymsdown::add_filter()
+bibliography: ../../../PhD.bib
 link-citations: true
 linkcolor: blue
-
-header-includes:
-  - \usepackage{rotating}
-  - \usepackage{float}
-
 execute:
   echo: false
-
 editor_options: 
   chunk_output_type: console
+filters:
+  - /home/lorenzo/R/x86_64-pc-linux-gnu-library/4.1/acronymsdown/parse-acronyms.lua
+acronyms:
+  insert_links: false
+  sorting: alphabetical
+  keys:
+    - key: edc
+      shortname: EDC
+      longname: endocrine disruptor
+    - key: edcs
+      shortname: EDCs
+      longname: endocrine disruptors
+    - key: mtp
+      shortname: MTP
+      longname: modified treatment policy
+    - key: mtps
+      shortname: MTPs
+      longname: modified treatment policies
+    - key: tmle
+      shortname: TMLE
+      longname: Targeted Minimum Loss-Based Estimator
+    - key: sl
+      shortname: SL
+      longname: Super Learner
+    - key: cpm
+      shortname: CPM
+      longname: Raven’s Coloured Progressive Matrices
+    - key: ant
+      shortname: ANT
+      longname: Attention Network Test
+    - key: cbcl
+      shortname: CBCL
+      longname: Child Behavior Checklist
+    - key: npsem
+      shortname: NPSEM
+      longname: non-parametric Structural Equation Model
+    - key: prs
+      shortname: PRS
+      longname: polygenic risk score
+    - key: mr
+      shortname: MR
+      longname: Mendelian Randomization
 ---
@@
 ```{r}
+#| echo: false
+#| include: false
+
+source("../../../all_phd/info_phd.R")
+source("../../../all_phd/triangulation_methods.R")
+infos <- info_phd()
+methods <- describe_triangulation_methods()
+```
+
+```{r}
 # Type of adjustment set from DAG
@@
 
+\newpage
+
 \textcolor{red}{Most of the following was **copy-pasted** from published papers.}
@@
 <!-- ####################################################################### -->
-# Step 0: Formulate the research question(s)
+# Formulate the research question(s)
 <!-- ####################################################################### -->
 
-The **aim** of the present study was to research the...
+The **aim** of the present study is to research the short-term effects of postnatal exposure to non-persistent \acr{edcs} on neurodevelopment and neurobehavior in childhood, and how the metabolome and the proteome might mediate these effects.
+We will strengthen the estimated effects by making use of the principles and criteria of triangulation [@LawlorTillingDaveySmith:2016].
+To estimate these effects, we will primarily rely on the use of \acr{mtps} [@munozPopulationInterventionCausal2012; @haneuseEstimationEffectInterventions2013; @diazNonparametricCausalEffects2021] in combination with \acr{tmle} [@vanderlaanTargetedMinimumLossBased2018], to avoid relying on arbitrary parametric assumptions.
+We will consider both additive and multiplicative shifts of the exposures' distribution.
 
-The study **population** was based on..., consisting of $\text{N}=$... Inclusion criteria were:...
+The primary study **population** is based on the HELIX sub-cohort, consisting of $\text{N}=1200$ mother-child pairs from six existing European birth cohorts [@vrijheidHumanEarlylifeExposome2014].
+We will replicate the obtained results in the HELIX Child Panel, consisting of $\text{N}=150$ children followed twice for one week.
+**Exposures** consisted of non-persistent \acr{edcs} (phenols, phthalates, and organophosphate compounds), measured in childhood in a pool of two urine samples.
+A single-spot blood sample, collected during the visit, was used for serum metabolomics and plasma proteomics.
+**Outcomes** related to childhood neurodevelopment and neurobehavior, included:
 
-Measured **exposures** consisted of...
-
-The **primary outcome** was... **Secondary outcomes** included...
+- \acr{cpm}, for assessing non-verbal intelligence.
+- Computerised n-back test, for assessing working memory.
+- \acr{ant}.
+- \acr{cbcl}, for assessing behavioural and emotional problems.
 
@@
 
+\textcolor{red}{TODO}
+
 <!-- ####################################################################### -->
-# Step 1: Define a realistic statistical model
+# Define a realistic statistical model
 <!-- ####################################################################### -->
@@
 and $Y_t$ the clinical outcome of interest. For each subject $i$ and each time index $t$, 
-we assumed its observed data $O_{ti} = (W_{ti}, A_{ti}, Y_{ti})$ were generated by sampling from a distribution 
+we assume its observed data $O_{ti} = (W_{ti}, A_{ti}, Y_{ti})$ were generated by sampling from a distribution 
 $\mathbb{P}_{0,t}$ compatible with the causal model specified in @eq-npsem.
@@
 <!-- ####################################################################### -->
-# Step 2: Specify a causal model and causal quantity of interest
+# Specify a causal model and causal quantity of interest
 <!-- ####################################################################### -->
 
-We specified the following nonparametric structural equation model (NPSEM) to present the data generating process, 
+We specify the following \acr{npsem} to present the data generating process, 
 including its measured confounders $W_t$, exposures $A_t$, and outcome $Y_t$, for each time index $t$:
@@
 
-where $(f_{W{_t}}, f_{A_{t}}, f_{Y_{t}})$ were the nonparametric structural equations, 
-and $(U_{W_{t}}, U_{A_{t}}, U_{Y_{t}})$ were the unmeasured factors contributing the confounders, exposures, and outcome, respectively. 
+where $(f_{W{_t}}, f_{A_{t}}, f_{Y_{t}})$ are the non-parametric structural equations, 
+and $(U_{W_{t}}, U_{A_{t}}, U_{Y_{t}})$ are the unmeasured factors contributing the confounders, exposures, and outcome, respectively. 
 The corresponding causal graph (without indexes for ease of notation) is given in @fig-causal-graph.
@@
 
-For each time index, we specified our causal parameter $\phi$ as...
+In order to test the \acr{npsem}’s implications (conditional independencies), we will perform simulations from the DAGs, checking assumptions and data consistency, using the `dagitty` R package.
 
-As **primary analysis**, we estimated....
+Following the \acr{mtp} framework, we will generate counterfactual outcomes $Y^d = f_Y (W, A^d, U_Y)$ by intervening on the causal model to shift the observed exposures by some user-specified function $A^d = d(\cdot)$.
+We will focus on simple interventions to shift the observed exposure $A$ by an additive constant $c$ or multiplicative constant $k$: $d(A,c) = c+A$ and $d(A,k) = k \times A$, respectively.
 
-In the following sections, we present the simplified DAG specific to each research question.
+We specify our causal parameter $\phi$ as the difference in the expected counterfactual outcome under the shifted exposures and the expected outcome under the observed exposures:
 
-## Research question 1:...
+$$
+\phi^{\Delta} = \mathbb{E}[Y^d] - \mathbb{E}[Y],
+$$
+
+where the expectation is over the individuals of our target population.
+
+As **primary analyses**, we will estimate the short-term effects of childhood exposure to individual non-persistent \acr{edcs} in relation to one or more neurodevelopmental outcomes of interest.
+We will further explore the potential mediating role of the serum metabolome and the plasma proteome.
+
+In the following sections, we present a brief summary of each research question.
+
+## Research question 1: What are the short-term effects of childhood exposure to \acr{edcs} on the outcomes?
+
+We will estimate the short-term effects of childhood exposure to non-persistent \acr{edcs} on the outcomes of interest measured in childhood.
+We will perform an ExWAS kind of analysis with the dependent variable $y$ being the outcome, and the independent variable $x_i$ being the levels of \acr{edc} $i$.
+We will assume that all associations are confounded by the same set of confounders, independently of the \acr{edc} considered.
 
@@
 
-## Research question 2:...
+## Research question 2: What are the short-term effects of childhood exposure to \acr{edcs} on metabolites and proteins?
 
-```{r}
-#| fig.cap: Simplified DAG for research question 2.
-sdag2 <- ggdag::dagify(X_ti ~ W_t, 
-                       Y ~ W_t + X_ti, 
-                       exposure = "X_ti", outcome = "Y") %>%
-  ggdag::tidy_dagitty() %>%
-  ggdag::ggdag(text_size = 2, node_size = 10) + ggdag::theme_dag() +
-  ggplot2::ggtitle("RQ2")
-```
+We will estimate the short-term effects of childhood exposure to non-persistent \acr{edcs} on metabolites and proteins measured in childhood.
+We will perform an MWAS analysis with the dependent variable $m_i$ being the metabolite or protein $i$, and the independent variable $x_j$ being the levels of \acr{edc} $j$.
+We will assume that all associations are confounded by the same set of confounders, independently of the \acr{edc} considered.
 
-## Research question 3:...
+## Research question 3: What are the short-term effects of childhood *exposure* to metabolites and proteins on the outcomes?
 
-```{r}
-#| fig.cap: Simplified DAG for research question 3.
-sdag3 <- ggdag::dagify(X_ti ~ W_t, 
-                       Y ~ W_t + X_ti, 
-                       exposure = "X_ti", outcome = "Y") %>%
-  ggdag::tidy_dagitty() %>%
-  ggdag::ggdag(text_size = 2, node_size = 10) + ggdag::theme_dag() +
-  ggplot2::ggtitle("RQ3")
-```
-
-```{r}
-ggpubr::ggarrange(sdag1, sdag2, sdag3, 
-                  ncol = 3, nrow = 1)
-```
+We will estimate the short-term effects of childhood *exposure* to metabolites and proteins on the outcomes of interest measured in childhood.
+We will perform an MWAS analysis with the dependent variable $y$ being the outcome, and the independent variable $m_i$ being the metabolite or protein $i$.
+We will assume that all associations are confounded by the same set of confounders, independently of the \acr{edc} considered.
 
 <!-- ####################################################################### -->
-# Step 3: Identification and the statistical estimand
+# Identification and the statistical estimand
 <!-- ####################################################################### -->
 
-For the causal parameter $\phi$, which involves a summary measure of the distribution of counterfactuals, 
-to be identified in terms of the observed data distribution, several assumptions would be required:
+For the causal parameter $\phi^{\Delta}$, which involves a summary measure of the distribution of counterfactuals, to be identified in terms of the observed data distribution, several assumptions would be required:
 
-* **No unmeasured confounding**: There are no unmeasured common causes of the exposure and the subsequent outcome.
-In our context, this assumption would be violated if, for example, an unmeasured variable influenced both the exposure levels and the outcome.
+* **No unmeasured confounding**: There are no unmeasured common causes of the exposure and the subsequent outcome, which we can formalize as $Y^d \perp \!\!\!
+\perp A|W$.
+In our context, this assumption would be violated if, for example, an unmeasured variable influences both the exposure levels and the clinical outcome.
 We cannot guarantee that this assumption holds, and therefore limit our interpretations to statistical associations rather than causal effects.
-* **Positivity**:....
+* **Positivity**: If $(a, w)$ is within the support of $A, W$, then $(d(a,c), w)$ for additive shifts and $(d(a,k), w)$ for multiplicative shifts must also be within the support of $A,W$.
+In practice, this means that for any given time index and set of adjustment covariates, there is a positive probability of finding a subject with the same covariate values and a exposure level matching the shifted value.
+We will attempt to improve plausibility of this assumption by considering small shifts, while recognizing that this approach makes our causal effect data-adaptive.
 * **Independence** of subjects. This assumption also implies **no interference**: the exposure level $a$ for a given subject does not affect the outcomes of the other subjects.
-* **Consistency**: If $A = a$ for any subject, then $Y(a) = Y$. This means that the counterfactual outcome for a subject with its observed exposure level is the observed outcome.
+* **Consistency**: If $A = a$ for any subject, then $Y(a) = Y$, and hence the full observed set of outcomes when $A^d = A$ is simply $Y(A^d) = Y$.
+This means that the counterfactual outcome for a subject with its observed exposure level is the observed outcome.
 * **Time-ordering**: The confounders $W$ precede the exposure $A$, which also precedes the outcome $Y$.
@@
 If these identifiability assumptions held, we could specify and focus our estimation efforts on a statistical estimand that equals the wished-for causal effect.
-In the (likely) case that they are not satisfied, we could still specify and focus our estimation efforts on a statistical estimand that is as close as possible to the causal parameter.
-Factoring the joint distribution of the observed data $\mathbb{P}_0$ into $\mathbb{P}_0 (O) = \mathbb{P}_0(Y|A,W) \mathbb{P}_0(A|W) \mathbb{P}_0(W)$, it can be shown that the statistical estimand corresponding to expected counterfactual outcome..., is given by
+In the likely case that they are not satisfied, we could still specify and focus our estimation efforts on a statistical estimand that is as close as possible to the causal parameter.
+Factoring the joint distribution of the observed data $\mathbb{P}_0$ into $\mathbb{P}_0 (O) = \mathbb{P}_0(Y|A,W) \mathbb{P}_0(A|W) \mathbb{P}_0(W)$, it can be shown that the statistical estimand corresponding to expected counterfactual outcome under shift $d$, $\mathbb{E}[Y^d]$, is given by
 
 $$
-..., 
+\psi_0 (A^d) = \int \mathbb{E} (Y | A=a^d, W=w) dF_{A,W}(a, w), 
 $$
 
-Therefore, our statistical estimand of interest was
+with $dF_{A,W}(a, w)$ as the joint density of received exposures $A$ and covariate levels $W$ being integrated over.
+We refer to $\psi_0 (A^d)$ as the *shift parameter*.
+Under no shift, the expected outcome was identified as $\psi_0 (A^d) = \mathbb{E} (Y)$.
+Therefore, our statistical estimand of interest, corresponding to the expected difference in the outcome under shifted and observed exposure levels, is
 
 $$
-....
+\psi_0^{\Delta} = \psi_0 (A^d) - \psi_0 (A) = \int \mathbb{E} (Y | A=a^d, W=w)dF_{A,W} (a,w) - \mathbb{E} (Y).
 $$
@@
 <!-- ####################################################################### -->
-# Step 4: Estimation from data and statistical inference
+# Estimation from data and statistical inference
 <!-- ####################################################################### -->
@@
 <!-- ####################################################################### -->
-# Step 5: Interpretation and sensitivity analyses to inform a substantive conclusion
+# Triangulation: Improving causal inference in aetiological epidemiology
+<!-- ####################################################################### -->
+
+From [@LawlorTillingDaveySmith:2016]:
+
+> *Triangulation is the practice of obtaining more reliable answers to research questions through integrating results from several different approaches, where each approach has different key sources of potential bias that are unrelated to each other.
+With respect to causal questions in aetiological epidemiology, if the results of different approaches all point to the same conclusion, this strengthens confidence in the finding.
+This is particularly the case when the key sources of bias of some of the approaches would predict that findings would point in opposite directions if they were due to such biases.*
+
+@tbl-triangulation-paper3 summarizes the approaches that we will use to perform triangulation, based on [@LawlorTillingDaveySmith:2016].
+
+{{< include ../../../all_phd/tab_triangulation_p3.qmd >}}
+
+For each research question, and each approach, we will perform the following analyses:
+
+- `outcome ~ chemical`
+  - Regression analysis.
+We will use \acr{tmle} in combination with \acr{sl}, and a library of estimators including both simple parametric models (e.g., `glm`) and data-adaptive semi-parametric models.
+The models will be adjusted for relevant confounders, as identified with `dagitty`.
+We hypothesize the presence of residual confounding, especially due to genetic and parental socio-economic factors, which would result in **exaggeration** of any true causal effect.
+  - Outcome negative control study.
+  - Literature search.
+  - Qualitative \acr{prs}.
+- `omic ~ chemical`
+  - Regression analysis.
+  - Cross-cohort comparison. We will compare the regression models' results between subjects of different ethnic origins.
+  - Literature search.
+  - Qualitative \acr{prs}.
+- `outcome ~ omic`
+  - Regression analysis.
+  - Cross-cohort comparison.
+  - \acr{mr}.
+
+# Mendelian Randomization: Assumptions
+
+<!-- ####################################################################### -->
+# Interpretation and sensitivity analyses to inform a substantive conclusion
 <!-- ####################################################################### -->

### 2023-03-07 docs/causal_roadmap.qmd
Kept working causal roadmap paper 3

@@
   pdf:
-    documentclass: report
     pdf-engine: pdflatex
+    number-sections: true
     toc: true
@@
     margin-right: 15mm
-    margin-bottom: 15mm
+    margin-bottom: 20mm
     margin-top: 15mm
@@
       longname: Mendelian Randomization
+    - key: se
+      shortname: SE
+      longname: socio-economic
+    - key: iv
+      shortname: IV
+      longname: instrumental variable
+    - key: ivs
+      shortname: IVs
+      longname: instrumental variables
+    - key: ctd
+      shortname: CTD
+      longname: Comparative Toxicogenomics Database
+    - key: cv
+      shortname: CV
+      longname: cross-validation
 ---
@@
 type_effect <- "direct"
+
+# Number of folds for Cross-Validation (CV)
+cv.folds <- 10
 ```
@@
 
+To estimate the expected outcome under the observed exposure $\psi_0 (A) = \mathbb{E} (Y)$, we will use the empirical mean outcome.
+For estimating the shift parameter $\psi_0 (A^d)$, we will use \acr{tmle}, which typically uses the factorization of the observed data distribution into an outcome regression $\bar{Q}(A,W) = \mathbb{E} (Y|A,W)$ and an intervention mechanism $g(A|W) = P(A|W)$.
+Then, an initial estimate of the outcome regression is updated by a fluctuation that is a function of the intervention mechanism.
+In this study, we will use the implementation of [@diazNonparametricCausalEffects2021].
+Here, we summarize the algorithm for the shift parameter $\psi_0 (A^d)$.
+We first define the density ratio
+
+$$
+r(a,w) = \frac{g^d(a|w)}{g(a|w)}.
+$$
+
+Estimation of the conditional density $g$ is done by recasting it as a classification problem. It can be shown that
+
+$$
+r(a,w) = \frac{P(\Lambda=1|a,w)}{1-P(\Lambda=1|a,w)},
+$$
+
+with $\Lambda$ indicating whether each subject received the natural or shifted value of treatment.
+We refer to $P(\Lambda=1|a,w)$ as $\bar{\lambda}$.
+To implement estimation of $\bar{\lambda}$ and $\bar{Q}$ flexibly, we will use the ensemble machine learning algorithm \acr{sl}.
+In the following, we present the steps necessary to estimate $\psi_0 (A^d)$ and the associational parameter $\psi_0^{\Delta}$.
+We note that our bounded continuous outcome will be **scaled** to $[0,1]$, and that to handle missing covariates $W_{i,j}$, we will redefine the data to include $\delta_{ij}$, where $\delta_{ij}=1$ if the covariate was observed and $0$ otherwise.
+
+1. Define and calculate the density ratio under the specified shift $\hat{r} (a_i, w_i)$ for each subject $i = 1, \dots, n$ from $\bar{\lambda}$ estimated via \acr{sl}.
+1. Generate initial conditional expectations for each subject, denoted $\bar{Q}(A,W)$.
+1. Fit the following logistic regression on the observed (scaled) $Y$ values, using the logit of the initial estimates as an offset
+$$
+\text{logit} (\bar{Q}^*(A,W)) = \text{logit} (\bar{Q}(A,W)) + \epsilon
+$$
+with weights $\hat{r} (a_i, w_i)$, calculating the estimated intercept $\hat{\epsilon}$, where $\bar{Q}^*(A,W)$ is now a targeted estimate of the conditional mean outcome.
+1. Use the resulting $\hat{\epsilon}$ value to generate targeted predictions under a shift, $\bar{Q}^*(A^d,W)$:
+$$
+\bar{Q}^*(A^d,W) = \text{logit}^{-1} \left[ \text{logit} (\bar{Q} (A^d, W)) + \hat{\epsilon} \right].
+$$
+1. With the updated estimates $\bar{Q}^*(A^d,W)$ in hand (after unscaling), define:
+$$
+\hat{\psi}_{\text{tmle}} = \frac{1}{n} \sum_{i=1}^n \bar{Q}^*(A^d,W).
+$$
+
+Using the empirical mean $\hat{Y} = \frac{1}{n} \sum_{i=1}^nY_i$, to estimate $\psi_0(A) = \mathbb{E}(Y)$, we obtain an estimate of the statistical association parameter $\psi_0^{\Delta} = \psi_0(A^d) = \psi_0(A)$ with
+$$
+\hat{\psi}^{\Delta} = \hat{\psi}_{\text{tmle}} - \bar{Y}.
+$$
+
+The $95\%$ Wald-style confidence interval $\hat{\psi}^{\Delta} \pm 1.96 \sqrt{Var(\hat{\psi}^{\Delta})}$ is obtained from the influence curve.
+
+Using `r I(cv.folds)`-fold \acr{cv}, \acr{sl} builds the best weighted combination of predictions from the algorithms listed in Table \@ref(tab:lmtp-specs).
+In case of repeated measures of the exposures, an `ID` variable will be used to make sure that the `r I(cv.folds)`-fold \acr{cv} splits will keep observations from the same individuals in the same split.
+
 <!-- ####################################################################### -->
@@
 The models will be adjusted for relevant confounders, as identified with `dagitty`.
-We hypothesize the presence of residual confounding, especially due to genetic and parental socio-economic factors, which would result in **exaggeration** of any true causal effect.
+We hypothesize the presence of residual confounding, especially due to genetic and parental \acr{se} factors, which would result in **exaggeration** of any true causal effect.
   - Outcome negative control study.
+We will try to identify outcomes in the same cohorts that would be affected by confounders that are relevant in the association of \acr{edcs} exposure to the clinical outcome of interest, but for which a biological/causal effect of the exposures is unlikely.
   - Literature search.
-  - Qualitative \acr{prs}.
+For each of the `outcome ~ chemical` associations, we will perform a literature search to compare the obtained results.
+We will make use of PubMed, using as keywords the clinical outcome and the *common* name of each chemical, and the \acr{ctd} [@DavisWiegersJohnson:2023].
+  - Qualitative \acr{prs}. \textcolor{red}{TODO}
 - `omic ~ chemical`
-  - Regression analysis.
-  - Cross-cohort comparison. We will compare the regression models' results between subjects of different ethnic origins.
-  - Literature search.
-  - Qualitative \acr{prs}.
+  - Regression analysis. As above.
+  - Cross-cohort comparison.
+We will compare the regression models' results between subjects of different ethnic origins.
+We postulate that \acr{se} factors represent the main confounders for this association.
+Based on the associations been \acr{se} factors (e.g., maternal education) and the omic markers, we will compare the results of `omic ~ chemical` with the expected modified associations (e.g., weaker or stronger).
+  - Literature search. As above.
+  - Qualitative \acr{prs}. \textcolor{red}{TODO}
 - `outcome ~ omic`
-  - Regression analysis.
-  - Cross-cohort comparison.
+  - Regression analysis. As above.
+  - Cross-cohort comparison. As above.
   - \acr{mr}.
+We will make use, ideally, of a weighted allele score of genetic variants known to be robustly associated with the omic markers as an \acr{IV}.
+We will use methods, including sensitivity analysis, to explore the possibility of bias due to: (i) weak instruments, and (ii) violation of the exclusion restriction criteria.
+Based on these, we will postulate whether the obtained results might be biased towards the null or not.
+By using \acr{mr}, we assume that this approach tests *exposure* to omic markers across the whole of childhood.
 
-# Mendelian Randomization: Assumptions
+## \acr{mr}: Assumptions
+
+\acr{mr} requires three \acr{iv} assumptions (taken from [Wikipedia](https://en.wikipedia.org/wiki/Mendelian_randomization)):
+
+- The genetic variant(s) used as instruments for the exposure is associated with the exposure (*relevance* assumption).
+- There are no confounders of the association between the genetic variant(s) and the outcome (*independence* or *exchangeability* assumption).
+- There is no pathway between the genetic variant(s) and the outcome, other than through the exposure (*exclusion restriction* or *no horizontal pleiotropy* assumption).
 
@@
 
-Highlight possible **limitations**:
+{{< include ../../../all_phd/app-rel-ci.qmd >}}
 
-* When considering only one exposure, we do not consider the effect of the others on that exposure and the outcome;
-* When we consider only one outcome, we do not consider the effect of the other outcomes on that one;
-* Although the interpretation of $\psi^{\text{causal}}$ is clear, how closely $\psi^{\text{stat}}$ matches it merits discussion.
-One considers the plausibility of each of the identifying assumptions in turn.
+---
 
-We performed **sensitivity analysis** to measure robustness to unmeasured confounders, based on the following considerations:
+{{< include ../../../all_phd/app-rel-si.qmd >}}
 
-* Broadly, an open question is whether the estimated treatment effect is biased due to confounding by unmeasured covariates.
-We can examine how the substantive conclusion would be impacted under a range of presumed causal bias, $\delta = \psi^{\text{causal}} - \psi^{\text{stat}}$.
-That is, how *strong* would a particular confounder (or group of confounders) have to be in order to change the conclusions of this study?
-In a worst case scenario, how vulnerable are the study's results to many or all unobserved confounders acting **together**, possibly non-linearly?
-The exercise illustrates how the effect estimates, and confidence interval bounds change, depending on the magnitude and direction of the hypothesized gap;
-* We implemented the following methods to conduct sensitivity analysis:
-  * Selection bias and stratification...
-  * Placebo treatment (replacement of the actual treatment with a random variable);
-  * We performed leave-1-out analyses to assess the influence of single subjects on the effect of interest.
+---
 
-We took into consideration the following points to *validate* our results:
-
-* We evaluated whether the obtained results are of any relevance from a public health point of view.
+{{< include ../../../all_phd/app-replicability.qmd >}}
 

### 2023-03-08 docs/causal_roadmap.qmd
Finished first draft of roadmap.

@@
       - \usepackage{float}
+      - \usepackage{typearea}
     pandoc_args:
@@
 acronyms:
+  fromfile: ../../../all_phd/acronyms.yml
   insert_links: false
   sorting: alphabetical
-  keys:
-    - key: edc
-      shortname: EDC
-      longname: endocrine disruptor
-    - key: edcs
-      shortname: EDCs
-      longname: endocrine disruptors
-    - key: mtp
-      shortname: MTP
-      longname: modified treatment policy
-    - key: mtps
-      shortname: MTPs
-      longname: modified treatment policies
-    - key: tmle
-      shortname: TMLE
-      longname: Targeted Minimum Loss-Based Estimator
-    - key: sl
-      shortname: SL
-      longname: Super Learner
-    - key: cpm
-      shortname: CPM
-      longname: Raven’s Coloured Progressive Matrices
-    - key: ant
-      shortname: ANT
-      longname: Attention Network Test
-    - key: cbcl
-      shortname: CBCL
-      longname: Child Behavior Checklist
-    - key: npsem
-      shortname: NPSEM
-      longname: non-parametric Structural Equation Model
-    - key: prs
-      shortname: PRS
-      longname: polygenic risk score
-    - key: mr
-      shortname: MR
-      longname: Mendelian Randomization
-    - key: se
-      shortname: SE
-      longname: socio-economic
-    - key: iv
-      shortname: IV
-      longname: instrumental variable
-    - key: ivs
-      shortname: IVs
-      longname: instrumental variables
-    - key: ctd
-      shortname: CTD
-      longname: Comparative Toxicogenomics Database
-    - key: cv
-      shortname: CV
-      longname: cross-validation
 ---
@@
 
-\textcolor{red}{Most of the following was **copy-pasted** from published papers.}
+Most of the following was **copy-pasted** from published papers.
 
@@
 
-The **aim** of the present study is to research the short-term effects of postnatal exposure to non-persistent \acr{edcs} on neurodevelopment and neurobehavior in childhood, and how the metabolome and the proteome might mediate these effects.
+The **aim** of the present study is to research the short-term effects of postnatal exposures to non-persistent \acr{edcs} on neurodevelopment and neurobehavior in childhood, and how the metabolome and the proteome might mediate these effects.
 We will strengthen the estimated effects by making use of the principles and criteria of triangulation [@LawlorTillingDaveySmith:2016].
@@
 
-\textcolor{red}{TODO}
+{{< include ../../../all_phd/descriptive_stats.qmd >}}
 
@@
 
+```{r}
+rq1.title <- "What are the short-term effects of childhood exposure to EDCs on the outcomes?"
+rq2.title <- "What are the short-term effects of childhood exposure to EDCs on metabolites and proteins?"
+rq3.title <- "What are the short-term effects of childhood *exposure* to metabolites and proteins on the outcomes?"
+```
+
 We specify the following \acr{npsem} to present the data generating process, 
@@
 and $(U_{W_{t}}, U_{A_{t}}, U_{Y_{t}})$ are the unmeasured factors contributing the confounders, exposures, and outcome, respectively. 
-The corresponding causal graph (without indexes for ease of notation) is given in @fig-causal-graph.
+The corresponding causal graph (without indexes for ease of notation) is shown in @fig-causal-graph.
 
@@
 #| label: fig-causal-graph
-#| fig.cap: Causal graph corresponding to structural equations.
+#| fig.cap: Causal graph corresponding to the structural equations.
 #| fig.height: 3.0
@@
 
-## Research question 1: What are the short-term effects of childhood exposure to \acr{edcs} on the outcomes?
+## Research question 1: `r rq1.title`
 
@@
 
-## Research question 2: What are the short-term effects of childhood exposure to \acr{edcs} on metabolites and proteins?
+## Research question 2: `r rq2.title`
 
 We will estimate the short-term effects of childhood exposure to non-persistent \acr{edcs} on metabolites and proteins measured in childhood.
-We will perform an MWAS analysis with the dependent variable $m_i$ being the metabolite or protein $i$, and the independent variable $x_j$ being the levels of \acr{edc} $j$.
+We will perform an XWAS analysis with the dependent variable $m_i$ being the metabolite or protein $i$, and the independent variable $x_j$ being the levels of \acr{edc} $j$.
 We will assume that all associations are confounded by the same set of confounders, independently of the \acr{edc} considered.
 
-## Research question 3: What are the short-term effects of childhood *exposure* to metabolites and proteins on the outcomes?
+## Research question 3: `r rq3.title`
 
 We will estimate the short-term effects of childhood *exposure* to metabolites and proteins on the outcomes of interest measured in childhood.
-We will perform an MWAS analysis with the dependent variable $y$ being the outcome, and the independent variable $m_i$ being the metabolite or protein $i$.
-We will assume that all associations are confounded by the same set of confounders, independently of the \acr{edc} considered.
+We will perform an XWAS analysis with the dependent variable $y$ being the outcome, and the independent variable $m_i$ being the metabolite or protein $i$.
+We will assume that all associations are confounded by the same set of confounders, independently of the biomarker considered.
 
@@
 
-* **No unmeasured confounding**: There are no unmeasured common causes of the exposure and the subsequent outcome, which we can formalize as $Y^d \perp \!\!\!
+* **No unmeasured confounding**.
+There are no unmeasured common causes of the exposure and the subsequent outcome, which we can formalize as $Y^d \perp \!\!\!
 \perp A|W$.
@@
 We cannot guarantee that this assumption holds, and therefore limit our interpretations to statistical associations rather than causal effects.
-* **Positivity**: If $(a, w)$ is within the support of $A, W$, then $(d(a,c), w)$ for additive shifts and $(d(a,k), w)$ for multiplicative shifts must also be within the support of $A,W$.
+* **Positivity**.
+If $(a, w)$ is within the support of $A, W$, then $(d(a,c), w)$ for additive shifts and $(d(a,k), w)$ for multiplicative shifts must also be within the support of $A,W$.
 In practice, this means that for any given time index and set of adjustment covariates, there is a positive probability of finding a subject with the same covariate values and a exposure level matching the shifted value.
@@
 * **Independence** of subjects. This assumption also implies **no interference**: the exposure level $a$ for a given subject does not affect the outcomes of the other subjects.
-* **Consistency**: If $A = a$ for any subject, then $Y(a) = Y$, and hence the full observed set of outcomes when $A^d = A$ is simply $Y(A^d) = Y$.
+* **Consistency**.
+If $A = a$ for any subject, then $Y(a) = Y$, and hence the full observed set of outcomes when $A^d = A$ is simply $Y(A^d) = Y$.
 This means that the counterfactual outcome for a subject with its observed exposure level is the observed outcome.
-* **Time-ordering**: The confounders $W$ precede the exposure $A$, which also precedes the outcome $Y$.
+* **Time-ordering**. The confounders $W$ precede the exposure $A$, which also precedes the outcome $Y$.
 
@@
 We refer to $P(\Lambda=1|a,w)$ as $\bar{\lambda}$.
-To implement estimation of $\bar{\lambda}$ and $\bar{Q}$ flexibly, we will use the ensemble machine learning algorithm \acr{sl}.
+To implement estimation of $\bar{\lambda}$ and $\bar{Q}$ flexibly, we will use the ensemble machine learning algorithm \acr{sl} [@vanderlaanSuperLearner2007; @naimiStackedGeneralizationIntroduction2018].
 In the following, we present the steps necessary to estimate $\psi_0 (A^d)$ and the associational parameter $\psi_0^{\Delta}$.
@@
 
-Using the empirical mean $\hat{Y} = \frac{1}{n} \sum_{i=1}^nY_i$, to estimate $\psi_0(A) = \mathbb{E}(Y)$, we obtain an estimate of the statistical association parameter $\psi_0^{\Delta} = \psi_0(A^d) = \psi_0(A)$ with
+Using the empirical mean $\hat{Y} = \frac{1}{n} \sum_{i=1}^nY_i$, to estimate $\psi_0(A) = \mathbb{E}(Y)$, we obtain an estimate of the statistical association parameter $\psi_0^{\Delta} = \psi_0(A^d) - \psi_0(A)$ with
 $$
@@
 
-Using `r I(cv.folds)`-fold \acr{cv}, \acr{sl} builds the best weighted combination of predictions from the algorithms listed in Table \@ref(tab:lmtp-specs).
+Using `r I(cv.folds)`-fold \acr{cv}, \acr{sl} builds the best weighted combination of predictions from the algorithms listed in \@ref(tab:lmtp-specs).
 In case of repeated measures of the exposures, an `ID` variable will be used to make sure that the `r I(cv.folds)`-fold \acr{cv} splits will keep observations from the same individuals in the same split.
@@
 - `outcome ~ chemical`
-  - Regression analysis.
+  - **Regression analysis**.
 We will use \acr{tmle} in combination with \acr{sl}, and a library of estimators including both simple parametric models (e.g., `glm`) and data-adaptive semi-parametric models.
@@
 We hypothesize the presence of residual confounding, especially due to genetic and parental \acr{se} factors, which would result in **exaggeration** of any true causal effect.
-  - Outcome negative control study.
+  - **Outcome negative control study**.
 We will try to identify outcomes in the same cohorts that would be affected by confounders that are relevant in the association of \acr{edcs} exposure to the clinical outcome of interest, but for which a biological/causal effect of the exposures is unlikely.
-  - Literature search.
+  - **Literature search**.
 For each of the `outcome ~ chemical` associations, we will perform a literature search to compare the obtained results.
 We will make use of PubMed, using as keywords the clinical outcome and the *common* name of each chemical, and the \acr{ctd} [@DavisWiegersJohnson:2023].
-  - Qualitative \acr{prs}. \textcolor{red}{TODO}
+  - **Qualitative \acr{prs}**. \textcolor{red}{TODO}
 - `omic ~ chemical`
-  - Regression analysis. As above.
-  - Cross-cohort comparison.
+  - **Regression analysis**. As above.
+  - **Cross-cohort comparison**.
 We will compare the regression models' results between subjects of different ethnic origins.
 We postulate that \acr{se} factors represent the main confounders for this association.
-Based on the associations been \acr{se} factors (e.g., maternal education) and the omic markers, we will compare the results of `omic ~ chemical` with the expected modified associations (e.g., weaker or stronger).
-  - Literature search. As above.
-  - Qualitative \acr{prs}. \textcolor{red}{TODO}
+Based on the associations between \acr{se} factors (e.g., maternal education) and the omic markers, we will compare the results of `omic ~ chemical` with the expected modified associations (e.g., weaker or stronger).
+  - **Literature search**. As above.
+  - **Qualitative \acr{prs}**. \textcolor{red}{TODO}
 - `outcome ~ omic`
-  - Regression analysis. As above.
-  - Cross-cohort comparison. As above.
-  - \acr{mr}.
+  - **Regression analysis**. As above.
+  - **Cross-cohort comparison**. As above.
+  - **\acr{mr}**.
 We will make use, ideally, of a weighted allele score of genetic variants known to be robustly associated with the omic markers as an \acr{IV}.
 We will use methods, including sensitivity analysis, to explore the possibility of bias due to: (i) weak instruments, and (ii) violation of the exclusion restriction criteria.
-Based on these, we will postulate whether the obtained results might be biased towards the null or not.
+Based on these, we will postulate whether the obtained results might be biased towards the **null** or not.
 By using \acr{mr}, we assume that this approach tests *exposure* to omic markers across the whole of childhood.
 
-## \acr{mr}: Assumptions
+## Mendelian Randomization: Assumptions
 
@@
 
+## Reliable Causal Inference
+
 {{< include ../../../all_phd/app-rel-ci.qmd >}}
 
----
+## Reliable Statistical Inference
 
@@
 
----
+## Replicability
 
@@
 
-## Checklist A: Reproducibility
+## Data Checking
+
+{{< include ../../../all_phd/app-data-checking.qmd >}}
+
+## Checklist A: Replicability
 
@@
 
-### Research question 1:...
+### Research question 1: `r rq1.title`
 
 ```{r}
-#| fig.cap: DAG for Research question 1.
-#| fig.height: 10.0
-#| fig.width: 18.0
-dag_rq1 <- dags$chem_to_out %>%
+#| fig-cap: DAG for research question 1.
+#| fig-align: center
+#| fig-width: 30
+#| fig-height: 40
+#| out-extra: "keepaspectratio=true"
+dag_rq <- dags$chem_to_out %>%
   dagitty::dagitty()
-dag_rq1 %>% ggdag::tidy_dagitty() %>%
-  ggdag::ggdag_status(text_size = 10, text_col = "black") +
-  ggdag::theme_dag() + ggdag::scale_adjusted() +
+dag_rq %>%
+  ggdag::tidy_dagitty() %>%
+  ggdag::ggdag_status(text_size = 12, text_col = "black") +
+  ggdag::scale_adjusted() +
+  ggdag::theme_dag() +
   ggplot2::theme(legend.text = ggplot2::element_text(size = 20))
+
+adj_sets <- dagitty::adjustmentSets(x = dag_rq, 
+                                    type = type_mas, effect = type_effect)
 ```
 
-The `r I(type_mas)` sufficient adjustment sets for estimating the `r I(type_effect)` effect of X on Y are:
+There are a total of `r length(adj_sets)` `r I(type_mas)` sufficient adjustment sets for estimating the `r I(type_effect)` effect. The first one is:
 
 ```{r}
-dagitty::adjustmentSets(x = dag_rq1, 
+adj_sets[[1]]
+```
+
+### Research question 2: `r rq2.title`
+
+```{r}
+#| fig-cap: DAG for research question 2.
+#| fig-align: center
+#| fig-width: 30
+#| fig-height: 40
+#| out-extra: "keepaspectratio=true"
+dag_rq <- dags$chem_to_out %>%
+  dagitty::dagitty()
+dag_rq %>%
+  ggdag::tidy_dagitty() %>%
+  ggdag::ggdag_status(text_size = 12, text_col = "black") +
+  ggdag::scale_adjusted() +
+  ggdag::theme_dag() +
+  ggplot2::theme(legend.text = ggplot2::element_text(size = 20))
+
+adj_sets <- dagitty::adjustmentSets(x = dag_rq, 
                                     type = type_mas, effect = type_effect)
+```
+
+There are a total of `r length(adj_sets)` `r I(type_mas)` sufficient adjustment sets for estimating the `r I(type_effect)` effect. The first one is:
+
+```{r}
+adj_sets[[1]]
+```
+
+### Research question 3: `r rq3.title`
+
+```{r}
+#| fig-cap: DAG for research question 3.
+#| fig-align: center
+#| fig-width: 30
+#| fig-height: 40
+#| out-extra: "keepaspectratio=true"
+dag_rq <- dags$chem_to_out %>%
+  dagitty::dagitty()
+dag_rq %>%
+  ggdag::tidy_dagitty() %>%
+  ggdag::ggdag_status(text_size = 12, text_col = "black") +
+  ggdag::scale_adjusted() +
+  ggdag::theme_dag() +
+  ggplot2::theme(legend.text = ggplot2::element_text(size = 20))
+
+adj_sets <- dagitty::adjustmentSets(x = dag_rq, 
+                                    type = type_mas, effect = type_effect)
+```
+
+There are a total of `r length(adj_sets)` `r I(type_mas)` sufficient adjustment sets for estimating the `r I(type_effect)` effect. The first one is:
+
+```{r}
+adj_sets[[1]]
 ```

### 2023-03-09 docs/causal_roadmap.qmd
Added unique file for acronyms and authors. Fixed problem of list of acronyms before TOC

@@
   - Neurobehavior
-author:
-  - name: "Lorenzo Fabbri"
-    number: 1
-    email: lorenzo.fabbri@isglobal.org
-    orcid: 0000-0003-3031-322X
-    attributes:
-      corresponding: true
-    affiliations:
-      - ref: isg
-      - ref: upf
-affiliations:
-  - id: isg
-    name: Barcelona Institute for Global Health (ISGlobal)
-    city: Barcelona
-    state: Spain
-  - id: upf
-    name: Pompeu Fabra University
-    city: Barcelona
-    state: Spain
+metadata-files:
+  - ../../../all_phd/authors.yml
 format:
@@
     number-sections: true
-    toc: true
+    toc: false
     margin-left: 15mm
@@
   - /home/lorenzo/R/x86_64-pc-linux-gnu-library/4.1/acronymsdown/parse-acronyms.lua
+  - ../../../all_phd/scholarly-metadata.lua
+  - ../../../all_phd/author-info-blocks.lua
 acronyms:
@@
   insert_links: false
+  insert_loa: false
   sorting: alphabetical
 ---
+
+```{=latex}
+\setcounter{tocdepth}{4}
+\tableofcontents
+```
+
+\printacronyms
 

### 2023-03-13 docs/causal_roadmap.qmd
Cleaned causal roadmap

@@
   echo: false
+  warning: false
+  error: false
+  message: false
 editor_options: 
@@
 
-\printacronyms
-
 ```{r}
-#| label: setup
-#| message: false
-library(tidyverse)
-```
-
-```{r}
-#| echo: false
 #| include: false
 
+here::i_am("causal_roadmap.qmd")
 source("../../../all_phd/info_phd.R")
@@
 methods <- describe_triangulation_methods()
+
+library(tidyverse)
 ```
@@
 ```{r}
-#| message: false
 source("../DAGs/dag_v1.R")
@@
 
-\newpage
+<!-- ####################################################################### -->
+# References {.unnumbered}
+<!-- ####################################################################### -->
 
-<!-- ####################################################################### -->
-# Bibliography
-<!-- ####################################################################### -->
+::: {#refs}
+:::

### 2023-03-14 docs/causal_roadmap.qmd
Added details for triangulation methods for different research questions.

@@
   - **Regression analysis**.
-We will use \acr{tmle} in combination with \acr{sl}, and a library of estimators including both simple parametric models (e.g., `glm`) and data-adaptive semi-parametric models.
-The models will be adjusted for relevant confounders, as identified with `dagitty`.
+We will use \acr{tmle} in combination with \acr{sl}, and a library of estimators including both simple parametric models (i.e., `glm`), and data-adaptive semi-parametric models.
+Specifically, we will make use of the `lmtp` R package, in combination with \acr{tmle} and \acr{sl}.
+We will also compare the results obtained with \acr{tmle} with those obtained with a \acr{sdre}.
+@tbl-lmtp-specs provides detailed information about the specifications used for `lmtp`.
+\textcolor{red}{TODO: specify the treatment and reference policies.} The models will be adjusted for relevant confounders, as identified with `dagitty`.
 We hypothesize the presence of residual confounding, especially due to genetic and parental \acr{se} factors, which would result in **exaggeration** of any true causal effect.
@@
 We will try to identify outcomes in the same cohorts that would be affected by confounders that are relevant in the association of \acr{edcs} exposure to the clinical outcome of interest, but for which a biological/causal effect of the exposures is unlikely.
-  - **Literature search**.
-For each of the `outcome ~ chemical` associations, we will perform a literature search to compare the obtained results.
-We will make use of PubMed, using as keywords the clinical outcome and the *common* name of each chemical, and the \acr{ctd} [@DavisWiegersJohnson:2023].
-  - **Qualitative \acr{prs}**. \textcolor{red}{TODO}
+We postulate that \acr{se} factors represent the main confounders for this association.
+We identified outcomes that are associated to \acr{se} factors, but that are not causally associated to the chemicals: having a car, and whether the child has their own room.
+\textcolor{red}{A possible issue is represented by the fact that these outcomes are measured on different scales.} We hypothesized that a higher \acr{se} position is positively associated to, e.g., having a car, and negatively associated to the chemicals' concentrations.
+This would result in a (confounded) **inverse** relationship.
+We will thus test the assumptions regarding the relation between the \acr{se} factors with each outcome, and check whether these associations are in the anticipated direction.
+  - **Literature search**. For each `outcome ~ chemical` association, we will perform a literature search to compare the obtained results. More precisely, we will adopt the following strategies:
+    - PubChem [@KimChenCheng:2023].
+We will search each chemical, and report the results from the sections *Associated Disorders and Diseases* and *Chemical-Disease Co-Occurrences in Literature*.
+We will report the total number of hits for each section, the number of hits relevant for our outcome(s), and the number of articles for each hit.
+    - The \acr{ctd} [@DavisWiegersJohnson:2023].
+We will search each chemical, and report the results from the section *Diseases*.
+We will specifically look for *curated* associations.
+We will report the total number of hits, the number of hits relevant for our outcome(s), and the detailed information for each of these.
 - `omic ~ chemical`
@@
   - **Literature search**. As above.
-  - **Qualitative \acr{prs}**. \textcolor{red}{TODO}
 - `outcome ~ omic`
   - **Regression analysis**. As above.
-  - **Cross-cohort comparison**. As above.
+  - **Cross-cohort comparison**.
+We postulate that \acr{se} factors represent the main confounders for this association.
+We will thus test the association between \acr{se} factors and the omics, separately for the two sub-populations (e.g., white British and Pakistani in BiB [@WrightSmallRaynor:2013]).
+If this association is smaller among Pakistani, if the association among white British were due to residual \acr{se} factor confounding, we would expect a **weaker** association among Pakistani.
   - **\acr{mr}**.
-We will make use, ideally, of a weighted allele score of genetic variants known to be robustly associated with the omic markers as an \acr{IV}.
+A simplified \acr{dag} is shown in @fig-dag-mr.
+We will make use, ideally, of a weighted allele score of genetic variants known to be robustly associated with the omic markers as an \acr{iv}.
 We will use methods, including sensitivity analysis, to explore the possibility of bias due to: (i) weak instruments, and (ii) violation of the exclusion restriction criteria.
@@
 By using \acr{mr}, we assume that this approach tests *exposure* to omic markers across the whole of childhood.
+Specifically, we will perform the following analyses:
+    - We will use the *standard* \acr{ivw} [@BurgessButterworthThompson:2013; @BowdenDelGrecoMMinelli:2019] method for primary \acr{mr} analyses.
+    - We will use MR-Egger [@BowdenDaveySmithBurgess:2015], weighted median [@BowdenSmithHaycock:2016], and MR-PRESSO [@VerbanckChenNeale:2018] for sensitivity analyses.
+
+    We will make use of \acr{2smr} [@DaveySmithEbrahim:2003; @PierceBurgess:2013; @DaveySmithHemani:2014].
+We will thus employ results from two different types of \acr{gwass}: a exposure \acr{gwas}, using The Metabolomics GWAS Server [@SuhreShinPetersen:2011; @ShinFaumanPetersen:2014], and a outcome \acr{gwas}, using the OpenGWAS project [@10.7554/eLife.34408; @ElsworthLyonAlexander:2020].
+The effects of the selected genetic variants will then be harmonized.
+
+```{r}
+#| fig-cap: Simplified DAG for MR analyses of `outcome ~ omic` associations.
+#| label: fig-dag-mr
+#| fig-width: 5
+#| fig-height: 4
+coords <- tibble::tribble(
+  ~name, ~x, ~y, 
+  "neuro", 1, 0, 
+  "metab", 0.5, 0, 
+  "SNP", 0, 0, 
+  "C", 0.75, 0.1
+)
+ggdag::dagify(
+  neuro ~ metab + C, 
+  metab ~ SNP + C, 
+  exposure = "metab", 
+  outcome = "neuro", 
+  coords = coords
+) %>%
+  ggdag::ggdag(., layout = "circle") +
+  ggdag::geom_dag_point(col = "gray") +
+  ggdag::geom_dag_edges() +
+  ggdag::geom_dag_text(col = "black") +
+  ggplot2::theme(axis.title = ggplot2::element_blank(), 
+                 axis.ticks = ggplot2::element_blank(), 
+                 axis.text = ggplot2::element_blank())
+```
 
@@
 
-## Checklist B1: `xxx` package specifications
+## Checklist B1: `lmtp` package specifications
 
@@
 |:-------------|:------------|:------------------|:------------|
-| ... | ... | ... | ... |
+| `learners_outcome` | `SL.mean` | N | |
+| `learners_outcome` | `SL.glm` | Y | |
+| `learners_outcome` | `SL.gam` | N | |
+| `learners_outcome` | `SL.hal9001` | N | |
+| `learners_outcome` | `SL.bart` | N | |
+| `learners_trt` | `SL.mean` | N | |
+| `learners_trt` | `SL.glm` | Y | |
+| `learners_trt` | `SL.gam` | N | |
+| `learners_trt` | `SL.hal9001` | N | |
+| `learners_trt` | `SL.bart` | N | |
+| `folds` | 10 | Y | The number of folds to be used for cross-fitting |
+| `.learners_outcome_folds` | 10 | Y | The number of \acr{cv} folds for `learners_outcome` |
+| `.learners_trt_folds` | 10 | Y | The number of \acr{cv} folds for `learners_trt` |
+
+: Specifications for the `lmtp` R package. {#tbl-lmtp-specs}
 

### 2023-03-14 code/dictionaries.R

@@ -1,1 +1,5 @@
 # Script containing dictionaries (e.g., exposures)
+
+################################################################################
+# Function that returns all necessary HELIX variable names
+################################################################################

### 2023-03-23 docs/causal_roadmap.qmd
Kept adding details to roadmap for triangulation. Started modifying files for tables and manuscript to use Quarto things

@@
 We will assume that all associations are confounded by the same set of confounders, independently of the \acr{edc} considered.
+We will further stratify the analyses by sex, to account for potential sex-specific effects of the chemicals.
 
@@
     - We will use MR-Egger [@BowdenDaveySmithBurgess:2015], weighted median [@BowdenSmithHaycock:2016], and MR-PRESSO [@VerbanckChenNeale:2018] for sensitivity analyses.
+For *secondary* sensitivity analyses, we will use: Steiger filtering, Cochran's Q-statistics (to check for heterogeneity), and the Egger intercept (to check for horizontal pleiotropy) \textcolor{red}{(citations needed)}.
 
@@
 The effects of the selected genetic variants will then be harmonized.
+We will fit simple regression models to examine whether the selected \acr{ivs} predict the relevant nodes of the \acr{dag} (i.e., exposures and outcomes), using the F-statistics.
+Furthermore, we will fit simple regression models of the *form* `confounder ~ SNP`, adjusted for the *top* genetic \acr{pcs}, to test whether the \acr{snps} are separately associated with the confounders.
+As done in previous works, we will standardized all the considered outcomes.
+The results (both of the main and the sensitivity analyses) will be presented by means of forest plots, for the *significant* associations.
+
+    In the case of \acr{1smr}, we will make use of the following criteria to select the \acr{snps} for both the exposures and the outcomes:
+      - We will favour *cis*-variants.
+      - We will *a priori* define a p-value threshold.
+      - We will perform \acr{ld} clumping of the genetic variants.
+      - We will *a priori* define a \acr{maf} threshold.
+    Similarly, the models will be adjusted for the *top* genetic \acr{pcs}.
 
@@
 
+\textcolor{red}{Is it possible to include models from the `brms` R package?}
+
 | **Argument** | **Setting** | **Default** (Y/N) | **Comment** |
@@
 ```{r}
-source("../DAGs/dag_v1.R")
-source("../code/dags.R")
+source("../DAGs/dag_v2.R")
 
-dags <- dag_v1()
+dags <- dag_v2()
 ```
@@
 #| out-extra: "keepaspectratio=true"
-dag_rq <- dags$chem_to_out %>%
-  dagitty::dagitty()
-dag_rq %>%
-  ggdag::tidy_dagitty() %>%
-  ggdag::ggdag_status(text_size = 12, text_col = "black") +
-  ggdag::scale_adjusted() +
-  ggdag::theme_dag() +
-  ggplot2::theme(legend.text = ggplot2::element_text(size = 20))
+dag_rq <- dags$chem_to_out
+
+myphd::visualize_dag(dag_rq)
+# dag_rq %>%
+#   ggdag::tidy_dagitty() %>%
+#   ggdag::node_status() %>%
+#   ggplot2::ggplot(ggplot2::aes(x, y, 
+#                                xend = xend, yend = yend, 
+#                                color = status)) +
+#   ggdag::geom_dag_edges() +
+#   ggdag::geom_dag_point() +
+#   geom_dag_text_repel() +
+#   ggokabeito::scale_color_okabe_ito(na.value = "grey90") +
+#   theme_dag() +
+#   ggplot2::coord_cartesian(clip = "off")
 
@@
 #| out-extra: "keepaspectratio=true"
-dag_rq <- dags$chem_to_out %>%
-  dagitty::dagitty()
-dag_rq %>%
-  ggdag::tidy_dagitty() %>%
-  ggdag::ggdag_status(text_size = 12, text_col = "black") +
-  ggdag::scale_adjusted() +
-  ggdag::theme_dag() +
-  ggplot2::theme(legend.text = ggplot2::element_text(size = 20))
-
-adj_sets <- dagitty::adjustmentSets(x = dag_rq, 
-                                    type = type_mas, effect = type_effect)
 ```
@@
 ```{r}
-adj_sets[[1]]
 ```

### 2023-03-23 output/paper/tables.qmd
Kept adding details to roadmap for triangulation. Started modifying files for tables and manuscript to use Quarto things

@@
 ---
-title: "Tables"
-output:
-  bookdown::pdf_document2
+title: "Paper 3 - Tables"
+format:
+  docx:
+    reference-doc: reference_document.docx
+  pdf:
+    pdf-engine: pdflatex
+    number-sections: true
+    toc: false
+    margin-left: 15mm
+    margin-right: 15mm
+    margin-bottom: 20mm
+    margin-top: 15mm
+    include-in-header:
+      - \usepackage{rotating}
+      - \usepackage{float}
+      - \usepackage{typearea}
+link-citations: true
+linkcolor: blue
+execute:
+  echo: false
+  warning: false
+  error: false
+  message: false
+editor_options: 
+  chunk_output_type: console
+project:
+  execute_dir: project
 ---
@@
 ```{r}
-knitr::opts_chunk$set(echo = FALSE, 
-                      warning = FALSE, message = FALSE, 
-                      cache = TRUE)
+#| include: false
+here::i_am("tables.qmd")
+source("../../../../all_phd/edcs.R")
+library(tidyverse)
 ```
@@
 <!-- ####################################################################### -->
-<!-- Study populations -->
+# Study populations
+<!-- ####################################################################### -->
 ```{r}
@@
 ```
-<!-- ####################################################################### -->
 
 <!-- ####################################################################### -->
-<!-- Exposures description -->
+# Exposures
+<!-- ####################################################################### -->
 ```{r}
-
+#| label: tbl-chem-info
+#| tbl-cap: Information about the non-persistent EDCs.
+edcs_info <- edcs_information() %>%
+  dplyr::select(-c(chem_id, smiles))
+gt::gt(edcs_info, groupname_col = "class") %>%
+  gt::cols_label(
+    full_name = "Compound", 
+    short_name = "Symbol", 
+    #smiles = "SMILES", 
+    pubchem_cid = "PubChem CID", 
+    ctd_id = "CTD ID", 
+    exposome_explorer_id = "Exposome Explore ID", 
+    parental_compound = "Parental compound"
+  ) %>%
+  gt::tab_style(
+    locations = gt::cells_column_labels(columns = everything()), 
+    style = list(
+      gt::cell_borders(sides = "bottom", weight = gt::px(3)), 
+      gt::cell_text(weight = "bold")
+    )
+  ) %>%
+  gt::tab_style(
+    locations = gt::cells_row_groups(group = everything()), 
+    style = list(gt::cell_text(weight = "bold"))
+  )
 ```
-<!-- ####################################################################### -->

### 2023-03-23 DAGs/dag_v2.R

@@ -1,0 +1,62 @@
+# DAGs based on research questions defined in the Causal Roadmap
+
+dag_v2 <- function() {
+  ##############################################################################
+  chem_to_out <- ggdag::dagify(
+    airPollution_child ~ cohort + familySEP + season_visit, 
+    airPollution_preg ~ cohort + maternalSEP_preg + paternalSEP_preg, 
+    breastfeeding ~ ethnicity_mother + maternalSEP_preg, 
+    bw ~ airPollution_preg + breastfeeding + cohort + ethnicity_child + ethnicity_mother + maternalAlcohol_preg + maternalDiet_preg + maternalIodine_preg + maternalIron_preg + maternalSEP_preg + maternalSmoking_preg + maternal_folicAcid_preg + otherChemicals_preg + paternalSEP_preg + paternalSmoking_preg + sex_child, 
+    chemical ~ chemicals_SNPs + child_diet + cohort + creatinine + familySEP + season_visit + time_lastMeal + water_child, 
+    chemicals_SNPs ~ ethnicity_child + ethnicity_mother, 
+    child_alcohol ~ age_child + ethnicity_child + familySEP + sex_child, 
+    child_depression ~ age_child + airPollution_child + airPollution_preg + breastfeeding + child_alcohol + child_diet + child_smoking + cohort + ethnicity_child + ethnicity_mother + familySEP + maternalAlcohol_preg + maternalIodine_preg + maternalIron_preg + maternalSmoking_preg + maternal_folicAcid_preg + otherChemicals_child + otherChemicals_preg + paternalSEP_preg + paternalSmoking_preg + sex_child, 
+    child_diet ~ age_child + cohort + ethnicity_mother + familySEP + sex_child, 
+    child_smoking ~ age_child + ethnicity_child + familySEP + sex_child, 
+    creatinine ~ age_child + child_diet + fatMass_child + sex_child + time_lastMeal, 
+    edu_child ~ age_child + cohort + ethnicity_mother + familySEP, 
+    envFactors_visit ~ season_visit, 
+    familySEP ~ ethnicity_mother, 
+    fatMass_child ~ age_child + airPollution_child + airPollution_preg + chemical + child_alcohol + child_diet + child_smoking + ethnicity_child + ethnicity_mother + familySEP + maternalAlcohol_preg + maternalDiet_preg + maternalSmoking_preg + otherChemicals_child + otherChemicals_preg + paternalSmoking_preg + sex_child, 
[…]

### 2023-03-24 docs/causal_roadmap.qmd
Switched to myphd R package for DAGs. Started working on DAG for RQ3

@@
     - We will use MR-Egger [@BowdenDaveySmithBurgess:2015], weighted median [@BowdenSmithHaycock:2016], and MR-PRESSO [@VerbanckChenNeale:2018] for sensitivity analyses.
-For *secondary* sensitivity analyses, we will use: Steiger filtering, Cochran's Q-statistics (to check for heterogeneity), and the Egger intercept (to check for horizontal pleiotropy) \textcolor{red}{(citations needed)}.
+For *secondary* sensitivity analyses, we will use: Steiger filtering and/or bi-directional \acr{mr}, Cochran's Q-statistics (to check for heterogeneity), and the Egger intercept (to check for horizontal pleiotropy [^1]) \textcolor{red}{(citations needed)}.
+
+    [^1]: In order to address the issue of horizontal pleiotropy, and to study the joint effects of multiple traits, another possibility is represented by the use of univariable \acr{mr} for screening purposes, followed by multivariable (with clumping for \acr{ld}) \acr{mr} [@DecinaHopkinsBowden:2023].
 
@@
     Similarly, the models will be adjusted for the *top* genetic \acr{pcs}.
+
+  From the metabolomics dataset, we will remove derived ratios and composite measures, and we will consider only the total lipids of each subclass [@DecinaHopkinsBowden:2023].
+Among the remaining metabolomic features, we will check their genetic correlation.
 
@@
 dag_rq <- dags$chem_to_out
-
 myphd::visualize_dag(dag_rq)
-# dag_rq %>%
-#   ggdag::tidy_dagitty() %>%
-#   ggdag::node_status() %>%
-#   ggplot2::ggplot(ggplot2::aes(x, y, 
-#                                xend = xend, yend = yend, 
-#                                color = status)) +
-#   ggdag::geom_dag_edges() +
-#   ggdag::geom_dag_point() +
-#   geom_dag_text_repel() +
-#   ggokabeito::scale_color_okabe_ito(na.value = "grey90") +
-#   theme_dag() +
-#   ggplot2::coord_cartesian(clip = "off")
-
 adj_sets <- dagitty::adjustmentSets(x = dag_rq, 

### 2023-03-24 DAGs/dag_v2.R

@@ -38,19 +38,19 @@ dag_v2 <- function() {
     water_preg ~ cohort + ethnicity_mother + maternalSEP_preg + paternalSEP_preg, 
     exposure = "chemical", 
     outcome = "intelligence"
-    # coords = list(
-    #   x = c(chemical = 0, intelligence = 1), 
-    #   y = c(chemical = 0, intelligence = 1)
-    # )
   )
   ##############################################################################
   
   ##############################################################################
-  chem_to_marker <- ''
+  chem_to_marker <- ggdag::dagify(
+    
+  )
   ##############################################################################
   
   ##############################################################################
-  marker_to_out <- ''
+  marker_to_out <- ggdag::dagify(
+    
+  )
   ##############################################################################
   
   return(list(

### 2023-03-27 docs/causal_roadmap.qmd
Added DAG for bimoarker to outcome

@@
 
+<!-- ####################################################################### -->
 ## Data Checking
+<!-- ####################################################################### -->
 
@@
 
+<!-- ####################################################################### -->
 ## Checklist A: Replicability
+<!-- ####################################################################### -->
 
@@
 
+<!-- ####################################################################### -->
+## Gene-Environment Interactions
+<!-- ####################################################################### -->
+
+Genetic polymorphisms relevant for the **chemicals**:
+
+* \acr{opps}:
+  * The \acr{snps} of the *PON1* (paraoxonase 1) gene, known to influence the ability to metabolize and detoxify \acr{opps}: `rs854560` and `rs662`. Reference: https://doi.org/10.1289/ehp.1408976.
+  * The \acr{snps} of the genes coding for \acr{opps}-metabolizing enzymes: `rs1803274`, `rs1799807`, `rs662`, `rs854560`, `rs705379`, `rs4244285`, `rs3892097`, and `rs2177180`.
+Reference: https://doi.org/10.1016/j.chemosphere.2015.07.079.
+* Phenols:
+  * The \acr{snp} of the *UGT2B15* gene: `rs1902023`. Reference: https://doi.org/10.1016/j.envint.2020.106172.
+* Phthalates:
+  *  The \acr{snps} of the cytochrome P450 (*CYPs*) and UDP-glucuronosyltransferase (*UGTs*) genes: `rs1057910`, `rs1799853`, `rs4244285`, `rs12248560`, `rs3892097`, `rs1902023`, and `rs11692021`.
+Reference: https://doi.org/10.1016/j.envint.2021.107046.
+  * The \acr{snps} of the *PPAR*$\alpha$ gene: `rs75525202`, `rs4823902`, `rs4253690`, `rs12330015`, `rs5766698`, `rs115250492`, and `rs1800246.` Reference: https://doi.org/10.3389/fgene.2022.855544.
+Not sure whether these \acr{snps} should be considered in the list below.
+
+Genetic polymorphisms relevant for the **outcomes**:
+
+<!-- ####################################################################### -->
 ## DAGs
+<!-- ####################################################################### -->
 
@@
 
+<!-- ####################################################################### -->
 ### Research question 1: `r rq1.title`
+<!-- ####################################################################### -->
 
@@
 
+<!-- ####################################################################### -->
 ### Research question 2: `r rq2.title`
+<!-- ####################################################################### -->
 
@@
 
+<!-- ####################################################################### -->
 ### Research question 3: `r rq3.title`
+<!-- ####################################################################### -->
 

### 2023-03-27 DAGs/dag_v2.R

@@ -7,7 +7,7 @@ dag_v2 <- function() {
     airPollution_preg ~ cohort + maternalSEP_preg + paternalSEP_preg, 
     breastfeeding ~ ethnicity_mother + maternalSEP_preg, 
     bw ~ airPollution_preg + breastfeeding + cohort + ethnicity_child + ethnicity_mother + maternalAlcohol_preg + maternalDiet_preg + maternalIodine_preg + maternalIron_preg + maternalSEP_preg + maternalSmoking_preg + maternal_folicAcid_preg + otherChemicals_preg + paternalSEP_preg + paternalSmoking_preg + sex_child, 
-    chemical ~ chemicals_SNPs + child_diet + cohort + creatinine + familySEP + season_visit + time_lastMeal + water_child, 
+    chemical ~ chemicals_SNPs + child_diet + cohort + familySEP + season_visit + time_lastMeal + water_child, 
     chemicals_SNPs ~ ethnicity_child + ethnicity_mother, 
     child_alcohol ~ age_child + ethnicity_child + familySEP + sex_child, 
     child_depression ~ age_child + airPollution_child + airPollution_preg + breastfeeding + child_alcohol + child_diet + child_smoking + cohort + ethnicity_child + ethnicity_mother + familySEP + maternalAlcohol_preg + maternalIodine_preg + maternalIron_preg + maternalSmoking_preg + maternal_folicAcid_preg + otherChemicals_child + otherChemicals_preg + paternalSEP_preg + paternalSmoking_preg + sex_child, 
@@ -49,7 +49,42 @@ dag_v2 <- function() {
   
   ##############################################################################
   marker_to_out <- ggdag::dagify(
-    
+    airPollution_child ~ cohort + familySEP + season_visit
+    airPollution_preg ~ cohort + maternalSEP_preg + paternalSEP_preg
+    biomarker ~ age_child + airPollution_child + chemical + child_alcohol + child_diet + child_smoking + cohort + creatinine + ethnicity_child + ethnicity_mother + familySEP + fatMass_child + otherChemicals_child + season_visit + sex_child + time_lastMeal
+    breastfeeding ~ ethnicity_mother + maternalSEP_preg
[…]

### 2023-04-06 docs/causal_roadmap.qmd
Started working on dictionaries for variables from DAGs to HELIX data and labels variables from HELIX codebooks. Added info to causal roadmap, including summary diagram of research questions and trinagulation

@@
 We will consider both additive and multiplicative shifts of the exposures' distribution.
+
+@fig-diagram summarizes the research questions and adopted methods to answer them.
+In case of regression analysis using \acr{mtps}, the causal quantity of interest, $\phi^{\Delta, i} = \mathbb{E}[Y^{d_i}] - \mathbb{E}[Y]$, will be *adapted* to the research question (i.e., different policies based on the type of exposure).
+
+```{mermaid}
+%%| fig-cap: Diagram summarizing research questions and methods.
+%%| label: fig-diagram
+flowchart LR
+  paper[Paper 3] --> rq1("RQ1
+                          neuro~chemical")
+  paper          --> rq2("RQ2
+                          met~chemical")
+  paper          --> rq3("RQ3
+                          neuro~met")
+  rq1            --> rqReg("Observational study.
+                            Method: Regression with MTPs.
+                            Causal estimand: phi.
+                            Uncertainty quantification: bootstrapping.")
+  rq1            --> rq1outNeg("Outcome negative control.
+                                Method: Regression with GLMs/GAMs.
+                                Uncertainty quantification: bootstrapping.")
+  rq1            --> rqLit("Literature search")
+
+  rq2            --> rqReg
+  rq2            --> rqCrossCohort("Cross-cohort comparison.
+                                    Method: Regression with GLMs/GAMs.
+                                    Uncertainty quantification: bootstrapping.")
+  rq2            --> rqLit
+
+  rq3            --> rqReg
+  rq3            --> rqCrossCohort
+  rq3            --> rq3mr("Mendelian Randomization.
+                            Causal estimand: .")
+```
 
@@
 
-Using `r I(cv.folds)`-fold \acr{cv}, \acr{sl} builds the best weighted combination of predictions from the algorithms listed in \@ref(tab:lmtp-specs).
+Using `r I(cv.folds)`-fold \acr{cv}, \acr{sl} builds the best weighted combination of predictions from the algorithms listed in @tbl-lmtp-specs.
 In case of repeated measures of the exposures, an `ID` variable will be used to make sure that the `r I(cv.folds)`-fold \acr{cv} splits will keep observations from the same individuals in the same split.
@@
 Reference: https://doi.org/10.1016/j.envint.2021.107046.
-  * The \acr{snps} of the *PPAR*$\alpha$ gene: `rs75525202`, `rs4823902`, `rs4253690`, `rs12330015`, `rs5766698`, `rs115250492`, and `rs1800246.` Reference: https://doi.org/10.3389/fgene.2022.855544.
+  * The \acr{snps} of the *PPAR*$\alpha$ gene: `rs75525202`, `rs4823902`, `rs4253690`, `rs12330015`, `rs5766698`, `rs115250492`, and `rs1800246`.
+Reference: https://doi.org/10.3389/fgene.2022.855544.
 Not sure whether these \acr{snps} should be considered in the list below.
@@
 #| out-extra: "keepaspectratio=true"
+dag_rq <- dags$marker_to_out
+myphd::visualize_dag(dag_rq)
+adj_sets <- dagitty::adjustmentSets(x = dag_rq, 
+                                    type = type_mas, effect = type_effect)
 ```
@@
 ```{r}
+adj_sets[[1]]
 ```

### 2023-04-06 output/paper/figures.qmd
Started working on dictionaries for variables from DAGs to HELIX data and labels variables from HELIX codebooks. Added info to causal roadmap, including summary diagram of research questions and trinagulation

@@
 ---
-title: "Figures"
-output:
-  bookdown::pdf_document2
+title: "Paper 3 - Figures"
+format:
+  pdf:
+    pdf-engine: pdflatex
+    toc: false
+    include-in-header:
+      - \usepackage{rotating}
+      - \usepackage{float}
+      - \usepackage{typearea}
+link-citations: true
+linkcolor: blue
+execute:
+  echo: false
+  warning: false
+  error: false
+  message: false
+editor_options: 
+  chunk_output_type: console
+project:
+  execute_dir: project
 ---
@@
 ```{r}
-knitr::opts_chunk$set(echo = FALSE, 
-                      warning = FALSE, message = FALSE, 
-                      cache = TRUE)
+#| include: false
+here::i_am("figures.qmd")
+source("../../code/ctd.R")
+library(tidyverse)
 ```
@@
 <!-- ####################################################################### -->
-<!-- Workflow -->
+# CTD Results
+<!-- ####################################################################### -->
 ```{r}
+#| label: fig-ctd-res
+#| fig.cap: Barplot CTD.
+#| fig.height: 10
+#| fig.width: 10
+path <- "../../results/files/CTD_chem_diseases_1680702221550_05042023.csv"
+ctd <- loadCTD(path = path, 
+               filter_evidence = TRUE)
+plotCTD(ctd)
 ```
-<!-- ####################################################################### -->

### 2023-04-06 output/paper/manuscript.qmd
Started working on dictionaries for variables from DAGs to HELIX data and labels variables from HELIX codebooks. Added info to causal roadmap, including summary diagram of research questions and trinagulation

@@
 ---
-title: "Paper X"
-keywords:
-  - Causal Inference
-  - Epidemiology
-author:
-  - name: "Lorenzo Fabbri"
-    number: 1
-    email: lorenzo.fabbri@isglobal.org
-    orcid: 0000-0003-3031-322X
-    attributes:
-      corresponding: true
-    affiliations:
-      - ref: isg
-      - ref: upf
-affiliations:
-  - id: isg
-    name: ISGlobal
-    city: Barcelona
-    state: Spain
-  - id: upf
-    name: UPF
-    city: Barcelona
-    state: Spain
+metadata-files:
+  - ../../../../all_phd/authors.yml
 format:
@@
     pdf-engine: pdflatex
+    number-sections: true
     toc: false
-  # docx:
-#csl: journal.csl
-bibliography: bibliography.bib
+    include-in-header:
+      - \usepackage{rotating}
+      - \usepackage{float}
+      - \usepackage{typearea}
+    pandoc_args:
+      - !expr acronymsdown::add_filter()
+bibliography: ../../../../PhD.bib
 link-citations: true
 linkcolor: blue
-header-includes:
-  - \usepackage{setspace}\doublespacing
-  - \usepackage[switch, pagewise, running]{lineno}
-  - \linenumbers
-  - \renewcommand\linenumberfont{\normalfont\small}
-  - \usepackage{rotating}
-  - \usepackage{float}
+execute:
+  echo: false
+  warning: false
+  error: false
+  message: false
+editor_options: 
+  chunk_output_type: console
+filters:
+  - /home/lorenzo/R/x86_64-pc-linux-gnu-library/4.1/acronymsdown/parse-acronyms.lua
+  - ../../../../all_phd/scholarly-metadata.lua
+  - ../../../../all_phd/author-info-blocks.lua
+acronyms:
+  fromfile: ../../../../all_phd/acronyms.yml
+  insert_links: false
+  insert_loa: false
+  sorting: alphabetical
 abstract: |
- \singlespacing
  **Background**
@@
 
-\twocolumn
-\doublespacing
-\pagewiselinenumbers
+```{r}
+#| include: false
+
+here::i_am("manuscript.qmd")
+source("../../../../all_phd/info_phd.R")
+infos <- info_phd()
+```
+
+---
+title: "`r infos$paper3$title`"
+---
 
@@
 \newpage
-\onecolumn
-\singlespacing
 

### 2023-04-06 output/paper/tables.qmd
Started working on dictionaries for variables from DAGs to HELIX data and labels variables from HELIX codebooks. Added info to causal roadmap, including summary diagram of research questions and trinagulation

@@
     reference-doc: reference_document.docx
-  pdf:
-    pdf-engine: pdflatex
-    number-sections: true
-    toc: false
-    margin-left: 15mm
-    margin-right: 15mm
-    margin-bottom: 20mm
-    margin-top: 15mm
-    include-in-header:
-      - \usepackage{rotating}
-      - \usepackage{float}
-      - \usepackage{typearea}
 link-citations: true

### 2023-04-06 code/ctd.R

@@ -1,0 +1,46 @@
+# Functions to load and analyse data 
+# from the Comparative Toxicogenomics Database (CTD).
+
+loadCTD <- function(path, filter_evidence) {
+  #path <- "results/files/CTD_chem_diseases_1680702221550_05042023.csv"
+  ctd <- read.csv(path) |>
+    tibble::as_tibble()
+  
+  if (filter_evidence == TRUE) {
+    ctd <- ctd |>
+      dplyr::filter(DirectEvidence %in% c("marker/mechanism", 
+                                          "therapeutic", 
+                                          "marker/mechanism|therapeutic"))
+  }
+  
+  ctd <- ctd |>
+    tidyr::separate_longer_delim(DiseaseCategories, 
+                                 delim = "|")
+  ctd$DiseaseCategories <- factor(ctd$DiseaseCategories, 
+                                  levels = sort(unique(ctd$DiseaseCategories), 
+                                                decreasing = TRUE))
+  
+  return(ctd)
+}
+
+describeCTD <- function(dat) {
+  
+}
+
+plotCTD <- function(dat) {
+  plt <- dat |>
+    dplyr::filter(DiseaseCategories != "") |>
+    dplyr::group_by(DiseaseCategories) |>
+    ggplot2::ggplot(mapping = ggplot2::aes(
+      x = DiseaseCategories, 
+      fill = factor(ChemicalName)
+    )) +
+    ggplot2::geom_bar() +
+    ggplot2::coord_flip() +
+    ggplot2::theme_minimal() +
+    ggplot2::labs(x = "Disease categories") +
+    colorspace::scale_fill_discrete_qualitative(name = "Chemicals", 
+                                                palette = "Dynamic")
+  
+  return(plt)
+}

### 2023-04-06 code/dags.R

@@ -1,1 +1,0 @@
-# Code related to DAGs: visualization, testing of dependencies

### 2023-04-06 code/dictionaries.R

@@ -1,5 +1,72 @@
 # Script containing dictionaries (e.g., exposures)
 
 ################################################################################
-# Function that returns all necessary HELIX variable names
+# Labels for HELIX variables
 ################################################################################
+process_one_sheet <- function(sheet_id) {
+  url <- "https://docs.google.com/spreadsheets/d/1aETxd2bzjhkbOGDLRgEPVjaWu1iyOt0dBGNV2QHTEv4/edit?usp=sharing"
+  google_sheet <- googlesheets4::read_sheet(ss = url, sheet = sheet_id) |>
+    dplyr::rename(var_name = `Variable name`, 
+                  label = Description) |>
+    dplyr::select(var_name, label)
+  
+  return(google_sheet)
+}
+
+dict_vars_labels <- function() {
+  vars_rq1 <- process_one_sheet(sheet_id = "RQ1")
+  vars_rq2 <- process_one_sheet(sheet_id = "RQ2")
+  vars_rq3 <- process_one_sheet(sheet_id = "RQ3")
+  
+  return(list(
+    vars_rq1 = vars_rq1, 
+    vars_rq2 = vars_rq2, 
+    vars_rq3 = vars_rq3
+  ))
+}
+
+################################################################################
+# Mapping between DAGs nodes and HELIX data
+################################################################################
+dict_mapping_vars <- function() {
+  vars_dag_rq1 <- list(
+    age_child = c("hs_age_years"), 
+    airPollution_child = c("hs_no2_wk_hs_t", 
+                           "hs_nox_wk_hs_t", 
+                           "hs_pm10_wk_hs_t", 
+                           "hs_pm25_wk_hs_t"), 
+    airPollution_preg = c(), 
+    breastfeeding = c(), 
+    bw = c(), 
+    child_alcohol = c(), 
+    child_diet = c(), 
+    child_smoking = c(), 
+    edu_child = c(), 
+    familySEP = c(), 
+    gestational_age = c(), 
+    intelligence_SNPs = c(), 
+    maternalAlcohol_preg = c(), 
+    maternalDiet_preg = c(), 
+    maternalIodine_preg = c(), 
+    maternalIron_preg = c(), 
+    maternalSEP_preg = c(), 
+    maternalSmoking_preg = c(), 
+    maternal_folicAcid_preg = c(), 
+    neuropsychologicalDiagnosis_child = c(), 
[…]

### 2023-04-12 docs/causal_roadmap.qmd
Removed some nodes from DAGs. Fixed some things in roadmap (style mainly). Added more HELIX variables to dictionary mapping.

@@
 
-- \acr{cpm}, for assessing non-verbal intelligence.
-- Computerised n-back test, for assessing working memory.
-- \acr{ant}.
-- \acr{cbcl}, for assessing behavioural and emotional problems.
+- \acr{cpm}, for assessing non-verbal intelligence [@RavenCourtRaven:1998]. It includes the following outcomes: sum of the right answers, response time in minutes.
+- Computerised n-back test, for assessing working memory [@Kirchner:1958]. The recommended outcome is $d^{\prime}$.
+- \acr{ant}, to provide a measure of the efficiency of three different functions of attention [@RuedaFanMcCandliss:2004].
+The recommended outcome is \acr{hitrtse}, a measure of response speed consistency.
+- \acr{cbcl}, for assessing behavioural and emotional problems [@Tm:2001].
+The recommended outcomes are: internalizing scales-emotional problems, externalizing scales-behavioral problems, and total problems.
 
@@
 * \acr{opps}:
-  * The \acr{snps} of the *PON1* (paraoxonase 1) gene, known to influence the ability to metabolize and detoxify \acr{opps}: `rs854560` and `rs662`. Reference: https://doi.org/10.1289/ehp.1408976.
+  * The \acr{snps} of the *PON1* (paraoxonase 1) gene, known to influence the ability to metabolize and detoxify \acr{opps}: `rs854560` and `rs662`. [Reference](https://doi.org/10.1289/ehp.1408976).
   * The \acr{snps} of the genes coding for \acr{opps}-metabolizing enzymes: `rs1803274`, `rs1799807`, `rs662`, `rs854560`, `rs705379`, `rs4244285`, `rs3892097`, and `rs2177180`.
-Reference: https://doi.org/10.1016/j.chemosphere.2015.07.079.
+[Reference](https://doi.org/10.1016/j.chemosphere.2015.07.079).
 * Phenols:
-  * The \acr{snp} of the *UGT2B15* gene: `rs1902023`. Reference: https://doi.org/10.1016/j.envint.2020.106172.
+  * The \acr{snp} of the *UGT2B15* gene: `rs1902023`. [Reference](https://doi.org/10.1016/j.envint.2020.106172).
 * Phthalates:
   *  The \acr{snps} of the cytochrome P450 (*CYPs*) and UDP-glucuronosyltransferase (*UGTs*) genes: `rs1057910`, `rs1799853`, `rs4244285`, `rs12248560`, `rs3892097`, `rs1902023`, and `rs11692021`.
-Reference: https://doi.org/10.1016/j.envint.2021.107046.
+[Reference](https://doi.org/10.1016/j.envint.2021.107046).
   * The \acr{snps} of the *PPAR*$\alpha$ gene: `rs75525202`, `rs4823902`, `rs4253690`, `rs12330015`, `rs5766698`, `rs115250492`, and `rs1800246`.
-Reference: https://doi.org/10.3389/fgene.2022.855544.
+[Reference](https://doi.org/10.3389/fgene.2022.855544).
 Not sure whether these \acr{snps} should be considered in the list below.

### 2023-04-12 code/dictionaries.R

@@ -26,7 +26,6 @@ process_one_sheet <- function(sheet_id) {
     ))
   
   # Some variables from the codebooks have the wrong type
-  google_sheet[google_sheet$var_name == "e3_alcpreg_g", "type"] <- "integer"
   google_sheet[google_sheet$var_name == "hs_tob", "type"] <- "categorical"
   google_sheet[google_sheet$var_name == "e3_edum", "type"] <- "categorical"
   google_sheet[google_sheet$var_name == "e3_eduf", "type"] <- "categorical"
@@ -41,6 +40,7 @@ process_one_sheet <- function(sheet_id) {
   google_sheet[google_sheet$var_name == "e3_bw", "type"] <- "numerical"
   google_sheet[google_sheet$var_name == "e3_bhc", "type"] <- "numerical"
   google_sheet[google_sheet$var_name == "e3_bl", "type"] <- "numerical"
+  google_sheet[google_sheet$var_name == "e3_alcpreg_g", "type"] <- "integer"
   google_sheet[google_sheet$var_name == "hs_total_fish", "type"] <- "integer"
   google_sheet[google_sheet$var_name == "hs_total_veg", "type"] <- "integer"
   google_sheet[google_sheet$var_name == "hs_total_fruits", "type"] <- "integer"
@@ -228,17 +228,23 @@ dict_mapping_vars <- function() {
                            "hs_nox_wk_hs_t", 
                            "hs_pm10_wk_hs_t", 
                            "hs_pm25_wk_hs_t"), 
-    ############################################################################
-    airPollution_preg = c("NULL"), 
-    ############################################################################
+    airPollution_preg = c("h_no2_ratio_t1", 
+                          "h_no2_ratio_t2", 
+                          "h_no2_ratio_t3", 
+                          "h_nox_ratio_t1", 
+                          "h_nox_ratio_t2", 
+                          "h_nox_ratio_t3", 
+                          "h_pm10_ratio_t1", 
+                          "h_pm10_ratio_t2", 
+                          "h_pm10_ratio_t3", 
+                          "h_pm25_ratio_t1", 
+                          "h_pm25_ratio_t2", 
+                          "h_pm25_ratio_t3"), 
     breastfeeding = c("hs_bf", 
                       "hs_bfdur"), 
     bw = c("e3_bw", 
            "e3_bhc", 
            "e3_bl"), 
-    ############################################################################
-    child_alcohol = c("NULL"), 
[…]

### 2023-04-12 docs/causal_roadmap.qmd
Started writing skeletons for functions for RQs.

@@
 
-Table with random seed, names, description, and version numbers of all software packages.
+Table with random seed, names, version numbers, and description of the main software packages.
 
@@
 | `R` | `r I(paste0(R.version$major, ".", R.version$minor))` | Statistical programming environment |
+| `tlverse` | `r packageVersion("tlverse")` | Targeted Learning software ecosystem in R |
+| `lmtp` | `r packageVersion("lmtp")` | Non-parametric Causal Effects of Feasible Interventions Based on Modified Treatment Policies |
+| `sl3` | `r packageVersion("sl3")` | Super Learner ensemble machine learning |
+| `labelled` | `r packageVersion("labelled")` | Manipulate metadata |
+| `dlookr` | `r packageVersion("dlookr")` | Diagnose, explore, and transform data |
+| `Hmisc` | `r packageVersion("Hmisc")` | Harrell Miscellaneous |
+| `dagitty` | `r packageVersion("dagitty")` | Graphical Analysis of Structural Causal Models |
+| `qreport` | `r packageVersion("qreport")` | Facilitating the composition of Quarto reports |
 | ... | ... | ... |

### 2023-04-12 code/dictionaries.R

@@ -1,6 +1,43 @@
 # Script containing dictionaries (e.g., exposures)
 
 ################################################################################
+# Parameters for running analyses
+################################################################################
+params_analyses <- function() {
+  # Common parameters
+  learners_outcome <- c("SL.mean", "SL.glm", 
+                        "SL.gam", "SL.hal9001", "SL.bart")
+  learners_exposure <- learners_outcome
+  estimator <- "tmle"
+  folds <- 10
+  folds_outcome <- 10
+  folds_exposure <- 10
+  density_ratio_trim <- 0.995
+  markov_assumption <- Inf
+  
+  rq1 <- list(
+    learners_outcome = learners_outcome, 
+    learners_trt = learners_exposure, 
+    estimator = estimator, 
+    folds = folds, 
+    .learners_outcome_folds = folds_outcome, 
+    .learners_trt_folds = folds_exposure, 
+    .trim = density_ratio_trim, 
+    k = markov_assumption
+  ) # End dictionary parameters RQ1
+  rq2 <- rq1
+  rq3 <- rq1
+  
+  ret <- list(
+    rq1 = rq1, 
+    rq2 = rq2, 
+    rq3 = rq3
+  )
+  
+  return(ret)
+}
+
+################################################################################
 # Labels for HELIX variables
 ################################################################################
 process_one_sheet <- function(sheet_id) {

### 2023-04-17 output/presentations/cep_internal_230419/presentation.qmd
Kept working on all data requests. Started writing presentation for internal CEP meeting.

---
title: "CEP Group Meeting"
date: 2023-04-19
author: Lorenzo Fabbri
format:
  revealjs:
    footer: "Lorenzo Fabbri"
    toc: true
    toc_float: true
    toc-depth: 2
    toc-location: body
link-citations: true
linkcolor: blue
execute:
  echo: false
  warning: false
  error: false
  message: false
editor_options: 
  chunk_output_type: console
---

# Overview

## The Research Questions

# Triangulation

# Reliability and Replicability

### 2023-04-19 output/presentations/cep_internal_230419/presentation.qmd
Almost finished full data request for HELIX

@@
 author: Lorenzo Fabbri
+institute: ISGlobal
 format:
   revealjs:
-    footer: "Lorenzo Fabbri"
-    toc: true
+    toc: false
     toc_float: true
@@
     toc-location: body
+    embed-resources: true
+    width: 1600
+    height: 800
+    css: style.css
+    scrollable: false
+    smaller: true
+bibliography: ../../../../../PhD.bib
 link-citations: true
@@
 
-# Overview
+```{r}
+#| include: false
+
+here::i_am("presentation.qmd")
+source("../../../../../all_phd/info_phd.R")
+source("../../../../../all_phd/triangulation_methods.R")
+infos <- info_phd()
+methods <- describe_triangulation_methods()
+```
+
+## Overview
+
+::: {.v-center-container}
+- **Aim**: to research the *short-term* effects of postnatal exposures to non-persistent EDCs on neurodevelopment and neurobehavior in childhood, and how the metabolome and the proteome might *mediate* these effects.
+- **How**: by making use of the principles and criteria of triangulation [@LawlorTillingDaveySmith:2016].
+We will primarily rely on the use of MTPs [@munozPopulationInterventionCausal2012; @haneuseEstimationEffectInterventions2013; @diazNonparametricCausalEffects2021] in combination with TMLE [@vanderLaanBenkeserSofrygin:2018].
+- **Population**: the HELIX sub-cohort, consisting of $\text{N}=1200$ mother-child pairs.
+- **Exposures**: non-persistent EDCs (phenols, phthalates, and organophosphate compounds), measured in childhood in a pool of two urine samples.
+- **Outcomes**
+  - Raven’s Coloured Progressive Matrices, for assessing non-verbal intelligence.
+  - Computerised n-back test, for assessing working memory.
+  - Attention Network Test, to provide a measure of the efficiency of three different functions of attention.
+  - Child Behavior Checklist, for assessing behavioural and emotional problems.
+- **Mediators**: serum and urine metabolome, plasma proteome.
+:::
 
@@
 
-# Triangulation
+:::: {.columns style='display: flex !important;'}
 
-# Reliability and Replicability
+::: {.column style='display: flex; justify-content: center; align-items: center;'}
+```{mermaid}
+%%| fig-cap: Simplified DAG.
+%%| label: fig-diag-dag
+%%| fig-align: center
+flowchart LR
+  exposure(chemical) --> mediator(metabolite)
+  mediator --> outcome(neurodevelopment)
+  exposure --> outcome
+```
+:::
+
+::: {.column style='justify-content: center; align-items: center;'}
+- *What are the short-term effects of childhood exposure to EDCs on the outcomes?*
+
+A ExWAS kind of analysis with the dependent variable $y$ being the outcome, and the independent variable $x_i$ being the levels of EDC $i$.
+
+- *What are the short-term effects of childhood exposure to EDCs on metabolites and proteins?*
+
+A XWAS analysis with the dependent variable $m_i$ being the metabolite or protein $i$, and the independent variable $x_j$ being the levels of EDC $j$.
+
+- *What are the short-term effects of childhood exposure to metabolites and proteins on the outcomes?*
+
+A XWAS analysis with the dependent variable $y$ being the outcome, and the independent variable $m_i$ being the metabolite or protein $i$.
+:::
+
+::::
+
+## Triangulation
+
+:::: {.columns style='display: flex !important;'}
+
+::: {.column style='display: flex; justify-content: center; align-items: center;'}
+```{mermaid}
+%%| fig-cap: Diagram summarizing research questions and methods.
+%%| label: fig-diagram
+flowchart LR
+  paper[Paper 3] --> rq1("RQ1
+                          neuro~chemical")
+  paper          --> rq2("RQ2
+                          met~chemical")
+  paper          --> rq3("RQ3
+                          neuro~met")
+  rq1            --> rqReg("Observational study.
+                            Method: Regression with MTPs.
+                            Causal estimand: phi.
+                            Uncertainty quantification: bootstrapping.")
+  rq1            --> rq1outNeg("Outcome negative control.
+                                Method: Regression with GLMs/GAMs.
+                                Uncertainty quantification: bootstrapping.")
+  rq1            --> rqLit("Literature search")
+
+  rq2            --> rqReg
+  rq2            --> rqCrossCohort("Cross-cohort comparison.
+                                    Method: Regression with GLMs/GAMs.
+                                    Uncertainty quantification: bootstrapping.")
+  rq2            --> rqLit
+
+  rq3            --> rqReg
+  rq3            --> rqCrossCohort
+  rq3            --> rq3mr("Mendelian Randomization.
+                            Causal estimand: .")
+```
+:::
+
+::: {.column .r-fit-text}
+- _Chemical → outcome_
+  - Multivariate analysis in observational data: application of multivariable regression to observational data ($\phi^{\Delta, i} = \mathbb{E}[Y^{d_i}] - \mathbb{E}[Y]$).
+  - Outcome negative control study: aims to reproduce the same conditions as the real study, but using a different outcome not plausibly causally related to the exposure.
+  - Literature search: aims to validate the obtained results with data from the published scientific literature.
+- _Chemical → omic_
+  - Multivariate analysis in observational data
+  - Cross-cohort comparison: compares results between two or more populations in different contexts that result in different confounding structures.
+  - Literature search
+- _Omic → outcome_
+  - Multivariate analysis in observational data
+  - Cross-cohort comparison
+  - Mendelian Randomization: instrumental variable is one or more genetic variant(s) that have been shown to robustly relate to the exposure.
+:::
+
+::::
+
+## Triangulation: Details
+
+::: {.panel-tabset}
+### `outcome ~ chemical`
+- **Regression analysis**.
+We will use TMLE in combination with SL, and a library of estimators including both simple parametric models (i.e., `glm`), and data-adaptive semi-parametric models.
+Specifically, we will make use of the `lmtp` R package, in combination with TMLE and SL.
+We hypothesize the presence of residual confounding, especially due to genetic and parental \acr{se} factors, which would result in **exaggeration** of any true causal effect.
+- **Outcome negative control study**.
+We postulate that SE factors represent the main confounders for this association.
+We identified outcomes that are associated to SE factors, but that are not causally associated to the chemicals: having a car, and whether the child has their own room.
+We will thus test the assumptions regarding the relation between the SE factors with each outcome, and check whether these associations are in the anticipated direction.
+- **Literature search**. For each `outcome ~ chemical` association, we will perform a literature search to compare the obtained results.
+
+### `omic ~ chemical`
+- **Regression analysis**. As above.
+- **Cross-cohort comparison**.
+We will compare the regression models' results between subjects of different ethnic origins.
+We postulate that SE factors represent the main confounders for this association.
+Based on the associations between SE factors (e.g., maternal education) and the omic markers, we will compare the results of `omic ~ chemical` with the expected modified associations (e.g., weaker or stronger).
+- **Literature search**. As above.
+
+### `outcome ~ omic`
+- **Regression analysis**. As above.
+- **Cross-cohort comparison**.
+We postulate that SE factors represent the main confounders for this association.
+We will thus test the association between SE factors and the omics, separately for the two sub-populations (e.g., white British and Pakistani in BiB [@WrightSmallRaynor:2013]).
+If this association is smaller among Pakistani, if the association among white British were due to residual SE factor confounding, we would expect a **weaker** association among Pakistani.
+- **MR**.
+We will make use, ideally, of a weighted allele score of genetic variants known to be robustly associated with the omic markers as an IV.
+We will use methods, including sensitivity analysis, to explore the possibility of bias due to: (i) weak instruments, and (ii) violation of the exclusion restriction criteria.
+We will make use of 2SMR [@DaveySmithEbrahim:2003; @PierceBurgess:2013; @DaveySmithHemani:2014].
+We will thus employ results from two different types of GWASs: a exposure GWAS, using The Metabolomics GWAS Server [@SuhreShinPetersen:2011; @ShinFaumanPetersen:2014], and a outcome GWAS, using the OpenGWAS project [@10.7554/eLife.34408; @ElsworthLyonAlexander:2020].
+:::
+
+## Bibliography

### 2023-05-03 docs/causal_roadmap.qmd
Checked some references.

@@
 We will strengthen the estimated effects by making use of the principles and criteria of triangulation [@LawlorTillingDaveySmith:2016].
-To estimate these effects, we will primarily rely on the use of \acr{mtps} [@munozPopulationInterventionCausal2012; @haneuseEstimationEffectInterventions2013; @diazNonparametricCausalEffects2021] in combination with \acr{tmle} [@vanderlaanTargetedMinimumLossBased2018], to avoid relying on arbitrary parametric assumptions.
+To estimate these effects, we will primarily rely on the use of \acr{mtps} [@munozPopulationInterventionCausal2012; @haneuseEstimationEffectInterventions2013; @diazNonparametricCausalEffects2021] in combination with \acr{tmle} [@vanderLaanBenkeserSofrygin:2018], to avoid relying on arbitrary parametric assumptions.
 We will consider both additive and multiplicative shifts of the exposures' distribution.
@@
 We refer to $P(\Lambda=1|a,w)$ as $\bar{\lambda}$.
-To implement estimation of $\bar{\lambda}$ and $\bar{Q}$ flexibly, we will use the ensemble machine learning algorithm \acr{sl} [@vanderlaanSuperLearner2007; @naimiStackedGeneralizationIntroduction2018].
+To implement estimation of $\bar{\lambda}$ and $\bar{Q}$ flexibly, we will use the ensemble machine learning algorithm \acr{sl} [@LaanPolleyHubbard:2007; @NaimiBalzer:2018].
 In the following, we present the steps necessary to estimate $\psi_0 (A^d)$ and the associational parameter $\psi_0^{\Delta}$.

### 2023-05-22 docs/diary.qmd
Updated diary. Error when rendering causal roadmap.

---
title: "Diary"
author: "Lorenzo Fabbri"
date-format: "DD/MM/YYYY"
format:
  pdf
editor_options: 
  chunk_output_type: console
---

# 28/05/2023

Remaining doubts:

* Should I include the covariates when imputing the exposures?
* Should I impute the covariates by cohort?

## Missing values by variables

* There is a clear pattern for some covariates and some cohorts.
* Possible solutions:
  * Drop the variables with a percentage of missing values greater than a pre-specified threshold, by cohort. Perform analyses by cohort, adjusting for different covariates, and pool the results.
  * Drop the variables with a percentage of missing values greater than a pre-specified threshold, by cohort and overall. Perform a single analysis. Then, perform sensitivity analyses by keeping those variables and dropping the cohorts with a *high* fraction of missing values.

## Missing values by cases

* `KAN` participants seem to be the ones with the largest fraction of missing values.
* Possible solutions:
  * Drop these cases and perform the analyses. Then perform sensitivity analysis by including these cases and imputing the missing values.

### 2023-05-26 docs/causal_roadmap.qmd
Updated diary. Error when rendering causal roadmap.
Kinda fixed acronyms problem.

@@
     include-in-header:
-      - \usepackage{rotating}
-      - \usepackage{float}
-      - \usepackage{typearea}
-    pandoc_args:
-      - !expr acronymsdown::add_filter()
+      text: |
+        \usepackage{rotating}
+        \usepackage{float}
+        \usepackage{typearea}
 bibliography: ../../../PhD.bib
@@
 filters:
-  - /home/lorenzo/R/x86_64-pc-linux-gnu-library/4.1/acronymsdown/parse-acronyms.lua
   - ../../../all_phd/scholarly-metadata.lua
   - ../../../all_phd/author-info-blocks.lua
+  - acronyms
 acronyms:
@@
 
-dags <- dag_v2()
+dags <- dags()
 ```

### 2023-05-26 docs/causal_roadmap.qmd
Fixed issues with authors names in PDF and acronyms.

@@
 filters:
-  - ../../../all_phd/scholarly-metadata.lua
-  - ../../../all_phd/author-info-blocks.lua
   - acronyms
+  - authors-block
 acronyms:

### 2023-05-26 docs/causal_roadmap.qmd
Using common lua filters for Quarto.

@@
 filters:
-  - acronyms
-  - authors-block
+  - ../../../all_phd/_extensions/acronyms/parse-acronyms.lua
+  - ../../../all_phd/_extensions/authors-block/authors-block.lua
 acronyms:

### 2023-05-26 output/paper/figures.qmd
Using common lua filters for Quarto.

--- 2023-01-12 output/paper/figures.qmd
@@
 ---
-title: "Figures"
-output:
-  bookdown::pdf_document2
+title: "Paper 3 - Figures"
+format:
+  pdf:
+    pdf-engine: pdflatex
+    toc: false
+    include-in-header:
+      text: |
+        \usepackage{rotating}
+        \usepackage{float}
+        \usepackage{typearea}
+link-citations: true
+linkcolor: blue
+execute:
+  echo: false
+  warning: false
+  error: false
+  message: false
+editor_options: 
+  chunk_output_type: console
+project:
+  execute_dir: project
 ---
@@
 ```{r}
-knitr::opts_chunk$set(echo = FALSE, 
-                      warning = FALSE, message = FALSE, 
-                      cache = TRUE)
+#| include: false
+here::i_am("figures.qmd")
+
+library(tidyverse)
 ```
@@
 <!-- ####################################################################### -->
-<!-- Workflow -->
+# CTD Results
+<!-- ####################################################################### -->
 ```{r}
+#| label: fig-ctd-res
+#| fig.cap: Barplot CTD.
+#| fig.height: 10
+#| fig.width: 10
+path <- "../../results/files/CTD_chem_diseases_1680702221550_05042023.csv"
+ctd <- myphd::load_ctd(path = path, 
+                       filter_evidence = TRUE)
+myphd::plot_ctd(ctd)
 ```
-<!-- ####################################################################### -->

### 2023-05-26 output/paper/manuscript.qmd
Using common lua filters for Quarto.

@@
     include-in-header:
-      - \usepackage{rotating}
-      - \usepackage{float}
-      - \usepackage{typearea}
-    pandoc_args:
-      - !expr acronymsdown::add_filter()
+      text: |
+        \usepackage{rotating}
+        \usepackage{float}
+        \usepackage{typearea}
 bibliography: ../../../../PhD.bib
@@
 filters:
-  - /home/lorenzo/R/x86_64-pc-linux-gnu-library/4.1/acronymsdown/parse-acronyms.lua
-  - ../../../../all_phd/scholarly-metadata.lua
-  - ../../../../all_phd/author-info-blocks.lua
+  - ../../../../all_phd/_extensions/acronyms/parse-acronyms.lua
+  - ../../../../all_phd/_extensions/authors-block/authors-block.lua
 acronyms:

### 2023-05-26 output/paper/tables.qmd
Using common lua filters for Quarto.

@@
 here::i_am("tables.qmd")
-source("../../../../all_phd/edcs.R")
+
 library(tidyverse)
@@
 #| tbl-cap: Information about the non-persistent EDCs.
-edcs_info <- edcs_information() %>%
+edcs_info <- myphd::edcs_information() %>%
   dplyr::select(-c(chem_id, smiles))

### 2023-06-05 _targets_rq01.R
Run tests analyses using fixed myphd package.

@@ -60,5 +60,23 @@ list(
                                         Sys.getenv("TAR_PROJECT"), 
                                         "/out_"
                                       ))
-  ) # End viz_miss_out target
+  ), # End viz_miss_out target
+  targets::tar_target(
+    name = desc_data_covars, 
+    command = myphd::describe_data(dat = load_dat$covariates, 
+                                   id_var = id_var, 
+                                   grouping_var = grouping_var)
+  ), # End desc_data_covars target
+  targets::tar_target(
+    name = desc_data_exp, 
+    command = myphd::describe_data(dat = load_dat$exposures, 
+                                   id_var = id_var, 
+                                   grouping_var = grouping_var)
+  ), # End desc_data_exp target
+  targets::tar_target(
+    name = desc_data_out, 
+    command = myphd::describe_data(dat = load_dat$outcome, 
+                                   id_var = id_var, 
+                                   grouping_var = grouping_var)
+  ) # End desc_data_out target
 )

### 2023-06-05 code/dictionaries.R

@@ -35,16 +35,17 @@ params <- function(is_hpc) {
   })
   ##############################################################################
   
+  outcome_rq1 <- "hs_correct_raven"
   variables <- list(
     identifier = "HelixID", 
     strategy_select_adj_set = "minimize_missings", 
     rq01 = list(
-      outcome = "hs_ADHD_raw", 
+      outcome = outcome_rq1, 
       outcome_negative = "", 
       exposures = "cadj"
     ), 
     rq1 = list(
-      outcome = "hs_ADHD_raw", 
+      outcome = outcome_rq1, 
       outcome_negative = "", 
       exposures = "cadj"
     ), 
@@ -55,7 +56,7 @@ params <- function(is_hpc) {
         threshold_overall = 40
       ), 
       standardization = list(
-        do = TRUE, 
+        do = FALSE, 
         center = TRUE, 
         scale = TRUE
       )

### 2023-07-13 output/presentations/cep_internal_230726/presentation.qmd
Started presentation internal meeting. Reformatted parameters for analyses.

--- 2023-04-17 output/presentations/cep_internal_230419/presentation.qmd
@@
 title: "CEP Group Meeting"
-date: 2023-04-19
+date: 2023-07-26
 author: Lorenzo Fabbri
+institute: ISGlobal
+
 format:
-  revealjs:
-    footer: "Lorenzo Fabbri"
-    toc: true
-    toc_float: true
+  beamer:
+    pdf-engine: xelatex
+    fontsize: "9pt"
+    toc: false
     toc-depth: 2
-    toc-location: body
+
+slide-number: true
+
+bibliography: ../../../../../PhD.bib
 link-citations: true
 linkcolor: blue
+
 execute:
@@
   message: false
+
 editor_options: 
@@
 
-# Overview
+```{r}
+#| include: false
+here::i_am("presentation.qmd")
+```
+
+## Overview
+
+- **Aim**: to research the *short-term* effects of postnatal exposures to non-persistent EDCs on neurodevelopment and neurobehavior in childhood, and how the metabolome and the proteome might *mediate* these effects.
+- **How**: by making use of the principles and criteria of triangulation. We will primarily rely on the use of MTPs in combination with TMLE.
+- **Population**: the HELIX sub-cohort, consisting of $\text{N}=1200$ mother-child pairs.
+- **Exposures**: non-persistent EDCs (phenols, phthalates, and organophosphate compounds), measured in childhood in a pool of two urine samples.
+- **Outcomes**
+  - Raven’s Coloured Progressive Matrices, for assessing non-verbal intelligence.
+  - Computerised n-back test, for assessing working memory.
+  - Attention Network Test, to provide a measure of the efficiency of three different functions of attention.
+  - Child Behavior Checklist, for assessing behavioural and emotional problems.
+- ***Mediators***: serum and urine metabolome, plasma proteome.
 
@@
 
-# Triangulation
+```{mermaid}
+%%| fig-cap: Simplified DAG.
+%%| label: fig-diag-dag
+%%| fig-align: center
+flowchart LR
+  exposure(chemical) --> mediator(metabolite)
+  mediator --> outcome(neurodevelopment)
+  exposure --> outcome
+```
 
-# Reliability and Replicability
+## Triangulation
+
+```{mermaid}
+%%| fig-cap: Diagram summarizing research questions and methods.
+%%| label: fig-diagram
+%%| fig-align: center
+%%| fig-height: 3
+flowchart LR
+  paper[Paper 3] --> rq1("RQ1
+                          neuro~chemical")
+  paper          --> rq2("RQ2
+                          met~chemical")
+  paper          --> rq3("RQ3
+                          neuro~met")
+  rq1            --> rqReg("Observational study.
+                            Method: Regression with MTPs.
+                            Causal estimand: phi.
+                            Uncertainty quantification: bootstrapping.")
+  rq1            --> rq1outNeg("Outcome negative control.
+                                Method: Regression with GLMs/GAMs.
+                                Uncertainty quantification: bootstrapping.")
+  rq1            --> rqLit("Literature search")
+
+  rq2            --> rqReg
+  rq2            --> rqCrossCohort("Cross-cohort comparison.
+                                    Method: Regression with GLMs/GAMs.
+                                    Uncertainty quantification: bootstrapping.")
+  rq2            --> rqLit
+
+  rq3            --> rqReg
+  rq3            --> rqCrossCohort
+  rq3            --> rq3mr("Mendelian Randomization.
+                            Causal estimand: .")
+```

### 2023-07-13 code/dictionaries.R

@@ -112,46 +112,30 @@ params <- function(is_hpc) {
 # Parameters for running analyses
 ################################################################################
 params_analyses <- function() {
-  # Common parameters
-  ## lmtp
-  learners_outcome <- c("SL.mean", "SL.glm")
-  learners_exposure <- c("SL.glm")
-  estimator <- "sdr"
-  folds <- 3
-  folds_outcome <- 1
-  folds_exposure <- 1
-  density_ratio_trim <- 0.995
-  markov_assumption <- Inf
-  shift_type <- "mul"
-  shift_amount <- 0.01
-  
-  ## WeightIt, Cobalt, marginaleffects
-  method_weightit <- "glm"
-  method_marginal <- "glm"
-  weights_trim <- 0.95
-  use_kernel <- FALSE
-  sl_discrete <- FALSE
-  sl_lib <- FALSE
-  
   rq1 <- list(
     # lmtp
-    learners_outcome = learners_outcome, 
-    learners_trt = learners_exposure, 
-    estimator = estimator, 
-    folds = folds, 
-    .learners_outcome_folds = folds_outcome, 
-    .learners_trt_folds = folds_exposure, 
-    .trim = density_ratio_trim, 
-    k = markov_assumption, 
-    shift_type = shift_type, 
-    shift_amount = shift_amount, 
+    learners_outcome = c("SL.mean", "SL.glm"), 
+    learners_trt = c("SL.glm"), 
+    estimator = "sdr", 
+    folds = 3, 
+    .learners_outcome_folds = 1, 
+    .learners_trt_folds = 1, 
+    .trim = 0.995, 
+    k = Inf, 
+    shift_type = "mul", 
+    shift_amount = 0.01, 
+    ############################################################################
     # WeightIt, Cobalt, marginaleffects
-    method_weightit = method_weightit, 
-    method_marginal = method_marginal, 
-    weights_trim = weights_trim, 
-    use_kernel = use_kernel, 
-    sl_discrete = sl_discrete, 
-    sl_lib = sl_lib
+    method_weightit = "glm", 
+    weights_trim = 0.95, 
+    use_kernel = FALSE, 
+    sl_discrete = FALSE, 
+    sl_lib = FALSE, 
+    method_marginal = "glm", 
+    family_marginal = "binomial", 
+    add_inter_exposure = TRUE, 
+    add_splines_exposure = TRUE, 
+    df_spline 3
   ) # End dictionary parameters RQ1
   rq2 <- rq1
   rq3 <- rq1

### 2023-07-13 code/research_questions/utils.R

@@ -326,7 +326,12 @@ rq_fit_model_weighted <- function(dat, weights) {
         covariates = list_covariates, 
         weights = weights_exposure, 
         method = params_ana$method_marginal, 
-        method_args = c()
+        method_args = c(
+          family = params_ana$family_marginal, 
+          add_inter_exposure = params_ana$add_inter_exposure, 
+          add_splines_exposure = params_ana$add_splines_exposure, 
+          df_splines = params_ana$df_splines
+        )
       )
     }) # End loop over exposures
   }) # End progress bar

### 2023-07-21 output/presentations/cep_internal_230726/presentation.qmd
Added DAG for RQ2, and corresponding file. Worked more on presentation for group meeting.

@@
 title: "CEP Group Meeting"
+subtitle: "Update on *paper 3* of my PhD"
 date: 2023-07-26
@@
     toc-depth: 2
+geometry:
+  - left=10mm
+  - right=10mm
 
@@
 
-- **Aim**: to research the *short-term* effects of postnatal exposures to non-persistent EDCs on neurodevelopment and neurobehavior in childhood, and how the metabolome and the proteome might *mediate* these effects.
-- **How**: by making use of the principles and criteria of triangulation. We will primarily rely on the use of MTPs in combination with TMLE.
+- **Aim**: to research the *short-term* effects of childhood exposures to non-persistent EDCs on neurodevelopment in childhood, and how the metabolome **might** *mediate* these effects.
+- **How**: by making use of the principles and criteria of triangulation.
 - **Population**: the HELIX sub-cohort, consisting of $\text{N}=1200$ mother-child pairs.
 - **Exposures**: non-persistent EDCs (phenols, phthalates, and organophosphate compounds), measured in childhood in a pool of two urine samples.
-- **Outcomes**
-  - Raven’s Coloured Progressive Matrices, for assessing non-verbal intelligence.
-  - Computerised n-back test, for assessing working memory.
-  - Attention Network Test, to provide a measure of the efficiency of three different functions of attention.
-  - Child Behavior Checklist, for assessing behavioural and emotional problems.
-- ***Mediators***: serum and urine metabolome, plasma proteome.
+- **Outcome**: Attention Network Test, to provide a measure of the efficiency of three different functions of attention.
+- ***Mediators***: urine metabolites (focus on corticosteroids).
 
@@
 ```{mermaid}
-%%| fig-cap: Simplified DAG.
+%%| fig-cap: Simplified *DAG*.
 %%| label: fig-diag-dag
@@
 ```
+
+Issues and limitations:
+
+- The effect of exposure to these chemicals on neurodevelopment has been studied already, although there are clear issues with that study (e.g., wrong model, DAG **not** tailored to the exposures of interest).
+- The effect of exposure to these chemicals on these **newly** measured metabolites could be identifiable, although they were measured in the same samples.
+This question poses some challenges since they were all measured in urine (inter-individual variability in urine dilution).
+- The effect of these **newly** measured metabolites on neurodevelopment could be identifiable, since they were measure in the urine samples (\textcolor{red}{which aliquot?}) collected before the visit.
 
@@
                             Causal estimand: .")
+  style rq2 fill:#7faac2
+  style rq3 fill:#7faac2
 ```
+
+## `myphd` R package
+
+![`myphd`](artwork/myphd.png){width=300}
+
+- URL manual: https://isglobal-cep.github.io/myphd/
+- To install it: `remotes::install_github("isglobal-cep/myphd")`
+
+:::: {.columns align=center}
+
+::: {.column width="33%" align=top}
+```{r, eval=FALSE, echo=TRUE}
+myphd::create_formula(
+  dat,
+  outcome,
+  exposure,
+  covariates,
+  method,
+  add_inter_exposure,
+  add_splines_exposure,
+  df_splines,
+  threshold_smooth,
+  threshold_k
+)
+```
+:::
+
+::: {.column width="33%" align=top}
+```{r, eval=FALSE, echo=TRUE}
+myphd::describe_data(
+  dat,
+  id_var,
+  grouping_var
+)
+```
+:::
+
+::: {.column width="33%" align=top}
+```{r, eval=FALSE, echo=TRUE}
+myphd::preproc_data(
+  dat,
+  outcome = NULL,
+  dic_steps,
+  id_var,
+  by_var
+)
+```
+:::
+
+::::
+
+## Current setup
+
+Variables:
+
+- Clinical outcome: Attention Network Test (`hs_hitrtse`, Hit Reaction Time Standard Error).
+
+Methods:
+
+- Weights estimation (`chemical ~ covariates`): energy balancing.
+- Effect estimation (`outcome ~ chemical + covariates`): `glm` with natural splines for exposure.
+
+## Population description: correlations
+
+![`Correlation numerical variables.`](artwork/corr.png)
+
+## Preliminary results
+
+## Next steps

### 2023-07-21 DAGs/dag_v1.R

@@ -183,7 +183,84 @@ dag_raw <- function() {
   ##############################################################################
   
   ##############################################################################
-  chem_to_marker <- ''
+  chem_to_marker <- 'dag {
+  bb="0,0,1,1"
+  age_child [pos="0.403,0.100"]
+  airPollution_child [pos="0.148,0.779"]
+  biomarker [outcome,pos="0.889,0.490"]
+  characteristics_child [pos="0.403,0.169"]
+  chemical [exposure,pos="0.154,0.487"]
+  child_diet [pos="0.417,0.647"]
+  child_smoking [pos="0.178,0.107"]
+  cohort [pos="0.608,0.188"]
+  creatinine [pos="0.508,0.573"]
+  ethnicity_child [pos="0.262,0.103"]
+  familySEP [pos="0.602,0.089"]
+  maternalAlcohol_preg [pos="0.829,0.694"]
+  maternalSmoking_preg [pos="0.832,0.771"]
+  otherChemicals_child [pos="0.150,0.695"]
+  season_visit [pos="0.511,0.174"]
+  sex_child [pos="0.337,0.099"]
+  time_lastMeal [pos="0.507,0.648"]
+  age_child -> biomarker
+  age_child -> characteristics_child
+  age_child -> creatinine
+  airPollution_child -> biomarker
+  airPollution_child -> characteristics_child
+  characteristics_child -> biomarker
+  characteristics_child -> chemical
+  characteristics_child -> creatinine
+  chemical -> biomarker
+  child_diet -> biomarker
+  child_diet -> characteristics_child
+  child_diet -> chemical
+  child_diet -> otherChemicals_child
+  child_smoking -> biomarker
+  child_smoking -> characteristics_child
+  child_smoking -> otherChemicals_child
+  cohort -> airPollution_child
+  cohort -> biomarker
+  cohort -> characteristics_child
+  cohort -> chemical
+  cohort -> child_diet
+  cohort -> child_smoking
+  cohort -> familySEP
+  cohort -> otherChemicals_child
+  creatinine -> biomarker
+  creatinine -> chemical
+  ethnicity_child -> airPollution_child
+  ethnicity_child -> biomarker
+  ethnicity_child -> characteristics_child
+  ethnicity_child -> chemical
+  ethnicity_child -> child_diet
+  ethnicity_child -> child_smoking
+  ethnicity_child -> creatinine
+  ethnicity_child -> otherChemicals_child
[…]

### 2023-07-21 DAGs/dag_v2.R

@@ -30,7 +30,17 @@ dags <- function() {
   
   ##############################################################################
   chem_to_marker <- ggdag::dagify(
-    
+    airPollution_child ~ cohort + ethnicity_child + familySEP + season_visit,
+    biomarker ~ age_child + airPollution_child + characteristics_child + chemical + child_diet + child_smoking + cohort + creatinine + ethnicity_child + familySEP + maternalAlcohol_preg + maternalSmoking_preg + otherChemicals_child + season_visit + sex_child + time_lastMeal,
+    characteristics_child ~ age_child + airPollution_child + child_diet + child_smoking + cohort + ethnicity_child + familySEP + maternalAlcohol_preg + maternalSmoking_preg + otherChemicals_child + sex_child,
+    chemical ~ characteristics_child + child_diet + cohort + creatinine + ethnicity_child + familySEP + season_visit + time_lastMeal,
+    child_diet ~ cohort + ethnicity_child + familySEP + sex_child,
+    child_smoking ~ cohort + ethnicity_child + familySEP + sex_child,
+    creatinine ~ age_child + characteristics_child + ethnicity_child + sex_child,
+    familySEP ~ cohort,
+    otherChemicals_child ~ child_diet + child_smoking + cohort + ethnicity_child + familySEP + season_visit,
+    exposure = "chemical",
+    outcome = "biomarker"
   )
   ##############################################################################
   

### 2023-07-21 output/presentations/cep_internal_230726/code.R

@@ -1,0 +1,60 @@
+source("code/dictionaries.R")
+source("DAGs/dag_v2.R")
+source("code/data.R")
+source("code/research_questions/utils.R")
+source("code/research_questions/rq1.R")
+
+Sys.setenv(is_hpc = FALSE)
+# Custom path to _targets for different research questions
+path_store <- ifelse(
+  Sys.getenv("is_hpc"), 
+  "/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/lorenzoF/data/data_paper3/_targets/_targetsRQ", 
+  "~/mounts/rstudioserver/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/lorenzoF/data/data_paper3/_targets/_targetsRQ"
+)
+Sys.setenv(path_store = path_store)
+# Path to store results
+path_store_res <- ifelse(
+  Sys.getenv("is_hpc"), 
+  "/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/lorenzoF/data/data_paper3/results/", 
+  "~/mounts/rstudioserver/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/lorenzoF/data/data_paper3/results/"
+)
+Sys.setenv(path_store_res = path_store_res)
+Sys.setenv(TAR_PROJECT = "rq1")
+store <- paste0(path_store, "1")
+################################################################################
+
+params_dag <- list(
+  type = "minimal", 
+  effect = "total"
+)
+dag <- load_dag(dags = dags(), 
+                exposure = "chemical", 
+                outcome = "intelligence", 
+                params_dag = params_dag)
+load_dat <- rq_load_data(ids_other_covars = c(), 
+                         res_dag = dag)
+preproc_dat <- rq1_prepare_data(dat = load_dat)
+df <- dplyr::inner_join(
+  preproc_dat$outcome, 
+  preproc_dat$covariates, 
+  by = "HelixID"
+) |>
+  myphd::extract_cohort(id_var = "HelixID")
+################################################################################
+
+vars_to_desc <- c(
+  "HelixID", 
+  "hs_age_years", "cohort", 
+  "e3_bw", "e3_gac", 
+  "hs_head_circ", "hs_waist", 
+  "hs_hitrtse"
+)
+desc_dat <- myphd::describe_data(df |>
+                                   dplyr::select(dplyr::all_of(vars_to_desc)), 
+                                 id_var = "HelixID", 
[…]

### 2023-07-25 output/presentations/cep_internal_230726/presentation.qmd
Run full pipeline for cortisone for presentation CEP.

@@
 institute: ISGlobal
-
 format:
@@
   - right=10mm
-
 slide-number: true
-
 bibliography: ../../../../../PhD.bib
@@
 linkcolor: blue
-
 execute:
@@
   message: false
-
 editor_options: 
@@
 ## Next steps
+
+- `lmtp` will be the focus of my research stay with ... at ...

### 2023-07-25 _targets_rqX.R

@@ -42,7 +42,7 @@ list(
   targets::tar_target(
     name = weights, 
     command = rq_estimate_weights(dat = preproc_dat, 
-                                  save_results = FALSE, 
+                                  save_results = TRUE, 
                                   parallel = FALSE)
   ), # End weights target
   targets::tar_target(

### 2023-07-25 code/utils.R

@@ -288,6 +288,19 @@ rq_estimate_weights <- function(dat, save_results, parallel, workers) {
     } else {
       future::plan(future::sequential())
     }
+    
+    invisible(
+      lapply(c("figures", "tables"), function(x) {
+        path_save_res <- paste0(
+          "results/", x, "/", 
+          Sys.getenv("TAR_PROJECT")
+        )
+        if (!dir.exists(path_save_res)) {
+          dir.create(path_save_res)
+        }
+      })
+    )
+    
     furrr::future_map(names(balance), function(x) {
       ### bal.plot
       .path <- paste0(
@@ -302,7 +315,7 @@ rq_estimate_weights <- function(dat, save_results, parallel, workers) {
                       gridExtra::marrangeGrob(grobs = balance[[x]]$graph,
                                               nrow = 1,
                                               ncol = 1),
-                      dpi = 480, type = "cairo")
+                      dpi = 480)
       
       ### bal.tab
       .path <- paste0(
@@ -315,10 +328,14 @@ rq_estimate_weights <- function(dat, save_results, parallel, workers) {
       )
       tab <- balance[[x]]$tab$Balance |>
         as.data.frame() |>
-        tibble::rownames_to_column(var = "variable")
-      colnames(tab) <- c("variable", "type",
-                         "unadj. correlation", "adj. correlation",
-                         "threshold", "adj. KS")
+        tibble::rownames_to_column(var = "variable") |>
+        dplyr::rename(
+          variable = variable, 
+          type = Type, 
+          `unadj. correlation` = `Corr.Un`, 
+          `adj. correlation` = `Corr.Adj`, 
+          threshold = `R.Threshold`
+        )
       tab <- tab |>
         dplyr::arrange(dplyr::desc(`adj. correlation`),
                        variable) |>
@@ -347,10 +364,10 @@ rq_estimate_weights <- function(dat, save_results, parallel, workers) {
       ".pdf"
     )
     ggplot2::ggsave(.path,
-                    gridExtra::marrangeGrob(grobs = sapply(balance, `[[`, "love"),
+                    gridExtra::marrangeGrob(grobs = lapply(balance, `[[`, "love"),
                                             nrow = 1,
                                             ncol = 1),
-                    dpi = 480, height = 6, type = "cairo")
+                    dpi = 480, height = 6)
   } # End save_results
   
   return(list(
[…]

### 2023-07-25 run.R

@@ -27,4 +27,5 @@ store <- paste0(path_store, rq)
 targets::tar_make(script = "_targets_rqX.R", 
                   store = store)
 
-targets::tar_load_everything(store = store)
+#targets::tar_destroy(destroy = "all", store = store)
+#targets::tar_load_everything(store = store)

### 2023-07-25 output/presentations/cep_internal_230726/presentation.qmd
Finished presentation for CEP meeting.

@@
 here::i_am("presentation.qmd")
+
+library(patchwork)
 ```
@@
 - URL manual: https://isglobal-cep.github.io/myphd/
-- To install it: `remotes::install_github("isglobal-cep/myphd")`
+- To install it: `remotes::install_github("isglobal-cep/myphd")`.
 
@@
 
-![`Correlation numerical variables.`](artwork/corr.png)
+![Correlation numerical variables.](artwork/corr.png)
 
-## Preliminary results
+## Preliminary results: weights
+
+```{r}
+targets::tar_load(c("weights", 
+                    "weighted_fits", 
+                    "marginal"), 
+                  store = paste0(
+                    "~/mounts/rstudioserver/PROJECTES/HELIX_OMICS/",  
+                    "DATA_PREVIOUS_MIGRATION/lorenzoF/", 
+                    "data/data_paper3/_targets/_targetsRQ2/"
+                  ))
+```
+
+```{r}
+weights$balance$hs_bpa_c$love
+```
+
+## Preliminary results: weighted G-computation
+
+- Marginal effects of exposure to non-persistent EDCs on Cortisone: comparison `Q3 - Q1`.
+
+![Effect estimates for EDCs and Cortisone.](artwork/effects_cortisone.png){width=130%, fig-align="center"}
+
+```{r}
+gtsummary::tbl_summary(weighted_fits$fits$hs_bpa_c$dat |>
+                         dplyr::rename(BPA = hs_bpa_c, 
+                                       cortisone = Cortisone_E), 
+                       include = c("BPA", "cortisone"), 
+                       by = "cohort") |>
+  gtsummary::add_overall() |>
+  gtsummary::as_kable_extra(booktabs = TRUE) |>
+  kableExtra::kable_styling(latex_options = "scale_down")
+```
+
+## Preliminary results: ADRF and AMEF
+
+```{r}
+marginal$marginal_effects$hs_bpa_c$adrf / 
+  marginal$marginal_effects$hs_bpa_c$amef
+```
 
@@
 
-- `lmtp` will be the focus of my research stay with ... at ...
+- Dealing with creatinine.
+
+![Comparison of effect estimation.](artwork/comp_creatinineAdj.png)
+
+- Developing a more appropriate pipeline to deal with censored exposures and metabolites.
+
+- Dealing with more appropriate modeling strategies?
+
+  - `lmtp` will be the focus of my research stay (*Mixtures done right*) under the supervision of Prof.
+Jessica G.
+Young and Prof.
+Ivan Diaz (who formalized causal identification of interventions that depend on the natural value of treatment).

### 2023-08-01 docs/causal_roadmap.qmd
Trying to use targets to compile documents for paper.

--- 2023-05-26 docs/causal_roadmap.qmd
@@
   - ../../../all_phd/authors.yml
+
 format:
@@
         \usepackage{typearea}
+
 bibliography: ../../../PhD.bib
@@
 linkcolor: blue
+
 execute:
@@
   chunk_output_type: console
+
 filters:
-  - acronyms
-  - authors-block
+  - ../../../all_phd/_extensions/acronyms/parse-acronyms.lua
+  - ../../../all_phd/_extensions/authors-block/authors-block.lua
 acronyms:
@@
 #| include: false
+source("../../../all_phd/info_phd.R")
+source("../../../all_phd/triangulation_methods.R")
 
 here::i_am("causal_roadmap.qmd")
-source("../../../all_phd/info_phd.R")
-source("../../../all_phd/triangulation_methods.R")
 infos <- info_phd()
 methods <- describe_triangulation_methods()
-
-library(tidyverse)
 ```

### 2023-08-01 output/paper/figures.qmd
Trying to use targets to compile documents for paper.

@@
 title: "Paper 3 - Figures"
+
 format:
@@
         \usepackage{typearea}
+
 link-citations: true
 linkcolor: blue
+
 execute:
@@
   chunk_output_type: console
-project:
-  execute_dir: project
 ---
-
-```{r}
-#| include: false
-here::i_am("figures.qmd")
-
-library(tidyverse)
-```
 

### 2023-08-01 output/paper/manuscript.qmd
Trying to use targets to compile documents for paper.

@@
   - ../../../../all_phd/authors.yml
+
 format:
   pdf:
-    pdf-engine: pdflatex
+    pdf-engine: xelatex
     number-sections: true
@@
         \usepackage{typearea}
+
 bibliography: ../../../../PhD.bib
@@
   chunk_output_type: console
+
 filters:
@@
   sorting: alphabetical
+
 abstract: |
@@
 
-```{r}
-#| include: false
-
-here::i_am("manuscript.qmd")
-source("../../../../all_phd/info_phd.R")
-infos <- info_phd()
-```
-
 ---
-title: "`r infos$paper3$title`"
+title: "Some Title"
 ---
@@
 <!-- ####################################################################### -->
-<!-- INTRODUCTION -->
 # Introduction {#intro}
 ```{r}
-#| child: sections/introduction.qmd
+#| child: output/paper/sections/introduction.qmd
 ```
@@
 <!-- ####################################################################### -->
-<!-- METHODS -->
 # Methods {#methods}
 ```{r}
-#| child: sections/methods.qmd
+#| child: output/paper/sections/methods.qmd
 ```
@@
 <!-- ####################################################################### -->
-<!-- RESULTS -->
 # Results {#res}
 ```{r}
-#| child: sections/results.qmd
+#| child: output/paper/sections/results.qmd
 ```
@@
 <!-- ####################################################################### -->
-<!-- DISCUSSION -->
 # Discussion {#discussion}
 ```{r}
-#| child: sections/discussion.qmd
+#| child: output/paper/sections/discussion.qmd
 ```
@@
 
-# References
+<!-- ####################################################################### -->
+# References {.unnumbered}
+<!-- ####################################################################### -->
+
+::: {#refs}
+:::

### 2023-08-01 output/paper/tables.qmd
Trying to use targets to compile documents for paper.

@@
 title: "Paper 3 - Tables"
+
 format:
@@
     reference-doc: reference_document.docx
+
 link-citations: true
 linkcolor: blue
+
 execute:
@@
   chunk_output_type: console
-project:
-  execute_dir: project
 ---
-
-```{r}
-#| include: false
-here::i_am("tables.qmd")
-
-library(tidyverse)
-```
 
@@
 #| tbl-cap: Information about the non-persistent EDCs.
-edcs_info <- myphd::edcs_information() %>%
+edcs_info <- myphd::edcs_information() |>
   dplyr::select(-c(chem_id, smiles))
-gt::gt(edcs_info, groupname_col = "class") %>%
+gt::gt(edcs_info, groupname_col = "class") |>
   gt::cols_label(
@@
     short_name = "Symbol", 
-    #smiles = "SMILES", 
     pubchem_cid = "PubChem CID", 
@@
     parental_compound = "Parental compound"
-  ) %>%
+  ) |>
   gt::tab_style(
@@
     )
-  ) %>%
+  ) |>
   gt::tab_style(
-    locations = gt::cells_row_groups(group = everything()), 
+    locations = gt::cells_row_groups(group = dplyr::everything()), 
     style = list(gt::cell_text(weight = "bold"))

### 2023-08-01 output/paper/_targets_paper.R

@@ -1,30 +1,29 @@
 targets::tar_option_set(
   format = "qs"
 )
-options(clustermq.scheduler = "future")
 
 list(
   tarchetypes::tar_quarto(
     name = figures, 
-    path = "figures.qmd", 
+    path = "output/paper/figures.qmd", 
     quiet = FALSE
   ), # End target manuscript
   ##############################################################################
   tarchetypes::tar_quarto(
     name = tables, 
-    path = "tables.qmd", 
+    path = "output/paper/tables.qmd", 
     quiet = FALSE
   ), # End target manuscript
   ##############################################################################
   tarchetypes::tar_quarto(
     name = manuscript, 
-    path = "manuscript.qmd", 
+    path = "output/paper/manuscript.qmd", 
     quiet = FALSE
   ), # End target manuscript
   ##############################################################################
   tarchetypes::tar_quarto(
     name = si, 
-    path = "si.qmd", 
+    path = "output/paper/si.qmd", 
     quiet = FALSE
   ) # End target manuscript
   ##############################################################################

### 2023-08-02 output/paper/figures.qmd
Using targets to build paper. Checking that labels levels factors are displayed in summary tables.

@@
   pdf:
-    pdf-engine: pdflatex
+    pdf-engine: lualatex
     toc: false
@@
 #| fig.width: 10
-path <- "../../results/files/CTD_chem_diseases_1680702221550_05042023.csv"
+path <- "results/files/CTD_chem_diseases_1680702221550_05042023.csv"
 ctd <- myphd::load_ctd(path = path, 

### 2023-08-02 output/paper/manuscript.qmd
Using targets to build paper. Checking that labels levels factors are displayed in summary tables.

@@
   pdf:
-    pdf-engine: xelatex
+    pdf-engine: lualatex
     number-sections: true
@@
 <!-- ####################################################################### -->
-# Introduction {#intro}
+# Introduction {#sec-intro}
 ```{r}
@@
 <!-- ####################################################################### -->
-# Methods {#methods}
+# Methods {#sec-methods}
 ```{r}
@@
 <!-- ####################################################################### -->
-# Results {#res}
+# Results {#sec-res}
 ```{r}
@@
 <!-- ####################################################################### -->
-# Discussion {#discussion}
+# Discussion {#sec-discussion}
 ```{r}

### 2023-08-02 output/paper/sections/methods.qmd
Using targets to build paper. Checking that labels levels factors are displayed in summary tables.

## Study population {#sec-study-population}

## Exposure assessment {#sec-exp-assessment}

## Metabolomics {#sec-metabolomics}

## Neurodevelopment {#sec-neurodevelopment}

## Covariates {#sec-covars}

## Statistical analysis {#sec-stat-analysis}

### 2023-08-02 output/paper/si.qmd
Using targets to build paper. Checking that labels levels factors are displayed in summary tables.

--- 2023-08-01 output/paper/tables.qmd
@@
 ---
-title: "Paper 3 - Tables"
+title: "Paper 3 - Supplementary Information"
 
@@
 ---
-
-<!-- ####################################################################### -->
-# Study populations
-<!-- ####################################################################### -->
-```{r}
-
-```
-
-<!-- ####################################################################### -->
-# Exposures
-<!-- ####################################################################### -->
-```{r}
-#| label: tbl-chem-info
-#| tbl-cap: Information about the non-persistent EDCs.
-edcs_info <- myphd::edcs_information() |>
-  dplyr::select(-c(chem_id, smiles))
-gt::gt(edcs_info, groupname_col = "class") |>
-  gt::cols_label(
-    full_name = "Compound", 
-    short_name = "Symbol", 
-    pubchem_cid = "PubChem CID", 
-    ctd_id = "CTD ID", 
-    exposome_explorer_id = "Exposome Explore ID", 
-    parental_compound = "Parental compound"
-  ) |>
-  gt::tab_style(
-    locations = gt::cells_column_labels(columns = everything()), 
-    style = list(
-      gt::cell_borders(sides = "bottom", weight = gt::px(3)), 
-      gt::cell_text(weight = "bold")
-    )
-  ) |>
-  gt::tab_style(
-    locations = gt::cells_row_groups(group = dplyr::everything()), 
-    style = list(gt::cell_text(weight = "bold"))
-  )
-```

### 2023-08-02 code/dictionaries.R

@@ -7,29 +7,13 @@ params <- function(is_hpc) {
   if (is_hpc == TRUE) {
     common_path <- "../../../../"
   } else {
-    common_path <- "~/mounts/rstudioserver/PROJECTES/HELIX_OMICS/"
+    common_path <- "~/mounts/rstudioserver/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/lorenzoF"
   }
   
   paths <- list(
     # Raw
-    path_covariates_post = "", 
-    path_covariates_preg = "", 
-    path_snps = "", 
-    path_exposures_backtrans = "data_raw/exposome.backtransf/child/preg_8y/20180901/20180919_v2_3_backtransformed.Rdata", 
-    path_exposures_post_raw = "data_final/exposome/child/preg_8y/v2_3_20180601/exppostnatal_raw_v2_3.RData", 
-    path_exposures_preg_raw = "data_final/exposome/child/preg_8y/v2_3_20180601/exppregnancy_raw_v2_3.RData", 
-    path_exposures_post_final = "data_final/exposome/child/preg_8y/v2_3_20180601/exppostnatal_v2_3.RData", 
-    path_exposures_preg_final = "data_final/exposome/child/preg_8y/v2_3_20180601/exppregnancy_v2_3.RData", 
-    path_metabolome_serum = "data_final/metab/child/8y/serum.urine_Biocrates.NMR_QChelix_20170101/metab_serum_subcohort_v3.RData", 
-    path_metabolome_urine = "data_final/metab/child/8y/serum.urine_Biocrates.NMR_QChelix_20170101/metab_urine_subcohort_v3.RData", 
-    path_dat_request = "DATA_PREVIOUS_MIGRATION/lorenzoF/data/data_paper3/requests/AP136/HELIX_AP_136_request_updated12jun.2023.csv", 
-    path_all_steroids = "DATA_PREVIOUS_MIGRATION/lorenzoF/data/data_paper3/requests/AP136/steroids/", 
-    
-    # Processed during analyses
-    path_exposures_post_procme = "DATA_PREVIOUS_MIGRATION/lorenzoF/data/data_paper3/processed/exposures/", 
-    path_exposures_preg_procme = "DATA_PREVIOUS_MIGRATION/lorenzoF/data/data_paper3/processed/exposures/", 
-    path_metabolome_serum = "DATA_PREVIOUS_MIGRATION/lorenzoF/data/data_paper3/processed/omics/", 
[…]

### 2023-08-02 output/paper/run_paper.R

@@ -1,0 +1,10 @@
+Sys.setenv(is_hpc = TRUE)
+store <- ifelse(
+  Sys.getenv("is_hpc"), 
+  "/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/lorenzoF/data/data_paper3/_targets/_targets_paper", 
+  "~/mounts/rstudioserver/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/lorenzoF/data/data_paper3/_targets/_targets_paper"
+)
+Sys.setenv(store = store)
+
+targets::tar_make(script = "output/paper/_targets_paper.R", 
+                  store = store)

### 2023-08-02 run.R

@@ -17,7 +17,7 @@ path_store_res <- ifelse(
 Sys.setenv(path_store_res = path_store_res)
 ################################################################################
 
-rq <- "3"
+rq <- "2"
 Sys.setenv(TAR_PROJECT = paste0("rq0", rq))
 store <- paste0(path_store, paste("0", rq))
 targets::tar_make(script = "_targets_rq0X.R", 

### 2023-08-09 output/paper/sections/discussion.qmd
Added function to plot densities of variables by grouping variable and with CDF. Working on avg_comparisons with  term and with functionality to compute quantiles by grouping variables since for instance chemical levels vary a lot by cohort.

List of possible limitations:

* The percentiles of the exposures might differ considerably by cohort. When using `avg_comparisons`, we do not consider this. This might cause cohort-dependent positivity issues (?).

### 2023-08-09 code/dictionaries.R

@@ -249,7 +249,8 @@ params_analyses <- function() {
     add_splines_exposure = TRUE, 
     df_splines = 3, 
     threshold_smooth = 10, 
-    threshold_k = 3
+    threshold_k = 3, 
+    type_avg_comparison = c(0.1, 0.9)
   ) # End dictionary parameters RQ1
   rq2 <- rq1
   rq3 <- rq1

### 2023-08-09 code/utils.R

@@ -522,6 +522,7 @@ rq_fit_model_weighted <- function(dat, outcome,
 rq_estimate_marginal_effects <- function(fits, parallel, workers) {
   rq <- Sys.getenv("TAR_PROJECT")
   params_dat <- params(is_hpc = Sys.getenv("is_hpc"))
+  params_ana <- params_analyses()[[rq]]
   
   # Loop over the fitted models to estimate marginal effects
   if (parallel == TRUE) {
@@ -545,8 +546,8 @@ rq_estimate_marginal_effects <- function(fits, parallel, workers) {
       # G-computation (ADRF)
       ## Values of exposure for counterfactual predictions, based on quantiles
       values <- with(dat, seq(
-        quantile(get(exposure), 0.1), 
-        quantile(get(exposure), 0.9), 
+        quantile(get(exposure), params_ana$type_avg_comparison[1]), 
+        quantile(get(exposure), params_ana$type_avg_comparison[2]), 
         length.out = 50
       ))
       gcomp <- eval(
@@ -637,19 +638,26 @@ rq_estimate_marginal_effects <- function(fits, parallel, workers) {
       
       ############################################################################
       # Comparisons (marginal estimates)
+      ## Create dataframe with `high` and `low` values for exposure
+      df_comparisons <- myphd::create_df_marginal_comparisons(
+        dat = dat, 
+        var = exposure, 
+        percentiles = params_ana$type_avg_comparison, 
+        group_var = "cohort"
+      )
+      
       avg_comp <- eval(
         parse(
           text = glue::glue(
             "marginaleffects::avg_comparisons(
             model = mod, 
             variables = list(
-              {exposure} = {glue::double_quote(type)}
+              {exposure} = df_comparisons
             ), 
             wts = weights, 
             vcov = {glue::double_quote(vcov)}
           )", 
             exposure = exposure, 
-            type = "iqr", 
             vcov = "HC3"
           )
         )

### 2023-08-18 output/paper/tables.qmd
Replaced tidyr/dplyr with tidylogs wherever possible. Tested new myphd package. Checking correctness code.

@@
 edcs_info <- myphd::edcs_information() |>
-  dplyr::select(-c(chem_id, smiles))
+  tidylog::select(-c(chem_id, smiles))
 gt::gt(edcs_info, groupname_col = "class") |>

### 2023-08-18 code/plot.R

@@ -12,13 +12,13 @@ viz_overlap_quantiles <- function(dat, id_var, group_var,
                                   lower_percentile, upper_percentile) {
   # Process data
   dat_proc <- dat |>
-    dplyr::select(-dplyr::any_of(id_var))
+    tidylog::select(-dplyr::any_of(id_var))
   vars <- setdiff(colnames(dat_proc), group_var)
   
   # Plot
   ret <- lapply(vars, function(x) {
     dat_proc |>
-      dplyr::select(dplyr::all_of(c(x, group_var))) |>
+      tidylog::select(dplyr::all_of(c(x, group_var))) |>
       ggplot2::ggplot(ggplot2::aes(x = .data[[x]], 
                                    y = .data[[group_var]], 
                                    fill = 0.5 - abs(0.5 - stat(ecdf)))) +
@@ -41,10 +41,10 @@ viz_overlap_quantiles <- function(dat, id_var, group_var,
       ggplot2::scale_fill_viridis_c(name = "Tail probability", 
                                     direction = -1) +
       ggplot2::labs(caption = glue::glue(
-                      "Lower percentile: {lower}th. Upper percentile: {upper}th.", 
-                      lower = lower_percentile * 100, 
-                      upper = upper_percentile * 100
-                    ))
+        "Lower percentile: {lower}th. Upper percentile: {upper}th.", 
+        lower = lower_percentile * 100, 
+        upper = upper_percentile * 100
+      ))
   }) # End loop over variables to plot
   
   return(ret)
@@ -60,15 +60,15 @@ viz_clinical_outcome <- function() {
   dat <- myphd::extract_cohort(dat, id_var = "HelixID")
   outcome <- vars_of_interest()$outcomes
   dat_sel <- dat |>
-    dplyr::select(
+    tidylog::select(
       cohort, 
       hs_age_years, e3_sex, 
       dplyr::all_of(outcome)
     ) |>
-    dplyr::mutate(
+    tidylog::mutate(
       hs_age_years = as.factor(round(hs_age_years, 0))
     ) |>
-    dplyr::rename(
+    tidylog::rename(
       age = hs_age_years, 
       sex = e3_sex, 
       hrt = {{outcome}}
@@ -112,25 +112,26 @@ viz_clinical_outcome <- function() {
 viz_desc_vars <- function(dat, vars, fct_levels) {
[…]

### 2023-09-08 output/paper/figures.qmd
Added details for paper if published in EHP. Still no line numbering in DOCX nor PDF.

@@
 ---
-title: "Paper 3 - Figures"
+title: "Figures"
 
@@
 
-<!-- ####################################################################### -->
+- Number figures according to the order in which they are first mentioned in the main text.
+- Clearly label all axes, giving both the measure and the unit of measurement.
+- Not larger than 7.5 inches in width.
+- Save and submit each main text figure as an individual file (PDF with fonts embedded). Include the figure number in the file name of each figure.
+- Provide main text figure captions on a new page of the main text after the tables.
+- Include a title for the entire figure and descriptors for each panel.
+
 # CTD Results
-<!-- ####################################################################### -->
 ```{r}
@@
 #| fig.width: 10
+
 path <- "results/files/CTD_chem_diseases_1680702221550_05042023.csv"

### 2023-09-08 output/paper/manuscript.qmd
Added details for paper if published in EHP. Still no line numbering in DOCX nor PDF.

--- 2023-08-01 output/paper/manuscript.qmd
@@
 format:
+  docx:
+    reference-doc: reference_document.docx
   pdf:
-    pdf-engine: xelatex
+    pdf-engine: lualatex
     number-sections: true
@@
         \usepackage{typearea}
-
-bibliography: ../../../../PhD.bib
-link-citations: true
-linkcolor: blue
-
-execute:
-  echo: false
-  warning: false
-  error: false
-  message: false
-editor_options: 
-  chunk_output_type: console
-
-filters:
-  - ../../../../all_phd/_extensions/acronyms/parse-acronyms.lua
-  - ../../../../all_phd/_extensions/authors-block/authors-block.lua
-acronyms:
-  fromfile: ../../../../all_phd/acronyms.yml
-  insert_links: false
-  insert_loa: false
-  sorting: alphabetical
+        \usepackage{lineno}
 
@@
  **Discussion**
+
+bibliography: bibliography.bib
+csl: ama11.csl
+
+link-citations: true
+linkcolor: blue
+
+editor_options: 
+  chunk_output_type: console
+
+filters:
+  - ../../../../all_phd/_extensions/acronyms/parse-acronyms.lua
+  - ../../../../all_phd/_extensions/authors-block/authors-block.lua
+  - ../../../../all_phd/_extensions/wordcount/citeproc.lua
+  - ../../../../all_phd/_extensions/wordcount/wordcount.lua
+
+acronyms:
+  fromfile: ../../../../all_phd/acronyms.yml
+  insert_links: false
+  insert_loa: false
+  sorting: alphabetical
+
+execute:
+  echo: false
+  warning: false
+  error: false
+  message: false
 ---
@@
 ---
-title: "Some Title"
+title: "Some Fancy Title"
 ---
 
-<!-- ####################################################################### -->
-# Introduction {#intro}
+- The title should be less or equal than 300 characters.
+It should indicate the study design, the subject of the paper, information regarding exposures and outcomes assessed, and whether the study was observational or experimental.
+- The suggested length of the abstract is less or equal than 300 words.
+- The suggested length is <7,000 words, excluding abstract, references, tables, figure captions, acknowledgments, and Supplementary Material.
+  - Concise sub-headings should be less or equal than 8 words, and they should be used to organize information rather than summarize the results.
+  - In-text citations with superscript numbers: outside periods and commas, but inside colons and semicolons.
+
+# Introduction {#sec-intro}
 ```{r}
@@
 ```
-<!-- ####################################################################### -->
 
-<!-- ####################################################################### -->
-# Methods {#methods}
+# Methods {#sec-methods}
 ```{r}
@@
 ```
-<!-- ####################################################################### -->
 
-<!-- ####################################################################### -->
-# Results {#res}
+# Results {#sec-res}
 ```{r}
@@
 ```
-<!-- ####################################################################### -->
 
-<!-- ####################################################################### -->
-# Discussion {#discussion}
+# Discussion {#sec-discussion}
 ```{r}
@@
 ```
-<!-- ####################################################################### -->
 
@@
 
-<!-- ####################################################################### -->
+# Funding
+
+- Give the source of funding and the role of the funders for the present study and, if applicable, for the original study on which the present article is based.
+
+\newpage
+
 # References {.unnumbered}
-<!-- ####################################################################### -->
 

### 2023-09-08 output/paper/sections/discussion.qmd
Added details for paper if published in EHP. Still no line numbering in DOCX nor PDF.

@@
-List of possible limitations:
+## Key results {#sec-disc-res-key}
 
-* The percentiles of the exposures might differ considerably by cohort. When using `avg_comparisons`, we do not consider this. This might cause cohort-dependent positivity issues (?).
+- Summarise key results with reference to study objectives.
+- Provide a review of the relevant literature to put the study findings into context.
+  - It should be complete and balanced, including inconsistent results.
+  - It should include, for each source, sufficient details: study design, sample size, population, specific exposures and outcomes.
+
+## Limitations {#sec-limitations}
+
+- Discuss limitations of the study, taking into account sources of potential bias or imprecision.
+- Discuss both direction and magnitude of any potential bias.
+
+## Interpretation {#sec-interpretation}
+
+- End with a summary of the key findings and their implications for the study hypotheses, future research, and policy.
+- Give a cautious overall interpretation of results considering objectives, limitations, multiplicity of analyses, results from similar studies, and other relevant evidence.
+
+## Generalisability {#sec-general}
+
+- Discuss the generalisability (external validity) of the study results.

### 2023-09-08 output/paper/sections/introduction.qmd
Added details for paper if published in EHP. Still no line numbering in DOCX nor PDF.

## Background and rationale {#sec-background}

- Brief review of the literature to summarize current knowledge.
  - Acknowledge inconsistencies.
  - For each study, indicate whether it was observational or experimental, and note key characteristics of study populations or experimental models.
- Explain the scientific background and rationale for the investigation being reported.
  - Identify knowledge gaps addressed by the current study.
- Provide context for the study: include information on exposures and outcomes, and why they are relevant to environmental health.

## Objectives {#sec-objectives}

- Provide a clear description of the study hypotheses/aims/objectives, and eventually an overview of the approach used to address them.

### 2023-09-08 output/paper/sections/methods.qmd
Added details for paper if published in EHP. Still no line numbering in DOCX nor PDF.

@@
-## Study population {#sec-study-population}
+## Study design {#sec-design}
 
-## Exposure assessment {#sec-exp-assessment}
+- Present key elements of study design
 
-## Metabolomics {#sec-metabolomics}
+## Setting {#sec-setting}
 
-## Neurodevelopment {#sec-neurodevelopment}
+- Describe the setting, locations, and relevant dates, including periods of recruitment, exposure, follow-up, and data collection.
 
-## Covariates {#sec-covars}
+## Participants {#sec-participants}
 
-## Statistical analysis {#sec-stat-analysis}
+- Cohort study: eligibility criteria, and the sources and methods of selection of participants. Describe methods of follow-up.
+- Cross-sectional study: give eligibility criteria, and the sources and methods of selection of participants.
+- Describe informed consent protocols.
+- Report how and by whom *race* or *ethnicity* was defined, and why this information was included in the study design. Disaggregate race and ethnicity data to the fullest extent possible.
+
+## Variables {#sec-vars}
+
+- Clearly define all outcomes, exposures, predictors, potential confounders, and effect modifiers.
+- Explain the rationale for treating race as an exposure, confounder, effect modifier, or other type of variable in analyses.
+
+### Confounders {#sec-confounders}
+
+### Endocrine disrupting chemicals {#sec-edcs}
+
+### Corticosteroids {#sec-steroids}
+
+### Neurodevelopment {#sec-neurodevelopment}
+
+## Data sources and measurement {#sec-dat-sources}
+
+- For each variable of interest, give sources of data and details of methods of assessment (measurement).
+
+## Bias {#sec-bias}
+
+- Describe any efforts to address potential sources of bias.
+
+## Study size {#sec-size}
+
+- Explain how the study size was arrived at.
+
+## Quantitative variables {#sec-quant-vars}
+
+- Explain how quantitative variables were handled in the analyses. If applicable, describe which groupings were chosen and why.
+
+## Statistical methods {#sec-stat-methods}
+
+- Methods for selecting potential confounders (provide DAGs).
+- Describe all statistical methods with assumptions, including those used to control for confounding.
+- Describe any methods used to examine subgroups and interactions.
+- Explain how missing data were addressed.
+- Cohort study: explain how loss to follow-up was addressed.
+- Cross-sectional study: describe analytical methods taking account of sampling strategy.
+- Describe any sensitivity analyses.
+- When referring to previous publications for methods' details, include a brief description of the approach, key assumptions and limitations, and any deviation.
+- Names and version numbers for the used software packages, including non-data arguments if deviating from the default ones.

### 2023-09-08 output/paper/sections/results.qmd
Added details for paper if published in EHP. Still no line numbering in DOCX nor PDF.

## Participants {#sec-res-participants}

- Report number of individuals at each stage of study: numbers potentially eligible, examined for eligibility, confirmed eligible, included in the study, completing follow-up, and analysed.
- Give reasons for non-participation at each stage.
- Consider use of a flow diagram.

## Descriptive data {#sec-res-pop-desc}

- Give characteristics of study participants (demographic, clinical, social) and information on exposures and potential confounders.
  - Include all covariates included in primary or secondary analyses.
  - Provide detailed information about exposures' distributions: minimum and maximum, percentiles, and number of observations above/below the assay's LOD and LOQ.
- Indicate number of participants with missing data for each variable of interest.
- Cohort study: summarise follow-up time (average and total amount).

## Outcome data {#sec-res-out-desc}

- Cohort study: report numbers of outcome events or summary measures over time.
- Cross-sectional study: report numbers of outcome events or summary measures.

## Main results {#sec-res-main}

- All results on which study conclusions or inferences are based, including null findings and results of secondary or sensitivity analyses, must be reported. Use of sub-headings that describe the nature of the results (but no declarative statements).
  - Provide a clear and concise description of all findings without extrapolating beyond the study results.
  - Do not limit results to those *statistically significant* or that support the study hypotheses. Avoid using statistical significance testing as the sole or primary criterion for interpreting the obtained results. If significance testing or *p*-values are used, report numeric *p*-values, rounded to 1-2 digits, for all results.
- Give unadjusted and confounder-adjusted estimates and their precision (e.g., $95\%$ confidence interval). Make clear which confounders were adjusted for and why they were included. Include the number of observations for each analysis after accounting for missing data. Include numeric data within figures (e.g., forest plots), or provide tables with corresponding numeric data for all figures.
- Report category boundaries when continuous variables were categorized.

## Other analyses {#sec-res-other}

- Report other analyses done (e.g., analyses of subgroups and interactions, and sensitivity analyses).

### 2023-09-08 output/paper/si.qmd
Added details for paper if published in EHP. Still no line numbering in DOCX nor PDF.

@@
 ---
-title: "Paper 3 - Supplementary Information"
+title: "Supplementary Material"
 
@@
     reference-doc: reference_document.docx
+
+bibliography: bibliography.bib
+csl: ama11.csl
 
@@
 ---
+
+- Should be limited to results of secondary analyses and background details needed to ensure transparency:
+  - Questionnaires.
+  - Tables and figures with results of sensitivity analyses.
+  - Tables with numeric data corresponding to results shown in the main text figures.
+  - DAGs used to select model covariates.
+- Text description of study methods should NOT go in the SM.
+- Order the content: text, supplemental tables, and supplemental figures.
+- Provide the title for each figure below the figure on the same manuscript page as the figure itself.
+- Provide the title for each table above the table and the notes below the table.
+- Number supplemental tables and figures separately according to the order in which they are first mentioned in the main text, using an "S" prefix.
+- Conclude the SM with a list of references, even if they are also cited in the main paper.

### 2023-09-08 output/paper/tables.qmd
Added details for paper if published in EHP. Still no line numbering in DOCX nor PDF.

@@
 ---
-title: "Paper 3 - Tables"
+title: "Tables"
 
@@
 
-<!-- ####################################################################### -->
+- Begin each table on a new page after the list of references.
+- Tables may not contain parts.
+- List footnotes after the general one, one per line and indicated with a lowercase italicized letter.
+
 # Study populations
-<!-- ####################################################################### -->
 ```{r}
@@
 
-<!-- ####################################################################### -->
 # Exposures
-<!-- ####################################################################### -->
 ```{r}
@@
 #| tbl-cap: Information about the non-persistent EDCs.
+
 edcs_info <- myphd::edcs_information() |>

### 2023-09-08 output/paper/run_paper.R

@@ -1,4 +1,4 @@
-Sys.setenv(is_hpc = TRUE)
+Sys.setenv(is_hpc = FALSE)
 store <- ifelse(
   Sys.getenv("is_hpc"), 
   "/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/lorenzoF/data/data_paper3/_targets/_targets_paper", 

### 2023-09-13 output/paper/figures.qmd
Working on structure of paper. Issues with cross-referencing tables in Quarto.

@@
     pdf-engine: lualatex
+    number-sections: true
     toc: false
@@
       text: |
+        \usepackage{lineno}
         \usepackage{rotating}
@@
         \usepackage{typearea}
+    include-before-body:
+      text: |
+        \linenumbers
 
@@
 linkcolor: blue
+
+editor_options: 
+  chunk_output_type: console
 
@@
   message: false
-editor_options: 
-  chunk_output_type: console
 ---
@@
 - Provide main text figure captions on a new page of the main text after the tables.
+- Provide the title for each figure below the figure on the same manuscript page as the figure itself.
 - Include a title for the entire figure and descriptors for each panel.
 
-# CTD Results
-```{r}
-#| label: fig-ctd-res
-#| fig.cap: Barplot CTD.
-#| fig.height: 10
-#| fig.width: 10
+# Participants
 
-path <- "results/files/CTD_chem_diseases_1680702221550_05042023.csv"
-ctd <- myphd::load_ctd(path = path, 
-                       filter_evidence = TRUE)
-myphd::plot_ctd(ctd)
-```
+- Flow diagram: number of individuals at each stage of study.
+Report numbers potentially eligible, examined for eligibility, confirmed eligible, included in the study, completing follow-up, and analysed.
+
+# Main results
+
+- Forest plot for RQ2 (group by chemical class, and for each chemical group bars for metabolites) and RQ3.

### 2023-09-13 output/paper/manuscript.qmd
Working on structure of paper. Issues with cross-referencing tables in Quarto.

@@
 ---
+title: "Some Fancy Title"
+
 metadata-files:
@@
     toc: false
+    keep-tex: true
     include-in-header:
       text: |
+        \usepackage{lineno}
         \usepackage{rotating}
@@
         \usepackage{typearea}
-        \usepackage{lineno}
+    include-before-body:
+      text: |
+        \linenumbers
 
-abstract: |
+bibliography: bibliography.bib
+csl: ama11.csl
+
+link-citations: true
+linkcolor: blue
+
+filters:
+  - authors-block
+  - ../../../../all_phd/_extensions/acronyms/parse-acronyms.lua
+  - ../../../../all_phd/_extensions/wordcount/citeproc.lua
+  - ../../../../all_phd/_extensions/wordcount/wordcount.lua
+
+acronyms:
+  fromfile: ../../../../all_phd/acronyms.yml
+  insert_links: false
+  insert_loa: false
+  sorting: alphabetical
+
+editor_options: 
+  chunk_output_type: console
+
+execute:
+  echo: false
+  warning: false
+  error: false
+  message: false
+---
+
+# Abstract {.unnumbered}
 **Background**
@@
 
-bibliography: bibliography.bib
-csl: ama11.csl
-
-link-citations: true
-linkcolor: blue
-
-editor_options: 
-  chunk_output_type: console
-
-filters:
-  - ../../../../all_phd/_extensions/acronyms/parse-acronyms.lua
-  - ../../../../all_phd/_extensions/authors-block/authors-block.lua
-  - ../../../../all_phd/_extensions/wordcount/citeproc.lua
-  - ../../../../all_phd/_extensions/wordcount/wordcount.lua
-
-acronyms:
-  fromfile: ../../../../all_phd/acronyms.yml
-  insert_links: false
-  insert_loa: false
-  sorting: alphabetical
-
-execute:
-  echo: false
-  warning: false
-  error: false
-  message: false
----
-
-```{=latex}
-\setcounter{tocdepth}{4}
-\tableofcontents
-```
-
----
-title: "Some Fancy Title"
----
+\newpage
 

### 2023-09-13 output/paper/manuscript.tex
Working on structure of paper. Issues with cross-referencing tables in Quarto.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{lineno}
\usepackage{rotating}
\usepackage{float}
\usepackage{typearea}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Some Fancy Title},
  pdfauthor={, and },
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Some Fancy Title}
\author{Lorenzo Fabbri\textsuperscript{1,2,*} \and Martine
Vrijheid\textsuperscript{1,2}}
\date{}

\begin{document}
\maketitle
\linenumbers

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, boxrule=0pt, breakable, sharp corners, frame hidden]}{\end{tcolorbox}}\fi

\textsuperscript{1} Barcelona Institute for Global Health (ISGlobal),
Barcelona, Spain\\
\textsuperscript{2} Pompeu Fabra University, Barcelona, Spain

\textsuperscript{*} Correspondence:
\href{mailto:lorenzo.fabbri@isglobal.org}{Lorenzo Fabbri
\textless{}lorenzo.fabbri@isglobal.org\textgreater{}}

\hypertarget{abstract}{%
\section*{Abstract}\label{abstract}}
\addcontentsline{toc}{section}{Abstract}

\textbf{Background}

\textbf{Objectives}

\textbf{Methods}

\textbf{Results}

\textbf{Discussion}

\newpage

\begin{itemize}
\tightlist
\item
  The title should be less or equal than 300 characters. It should
  indicate the study design, the subject of the paper, information
  regarding exposures and outcomes assessed, and whether the study was
  observational or experimental.
\item
  The suggested length of the abstract is less or equal than 300 words.
\item
  The suggested length is \textless7,000 words, excluding abstract,
  references, tables, figure captions, acknowledgments, and
  Supplementary Material.

  \begin{itemize}
  \tightlist
  \item
    Concise sub-headings should be less or equal than 8 words, and they
    should be used to organize information rather than summarize the
    results.
  \item
    In-text citations with superscript numbers: outside periods and
    commas, but inside colons and semicolons.
  \end{itemize}
\end{itemize}

\hypertarget{sec-intro}{%
\section{Introduction}\label{sec-intro}}

\hypertarget{sec-background}{%
\subsection{Background and rationale}\label{sec-background}}

\begin{itemize}
\tightlist
\item
  Brief review of the literature to summarize current knowledge.

  \begin{itemize}
  \tightlist
  \item
    Acknowledge inconsistencies.
  \item
    For each study, indicate whether it was observational or
    experimental, and note key characteristics of study populations or
    experimental models.
  \end{itemize}
\item
  Explain the scientific background and rationale for the investigation
  being reported.

  \begin{itemize}
  \tightlist
  \item
    Identify knowledge gaps addressed by the current study.
  \end{itemize}
\item
  Provide context for the study: include information on exposures and
  outcomes, and why they are relevant to environmental health.
\end{itemize}

\hypertarget{sec-objectives}{%
\subsection{Objectives}\label{sec-objectives}}

\begin{itemize}
\tightlist
\item
  Provide a clear description of the study hypotheses/aims/objectives,
  and eventually an overview of the approach used to address them.
\end{itemize}

\hypertarget{sec-methods}{%
\section{Methods}\label{sec-methods}}

\hypertarget{sec-design}{%
\subsection{Study design}\label{sec-design}}

\begin{itemize}
\tightlist
\item
  Present key elements of study design
\end{itemize}

\hypertarget{sec-setting}{%
\subsection{Setting}\label{sec-setting}}

\begin{itemize}
\tightlist
\item
  Describe the setting, locations, and relevant dates, including periods
  of recruitment, exposure, follow-up, and data collection.
\end{itemize}

\hypertarget{sec-participants}{%
\subsection{Participants}\label{sec-participants}}

\begin{itemize}
\tightlist
\item
  Cohort study: eligibility criteria, and the sources and methods of
  selection of participants. Describe methods of follow-up.
\item
  Cross-sectional study: give eligibility criteria, and the sources and
  methods of selection of participants.
\item
  Describe informed consent protocols.
\item
  Report how and by whom \emph{race} or \emph{ethnicity} was defined,
  and why this information was included in the study design.
  Disaggregate race and ethnicity data to the fullest extent possible.
\end{itemize}

\hypertarget{sec-vars}{%
\subsection{Variables}\label{sec-vars}}

\begin{itemize}
\tightlist
\item
  Clearly define all outcomes, exposures, predictors, potential
  confounders, and effect modifiers.
\item
  Explain the rationale for treating race as an exposure, confounder,
  effect modifier, or other type of variable in analyses.
\end{itemize}

\hypertarget{sec-confounders}{%
\subsubsection{Confounders}\label{sec-confounders}}

\hypertarget{sec-edcs}{%
\subsubsection{Endocrine disrupting chemicals}\label{sec-edcs}}

\hypertarget{sec-steroids}{%
\subsubsection{Corticosteroids}\label{sec-steroids}}

\hypertarget{sec-neurodevelopment}{%
\subsubsection{Neurodevelopment}\label{sec-neurodevelopment}}

\hypertarget{sec-dat-sources}{%
\subsection{Data sources and measurement}\label{sec-dat-sources}}

\begin{itemize}
\tightlist
\item
  For each variable of interest, give sources of data and details of
  methods of assessment (measurement).
\end{itemize}

\hypertarget{sec-bias}{%
\subsection{Bias}\label{sec-bias}}

\begin{itemize}
\tightlist
\item
  Describe any efforts to address potential sources of bias.
\end{itemize}

\hypertarget{sec-size}{%
\subsection{Study size}\label{sec-size}}

\begin{itemize}
\tightlist
\item
  Explain how the study size was arrived at.
\end{itemize}

\hypertarget{sec-quant-vars}{%
\subsection{Quantitative variables}\label{sec-quant-vars}}

\begin{itemize}
\tightlist
\item
  Explain how quantitative variables were handled in the analyses. If
  applicable, describe which groupings were chosen and why.
\end{itemize}

\hypertarget{sec-stat-methods}{%
\subsection{Statistical methods}\label{sec-stat-methods}}

\begin{itemize}
\tightlist
\item
  Methods for selecting potential confounders (provide DAGs).
\item
  Describe all statistical methods with assumptions, including those
  used to control for confounding.

  \begin{itemize}
  \tightlist
  \item
    Description of outcome model, weighting method, estimand, and
    balance assessment.
  \item
    Description of method used to estimate effects (e.g.,
    g-computation).
  \item
    Description of method used for SE and CI.
  \end{itemize}
\item
  Describe any methods used to examine subgroups and interactions.
\item
  Explain how missing data were addressed.
\item
  Cohort study: explain how loss to follow-up was addressed.
\item
  Cross-sectional study: describe analytical methods taking account of
  sampling strategy.
\item
  Describe any sensitivity analyses.
\item
  When referring to previous publications for methods' details, include
  a brief description of the approach, key assumptions and limitations,
  and any deviation.
\item
  Names and version numbers for the used software packages, including
  non-data arguments if deviating from the default ones.
\end{itemize}

\hypertarget{sec-res}{%
\section{Results}\label{sec-res}}

\hypertarget{sec-res-participants}{%
\subsection{Participants}\label{sec-res-participants}}

\begin{itemize}
\tightlist
\item
  Give reasons for non-participation at each stage.
\end{itemize}

\hypertarget{sec-res-pop-desc}{%
\subsection{Descriptive data}\label{sec-res-pop-desc}}

\hypertarget{sec-res-out-desc}{%
\subsection{Outcome data}\label{sec-res-out-desc}}

\hypertarget{sec-res-main}{%
\subsection{Main results}\label{sec-res-main}}

\begin{itemize}
\tightlist
\item
  All results on which study conclusions or inferences are based,
  including null findings and results of secondary or sensitivity
  analyses, must be reported. Use of sub-headings that describe the
  nature of the results (but no declarative statements).

  \begin{itemize}
  \tightlist
  \item
    Provide a clear and concise description of all findings without
    extrapolating beyond the study results.
  \item
    Do not limit results to those \emph{statistically significant} or
    that support the study hypotheses. Avoid using statistical
    significance testing as the sole or primary criterion for
    interpreting the obtained results. If significance testing or
    \emph{p}-values are used, report numeric \emph{p}-values, rounded to
    1-2 digits, for all results.
  \end{itemize}
\item
  Give unadjusted and confounder-adjusted estimates and their precision
  (e.g., \(95\%\) confidence interval). Make clear which confounders
  were adjusted for and why they were included. Include the number of
  observations for each analysis after accounting for missing data.
  Include numeric data within figures (e.g., forest plots), or provide
  tables with corresponding numeric data for all figures.

  \begin{itemize}
  \tightlist
  \item
    \href{https://vincentarelbundock.github.io/marginaleffects/articles/tables.html}{\texttt{marginaleffects}
    tables}.
  \end{itemize}
\item
  Report category boundaries when continuous variables were categorized.
\end{itemize}

\hypertarget{sec-res-other}{%
\subsection{Other analyses}\label{sec-res-other}}

\begin{itemize}
\tightlist
\item
  Report other analyses done (e.g., analyses of subgroups and
  interactions, and sensitivity analyses).
\end{itemize}

\textbf{?@tbl-chem-info}

\hypertarget{sec-discussion}{%
\section{Discussion}\label{sec-discussion}}

\hypertarget{sec-disc-res-key}{%
\subsection{Key results}\label{sec-disc-res-key}}

\begin{itemize}
\tightlist
\item
  Summarise key results with reference to study objectives.
\item
  Provide a review of the relevant literature to put the study findings
  into context.

  \begin{itemize}
  \tightlist
  \item
    It should be complete and balanced, including inconsistent results.
  \item
    It should include, for each source, sufficient details: study
    design, sample size, population, specific exposures and outcomes.
  \end{itemize}
\end{itemize}

\hypertarget{sec-limitations}{%
\subsection{Limitations}\label{sec-limitations}}

\begin{itemize}
\tightlist
\item
  Discuss limitations of the study, taking into account sources of
  potential bias or imprecision.
\item
  Discuss both direction and magnitude of any potential bias.
\end{itemize}

\hypertarget{sec-interpretation}{%
\subsection{Interpretation}\label{sec-interpretation}}

\begin{itemize}
\tightlist
\item
  End with a summary of the key findings and their implications for the
  study hypotheses, future research, and policy.
\item
  Give a cautious overall interpretation of results considering
  objectives, limitations, multiplicity of analyses, results from
  similar studies, and other relevant evidence.
\end{itemize}

\hypertarget{sec-general}{%
\subsection{Generalisability}\label{sec-general}}

\begin{itemize}
\tightlist
\item
  Discuss the generalisability (external validity) of the study results.
\end{itemize}

\newpage

\hypertarget{funding}{%
\section{Funding}\label{funding}}

\begin{itemize}
\tightlist
\item
  Give the source of funding and the role of the funders for the present
  study and, if applicable, for the original study on which the present
  article is based.
\end{itemize}

\newpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\end{CSLReferences}

\end{document}

### 2023-09-13 output/paper/si.qmd
Working on structure of paper. Issues with cross-referencing tables in Quarto.

@@
 
+editor_options: 
+  chunk_output_type: console
+
 execute:
@@
 - Text description of study methods should NOT go in the SM.
-- Order the content: text, supplemental tables, and supplemental figures.
+- Number supplemental tables and figures separately according to the order in which they are first mentioned in the main text, using an "S" prefix.
+
+# Supplementary information
+
+# Supplementary tables
+
+- Provide the title for each table above the table and the notes below the table.
+
+## Descriptive data
+
+### Study populations
+
+- Detailed characteristics of study participants (demographic, clinical, social), by cohort and overall.
+
+# Supplementary figures
+
 - Provide the title for each figure below the figure on the same manuscript page as the figure itself.
-- Provide the title for each table above the table and the notes below the table.
-- Number supplemental tables and figures separately according to the order in which they are first mentioned in the main text, using an "S" prefix.
-- Conclude the SM with a list of references, even if they are also cited in the main paper.
+
+## Main results
+
+- ADRF and AMEF for RQ2 and RQ3.
+- Whole distribution (i.e., density) of e.g., risk differences (instead of marginalized estimates), with lines for mean and median.
+
+# References {.unnumbered}
+
+::: {#refs}
+:::

### 2023-09-13 output/paper/tables.qmd
Working on structure of paper. Issues with cross-referencing tables in Quarto.

@@
 
+tbl-cap-location: top
 link-citations: true
 linkcolor: blue
+
+editor_options: 
+  chunk_output_type: console
 
@@
 - Tables may not contain parts.
+- Provide the title for each table above the table and the notes below the table.
 - List footnotes after the general one, one per line and indicated with a lowercase italicized letter.
 
-# Study populations
-```{r}
+# Descriptive data
 
-```
+## Study populations
 
-# Exposures
+- Basic characteristics of study participants (demographic, clinical, social), by cohort and overall.
+  - Include all covariates included in primary or secondary analyses.
+  - Indicate number of participants with missing data for each variable of interest.
+  - Report numbers of outcome events or summary measures.
+
+## Endocrine disruptors
+
+- Levels of unprocessed biomarkers (chemicals), with IDs.
+- Report: minimum and maximum, percentiles, and number of observations above/below the assay's LOD and LOQ.
+
 ```{r}
 #| label: tbl-chem-info
-#| tbl-cap: Information about the non-persistent EDCs.
+#| tbl-cap: "Information about the non-persistent EDCs."
 
@@
   tidylog::select(-c(chem_id, smiles))
-gt::gt(edcs_info, groupname_col = "class") |>
+gt::gt(edcs_info,
+       groupname_col = "class") |>
+  gt::tab_header(
+    title = "Information about the non-persistent EDCs."
+  ) |>
   gt::cols_label(
@@
 ```
+
+## Corticosteroids
+
+- Levels of unprocessed biomarkers (metabolites), with IDs.
+
+# Main results
+
+- RQ2: group chemicals by class, and for each chemical report results with metabolites (`avg_comparisons` with estimate, SE, s-value, CI).
+- RQ3: report results with outcome (`avg_comparisons` with estimate, SE, s-value, CI).
+
+# Other analyses

### 2023-09-13 output/paper/manuscript.tex
Tried pre-release of Quarto 1.4 to fix issue with cross-ref of tables but it fails to render PDF. Went back to stable release.

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, boxrule=0pt, breakable, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, boxrule=0pt, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 
-\textbf{?@tbl-chem-info}
-
 \hypertarget{sec-discussion}{%

### 2023-09-19 output/paper/manuscript.tex
Nothing.

--- 2023-09-13 output/paper/manuscript.tex
@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, boxrule=0pt, breakable, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, interior hidden, sharp corners, boxrule=0pt, breakable, enhanced]}{\end{tcolorbox}}\fi
 

### 2023-10-03 output/paper/sections/methods.qmd
Working on structure of paper. Issues with cross-referencing tables in Quarto.
Just checking pipeline works.

@@
 - Describe all statistical methods with assumptions, including those used to control for confounding.
-- Describe any methods used to examine subgroups and interactions.
+  - Description of outcome model, weighting method, estimand, and balance assessment.
+  - Description of method used to estimate effects (e.g., g-computation).
+  - Description of method used for SE and CI.
+- Describe any methods used to examine subgroups and interactions (sub-group analysis or moderation analysis or analysis of effect-modification).
 - Explain how missing data were addressed.

### 2023-10-03 code/data.R

@@ -18,7 +18,7 @@ create_steroid_scores <- function(dat) {
         X6OHF, X5a20acortol, X5a20bcortol, X5b20acortol, X5b20bcortol
       ))) / `F`,
       cortisone_production = rowSums(dplyr::across(c(
-        E, X20aDHE, X20bDHE, X5aTHE, X5bTHE, X6OHE,
+        CortisoneE, X20aDHE, X20bDHE, X5aTHE, X5bTHE, X6OHE,
         X5b20acortolone, X5b20bcortolone
       ))),
       X11bHSD = cortisone_production / cortisol_production

### 2023-10-03 code/dictionaries.R

@@ -103,7 +103,8 @@ params <- function(is_hpc) {
     ),
     categorical = c(
       "e3_sex", "h_ethnicity_spiro",
-      "cohort"
+      "cohort",
+      "hs_sample_c"
     )
   )
 
@@ -136,7 +137,7 @@ params <- function(is_hpc) {
       preproc_exposures = list(
         missings = list(
           threshold_within = 40,
-          threshold_overall = 40,
+          threshold_overall = 30,
           use_additional_covariates = FALSE,
           selected_covariates = NULL,
           method_imputation = "vim.knn",
@@ -162,7 +163,7 @@ params <- function(is_hpc) {
         ),
         missings = list(
           threshold_within = 40,
-          threshold_overall = 40,
+          threshold_overall = 30,
           use_additional_covariates = FALSE,
           selected_covariates = NULL,
           method_imputation = "vim.knn",
@@ -183,7 +184,7 @@ params <- function(is_hpc) {
       preproc_covars = list(
         missings = list(
           threshold_within = 40,
-          threshold_overall = 40,
+          threshold_overall = 30,
           use_additional_covariates = FALSE,
           selected_covariates = NULL,
           method_imputation = "vim.knn",
@@ -205,7 +206,7 @@ params <- function(is_hpc) {
         ),
         missings = list(
           threshold_within = 40,
-          threshold_overall = 40,
+          threshold_overall = 30,
           use_additional_covariates = FALSE,
           selected_covariates = NULL,
           method_imputation = "vim.knn",
@@ -223,7 +224,7 @@ params <- function(is_hpc) {
       preproc_outcome = list(
         missings = list(
           threshold_within = 40,
-          threshold_overall = 40,
+          threshold_overall = 30,
           use_additional_covariates = FALSE,
           selected_covariates = NULL,
           method_imputation = "vim.knn",
@@ -237,7 +238,7 @@ params <- function(is_hpc) {
       preproc_covars = list(
         missings = list(
           threshold_within = 40,
-          threshold_overall = 40,
+          threshold_overall = 30,
           use_additional_covariates = FALSE,
           selected_covariates = NULL,
           method_imputation = "vim.knn",

### 2023-10-03 code/utils.R

@@ -188,6 +188,18 @@ rq_load_data <- function(res_dag) {
       params_dat$variables[[rq]]$outcome
     )
 
+  ## Description of metabolites
+  if (rq == "rq2") {
+    mets_to_select <- params_dat$variables[[rq]]$outcome
+  } else if (rq == "rq3") {
+    mets_to_select <- params_dat$variables[[rq]]$exposures
+  }
+  dat$metab_desc <- dat$metab_desc |>
+    tidylog::select(
+      params_dat$variables$identifier,
+      dplyr::all_of(mets_to_select)
+    )
+
   ## Covariates
   if (length(res_dag$adjustment_sets) > 1) {
     adj_set <- select_adjustment_set(
@@ -249,7 +261,7 @@ rq_prepare_data <- function(dat) {
   steps_covars <- params_dat$steps[[rq]]$preproc_covars
   steps_exposures <- params_dat$steps[[rq]]$preproc_exposures
   steps_outcome <- params_dat$steps[[rq]]$preproc_outcome
-  
+
   # Create folders to store results
   invisible(lapply(c("figures", "tables"), function(x) {
     path_save_res <- paste0(
@@ -288,6 +300,15 @@ rq_prepare_data <- function(dat) {
     by_var = "cohort"
   ) |>
     dplyr::select(-cohort)
+  ## Eventually add scores and pre-process them
+  if (rq == "rq3") {
+    cols_to_remove <- setdiff(
+      colnames(dat$exposures),
+      params_dat$variables$identifier
+    )
+    dat$exposures <- create_steroid_scores(dat = dat$exposures) |>
+      dplyr::select(-dplyr::all_of(cols_to_remove))
+  }
 
   # Process outcome
   dat$outcome <- myphd::preproc_data(
@@ -307,6 +328,15 @@ rq_prepare_data <- function(dat) {
     by_var = "cohort"
   ) |>
     dplyr::select(-cohort)
+  ## Eventually add scores and pre-process them
+  if (rq == "rq2") {
+    cols_to_remove <- setdiff(
+      colnames(dat$outcome),
+      params_dat$variables$identifier
+    )
+    dat$outcome <- create_steroid_scores(dat = dat$outcome) |>
+      dplyr::select(-dplyr::all_of(cols_to_remove))
+  }
 
   return(dat)
 } # End function rq_prepare_data
[…]

### 2023-10-05 DAGs/dag_v2.R
Modified pipelines to handle ratios as outcome or exposure correctly.

@@ -7,17 +7,18 @@ dags <- function() {
     breastfeeding ~ ethnicity_mother + maternalSEP_preg + paternalSEP_preg,
     bw ~ cohort + ethnicity_child + ethnicity_mother + gestational_age + maternalAlcohol_preg + maternalSEP_preg + maternalSmoking_preg + paternalSEP_preg,
     characteristics_child ~ age_child + bw + child_diet + child_smoking + cohort + ethnicity_child + ethnicity_mother + familySEP + gestational_age + maternalAlcohol_preg + maternalDiet_preg + maternalSEP_preg + maternalSmoking_preg + paternalSEP_preg + sex_child,
-    chemical ~ characteristics_child + child_diet + cohort + creatinine + ethnicity_child + familySEP + season_visit + sex_child + time_lastMeal,
+    chemical ~ type_sample + characteristics_child + child_diet + cohort + creatinine + ethnicity_child + familySEP + season_visit + sex_child + time_lastMeal,
     child_diet ~ cohort + ethnicity_child + ethnicity_mother + familySEP + sex_child,
     child_smoking ~ ethnicity_child + familySEP + sex_child,
-    creatinine ~ age_child + characteristics_child + child_smoking + ethnicity_child + familySEP + sex_child,
+    creatinine ~ age_child + characteristics_child + child_smoking + ethnicity_child + familySEP + sex_child + cohort + type_sample,
+    type_sample ~ age_child + sex_child,
     familySEP ~ ethnicity_mother + maternalSEP_preg + paternalSEP_preg,
     maternalAlcohol_preg ~ ethnicity_mother + maternalSEP_preg + paternalSEP_preg,
     maternalDiet_preg ~ ethnicity_mother + maternalSEP_preg + paternalSEP_preg,
     maternalSEP_preg ~ ethnicity_mother,
     maternalSmoking_preg ~ ethnicity_mother + maternalSEP_preg + paternalSEP_preg,
-    neuropsychologicalDiagnosis_child ~ maternalSEP_preg + sex_child,
-    outcome ~ age_child + biomarker + breastfeeding + characteristics_child + chemical + child_diet + child_smoking + cohort + creatinine + envFactors_visit + ethnicity_child + ethnicity_mother + familySEP + maternalAlcohol_preg + maternalDiet_preg + maternalSEP_preg + maternalSmoking_preg + neuropsychologicalDiagnosis_child + paternalSEP_preg + sex_child,
[…]

### 2023-10-05 _targets_rqX.R

@@ -1,7 +1,6 @@
 source("DAGs/dag_v2.R")
 source("code/dictionaries.R")
 source("code/data.R")
-source("code/mtps.R")
 source("code/utils.R")
 
 targets::tar_option_set(

### 2023-10-05 code/dictionaries.R

@@ -104,13 +104,14 @@ params <- function(is_hpc) {
     categorical = c(
       "e3_sex", "h_ethnicity_spiro",
       "cohort",
-      "hs_sample_c"
+      "hs_sample_c",
+      "hs_tob"
     )
   )
 
   variables <- list(
     identifier = "HelixID",
-    strategy_select_adj_set = "smallest",
+    strategy_select_adj_set = "minimize_missings",
     ############################################################################
     rq1 = list(
       outcome = clinical_outcomes,
@@ -142,14 +143,14 @@ params <- function(is_hpc) {
           selected_covariates = NULL,
           method_imputation = "vim.knn",
           k = 5
-        ),
-        creatinine = list(
-          method = "cas",
-          method_fit_args = list(family = gaussian(link = "identity")),
-          creatinine_covariates_names = creatinine_covariates_names,
-          creatinine_name = "hs_creatinine_cg",
-          path_save_res = "results/figures/rq2/model_check_creatinine_exp.png"
         )
+        # creatinine = list(
+        #   method = "cas",
+        #   method_fit_args = list(family = gaussian(link = "identity")),
+        #   creatinine_covariates_names = creatinine_covariates_names,
+        #   creatinine_name = "hs_creatinine_cg",
+        #   path_save_res = "results/figures/rq2/model_check_creatinine_exp.png"
+        # )
       ),
       # End preproc_exposures
       preproc_outcome = list(
@@ -168,17 +169,14 @@ params <- function(is_hpc) {
           selected_covariates = NULL,
           method_imputation = "vim.knn",
           k = 5
-        ),
-        creatinine = list(
-          method = "cas",
-          method_fit_args = list(family = gaussian(link = "identity")),
-          creatinine_covariates_names = creatinine_covariates_names,
-          creatinine_name = "hs_creatinine_cg",
-          path_save_res = "results/figures/rq2/model_check_creatinine_out.png"
-        ),
-        transform = list(
-          transformation_fun = log
         )
+        # creatinine = list(
+        #   method = "cas",
+        #   method_fit_args = list(family = gaussian(link = "identity")),
+        #   creatinine_covariates_names = creatinine_covariates_names,
+        #   creatinine_name = "hs_creatinine_cg",
+        #   path_save_res = "results/figures/rq2/model_check_creatinine_out.png"
+        # )
       ),
       # End preproc_outcome
       preproc_covars = list(
@@ -211,14 +209,14 @@ params <- function(is_hpc) {
[…]

### 2023-10-05 code/utils.R

@@ -83,7 +83,8 @@ select_adjustment_set <- function(dat, meta, res_dag, strategy) {
     "minimize_missings" = all_as[[myphd::minimize_missings(
       dat = dat,
       meta = meta,
-      adjustment_sets = all_as
+      adjustment_sets = all_as,
+      by_var = "cohort"
     )]]
   )
 
@@ -308,6 +309,14 @@ rq_prepare_data <- function(dat) {
     )
     dat$exposures <- create_steroid_scores(dat = dat$exposures) |>
       dplyr::select(-dplyr::all_of(cols_to_remove))
+    
+    # Pre-process scores
+    dat$exposures <- myphd::handle_transformation(
+      dat = dat$exposures,
+      id_var = params_dat$variables$identifier,
+      by_var = "cohort",
+      transformation_fun = log
+    )
   }
 
   # Process outcome
@@ -332,9 +341,21 @@ rq_prepare_data <- function(dat) {
   if (rq == "rq2") {
     cols_to_remove <- setdiff(
       colnames(dat$outcome),
-      params_dat$variables$identifier
+      c(params_dat$variables$identifier,
+        "F", "cortisol_production")
     )
-    dat$outcome <- create_steroid_scores(dat = dat$outcome) |>
+    dat$outcome <- create_steroid_scores(dat = dat$outcome)
+    
+    # Pre-process scores
+    dat$outcome <- myphd::handle_transformation(
+      dat = dat$outcome,
+      id_var = params_dat$variables$identifier,
+      by_var = "cohort",
+      transformation_fun = log
+    )
+    
+    # Remove single metabolites
+    dat$outcome <- dat$outcome |>
       dplyr::select(-dplyr::all_of(cols_to_remove))
   }
 
@@ -624,13 +645,23 @@ rq_fit_model_weighted <- function(dat, outcome,
     )
     return()
   }
-
-  # Eventually remove covariates that should not be included in outcome models
-  if (rq == "rq3") {
+  
+  # Eventually add control for denominator of outcome
+  if (rq == "rq2") {
+    den <- switch (outcome,
+      "cortisol_metabolism" = "F",
+      "X11bHSD" = "cortisol_production",
+      NULL
+    )
     dat$covariates <- dat$covariates |>
-      dplyr::select(-dplyr::any_of(c(
-        "hs_creatinine_cg", "hs_sample_c"
-      )))
+      dplyr::full_join(
+        dat$outcome |>
+          dplyr::select(
[…]

### 2023-10-09 _targets_rq0X.R
Trying raw targets.

@@ -9,136 +9,229 @@ targets::tar_option_set(
 )
 
 params_dag <- list(
-  type = "minimal", 
+  type = "minimal",
   effect = "total"
 )
 id_var <- "HelixID"
 by_var <- "cohort"
-exposure <- switch(Sys.getenv("TAR_PROJECT"), 
-                   "rq01" = "chemical", 
-                   "rq02" = "chemical", 
-                   "rq03" = "biomarker")
-outcome <- switch(Sys.getenv("TAR_PROJECT"), 
-                  "rq01" = "outcome", 
-                  "rq02" = "biomarker", 
-                  "rq03" = "outcome")
+exposure <- switch(Sys.getenv("TAR_PROJECT"),
+  "rq01" = "chemical",
+  "rq02" = "chemical",
+  "rq03" = "biomarker"
+)
+outcome <- switch(Sys.getenv("TAR_PROJECT"),
+  "rq01" = "outcome",
+  "rq02" = "biomarker",
+  "rq03" = "outcome"
+)
 
 list(
-  targets::tar_target(
-    name = dag, 
-    command = load_dag(dags = dags(), 
-                       exposure = exposure, 
-                       outcome = outcome, 
-                       params_dag = params_dag)
-  ), # End dag target
-  targets::tar_target(
-    name = load_dat, 
-    command = rq_load_data(res_dag = dag)
-  ), # End load_dat target
-  targets::tar_target(
-    name = viz_desc_chems, 
-    command = viz_desc_vars(
-      dat = load_dat_request()$dat, 
-      vars = vars_of_interest(append_to_chem = "cdesc")$chemicals, 
-      fct_levels = c(1, 2, 3, 4), 
-      is_chem = TRUE
+  targets::tar_target_raw(
+    name = paste0(Sys.getenv("TAR_PROJECT"), "_dag"),
+    command = expression(
+      load_dag(
+        dags = dags(),
+        exposure = exposure,
+        outcome = outcome,
+        params_dag = params_dag
+      )
+    )
+  ),
+  # End dag target
+  targets::tar_target_raw(
+    name = paste0(Sys.getenv("TAR_PROJECT"), "_load_dat"),
+    command = substitute(
+      rq_load_data(res_dag = dag),
+      env = list(dag = as.symbol(paste0(Sys.getenv("TAR_PROJECT"), "_dag")))
+    )
+  ),
+  # End load_dat target
+  targets::tar_target_raw(
[…]

### 2023-10-09 _targets_rqX.R

@@ -62,7 +62,7 @@ list(
   targets::tar_target(
     name = weights, 
     command = rq_estimate_weights(dat = preproc_dat, 
-                                  save_results = TRUE, 
+                                  save_results = FALSE, 
                                   parallel = FALSE, 
                                   workers = 10)
   ), # End weights target

### 2023-10-09 code/utils.R

@@ -238,6 +238,10 @@ rq_load_data <- function(res_dag) {
     ))
   
   ## Metadata
+  warning(
+    "Add meta to dat.",
+    call. = TRUE
+  )
   dat$meta <- dat_request$meta
 
   return(dat)

### 2023-10-14 output/paper/manuscript.tex
Just tried to compile paper.

@@
 \makeatother
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
-  \fi
-  % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 \KOMAoption{captions}{tableheading}
-\makeatletter
-\makeatother
-\makeatletter
-\makeatother
 \makeatletter
@@
 \makeatletter
+\makeatother
+\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
 \@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
-\makeatother
-\makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
-\makeatother
-\makeatletter
-\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
 \makeatother
@@
 \hypersetup{
-  pdftitle={Some Fancy Title},
+  pdftitle={Some Title},
   pdfauthor={, and },
@@
 
-\title{Some Fancy Title}
+\title{Some Title}
 \author{Lorenzo Fabbri\textsuperscript{1,2,*} \and Martine
@@
 \linenumbers
-
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, interior hidden, sharp corners, boxrule=0pt, breakable, enhanced]}{\end{tcolorbox}}\fi
 
@@
 
-\hypertarget{abstract}{%
-\section*{Abstract}\label{abstract}}
+\section*{Abstract}\label{abstract}
 \addcontentsline{toc}{section}{Abstract}
@@
 
-\hypertarget{sec-intro}{%
-\section{Introduction}\label{sec-intro}}
+\section{Introduction}\label{sec-intro}
 
-\hypertarget{sec-background}{%
-\subsection{Background and rationale}\label{sec-background}}
+\subsection{Background and rationale}\label{sec-background}
 
@@
 
-\hypertarget{sec-objectives}{%
-\subsection{Objectives}\label{sec-objectives}}
+\subsection{Objectives}\label{sec-objectives}
 
@@
 
-\hypertarget{sec-methods}{%
-\section{Methods}\label{sec-methods}}
+\section{Methods}\label{sec-methods}
 
-\hypertarget{sec-design}{%
-\subsection{Study design}\label{sec-design}}
+\subsection{Study design}\label{sec-design}
 
@@
 
-\hypertarget{sec-setting}{%
-\subsection{Setting}\label{sec-setting}}
+\subsection{Setting}\label{sec-setting}
 
@@
 
-\hypertarget{sec-participants}{%
-\subsection{Participants}\label{sec-participants}}
+\subsection{Participants}\label{sec-participants}
 
@@
 
-\hypertarget{sec-vars}{%
-\subsection{Variables}\label{sec-vars}}
+\subsection{Variables}\label{sec-vars}
 
@@
 
-\hypertarget{sec-confounders}{%
-\subsubsection{Confounders}\label{sec-confounders}}
+\subsubsection{Confounders}\label{sec-confounders}
 
-\hypertarget{sec-edcs}{%
-\subsubsection{Endocrine disrupting chemicals}\label{sec-edcs}}
+\subsubsection{Endocrine disrupting chemicals}\label{sec-edcs}
 
-\hypertarget{sec-steroids}{%
-\subsubsection{Corticosteroids}\label{sec-steroids}}
+\subsubsection{Corticosteroids}\label{sec-steroids}
 
-\hypertarget{sec-neurodevelopment}{%
-\subsubsection{Neurodevelopment}\label{sec-neurodevelopment}}
+\subsubsection{Neurodevelopment}\label{sec-neurodevelopment}
 
-\hypertarget{sec-dat-sources}{%
-\subsection{Data sources and measurement}\label{sec-dat-sources}}
+\subsection{Data sources and measurement}\label{sec-dat-sources}
 
@@
 
-\hypertarget{sec-bias}{%
-\subsection{Bias}\label{sec-bias}}
+\subsection{Bias}\label{sec-bias}
 
@@
 
-\hypertarget{sec-size}{%
-\subsection{Study size}\label{sec-size}}
+\subsection{Study size}\label{sec-size}
 
@@
 
-\hypertarget{sec-quant-vars}{%
-\subsection{Quantitative variables}\label{sec-quant-vars}}
+\subsection{Quantitative variables}\label{sec-quant-vars}
 
@@
 
-\hypertarget{sec-stat-methods}{%
-\subsection{Statistical methods}\label{sec-stat-methods}}
+\subsection{Statistical methods}\label{sec-stat-methods}
 
@@
 \item
-  Describe any methods used to examine subgroups and interactions.
+  Describe any methods used to examine subgroups and interactions
+  (sub-group analysis or moderation analysis or analysis of
+  effect-modification).
 \item
@@
 
-\hypertarget{sec-res}{%
-\section{Results}\label{sec-res}}
+\section{Results}\label{sec-res}
 
-\hypertarget{sec-res-participants}{%
-\subsection{Participants}\label{sec-res-participants}}
+\subsection{Participants}\label{sec-res-participants}
 
@@
 
-\hypertarget{sec-res-pop-desc}{%
-\subsection{Descriptive data}\label{sec-res-pop-desc}}
+\subsection{Descriptive data}\label{sec-res-pop-desc}
 
-\hypertarget{sec-res-out-desc}{%
-\subsection{Outcome data}\label{sec-res-out-desc}}
+\subsection{Outcome data}\label{sec-res-out-desc}
 
-\hypertarget{sec-res-main}{%
-\subsection{Main results}\label{sec-res-main}}
+\subsection{Main results}\label{sec-res-main}
 
@@
 
-\hypertarget{sec-res-other}{%
-\subsection{Other analyses}\label{sec-res-other}}
+\subsection{Other analyses}\label{sec-res-other}
 
@@
 
-\hypertarget{sec-discussion}{%
-\section{Discussion}\label{sec-discussion}}
+\section{Discussion}\label{sec-discussion}
 
-\hypertarget{sec-disc-res-key}{%
-\subsection{Key results}\label{sec-disc-res-key}}
+\subsection{Key results}\label{sec-disc-res-key}
 
@@
 
-\hypertarget{sec-limitations}{%
-\subsection{Limitations}\label{sec-limitations}}
+\subsection{Limitations}\label{sec-limitations}
 
@@
 
-\hypertarget{sec-interpretation}{%
-\subsection{Interpretation}\label{sec-interpretation}}
+Some limitations:
+
+\begin{itemize}
+\tightlist
+\item
+  Cross-sectional study.
+\item
+  Chemicals measured in night and morning samples, whereas metabolites
+  (the outcome) were measured only in night samples.
+\item
+  Cortisol measured at night, when should be lowest.
+\item
+  Change of estimand when trimming weights.
+\item
+  Model misspecification.
+\item
+  Mixtures effect.
+\item
+  Residual confounding.
+\item
+  Some confounders were not used since large percentage of missing
+  values.
+\item
+  Multiple comparisons.
+\end{itemize}
+
+\subsection{Interpretation}\label{sec-interpretation}
 
@@
 
-\hypertarget{sec-general}{%
-\subsection{Generalisability}\label{sec-general}}
+\subsection{Generalisability}\label{sec-general}
 
@@
 
-\hypertarget{funding}{%
-\section{Funding}\label{funding}}
+\section{Funding}\label{funding}
 
@@
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
+\section*{References}\label{references}
 \addcontentsline{toc}{section}{References}
 
-\hypertarget{refs}{}
-\begin{CSLReferences}{0}{0}
-\end{CSLReferences}
-
 \end{document}

### 2023-10-16 output/paper/tables.qmd
Tried pre-release of Quarto 1.4 to fix issue with cross-ref of tables but it fails to render PDF. Went back to stable release.
Modified pipelines to handle ratios as outcome or exposure correctly.
Trying to solve cross ref with gt and docx.
How the fuck do I cross ref across files?

--- 2023-09-08 output/paper/tables.qmd
@@
 
+editor_options:
+  chunk_output_type: console
+
 execute:
@@
   tidylog::select(-c(chem_id, smiles))
-gt::gt(edcs_info, groupname_col = "class") |>
+gt::gt(edcs_info,
+       groupname_col = "class") |>
   gt::cols_label(
@@
 ```
+
+## Corticosteroids
+
+- Levels of unprocessed biomarkers (metabolites), with IDs.
+- Table comparing subjects with and without measurement of steroids.
+
+# Main results
+
+- RQ2: group chemicals by class, and for each chemical report results with metabolites (`avg_comparisons` with estimate, SE, s-value, CI).
+- RQ3: report results with outcome (`avg_comparisons` with estimate, SE, s-value, CI).
+
+# Other analyses

### 2023-10-16 output/paper/manuscript.qmd
Nothing.
Just tried to compile paper.
How the fuck do I cross ref across files?
I think I fixed the cross-referencing with external files and tables at the end.

@@
 ---
-title: "Some Fancy Title"
+title: "Some Title"
 
@@
 # Introduction {#sec-intro}
-```{r}
-#| child: output/paper/sections/introduction.qmd
-```
+
+{{< include sections/introduction.qmd >}}
 
 # Methods {#sec-methods}
-```{r}
-#| child: output/paper/sections/methods.qmd
-```
+
+{{< include sections/methods.qmd >}}
 
 # Results {#sec-res}
-```{r}
-#| child: output/paper/sections/results.qmd
-```
+
+{{< include sections/results.qmd >}}
 
 # Discussion {#sec-discussion}
-```{r}
-#| child: output/paper/sections/discussion.qmd
-```
+
+{{< include sections/discussion.qmd >}}
 
@@
 :::
+
+{{< embed tables.qmd >}}

### 2023-10-16 output/paper/manuscript.tex
How the fuck do I cross ref across files?
I think I fixed the cross-referencing with external files and tables at the end.

@@
 
+Table~\ref{tbl-chem-info}
+
 \subsection{Outcome data}\label{sec-res-out-desc}
@@
 \end{itemize}
-
-\textbf{?@tbl-chem-info}
 
@@
 
+\section{Tables}\label{tables}
+
+\begin{itemize}
+\tightlist
+\item
+  Begin each table on a new page after the list of references.
+\item
+  Tables may not contain parts.
+\item
+  Provide the title for each table above the table and the notes below
+  the table.
+\item
+  List footnotes after the general one, one per line and indicated with
+  a lowercase italicized letter.
+\end{itemize}
+
+\section{Descriptive data}\label{descriptive-data}
+
+\subsection{Study populations}\label{study-populations}
+
+\begin{itemize}
+\tightlist
+\item
+  Basic characteristics of study participants (demographic, clinical,
+  social), by cohort and overall.
+
+  \begin{itemize}
+  \tightlist
+  \item
+    Include all covariates included in primary or secondary analyses.
+  \item
+    Indicate number of participants with missing data for each variable
+    of interest.
+  \item
+    Report numbers of outcome events or summary measures.
+  \end{itemize}
+\end{itemize}
+
+\subsection{Endocrine disruptors}\label{endocrine-disruptors}
+
+\begin{itemize}
+\tightlist
+\item
+  Levels of unprocessed biomarkers (chemicals), with IDs.
+\item
+  Report: minimum and maximum, percentiles, and number of observations
+  above/below the assay's LOD and LOQ.
+\end{itemize}
+
+\begin{table}
+
+\caption{\label{tbl-chem-info}Information about the non-persistent
+EDCs.}
+
+\centering{
+
+\begin{longtable}[]{@{}llllll@{}}
+\toprule\noalign{}
+Compound & Symbol & PubChem CID & CTD ID & Exposome Explore ID &
+Parental compound \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+OP pesticide metabolites & & & & & \\
+diethyl dithiophosphate & DEDTP & 9274 & C000654497 & NA & NA \\
+diethyl phosphate & DEP & 654 & C034789 & NA & NA \\
+diethyl thiophosphate & DETP & 3683036 & C035638 & NA & NA \\
+dimethyl dithiophosphate & DMDTP & NA & Unsure & NA & NA \\
+dimethyl phosphate & DMP & 13134 & C007477 & NA & NA \\
+dimethyl thiophosphate & DMTP & 168140 & C040340 & NA & NA \\
+Phenols & & & & & \\
+bisphenol A & BPA & 6623 & C006780 & 1418 & NA \\
+ethyl-paraben & ETPA & 8434 & C012313 & 1422 & NA \\
+methyl-paraben & MEPA & 7456 & C015358 & 1421 & NA \\
+n‑butyl‑paraben & BUPA & 7184 & C038091 & 1424 & NA \\
+oxybenzone & OXBE & 4632 & C005290 & 1419 & NA \\
+propyl-paraben & PRPA & 7175 & C006068 & 1423 & NA \\
+triclosan & TRCS & 5564 & D014260 & 1420 & NA \\
+Phthalate metabolites & & & & & \\
+mono benzyl phthalate & MBzP & 31736 & C103325 & 1397 & BzBP \\
+mono‑2‑ethyl 5‑carboxypentyl phthalate & MECPP & 148386 & C051450 & 1403
+& DEHP \\
+mono‑2‑ethyl‑5‑hydroxyhexyl phthalate & MEHHP & 170295 & C479069 & 1402
+& DEHP \\
+mono‑2‑ethyl‑5‑oxohexyl phthalate & MEOHP & 119096 & C080276 & 1401 &
+DEHP \\
+mono‑2‑ethylhexyl phthalate & MEHP & 21924291 & C016599 & Unsure &
+DEHP \\
+mono‑4‑methyl‑7‑hydroxyoctyl phthalate & oh-MiNP & 102401880 & NA & 1451
+& NA \\
+mono‑4‑methyl‑7‑oxooctyl phthalate & oxo-MiNP & 102401881 & NA & 1492 &
+NA \\
+mono‑iso‑butyl phthalate & MiBP & 92272 & C575690 & 1399 & DiBP \\
+mono‑n‑butyl phthalate & MnBP & 8575 & C028577 & 1398 & NA \\
+monoethyl phthalate & MEP & 75318 & C581825 & 1396 & DEP \\
+\end{longtable}
+
+}
+
+\end{table}%
+
+\subsection{Corticosteroids}\label{corticosteroids}
+
+\begin{itemize}
+\tightlist
+\item
+  Levels of unprocessed biomarkers (metabolites), with IDs.
+\item
+  Table comparing subjects with and without measurement of steroids.
+\end{itemize}
+
+\section{Main results}\label{main-results}
+
+\begin{itemize}
+\tightlist
+\item
+  RQ2: group chemicals by class, and for each chemical report results
+  with metabolites (\texttt{avg\_comparisons} with estimate, SE,
+  s-value, CI).
+\item
+  RQ3: report results with outcome (\texttt{avg\_comparisons} with
+  estimate, SE, s-value, CI).
+\end{itemize}
+
+\section{Other analyses}\label{other-analyses}
+
 \end{document}

### 2023-10-16 output/paper/sections/results.qmd
I think I fixed the cross-referencing with external files and tables at the end.

@@
 
-- Report number of individuals at each stage of study: numbers potentially eligible, examined for eligibility, confirmed eligible, included in the study, completing follow-up, and analysed.
 - Give reasons for non-participation at each stage.
-- Consider use of a flow diagram.
 
@@
 
-- Give characteristics of study participants (demographic, clinical, social) and information on exposures and potential confounders.
-  - Include all covariates included in primary or secondary analyses.
-  - Provide detailed information about exposures' distributions: minimum and maximum, percentiles, and number of observations above/below the assay's LOD and LOQ.
-- Indicate number of participants with missing data for each variable of interest.
-- Cohort study: summarise follow-up time (average and total amount).
+@tbl-chem-info
 
 ## Outcome data {#sec-res-out-desc}
-
-- Cohort study: report numbers of outcome events or summary measures over time.
-- Cross-sectional study: report numbers of outcome events or summary measures.
 
@@
 Include numeric data within figures (e.g., forest plots), or provide tables with corresponding numeric data for all figures.
+  - [`marginaleffects` tables](https://vincentarelbundock.github.io/marginaleffects/articles/tables.html).
 - Report category boundaries when continuous variables were categorized.

### 2023-10-24 _targets_raw_rqX.R
Tested pipeline w/ vector of id_val.

@@ -64,7 +64,7 @@ list(
     name = paste0(rq, "_load_dat"),
     command = substitute(
       rq_load_data(res_dag = dag,
-                   remove_kanc = TRUE),
+                   remove_kanc = FALSE),
       env = list(dag = as.symbol(paste0(rq, "_dag")))
     )
   ),
@@ -84,7 +84,7 @@ list(
     command = substitute(
       rq_estimate_weights(
         dat = dat,
-        save_results = FALSE,
+        save_results = TRUE,
         parallel = FALSE,
         workers = 10
       ),
@@ -116,11 +116,11 @@ list(
       name = paste0(rq, "_marginal"),
       command = substitute(
         rq_estimate_marginal_effects(
-          fits = fits$fits,
+          fits = all_fits$fits,
           parallel = TRUE,
           workers = 3
         ),
-        env = list(fits = as.symbol(paste0(rq, "_weighted_fits")))
+        env = list(all_fits = as.symbol(paste0(rq, "_weighted_fits")))
       )
     ) # End marginal target
   ) # End loop over outcomes

### 2023-10-24 _targets_rqX.R

@@ -58,7 +58,7 @@ list(
   targets::tar_target(
     name = load_dat, 
     command = rq_load_data(res_dag = dag,
-                           remove_kanc = TRUE)
+                           remove_kanc = FALSE)
   ), # End load_dat target
   ##############################################################################
   targets::tar_target(

### 2023-10-24 code/data.R

@@ -154,7 +154,7 @@ load_steroids <- function() {
     dd <- dd |>
       dplyr::mutate(
         # Creatinine to g/L as in HELIX, instead of mcrM/mL.
-        Creatinine = Creatinine * 1e-6 * 113.12,
+        creatinine_to_helix = Creatinine * 1e-6 * 113.12,
       )
     ############################################################################
 
@@ -204,7 +204,7 @@ load_steroids <- function() {
     msg = "Mismatched number of rows in metabolomics and description data."
   )
   assertthat::assert_that(
-    ncol(metabs) == ncol(cdescs),
+    ncol(metabs) == ncol(cdescs) + 1,
     msg = "Mismatched number of columns in metabolomics and description data."
   )
 

### 2023-10-24 code/dictionaries.R

@@ -65,8 +65,8 @@ vars_of_interest <- function(append_to_chem = NULL) {
     "Etio"
   )
   outcomes <- c(
-    "hs_hitrtse",
-    "hs_dcolors3"
+    "hs_hitrtse"
+    #"hs_dcolors3"
   )
 
   return(list(
@@ -207,7 +207,7 @@ params <- function(is_hpc) {
       # End preproc_exposures
       preproc_outcome = list(
         llodq = list(
-          id_val = 2,
+          id_val = c(2, 4),
           method = "replace",
           creatinine_threshold = NULL,
           threshold_within = threshold_within_lod,
@@ -247,7 +247,7 @@ params <- function(is_hpc) {
     rq3 = list(
       preproc_exposures = list(
         llodq = list(
-          id_val = 2,
+          id_val = c(2, 4),
           method = "replace",
           creatinine_threshold = NULL,
           threshold_within = threshold_within_lod,

### 2023-10-24 code/results.R

@@ -8,7 +8,7 @@ num_digits_est <- 3
 num_digits_sig <- 3
 ################################################################################
 
-for (rq in c("2", "3")) {
+for (rq in c("1", "2", "3")) {
   # Exploratory
   targets::tar_load_everything(
     store = paste0(path_store, paste0("0", rq))
@@ -16,6 +16,16 @@ for (rq in c("2", "3")) {
   
   # Analyses
   targets::tar_load_everything(
+    store = paste0(path_store, rq)
+  )
+}
+
+for (rq in c("1", "2", "3")) {
+  # Marginal effects
+  targets::tar_load(
+    dplyr::contains(paste0(
+      "rq", rq, "_marginal_"
+    )),
     store = paste0(path_store, rq)
   )
 }
@@ -34,10 +44,14 @@ sensitivity_analyses <- function(which_sens, fits, rq) {
 ################################################################################
 
 tidy_res_weighted_fits <- function(res_list, rq) {
-  names_ <- gsub("weighted_fits_", "", names(res_list))
+  names_ <- gsub(paste0("rq", rq, "_weighted_fits_"),
+                 "",
+                 names(res_list))
   
   weights_ <- lapply(1:1, function(idx) {
-    outcome <- gsub("weighted_fits_", "", names_[idx])
+    outcome <- gsub(paste0("rq", rq, "_weighted_fits_"),
+                    "",
+                    names_[idx])
     tmp <- lapply(res_list[[idx]]$fits, "[[", "weights")
     tmp <- lapply(seq_along(tmp), function(idx2) {
       ret <- tibble::tibble(tmp[[idx2]])
@@ -77,7 +91,7 @@ tidy_res_weighted_fits <- function(res_list, rq) {
     ) |>
     dplyr::select(-name)
   
-  if (rq == "2") {
+  if (rq %in% c("1", "2")) {
     info_edcs <- myphd::edcs_information() |>
       tibble::as_tibble()
     weights_ <- weights_ |>
@@ -158,12 +172,10 @@ tidy_res_weighted_fits <- function(res_list, rq) {
 }
 ################################################################################
 
-tidy_res_marginaleffects <- function(rq) {
+tidy_res_meffects <- function(marginal_effects, rq) {
+  #marginal_effects <- mget(ls(pattern = "marginal_*"))
   names_ <- gsub("marginal_", "",
-                 ls(pattern = "marginal_*"))
-  marginal_effects <- mget(
[…]

### 2023-10-24 run_loop.R

@@ -1,4 +1,4 @@
-Sys.setenv(is_hpc = FALSE)
+Sys.setenv(is_hpc = TRUE)
 # Custom path to _targets for different research questions
 path_store <- ifelse(
   Sys.getenv("is_hpc"), 
@@ -16,10 +16,10 @@ Sys.setenv(path_store_res = path_store_res)
 
 for (rq in c("1", "2", "3")) {
   # Exploratory
-  Sys.setenv(TAR_PROJECT = paste0("rq0", rq))
-  store <- paste0(path_store, paste0("0", rq))
-  targets::tar_make(script = "_targets_rq0X.R", 
-                    store = store)
+  # Sys.setenv(TAR_PROJECT = paste0("rq0", rq))
+  # store <- paste0(path_store, paste0("0", rq))
+  # targets::tar_make(script = "_targets_rq0X.R", 
+  #                   store = store)
   # Analyses
   Sys.setenv(TAR_PROJECT = paste0("rq", rq))
   store <- paste0(path_store, rq)

### 2023-10-26 output/paper/figures.qmd
Modofied code to plot results of weights and marginal effects. Saving them to file too.

@@
 title: "Figures"
-
 format:
@@
         \linenumbers
-
 link-citations: true
 linkcolor: blue
-
 editor_options: 
   chunk_output_type: console
-
 execute:
@@
 - Clearly label all axes, giving both the measure and the unit of measurement.
-- Not larger than 7.5 inches in width.
-- Save and submit each main text figure as an individual file (PDF with fonts embedded). Include the figure number in the file name of each figure.
-- Provide main text figure captions on a new page of the main text after the tables.
-- Provide the title for each figure below the figure on the same manuscript page as the figure itself.
-- Include a title for the entire figure and descriptors for each panel.
+- Smallest type size: 8pt.
+- Maximum width: 7 inches in portrait, 9 inches in landscape mode.
+- Multi-panel figures: capital letters followed by closing parenthesis, completely above and to the left of the panel.
+- Use log-scale when plotting relative measures of effects.
+- Figure titles and legends should be submitted as single paragraph.
+- The text of figure legends typed after the reference list.
+- Save and submit each main text figure as an individual file (PDF with fonts embedded).
 

### 2023-10-26 code/plot.R

@@ -1,3 +1,342 @@
+#' Title
+#'
+#' @param res_list 
+#' @param rq 
+#'
+#' @return
+#' @export
+tidy_res_weighted_fits <- function(res_list, rq) {
+  names_ <- gsub(paste0("rq", rq, "_weighted_fits_"),
+                 "",
+                 names(res_list))
+  
+  weights_ <- lapply(1:1, function(idx) {
+    outcome <- gsub(paste0("rq", rq, "_weighted_fits_"),
+                    "",
+                    names_[idx])
+    tmp <- lapply(res_list[[idx]]$fits, "[[", "weights")
+    tmp <- lapply(seq_along(tmp), function(idx2) {
+      ret <- tibble::tibble(tmp[[idx2]])
+      colnames(ret) <- names(tmp)[idx2]
+      return(ret)
+    }) |>
+      dplyr::bind_cols()
+    tmp <- tmp |>
+      dplyr::mutate(
+        outcome = names_[[idx]]
+      ) |>
+      dplyr::relocate(outcome)
+  }) |> # End loop extract balancing weights
+    dplyr::bind_rows()
+  weights_wide <- weights_
+  
+  # Tidy data
+  weights_ <- weights_ |>
+    dplyr::mutate(
+      outcome = gsub("_", " ", outcome)
+    ) |>
+    tidyr::pivot_longer(
+      cols = -c(outcome)
+    ) |>
+    dplyr::mutate(
+      variable = name,
+      outcome = replace(
+        outcome, outcome == "x11bhsd", "11bhsd"
+      ),
+      variable = replace(
+        variable, variable == "x11bhsd", "11bhsd"
+      ),
+      variable = gsub("hs_", "", variable),
+      variable = gsub("_c", "", variable),
+      variable = gsub("_", " ", variable),
+      outcome = gsub("_", " ", outcome)
+    ) |>
+    dplyr::select(-name)
+  
+  if (rq %in% c("1", "2")) {
+    info_edcs <- myphd::edcs_information() |>
+      tibble::as_tibble()
+    weights_ <- weights_ |>
+      dplyr::left_join(
+        info_edcs |>
+          dplyr::select(chem_id, class),
+        by = c("variable" = "chem_id")
+      )
+  } else {
+    weights_ <- weights_ |>
+      dplyr::mutate(
+        class = c("TMP")
+      )
+  }
+  
+  # Plot distribution of weights
+  plot <- weights_ |>
+    ggplot2::ggplot(
+      ggplot2::aes(
+        x = value,
[…]

### 2023-10-26 run_loop.R

@@ -16,10 +16,13 @@ Sys.setenv(path_store_res = path_store_res)
 
 for (rq in c("1", "2", "3", "4")) {
   # Exploratory
-  # Sys.setenv(TAR_PROJECT = paste0("rq0", rq))
-  # store <- paste0(path_store, paste0("0", rq))
-  # targets::tar_make(script = "_targets_rq0X.R", 
-  #                   store = store)
+  if (rq != "4") {
+    Sys.setenv(TAR_PROJECT = paste0("rq0", rq))
+    store <- paste0(path_store, paste0("0", rq))
+    targets::tar_make(script = "_targets_rq0X.R",
+                      store = store)
+  }
+  
   # Analyses
   Sys.setenv(TAR_PROJECT = paste0("rq", rq))
   store <- paste0(path_store, rq)

### 2023-10-27 output/paper/tables.qmd
Kept working on functions to process results. Now table 1 for paper and cleaned previous results.

@@
 
-- Begin each table on a new page after the list of references.
+- Begin each table on a new page after the list of references, numbered with Arabic numerals.
+- Leave blank spaces for no entry.
 - Tables may not contain parts.
 - Provide the title for each table above the table and the notes below the table.
-- List footnotes after the general one, one per line and indicated with a lowercase italicized letter.
+- Order footnotes: abbreviations listed alphabetically and separated by semicolons; others preceded by superscript lowercase letter.
 

### 2023-10-27 DAGs/dag_v2.R

@@ -18,7 +18,7 @@ dags <- function() {
     maternalSEP_preg ~ ethnicity_mother,
     maternalSmoking_preg ~ ethnicity_mother + maternalSEP_preg + paternalSEP_preg,
     neuropsychologicalDiagnosis_child ~ ethnicity_mother + maternalSEP_preg + sex_child + ethnicity_child + breastfeeding + bw + gestational_age + maternalAlcohol_preg + maternalDiet_preg + maternalSmoking_preg + paternalSEP_preg,
-    outcome ~ age_child + biomarker + breastfeeding + characteristics_child + chemical + child_diet + child_smoking + cohort + envFactors_visit + ethnicity_child + ethnicity_mother + familySEP + maternalAlcohol_preg + maternalDiet_preg + maternalSEP_preg + maternalSmoking_preg + neuropsychologicalDiagnosis_child + paternalSEP_preg + sex_child,
+    outcome ~ creatinine + age_child + biomarker + breastfeeding + characteristics_child + chemical + child_diet + child_smoking + cohort + envFactors_visit + ethnicity_child + ethnicity_mother + familySEP + maternalAlcohol_preg + maternalDiet_preg + maternalSEP_preg + maternalSmoking_preg + neuropsychologicalDiagnosis_child + paternalSEP_preg + sex_child,
     qualityTesting_child ~ envFactors_visit,
     exposure = "chemical",
     outcome = "outcome"
@@ -42,7 +42,7 @@ dags <- function() {
     maternalSEP_preg ~ ethnicity_mother,
     maternalSmoking_preg ~ ethnicity_mother + maternalSEP_preg + paternalSEP_preg,
     neuropsychologicalDiagnosis_child ~ ethnicity_mother + maternalSEP_preg + sex_child + ethnicity_child + breastfeeding + bw + gestational_age + maternalAlcohol_preg + maternalDiet_preg + maternalSmoking_preg + paternalSEP_preg,
-    outcome ~ age_child + biomarker + breastfeeding + characteristics_child + chemical + child_diet + child_smoking + cohort + envFactors_visit + ethnicity_child + ethnicity_mother + familySEP + maternalAlcohol_preg + maternalDiet_preg + maternalSEP_preg + maternalSmoking_preg + neuropsychologicalDiagnosis_child + paternalSEP_preg + sex_child,
[…]

### 2023-10-27 _targets_rq0X.R

@@ -64,7 +64,16 @@ list(
       viz_desc_vars(
         dat = myphd::extract_cohort(load_steroids()$desc,
           id_var = id_var
-        ),
+        ) |>
+          tidylog::mutate(
+            cohort = dplyr::case_when(
+              cohort == "EDE" ~ "EDEN",
+              cohort == "KAN" ~ "KANC",
+              cohort == "MOB" ~ "MOBA",
+              cohort == "RHE" ~ "RHEA",
+              .default = cohort
+            )
+          ),
         vars = vars_of_interest()$metabolites |>
           paste0("_cdesc"),
         fct_levels = c(1, 2, 3, 4),
@@ -155,7 +164,19 @@ list(
       myphd::describe_data(
         dat = myphd::extract_cohort(dat$exposures,
           id_var = id_var
-        ),
+        ) |>
+          tidylog::mutate(
+            cohort = dplyr::case_when(
+              cohort == "EDE" ~ "EDEN",
+              cohort == "KAN" ~ "KANC",
+              cohort == "MOB" ~ "MOBA",
+              cohort == "RHE" ~ "RHEA",
+              .default = cohort
+            )
+          ) |>
+          tidylog::rename_with(
+            .fn = \(x) gsub("hs_|_c", "", x)
+          ),
         id_var = id_var,
         by_var = by_var
       ),
@@ -170,7 +191,19 @@ list(
       myphd::describe_data(
         dat = myphd::extract_cohort(dat$outcome,
           id_var = id_var
-        ),
+        ) |>
+          tidylog::mutate(
+            cohort = dplyr::case_when(
+              cohort == "EDE" ~ "EDEN",
+              cohort == "KAN" ~ "KANC",
+              cohort == "MOB" ~ "MOBA",
+              cohort == "RHE" ~ "RHEA",
+              .default = cohort
+            )
+          ) |>
+          tidylog::rename_with(
+            .fn = \(x) gsub("^X", "", x)
+          ),
         id_var = id_var,
         by_var = by_var
       ),

### 2023-10-27 code/data.R

@@ -152,7 +152,7 @@ load_steroids <- function() {
     
     ## Convert units of creatinine. Note: ng/mL = mcrg/L (metabolites).
     dd <- dd |>
-      dplyr::mutate(
+      tidylog::mutate(
         # Creatinine to g/L as in HELIX, instead of mcrM/mL.
         creatinine_to_helix = Creatinine * 1e-6 * 113.12,
       )
@@ -307,19 +307,19 @@ load_dat_request <- function() {
         \(x) factor(x)
       ),
       # Correct variable for ethnicity of the children
-      h_ethnicity_c = factor(
-        h_ethnicity_c,
-        levels = c(
-          "African",
-          "Asian",
-          "Caucasian",
-          "Native American",
-          "Other",
-          "Pakistani",
-          "White non-European"
-        ),
-        labels = c(1, 2, 3, 4, 5, 6, 7)
-      ),
+      # h_ethnicity_c = factor(
+      #   h_ethnicity_c,
+      #   levels = c(
+      #     "African",
+      #     "Asian",
+      #     "Caucasian",
+      #     "Native American",
+      #     "Other",
+      #     "Pakistani",
+      #     "White non-European"
+      #   ),
+      #   labels = c(1, 2, 3, 4, 5, 6, 7)
+      # ),
       e3_edum = factor(e3_edum,
         levels = c(
           "primary school",

### 2023-10-27 code/plot.R

@@ -22,7 +22,7 @@ tidy_res_weighted_fits <- function(res_list, rq) {
     }) |>
       dplyr::bind_cols()
     tmp <- tmp |>
-      dplyr::mutate(
+      tidylog::mutate(
         outcome = names_[[idx]]
       ) |>
       dplyr::relocate(outcome)
@@ -32,13 +32,13 @@ tidy_res_weighted_fits <- function(res_list, rq) {
   
   # Tidy data
   weights_ <- weights_ |>
-    dplyr::mutate(
+    tidylog::mutate(
       outcome = gsub("_", " ", outcome)
     ) |>
     tidyr::pivot_longer(
       cols = -c(outcome)
     ) |>
-    dplyr::mutate(
+    tidylog::mutate(
       variable = name,
       outcome = replace(
         outcome, outcome == "x11bhsd", "11bhsd"
@@ -51,20 +51,20 @@ tidy_res_weighted_fits <- function(res_list, rq) {
       variable = gsub("_", " ", variable),
       outcome = gsub("_", " ", outcome)
     ) |>
-    dplyr::select(-name)
+    tidylog::select(-name)
   
   if (rq %in% c("1", "2")) {
     info_edcs <- myphd::edcs_information() |>
       tibble::as_tibble()
     weights_ <- weights_ |>
-      dplyr::left_join(
+      tidylog::left_join(
         info_edcs |>
-          dplyr::select(chem_id, class),
+          tidylog::select(chem_id, class),
         by = c("variable" = "chem_id")
       )
   } else {
     weights_ <- weights_ |>
-      dplyr::mutate(
+      tidylog::mutate(
         class = c("TMP")
       )
   }
@@ -111,7 +111,7 @@ tidy_res_weighted_fits <- function(res_list, rq) {
     "hs_|_c", "", colnames(weights_wide)
   )
   weights_wide <- weights_wide |>
-    dplyr::select(-outcome)
+    tidylog::select(-outcome)
   tbl <- c("{median} ({p25}, {p75})", "{min}, {max}") |>
     purrr::map(
       ~ weights_wide |>
@@ -172,12 +172,12 @@ tidy_res_meffects <- function(marginal_effects, rq) {
     if (length(x$marginal_effects) == 0) return(NULL)
     df <- lapply(x$marginal_effects, "[[", "comparisons") |>
       dplyr::bind_rows() |>
-      dplyr::mutate(
+      tidylog::mutate(
         outcome = outcome,
[…]

### 2023-10-27 code/results.R

@@ -1,5 +1,6 @@
 source("DAGs/dag_v2.R")
 source("code/dictionaries.R")
+source("code/data.R")
 source("code/plot.R")
 
 is_hpc = TRUE
@@ -9,7 +10,140 @@ path_store <- ifelse(
   "~/mounts/rstudioserver/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/lorenzoF/data/data_paper3/_targets/_targetsRQ"
 )
 num_digits_est <- 3
-num_digits_sig <- 3
+num_digits_sig <- 2
+################################################################################
+
+# Table 1 (description population and outcome)
+tbl_desc_pop <- function(path_store) {
+  # Load data request (all variables)
+  params_dat <- params(is_hpc = Sys.getenv("is_hpc"))
+  dat_request <- load_dat_request()
+  dat_request$dat <- dat_request$dat |>
+    tidylog::mutate(
+      cohort = dplyr::case_when(
+        cohort == "EDE" ~ "EDEN",
+        cohort == "KAN" ~ "KANC",
+        cohort == "MOB" ~ "MOBA",
+        cohort == "RHE" ~ "RHEA",
+        .default = cohort
+      )
+    )
+  ## Handle time variables
+  cols_to_season <- c("hs_date_neu", "e3_cbirth")
+  dat_request$dat <- myphd::convert_time_season(
+      dat = dat_request$dat,
+      cols = cols_to_season
+    ) |>
+    tidylog::mutate(dplyr::across(
+      dplyr::any_of(cols_to_season),
+      \(x) factor(x)
+    ))
+  
+  # Extract confounders by RQ
+  mapped_adj_sets <- list()
+  for (rq in c("1", "2", "3")) {
+    targets::tar_load(
+      paste0("rq", rq, "_load_dat"),
+      store = paste0(path_store, rq)
+    )
+    mapped_adj_sets[[paste0("rq",
+                            rq)]] <- get(
+                              paste0("rq", rq, "_load_dat")
+                            )$mapping_covariates
+  }
+  
+  # Select confounders and clinical outcome in data request
+  all_confounders <- mapped_adj_sets |>
+    unlist() |>
+    unname() |>
+    unique()
+  df <- dat_request$dat |>
+    tidylog::select(
+      HelixID,
+      params_dat$variables$rq1$outcome,
+      dplyr::any_of(all_confounders)
+    )
+  ## Must add creatinine from IMIM and remove chemicals
+  df <- df |>
+    tidylog::left_join(
[…]

### 2023-10-27 code/utils.R

@@ -336,7 +336,7 @@ rq_prepare_data <- function(dat) {
     id_var = params_dat$variables$identifier,
     by_var = "cohort"
   ) |>
-    dplyr::select(-cohort)
+    tidylog::select(-cohort)
   ## Eventually add scores and pre-process them
   if (rq %in% c("rq3", "rq4")) {
     cols_to_remove <- setdiff(
@@ -344,7 +344,7 @@ rq_prepare_data <- function(dat) {
       params_dat$variables$identifier
     )
     dat$exposures <- create_steroid_scores(dat = dat$exposures) |>
-      dplyr::select(-dplyr::all_of(cols_to_remove))
+      tidylog::select(-dplyr::all_of(cols_to_remove))
     
     # Pre-process scores
     dat$exposures <- myphd::handle_transformation(
@@ -372,7 +372,7 @@ rq_prepare_data <- function(dat) {
     id_var = params_dat$variables$identifier,
     by_var = "cohort"
   ) |>
-    dplyr::select(-cohort)
+    tidylog::select(-cohort)
   ## Eventually add scores and pre-process them
   if (rq == "rq2") {
     cols_to_remove <- setdiff(
@@ -392,7 +392,7 @@ rq_prepare_data <- function(dat) {
     
     # Remove single metabolites
     dat$outcome <- dat$outcome |>
-      dplyr::select(-dplyr::all_of(cols_to_remove))
+      tidylog::select(-dplyr::all_of(cols_to_remove))
   }
   
   # Process metadata
@@ -613,9 +613,9 @@ rq_fit_model_weighted <- function(dat, outcome,
       NULL
     )
     dat$covariates <- dat$covariates |>
-      dplyr::full_join(
+      tidylog::full_join(
         dat$outcome |>
-          dplyr::select(
+          tidylog::select(
             HelixID,
             dplyr::all_of(den)
           ),

### 2023-11-07 output/paper/figures.qmd
Modified pipeline so that targets for processing results do not actually produce eg plots but only ready to use tibbles for plots and tables. Tested pipeline to create paper but I think I still have to manually reference tables and figures in manuscript.

@@
 
+```{r}
+source("code/proc_res.R")
+```
+
+```{r vars}
+store <- ifelse(
+  Sys.getenv("is_hpc"), 
+  paste0(
+    "/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/",
+    "lorenzoF/data/data_paper3/_targets/_targets_res"
+  ), 
+  "~/mounts/rstudioserver/data/data_paper3/_targets/_targets_res"
+)
+
+num_digits_est = 3
+num_digits_sig = 2
+```
+
 - Number figures according to the order in which they are first mentioned in the main text.
@@
 - Forest plot for RQ2 (group by chemical class, and for each chemical group bars for metabolites) and RQ3.
+
+## Marginal comparisons
+
+```{r}
+targets::tar_load(
+  dplyr::matches("marginal_comparisons_(\\d)$"),
+  store = store
+)
+```
+
+```{r marginal-1}
+#| label: fig-marginal-1
+#| fig-cap: Average comparisons
+#| fig-width: 6
+#| fig-height: 7
+
+tidy_res_meffects(
+  df = marginal_comparisons_1,
+  sa_var = NULL,
+  which_res = "comparisons",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)$plot
+```
+
+```{r marginal-2}
+#| label: fig-marginal-2
+#| fig-cap: Average comparisons
+#| fig-width: 10
+#| fig-height: 10
+
+tidy_res_meffects(
+  df = marginal_comparisons_2,
+  sa_var = NULL,
+  which_res = "comparisons",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)$plot
+```
+
+```{r marginal-3}
+#| label: fig-marginal-3
+#| fig-cap: Average comparisons
+#| fig-width: 5
+#| fig-height: 5
+
+tidy_res_meffects(
+  df = marginal_comparisons_3,
+  sa_var = NULL,
+  which_res = "comparisons",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)$plot
+```

### 2023-11-07 output/paper/si.qmd
Modified pipeline so that targets for processing results do not actually produce eg plots but only ready to use tibbles for plots and tables. Tested pipeline to create paper but I think I still have to manually reference tables and figures in manuscript.

@@
     reference-doc: reference_document.docx
+
+crossref:
+  fig-title: Supplementary Figure
+  tbl-title: Supplementary Table
 
@@
 
+```{r}
+source("code/proc_res.R")
+```
+
+```{r vars}
+store <- ifelse(
+  Sys.getenv("is_hpc"), 
+  paste0(
+    "/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/",
+    "lorenzoF/data/data_paper3/_targets/_targets_res"
+  ), 
+  "~/mounts/rstudioserver/data/data_paper3/_targets/_targets_res"
+)
+
+num_digits_est = 3
+num_digits_sig = 2
+```
+
 - Should be limited to results of secondary analyses and background details needed to ensure transparency:
@@
 # Supplementary information
+
+## Directed Acyclic Graphs
+
+```{r dag-1}
+targets::tar_load("all_dags", store = store)
+cat(all_dags$chem_to_out)
+```
+
+```{r dag-2}
+cat(all_dags$chem_to_marker)
+```
+
+```{r dag-3}
+cat(all_dags$marker_to_out)
+```
 
@@
 
+```{r codebook-1}
+#| label: tbl-codebook-1
+#| tbl-cap: Codebook for covariates used in RQ1
+
+targets::tar_load(
+  dplyr::contains("codebook_"),
+  store = store
+)
+codebook_1
+```
+
+```{r codebook-2}
+#| label: tbl-codebook-2
+#| tbl-cap: Codebook for covariates used in RQ2
+
+codebook_2
+```
+
+```{r codebook-3}
+#| label: tbl-codebook-3
+#| tbl-cap: Codebook for covariates used in RQ3
+
+codebook_3
+```
+
+## Other results
+
+### Balancing weights for effect modification
+
+```{r}
+targets::tar_load(
+  dplyr::matches("weights_(\\d)SA$"),
+  store = store
+)
+```
+
+```{r weights-1sa}
+#| label: tbl-weights-1sa
+#| tbl-cap: Weights
+
+tidy_res_weighted_fits(
+  dat_tbl = weights_1SA$dat_tbl,
+  dat_plt = weights_1SA$dat_plt,
+  sa_var = "e3_sex"
+)$table
+```
+
+```{r weights-2sa}
+#| label: tbl-weights-2sa
+#| tbl-cap: Weights
+
+tidy_res_weighted_fits(
+  dat_tbl = weights_2SA$dat_tbl,
+  dat_plt = weights_2SA$dat_plt,
+  sa_var = "e3_sex"
+)$table
+```
+
+```{r weights-3sa}
+#| label: tbl-weights-3sa
+#| tbl-cap: Weights
+
+tidy_res_weighted_fits(
+  dat_tbl = weights_3SA$dat_tbl,
+  dat_plt = weights_3SA$dat_plt,
+  sa_var = "e3_sex"
+)$table
+```
+
 # Supplementary figures
@@
 
+## Other results
+
+### Marginal comparisons for effect modification
+
+```{r}
+targets::tar_load(
+  dplyr::matches("marginal_comparisons_(\\d)SA$"),
+  store = store
+)
+```
+
+```{r marginal-1sa}
+#| label: fig-marginal-1sa
+#| fig-cap: Average comparisons for effect modification
+#| fig-width: 6
+#| fig-height: 7
+
+tidy_res_meffects(
+  df = marginal_comparisons_1SA,
+  sa_var = "e3_sex",
+  which_res = "comparisons",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)$plot
+```
+
+```{r marginal-2sa}
+#| label: fig-marginal-2sa
+#| fig-cap: Average comparisons for effect modification
+#| fig-width: 10
+#| fig-height: 10
+
+tidy_res_meffects(
+  df = marginal_comparisons_2SA,
+  sa_var = "e3_sex",
+  which_res = "comparisons",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)$plot
+```
+
+```{r marginal-3sa}
+#| label: fig-marginal-3sa
+#| fig-cap: Average comparisons for effect modification
+#| fig-width: 5
+#| fig-height: 5
+
+tidy_res_meffects(
+  df = marginal_comparisons_3SA,
+  sa_var = "e3_sex",
+  which_res = "comparisons",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)$plot
+```
+
 # References {.unnumbered}

### 2023-11-07 output/paper/tables.qmd
Modified pipeline so that targets for processing results do not actually produce eg plots but only ready to use tibbles for plots and tables. Tested pipeline to create paper but I think I still have to manually reference tables and figures in manuscript.

@@
 
+```{r}
+source("code/proc_res.R")
+```
+
+```{r vars}
+store <- ifelse(
+  Sys.getenv("is_hpc"), 
+  paste0(
+    "/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/",
+    "lorenzoF/data/data_paper3/_targets/_targets_res"
+  ), 
+  "~/mounts/rstudioserver/data/data_paper3/_targets/_targets_res"
+)
+
+num_digits_est = 3
+num_digits_sig = 2
+```
+
 - Begin each table on a new page after the list of references, numbered with Arabic numerals.
@@
 
+```{r pop-desc}
+#| label: tbl-pop-desc
+#| tbl-cap: Population description
+
+targets::tar_load("table1", store = store)
+table1
+```
+
 ## Endocrine disruptors
@@
 
-```{r}
-#| label: tbl-chem-info
-#| tbl-cap: "Information about the non-persistent EDCs."
+```{r edc-desc}
+#| label: tbl-edc-desc
+#| tbl-cap: Levels of unprocessed chemicals
 
-edcs_info <- myphd::edcs_information() |>
-  tidylog::select(-c(chem_id, smiles))
-gt::gt(edcs_info,
-       groupname_col = "class") |>
-  gt::cols_label(
-    full_name = "Compound",
-    short_name = "Symbol",
-    pubchem_cid = "PubChem CID",
-    ctd_id = "CTD ID",
-    exposome_explorer_id = "Exposome Explore ID",
-    parental_compound = "Parental compound"
-  ) |>
-  gt::tab_style(
-    locations = gt::cells_column_labels(columns = everything()),
-    style = list(
-      gt::cell_borders(sides = "bottom", weight = gt::px(3)),
-      gt::cell_text(weight = "bold")
-    )
-  ) |>
-  gt::tab_style(
-    locations = gt::cells_row_groups(group = dplyr::everything()),
-    style = list(gt::cell_text(weight = "bold"))
-  )
+targets::tar_load("desc_chems_mets", store = store)
+desc_chems_mets$desc_chems
 ```
@@
 
+```{r met-desc}
+#| label: tbl-met-desc
+#| tbl-cap: Levels of unprocessed corticosteroids
+
+desc_chems_mets$desc_mets
+```
+
 # Main results
 
-- RQ2: group chemicals by class, and for each chemical report results with metabolites (`avg_comparisons` with estimate, SE, s-value, CI).
-- RQ3: report results with outcome (`avg_comparisons` with estimate, SE, s-value, CI).
+## Balancing weights
+
+```{r}
+targets::tar_load(
+  dplyr::matches("weights_(\\d)$"),
+  store = store
+)
+```
+
+```{r weights-1}
+#| label: tbl-weights-1
+#| tbl-cap: Weights
+
+tidy_res_weighted_fits(
+  dat_tbl = weights_1$dat_tbl,
+  dat_plt = weights_1$dat_plt,
+  sa_var = NULL
+)$table
+```
+
+```{r weights-2}
+#| label: tbl-weights-2
+#| tbl-cap: Weights
+
+tidy_res_weighted_fits(
+  dat_tbl = weights_2$dat_tbl,
+  dat_plt = weights_2$dat_plt,
+  sa_var = NULL
+)$table
+```
+
+```{r weights-3}
+#| label: tbl-weights-3
+#| tbl-cap: Weights
+
+tidy_res_weighted_fits(
+  dat_tbl = weights_3$dat_tbl,
+  dat_plt = weights_3$dat_plt,
+  sa_var = NULL
+)$table
+```
+
+## Marginal comparisons
+
+```{r}
+targets::tar_load(
+  dplyr::matches("marginal_comparisons_(\\d)$"),
+  store = store
+)
+```
+
+```{r marginal-1}
+#| label: tbl-marginal-1
+#| tbl-cap: Average comparisons
+
+tidy_res_meffects(
+  df = marginal_comparisons_1,
+  sa_var = NULL,
+  which_res = "comparisons",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)$table
+```
+
+```{r marginal-2}
+#| label: tbl-marginal-2
+#| tbl-cap: Average comparisons
+
+tidy_res_meffects(
+  df = marginal_comparisons_2,
+  sa_var = NULL,
+  which_res = "comparisons",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)$table
+```
+
+```{r marginal-3}
+#| label: tbl-marginal-3
+#| tbl-cap: Average comparisons
+
+tidy_res_meffects(
+  df = marginal_comparisons_3,
+  sa_var = NULL,
+  which_res = "comparisons",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)$table
+```
 
 # Other analyses
+
+## Marginal hypothesis
+
+```{r}
+targets::tar_load(
+  dplyr::matches("marginal_hypothesis_(\\d)SA$"),
+  store = store
+)
+```
+
+```{r hypothesis-1}
+#| label: tbl-hypothesis-1
+#| tbl-cap: Hypothesis testing for effect modification
+
+tidy_res_meffects(
+  df = marginal_hypothesis_1SA,
+  sa_var = "e3sex",
+  which_res = "hypothesis",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)$table
+```
+
+```{r hypothesis-2}
+#| label: tbl-hypothesis-2
+#| tbl-cap: Hypothesis testing for effect modification
+
+tidy_res_meffects(
+  df = marginal_hypothesis_2SA,
+  sa_var = "e3sex",
+  which_res = "hypothesis",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)$table
+```
+
+```{r hypothesis-3}
+#| label: tbl-hypothesis-3
+#| tbl-cap: Hypothesis testing for effect modification
+
+tidy_res_meffects(
+  df = marginal_hypothesis_3SA,
+  sa_var = "e3sex",
+  which_res = "hypothesis",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)$table
+```

### 2023-11-07 _targets_res.R

@@ -9,7 +9,7 @@ targets::tar_option_set(
 path_store <- Sys.getenv("path_store")
 num_digits_est = 3
 num_digits_sig = 2
-sa_var <- "e3_sex"
+sa_var_ <- "e3_sex"
 
 list(
   targets::tar_target(
@@ -26,62 +26,332 @@ list(
   ), # End desc_chems_mets target
   ##############################################################################
   tarchetypes::tar_map(
-    list(rq = c("1", "2", "3", "1SA", "2SA", "3SA")),
+    list(rq_ = c("1", "2", "3", "1SA", "2SA", "3SA")),
     targets::tar_target(
       name = "weights",
       command = {
         targets::tar_load(
-          dplyr::contains(paste0("rq", substr(rq, 1, 1), "_weighted_fits_")),
-          store = paste0(path_store, rq)
+          dplyr::contains(paste0("rq", substr(rq_, 1, 1), "_weighted_fits_")),
+          store = paste0(path_store, rq_)
         )
-        tidy_res_weighted_fits(
-          res_list = mget(ls(
-            pattern = paste0("rq", substr(rq, 1, 1), "_weighted_fits_*")
-          )),
-          rq = substr(rq, 1, 1),
-          sa_var = if (grepl("SA$", rq)) sa_var else NULL
+        res_list <- mget(ls(
+          pattern = paste0("rq", substr(rq_, 1, 1), "_weighted_fits_*")
+        ))
+        rq <- substr(rq_, 1, 1)
+        sa_var <- if (grepl("SA$", rq_)) sa_var_ else NULL
+        names_ <- gsub(paste0("rq", rq, "_weighted_fits_"),
+                       "",
+                       names(res_list))
+        weights_ <- lapply(1:1, function(idx) {
+          outcome <- gsub(paste0("rq", rq, "_weighted_fits_"),
+                          "",
+                          names_[idx])
+          tmp <- lapply(res_list[[idx]]$fits, "[[", "weights")
+          tmp <- lapply(seq_along(tmp), function(idx2) {
+            ret <- tibble::tibble(tmp[[idx2]])
+            colnames(ret) <- names(tmp)[idx2]
+            return(ret)
+          }) |>
+            dplyr::bind_cols()
+          # Check whether is SA and eventually add effect modifier
+          if (!is.null(sa_var)) {
+            tmp <- tmp |>
+              tidylog::mutate(
+                modifier = res_list[[1]]$fits[[1]]$dat[[sa_var]]
[…]

### 2023-11-07 code/utils.R

@@ -323,7 +323,7 @@ rq_prepare_data <- function(dat) {
   steps_outcome <- params_dat$steps[[rq]]$preproc_outcome
 
   # Create folders to store results
-  invisible(lapply(c("figures", "tables"), function(x) {
+  invisible(lapply(c("figures"), function(x) {
     path_save_res <- paste0(
       "results/", x, "/", rq
     )

### 2023-11-08 output/paper/figures.qmd
Checked compliance to guidenlines for figures and tables.

@@
 
-- Number figures according to the order in which they are first mentioned in the main text.
-- Clearly label all axes, giving both the measure and the unit of measurement.
-- Smallest type size: 8pt.
-- Maximum width: 7 inches in portrait, 9 inches in landscape mode.
-- Multi-panel figures: capital letters followed by closing parenthesis, completely above and to the left of the panel.
 - Use log-scale when plotting relative measures of effects.
 - Figure titles and legends should be submitted as single paragraph.
-- The text of figure legends typed after the reference list.
-- Save and submit each main text figure as an individual file (PDF with fonts embedded).
-
-# Participants
-
-- Flow diagram: number of individuals at each stage of study.
-Report numbers potentially eligible, examined for eligibility, confirmed eligible, included in the study, completing follow-up, and analysed.
 
 # Main results
-
-- Forest plot for RQ2 (group by chemical class, and for each chemical group bars for metabolites) and RQ3.
 
@@
 #| fig-cap: Average comparisons
-#| fig-width: 10
+#| fig-width: 9
 #| fig-height: 10

### 2023-11-08 output/paper/manuscript.qmd
Modified pipeline so that targets for processing results do not actually produce eg plots but only ready to use tibbles for plots and tables. Tested pipeline to create paper but I think I still have to manually reference tables and figures in manuscript.
Checked compliance to guidenlines for figures and tables.

@@
 
-- The title should be less or equal than 300 characters.
-It should indicate the study design, the subject of the paper, information regarding exposures and outcomes assessed, and whether the study was observational or experimental.
-- The suggested length of the abstract is less or equal than 300 words.
-- The suggested length is <7,000 words, excluding abstract, references, tables, figure captions, acknowledgments, and Supplementary Material.
-  - Concise sub-headings should be less or equal than 8 words, and they should be used to organize information rather than summarize the results.
-  - In-text citations with superscript numbers: outside periods and commas, but inside colons and semicolons.
+- Authorship must follow the ICMJE's criteria.
+Author names should be listed in ScholarOne and author contributions should be detailed in the cover letter ("Author A...").
+Neither names nor contributions should appear in the blinded manuscript.
+Do not include conflicts of interest.
+- The use of the word "effect(s)" as a proxy for "association(s)" is discouraged.
+- The length of the abstract must be less or equal than 200 words, and should be unstructured, stating the research questions, the methods used, and the results and conclusions of the research.
+- The length must be <4,000 words, excluding abstract, references, tables, figure captions, acknowledgments, and Supplementary Material.
+  - Use the AMA format.
+  - In-text citations: full-sized Arabic numerals in parentheses within the sentence.
+- Submission: SER members receive a 10% discount on the fees per page.
+This should be requested in the cover letter.
+Also 20% discount for Open Access charges.
+If uploading a single file (document, tables, figures, SM), designate the file as "Main document - anonymous".
 
 # Introduction {#sec-intro}
+
+**Do not use "Introduction" as a heading**.
 
@@
 
-- Give the source of funding and the role of the funders for the present study and, if applicable, for the original study on which the present article is based.
-
 \newpage
@@
 :::
-
-{{< embed tables.qmd >}}

### 2023-11-08 output/paper/sections/methods.qmd
Tested pipeline w/ vector of id_val.
Checked compliance to guidenlines for figures and tables.

@@
 ### Confounders {#sec-confounders}
+
+- For RQ1 I used creatinine values from HELIX. For RQ3 the ones from the steroids dataset. For RQ2, I included in the model both variables.
 
@@
 
+- In RQ2, I included the logarithm of the denominator in the RHS. In RQ3, I used the logarithm of the ratio in the RHS.
+
 - Methods for selecting potential confounders (provide DAGs).
+- Whenever possible, mathematical equations should be written on a single line. For multiplication, use a times sign.
 - Describe all statistical methods with assumptions, including those used to control for confounding.

### 2023-11-08 output/paper/si.qmd
Checked compliance to guidenlines for figures and tables.

@@
 ```{r}
+source("code/dictionary.R")
+source("code/data.R")
 source("code/proc_res.R")
@@
 
+id_var <- "HelixID"
 num_digits_est = 3
@@
 
-- Should be limited to results of secondary analyses and background details needed to ensure transparency:
-  - Questionnaires.
-  - Tables and figures with results of sensitivity analyses.
-  - Tables with numeric data corresponding to results shown in the main text figures.
-  - DAGs used to select model covariates.
-- Text description of study methods should NOT go in the SM.
-- Number supplemental tables and figures separately according to the order in which they are first mentioned in the main text, using an "S" prefix.
+- Files for supplementary data must be accompanied by a summary of the file names and types.
 
@@
 
-- Provide the title for each table above the table and the notes below the table.
-
 ## Descriptive data
@@
 ### Study populations
-
-- Detailed characteristics of study participants (demographic, clinical, social), by cohort and overall.
 
@@
 codebook_3
+```
+
+### Description of chemicals
+
+```{r info-chems}
+#| label: tbl-info-chems
+#| tbl-cap: Information about non-persistent EDCs
+
+edcs_info <- myphd::edcs_information() |>
+  tidylog::select(-c(chem_id, smiles, ctd_id, exposome_explorer_id)) |>
+  tidylog::mutate(
+    var = janitor::make_clean_names(
+      edcs_info$short_name,
+      replace = c(`-` = ""),
+      case = "none"
+    ) |>
+      tolower()
+  ) |>
+  tidylog::relocate(var, .after = short_name)
+gt::gt(edcs_info,
+       groupname_col = "class") |>
+  gt::sub_values(
+    values = c("NA"),
+    replacement = ""
+  ) |>
+  gt::cols_label(
+    full_name = "Compound",
+    short_name = "Symbol",
+    var = "Variable name",
+    pubchem_cid = "PubChem CID",
+    parental_compound = "Parental compound"
+  ) |>
+  gt::tab_style(
+    locations = gt::cells_row_groups(group = dplyr::everything()),
+    style = list(gt::cell_text(weight = "bold"))
+  )
+```
+
+### Description of metabolites
+
+```{r info-mets}
+#| label: tbl-info-mets
+#| tbl-cap: Information about metabolites
+
+mets_info <- readODS::read_ods(
+  path = "docs/steroids.ods"
+) |>
+  tidylog::select(analyte, acronym, type, hmdb_id, cas_number)
+gt::gt(mets_info,
+       groupname_col = "type") |>
+  gt::sub_values(
+    values = c("NA"),
+    replacement = ""
+  ) |>
+  gt::cols_label(
+    analyte = "Metabolite",
+    acronym = "Symbol",
+    hmdb_id = "HMDB ID",
+    cas_number = "CAS number"
+  ) |>
+  gt::tab_style(
+    locations = gt::cells_row_groups(group = dplyr::everything()),
+    style = list(gt::cell_text(weight = "bold"))
+  ) |>
+  gt::tab_footnote(
+    footnote = "Human Metabolome Database",
+    locations = gt::cells_column_labels(
+      columns = hmdb_id
+    )
+  ) |>
+  gt::tab_footnote(
+    footnote = "Chemical Abstracts Service",
+    locations = gt::cells_column_labels(
+      columns = cas_number
+    )
+  )
 ```
@@
 
-- Provide the title for each figure below the figure on the same manuscript page as the figure itself.
+## Descriptive data
+
+### Description of chemicals
+
+```{r desc-chems-lod}
+#| label: fig-desc-chems
+#| fig-cap: Description of chemicals
+#| fig-width: 6
+#| fig-height: 7
+
+viz_desc_vars(
+  dat = load_dat_request()$dat,
+  vars = vars_of_interest(append_to_chem = "cdesc")$chemicals,
+  fct_levels = c(1, 2, 3, 4),
+  is_chem = TRUE
+)
+```
+
+### Description of metabolites
+
+```{r desc-mets-lod}
+#| label: fig-desc-mets
+#| fig-cap: Description of metabolites
+#| fig-width: 6
+#| fig-height: 7
+
+viz_desc_vars(
+  dat = myphd::extract_cohort(load_steroids()$desc,
+                              id_var = id_var) |>
+    tidylog::mutate(
+      cohort = dplyr::case_when(
+        cohort == "EDE" ~ "EDEN",
+        cohort == "KAN" ~ "KANC",
+        cohort == "MOB" ~ "MOBA",
+        cohort == "RHE" ~ "RHEA",
+        .default = cohort
+      )
+    ),
+  vars = vars_of_interest()$metabolites |>
+    paste0("_cdesc"),
+  fct_levels = c(1, 2, 3, 4),
+  is_chem = FALSE
+)
+```
 
 ## Main results
-
-- ADRF and AMEF for RQ2 and RQ3.
-- Whole distribution (i.e., density) of e.g., risk differences (instead of marginalized estimates), with lines for mean and median.
 
@@
 #| fig-cap: Average comparisons for effect modification
-#| fig-width: 10
+#| fig-width: 9
 #| fig-height: 10

### 2023-11-08 output/paper/tables.qmd
Checked compliance to guidenlines for figures and tables.

@@
 
-- Begin each table on a new page after the list of references, numbered with Arabic numerals.
-- Leave blank spaces for no entry.
-- Tables may not contain parts.
-- Provide the title for each table above the table and the notes below the table.
-- Order footnotes: abbreviations listed alphabetically and separated by semicolons; others preceded by superscript lowercase letter.
-
 # Descriptive data
@@
 ## Study populations
-
-- Basic characteristics of study participants (demographic, clinical, social), by cohort and overall.
-  - Include all covariates included in primary or secondary analyses.
-  - Indicate number of participants with missing data for each variable of interest.
-  - Report numbers of outcome events or summary measures.
 
@@
 
-- Levels of unprocessed biomarkers (chemicals), with IDs.
-- Report: minimum and maximum, percentiles, and number of observations above/below the assay's LOD and LOQ.
-
 ```{r edc-desc}
@@
 ## Corticosteroids
-
-- Levels of unprocessed biomarkers (metabolites), with IDs.
-- Table comparing subjects with and without measurement of steroids.
 

### 2023-11-09 output/paper/manuscript.tex
Modified pipeline so that targets for processing results do not actually produce eg plots but only ready to use tibbles for plots and tables. Tested pipeline to create paper but I think I still have to manually reference tables and figures in manuscript.
Kept working on processing results for paper.

--- 2023-10-14 output/paper/manuscript.tex
@@
 \item
-  The title should be less or equal than 300 characters. It should
-  indicate the study design, the subject of the paper, information
-  regarding exposures and outcomes assessed, and whether the study was
-  observational or experimental.
+  Authorship must follow the ICMJE's criteria. Author names should be
+  listed in ScholarOne and author contributions should be detailed in
+  the cover letter (``Author A\ldots{}''). Neither names nor
+  contributions should appear in the blinded manuscript. Do not include
+  conflicts of interest.
 \item
-  The suggested length of the abstract is less or equal than 300 words.
+  The use of the word ``effect(s)'' as a proxy for ``association(s)'' is
+  discouraged.
 \item
-  The suggested length is \textless7,000 words, excluding abstract,
+  The length of the abstract must be less or equal than 200 words, and
+  should be unstructured, stating the research questions, the methods
+  used, and the results and conclusions of the research.
+\item
+  The length must be \textless4,000 words, excluding abstract,
   references, tables, figure captions, acknowledgments, and
@@
   \item
-    Concise sub-headings should be less or equal than 8 words, and they
-    should be used to organize information rather than summarize the
-    results.
+    Use the AMA format.
   \item
-    In-text citations with superscript numbers: outside periods and
-    commas, but inside colons and semicolons.
+    In-text citations: full-sized Arabic numerals in parentheses within
+    the sentence.
   \end{itemize}
+\item
+  Submission: SER members receive a 10\% discount on the fees per page.
+  This should be requested in the cover letter. Also 20\% discount for
+  Open Access charges. If uploading a single file (document, tables,
+  figures, SM), designate the file as ``Main document - anonymous''.
 \end{itemize}
@@
 \section{Introduction}\label{sec-intro}
+
+\textbf{Do not use ``Introduction'' as a heading}.
 
@@
 \begin{itemize}
-\tightlist
 \item
@@
   effect modifier, or other type of variable in analyses.
+\item
+  Units (in \texttt{log}):
+
+  \begin{itemize}
+  \tightlist
+  \item
+    \texttt{hitrtse} in \texttt{ms}.
+  \item
+    EDCs in \texttt{microg/L}.
+  \item
+    Cortisol production and cortisone production in \texttt{ng/mL}.
+  \item
+    Cortisol metabolism and 11bHSD in \texttt{NA}.
+  \end{itemize}
 \end{itemize}
@@
 \subsubsection{Confounders}\label{sec-confounders}
+
+\begin{itemize}
+\tightlist
+\item
+  For RQ1 I used creatinine values from HELIX. For RQ3 the ones from the
+  steroids dataset. For RQ2, I included in the model both variables.
+\end{itemize}
 
@@
 \begin{itemize}
-\tightlist
+\item
+  In RQ2, I included the logarithm of the denominator in the RHS. In
+  RQ3, I used the logarithm of the ratio in the RHS.
 \item
   Methods for selecting potential confounders (provide DAGs).
+\item
+  Whenever possible, mathematical equations should be written on a
+  single line. For multiplication, use a times sign.
 \item
@@
     \emph{p}-values are used, report numeric \emph{p}-values, rounded to
-    1-2 digits, for all results.
+    1-2 digits, for all results. Use an uppercase italic letter ``P'',
+    and the values should not be bolded. Indicate whether are 1- or
+    2-sided.
   \end{itemize}
@@
 
-\textbf{?@tbl-chem-info}
-
 \section{Discussion}\label{sec-discussion}
@@
   \end{itemize}
+\end{itemize}
+
+\subsubsection{What does the literature
+say?}\label{what-does-the-literature-say}
+
+\begin{itemize}
+\tightlist
+\item
+  EDCs and neurodevelopment (ANT).
+\item
+  EDCs and corticosteroids.
+\item
+  Corticosteroids and neurodevelopment (ANT).
 \end{itemize}
@@
 
-\begin{itemize}
-\tightlist
-\item
-  Give the source of funding and the role of the funders for the present
-  study and, if applicable, for the original study on which the present
-  article is based.
-\end{itemize}
-
 \newpage

### 2023-11-09 output/paper/si.qmd
Kept working on processing results for paper.

@@
 ```{r}
-source("code/dictionary.R")
+source("code/dictionaries.R")
 source("code/data.R")
@@
     var = janitor::make_clean_names(
-      edcs_info$short_name,
+      short_name,
       replace = c(`-` = ""),
@@
   is_chem = TRUE
-)
+)$plot
 ```
@@
   is_chem = FALSE
+)$plot
+```
+
+## Main results
+
+### Average dose-response functions for significant results
+
+```{r}
+targets::tar_load(
+  dplyr::matches("marginal_gcomp_(\\d)$"),
+  store = store
 )
@@
 
-## Main results
+```{r adrfs-rq3}
+#| label: fig-adrfs-rq3
+#| fig-cap: Average dose-response functions for RQ3
+#| fig-subcap:
+#|   - "Cortisol production"
+#|   - "11bHSD"
+#| fig-width: 6
+#| fig-height: 7
+#| layout-ncol: 2
+
+exposures_of_interest <- c(
+  "cortisol production",
+  "X11bHSD"
+)
+ret_adrfs <- lapply(
+  exposures_of_interest, function(x) {
+    plot_adrf(
+      df_preds = marginal_gcomp_3 |>
+        tidylog::filter(
+          variable == x
+        )
+    )
+  }
+)
+ret_adrfs[[1]]
+ret_adrfs[[2]]
+```
 

### 2023-11-09 _targets_res.R

@@ -28,6 +28,17 @@ list(
   tarchetypes::tar_map(
     list(rq_ = c("1", "2", "3", "1SA", "2SA", "3SA")),
     targets::tar_target(
+      name = "balance",
+      command = {
+        if (!grepl("SA$", rq_)) {
+          load_tidy_res_weighting(
+            path_store = path_store,
+            rq = rq_
+          )
+        }
+      }#########################################################################
+    ), # End balance target
+    targets::tar_target(
       name = "weights",
       command = load_res_weighted_fits(
         path_store = path_store,
@@ -43,6 +54,19 @@ list(
         sa_var = if (grepl("SA$", rq_)) sa_var_ else NULL,
         which_res = "comparisons"
       )
+    ), # End marginal_comparisons target
+    targets::tar_target(
+      name = "marginal_gcomp",
+      command = {
+        if (!grepl("SA$", rq_)) {
+          load_res_meffects(
+            path_store = path_store,
+            rq = rq_,
+            sa_var = if (grepl("SA$", rq_)) sa_var_ else NULL,
+            which_res = "gcomp"
+          )
+        }
+      }#########################################################################
     ), # End marginal_comparisons target
     targets::tar_target(
       name = "marginal_hypothesis",

### 2023-11-09 code/utils.R

@@ -791,6 +791,13 @@ rq_estimate_marginal_effects <-
     params_dat <- params(is_hpc = Sys.getenv("is_hpc"))
     params_ana <- params_analyses()[[rq]]
     
+    old_by <- by
+    if (is.null(by)) {
+      by <- TRUE
+    } else {
+      by <- glue::double_quote(by)
+    }
+    
     # Loop over the fitted models to estimate marginal effects
     if (parallel == TRUE) {
       future::plan(future::multisession,
@@ -822,35 +829,13 @@ rq_estimate_marginal_effects <-
               {exposure} = values
             ),
             wts = weights,
-            vcov = {glue::double_quote(vcov)}
+            vcov = {vcov}
           )",
           exposure = exposure,
-          vcov = "HC3"
+          vcov = "~ cohort"
         )
-      )) # End G-computation (avg_predictions)
-      
-      # Plot ADRF
-      # adrf <- gcomp |>
-      #   ggplot2::ggplot(ggplot2::aes(x = .data[[exposure]])) +
-      #   ggplot2::geom_point(ggplot2::aes(y = estimate)) +
-      #   ggplot2::geom_line(ggplot2::aes(y = estimate),
-      #     linewidth = 0.2
-      #   ) +
-      #   ggplot2::geom_ribbon(ggplot2::aes(
-      #     ymin = conf.low,
-      #     ymax = conf.high
-      #   ), alpha = 0.2) +
-      #   ggplot2::geom_rug(
-      #     mapping = ggplot2::aes(x = .data[[exposure]]),
-      #     data = dat,
-      #     inherit.aes = FALSE
-      #   ) +
-      #   ggplot2::scale_x_continuous(limits = c(
-      #     values[1],
-      #     values[length(values)]
-      #   )) +
-      #   ggplot2::labs(x = exposure, y = "E[Y|A]") +
-      #   ggplot2::theme_minimal()
+      )) |> # End G-computation (avg_predictions)
+        marginaleffects::tidy()
       ########################################################################
       
       ########################################################################
@@ -872,33 +857,6 @@ rq_estimate_marginal_effects <-
       #     vcov = "HC3"
       #   )
       # )) # End slopes (avg_slopes)
-      
-      # Plot AMEF
-      # amef <- slopes |>
-      #   ggplot2::ggplot(ggplot2::aes(x = .data[[exposure]])) +
-      #   ggplot2::geom_point(ggplot2::aes(y = estimate)) +
[…]

### 2023-11-10 output/paper/tables.qmd
Kept working on processing results for paper.
Added target to produce figures for overlap exposures. Fixed tables for obs numbers afrer weighting and trimming.

@@
 ## Balancing weights
+
+```{r}
+targets::tar_load(
+  dplyr::matches("balance_(\\d)$"),
+  store = store
+)
+```
+
+```{r balance-1}
+#| label: tbl-balance-1
+#| tbl-cap: Effective number of observations
+
+balance_1 |>
+  tidylog::group_by(class) |>
+  dplyr::arrange(dplyr::desc(Adjusted)) |>
+  tidylog::ungroup() |>
+  gt::gt(groupname_col = "class") |>
+  gt::fmt_number(n_sigfig = 4) |>
+  gt::tab_footnote(
+    footnote = "N.",
+    locations = gt::cells_column_labels(
+      columns = c(Unadjusted, Adjusted)
+    )
+  ) |>
+  gt::tab_footnote(
+    footnote = "Truncated weights.",
+    locations = gt::cells_column_labels(columns = Adjusted)
+  )
+```
+
+```{r balance-2}
+#| label: tbl-balance-2
+#| tbl-cap: Effective number of observations
+
+balance_2 |>
+  tidylog::group_by(class) |>
+  dplyr::arrange(dplyr::desc(Adjusted)) |>
+  tidylog::ungroup() |>
+  gt::gt(groupname_col = "class") |>
+  gt::fmt_number(n_sigfig = 4) |>
+  gt::tab_footnote(
+    footnote = "N.",
+    locations = gt::cells_column_labels(
+      columns = c(Unadjusted, Adjusted)
+    )
+  ) |>
+  gt::tab_footnote(
+    footnote = "Truncated weights.",
+    locations = gt::cells_column_labels(columns = Adjusted)
+  )
+```
+
+```{r balance-3}
+#| label: tbl-balance-3
+#| tbl-cap: Effective number of observations
+
+balance_3 |>
+  dplyr::arrange(dplyr::desc(Adjusted)) |>
+  gt::gt() |>
+  gt::fmt_number(n_sigfig = 4) |>
+  gt::tab_footnote(
+    footnote = "N.",
+    locations = gt::cells_column_labels(
+      columns = c(Unadjusted, Adjusted)
+    )
+  ) |>
+  gt::tab_footnote(
+    footnote = "Truncated weights.",
+    locations = gt::cells_column_labels(columns = Adjusted)
+  )
+```
 

### 2023-11-10 code/proc_res.R

@@ -269,11 +269,28 @@ load_tidy_res_weighting <- function(path_store, rq) {
         values_from = "N"
       ) |>
       tidylog::mutate(
-        exposure = names(obj)[[idx]]
+        exposure = names(obj)[[idx]],
+        exposure = gsub("hs_|_c", "", exposure),
+        exposure = gsub("_", " ", exposure),
+        exposure = replace(
+          exposure, tolower(exposure) == "x11bhsd", "11bhsd"
+        )
       ) |>
       tidylog::relocate(exposure)
   }) |>
     dplyr::bind_rows()
+  
+  ## Eventually add information on EDCs (chemical classes)
+  if (rq %in% c("1", "2")) {
+    info_edcs <- myphd::edcs_information() |>
+      tibble::as_tibble()
+    info <- info |>
+      tidylog::left_join(
+        info_edcs |>
+          tidylog::select(chem_id, class),
+        by = c("exposure" = "chem_id")
+      )
+  }
   
   return(info)
 } # End function load_tidy_res_weighting
@@ -843,10 +860,20 @@ viz_overlap_quantiles <- function(dat,
                                   id_var,
                                   group_var,
                                   lower_percentile,
-                                  upper_percentile) {
+                                  upper_percentile,
+                                  rq) {
   # Process data
   dat_proc <- dat |>
     tidylog::select(-dplyr::any_of(id_var))
+  colnames(dat_proc) <- colnames(dat_proc) |>
+    stringr::str_replace_all(
+      pattern = "hs_|_c",
+      replacement = ""
+    ) |>
+    stringr::str_replace_all(
+      pattern = "_",
+      replacement = " "
+    )
   vars <- setdiff(colnames(dat_proc), group_var)
   
   # Plot
@@ -878,16 +905,31 @@ viz_overlap_quantiles <- function(dat,
         name = "Tail probability",
         direction = -1
       ) +
+      ggplot2::xlim(
+        min(dat_proc[[x]]),
+        quantile(dat_proc[[x]], upper_percentile)
+      ) +
       ggplot2::labs(
         caption = glue::glue(
-          "Lower percentile: {lower}th. Upper percentile: {upper}th.",
+          "Red bars: {lower}th and {upper}th percentiles.
+          Figure truncated at the {upper}th percentile of the exposure.",
           lower = lower_percentile * 100,
           upper = upper_percentile * 100
         )
       )
   }) # End loop over variables to plot
   
[…]

### 2023-11-10 code/utils.R

@@ -12,18 +12,18 @@ load_dag <- function(dags, exposure, outcome, params_dag) {
   ret <- list()
   rq <- Sys.getenv("TAR_PROJECT")
   which_dag <- switch(rq,
-    "rq01" = "chem_to_out",
-    "rq1" = "chem_to_out",
-    "rq02" = "chem_to_marker",
-    "rq2" = "chem_to_marker",
-    "rq03" = "marker_to_out",
-    "rq3" = "marker_to_out",
-    "rq04" = "marker_to_chem",
-    "rq4" = "marker_to_chem"
+                      "rq01" = "chem_to_out",
+                      "rq1" = "chem_to_out",
+                      "rq02" = "chem_to_marker",
+                      "rq2" = "chem_to_marker",
+                      "rq03" = "marker_to_out",
+                      "rq3" = "marker_to_out",
+                      "rq04" = "marker_to_chem",
+                      "rq4" = "marker_to_chem"
   )
-
+  
   dag <- dags[[which_dag]]
-
+  
   # Find the adjustment set(s) of interest
   adjustment_set <- dagitty::adjustmentSets(
     x = dag,
@@ -32,7 +32,7 @@ load_dag <- function(dags, exposure, outcome, params_dag) {
     type = params_dag$type,
     effect = params_dag$effect
   )
-
+  
   # Identify precision covariates to add to adjustment set
   prec_covars <- dagitty::parents(
     x = dag,
@@ -49,7 +49,7 @@ load_dag <- function(dags, exposure, outcome, params_dag) {
     ),
     parents_exposure
   )
-
+  
   return(list(
     dag = dag,
     adjustment_sets = adjustment_set,
@@ -76,20 +76,20 @@ select_adjustment_set <- function(dat, meta, res_dag, strategy) {
   lengths_as <- lapply(all_as, length) |>
     unname() |>
     unlist()
-
+  
   ret <- switch(strategy,
-    "first" = all_as[[1]],
-    "smallest" = all_as[[which.min(lengths_as)]],
-    "largest" = all_as[[which.max(lengths_as)]],
-    "random" = sample(all_as, size = 1),
-    "minimize_missings" = all_as[[myphd::minimize_missings(
-      dat = dat,
-      meta = meta,
-      adjustment_sets = all_as,
-      by_var = "cohort"
-    )]]
+                "first" = all_as[[1]],
+                "smallest" = all_as[[which.min(lengths_as)]],
[…]

### 2023-11-13 output/paper/manuscript.qmd
Started writing methods for paper.

@@
 filters:
+  - acronyms
   - authors-block
-  - ../../../../all_phd/_extensions/acronyms/parse-acronyms.lua
   - ../../../../all_phd/_extensions/wordcount/citeproc.lua

### 2023-11-13 output/paper/manuscript.tex
Started writing methods for paper.

@@
 \makeatother
+% definitions for citeproc citations
+\NewDocumentCommand\citeproctext{}{}
+\NewDocumentCommand\citeproc{mm}{%
+  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
+\makeatletter
+ % allow citations to break across lines
+ \let\@cite@ofmt\@firstofone
+ % avoid brackets around text for \cite:
+ \def\@biblabel#1{}
+ \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
+\makeatother
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
+ {\begin{list}{}{%
+  \setlength{\itemindent}{0pt}
+  \setlength{\leftmargin}{0pt}
+  \setlength{\parsep}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+   \setlength{\leftmargin}{\cslhangindent}
+   \setlength{\itemindent}{-1\cslhangindent}
+  \fi
+  % set entry spacing
+  \setlength{\itemsep}{#2\baselineskip}}}
+ {\end{list}}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{\hfill\break#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 
-\subsection{Study design}\label{sec-design}
+\subsection{Study population and design}\label{sec-design}
 
-\begin{itemize}
-\tightlist
-\item
-  Present key elements of study design
-\end{itemize}
+The Human Early-Life Exposome (HELIX) is an ongoing project which aims
+to characterize early-life exposures and their potential association
+with endogenous biomarkers and health
+outcomes\textsuperscript{\citeproc{ref-VrijheidSlamaRobinson:2014}{1}}.
+It consists of six existing population-based birth cohort studies across
+Europe: BiB (Born in Bradford,
+UK)\textsuperscript{\citeproc{ref-WrightSmallRaynor:2013}{2}}, EDEN
+(Study of determinants of pre- and postnatal developmental,
+France)\textsuperscript{\citeproc{ref-HeudeForhanSlama:2016}{3}}, INMA
+(Environment and Childhood,
+Spain)\textsuperscript{\citeproc{ref-GuxensBallesterEspada:2012}{4}},
+KANC (Kaunas Cohort,
+Lithuania)\textsuperscript{\citeproc{ref-GrazulevicieneDanileviciuteNadisauskiene:2009}{5}},
+MoBa (The Norwegian Mother and Child Cohort Study,
+Norway)\textsuperscript{\citeproc{ref-MagnusIrgensHaug:2006}{6}}, and
+Rhea (Mother--Child Cohort in Crete,
+Greece)\textsuperscript{\citeproc{ref-ChatziPlanaDaraki:2009}{7}}, for a
+total of 32,000 mother-child pairs. A HELIX subcohort of 1,200
+mother-child pairs was fully characterized for the external and internal
+exposome, including exposure and omics biomarkers during childhood.
+Eligibility criteria for inclusion in the HELIX subcohort included: a)
+age 6-11 years, with a preference for 7-9 years; b) availability of
+sufficient stored pregnancy blood and urine samples; c) availability of
+complete address history from first to last follow-up; d) no serious
+health problems, which might affect the results of the clinical testing.
+Further information can be found
+in\textsuperscript{\citeproc{ref-MaitreBontCasas:2018}{8}}.
 
-\subsection{Setting}\label{sec-setting}
-
-\begin{itemize}
-\tightlist
-\item
-  Describe the setting, locations, and relevant dates, including periods
-  of recruitment, exposure, follow-up, and data collection.
-\end{itemize}
+Ethical permission was obtained from the relevant authorities in the
+corresponding country.
 
@@
 
+\phantomsection\label{refs}
+\begin{CSLReferences}{0}{1}
+\bibitem[\citeproctext]{ref-VrijheidSlamaRobinson:2014}
+\CSLLeftMargin{1. }%
+\CSLRightInline{Vrijheid M, Slama R, Robinson O, et al. The human
+early-life exposome ({HELIX}): Project rationale and design.
+\emph{Environ Health Perspect}. 2014;122(6):535-544.
+doi:\href{https://doi.org/10.1289/ehp.1307204}{10.1289/ehp.1307204}}
+
+\bibitem[\citeproctext]{ref-WrightSmallRaynor:2013}
+\CSLLeftMargin{2. }%
+\CSLRightInline{Wright J, Small N, Raynor P, et al. Cohort {Profile}:
+{The Born} in {Bradford} multi-ethnic family cohort study.
+\emph{International Journal of Epidemiology}. 2013;42(4):978-991.
+doi:\href{https://doi.org/10.1093/ije/dys112}{10.1093/ije/dys112}}
+
+\bibitem[\citeproctext]{ref-HeudeForhanSlama:2016}
+\CSLLeftMargin{3. }%
+\CSLRightInline{Heude B, Forhan A, Slama R, et al. Cohort {Profile}:
+{The EDEN} mother-child cohort on the prenatal and early postnatal
+determinants of child health and development. \emph{International
+Journal of Epidemiology}. 2016;45(2):353-363.
+doi:\href{https://doi.org/10.1093/ije/dyv151}{10.1093/ije/dyv151}}
+
+\bibitem[\citeproctext]{ref-GuxensBallesterEspada:2012}
+\CSLLeftMargin{4. }%
+\CSLRightInline{Guxens M, Ballester F, Espada M, et al. Cohort
+{Profile}: {The INMA}---{INfancia} y {Medio Ambiente}---({Environment}
+and {Childhood}) {Project}. \emph{International Journal of
+Epidemiology}. 2012;41(4):930-940.
+doi:\href{https://doi.org/10.1093/ije/dyr054}{10.1093/ije/dyr054}}
+
+\bibitem[\citeproctext]{ref-GrazulevicieneDanileviciuteNadisauskiene:2009}
+\CSLLeftMargin{5. }%
+\CSLRightInline{Grazuleviciene R, Danileviciute A, Nadisauskiene R,
+Vencloviene J. Maternal {Smoking},{GSTM1} and {GSTT1 Polymorphism} and
+{Susceptibility} to {Adverse Pregnancy Outcomes}. \emph{International
+Journal of Environmental Research and Public Health}. 2009;6(3,
+3):1282-1297.
+doi:\href{https://doi.org/10.3390/ijerph6031282}{10.3390/ijerph6031282}}
+
+\bibitem[\citeproctext]{ref-MagnusIrgensHaug:2006}
+\CSLLeftMargin{6. }%
+\CSLRightInline{Magnus P, Irgens LM, Haug K, et al. Cohort profile: {The
+Norwegian Mother} and {Child Cohort Study} ({MoBa}). \emph{International
+Journal of Epidemiology}. 2006;35(5):1146-1150.
+doi:\href{https://doi.org/10.1093/ije/dyl170}{10.1093/ije/dyl170}}
+
+\bibitem[\citeproctext]{ref-ChatziPlanaDaraki:2009}
+\CSLLeftMargin{7. }%
+\CSLRightInline{Chatzi L, Plana E, Daraki V, et al. Metabolic {Syndrome}
+in {Early Pregnancy} and {Risk} of {Preterm Birth}. \emph{American
+Journal of Epidemiology}. 2009;170(7):829-836.
+doi:\href{https://doi.org/10.1093/aje/kwp211}{10.1093/aje/kwp211}}
+
+\bibitem[\citeproctext]{ref-MaitreBontCasas:2018}
+\CSLLeftMargin{8. }%
+\CSLRightInline{Maitre L, Bont J de, Casas M, et al. Human {Early Life
+Exposome} ({HELIX}) study: A {European} population-based exposome
+cohort. \emph{BMJ Open}. 2018;8(9):e021311.
+doi:\href{https://doi.org/10.1136/bmjopen-2017-021311}{10.1136/bmjopen-2017-021311}}
+
+\bibitem[\citeproctext]{ref-VrijheidSlamaRobinson:2014}
+\CSLLeftMargin{1. }%
+\CSLRightInline{Vrijheid M, Slama R, Robinson O, et al. The human
+early-life exposome ({HELIX}): Project rationale and design.
+\emph{Environ Health Perspect}. 2014;122(6):535-544.
+doi:\href{https://doi.org/10.1289/ehp.1307204}{10.1289/ehp.1307204}}
+
+\bibitem[\citeproctext]{ref-WrightSmallRaynor:2013}
+\CSLLeftMargin{2. }%
+\CSLRightInline{Wright J, Small N, Raynor P, et al. Cohort {Profile}:
+{The Born} in {Bradford} multi-ethnic family cohort study.
+\emph{International Journal of Epidemiology}. 2013;42(4):978-991.
+doi:\href{https://doi.org/10.1093/ije/dys112}{10.1093/ije/dys112}}
+
+\bibitem[\citeproctext]{ref-HeudeForhanSlama:2016}
+\CSLLeftMargin{3. }%
+\CSLRightInline{Heude B, Forhan A, Slama R, et al. Cohort {Profile}:
+{The EDEN} mother-child cohort on the prenatal and early postnatal
+determinants of child health and development. \emph{International
+Journal of Epidemiology}. 2016;45(2):353-363.
+doi:\href{https://doi.org/10.1093/ije/dyv151}{10.1093/ije/dyv151}}
+
+\bibitem[\citeproctext]{ref-GuxensBallesterEspada:2012}
+\CSLLeftMargin{4. }%
+\CSLRightInline{Guxens M, Ballester F, Espada M, et al. Cohort
+{Profile}: {The INMA}---{INfancia} y {Medio Ambiente}---({Environment}
+and {Childhood}) {Project}. \emph{International Journal of
+Epidemiology}. 2012;41(4):930-940.
+doi:\href{https://doi.org/10.1093/ije/dyr054}{10.1093/ije/dyr054}}
+
+\bibitem[\citeproctext]{ref-GrazulevicieneDanileviciuteNadisauskiene:2009}
+\CSLLeftMargin{5. }%
+\CSLRightInline{Grazuleviciene R, Danileviciute A, Nadisauskiene R,
+Vencloviene J. Maternal {Smoking},{GSTM1} and {GSTT1 Polymorphism} and
+{Susceptibility} to {Adverse Pregnancy Outcomes}. \emph{International
+Journal of Environmental Research and Public Health}. 2009;6(3,
+3):1282-1297.
+doi:\href{https://doi.org/10.3390/ijerph6031282}{10.3390/ijerph6031282}}
+
+\bibitem[\citeproctext]{ref-MagnusIrgensHaug:2006}
+\CSLLeftMargin{6. }%
+\CSLRightInline{Magnus P, Irgens LM, Haug K, et al. Cohort profile: {The
+Norwegian Mother} and {Child Cohort Study} ({MoBa}). \emph{International
+Journal of Epidemiology}. 2006;35(5):1146-1150.
+doi:\href{https://doi.org/10.1093/ije/dyl170}{10.1093/ije/dyl170}}
+
+\bibitem[\citeproctext]{ref-ChatziPlanaDaraki:2009}
+\CSLLeftMargin{7. }%
+\CSLRightInline{Chatzi L, Plana E, Daraki V, et al. Metabolic {Syndrome}
+in {Early Pregnancy} and {Risk} of {Preterm Birth}. \emph{American
+Journal of Epidemiology}. 2009;170(7):829-836.
+doi:\href{https://doi.org/10.1093/aje/kwp211}{10.1093/aje/kwp211}}
+
+\bibitem[\citeproctext]{ref-MaitreBontCasas:2018}
+\CSLLeftMargin{8. }%
+\CSLRightInline{Maitre L, Bont J de, Casas M, et al. Human {Early Life
+Exposome} ({HELIX}) study: A {European} population-based exposome
+cohort. \emph{BMJ Open}. 2018;8(9):e021311.
+doi:\href{https://doi.org/10.1136/bmjopen-2017-021311}{10.1136/bmjopen-2017-021311}}
+
+\end{CSLReferences}
+
 \end{document}

### 2023-11-13 output/paper/sections/methods.qmd
Kept working on processing results for paper.
Started writing methods for paper.

@@
-## Study design {#sec-design}
+## Study population and design {#sec-design}
 
-- Present key elements of study design
+The \acr{helix} is an ongoing project which aims to characterize early-life exposures and their potential association with endogenous biomarkers and health outcomes [@VrijheidSlamaRobinson:2014].
+It consists of six existing population-based birth cohort studies across Europe: BiB (Born in Bradford, UK) [@WrightSmallRaynor:2013], EDEN (Study of determinants of pre- and postnatal developmental, France) [@HeudeForhanSlama:2016], INMA (Environment and Childhood, Spain) [@GuxensBallesterEspada:2012], KANC (Kaunas Cohort, Lithuania) [@GrazulevicieneDanileviciuteNadisauskiene:2009], MoBa (The Norwegian Mother and Child Cohort Study, Norway) [@MagnusIrgensHaug:2006], and Rhea (Mother–Child Cohort in Crete, Greece) [@ChatziPlanaDaraki:2009], for a total of 32,000 mother-child pairs.
+A \acr{helix} subcohort of 1,200 mother-child pairs was fully characterized for the external and internal exposome, including exposure and omics biomarkers during childhood.
+Eligibility criteria for inclusion in the \acr{helix} subcohort included: a) age 6-11 years, with a preference for 7-9 years; b) availability of sufficient stored pregnancy blood and urine samples; c) availability of complete address history from first to last follow-up; d) no serious health problems, which might affect the results of the clinical testing.
+Further information can be found in @MaitreBontCasas:2018.
 
-## Setting {#sec-setting}
-
-- Describe the setting, locations, and relevant dates, including periods of recruitment, exposure, follow-up, and data collection.
-
-## Participants {#sec-participants}
-
-- Cohort study: eligibility criteria, and the sources and methods of selection of participants. Describe methods of follow-up.
-- Cross-sectional study: give eligibility criteria, and the sources and methods of selection of participants.
-- Describe informed consent protocols.
-- Report how and by whom *race* or *ethnicity* was defined, and why this information was included in the study design. Disaggregate race and ethnicity data to the fullest extent possible.
+Ethical permission was obtained from the relevant authorities in the corresponding country.
 
@@
 
-- Clearly define all outcomes, exposures, predictors, potential confounders, and effect modifiers.
-- Explain the rationale for treating race as an exposure, confounder, effect modifier, or other type of variable in analyses.
+- Units (in `log`):
+  - `hitrtse` in `ms`.
+  - EDCs in `microg/L`.
+  - Cortisol production and cortisone production in `ng/mL`.
+  - Cortisol metabolism and 11bHSD in `NA`.
 
 ### Confounders {#sec-confounders}
+
+For each research question, defined by a specific type of exposure and outcome, the minimal set of covariates for inclusion in the analyses was selected on the basis of a \acr{dag} built with DAGitty [@TextorvanderZanderGilthorpe:2016] and ggdag [@Barrett:2023].
+The sets of covariates were selected to estimate the total effect of the exposure on the outcome.
+Further, each minimal adjustment set was *augmented* with precision covariates, defined as the set of parents variable of the outcome that are not parents of the exposure.
+The adjustment sets are provided in the Supplementary Material as text files compatible with DAGitty.
 
@@
 
+Children were assessed between December 2013 and February 2016, and included neurological testing and urine collection.
+Urine samples of the night before and the first morning void on the day of the visit were combined to provide a more reliable exposure assessment.
+Non-persistent \acr{edcs} assessed in urine samples from children included phthalate metabolites, phenols, and \acr{op} pesticide metabolites.
+A list of the environmental chemicals determined in urine samples and used for the present study is given in @tbl-info-chems.
+The laboratory protocols for the analysis are described elsewhere [@HaugSakhiCequier:2018].
+
 ### Corticosteroids {#sec-steroids}
+
+Urine samples of the night before the day of the visit were used to measure levels of the corticosteroids.
+These included glucocorticosteroids, glucocorticosteroid metabolites, glucocorticosteroid precursors, glucocorticosteroid precursor metabolites, androgens, and androgen metabolites.
+A list of the corticosteroids determined in urine samples and used for the present study is given in @tbl-info-mets.
+
+To assess the levels of corticosteroids and their metabolites, LC-MS/MS analysis was applied at the Applied Metabolomics Research Group, IMIM (Hospital del Mar Medical Research Institute).
+The laboratory protocols for the analysis are described elsewhere [@MarcosRenauCasals:2014; @Gomez-GomezPozo:2020].
+Of the 1,004 urine samples, 980 children were matched to the \acr{helix} subcohort.
+
+Three additional markers, cortisol production, cortisol metabolism, cortisone production, and 11bHSD activity, were computed based on the following: cortisol production as the sum of cortisol and its metabolites (20aDHF, 20bDHF, 5bDHF, 5aTHF, 5bTHF, 6OHF, 5a20acortol, 5a20bcortol, 5b20acortol, 5b20bcortol), cortisol metabolism as the inverse of the ratio between cortisol and its metabolites, cortisone production as the sum of cortisone and its metabolites (20aDHE, 20bDHE, 5aTHE, 5bTHE, 6OHE, 5b20acortolone, b20bcortolone), and 11bHSD activity as the ratio between cortisone production and cortisol production.
+11bHSD activity gives a measure of conversion of cortisone to cortisol.
 
@@
 
-## Data sources and measurement {#sec-dat-sources}
-
-- For each variable of interest, give sources of data and details of methods of assessment (measurement).
-
-## Bias {#sec-bias}
-
-- Describe any efforts to address potential sources of bias.
-
-## Study size {#sec-size}
-
-- Explain how the study size was arrived at.
-
-## Quantitative variables {#sec-quant-vars}
-
-- Explain how quantitative variables were handled in the analyses. If applicable, describe which groupings were chosen and why.
+Neurodevelopmental outcomes were assessed with standardized, non-linguistic, and culturally blind computer tests, including the \acr{ant} [@RuedaFanMcCandliss:2004].
+Further information can be found in @MaitreBontCasas:2018.
+Briefly, it is a computerized test that provides a measure of efficiency in three different functions of attention: alerting, orienting, and executive attention.
+The outcome of interest for the present study is the \acr{ahitrtse} [@SunyerEsnaolaAlvarez-Pedrerol:2015], a measure of response speed consistency throughout the test.
+A high \acr{ahitrtse} indicates highly variable reactions, and is considered a measure of inattentiveness.
 
@@
 
+### Data pre-processing {#sec-dat-preproc}
+
+### Estimation of balancing weights {#sec-weights-est}
+
+### G-computation {#sec-gcomp}
+
 - In RQ2, I included the logarithm of the denominator in the RHS. In RQ3, I used the logarithm of the ratio in the RHS.
 
-- Methods for selecting potential confounders (provide DAGs).
-- Whenever possible, mathematical equations should be written on a single line. For multiplication, use a times sign.
-- Describe all statistical methods with assumptions, including those used to control for confounding.
+- Describe all statistical methods with assumptions.
   - Description of outcome model, weighting method, estimand, and balance assessment.

### 2023-11-14 output/paper/sections/methods.qmd
Finished first draft of Methods.

@@
 ## Variables {#sec-vars}
-
-- Units (in `log`):
-  - `hitrtse` in `ms`.
-  - EDCs in `microg/L`.
-  - Cortisol production and cortisone production in `ng/mL`.
-  - Cortisol metabolism and 11bHSD in `NA`.
 
@@
 Briefly, it is a computerized test that provides a measure of efficiency in three different functions of attention: alerting, orienting, and executive attention.
-The outcome of interest for the present study is the \acr{ahitrtse} [@SunyerEsnaolaAlvarez-Pedrerol:2015], a measure of response speed consistency throughout the test.
-A high \acr{ahitrtse} indicates highly variable reactions, and is considered a measure of inattentiveness.
+The outcome of interest for the present study is the \acr{hitrtse} [@SunyerEsnaolaAlvarez-Pedrerol:2015], a measure of response speed consistency throughout the test.
+A high \acr{hitrtse} indicates highly variable reactions, and is considered a measure of inattentiveness.
 
@@
 
+```{r params-preproc}
+threshold_within_lod <- 20
+threshold_overall_lod <- 10
+threshold_within_nan <- 40
+threshold_overall_nan <- 30
+knn_k <- 5
+```
+
 ### Data pre-processing {#sec-dat-preproc}
+
+Concentrations of the corticosteroids were classified as quantifiable, below the \acr{loq}, possible interference or out of range, and not detected.
+For each metabolite, we computed the fraction of values below the \acr{loq} and not detected, both within each cohort and overall.
+We proceeded to impute these values using half the value of the corresponding \acr{lloq}, for those metabolites that had less than `r threshold_within_lod`% of missings within each cohort and `r threshold_overall_lod`% of missings overall.
+The remaining missing values were imputed using kNN from the `VIM` R package [@KowarikTempl:2016], for those metabolites that had less than `r threshold_within_nan`% of remaining missings within each cohort and `r threshold_overall_nan`% of remaining missings overall.
+We used `r knn_k` nearest neighbors.
+We natural log-transformed them to improve model fit, assessed with posterior predictive checks.
+To do so, replicated data were simulated with the fitted models and compared to the observed data.
+We used the `check_predictions` function from the `performance` R package using the default arguments [@LudeckeBen-ShacharPatil:2021].
+Values of cortisol production and cortisone production were expressed in nanograms per millilitre, whereas values of cortisol metabolism and 11bHSD activity were unitless.
+
+Concentrations of the non-persistent \acr{edcs} were classified as quantifiable, below the \acr{lod}, possible interference or out of range, and not analysed.
+Concentrations below the \acr{lod} were singly imputed using a quantile regression approach for the imputation of left-censored missing data, as implemented in the `impute.QRILC` function from the `imputeLCMD` R package [@lazar2015imputelcmd].
+Chemicals with more than 70% of observations below the \acr{lod} were not considered in the present study.
+Remaining missing values were imputed similarly using kNN.
+Values of the chemicals were expressed in $\mu$grams per litre.
+
+Missing values in the clinical outcome were imputed similarly using kNN.
+Similarly, we natural log-transformed them to improve model fit, assessed with posterior predictive checks.
+Values of the clinical outcome were expressed in \acr{mss}.
+
+Missing values in the covariates were imputed similarly using kNN.
+Categorical covariates were imputed using the `maxCat` function, which chooses the level with the most occurrences.
+Creatinine values were expressed in grams per litre.
 
@@
 
+Stabilized balancing weights were estimated using the energy method available in the `WeightIt` R package [@Greifer:2023].
+This methods estimates weights by minimizing an energy statistic related to covariate balance [@HulingGreiferChen:2023], thus avoiding the need to specify a parametric model.
+Weights below the 0.1 and above the 0.9 quantiles were trimmed.
+Trimming might lead to decreased covariate balance and potentially change the estimand, but can also decrease the variability of the weights.
+Covariate balance was assessed using functionalities provided by the `cobalt` R package [@Greifer:2023a]. Specifically, we used *Love* plots to visualize covariate balance before and after adjusting.
+
 ### G-computation {#sec-gcomp}
 
-- In RQ2, I included the logarithm of the denominator in the RHS. In RQ3, I used the logarithm of the ratio in the RHS.
+We estimated average contrasts with the parametric g-formula, a method of standardization.
+The parametric g-formula involves the following steps: 1) fit a outcome model including both covariates and balancing weights; 2) create two new datasets identical to the original one but with the exposure shifted according to a user-specified intervention set by a deterministic function of the observed exposure levels; 3) use the outcome model to compute adjusted predictions in the two counterfactual datasets; 4) compute the difference between the means of the adjusted predictions in the counterfactual datasets.
+The causal parameter of interest $\phi$ was thus specified as the difference in the expected counterfactual outcomes under the shifted exposure levels $\left( \phi^\Delta = \mathbb{E} \left[ Y^{d_1} \right] - \mathbb{E} \left[ Y^{d_2} \right] \right)$.
+In order for this parameter to be identified, the usual causal identifiability conditions (no unmeasured confounding, positivity, and consistency) are required.
+Since these conditions are likely not satisfied, we focused on the estimation of a statistical estimand that is as close as possible to the causal parameter of interest.
 
-- Describe all statistical methods with assumptions.
-  - Description of outcome model, weighting method, estimand, and balance assessment.
-  - Description of method used to estimate effects (e.g., g-computation).
-  - Description of method used for SE and CI.
-- Describe any methods used to examine subgroups and interactions (sub-group analysis or moderation analysis or analysis of effect-modification).
-- Explain how missing data were addressed.
-- Cohort study: explain how loss to follow-up was addressed.
-- Cross-sectional study: describe analytical methods taking account of sampling strategy.
-- Describe any sensitivity analyses.
-- When referring to previous publications for methods' details, include a brief description of the approach, key assumptions and limitations, and any deviation.
-- Names and version numbers for the used software packages, including non-data arguments if deviating from the default ones.
+We fit the outcome model using the `glm` function and a Gaussian family with identity link from base R.
+The exposure variable was modeled using natural cubic splines with 3 degrees of freedom, to more flexibly capture the \acr{adrf}.
+When the outcome was a ratio, as was the case for cortisol metabolism and 11bHSD activity, we included the logarithm of its denominator, cortisol and cortisol production, respectively, as a control variable [@BartlettPartnoy:2020].
+
+To estimate the average contrasts, we used the `avg_comparisons` function from the `marginaleffects` R package [@Arel-Bundock:2023].
+The two counterfactual datasets were obtained by setting the exposures levels to 90th percentile ($d_1$) and the 10th percentile ($d_2$), for each cohort separately.
+The average contrasts were computed using the estimated balancing weights above.
+Robust standard errors were computed with the `sandwich` R package, using cohort as variable indicating clustering of observations [@Zeileis:2004; @ZeileisKollGraham:2020].
+
+We further estimated the \acr{adrf} using the `avg_predictions` function from the `marginaleffects` R package, examining 50 exposure values from the 10th to the 90th percentiles of the exposure.
+As done for the average contrasts, we included the estimated balancing weights and used cohort as a clustering variable when computing robust standard errors.
+
+### Effect-modification analysis
+
+We tested for possible effect-modification by sex.
+To do so, balancing weights were estimated separately for each level of the sex variable, and an interaction term between the exposure and sex was included in the outcome model.
+Similarly, the average contrasts were aggregated separately for each level of sex.
+We further tested for significance of the difference between the average contrasts of males and females.

### 2023-11-14 output/paper/manuscript.qmd
Added table LLOQs for steroids. Started kinda intro to paper.

@@
 ---
-title: "Some Title"
+title: "Estimating associations between childhood exposure to non-persistent endocrine disruptors, corticosteroids, and neurodevelopment: A study based on the parametric g-formula"
 
-metadata-files:
-  - ../../../../all_phd/authors.yml
+# metadata-files:
+#   - ../../../../all_phd/authors.yml
 
@@
   - acronyms
-  - authors-block
+  # - authors-block
   - ../../../../all_phd/_extensions/wordcount/citeproc.lua
@@
 # Abstract {.unnumbered}
-**Background**
-
-**Objectives**
-
-**Methods**
-
-**Results**
-
-**Discussion**
 
@@
 - The length must be <4,000 words, excluding abstract, references, tables, figure captions, acknowledgments, and Supplementary Material.
-  - Use the AMA format.
   - In-text citations: full-sized Arabic numerals in parentheses within the sentence.
@@
 If uploading a single file (document, tables, figures, SM), designate the file as "Main document - anonymous".
-
-# Introduction {#sec-intro}
-
-**Do not use "Introduction" as a heading**.
 

### 2023-11-14 output/paper/sections/introduction.qmd
Added table LLOQs for steroids. Started kinda intro to paper.

@@
-## Background and rationale {#sec-background}
-
-- Brief review of the literature to summarize current knowledge.
-  - Acknowledge inconsistencies.
-  - For each study, indicate whether it was observational or experimental, and note key characteristics of study populations or experimental models.
-- Explain the scientific background and rationale for the investigation being reported.
-  - Identify knowledge gaps addressed by the current study.
-- Provide context for the study: include information on exposures and outcomes, and why they are relevant to environmental health.
-
-## Objectives {#sec-objectives}
-
-- Provide a clear description of the study hypotheses/aims/objectives, and eventually an overview of the approach used to address them.
+In the present study, we estimated associations between 1) non-persistent \acr{edcs} and attention, 2) non-persistent \acr{edcs} and corticosteroids, and 3) corticosteroids and attention, using the parametric g-formula and average contrasts, in children of a large prospective birth cohort in Europe.

### 2023-11-14 output/paper/si.qmd
Added table LLOQs for steroids. Started kinda intro to paper.

@@
       columns = cas_number
+    )
+  )
+```
+
+```{r lloq-mets}
+#| label: tbl-lloq-mets
+#| tbl-cap: Lower limits of quantification for the corticosteroids
+
+lloq_mets <- load_steroids()$loq |>
+  tidylog::mutate(
+    metabolite = stringr::str_replace(
+      metabolite,
+      "^X", ""
+    )
+  ) |>
+  tidylog::rename(
+    lloq = "loq"
+  )
+gt::gt(lloq_mets |> dplyr::arrange(metabolite)) |>
+  gt::tab_footnote(
+    footnote = "Lower limit of quantification expressed in nanograms per millilitre.",
+    locations = gt::cells_column_labels(
+      columns = lloq
     )

### 2023-11-16 docs/causal_roadmap.qmd
Updated causal roadmap even if not using it for this paper.

--- 2023-08-01 docs/causal_roadmap.qmd
@@
   - Neurobehavior
-metadata-files:
-  - ../../../all_phd/authors.yml
-
 format:
@@
         \usepackage{typearea}
-
 bibliography: ../../../PhD.bib
@@
   chunk_output_type: console
-
 filters:
   - ../../../all_phd/_extensions/acronyms/parse-acronyms.lua
-  - ../../../all_phd/_extensions/authors-block/authors-block.lua
 acronyms:
@@
 <!-- ####################################################################### -->
-## Descriptive analyses
-<!-- ####################################################################### -->
-
-{{< include ../../../all_phd/descriptive_stats.qmd >}}
-
-<!-- ####################################################################### -->
 # Define a realistic statistical model
@@
 causal_graph <- ggdag::dagify(A ~ W + U_A, Y ~ W + A + U_Y, W ~ U_W, 
-                              exposure = "A", outcome = "Y") %>%
+                              exposure = "A", outcome = "Y") |>
   ggdag::tidy_dagitty()
@@
                        Y ~ W_t + X_ti, 
-                       exposure = "X_ti", outcome = "Y") %>%
-  ggdag::tidy_dagitty() %>%
+                       exposure = "X_ti", outcome = "Y") |>
+  ggdag::tidy_dagitty() |>
   ggdag::ggdag(text_size = 2, node_size = 10) + ggdag::theme_dag() +
@@
   coords = coords
-) %>%
-  ggdag::ggdag(., layout = "circle") +
+) |>
+  ggdag::ggdag(layout = "circle") +
   ggdag::geom_dag_point(col = "gray") +
@@
 
-## Reliable Causal Inference
-
-{{< include ../../../all_phd/app-rel-ci.qmd >}}
-
-## Reliable Statistical Inference
-
-{{< include ../../../all_phd/app-rel-si.qmd >}}
-
-## Replicability
-
-{{< include ../../../all_phd/app-replicability.qmd >}}
-
 <!-- ####################################################################### -->
@@
 <!-- ####################################################################### -->
-
-<!-- ####################################################################### -->
-## Data Checking
-<!-- ####################################################################### -->
-
-{{< include ../../../all_phd/app-data-checking.qmd >}}
 
@@
 | `R` | `r I(paste0(R.version$major, ".", R.version$minor))` | Statistical programming environment |
-| `tlverse` | `r packageVersion("tlverse")` | Targeted Learning software ecosystem in R |
 | `lmtp` | `r packageVersion("lmtp")` | Non-parametric Causal Effects of Feasible Interventions Based on Modified Treatment Policies |
-| `sl3` | `r packageVersion("sl3")` | Super Learner ensemble machine learning |
 | `labelled` | `r packageVersion("labelled")` | Manipulate metadata |
 | `dlookr` | `r packageVersion("dlookr")` | Diagnose, explore, and transform data |
-| `Hmisc` | `r packageVersion("Hmisc")` | Harrell Miscellaneous |
 | `dagitty` | `r packageVersion("dagitty")` | Graphical Analysis of Structural Causal Models |
-| `qreport` | `r packageVersion("qreport")` | Facilitating the composition of Quarto reports |
 | ... | ... | ... |
@@
 : Specifications for the `lmtp` R package. {#tbl-lmtp-specs}
-
-<!-- ####################################################################### -->
-## Gene-Environment Interactions
-<!-- ####################################################################### -->
-
-Genetic polymorphisms relevant for the **chemicals**:
-
-* \acr{opps}:
-  * The \acr{snps} of the *PON1* (paraoxonase 1) gene, known to influence the ability to metabolize and detoxify \acr{opps}: `rs854560` and `rs662`. [Reference](https://doi.org/10.1289/ehp.1408976).
-  * The \acr{snps} of the genes coding for \acr{opps}-metabolizing enzymes: `rs1803274`, `rs1799807`, `rs662`, `rs854560`, `rs705379`, `rs4244285`, `rs3892097`, and `rs2177180`.
-[Reference](https://doi.org/10.1016/j.chemosphere.2015.07.079).
-* Phenols:
-  * The \acr{snp} of the *UGT2B15* gene: `rs1902023`. [Reference](https://doi.org/10.1016/j.envint.2020.106172).
-* Phthalates:
-  *  The \acr{snps} of the cytochrome P450 (*CYPs*) and UDP-glucuronosyltransferase (*UGTs*) genes: `rs1057910`, `rs1799853`, `rs4244285`, `rs12248560`, `rs3892097`, `rs1902023`, and `rs11692021`.
-[Reference](https://doi.org/10.1016/j.envint.2021.107046).
-  * The \acr{snps} of the *PPAR*$\alpha$ gene: `rs75525202`, `rs4823902`, `rs4253690`, `rs12330015`, `rs5766698`, `rs115250492`, and `rs1800246`.
-[Reference](https://doi.org/10.3389/fgene.2022.855544).
-Not sure whether these \acr{snps} should be considered in the list below.
-
-Genetic polymorphisms relevant for the **outcomes**:
 

### 2023-11-16 output/paper/figures.qmd
Modified how paper is rendered, moving away from targets for that and simply calling quarto render on SI, tables, figures, and manuscript, and including sections of manuscript in manuscript.qmd. Still, cross ref does not work like this since they are rendered independently.

@@
 title: "Figures"
+
 format:
@@
         \linenumbers
+
 link-citations: true
 linkcolor: blue
+
 editor_options: 
   chunk_output_type: console
+
 execute:
@@
 ```{r}
-source("code/proc_res.R")
+here::i_am("output/paper/figures.qmd")
 ```
@@
 ```{r vars}
+source(here::here("code/dictionaries.R"))
+source(here::here("code/data.R"))
+source(here::here("code/proc_res.R"))
+
+is_hpc <- FALSE
 store <- ifelse(
-  Sys.getenv("is_hpc"), 
+  is_hpc, 
   paste0(
@@
 
+id_var <- "HelixID"
 num_digits_est = 3
 num_digits_sig = 2
+params <- params(is_hpc = is_hpc)
+threshold_within_lod <- params$steps$rq2$preproc_outcome$llodq$threshold_within
+threshold_overall_lod <- params$steps$rq2$preproc_outcome$llodq$threshold_overall
+threshold_within_nan <- params$steps$rq2$preproc_outcome$missings$threshold_within
+threshold_overall_nan <- params$steps$rq2$preproc_outcome$missings$threshold_overall
+knn_k <- params$steps$rq2$preproc_outcome$missings$k
 ```
 
-- Use log-scale when plotting relative measures of effects.
 - Figure titles and legends should be submitted as single paragraph.

### 2023-11-16 output/paper/manuscript.qmd
Modified how paper is rendered, moving away from targets for that and simply calling quarto render on SI, tables, figures, and manuscript, and including sections of manuscript in manuscript.qmd. Still, cross ref does not work like this since they are rendered independently.

@@
     toc: false
-    keep-tex: true
+    keep-tex: false
     include-in-header:
@@
 ---
+
+```{r}
+here::i_am("output/paper/manuscript.qmd")
+```
 
@@
 # References {.unnumbered}
-
-::: {#refs}
-:::

### 2023-11-16 output/paper/sections/methods.qmd
Added table LLOQs for steroids. Started kinda intro to paper.
Modified how paper is rendered, moving away from targets for that and simply calling quarto render on SI, tables, figures, and manuscript, and including sections of manuscript in manuscript.qmd. Still, cross ref does not work like this since they are rendered independently.

@@
+```{r vars}
+source(here::here("code/dictionaries.R"))
+
+params <- params(is_hpc = FALSE)
+threshold_within_lod <- params$steps$rq2$preproc_outcome$llodq$threshold_within
+threshold_overall_lod <- params$steps$rq2$preproc_outcome$llodq$threshold_overall
+threshold_within_nan <- params$steps$rq2$preproc_outcome$missings$threshold_within
+threshold_overall_nan <- params$steps$rq2$preproc_outcome$missings$threshold_overall
+knn_k <- params$steps$rq2$preproc_outcome$missings$k
+```
+
 ## Study population and design {#sec-design}
@@
 
-```{r params-preproc}
-threshold_within_lod <- 20
-threshold_overall_lod <- 10
-threshold_within_nan <- 40
-threshold_overall_nan <- 30
-knn_k <- 5
-```
-
 ### Data pre-processing {#sec-dat-preproc}
@@
 We proceeded to impute these values using half the value of the corresponding \acr{lloq}, for those metabolites that had less than `r threshold_within_lod`% of missings within each cohort and `r threshold_overall_lod`% of missings overall.
+Information about the \acr{lloq} for the corticosteroids is provided in @tbl-lloq-mets.
 The remaining missing values were imputed using kNN from the `VIM` R package [@KowarikTempl:2016], for those metabolites that had less than `r threshold_within_nan`% of remaining missings within each cohort and `r threshold_overall_nan`% of remaining missings overall.
@@
 Concentrations below the \acr{lod} were singly imputed using a quantile regression approach for the imputation of left-censored missing data, as implemented in the `impute.QRILC` function from the `imputeLCMD` R package [@lazar2015imputelcmd].
+Information about the lower limits of detection can be found in @HaugSakhiCequier:2018.
 Chemicals with more than 70% of observations below the \acr{lod} were not considered in the present study.
@@
 Robust standard errors were computed with the `sandwich` R package, using cohort as variable indicating clustering of observations [@Zeileis:2004; @ZeileisKollGraham:2020].
+For each outcome, we report the results as differences between average contrasts.
 

### 2023-11-16 output/paper/si.qmd
Modified how paper is rendered, moving away from targets for that and simply calling quarto render on SI, tables, figures, and manuscript, and including sections of manuscript in manuscript.qmd. Still, cross ref does not work like this since they are rendered independently.

@@
 ```{r}
-source("code/dictionaries.R")
-source("code/data.R")
-source("code/proc_res.R")
+here::i_am("output/paper/si.qmd")
 ```
@@
 num_digits_sig = 2
+params_ <- params(is_hpc = is_hpc)
+threshold_within_lod <- params_$steps$rq2$preproc_outcome$llodq$threshold_within
+threshold_overall_lod <- params_$steps$rq2$preproc_outcome$llodq$threshold_overall
+threshold_within_nan <- params_$steps$rq2$preproc_outcome$missings$threshold_within
+threshold_overall_nan <- params_$steps$rq2$preproc_outcome$missings$threshold_overall
+knn_k <- params_$steps$rq2$preproc_outcome$missings$k
 ```
@@
 mets_info <- readODS::read_ods(
-  path = "docs/steroids.ods"
+  path = here::here("docs", "steroids.ods")
 ) |>

### 2023-11-16 output/paper/tables.qmd
Modified how paper is rendered, moving away from targets for that and simply calling quarto render on SI, tables, figures, and manuscript, and including sections of manuscript in manuscript.qmd. Still, cross ref does not work like this since they are rendered independently.

@@
 ```{r}
-source("code/proc_res.R")
+here::i_am("output/paper/tables.qmd")
 ```
@@
 
+id_var <- "HelixID"
 num_digits_est = 3
 num_digits_sig = 2
+params <- params(is_hpc = is_hpc)
+threshold_within_lod <- params$steps$rq2$preproc_outcome$llodq$threshold_within
+threshold_overall_lod <- params$steps$rq2$preproc_outcome$llodq$threshold_overall
+threshold_within_nan <- params$steps$rq2$preproc_outcome$missings$threshold_within
+threshold_overall_nan <- params$steps$rq2$preproc_outcome$missings$threshold_overall
+knn_k <- params$steps$rq2$preproc_outcome$missings$k
 ```

### 2023-11-16 _targets_sa_rqX.R

@@ -25,15 +25,9 @@ outcome <- switch(rq,
 
 if (rq == "rq2") {
   tbl_outcomes <- tibble::tibble(
-    name = c("cortisol_production",
-             "cortisol_metabolism",
-             "cortisone_production",
-             "X11bHSD") |>
+    name = vars_of_interest()$new_metabolites |>
       stringr::str_to_lower(),
-    outcome = c("cortisol_production",
-                "cortisol_metabolism",
-                "cortisone_production",
-                "X11bHSD")
+    outcome = vars_of_interest()$new_metabolites
   )
 } else {
   tbl_outcomes <- tibble::tibble(

### 2023-11-16 code/data.R

@@ -6,22 +6,34 @@
 #' 
 #' @export
 create_steroid_scores <- function(dat) {
+  warning(
+    "Computation of `androgens_production` is wrong since AED is missing.",
+    call. = TRUE
+  )
+  
   dat_ret <- dat |>
     tidylog::mutate(
-      cortisol_production = rowSums(dplyr::across(c(
-        `F`, X20aDHF, X20bDHF, X5bDHF, X5aTHF, X5bTHF,
-        X6OHF, X5a20acortol, X5a20bcortol,
-        X5b20acortol, X5b20bcortol
-      ))), 
-      cortisol_metabolism = rowSums(dplyr::across(c(
-        X20aDHF, X20bDHF, X5bDHF, X5aTHF, X5bTHF,
-        X6OHF, X5a20acortol, X5a20bcortol, X5b20acortol, X5b20bcortol
-      ))) / `F`,
-      cortisone_production = rowSums(dplyr::across(c(
-        CortisoneE, X20aDHE, X20bDHE, X5aTHE, X5bTHE, X6OHE,
-        X5b20acortolone, X5b20bcortolone
-      ))),
-      X11bHSD = cortisone_production / cortisol_production
+      # Cortisol
+      cortisol_production = rowSums(dplyr::across(c(`F`, X20aDHF, X20bDHF, X5bDHF, X5aTHF, X5bTHF, X6OHF, X5a20acortol, X5a20bcortol, X5b20acortol, X5b20bcortol))), 
+      cortisol_metabolism = rowSums(dplyr::across(c(X20aDHF, X20bDHF, X5bDHF, X5aTHF, X5bTHF, X6OHF, X5a20acortol, X5a20bcortol, X5b20acortol, X5b20bcortol))) / `F`,
+      # Cortisone
+      cortisone_production = rowSums(dplyr::across(c(CortisoneE, X20aDHE, X20bDHE, X5aTHE, X5bTHE, X6OHE, X5b20acortolone, X5b20bcortolone))),
+      cortisone_metabolism = rowSums(dplyr::across(c(X20aDHE, X20bDHE, X5aTHE, X5bTHE, X6OHE, X5b20acortolone, X5b20bcortolone))) / CortisoneE,
+      # 11beta
+      X11bHSD = cortisone_production / cortisol_production,
+      # Global reductases
+      #global_reductase_f = rowSums(dplyr::across(c(X5aTHF, X5bTHF))) / `F`,
[…]

### 2023-11-16 code/proc_res.R

@@ -1061,7 +1061,6 @@ viz_desc_vars <- function(dat, vars, fct_levels, is_chem) {
     ggplot2::scale_fill_distiller(palette = "Blues", direction = 1) +
     ggplot2::coord_fixed() +
     ggplot2::facet_grid(cohort ~ .) +
-    ggplot2::labs(caption = "1: quantifiable, 2: <LOD, 3: interference or out of range, 4: no sample") +
     ggplot2::theme(
       axis.line = ggplot2::element_line(color = "black"),
       plot.background = ggplot2::element_blank(),
@@ -1074,6 +1073,14 @@ viz_desc_vars <- function(dat, vars, fct_levels, is_chem) {
         hjust = 1
       )
     )
+  
+  if (is_chem) {
+    plt <- plt +
+      ggplot2::labs(caption = "1: quantifiable, 2: <LOD, 3: interference or out of range, 4: not analysed")
+  } else {
+    plt <- plt +
+      ggplot2::labs(caption = "1: quantifiable, 2: <LOQ, 3: interference or out of range, 4: not detected")
+  }
   
   df_plot <- df_plot |>
     dplyr::arrange(

### 2023-11-16 code/utils.R

@@ -365,7 +365,9 @@ rq_prepare_data <- function(dat) {
     cols_to_remove <- setdiff(
       colnames(dat$exposures),
       c(params_dat$variables$identifier,
-        "F", "cortisol_production")
+        "F", "cortisol_production", "CortisoneE",
+        "X5bTHS", "X5aTHB", "X5bTHB",
+        "Etio", "X17HP", "PT")
     )
     dat$exposures <- create_steroid_scores(dat = dat$exposures)
     
@@ -406,7 +408,9 @@ rq_prepare_data <- function(dat) {
     cols_to_remove <- setdiff(
       colnames(dat$outcome),
       c(params_dat$variables$identifier,
-        "F", "cortisol_production")
+        "F", "cortisol_production", "CortisoneE",
+        "X5bTHS", "X5aTHB", "X5bTHB",
+        "Etio", "X17HP", "PT")
     )
     dat$outcome <- create_steroid_scores(dat = dat$outcome)
     
@@ -459,10 +463,7 @@ rq_estimate_weights <- function(dat, by = NULL, save_results,
   
   # Step 1: estimate weights for covariate balance
   if (rq %in% c("rq3", "rq4")) {
-    list_exposures <- c("cortisol_production",
-                        "cortisol_metabolism",
-                        "cortisone_production",
-                        "X11bHSD")
+    list_exposures <- vars_of_interest()$new_metabolites
   } else {
     list_exposures <- names(dat$exposures)
   }
@@ -643,7 +644,16 @@ rq_fit_model_weighted <- function(dat, outcome, by = c(),
   if (rq == "rq2") {
     den <- switch(outcome,
                   "cortisol_metabolism" = "F",
+                  "cortisone_metabolism" = "CortisoneE",
                   "X11bHSD" = "cortisol_production",
+                  "global_reductase_f" = "F",
+                  "global_reductase_e" = "CortisoneE",
+                  "cyp3a4" = "F",
+                  "X11hydroxylase" = "X5bTHS",
+                  "X17hydroxylase" = c("X5aTHB", "X5bTHB"),
+                  "X5a_reductase" = "Etio",
+                  "lyase" = c("X17HP", "PT"),
+                  "global_adrenal_function" = c("X17HP", "PT"),
                   NULL
     )
     dat$covariates <- dat$covariates |>
@@ -664,10 +674,7 @@ rq_fit_model_weighted <- function(dat, outcome, by = c(),
       outcome
     )))
   if (rq %in% c("rq3", "rq4")) {
[…]

### 2023-11-16 output/paper/render.R

@@ -1,0 +1,4 @@
+quarto::quarto_render(input = here::here("output/paper/figures.qmd"))
+quarto::quarto_render(input = here::here("output/paper/tables.qmd"))
+quarto::quarto_render(input = here::here("output/paper/manuscript.qmd"))
+quarto::quarto_render(input = here::here("output/paper/si.qmd"))

### 2023-11-21 output/paper/_figures.qmd
Added presentation for group meeting regarding paper 3. Moved individual sections manuscript directly in manuscript. Added include of SI and tables and figures in manuscript for cross referencing.

--- 2023-11-08 output/paper/figures.qmd
@@
----
-title: "Figures"
-format:
-  pdf:
-    pdf-engine: lualatex
-    number-sections: true
-    toc: false
-    include-in-header:
-      text: |
-        \usepackage{lineno}
-        \usepackage{rotating}
-        \usepackage{float}
-        \usepackage{typearea}
-    include-before-body:
-      text: |
-        \linenumbers
-link-citations: true
-linkcolor: blue
-editor_options: 
-  chunk_output_type: console
-execute:
-  echo: false
-  warning: false
-  error: false
-  message: false
----
+<!-- title: "Figures" -->
+
+<!-- format: -->
+<!--   docx: -->
+<!--     reference-doc: reference_document.docx -->
+<!--   pdf: -->
+<!--     pdf-engine: lualatex -->
+<!--     number-sections: true -->
+<!--     toc: false -->
+<!--     include-in-header: -->
+<!--       text: | -->
+<!--         \usepackage{lineno} -->
+<!--         \usepackage{rotating} -->
+<!--         \usepackage{float} -->
+<!--         \usepackage{typearea} -->
+<!--     include-before-body: -->
+<!--       text: | -->
+<!--         \linenumbers -->
+
+<!-- link-citations: true -->
+<!-- linkcolor: blue -->
+
+<!-- editor_options: -->
+<!--   chunk_output_type: console -->
+
+<!-- execute: -->
+<!--   echo: false -->
+<!--   warning: false -->
+<!--   error: false -->
+<!--   message: false -->
 
 ```{r}
-source("code/proc_res.R")
+here::i_am("output/paper/_figures.qmd")
 ```
 
-```{r vars}
-store <- ifelse(
-  Sys.getenv("is_hpc"), 
-  paste0(
-    "/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/",
-    "lorenzoF/data/data_paper3/_targets/_targets_res"
-  ), 
-  "~/mounts/rstudioserver/data/data_paper3/_targets/_targets_res"
-)
-
-num_digits_est = 3
-num_digits_sig = 2
-```
-
-- Use log-scale when plotting relative measures of effects.
 - Figure titles and legends should be submitted as single paragraph.

### 2023-11-21 output/paper/_si.qmd
Added presentation for group meeting regarding paper 3. Moved individual sections manuscript directly in manuscript. Added include of SI and tables and figures in manuscript for cross referencing.

--- 2023-11-16 output/paper/si.qmd
@@
----
-title: "Supplementary Material"
-
-format:
-  docx:
-    reference-doc: reference_document.docx
-
-crossref:
-  fig-title: Supplementary Figure
-  tbl-title: Supplementary Table
-
-bibliography: bibliography.bib
-csl: ama11.csl
-
-link-citations: true
-linkcolor: blue
-
-editor_options: 
-  chunk_output_type: console
-
-execute:
-  echo: false
-  warning: false
-  error: false
-  message: false
----
-
 ```{r}
-here::i_am("output/paper/si.qmd")
-```
-
-```{r vars}
-source(here::here("code/dictionaries.R"))
-source(here::here("code/data.R"))
-source(here::here("code/proc_res.R"))
-
-is_hpc <- FALSE
-store <- ifelse(
-  is_hpc, 
-  paste0(
-    "/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/",
-    "lorenzoF/data/data_paper3/_targets/_targets_res"
-  ), 
-  "~/mounts/rstudioserver/data/data_paper3/_targets/_targets_res"
-)
-
-id_var <- "HelixID"
-num_digits_est = 3
-num_digits_sig = 2
-params_ <- params(is_hpc = is_hpc)
-threshold_within_lod <- params_$steps$rq2$preproc_outcome$llodq$threshold_within
-threshold_overall_lod <- params_$steps$rq2$preproc_outcome$llodq$threshold_overall
-threshold_within_nan <- params_$steps$rq2$preproc_outcome$missings$threshold_within
-threshold_overall_nan <- params_$steps$rq2$preproc_outcome$missings$threshold_overall
-knn_k <- params_$steps$rq2$preproc_outcome$missings$k
+here::i_am("output/paper/_si.qmd")
 ```
@@
 ```{r codebook-1}
-#| label: tbl-codebook-1
+#| label: supptbl-codebook-1
 #| tbl-cap: Codebook for covariates used in RQ1
@@
 ```{r codebook-2}
-#| label: tbl-codebook-2
+#| label: supptbl-codebook-2
 #| tbl-cap: Codebook for covariates used in RQ2
@@
 ```{r codebook-3}
-#| label: tbl-codebook-3
+#| label: supptbl-codebook-3
 #| tbl-cap: Codebook for covariates used in RQ3
@@
 ```{r info-chems}
-#| label: tbl-info-chems
+#| label: supptbl-info-chems
 #| tbl-cap: Information about non-persistent EDCs
@@
 ```{r info-mets}
-#| label: tbl-info-mets
+#| label: supptbl-info-mets
 #| tbl-cap: Information about metabolites
@@
 ```{r lloq-mets}
-#| label: tbl-lloq-mets
+#| label: supptbl-lloq-mets
 #| tbl-cap: Lower limits of quantification for the corticosteroids
@@
 ```{r weights-1sa}
-#| label: tbl-weights-1sa
+#| label: supptbl-weights-1sa
 #| tbl-cap: Weights
@@
 ```{r weights-2sa}
-#| label: tbl-weights-2sa
+#| label: supptbl-weights-2sa
 #| tbl-cap: Weights
@@
 ```{r weights-3sa}
-#| label: tbl-weights-3sa
+#| label: supptbl-weights-3sa
 #| tbl-cap: Weights
@@
 ```{r desc-chems-lod}
-#| label: fig-desc-chems
+#| label: suppfig-desc-chems
 #| fig-cap: Description of chemicals
@@
 ```{r desc-mets-lod}
-#| label: fig-desc-mets
+#| label: suppfig-desc-mets
 #| fig-cap: Description of metabolites
@@
 ```{r adrfs-rq3}
-#| label: fig-adrfs-rq3
+#| label: suppfig-adrfs-rq3
 #| fig-cap: Average dose-response functions for RQ3
-#| fig-subcap:
-#|   - "Cortisol production"
-#|   - "11bHSD"
 #| fig-width: 6
 #| fig-height: 7
-#| layout-ncol: 2
 
@@
 ret_adrfs[[1]]
-ret_adrfs[[2]]
+#ret_adrfs[[2]]
 ```
@@
 ```{r marginal-1sa}
-#| label: fig-marginal-1sa
+#| label: suppfig-marginal-1sa
 #| fig-cap: Average comparisons for effect modification
@@
 ```{r marginal-2sa}
-#| label: fig-marginal-2sa
+#| label: suppfig-marginal-2sa
 #| fig-cap: Average comparisons for effect modification
@@
 ```{r marginal-3sa}
-#| label: fig-marginal-3sa
+#| label: suppfig-marginal-3sa
 #| fig-cap: Average comparisons for effect modification
@@
 ```
-
-# References {.unnumbered}
-
-::: {#refs}
-:::

### 2023-11-21 output/paper/_tables.qmd
Added presentation for group meeting regarding paper 3. Moved individual sections manuscript directly in manuscript. Added include of SI and tables and figures in manuscript for cross referencing.

--- 2023-11-10 output/paper/tables.qmd
@@
----
-title: "Tables"
+<!-- title: "Tables" -->
 
-format:
-  docx:
-    reference-doc: reference_document.docx
+<!-- format: -->
+<!--   docx: -->
+<!--     reference-doc: reference_document.docx -->
 
-link-citations: true
-linkcolor: blue
+<!-- link-citations: true -->
+<!-- linkcolor: blue -->
 
-editor_options:
-  chunk_output_type: console
+<!-- editor_options: -->
+<!--   chunk_output_type: console -->
 
-execute:
-  echo: false
-  warning: false
-  error: false
-  message: false
----
+<!-- execute: -->
+<!--   echo: false -->
+<!--   warning: false -->
+<!--   error: false -->
+<!--   message: false -->
 
 ```{r}
-source("code/proc_res.R")
-```
-
-```{r vars}
-store <- ifelse(
-  Sys.getenv("is_hpc"), 
-  paste0(
-    "/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/",
-    "lorenzoF/data/data_paper3/_targets/_targets_res"
-  ), 
-  "~/mounts/rstudioserver/data/data_paper3/_targets/_targets_res"
-)
-
-num_digits_est = 3
-num_digits_sig = 2
+here::i_am("output/paper/_tables.qmd")
 ```

### 2023-11-21 output/paper/manuscript.qmd
Added presentation for group meeting regarding paper 3. Moved individual sections manuscript directly in manuscript. Added include of SI and tables and figures in manuscript for cross referencing.

@@
     reference-doc: reference_document.docx
-  pdf:
-    pdf-engine: lualatex
-    number-sections: true
-    toc: false
-    keep-tex: false
-    include-in-header:
-      text: |
-        \usepackage{lineno}
-        \usepackage{rotating}
-        \usepackage{float}
-        \usepackage{typearea}
-    include-before-body:
-      text: |
-        \linenumbers
+    citeproc: false
+  # pdf:
+  #   pdf-engine: lualatex
+  #   number-sections: true
+  #   toc: false
+  #   keep-tex: false
+  #   include-in-header:
+  #     text: |
+  #       \usepackage{lineno}
+  #       \usepackage{rotating}
+  #       \usepackage{float}
+  #       \usepackage{typearea}
+  #   include-before-body:
+  #     text: |
+  #       \linenumbers
+
+crossref:
+  custom: 
+    - kind: float
+      ref-type: supptbl
+      latex-env: supptbl
+      prefix: Table S
+      name: Table S
+      space-before-numbering: false
+    - kind: float
+      ref-type: suppfig
+      latex-env: suppfig
+      prefix: Figure S
+      name: Figure S
+      space-before-numbering: false
 
@@
 
+```{r vars-methods}
+source(here::here("code/dictionaries.R"))
+source(here::here("code/data.R"))
+source(here::here("code/proc_res.R"))
+
+is_hpc <- FALSE
+store <- ifelse(
+  is_hpc, 
+  paste0(
+    "/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/",
+    "lorenzoF/data/data_paper3/_targets/_targets_res"
+  ), 
+  "~/mounts/rstudioserver/data/data_paper3/_targets/_targets_res"
+)
+id_var <- "HelixID"
+num_digits_est = 3
+num_digits_sig = 2
+params_ <- params(is_hpc = is_hpc)
+threshold_within_lod <- params_$steps$rq2$preproc_outcome$llodq$threshold_within
+threshold_overall_lod <- params_$steps$rq2$preproc_outcome$llodq$threshold_overall
+threshold_within_nan <- params_$steps$rq2$preproc_outcome$missings$threshold_within
+threshold_overall_nan <- params_$steps$rq2$preproc_outcome$missings$threshold_overall
+knn_k <- params_$steps$rq2$preproc_outcome$missings$k
+```
+
 # Abstract {.unnumbered}
 
-\newpage
+{{< pagebreak >}}
 
@@
 
-{{< include sections/introduction.qmd >}}
+The prevalence of several neurodevelopmental disorders has increased in the pediatric population [@GrandjeanLandrigan:2014], and multiple environmental pollutants play a role in the increased rates of these disorders [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
+\acr{edcs}, chemicals that are capable of interfering with the endocrine system, have been shown to have a role in childhood neurodevelopment [@Braun:2017].
+Although both pregnancy and early childhood are crucial stages of (neuro)development, most of the available literature is focused on the effects of prenatal exposure to \acr{edcs} on child neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
+Moreover, little is known about the biological mechanisms of action [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
+
+\acr{opps} might have a deleterious effect on neurodevelopment, although the few studies conducted on children and using a biological matrix for exposure assessment suffered from a series of limitations, including a small sample size and few \acr{opp} metabolites measured [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
+Phthalates and their metabolites have been associated to several adverse effects on neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
+Similarly, these studies focused on few phthalates metabolites and were conducted in small populations.
+On the other hand, the effects of exposure to \acr{bpa} on cognitive functions are still unclear [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
+
+Another major limitation of these epidemiological studies, especially when non-persistent chemicals are analyzed, is the reliance on single-spot urine samples for exposure assessment, which might not be representative of long-term exposures.
+
+Something about corticosteroids...
+
+In the present study, we estimated associations between 1) non-persistent \acr{edcs} and attention, 2) non-persistent \acr{edcs} and corticosteroids, and 3) corticosteroids and attention, using the parametric g-formula and average contrasts, in children of a large prospective birth cohort in Europe.
 
@@
 
-{{< include sections/methods.qmd >}}
+## Study population and design {#sec-design}
+
+The \acr{helix} is an ongoing project which aims to characterize early-life exposures and their potential association with endogenous biomarkers and health outcomes [@VrijheidSlamaRobinson:2014].
+It consists of six existing population-based birth cohort studies across Europe: BiB (Born in Bradford, UK) [@WrightSmallRaynor:2013], EDEN (Study of determinants of pre- and postnatal developmental, France) [@HeudeForhanSlama:2016], INMA (Environment and Childhood, Spain) [@GuxensBallesterEspada:2012], KANC (Kaunas Cohort, Lithuania) [@GrazulevicieneDanileviciuteNadisauskiene:2009], MoBa (The Norwegian Mother and Child Cohort Study, Norway) [@MagnusIrgensHaug:2006], and Rhea (Mother–Child Cohort in Crete, Greece) [@ChatziPlanaDaraki:2009], for a total of 32,000 mother-child pairs.
+A \acr{helix} subcohort of 1,200 mother-child pairs was fully characterized for the external and internal exposome, including exposure and omics biomarkers during childhood.
+Eligibility criteria for inclusion in the \acr{helix} subcohort included: a) age 6-11 years, with a preference for 7-9 years; b) availability of sufficient stored pregnancy blood and urine samples; c) availability of complete address history from first to last follow-up; d) no serious health problems, which might affect the results of the clinical testing.
+Further information can be found in @MaitreBontCasas:2018.
+
+Ethical permission was obtained from the relevant authorities in the corresponding country.
+
+## Variables {#sec-vars}
+
+### Confounders {#sec-confounders}
+
+For each research question, defined by a specific type of exposure and outcome, the minimal set of covariates for inclusion in the analyses was selected on the basis of a \acr{dag} built with DAGitty [@TextorvanderZanderGilthorpe:2016] and ggdag [@Barrett:2023].
+The sets of covariates were selected to estimate the total effect of the exposure on the outcome.
+Further, each minimal adjustment set was *augmented* with precision covariates, defined as the set of parents variable of the outcome that are not parents of the exposure.
+The adjustment sets are provided in the Supplementary Material as text files compatible with DAGitty.
+
+- For RQ1 I used creatinine values from HELIX. For RQ3 the ones from the steroids dataset. For RQ2, I included in the model both variables.
+
+### Endocrine disrupting chemicals {#sec-edcs}
+
+Children were assessed between December 2013 and February 2016, and included neurological testing and urine collection.
+Urine samples of the night before and the first morning void on the day of the visit were combined to provide a more reliable exposure assessment.
+Non-persistent \acr{edcs} assessed in urine samples from children included phthalate metabolites, phenols, and \acr{op} pesticide metabolites.
+A list of the environmental chemicals determined in urine samples and used for the present study is given in @supptbl-info-chems.
+The laboratory protocols for the analysis are described elsewhere [@HaugSakhiCequier:2018].
+
+### Corticosteroids {#sec-steroids}
+
+Urine samples of the night before the day of the visit were used to measure levels of the corticosteroids.
+These included glucocorticosteroids, glucocorticosteroid metabolites, glucocorticosteroid precursors, glucocorticosteroid precursor metabolites, androgens, and androgen metabolites.
+A list of the corticosteroids determined in urine samples and used for the present study is given in @supptbl-info-mets.
+
+To assess the levels of corticosteroids and their metabolites, LC-MS/MS analysis was applied at the Applied Metabolomics Research Group, IMIM (Hospital del Mar Medical Research Institute).
+The laboratory protocols for the analysis are described elsewhere [@MarcosRenauCasals:2014; @Gomez-GomezPozo:2020].
+Of the 1,004 urine samples, 980 children were matched to the \acr{helix} subcohort.
+
+Three additional markers, cortisol production, cortisol metabolism, cortisone production, and 11bHSD activity, were computed based on the following: cortisol production as the sum of cortisol and its metabolites (20aDHF, 20bDHF, 5bDHF, 5aTHF, 5bTHF, 6OHF, 5a20acortol, 5a20bcortol, 5b20acortol, 5b20bcortol), cortisol metabolism as the inverse of the ratio between cortisol and its metabolites, cortisone production as the sum of cortisone and its metabolites (20aDHE, 20bDHE, 5aTHE, 5bTHE, 6OHE, 5b20acortolone, b20bcortolone), and 11bHSD activity as the ratio between cortisone production and cortisol production.
+11bHSD activity gives a measure of conversion of cortisone to cortisol.
+
+### Neurodevelopment {#sec-neurodevelopment}
+
+Neurodevelopmental outcomes were assessed with standardized, non-linguistic, and culturally blind computer tests, including the \acr{ant} [@RuedaFanMcCandliss:2004].
+Further information can be found in @MaitreBontCasas:2018.
+Briefly, it is a computerized test that provides a measure of efficiency in three different functions of attention: alerting, orienting, and executive attention.
+The outcome of interest for the present study is the \acr{hitrtse} [@SunyerEsnaolaAlvarez-Pedrerol:2015], a measure of response speed consistency throughout the test.
+A high \acr{hitrtse} indicates highly variable reactions, and is considered a measure of inattentiveness.
+
+## Statistical methods {#sec-stat-methods}
+
+### Data pre-processing {#sec-dat-preproc}
+
+Concentrations of the corticosteroids were classified as quantifiable, below the \acr{loq}, possible interference or out of range, and not detected.
+For each metabolite, we computed the fraction of values below the \acr{loq} and not detected, both within each cohort and overall.
+We proceeded to impute these values using half the value of the corresponding \acr{lloq}, for those metabolites that had less than `r threshold_within_lod`% of missings within each cohort and `r threshold_overall_lod`% of missings overall.
+Information about the \acr{lloq} for the corticosteroids is provided in @supptbl-lloq-mets.
+The remaining missing values were imputed using kNN from the `VIM` R package [@KowarikTempl:2016], for those metabolites that had less than `r threshold_within_nan`% of remaining missings within each cohort and `r threshold_overall_nan`% of remaining missings overall.
+We used `r knn_k` nearest neighbors.
+We natural log-transformed them to improve model fit, assessed with posterior predictive checks.
+To do so, replicated data were simulated with the fitted models and compared to the observed data.
+We used the `check_predictions` function from the `performance` R package using the default arguments [@LudeckeBen-ShacharPatil:2021].
+Values of cortisol production and cortisone production were expressed in nanograms per millilitre, whereas values of cortisol metabolism and 11bHSD activity were unitless.
+
+Concentrations of the non-persistent \acr{edcs} were classified as quantifiable, below the \acr{lod}, possible interference or out of range, and not analysed.
+Concentrations below the \acr{lod} were singly imputed using a quantile regression approach for the imputation of left-censored missing data, as implemented in the `impute.QRILC` function from the `imputeLCMD` R package [@lazar2015imputelcmd].
+Information about the lower limits of detection can be found in @HaugSakhiCequier:2018.
+Chemicals with more than 70% of observations below the \acr{lod} were not considered in the present study.
+Remaining missing values were imputed similarly using kNN.
+Values of the chemicals were expressed in $\mu$grams per litre.
+
+Missing values in the clinical outcome were imputed similarly using kNN.
+Similarly, we natural log-transformed them to improve model fit, assessed with posterior predictive checks.
+Values of the clinical outcome were expressed in \acr{mss}.
+
+Missing values in the covariates were imputed similarly using kNN.
+Categorical covariates were imputed using the `maxCat` function, which chooses the level with the most occurrences.
+Creatinine values were expressed in grams per litre.
+
+### Estimation of balancing weights {#sec-weights-est}
+
+Stabilized balancing weights were estimated using the energy method available in the `WeightIt` R package [@Greifer:2023].
+This methods estimates weights by minimizing an energy statistic related to covariate balance [@HulingGreiferChen:2023], thus avoiding the need to specify a parametric model.
+Weights below the 0.1 and above the 0.9 quantiles were trimmed.
+Trimming might lead to decreased covariate balance and potentially change the estimand, but can also decrease the variability of the weights.
+Covariate balance was assessed using functionalities provided by the `cobalt` R package [@Greifer:2023a]. Specifically, we used *Love* plots to visualize covariate balance before and after adjusting.
+
+### G-computation {#sec-gcomp}
+
+We estimated average contrasts with the parametric g-formula, a method of standardization.
+The parametric g-formula involves the following steps: 1) fit a outcome model including both covariates and balancing weights; 2) create two new datasets identical to the original one but with the exposure shifted according to a user-specified intervention set by a deterministic function of the observed exposure levels; 3) use the outcome model to compute adjusted predictions in the two counterfactual datasets; 4) compute the difference between the means of the adjusted predictions in the counterfactual datasets.
+The causal parameter of interest $\phi$ was thus specified as the difference in the expected counterfactual outcomes under the shifted exposure levels $\left( \phi^\Delta = \mathbb{E} \left[ Y^{d_1} \right] - \mathbb{E} \left[ Y^{d_2} \right] \right)$.
+In order for this parameter to be identified, the usual causal identifiability conditions (no unmeasured confounding, positivity, and consistency) are required.
+Since these conditions are likely not satisfied, we focused on the estimation of a statistical estimand that is as close as possible to the causal parameter of interest.
+
+We fit the outcome model using the `glm` function and a Gaussian family with identity link from base R.
+The exposure variable was modeled using natural cubic splines with 3 degrees of freedom, to more flexibly capture the \acr{adrf}.
+When the outcome was a ratio, as was the case for cortisol metabolism and 11bHSD activity, we included the logarithm of its denominator, cortisol and cortisol production, respectively, as a control variable [@BartlettPartnoy:2020].
+
+To estimate the average contrasts, we used the `avg_comparisons` function from the `marginaleffects` R package [@Arel-Bundock:2023].
+The two counterfactual datasets were obtained by setting the exposures levels to 90th percentile ($d_1$) and the 10th percentile ($d_2$), for each cohort separately.
+The average contrasts were computed using the estimated balancing weights above.
+Robust standard errors were computed with the `sandwich` R package, using cohort as variable indicating clustering of observations [@Zeileis:2004; @ZeileisKollGraham:2020].
+For each outcome, we report the results as differences between average contrasts.
+
+We further estimated the \acr{adrf} using the `avg_predictions` function from the `marginaleffects` R package, examining 50 exposure values from the 10th to the 90th percentiles of the exposure.
+As done for the average contrasts, we included the estimated balancing weights and used cohort as a clustering variable when computing robust standard errors.
+
+### Effect-modification analysis
+
+We tested for possible effect-modification by sex.
+To do so, balancing weights were estimated separately for each level of the sex variable, and an interaction term between the exposure and sex was included in the outcome model.
+Similarly, the average contrasts were aggregated separately for each level of sex.
+We further tested for significance of the difference between the average contrasts of males and females.
 
@@
 
-{{< include sections/results.qmd >}}
+## Participants {#sec-res-participants}
+
+## Descriptive data {#sec-res-pop-desc}
+
+## Outcome data {#sec-res-out-desc}
+
+## Main results {#sec-res-main}
+
+- All results on which study conclusions or inferences are based, including null findings and results of secondary or sensitivity analyses, must be reported.
+Use of sub-headings that describe the nature of the results (but no declarative statements).
+  - Provide a clear and concise description of all findings without extrapolating beyond the study results.
+  - Do not limit results to those *statistically significant* or that support the study hypotheses.
+Avoid using statistical significance testing as the sole or primary criterion for interpreting the obtained results.
+If significance testing or *p*-values are used, report numeric *p*-values, rounded to 1-2 digits, for all results.
+Use an uppercase italic letter "P", and the values should not be bolded.
+Indicate whether are 1- or 2-sided.
+- Give unadjusted and confounder-adjusted estimates and their precision (e.g., $95\%$ confidence interval).
+Make clear which confounders were adjusted for and why they were included.
+Include the number of observations for each analysis after accounting for missing data.
+Include numeric data within figures (e.g., forest plots), or provide tables with corresponding numeric data for all figures.
+  - [`marginaleffects` tables](https://vincentarelbundock.github.io/marginaleffects/articles/tables.html).
+- Report category boundaries when continuous variables were categorized.
+
+## Other analyses {#sec-res-other}
+
+- Report other analyses done (e.g., analyses of subgroups and interactions, and sensitivity analyses).
 
@@
 
-{{< include sections/discussion.qmd >}}
+## Key results {#sec-disc-res-key}
 
-\newpage
+- Summarise key results with reference to study objectives.
+- Provide a review of the relevant literature to put the study findings into context.
+  - It should be complete and balanced, including inconsistent results.
+  - It should include, for each source, sufficient details: study design, sample size, population, specific exposures and outcomes.
+
+### What does the literature say?
+
+- EDCs and neurodevelopment (ANT).
+- EDCs and corticosteroids.
+- Corticosteroids and neurodevelopment (ANT).
+
+## Limitations {#sec-limitations}
+
+- Discuss limitations of the study, taking into account sources of potential bias or imprecision.
+- Discuss both direction and magnitude of any potential bias.
+
+Some limitations:
+
+- Cross-sectional study.
+- Chemicals measured in night and morning samples, whereas metabolites (the outcome) were measured only in night samples.
+- Cortisol measured at night, when should be lowest.
+- Change of estimand when trimming weights.
+- Model misspecification.
+- Mixtures effect.
+- Residual confounding.
+- Some confounders were not used since large percentage of missing values.
+- Multiple comparisons.
+
+## Interpretation {#sec-interpretation}
+
+- End with a summary of the key findings and their implications for the study hypotheses, future research, and policy.
+- Give a cautious overall interpretation of results considering objectives, limitations, multiplicity of analyses, results from similar studies, and other relevant evidence.
+
+## Generalisability {#sec-general}
+
+- Discuss the generalisability (external validity) of the study results.
+
+{{< pagebreak >}}
 
@@
 
-\newpage
+{{< pagebreak >}}
 
 # References {.unnumbered}
+
+::: {#refs}
+:::
+
+{{< pagebreak >}}
+
+{{< include _tables.qmd >}}
+
+{{< pagebreak >}}
+
+{{< include _figures.qmd >}}
+
+{{< pagebreak >}}
+
+{{< include _si.qmd >}}

### 2023-11-21 output/paper/sections_old/_introduction.qmd
Added presentation for group meeting regarding paper 3. Moved individual sections manuscript directly in manuscript. Added include of SI and tables and figures in manuscript for cross referencing.

--- 2023-11-14 output/paper/manuscript.qmd
@@
----
-title: "Estimating associations between childhood exposure to non-persistent endocrine disruptors, corticosteroids, and neurodevelopment: A study based on the parametric g-formula"
+The prevalence of several neurodevelopmental disorders has increased in the pediatric population [@GrandjeanLandrigan:2014], and multiple environmental pollutants play a role in the increased rates of these disorders [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
+\acr{edcs}, chemicals that are capable of interfering with the endocrine system, have been shown to have a role in childhood neurodevelopment [@Braun:2017].
+Although both pregnancy and early childhood are crucial stages of (neuro)development, most of the available literature is focused on the effects of prenatal exposure to \acr{edcs} on child neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
+Moreover, little is known about the biological mechanisms of action [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
 
-# metadata-files:
-#   - ../../../../all_phd/authors.yml
+\acr{opps} might have a deleterious effect on neurodevelopment, although the few studies conducted on children and using a biological matrix for exposure assessment suffered from a series of limitations, including a small sample size and few \acr{opp} metabolites measured [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
+Phthalates and their metabolites have been associated to several adverse effects on neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
+Similarly, these studies focused on few phthalates metabolites and were conducted in small populations.
+On the other hand, the effects of exposure to \acr{bpa} on cognitive functions are still unclear [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
 
-format:
-  docx:
-    reference-doc: reference_document.docx
-  pdf:
-    pdf-engine: lualatex
-    number-sections: true
-    toc: false
-    keep-tex: true
-    include-in-header:
-      text: |
-        \usepackage{lineno}
-        \usepackage{rotating}
-        \usepackage{float}
-        \usepackage{typearea}
-    include-before-body:
-      text: |
-        \linenumbers
+Another major limitation of these epidemiological studies, especially when non-persistent chemicals are analyzed, is the reliance on single-spot urine samples for exposure assessment, which might not be representative of long-term exposures.
 
-bibliography: bibliography.bib
-csl: ama11.csl
+Something about corticosteroids...
 
-link-citations: true
-linkcolor: blue
-
-filters:
-  - acronyms
-  # - authors-block
-  - ../../../../all_phd/_extensions/wordcount/citeproc.lua
-  - ../../../../all_phd/_extensions/wordcount/wordcount.lua
-
-acronyms:
-  fromfile: ../../../../all_phd/acronyms.yml
-  insert_links: false
-  insert_loa: false
-  sorting: alphabetical
-
-editor_options: 
-  chunk_output_type: console
-
-execute:
-  echo: false
-  warning: false
-  error: false
-  message: false
----
-
-# Abstract {.unnumbered}
-
-\newpage
-
-- Authorship must follow the ICMJE's criteria.
-Author names should be listed in ScholarOne and author contributions should be detailed in the cover letter ("Author A...").
-Neither names nor contributions should appear in the blinded manuscript.
-Do not include conflicts of interest.
-- The use of the word "effect(s)" as a proxy for "association(s)" is discouraged.
-- The length of the abstract must be less or equal than 200 words, and should be unstructured, stating the research questions, the methods used, and the results and conclusions of the research.
-- The length must be <4,000 words, excluding abstract, references, tables, figure captions, acknowledgments, and Supplementary Material.
-  - In-text citations: full-sized Arabic numerals in parentheses within the sentence.
-- Submission: SER members receive a 10% discount on the fees per page.
-This should be requested in the cover letter.
-Also 20% discount for Open Access charges.
-If uploading a single file (document, tables, figures, SM), designate the file as "Main document - anonymous".
-
-{{< include sections/introduction.qmd >}}
-
-# Methods {#sec-methods}
-
-{{< include sections/methods.qmd >}}
-
-# Results {#sec-res}
-
-{{< include sections/results.qmd >}}
-
-# Discussion {#sec-discussion}
-
-{{< include sections/discussion.qmd >}}
-
-\newpage
-
-# Funding
-
-\newpage
-
-# References {.unnumbered}
-
-::: {#refs}
-:::
+In the present study, we estimated associations between 1) non-persistent \acr{edcs} and attention, 2) non-persistent \acr{edcs} and corticosteroids, and 3) corticosteroids and attention, using the parametric g-formula and average contrasts, in children of a large prospective birth cohort in Europe.

### 2023-11-21 output/paper/sections_old/_methods.qmd
Added presentation for group meeting regarding paper 3. Moved individual sections manuscript directly in manuscript. Added include of SI and tables and figures in manuscript for cross referencing.

--- 2023-11-21 output/paper/manuscript.qmd
@@
----
-title: "Estimating associations between childhood exposure to non-persistent endocrine disruptors, corticosteroids, and neurodevelopment: A study based on the parametric g-formula"
-
-# metadata-files:
-#   - ../../../../all_phd/authors.yml
-
-format:
-  docx:
-    reference-doc: reference_document.docx
-    citeproc: false
-  # pdf:
-  #   pdf-engine: lualatex
-  #   number-sections: true
-  #   toc: false
-  #   keep-tex: false
-  #   include-in-header:
-  #     text: |
-  #       \usepackage{lineno}
-  #       \usepackage{rotating}
-  #       \usepackage{float}
-  #       \usepackage{typearea}
-  #   include-before-body:
-  #     text: |
-  #       \linenumbers
-
-crossref:
-  custom: 
-    - kind: float
-      ref-type: supptbl
-      latex-env: supptbl
-      prefix: Table S
-      name: Table S
-      space-before-numbering: false
-    - kind: float
-      ref-type: suppfig
-      latex-env: suppfig
-      prefix: Figure S
-      name: Figure S
-      space-before-numbering: false
-
-bibliography: bibliography.bib
-csl: ama11.csl
-
-link-citations: true
-linkcolor: blue
-
-filters:
-  - acronyms
-  # - authors-block
-  - ../../../../all_phd/_extensions/wordcount/citeproc.lua
-  - ../../../../all_phd/_extensions/wordcount/wordcount.lua
-
-acronyms:
-  fromfile: ../../../../all_phd/acronyms.yml
-  insert_links: false
-  insert_loa: false
-  sorting: alphabetical
-
-editor_options: 
-  chunk_output_type: console
-
-execute:
-  echo: false
-  warning: false
-  error: false
-  message: false
----
-
-```{r}
-here::i_am("output/paper/manuscript.qmd")
-```
-
-```{r vars-methods}
-source(here::here("code/dictionaries.R"))
-source(here::here("code/data.R"))
-source(here::here("code/proc_res.R"))
-
-is_hpc <- FALSE
-store <- ifelse(
-  is_hpc, 
-  paste0(
-    "/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/",
-    "lorenzoF/data/data_paper3/_targets/_targets_res"
-  ), 
-  "~/mounts/rstudioserver/data/data_paper3/_targets/_targets_res"
-)
-id_var <- "HelixID"
-num_digits_est = 3
-num_digits_sig = 2
-params_ <- params(is_hpc = is_hpc)
-threshold_within_lod <- params_$steps$rq2$preproc_outcome$llodq$threshold_within
-threshold_overall_lod <- params_$steps$rq2$preproc_outcome$llodq$threshold_overall
-threshold_within_nan <- params_$steps$rq2$preproc_outcome$missings$threshold_within
-threshold_overall_nan <- params_$steps$rq2$preproc_outcome$missings$threshold_overall
-knn_k <- params_$steps$rq2$preproc_outcome$missings$k
-```
-
-# Abstract {.unnumbered}
-
-{{< pagebreak >}}
-
-- Authorship must follow the ICMJE's criteria.
-Author names should be listed in ScholarOne and author contributions should be detailed in the cover letter ("Author A...").
-Neither names nor contributions should appear in the blinded manuscript.
-Do not include conflicts of interest.
-- The use of the word "effect(s)" as a proxy for "association(s)" is discouraged.
-- The length of the abstract must be less or equal than 200 words, and should be unstructured, stating the research questions, the methods used, and the results and conclusions of the research.
-- The length must be <4,000 words, excluding abstract, references, tables, figure captions, acknowledgments, and Supplementary Material.
-- In-text citations: full-sized Arabic numerals in parentheses within the sentence.
-- Submission: SER members receive a 10% discount on the fees per page.
-This should be requested in the cover letter.
-Also 20% discount for Open Access charges.
-If uploading a single file (document, tables, figures, SM), designate the file as "Main document - anonymous".
-
-The prevalence of several neurodevelopmental disorders has increased in the pediatric population [@GrandjeanLandrigan:2014], and multiple environmental pollutants play a role in the increased rates of these disorders [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
-\acr{edcs}, chemicals that are capable of interfering with the endocrine system, have been shown to have a role in childhood neurodevelopment [@Braun:2017].
-Although both pregnancy and early childhood are crucial stages of (neuro)development, most of the available literature is focused on the effects of prenatal exposure to \acr{edcs} on child neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
-Moreover, little is known about the biological mechanisms of action [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
-
-\acr{opps} might have a deleterious effect on neurodevelopment, although the few studies conducted on children and using a biological matrix for exposure assessment suffered from a series of limitations, including a small sample size and few \acr{opp} metabolites measured [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
-Phthalates and their metabolites have been associated to several adverse effects on neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
-Similarly, these studies focused on few phthalates metabolites and were conducted in small populations.
-On the other hand, the effects of exposure to \acr{bpa} on cognitive functions are still unclear [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
-
-Another major limitation of these epidemiological studies, especially when non-persistent chemicals are analyzed, is the reliance on single-spot urine samples for exposure assessment, which might not be representative of long-term exposures.
-
-Something about corticosteroids...
-
-In the present study, we estimated associations between 1) non-persistent \acr{edcs} and attention, 2) non-persistent \acr{edcs} and corticosteroids, and 3) corticosteroids and attention, using the parametric g-formula and average contrasts, in children of a large prospective birth cohort in Europe.
-
-# Methods {#sec-methods}
-
 ## Study population and design {#sec-design}
@@
 We further tested for significance of the difference between the average contrasts of males and females.
-
-# Results {#sec-res}
-
-## Participants {#sec-res-participants}
-
-## Descriptive data {#sec-res-pop-desc}
-
-## Outcome data {#sec-res-out-desc}
-
-## Main results {#sec-res-main}
-
-- All results on which study conclusions or inferences are based, including null findings and results of secondary or sensitivity analyses, must be reported.
-Use of sub-headings that describe the nature of the results (but no declarative statements).
-  - Provide a clear and concise description of all findings without extrapolating beyond the study results.
-  - Do not limit results to those *statistically significant* or that support the study hypotheses.
-Avoid using statistical significance testing as the sole or primary criterion for interpreting the obtained results.
-If significance testing or *p*-values are used, report numeric *p*-values, rounded to 1-2 digits, for all results.
-Use an uppercase italic letter "P", and the values should not be bolded.
-Indicate whether are 1- or 2-sided.
-- Give unadjusted and confounder-adjusted estimates and their precision (e.g., $95\%$ confidence interval).
-Make clear which confounders were adjusted for and why they were included.
-Include the number of observations for each analysis after accounting for missing data.
-Include numeric data within figures (e.g., forest plots), or provide tables with corresponding numeric data for all figures.
-  - [`marginaleffects` tables](https://vincentarelbundock.github.io/marginaleffects/articles/tables.html).
-- Report category boundaries when continuous variables were categorized.
-
-## Other analyses {#sec-res-other}
-
-- Report other analyses done (e.g., analyses of subgroups and interactions, and sensitivity analyses).
-
-# Discussion {#sec-discussion}
-
-## Key results {#sec-disc-res-key}
-
-- Summarise key results with reference to study objectives.
-- Provide a review of the relevant literature to put the study findings into context.
-  - It should be complete and balanced, including inconsistent results.
-  - It should include, for each source, sufficient details: study design, sample size, population, specific exposures and outcomes.
-
-### What does the literature say?
-
-- EDCs and neurodevelopment (ANT).
-- EDCs and corticosteroids.
-- Corticosteroids and neurodevelopment (ANT).
-
-## Limitations {#sec-limitations}
-
-- Discuss limitations of the study, taking into account sources of potential bias or imprecision.
-- Discuss both direction and magnitude of any potential bias.
-
-Some limitations:
-
-- Cross-sectional study.
-- Chemicals measured in night and morning samples, whereas metabolites (the outcome) were measured only in night samples.
-- Cortisol measured at night, when should be lowest.
-- Change of estimand when trimming weights.
-- Model misspecification.
-- Mixtures effect.
-- Residual confounding.
-- Some confounders were not used since large percentage of missing values.
-- Multiple comparisons.
-
-## Interpretation {#sec-interpretation}
-
-- End with a summary of the key findings and their implications for the study hypotheses, future research, and policy.
-- Give a cautious overall interpretation of results considering objectives, limitations, multiplicity of analyses, results from similar studies, and other relevant evidence.
-
-## Generalisability {#sec-general}
-
-- Discuss the generalisability (external validity) of the study results.
-
-{{< pagebreak >}}
-
-# Funding
-
-{{< pagebreak >}}
-
-# References {.unnumbered}
-
-::: {#refs}
-:::
-
-{{< pagebreak >}}
-
-{{< include _tables.qmd >}}
-
-{{< pagebreak >}}
-
-{{< include _figures.qmd >}}
-
-{{< pagebreak >}}
-
-{{< include _si.qmd >}}

### 2023-11-21 output/presentations/cep_internal_231122/presentation.qmd
Added presentation for group meeting regarding paper 3. Moved individual sections manuscript directly in manuscript. Added include of SI and tables and figures in manuscript for cross referencing.

---
author: Lorenzo Fabbri
title: "Estimating associations between childhood exposure to non-persistent endocrine disruptors, corticosteroids, and neurodevelopment: A study based on the parametric g-formula"

format:
  revealjs:
    embed-resources: true
    slide-number: true
    width: 1600
    height: 900
    theme: simple
    center: true

editor_options: 
  chunk_output_type: console
execute:
  echo: false
  warning: false
  error: false
  message: false
---

## Recap

* We want to estimate the following *effects* (in children from HELIX):
  * Non-persistent EDCs and neurodevelopment (using attention as *proxy*).
  * Non-persistent EDCs and corticosteroids.
  * Corticosteroids and neurodevelopment.
* For attention we used the HIT-RT SE from the Attention Network Test.
* We did not consider individual metabolites but sums and ratios.
* We want to do so using the parametric g-formula and average contrasts.

## The pipeline

1. Estimation of balancing weights: `exposure ~ covariates`.
1. G-computation:
    1. Fit a **weighted outcome model**: `outcome ~ exposure + covariates`.
    1. Create two copies of the original dataset, with the exposure **shifted**.
    1. Predict the outcomes in the two **counterfactual** datasets.
    1. Estimate the **estimand** $\phi^\Delta = \mathbb{E} [Y^{d_1}] - \mathbb{E} [Y^{d_2}]$.

In this case:

* Cohort-specific shifting of the exposure. $d_1$ is a deterministic function setting the exposure to its 90th percentile, $d_2$ to its 10th percentile.

We also looked into **effect-modification** by sex (weights estimated separately, interaction in outcome model, estimation of marginal effects taking this into account).

## Difficulties {.smaller}

It looks simple, but lots of literature searches:

* DAGs for three different research questions.
* New data (corticosteroids), and still awaiting answers, and ratios difficult to interpret.
* Estimation of balancing weights for **ratios** is (apparently) still an open question.
* Usage of more **flexible modeling** approaches to estimate the ADRF (natural cubic splines).
* Careful modeling of **ratios**; it's not that simple:
    * `log (num / den) ~ exposure + log den + covariates`.
    * `outcome ~ log (num / den)`.
* Creatinine coming from different sources.
* Proper modeling of creatinine to take into account dilution effects.
* Careful consideration of positivity violations when shifting the exposures (cohort-specific and smaller percentiles).
* Computation of **robust** standard errors for inference using `cohort` as clusters.
* Careful usage of tools to improve **replicability** across the entire project (from data loading to paper writing).

## Results

See files...

### 2023-11-21 _targets_res.R

@@ -7,8 +7,8 @@ targets::tar_option_set(
   format = "qs"
 )
 path_store <- Sys.getenv("path_store")
-num_digits_est = 3
-num_digits_sig = 2
+num_digits_est <- 3
+num_digits_sig <- 2
 sa_var_ <- "e3_sex"
 
 list(

### 2023-11-21 _targets_rqX.R

@@ -8,18 +8,18 @@ targets::tar_option_set(
 )
 
 params_dag <- list(
-  type = "minimal", 
+  type = "minimal",
   effect = "total"
 )
 rq <- Sys.getenv("TAR_PROJECT")
-exposure <- switch(rq, 
-                   "rq1" = "chemical", 
-                   "rq2" = "chemical", 
+exposure <- switch(rq,
+                   "rq1" = "chemical",
+                   "rq2" = "chemical",
                    "rq3" = "biomarker",
                    "rq4" = "biomarker")
-outcome <- switch(rq, 
-                  "rq1" = "outcome", 
-                  "rq2" = "biomarker", 
+outcome <- switch(rq,
+                  "rq1" = "outcome",
+                  "rq2" = "biomarker",
                   "rq3" = "outcome",
                   "rq4" = "chemical")
 
@@ -32,7 +32,7 @@ if (rq == "rq2") {
 } else {
   tbl_outcomes <- tibble::tibble(
     name = params(is_hpc = Sys.getenv("is_hpc"))$variables[[rq]]$outcome |>
-      stringr::str_to_lower(), 
+      stringr::str_to_lower(),
     outcome = params(is_hpc = Sys.getenv("is_hpc"))$variables[[rq]]$outcome
   )
 }

### 2023-11-21 output/paper/_tables.qmd
Started writing results.

@@
 #| label: tbl-pop-desc
-#| tbl-cap: Population description
+#| tbl-cap: Characteristics of the children included in the analyses, by cohort and overall
 
@@
 #| label: tbl-edc-desc
-#| tbl-cap: Levels of unprocessed chemicals
+#| tbl-cap: Levels of unprocessed chemicals, by cohort and overall
 
@@
 #| label: tbl-met-desc
-#| tbl-cap: Levels of unprocessed corticosteroids
+#| tbl-cap: Levels of unprocessed corticosteroids, by cohort and overall
 

### 2023-11-21 output/paper/manuscript.qmd
Started writing results.

@@
 
+In total, 1,297 children were included in the dataset.
+Of these, 976 had information also about the corticosteroids.
+The sample consisted of 55% girls.
+The median \acr{hitrtse} was 300 ms (\acr{iqr}, 231-368), with lower median values for EDEN, MOBA, and SAB, corresponding to the cohorts with older children.
+At the time of visit, the median age of the children was 8.06 years.
+The children were mostly Caucasian (90%) or Pakistani (6.2%).
+@tbl-pop-desc provides descriptive statistics for the outcome and covariates, by cohort and overall.
+Codebooks for the used covariates, by research question, are provided in @supptbl-codebook-1, @supptbl-codebook-2, and @supptbl-codebook-3.
+
 ## Descriptive data {#sec-res-pop-desc}
 
-## Outcome data {#sec-res-out-desc}
+Levels of unprocessed non-persistent \acr{edcs} and corticosteroids are presented in @tbl-edc-desc and @tbl-met-desc, respectively.
 

### 2023-11-22 output/paper/_si.qmd
Looked into ivreg for metabolites to chemicals.

@@
 
-```{r codebook-1}
-#| label: supptbl-codebook-1
-#| tbl-cap: Codebook for covariates used in RQ1
-
+```{r}
 targets::tar_load(
@@
 )
+```
+
+::: {#supptbl-codebook-1}
+```{r codebook-1}
 codebook_1
 ```
+Codebook for covariates used in RQ1
+:::
 
+::: {#supptbl-codebook-2}
 ```{r codebook-2}
-#| label: supptbl-codebook-2
-#| tbl-cap: Codebook for covariates used in RQ2
-
 codebook_2
 ```
+Codebook for covariates used in RQ2
+:::
 
+::: {#supptbl-codebook-3}
 ```{r codebook-3}
-#| label: supptbl-codebook-3
-#| tbl-cap: Codebook for covariates used in RQ3
-
 codebook_3
 ```
+Codebook for covariates used in RQ3
+:::
 
@@
 
+::: {#supptbl-info-chems}
 ```{r info-chems}
-#| label: supptbl-info-chems
-#| tbl-cap: Information about non-persistent EDCs
-
 edcs_info <- myphd::edcs_information() |>
@@
 ```
+Information about non-persistent EDCs
+:::
 
@@
 
+::: {#supptbl-info-mets}
 ```{r info-mets}
-#| label: supptbl-info-mets
-#| tbl-cap: Information about metabolites
-
 mets_info <- readODS::read_ods(
@@
 ```
+Information about metabolites
+:::
 
+::: {#supptbl-lloq-mets}
 ```{r lloq-mets}
-#| label: supptbl-lloq-mets
-#| tbl-cap: Lower limits of quantification for the corticosteroids
-
 lloq_mets <- load_steroids()$loq |>
@@
 ```
+Lower limits of quantification for the corticosteroids
+:::
 
@@
 
+::: {#supptbl-weights-1sa}
 ```{r weights-1sa}
-#| label: supptbl-weights-1sa
-#| tbl-cap: Weights
-
 tidy_res_weighted_fits(
@@
 ```
+Weights
+:::
 
+::: {#supptbl-weights-2sa}
 ```{r weights-2sa}
-#| label: supptbl-weights-2sa
-#| tbl-cap: Weights
-
 tidy_res_weighted_fits(
@@
 ```
+Weights
+:::
 
+::: {#supptbl-weights-3sa}
 ```{r weights-3sa}
-#| label: supptbl-weights-3sa
-#| tbl-cap: Weights
-
 tidy_res_weighted_fits(
@@
 ```
+Weights
+:::
 
@@
 
+::: {#suppfig-desc-chems}
 ```{r desc-chems-lod}
-#| label: suppfig-desc-chems
-#| fig-cap: Description of chemicals
 #| fig-width: 6
@@
 ```
+Description of chemicals
+:::
 
@@
 
+::: {#suppfig-desc-mets}
 ```{r desc-mets-lod}
-#| label: suppfig-desc-mets
-#| fig-cap: Description of metabolites
 #| fig-width: 6
@@
 ```
+Description of metabolites
+:::
 
@@
 
+::: {#suppfig-adrfs-rq3}
 ```{r adrfs-rq3}
-#| label: suppfig-adrfs-rq3
-#| fig-cap: Average dose-response functions for RQ3
 #| fig-width: 6
@@
 ```
+Average dose-response functions for RQ3
+:::
 
@@
 
+::: {#suppfig-marginal-1sa}
 ```{r marginal-1sa}
-#| label: suppfig-marginal-1sa
-#| fig-cap: Average comparisons for effect modification
 #| fig-width: 6
@@
 ```
+Average comparisons for effect modification
+:::
 
+::: {#suppfig-marginal-2sa}
 ```{r marginal-2sa}
-#| label: suppfig-marginal-2sa
-#| fig-cap: Average comparisons for effect modification
 #| fig-width: 9
@@
 ```
+Average comparisons for effect modification
+:::
 
+::: {#suppfig-marginal-3sa}
 ```{r marginal-3sa}
-#| label: suppfig-marginal-3sa
-#| fig-cap: Average comparisons for effect modification
 #| fig-width: 5
@@
 ```
+Average comparisons for effect modification
+:::

### 2023-11-22 output/paper/manuscript.qmd
Looked into ivreg for metabolites to chemicals.

@@
     reference-doc: reference_document.docx
-    citeproc: false
+    citeproc: true
   # pdf:
@@
   # - authors-block
-  - ../../../../all_phd/_extensions/wordcount/citeproc.lua
+  # - ../../../../all_phd/_extensions/wordcount/citeproc.lua
   - ../../../../all_phd/_extensions/wordcount/wordcount.lua
@@
 A list of the environmental chemicals determined in urine samples and used for the present study is given in @supptbl-info-chems.
+Briefly, we analyzed a total of 7 phenols, 4 non-specific organophosphate pesticide metabolites, and 10 phthalate metabolites originating from 6 distinct phthalate parent compounds.
 The laboratory protocols for the analysis are described elsewhere [@HaugSakhiCequier:2018].
@@
 
-Levels of unprocessed non-persistent \acr{edcs} and corticosteroids are presented in @tbl-edc-desc and @tbl-met-desc, respectively.
+Levels of unprocessed non-persistent \acr{edcs} and corticosteroids, by cohort and overall, are presented in @tbl-edc-desc and @tbl-met-desc, respectively.
 

### 2023-11-22 output/presentations/cep_internal_231122/presentation.qmd
Looked into ivreg for metabolites to chemicals.

@@
 
+## Even more difficulties
+
+Most likely there is reverse causality (assumption of time-ordering from DAGs is not satisfied).
+
+:::: {.columns}
+::: {.column}
+```{r}
+coords <- tibble::tribble(
+  ~name, ~x, ~y,
+  "edc", 0, 0,
+  "cortisol", 1, 0,
+  "U", 0.5, 0.5
+)
+dag <- ggdag::dagify(
+  cortisol ~ edc + U,
+  edc ~ U,
+  exposure = "edc",
+  outcome = "cortisol",
+  labels = c(
+    "edc" = "EDC",
+    "cortisol" = "cortisol",
+    "U" = "confounders"
+  ),
+  coords = coords
+)
+
+ggdag::ggdag(
+  dag,
+  text = FALSE,
+  use_labels = "label"
+) +
+  ggdag::theme_dag()
+```
+:::
+
+::: {.column}
+```{r}
+coords <- tibble::tribble(
+  ~name, ~x, ~y,
+  "edc", 1, 0,
+  "cortisol", 0, 0,
+  "U", 0.5, 0.5,
+  "iv", -0.5, 0
+)
+dag <- ggdag::dagify(
+  cortisol ~ iv + U,
+  edc ~ cortisol + U,
+  exposure = "cortisol",
+  outcome = "edc",
+  labels = c(
+    "edc" = "EDC",
+    "cortisol" = "cortisol",
+    "U" = "confounders",
+    "iv" = "SNP"
+  ),
+  coords = coords
+)
+
+ggdag::ggdag(
+  dag,
+  text = FALSE,
+  use_labels = "label"
+) +
+  ggdag::theme_dag()
+```
+:::
+::::
+
+* We cannot use 2-sample MR since there are no SNPs for the *outcome*.
+* We do not have single metabolites but sums and ratios.
+* I am not sure it makes sense...
+
 ## Results

### 2023-12-11 output/paper/_si.qmd
Transformed into R package cause I wanted to use roxygen2 to document functions but it does not seem to be working. Renamed directory code to R for same reason. Removed virtually all default arguments functions. Testing pipelines.

@@
   dat = myphd::extract_cohort(load_steroids()$desc,
-                              id_var = id_var) |>
+                              id_var = id_var,
+                              st = 1, en = 3) |>
     tidylog::mutate(
@@
     ),
-  vars = vars_of_interest()$metabolites |>
+  vars = vars_of_interest(append_to_chem = NULL)$metabolites |>
     paste0("_cdesc"),

### 2023-12-11 R/data.R

@@ -3,7 +3,7 @@
 #' @param dat Dataset containing original variables. A dataframe.
 #'
 #' @return A dataset with additional variables. A dataframe.
-#' 
+#'
 #' @export
 create_steroid_scores <- function(dat) {
   warning(
@@ -14,7 +14,7 @@ create_steroid_scores <- function(dat) {
   dat_ret <- dat |>
     tidylog::mutate(
       # Cortisol
-      cortisol_production = rowSums(dplyr::across(c(`F`, X20aDHF, X20bDHF, X5bDHF, X5aTHF, X5bTHF, X6OHF, X5a20acortol, X5a20bcortol, X5b20acortol, X5b20bcortol))), 
+      cortisol_production = rowSums(dplyr::across(c(`F`, X20aDHF, X20bDHF, X5bDHF, X5aTHF, X5bTHF, X6OHF, X5a20acortol, X5a20bcortol, X5b20acortol, X5b20bcortol))),
       cortisol_metabolism = rowSums(dplyr::across(c(X20aDHF, X20bDHF, X5bDHF, X5aTHF, X5bTHF, X6OHF, X5a20acortol, X5a20bcortol, X5b20acortol, X5b20bcortol))) / `F`,
       # Cortisone
       cortisone_production = rowSums(dplyr::across(c(CortisoneE, X20aDHE, X20bDHE, X5aTHE, X5bTHE, X6OHE, X5b20acortolone, X5b20bcortolone))),
@@ -22,22 +22,23 @@ create_steroid_scores <- function(dat) {
       # 11beta
       X11bHSD = cortisone_production / cortisol_production,
       # Global reductases
-      #global_reductase_f = rowSums(dplyr::across(c(X5aTHF, X5bTHF))) / `F`,
-      #global_reductase_e = X5bTHE / CortisoneE,
+      # global_reductase_f = rowSums(dplyr::across(c(X5aTHF, X5bTHF))) / `F`,
+      # global_reductase_e = X5bTHE / CortisoneE,
       # Others
       cyp3a4 = X6OHF / `F`,
       corticosterone_production = rowSums(dplyr::across(c(X5aTHB, X5bTHB, A, X17DOcortolone))),
-      #X11deoxycortisol_production = rowSums(dplyr::across(c(S, X5bDHS, X5bTHS))),
-      #X11hydroxylase = rowSums(dplyr::across(c(X5aTHF, X5bTHF, X5bTHE))) / X5bTHS,
[…]

### 2023-12-11 R/dictionaries.R

@@ -3,7 +3,7 @@
 ################################################################################
 # Parameters for entire paper
 ################################################################################
-vars_of_interest <- function(append_to_chem = NULL) {
+vars_of_interest <- function(append_to_chem) {
   chemicals <- c(
     "hs_mep_",
     "hs_mibp_",
@@ -71,21 +71,21 @@ vars_of_interest <- function(append_to_chem = NULL) {
     "cortisone_production",
     "cortisone_metabolism",
     "X11bHSD",
-    #"global_reductase_f",
-    #"global_reductase_e",
-    #"cyp3a4",
+    # "global_reductase_f",
+    # "global_reductase_e",
+    # "cyp3a4",
     "corticosterone_production"
-    #"X11deoxycortisol_production",
-    #"X11hydroxylase",
-    #"X17hydroxylase",
-    #"androgens_production",
-    #"X5a_reductase",
-    #"lyase",
-    #"global_adrenal_function"
+    # "X11deoxycortisol_production",
+    # "X11hydroxylase",
+    # "X17hydroxylase",
+    # "androgens_production",
+    # "X5a_reductase",
+    # "lyase",
+    # "global_adrenal_function"
   )
   outcomes <- c(
     "hs_hitrtse"
-    #"hs_dcolors3"
+    # "hs_dcolors3"
   )
   
   return(list(
@@ -95,6 +95,7 @@ vars_of_interest <- function(append_to_chem = NULL) {
     outcomes = outcomes
   ))
 } # End function vars_of_interest
+################################################################################
 
 ################################################################################
 # Parameters for entire paper
@@ -118,8 +119,8 @@ params <- function(is_hpc) {
   ##############################################################################
   
   chemicals <- vars_of_interest(append_to_chem = "c")$chemicals
-  metabolites <- vars_of_interest()$metabolites
-  clinical_outcomes <- vars_of_interest()$outcomes
+  metabolites <- vars_of_interest(append_to_chem = NULL)$metabolites
+  clinical_outcomes <- vars_of_interest(append_to_chem = NULL)$outcomes
   creatinine_covariates_names <- list(
     numerical = c(
       "hs_age_years", "hs_c_weight", "hs_c_height",
@@ -391,6 +392,7 @@ params <- function(is_hpc) {
     steps = steps
   ))
 } # End function params
+################################################################################
 
 ################################################################################
 # Parameters for running analyses
@@ -442,3 +444,4 @@ params_analyses <- function() {
   
   return(ret)
 } # End function params_analyses
+################################################################################

### 2023-12-11 _targets_rq0X.R

@@ -1,8 +1,8 @@
 source("DAGs/dag_v2.R")
-source("code/dictionaries.R")
-source("code/data.R")
-source("code/utils.R")
-source("code/proc_res.R")
+source("R/dictionaries.R")
+source("R/data.R")
+source("R/utils.R")
+source("R/proc_res.R")
 
 targets::tar_option_set(
   format = "qs"
@@ -62,8 +62,10 @@ list(
     name = paste0(rq, "_viz_desc_metabs"),
     command = expression(
       viz_desc_vars(
-        dat = myphd::extract_cohort(load_steroids()$desc,
-                                    id_var = id_var
+        dat = myphd::extract_cohort(
+          load_steroids()$desc,
+          id_var = id_var,
+          st = 1, en = 3
         ) |>
           tidylog::mutate(
             cohort = dplyr::case_when(
@@ -74,7 +76,7 @@ list(
               .default = cohort
             )
           ),
-        vars = vars_of_interest()$metabolites |>
+        vars = vars_of_interest(append_to_chem = NULL)$metabolites |>
           paste0("_cdesc"),
         fct_levels = c(1, 2, 3, 4),
         is_chem = FALSE
@@ -86,8 +88,10 @@ list(
     name = paste0(rq, "_viz_miss_covars"),
     command = substitute(
       myphd::explore_missings(
-        myphd::extract_cohort(dat$covariates,
-                              id_var = id_var
+        myphd::extract_cohort(
+          dat$covariates,
+          id_var = id_var,
+          st = 1, en = 3
         ),
         id_var = id_var,
         by_var = by_var,
@@ -106,8 +110,10 @@ list(
     name = paste0(rq, "_viz_miss_exp"),
     command = substitute(
       myphd::explore_missings(
-        myphd::extract_cohort(dat$exposures,
-                              id_var = id_var
+        myphd::extract_cohort(
+          dat$exposures,
+          id_var = id_var,
+          st = 1, en = 3
         ),
         id_var = id_var,
         by_var = by_var,
@@ -126,8 +132,10 @@ list(
     name = paste0(rq, "_viz_miss_out"),
     command = substitute(
       myphd::explore_missings(
-        myphd::extract_cohort(dat$outcome,
-                              id_var = id_var
+        myphd::extract_cohort(
+          dat$outcome,
+          id_var = id_var,
+          st = 1, en = 3
         ),
         id_var = id_var,
         by_var = by_var,
@@ -147,8 +155,10 @@ list(
[…]

### 2023-12-11 _targets_rqX.R

@@ -1,7 +1,7 @@
 source("DAGs/dag_v2.R")
-source("code/dictionaries.R")
-source("code/data.R")
-source("code/utils.R")
+source("R/dictionaries.R")
+source("R/data.R")
+source("R/utils.R")
 
 targets::tar_option_set(
   format = "qs"
@@ -25,9 +25,9 @@ outcome <- switch(rq,
 
 if (rq == "rq2") {
   tbl_outcomes <- tibble::tibble(
-    name = vars_of_interest()$new_metabolites |>
+    name = vars_of_interest(append_to_chem = NULL)$new_metabolites |>
       stringr::str_to_lower(),
-    outcome = vars_of_interest()$new_metabolites
+    outcome = vars_of_interest(append_to_chem = NULL)$new_metabolites
   )
 } else {
   tbl_outcomes <- tibble::tibble(
@@ -62,7 +62,12 @@ list(
   targets::tar_target_raw(
     name = paste0(rq, "_preproc_dat"),
     command = substitute(
-      rq_prepare_data(dat = dat, filter_panel = FALSE, type_sample_hcp = NULL),
+      rq_prepare_data(
+        dat = dat,
+        filter_panel = FALSE,
+        type_sample_hcp = NULL,
+        is_sa = FALSE
+      ),
       env = list(dat = as.symbol(paste0(rq, "_load_dat")))
     )
   ),
@@ -73,8 +78,9 @@ list(
     command = substitute(
       rq_estimate_weights(
         dat = dat,
+        by = NULL,
         include_selection_weights = FALSE,
-        save_results = TRUE,
+        save_results = FALSE,
         parallel = FALSE,
         workers = 10
       ),
@@ -91,6 +97,8 @@ list(
         rq_fit_model_weighted(
           dat = dat,
           outcome = outcome,
+          by = NULL,
+          is_panel = FALSE,
           weights = weights$estimated_weights,
           parallel = FALSE,
           workers = 10
@@ -107,6 +115,8 @@ list(
       command = substitute(
         rq_estimate_marginal_effects(
           fits = all_fits$fits,
+          by = NULL,
+          is_hcp = FALSE,
           parallel = TRUE,
           workers = 3
         ),

### 2023-12-11 _targets_sa_rqX.R

@@ -1,7 +1,7 @@
 source("DAGs/dag_v2.R")
-source("code/dictionaries.R")
-source("code/data.R")
-source("code/utils.R")
+source("R/dictionaries.R")
+source("R/data.R")
+source("R/utils.R")
 
 targets::tar_option_set(
   format = "qs"
@@ -25,9 +25,9 @@ outcome <- switch(rq,
 
 if (rq == "rq2") {
   tbl_outcomes <- tibble::tibble(
-    name = vars_of_interest()$new_metabolites |>
+    name = vars_of_interest(append_to_chem = NULL)$new_metabolites |>
       stringr::str_to_lower(),
-    outcome = vars_of_interest()$new_metabolites
+    outcome = vars_of_interest(append_to_chem = NULL)$new_metabolites
   )
 } else {
   tbl_outcomes <- tibble::tibble(
@@ -77,9 +77,9 @@ list(
     command = substitute(
       rq_estimate_weights(
         dat = dat,
-        include_selection_weights = FALSE,
         by = by,
-        save_results = TRUE,
+        include_selection_weights = FALSE,
+        save_results = FALSE,
         parallel = FALSE,
         workers = 10
       ),
@@ -97,6 +97,7 @@ list(
           dat = dat,
           outcome = outcome,
           by = c(by),
+          is_panel = FALSE,
           weights = weights$estimated_weights,
           parallel = FALSE,
           workers = 10
@@ -114,6 +115,7 @@ list(
         rq_estimate_marginal_effects(
           fits = all_fits$fits,
           by = by,
+          is_hcp = FALSE,
           parallel = TRUE,
           workers = 3
         ),

### 2023-12-11 archive/_targets_rqX.R

@@ -1,7 +1,7 @@
 source("DAGs/dag_v2.R")
-source("code/dictionaries.R")
-source("code/data.R")
-source("code/utils.R")
+source("R/dictionaries.R")
+source("R/data.R")
+source("R/utils.R")
 
 targets::tar_option_set(
   format = "qs"
@@ -69,7 +69,7 @@ list(
   targets::tar_target(
     name = weights, 
     command = rq_estimate_weights(dat = preproc_dat, 
-                                  save_results = TRUE, 
+                                  save_results = FALSE, 
                                   parallel = FALSE, 
                                   workers = 10)
   ), # End weights target

### 2023-12-14 output/paper/_figures.qmd
Moved some tables to SI. Added better captions for tables and figures.

@@
 
-# Main results
+# Figures for main results
 
@@
 #| label: fig-marginal-1
-#| fig-cap: Average comparisons
+#| fig-cap: Marginal comparison results for $hitrtse \sim edc$
 #| fig-width: 6
@@
 #| label: fig-marginal-2
-#| fig-cap: Average comparisons
+#| fig-cap: Marginal comparison results for $metabolite \sim edc$
 #| fig-width: 9
@@
 #| label: fig-marginal-3
-#| fig-cap: Average comparisons
+#| fig-cap: Marginal comparison results for $hitrtse \sim metabolite$
 #| fig-width: 5

### 2023-12-14 output/paper/_si.qmd
Moved some tables to SI. Added better captions for tables and figures.

--- 2023-11-22 output/paper/_si.qmd
@@
 
-## Descriptive data
+## Tables for descriptive data
 
@@
 ```
-Codebook for covariates used in RQ1
+Codebook for covariates used in $hitrtse \sim edc$
 :::
@@
 ```
-Codebook for covariates used in RQ2
+Codebook for covariates used in $metabolite \sim edc$
 :::
@@
 ```
-Codebook for covariates used in RQ3
+Codebook for covariates used in $hitrtse \sim metabolite$
 :::
@@
 
-## Other results
+## Tables for main results
+
+### Balancing weights
+
+```{r}
+targets::tar_load(
+  dplyr::matches("balance_(\\d)$"),
+  store = store
+)
+```
+
+::: {#supptbl-balance-1}
+```{r balance-1}
+balance_1 |>
+  tidylog::group_by(class) |>
+  dplyr::arrange(dplyr::desc(Adjusted)) |>
+  tidylog::ungroup() |>
+  gt::gt(groupname_col = "class") |>
+  gt::fmt_number(n_sigfig = 4) |>
+  gt::tab_footnote(
+    footnote = "N.",
+    locations = gt::cells_column_labels(
+      columns = c(Unadjusted, Adjusted)
+    )
+  ) |>
+  gt::tab_footnote(
+    footnote = "Truncated weights.",
+    locations = gt::cells_column_labels(columns = Adjusted)
+  )
+```
+Effective sample size before and after weights estimation for $hitrtse \sim edc$
+:::
+
+::: {#supptbl-balance-2}
+```{r balance-2}
+balance_2 |>
+  tidylog::group_by(class) |>
+  dplyr::arrange(dplyr::desc(Adjusted)) |>
+  tidylog::ungroup() |>
+  gt::gt(groupname_col = "class") |>
+  gt::fmt_number(n_sigfig = 4) |>
+  gt::tab_footnote(
+    footnote = "N.",
+    locations = gt::cells_column_labels(
+      columns = c(Unadjusted, Adjusted)
+    )
+  ) |>
+  gt::tab_footnote(
+    footnote = "Truncated weights.",
+    locations = gt::cells_column_labels(columns = Adjusted)
+  )
+```
+Effective sample size before and after weights estimation for $metabolite \sim edc$
+:::
+
+::: {#supptbl-balance-3}
+```{r balance-3}
+balance_3 |>
+  dplyr::arrange(dplyr::desc(Adjusted)) |>
+  gt::gt() |>
+  gt::fmt_number(n_sigfig = 4) |>
+  gt::tab_footnote(
+    footnote = "N.",
+    locations = gt::cells_column_labels(
+      columns = c(Unadjusted, Adjusted)
+    )
+  ) |>
+  gt::tab_footnote(
+    footnote = "Truncated weights.",
+    locations = gt::cells_column_labels(columns = Adjusted)
+  )
+```
+Effective sample size before and after weights estimation for $hitrtse \sim metabolite$
+:::
+
+```{r}
+targets::tar_load(
+  dplyr::matches("weights_(\\d)$"),
+  store = store
+)
+```
+
+::: {#supptbl-weights-1}
+```{r weights-1}
+tidy_res_weighted_fits(
+  dat_tbl = weights_1$dat_tbl,
+  dat_plt = weights_1$dat_plt,
+  sa_var = NULL
+)$table
+```
+Summary statistics of the estimated balancing weights for $hitrtse \sim edc$
+:::
+
+::: {#supptbl-weights-2}
+```{r weights-2}
+tidy_res_weighted_fits(
+  dat_tbl = weights_2$dat_tbl,
+  dat_plt = weights_2$dat_plt,
+  sa_var = NULL
+)$table
+```
+Summary statistics of the estimated balancing weights for $metabolite \sim edc$
+:::
+
+::: {#supptbl-weights-3}
+```{r weights-3}
+tidy_res_weighted_fits(
+  dat_tbl = weights_3$dat_tbl,
+  dat_plt = weights_3$dat_plt,
+  sa_var = NULL
+)$table
+```
+Summary statistics of the estimated balancing weights for $hitrtse \sim metabolite$
+:::
+
+## Tables for other results
 
@@
 ```
-Weights
+Summary statistics of the estimated balancing weights for $hitrtse \sim edc \times sex$
 :::
@@
 ```
-Weights
+Summary statistics of the estimated balancing weights for $metabolite \sim edc \times sex$
 :::
@@
 ```
-Weights
+Summary statistics of the estimated balancing weights for $hitrtse \sim metabolite \times sex$
 :::
@@
 
-## Descriptive data
+## Figures for descriptive data
 
@@
 ```
-Description of chemicals
+Classification of the measured non-persistent EDCs, by cohort.
 :::
@@
 ```
-Description of metabolites
+Classification of the measured corticosteroids, by cohort.
 :::
 
-## Main results
+## Figures for main results
 
@@
 ```
-Average dose-response functions for RQ3
+Average dose-response functions for...
 :::
 
-## Other results
+## Figures for other results
 
@@
 ```
-Average comparisons for effect modification
+Marginal comparison results for $hitrtse \sim edc \times sex$
 :::
@@
 ```
-Average comparisons for effect modification
+Marginal comparison results for $metabolite \sim edc \times sex$
 :::
@@
 ```
-Average comparisons for effect modification
+Marginal comparison results for $hitrtse \sim metabolite \times sex$
 :::

### 2023-12-14 output/paper/_tables.qmd
Nothing.
Moved some tables to SI. Added better captions for tables and figures.

@@
 
-# Descriptive data
+# Tables for descriptive data
 
@@
 
-# Main results
-
-## Balancing weights
-
-```{r}
-targets::tar_load(
-  dplyr::matches("balance_(\\d)$"),
-  store = store
-)
-```
-
-```{r balance-1}
-#| label: tbl-balance-1
-#| tbl-cap: Effective number of observations
-
-balance_1 |>
-  tidylog::group_by(class) |>
-  dplyr::arrange(dplyr::desc(Adjusted)) |>
-  tidylog::ungroup() |>
-  gt::gt(groupname_col = "class") |>
-  gt::fmt_number(n_sigfig = 4) |>
-  gt::tab_footnote(
-    footnote = "N.",
-    locations = gt::cells_column_labels(
-      columns = c(Unadjusted, Adjusted)
-    )
-  ) |>
-  gt::tab_footnote(
-    footnote = "Truncated weights.",
-    locations = gt::cells_column_labels(columns = Adjusted)
-  )
-```
-
-```{r balance-2}
-#| label: tbl-balance-2
-#| tbl-cap: Effective number of observations
-
-balance_2 |>
-  tidylog::group_by(class) |>
-  dplyr::arrange(dplyr::desc(Adjusted)) |>
-  tidylog::ungroup() |>
-  gt::gt(groupname_col = "class") |>
-  gt::fmt_number(n_sigfig = 4) |>
-  gt::tab_footnote(
-    footnote = "N.",
-    locations = gt::cells_column_labels(
-      columns = c(Unadjusted, Adjusted)
-    )
-  ) |>
-  gt::tab_footnote(
-    footnote = "Truncated weights.",
-    locations = gt::cells_column_labels(columns = Adjusted)
-  )
-```
-
-```{r balance-3}
-#| label: tbl-balance-3
-#| tbl-cap: Effective number of observations
-
-balance_3 |>
-  dplyr::arrange(dplyr::desc(Adjusted)) |>
-  gt::gt() |>
-  gt::fmt_number(n_sigfig = 4) |>
-  gt::tab_footnote(
-    footnote = "N.",
-    locations = gt::cells_column_labels(
-      columns = c(Unadjusted, Adjusted)
-    )
-  ) |>
-  gt::tab_footnote(
-    footnote = "Truncated weights.",
-    locations = gt::cells_column_labels(columns = Adjusted)
-  )
-```
-
-```{r}
-targets::tar_load(
-  dplyr::matches("weights_(\\d)$"),
-  store = store
-)
-```
-
-```{r weights-1}
-#| label: tbl-weights-1
-#| tbl-cap: Weights
-
-tidy_res_weighted_fits(
-  dat_tbl = weights_1$dat_tbl,
-  dat_plt = weights_1$dat_plt,
-  sa_var = NULL
-)$table
-```
-
-```{r weights-2}
-#| label: tbl-weights-2
-#| tbl-cap: Weights
-
-tidy_res_weighted_fits(
-  dat_tbl = weights_2$dat_tbl,
-  dat_plt = weights_2$dat_plt,
-  sa_var = NULL
-)$table
-```
-
-```{r weights-3}
-#| label: tbl-weights-3
-#| tbl-cap: Weights
-
-tidy_res_weighted_fits(
-  dat_tbl = weights_3$dat_tbl,
-  dat_plt = weights_3$dat_plt,
-  sa_var = NULL
-)$table
-```
+# Tables for main results
 
@@
 #| label: tbl-marginal-1
-#| tbl-cap: Average comparisons
+#| tbl-cap: Marginal comparison results for $hitrtse \sim edc$
 
@@
 #| label: tbl-marginal-2
-#| tbl-cap: Average comparisons
+#| tbl-cap: Marginal comparison results for $metabolite \sim edc$
 
@@
 #| label: tbl-marginal-3
-#| tbl-cap: Average comparisons
+#| tbl-cap: Marginal comparison results for $hitrtse \sim metabolite$
 
@@
 
-# Other analyses
+# Tables for other analyses
 
@@
 #| label: tbl-hypothesis-1
-#| tbl-cap: Hypothesis testing for effect modification
+#| tbl-cap: Hypothesis testing results of effect modification by sex for $hitrtse \sim edc$
 
@@
 #| label: tbl-hypothesis-2
-#| tbl-cap: Hypothesis testing for effect modification
+#| tbl-cap: Hypothesis testing results of effect modification by sex for $metabolite \sim edc$
 
@@
 #| label: tbl-hypothesis-3
-#| tbl-cap: Hypothesis testing for effect modification
+#| tbl-cap: Hypothesis testing results of effect modification by sex for $hitrtse \sim metabolite$
 

### 2023-12-14 output/paper/manuscript.qmd
Transformed into R package cause I wanted to use roxygen2 to document functions but it does not seem to be working. Renamed directory code to R for same reason. Removed virtually all default arguments functions. Testing pipelines.
Moved some tables to SI. Added better captions for tables and figures.

@@
 ```{r vars-methods}
-source(here::here("code/dictionaries.R"))
-source(here::here("code/data.R"))
-source(here::here("code/proc_res.R"))
+source(here::here("R/dictionaries.R"))
+source(here::here("R/data.R"))
+source(here::here("R/proc_res.R"))
 
@@
 
-- Discuss limitations of the study, taking into account sources of potential bias or imprecision.
-- Discuss both direction and magnitude of any potential bias.
-
 Some limitations:
@@
 - Multiple comparisons.
+
+These findings should be interpreted in light of the following limitations and strengths.
+
+Strengths of the present study include...
 

### 2023-12-14 output/paper/manuscript.qmd
Limitations and strenghts.

@@
 
-- All results on which study conclusions or inferences are based, including null findings and results of secondary or sensitivity analyses, must be reported.
-Use of sub-headings that describe the nature of the results (but no declarative statements).
-  - Provide a clear and concise description of all findings without extrapolating beyond the study results.
-  - Do not limit results to those *statistically significant* or that support the study hypotheses.
-Avoid using statistical significance testing as the sole or primary criterion for interpreting the obtained results.
-If significance testing or *p*-values are used, report numeric *p*-values, rounded to 1-2 digits, for all results.
-Use an uppercase italic letter "P", and the values should not be bolded.
-Indicate whether are 1- or 2-sided.
-- Give unadjusted and confounder-adjusted estimates and their precision (e.g., $95\%$ confidence interval).
-Make clear which confounders were adjusted for and why they were included.
-Include the number of observations for each analysis after accounting for missing data.
+- Use an uppercase italic letter "P", and the values should not be bolded. Indicate whether are 1- or 2-sided.
+- Include the number of observations for each analysis after accounting for missing data.
 Include numeric data within figures (e.g., forest plots), or provide tables with corresponding numeric data for all figures.
-  - [`marginaleffects` tables](https://vincentarelbundock.github.io/marginaleffects/articles/tables.html).
-- Report category boundaries when continuous variables were categorized.
 
@@
 
-### What does the literature say?
-
-- EDCs and neurodevelopment (ANT).
-- EDCs and corticosteroids.
-- Corticosteroids and neurodevelopment (ANT).
-
 ## Limitations {#sec-limitations}
-
-Some limitations:
-
-- Cross-sectional study.
-- Chemicals measured in night and morning samples, whereas metabolites (the outcome) were measured only in night samples.
-- Cortisol measured at night, when should be lowest.
-- Change of estimand when trimming weights.
-- Model misspecification.
-- Mixtures effect.
-- Residual confounding.
-- Some confounders were not used since large percentage of missing values.
-- Multiple comparisons.
 
@@
 
-Strengths of the present study include...
+Limitations include the cross-sectional design of the present study.
+Importantly, the non-persistent \acr{edcs} were measured in a pool of night and morning urine samples before the clinical visit, whereas the corticosteroids were measured in the night urine sample only.
+Although we included a wide range of confounders there is the possibility, as with other observational studies, of residual confounding, which might lead to a bias away from the null.
+Some of the confounders indicated in the adjustment sets had to be remove due to large fractions of missing values.
+There is further the possibility of misspecification of the outcome model, although we included a spline of the exposure to relax some of the linearity assumptions.
+The use of more data-adaptive learners was excluded due to the relatively small sample size.
+We finally acknowledge the possibility that some of chemicals might not act independently (mixture effect).
+Further research is thus warranted.
+
+Strengths of the present study include the use of pooled urine samples for chemical assessment, since it is known that these specific \acr{edcs} have very short half-lives [@CasasBasaganaSakhi:2018].
+We decided to model both the *treatment* mechanisms, for the estimation of balancing weights, and the outcomes, with traditional covariates adjustment, to try to obtain *doubly robust* effect estimates.
+Finally, we decided not to interpret our results by focusing on the estimated coefficients of the regression models, but by making use of the g-computation procedure and estimate average comparisons.
 
@@
 
-## Generalisability {#sec-general}
-
-- Discuss the generalisability (external validity) of the study results.
+In conclusion, in a study of...
 

### 2023-12-15 output/paper/_figures.qmd
Merged tables and figures for marginal effects into forest plots.

--- 2023-11-21 output/paper/_figures.qmd
@@
 
+```{r}
+# Function to align forest plot with table numerical results
+forest_plot <- function(obj, r = 4) {
+  # From table of numerical results to ggplot object
+  tbl_num_res <- obj$numerical_results |>
+    tidylog::mutate(
+      variable = forcats::fct_reorder2(
+        variable, estimate, class
+      )
+    ) |>
+    tidylog::select(-estimate) |>
+    ggplot2::ggplot(
+      ggplot2::aes(y = variable)
+    ) +
+    ggplot2::geom_text(
+      ggplot2::aes(
+        x = 0,
+        label = variable
+      ),
+      hjust = 0,
+      fontface = "bold"
+    ) +
+    ggplot2::geom_text(
+      ggplot2::aes(
+        x = 1,
+        label = val
+      ),
+      hjust = 0,
+      fontface = ifelse(
+        obj$numerical_results$val == "Marginal contrast (95% CI)",
+        "bold", "plain"
+      )
+    ) +
+    ggplot2::theme_void() +
+    ggplot2::theme(
+      strip.background = ggplot2::element_blank(),
+      strip.text = ggplot2::element_blank()
+    ) +
+    ggplot2::coord_cartesian(
+      xlim = c(0, 4)
+    )
+
+  # Assemble elements
+  ret <- patchwork::wrap_plots(
+    tbl_num_res, obj$plot
+  ) +
+    patchwork::plot_layout(
+      design = c(
+        patchwork::area(t = 0, l = 1, b = 3, r = r),
+        patchwork::area(t = 1, l = r, b = 3, r = r + 4)
+      )
+    )
+
+  return(ret)
+} # End function forest_plot
+```
+
 - Figure titles and legends should be submitted as single paragraph.
 
-# Main results
+# Figures for main results
 
@@
 #| label: fig-marginal-1
-#| fig-cap: Average comparisons
-#| fig-width: 6
-#| fig-height: 7
+#| fig-cap: Marginal comparison results for $hitrtse \sim edc$
+#| fig-width: 9
+#| fig-height: 10
 
-tidy_res_meffects(
+obj1 <- tidy_res_meffects(
   df = marginal_comparisons_1,
   sa_var = NULL,
+  outcome = "hitrtse",
   which_res = "comparisons",
@@
   num_digits_sig = num_digits_sig
-)$plot
+)
+forest_plot(obj = obj1)
 ```
 
-```{r marginal-2}
-#| label: fig-marginal-2
-#| fig-cap: Average comparisons
+```{r marginal-21}
+#| label: fig-marginal-2-cortisolprod
+#| fig-cap: Marginal comparison results for $cortisol production \sim edc$
 #| fig-width: 9
@@
 
-tidy_res_meffects(
+obj2 <- tidy_res_meffects(
   df = marginal_comparisons_2,
   sa_var = NULL,
+  outcome = "cortisol production",
   which_res = "comparisons",
@@
   num_digits_sig = num_digits_sig
-)$plot
+)
+forest_plot(obj = obj2)
+```
+
+```{r marginal-22}
+#| label: fig-marginal-2-cortisoneprod
+#| fig-cap: Marginal comparison results for $cortisone production \sim edc$
+#| fig-width: 9
+#| fig-height: 10
+
+obj2 <- tidy_res_meffects(
+  df = marginal_comparisons_2,
+  sa_var = NULL,
+  outcome = "cortisone production",
+  which_res = "comparisons",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)
+forest_plot(obj = obj2)
+```
+
+```{r marginal-23}
+#| label: fig-marginal-2-corticosteroneprod
+#| fig-cap: Marginal comparison results for $corticosterone production \sim edc$
+#| fig-width: 9
+#| fig-height: 10
+
+obj2 <- tidy_res_meffects(
+  df = marginal_comparisons_2,
+  sa_var = NULL,
+  outcome = "corticosterone production",
+  which_res = "comparisons",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)
+forest_plot(obj = obj2)
+```
+
+```{r marginal-24}
+#| label: fig-marginal-2-11b
+#| fig-cap: Marginal comparison results for $11bhsd \sim edc$
+#| fig-width: 9
+#| fig-height: 10
+
+obj2 <- tidy_res_meffects(
+  df = marginal_comparisons_2,
+  sa_var = NULL,
+  outcome = "11bhsd",
+  which_res = "comparisons",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)
+forest_plot(obj = obj2)
 ```
@@
 #| label: fig-marginal-3
-#| fig-cap: Average comparisons
-#| fig-width: 5
+#| fig-cap: Marginal comparison results for $hitrtse \sim metabolite$
+#| fig-width: 9
 #| fig-height: 5
 
-tidy_res_meffects(
+obj3 <- tidy_res_meffects(
   df = marginal_comparisons_3,
   sa_var = NULL,
+  outcome = "hitrtse",
   which_res = "comparisons",
@@
   num_digits_sig = num_digits_sig
-)$plot
+)
+forest_plot(obj = obj3, r = 5)
 ```

### 2023-12-15 output/paper/_si.qmd
Merged tables and figures for marginal effects into forest plots.

@@
   sa_var = "e3_sex",
+  outcome = NULL,
   which_res = "comparisons",
@@
 #| fig-width: 9
-#| fig-height: 10
+#| fig-height: 12
 
 tidy_res_meffects(
-  df = marginal_comparisons_2SA,
+  df = marginal_comparisons_2SA |>
+    tidylog::filter(
+      !grepl("metabolism", outcome)
+    ),
   sa_var = "e3_sex",
+  outcome = NULL,
   which_res = "comparisons",
@@
 ```{r marginal-3sa}
-#| fig-width: 5
+#| fig-width: 6
 #| fig-height: 5

### 2023-12-15 output/paper/_tables.qmd
Merged tables and figures for marginal effects into forest plots.

@@
 ```{r}
+#| eval: false
+#| echo: false
+
 targets::tar_load(
@@
 ```{r marginal-1}
+#| eval: false
+#| echo: false
 #| label: tbl-marginal-1
@@
 ```{r marginal-2}
+#| eval: false
+#| echo: false
 #| label: tbl-marginal-2
@@
 ```{r marginal-3}
+#| eval: false
+#| echo: false
 #| label: tbl-marginal-3
@@
   sa_var = "e3sex",
+  outcome = NULL,
   which_res = "hypothesis",

### 2023-12-15 R/proc_res.R

@@ -680,15 +680,37 @@ load_res_meffects <- function(path_store, rq, sa_var, which_res) {
 #'
 #' @param df
 #' @param sa_var
+#' @param outcome
 #' @param which_res
 #' @param num_digits_est
 #' @param num_digits_sig
 #'
 #' @return
 #' @export
-tidy_res_meffects <- function(df, sa_var,
+tidy_res_meffects <- function(df, sa_var, outcome,
                               which_res,
                               num_digits_est, num_digits_sig) {
+  if (!is.null(outcome)) {
+    df <- df |>
+      tidylog::filter(
+        outcome == .env$outcome
+      )
+  }
+  
+  df <- df |>
+    tidylog::mutate(
+      variable = as.character(variable),
+      variable = dplyr::case_match(
+        variable,
+        "corticosterone production" ~ "corticost. prod.",
+        "cortisol production" ~ "cortisol prod.",
+        "cortisone production" ~ "cortisone prod.",
+        "cortisol metabolism" ~ "cortisol met.",
+        "cortisone metabolism" ~ "cortisone met.",
+        .default = variable
+      )
+    )
+  
   # Forest plots side-by-side
   if (is.null(sa_var) | which_res == "hypothesis") {
     plot <- df |>
@@ -720,7 +742,6 @@ tidy_res_meffects <- function(df, sa_var,
     ggstance::position_dodgev(height = 0.5)
   }
   plot <- plot +
-    ggplot2::scale_y_discrete() +
     ggplot2::geom_point(
       mapping = ggplot2::aes(
         size = .data[["s.value"]]
@@ -737,27 +758,54 @@ tidy_res_meffects <- function(df, sa_var,
         xmax = conf.high
       ),
       position = position,
-      width = 0.0,
+      width = 0.3,
       linewidth = 0.3
     ) +
     ggplot2::geom_vline(
-      xintercept = 0.0
-    ) +
-    ggplot2::facet_wrap(
-      ggplot2::vars(outcome),
-      ncol = 2,
-      scales = "free_x"
+      xintercept = 0.0,
+      linetype = "longdash"
     ) +
     ggplot2::labs(
-      y = "exposure",
-      caption = "Size of the dots based on the S value."
+      x = "Marginal contrast"
     ) +
-    ggplot2::theme(
-      plot.caption = ggplot2::element_text(hjust = 0),
-      text = ggplot2::element_text(
-        size = 12
[…]

### 2023-12-19 output/paper/manuscript.qmd
Added list tables and figures for main results.
Merged tables and figures for marginal effects into forest plots.
Writing description results. Trying to merge forest plots by outcome.

@@
 
-- Use an uppercase italic letter "P", and the values should not be bolded. Indicate whether are 1- or 2-sided.
-- Include the number of observations for each analysis after accounting for missing data.
-Include numeric data within figures (e.g., forest plots), or provide tables with corresponding numeric data for all figures.
+The effective sample sizes before and after balancing weights estimation are presented in Tables [-@supptbl-balance-1; -@supptbl-balance-2; -@supptbl-balance-3], while basic summary statistics of the estimated balancing weights are presented in Tables [-@supptbl-weights-1; -@supptbl-weights-2; -@supptbl-weights-3].
+As expected, the median value of the weights for each exposure was close to $1.00$.
+
+@Fig-marginal-1 presents the forest plot for the marginal effects of the non-persistent \acr{edcs} on \acr{hitrtse}.
+Overall, a cohort-specific increase in the levels of the exposures from the 10th to the 90th percentiles was associated with a positive marginal contrast, indicating an increase in the values of \acr{hitrtse}.
+Most of the \acr{ci} included the null effect, though.
+Statistically significant associations were observed for the phenol MEPA (\acr{mc} = $0.042$, $95\%$ \acr{ci}: $0.013, 0.071$), and the phthalate metabolites oxo-MiNP (\acr{mc} = $0.023$, $95\%$ \acr{ci}: $0.003, 0.044$), oh-MiNP (\acr{mc} = $0.039$, $95\%$ \acr{ci}: $0.001, 0.076$), and MEHP (\acr{mc} = $0.036$, $95\%$ \acr{ci}: $0.008, 0.063$).
+The \acr{opp} DETP was strongly negatively associated with \acr{hitrtse} (\acr{mc} = $-0.026$, $95\%$ \acr{ci}: $-0.054, 0.001$).
+
+Figures [-@fig-marginal-2-cortisolprod; -@fig-marginal-2-cortisoneprod; -@fig-marginal-2-corticosteroneprod; -@fig-marginal-2-11b] present the forest plots for the marginal effects of the non-persistent \acr{edcs} on cortisol production, cortisone production, corticosterone production, and 11bHSD activity, respectively.
+
+@Fig-marginal-3 presents the forest plot for the marginal effects of the corticosteroids on \acr{hitrtse}.
+
+Results are presented as effect estimate of the marginal contrast and the $95\%$ \acr{ci}.
+
+suppfig-adrfs-rq...
 
@@
 
-{{< include _tables.qmd >}}
+<!-- {{< include _tables.qmd >}} -->
 
@@
 
-{{< include _si.qmd >}}
+<!-- {{< include _si.qmd >}} -->

### 2023-12-19 R/proc_res.R

@@ -802,7 +802,7 @@ tidy_res_meffects <- function(df, sa_var, outcome,
         strip.background = ggplot2::element_blank(),
         legend.position = "right",
         axis.ticks.y = ggplot2::element_blank(),
-        axis.line.x = ggplot2::element_line(size = 0.6),
+        axis.line.x = ggplot2::element_line(linewidth = 0.6),
         axis.ticks.length = ggplot2::unit(0.3, "cm")
       )
   }
@@ -875,6 +875,12 @@ tidy_res_meffects <- function(df, sa_var, outcome,
     ) |>
     gt::tab_footnote(
       footnote = "*Significant results."
+    ) |>
+    gt::tab_footnote(
+      footnote = "Estimate and 95% CI.",
+      locations = gt::cells_column_labels(
+        columns = names_
+      )
     )
   
   if (is.null(sa_var) & which_res == "comparisons") {

### 2023-12-19 output/paper/_figures.qmd
Forst plots with multi-outcome for RQ2.

--- 2023-11-21 output/paper/_figures.qmd
@@
 
+```{r}
+# Function to align forest plot with table numerical results
+forest_plot <- function(obj, r = 4) {
+  # From table of numerical results to ggplot object
+  tbl_num_res <- obj$numerical_results |>
+    tidylog::mutate(
+      variable = forcats::fct_reorder2(
+        variable, estimate, class
+      )
+    ) |>
+    tidylog::select(-estimate) |>
+    ggplot2::ggplot(
+      ggplot2::aes(
+        y = variable,
+        shape = outcome
+      )
+    ) +
+    ggplot2::geom_text(
+      ggplot2::aes(
+        x = 0,
+        label = variable
+      ),
+      hjust = 0,
+      fontface = "bold"
+    ) +
+    ggplot2::geom_text(
+      ggplot2::aes(
+        x = 1,
+        label = val
+      ),
+      hjust = 0,
+      position = ggstance::position_dodgev(height = 0.9)
+    ) +
+    ggplot2::theme_void() +
+    ggplot2::theme(
+      strip.background = ggplot2::element_blank(),
+      strip.text = ggplot2::element_blank(),
+      legend.position = "bottom",
+      legend.box = "vertical",
+      legend.margin = ggplot2::margin()
+    ) +
+    ggplot2::coord_cartesian(
+      xlim = c(0, 4)
+    )
+
+  # Assemble elements
+  ret <- patchwork::wrap_plots(
+    tbl_num_res, obj$plot
+  ) +
+    patchwork::plot_layout(
+      design = c(
+        patchwork::area(t = 0, l = 1, b = 3, r = r),
+        patchwork::area(t = 1, l = r, b = 3, r = r + 4)
+      )
+    )
+
+  return(ret)
+} # End function forest_plot
+```
+
 - Figure titles and legends should be submitted as single paragraph.
 
-# Main results
+# Figures for main results
 
@@
 #| label: fig-marginal-1
-#| fig-cap: Average comparisons
-#| fig-width: 6
-#| fig-height: 7
+#| fig-cap: Marginal comparison results for $hitrtse \sim edc$
+#| fig-width: 9
+#| fig-height: 10
 
-tidy_res_meffects(
+obj1 <- tidy_res_meffects(
   df = marginal_comparisons_1,
   sa_var = NULL,
+  outcome = c("hitrtse"),
   which_res = "comparisons",
@@
   num_digits_sig = num_digits_sig
-)$plot
+)
+forest_plot(obj = obj1)
 ```
@@
 #| label: fig-marginal-2
-#| fig-cap: Average comparisons
+#| fig-cap: Marginal comparison results for $production \sim edc$
 #| fig-width: 9
-#| fig-height: 10
+#| fig-height: 12
 
-tidy_res_meffects(
+obj2 <- tidy_res_meffects(
   df = marginal_comparisons_2,
   sa_var = NULL,
+  outcome = c(
+    "cortisol production",
+    "cortisone production",
+    "corticosterone production"
+  ),
   which_res = "comparisons",
@@
   num_digits_sig = num_digits_sig
-)$plot
+)
+forest_plot(obj = obj2)
 ```
@@
 #| label: fig-marginal-3
-#| fig-cap: Average comparisons
-#| fig-width: 5
+#| fig-cap: Marginal comparison results for $hitrtse \sim metabolite$
+#| fig-width: 9
 #| fig-height: 5
 
-tidy_res_meffects(
+obj3 <- tidy_res_meffects(
   df = marginal_comparisons_3,
   sa_var = NULL,
+  outcome = c("hitrtse"),
   which_res = "comparisons",

### 2023-12-20 output/paper/manuscript.qmd
Nothing.

@@
 
-Figures [-@fig-marginal-2-cortisolprod; -@fig-marginal-2-cortisoneprod; -@fig-marginal-2-corticosteroneprod; -@fig-marginal-2-11b] present the forest plots for the marginal effects of the non-persistent \acr{edcs} on cortisol production, cortisone production, corticosterone production, and 11bHSD activity, respectively.
+@Fig-marginal-2 presents the forest plot for the marginal effects of the non-persistent \acr{edcs} on cortisol production, cortisone production, corticosterone production, and 11bHSD activity, respectively.
 

### 2023-12-21 output/paper/_tables.qmd
Writing results.

@@
   sa_var = NULL,
+  outcome = NULL,
   which_res = "comparisons",

### 2023-12-21 output/paper/manuscript.qmd
Writing results.

@@
 
+```{r}
+format_res <- function(df, exposure, outcome) {
+  # df: variable, val, estimate, outcome, class
+  # exposure in variable, outcome in outcome
+
+  val <- df[df$variable == exposure & df$outcome == outcome, ]$val
+  val <- stringr::str_replace(
+    string = val,
+    pattern = "\\*",
+    replacement = ""
+  )
+  glue::glue(
+    "\\acr{{mc}} and $95\\%$ \\acr{{ci}}: {val}"
+  )
+}
+```
+
+```{r}
+#| echo: false
+
+targets::tar_load(
+  dplyr::matches("marginal_comparisons_(\\d)$"),
+  store = store
+)
+meffects1 <- tidy_res_meffects(
+  df = marginal_comparisons_1,
+  sa_var = NULL,
+  outcome = c("hitrtse"),
+  which_res = "comparisons",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)$numerical_results
+meffects2 <- tidy_res_meffects(
+  df = marginal_comparisons_2,
+  sa_var = NULL,
+  outcome = c(
+    "cortisol production",
+    "cortisone production",
+    "corticosterone production"
+  ),
+  which_res = "comparisons",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)$numerical_results
+meffects3 <- tidy_res_meffects(
+  df = marginal_comparisons_3,
+  sa_var = NULL,
+  outcome = c("hitrtse"),
+  which_res = "comparisons",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)$numerical_results
+```
+
 The effective sample sizes before and after balancing weights estimation are presented in Tables [-@supptbl-balance-1; -@supptbl-balance-2; -@supptbl-balance-3], while basic summary statistics of the estimated balancing weights are presented in Tables [-@supptbl-weights-1; -@supptbl-weights-2; -@supptbl-weights-3].
@@
 Overall, a cohort-specific increase in the levels of the exposures from the 10th to the 90th percentiles was associated with a positive marginal contrast, indicating an increase in the values of \acr{hitrtse}.
-Most of the \acr{ci} included the null effect, though.
-Statistically significant associations were observed for the phenol MEPA (\acr{mc} = $0.042$, $95\%$ \acr{ci}: $0.013, 0.071$), and the phthalate metabolites oxo-MiNP (\acr{mc} = $0.023$, $95\%$ \acr{ci}: $0.003, 0.044$), oh-MiNP (\acr{mc} = $0.039$, $95\%$ \acr{ci}: $0.001, 0.076$), and MEHP (\acr{mc} = $0.036$, $95\%$ \acr{ci}: $0.008, 0.063$).
-The \acr{opp} DETP was strongly negatively associated with \acr{hitrtse} (\acr{mc} = $-0.026$, $95\%$ \acr{ci}: $-0.054, 0.001$).
+Most of the \acr{ci}s included the null effect, though.
+Statistically significant associations were observed for the phenol MEPA (`r format_res(meffects1, "mepa", "hitrtse")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "hitrtse")`), oh-MiNP (`r format_res(meffects1, "ohminp", "hitrtse")`), and MEHP (`r format_res(meffects1, "mehp", "hitrtse")`).
+The \acr{opp} DETP was strongly negatively associated with \acr{hitrtse} (`r format_res(meffects1, "detp", "hitrtse")`).
 
-@Fig-marginal-2 presents the forest plot for the marginal effects of the non-persistent \acr{edcs} on cortisol production, cortisone production, corticosterone production, and 11bHSD activity, respectively.
+@Fig-marginal-2 presents the forest plot for the marginal effects of the non-persistent \acr{edcs} on cortisone production, cortisol production, and corticosterone production.
+Overall, a cohort-specific increase in the levels of the exposures from the 10th to the 90th percentiles was associated with a positive marginal contrast, indicating an increase in the total production of these metabolites, with the exception of BUPA, which was associated with negative marginal contrasts for all three outcomes, and MiBP, which was associated with a negative marginal contrast for total cortisone production only.
+The majority of the associations for the phenols and phthalate metabolites included the null.
+The phenol BPA showed the largest marginal contrasts across all three outcomes (cortisone production, `r format_res(meffects2, "bpa", "cortisone production")`; cortisol production, `r format_res(meffects1, "bpa", "cortisol production")`; corticosterone production, `r format_res(meffects1, "bpa", "corticosterone production")`).
 
@@
 
-<!-- {{< include _tables.qmd >}} -->
+{{< include _tables.qmd >}}
 
@@
 
-<!-- {{< include _si.qmd >}} -->
+{{< include _si.qmd >}}

### 2023-12-21 R/proc_res.R

@@ -810,7 +810,7 @@ tidy_res_meffects <- function(df, sa_var, outcome,
         axis.text.y = ggplot2::element_blank(),
         axis.ticks.y = ggplot2::element_blank(),
         axis.line.y = ggplot2::element_blank(),
-        axis.line.x = ggplot2::element_line(size = 0.6),
+        axis.line.x = ggplot2::element_line(linewidth = 0.6),
         axis.ticks.length = ggplot2::unit(0.3, "cm")
       )
   } else {

### 2023-12-21 output/paper/_si.qmd
Writing results.
Finished describing main results and modification by sex.

@@
 
-## Figures for main results
-
-### Average dose-response functions for significant results
-
-```{r}
-targets::tar_load(
-  dplyr::matches("marginal_gcomp_(\\d)$"),
-  store = store
-)
-```
-
-::: {#suppfig-adrfs-rq3}
-```{r adrfs-rq3}
-#| fig-width: 6
-#| fig-height: 7
-
-exposures_of_interest <- c(
-  "cortisol production",
-  "X11bHSD"
-)
-ret_adrfs <- lapply(
-  exposures_of_interest, function(x) {
-    plot_adrf(
-      df_preds = marginal_gcomp_3 |>
-        tidylog::filter(
-          variable == x
-        )
-    )
-  }
-)
-ret_adrfs[[1]]
-#ret_adrfs[[2]]
-```
-Average dose-response functions for...
-:::
-
 ## Figures for other results
@@
 :::
+
+### Sensitivity to exposure assessment

### 2023-12-21 output/paper/_tables.qmd
Finished describing main results and modification by sex.

--- 2023-12-15 output/paper/_tables.qmd
@@
 
-# Tables for main results
-
-## Marginal comparisons
-
-```{r}
-#| eval: false
-#| echo: false
-
-targets::tar_load(
-  dplyr::matches("marginal_comparisons_(\\d)$"),
-  store = store
-)
-```
-
-```{r marginal-1}
-#| eval: false
-#| echo: false
-#| label: tbl-marginal-1
-#| tbl-cap: Marginal comparison results for $hitrtse \sim edc$
-
-tidy_res_meffects(
-  df = marginal_comparisons_1,
-  sa_var = NULL,
-  which_res = "comparisons",
-  num_digits_est = num_digits_est,
-  num_digits_sig = num_digits_sig
-)$table
-```
-
-```{r marginal-2}
-#| eval: false
-#| echo: false
-#| label: tbl-marginal-2
-#| tbl-cap: Marginal comparison results for $metabolite \sim edc$
-
-tidy_res_meffects(
-  df = marginal_comparisons_2,
-  sa_var = NULL,
-  which_res = "comparisons",
-  num_digits_est = num_digits_est,
-  num_digits_sig = num_digits_sig
-)$table
-```
-
-```{r marginal-3}
-#| eval: false
-#| echo: false
-#| label: tbl-marginal-3
-#| tbl-cap: Marginal comparison results for $hitrtse \sim metabolite$
-
-tidy_res_meffects(
-  df = marginal_comparisons_3,
-  sa_var = NULL,
-  which_res = "comparisons",
-  num_digits_est = num_digits_est,
-  num_digits_sig = num_digits_sig
-)$table
-```
-
 # Tables for other analyses
 
-## Marginal hypothesis
+## Marginal hypotheses for effect modification
 

### 2023-12-21 output/paper/manuscript.qmd
Finished describing main results and modification by sex.

@@
 @Fig-marginal-3 presents the forest plot for the marginal effects of the corticosteroids on \acr{hitrtse}.
-
-Results are presented as effect estimate of the marginal contrast and the $95\%$ \acr{ci}.
-
-suppfig-adrfs-rq...
+All marginal contrasts included the null, with no clear indication of directionality of the effect.
 
@@
 
-- Report other analyses done (e.g., analyses of subgroups and interactions, and sensitivity analyses).
+### Effect modification by sex
+
+```{r}
+#| echo: false
+
+targets::tar_load(
+  dplyr::matches("marginal_hypothesis_(\\d)SA$"),
+  store = store
+)
+heffects1 <- tidy_res_meffects(
+  df = marginal_hypothesis_1SA,
+  sa_var = "e3sex",
+  outcome = NULL,
+  which_res = "hypothesis",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)$numerical_results
+heffects2 <- tidy_res_meffects(
+  df = marginal_hypothesis_2SA,
+  sa_var = "e3sex",
+  outcome = NULL,
+  which_res = "hypothesis",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)$numerical_results
+heffects3 <- tidy_res_meffects(
+  df = marginal_hypothesis_3SA,
+  sa_var = "e3sex",
+  outcome = NULL,
+  which_res = "hypothesis",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)$numerical_results
+```
+
+Basic summary statistics of the estimated balancing weights for effect modification are presented in Tables [-@supptbl-weights-1sa; -@supptbl-weights-2sa; -@supptbl-weights-3sa].
+As expected, the median value of the weights for each exposure was close to $1.00$.
+
+@Tbl-hypothesis-1 presents the results of a hypothesis test for the difference between estimates of the marginal contrasts for females and males, for each exposure.
+Significant differences were present for the phenol OXBE (`r format_res(heffects1, "oxbe", "hitrtse")`) and the phthalate metabolites MEP (`r format_res(heffects1, "mep", "hitrtse")`) and MbZP (`r format_res(heffects1, "mbzp", "hitrtse")`).
+The forest plot of the individual marginal contrasts is presented in @suppfig-marginal-1sa.
+
+@Tbl-hypothesis-2 presents the results of a hypothesis test for the difference between estimates of the marginal contrasts for females and males, for each exposure and outcome.
+Significant differences were present across all three classes of \acr{edcs} and for all outcomes.
+The largest differences were attributable to the \acr{opps} DMTP (cortisol production, `r format_res(heffects2, "dmtp", "cortisol production")`) and DETP (corticosterone production, (`r format_res(heffects2, "detp", "corticosterone production")`); cortisone production, (`r format_res(heffects2, "detp", "cortisone production")`)).
+The forest plots of the individual marginal contrasts are presented in @suppfig-marginal-2sa.
+
+@Tbl-hypothesis-3 presents the results of a hypothesis test for the difference between estimates of the marginal contrasts for females and males, for each exposure.
+Significant differences were present for cortisone production (`r format_res(heffects3, "cortisone prod.", "hitrtse")`) and corticosterone production (`r format_res(heffects3, "corticost.
+prod.", "hitrtse")`).
+Furthermore, for all exposures, the marginal contrasts had opposite sign (positive for males and negative for females).
+The forest plot of the individual marginal contrasts is presented in @suppfig-marginal-3sa.
+
+### Sensitivity to exposure assessment
 
@@
 
-# Funding
-
-{{< pagebreak >}}
-
 # References {.unnumbered}

### 2023-12-27 output/paper/_figures.qmd
Nothing.

@@
 ```
+
+**The figure titles and legends should be submitted as a single paragraph.
+The text of figure legends beyond the scale, arrow, panel locant, and axis information should not be included on the figures themselves but should be typed after the reference list.
+Each legend should be formatted as a separate paragraph and should include details on the place of the study, the time of the study, and the study population (if applicable).
+Define all figure abbreviations in the legend.**
 

### 2023-12-27 output/paper/_tables.qmd
Nothing.

@@
 ```
+
+**Table titles should give details on the place of the study, the time of the study, and the study population (if applicable).
+Example, “Baseline Characteristics of Infants With Initial and Follow-up Screening, London, United Kingdom, 2001–2003.” In the table body, leave blank spaces for no entry; avoid using dashes.
+Order of footnotes: 1) Abbreviations: (no footnote symbol, listed alphabetically, separated by semicolons); 2) other footnotes as necessary, each preceded by a superscript lowercase letter.
+Footnotes must be cited in order when reading from left to right and top to bottom.**
 

### 2023-12-27 output/paper/manuscript.qmd
Nothing.

@@
 
-## Key results {#sec-disc-res-key}
-
-- Summarise key results with reference to study objectives.
-- Provide a review of the relevant literature to put the study findings into context.
-  - It should be complete and balanced, including inconsistent results.
-  - It should include, for each source, sufficient details: study design, sample size, population, specific exposures and outcomes.
-
-## Limitations {#sec-limitations}
-
 These findings should be interpreted in light of the following limitations and strengths.
@@
 Finally, we decided not to interpret our results by focusing on the estimated coefficients of the regression models, but by making use of the g-computation procedure and estimate average comparisons.
-
-## Interpretation {#sec-interpretation}
-
-- End with a summary of the key findings and their implications for the study hypotheses, future research, and policy.
-- Give a cautious overall interpretation of results considering objectives, limitations, multiplicity of analyses, results from similar studies, and other relevant evidence.
 

### 2023-12-27 output/paper/manuscript.qmd
Discussion of results and intro with corticosteroids.

@@
     - kind: float
-      ref-type: supptbl
+      key: supptbl
       latex-env: supptbl
-      prefix: Table S
-      name: Table S
+      reference-prefix: Table S
       space-before-numbering: false
     - kind: float
-      ref-type: suppfig
+      key: suppfig
       latex-env: suppfig
-      prefix: Figure S
-      name: Figure S
+      reference-prefix: Figure S
       space-before-numbering: false
@@
 Do not include conflicts of interest.
-- The use of the word "effect(s)" as a proxy for "association(s)" is discouraged.
 - The length of the abstract must be less or equal than 200 words, and should be unstructured, stating the research questions, the methods used, and the results and conclusions of the research.
@@
 
-Another major limitation of these epidemiological studies, especially when non-persistent chemicals are analyzed, is the reliance on single-spot urine samples for exposure assessment, which might not be representative of long-term exposures.
+There is furthermore evidence that exposure to certain \acr{edcs}, specifically phthalates, might interfere with the \acr{hpa} axis and might interact with the glucocorticoid receptor [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023].
 
-Something about corticosteroids...
+Finally, the development of the fetal brain is regulated by, among others, glucocorticoids [@SunLiJin:2018], with regulation of their homeostasis being crucial for neurodevelopment in childhood [@SearsLiuLanphear:2023].
 
-In the present study, we estimated associations between 1) non-persistent \acr{edcs} and attention, 2) non-persistent \acr{edcs} and corticosteroids, and 3) corticosteroids and attention, using the parametric g-formula and average contrasts, in children of a large prospective birth cohort in Europe.
+Taken together, these results suggest that the negative influence of exposure to certain \acr{edcs} on neurodevelopmental outcomes, might be mediated, at least partially, by disruption of the \acr{hpa} axis' homeostasis.
+In the present study, we thus estimated associations between 1) non-persistent \acr{edcs} and attention, 2) non-persistent \acr{edcs} and corticosteroids, and 3) corticosteroids and attention, using the parametric g-formula and average contrasts, in children of a large prospective birth cohort in Europe.
 
@@
 
+The impact of exposure to (non-persistent) \acr{edcs} on human health has attracted considerable research interest.
+While research in this area has mainly investigated the effects of prenatal exposure on child neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022], little is still known about later-in-life exposure and childhood neurodevelopment.
+In this study, consisting of almost 1,000 children from 6 European birth cohorts, we observed that short-term childhood exposure to non-persistent \acr{edcs} had negative effects on \acr{hitrtse} and total production of cortisol, cortisone, and corticosterone, with the majority of the \acr{ci}s including the null effect.
+Increased production of these glucocorticoids did not seem to affect \acr{hitrtse}.
+Some of these effects differed for females and males, including significant differences for the effects of increased production of cortisone and corticosterone on \acr{hitrtse}.
+Specifically, an increased production of these glucocorticoids was associated with lower values of \acr{hitrtse} for females, and higher values for males.
+Taken together, these results suggest that these non-persistent \acr{edcs} might be responsible for perturbations of the \acr{hpa} axis' homeostasis, and that higher levels of these glucocorticoids might interfere with different functions of attention in a sex-specific manner.
+
+Our results are consistent with prior research that associated exposure to certain \acr{edcs} with higher levels of urinary cortisol [@KimLeeMoon:2018; @SearsLiuLanphear:2023].
+Contrary to these prior studies, we did find effect modification by sex.
+We are not aware of other epidemiological studies investigating phthalates metabolites, phenols, and \acr{opps}, in relation to urinary glucocorticoids in childhood.
+We are also not aware of prior studies specifically investigating the effects of exposure to \acr{edcs} and elevated levels of glucocorticoids in relation to \acr{hitrtse}.
+
+[] EDCs to HITRTSE
+[X] EDCs to CORT.
+[] CORT. to HITRTSE
+
+---
+
 These findings should be interpreted in light of the following limitations and strengths.

### 2023-12-28 output/paper/manuscript.qmd
Working on manuscript.

@@
 
-Our results are consistent with prior research that associated exposure to certain \acr{edcs} with higher levels of urinary cortisol [@KimLeeMoon:2018; @SearsLiuLanphear:2023].
+Our results are consistent with prior research that associated exposure to certain \acr{edcs} with higher levels of urinary and hair cortisol [@KimLeeMoon:2018; @SearsLiuLanphear:2023].
 Contrary to these prior studies, we did find effect modification by sex.

### 2023-12-29 docs/bib_edc_cort_neuro.qmd
Started doc for references EDCs, corticosteroids, and nuerodevelopment.

---
title: "Some references on EDCs, corticosteroids, and neurodevelopment"

format:
  pdf:
    pdf-engine: pdflatex
    number-sections: true
    number-depth: 2
    toc: true
    toc-depth: 2

bibliography: ../output/paper/bibliography.bib

link-citations: true
linkcolor: blue

execute:
  echo: false
  warning: false
  error: false
  message: false

editor_options:
  chunk_output_type: console
---

``` {r template}
#| echo: false
#| eval: false

## @

- Title: **.
- Exposures: .
- Medium exposure assessment: .
- Timing exposure assessment: .
- Outcomes: .
- Medium outcome assessment: .
- Timing outcome assessment: .
- Results:
    - .
- Sex-specific results: .
```

# EDCs and neurodevelopment

## @BouchardBellingerWright:2010

- Title: *Attention-Deficit/Hyperactivity Disorder and Urinary Metabolites of Organophosphate Pesticides*.
- Exposures: dialkylphosphate metabolites (biomarkers of organophosphate pesticide exposure).
- Medium exposure assessment: spot urine samples.
- Timing exposure assessment: ages 8-15.
- Outcomes: ADHD diagnostic status.
- Timing outcome assessment: ages 8-15.
- Results:
  - Exposure to OP pesticides might contribute to ADHD.
- Sex-specific results: NA.

## @CartierWarembourgLeManer-Idrissi:2016

- Title: *Organophosphate Insecticide Metabolites in Prenatal and Childhood Urine Samples and Intelligence Scores at 6 Years of Age: Results from the Mother–Child PELAGIE Cohort*.
- Exposures: dialkylphosphate metabolites (biomarkers of organophosphate pesticide exposure).
- Medium exposure assessment: spot urine samples.
- Timing exposure assessment: age 6.
- Outcomes: Wechsler Intelligence Scale for Children.
- Timing outcome assessment: age 6.
- Results:
  - No association between total DAP metabolites and WISC scores.
  - WISC working memory score inversely associated with levels of DE metabolites.
- Sex-specific results: NA.

## @FurlongHerringBuckley:2017

- Title: *Prenatal exposure to organophosphorus pesticides and childhood neurodevelopmental phenotypes*.
- Exposures: organophosphorus pesticides. Total diethylphosphate and total dimethylphosphate.
- Medium exposure assessment: spot urine samples.
- Timing exposure assessment: prenatal between 25 and 40 weeks of gestation.
- Outcomes: behavior, executive functioning, and IQ.
- Timing outcome assessment: ages 6–9.
- Results:
  - Dimethylphosphate metabolites negatively associated with Internalizing factor scores.
  - Dimethylphosphate metabolites positively associated with Executive Functioning factor scores.
  - Diethylphosphate metabolites negatively associated with Working Memory Index.
- Sex-specific results: no interactions by sex.

## @

- Title: **.
- Exposures: .
- Medium exposure assessment: .
- Timing exposure assessment: .
- Outcomes: .
- Medium outcome assessment: .
- Timing outcome assessment: .
- Results:
    - .
- Sex-specific results: .

## @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015

- Title: *Pre- and postnatal exposures to pesticides and neurodevelopmental effects in children living in agricultural communities from South-Eastern Spain*.
- Exposures: dialkylphosphate metabolites (biomarkers of organophosphate pesticide exposure).
- Medium exposure assessment: urine samples.
- Timing exposure assessment: ages 6-11.
- Outcomes: Wechsler Intelligence Scale for Children.
- Timing outcome assessment: ages 6-11.
- Results:
  - DAP levels inversely associated with performance on intelligence quotient and verbal comprehension domain.
- Sex-specific results: stronger associations in males.

## @HuangChenSu:2015

- Title: *Association of Exposure to Endocrine-Disrupting Chemicals During Adolescence With Attention-Deficit/Hyperactivity Disorder–Related Behaviors*.
- Exposures: seven metabolite of phthalate esters.
- Medium exposure assessment: urine samples.
- Timing exposure assessment: prenatal and at ages 2, 5, 8, and 11.
- Outcomes: Bayley and Wechsler tests for assessing neurocognitive functions and intelligence (IQ).
- Timing outcome assessment: ages 2, 5, 8, and 11.
- Results:
  - Children’s phthalate exposure (MEOHP and sum of DEHP metabolites) inversely associated with cognitive development.
- Sex-specific results: NA.

## @HuangTsaiChen:2017

- Title: *Intellectual evaluation of children exposed to phthalate-tainted products after the 2011 Taiwan phthalate episode*.
- Exposures: 5 phthalate metabolites.
- Medium exposure assessment: first-morning urine samples.
- Timing exposure assessment: ages 3-12.
- Outcomes: Wechsler tests for assessing the children's intelligence quotient.
- Timing outcome assessment: ages 3-12.
- Results:
  - MEOHP, MnBP, and MiBP inversely associated with verbal intelligence.
- Sex-specific results: NA.

## @KimHongShin:2017

- Title: *The effects of maternal and children phthalate exposure on the neurocognitive function of 6-year-old children*.
- Exposures: phthalate metabolites.
- Medium exposure assessment: urine samples.
- Timing exposure assessment: age 6
- Outcomes: IQ scores and continuous performance test variables.
- Timing outcome assessment: age 6.
- Results:
  - DEHP metabolites (including MEHHP and MEOHP) inversely associated with intelligence, attention, and response time variability.
- Sex-specific results: NA.

## @LiZhangKuang:2018

- Title: *Relationship between bisphenol A exposure and attention-deficit/ hyperactivity disorder: A case-control study for primary school children in Guangzhou, China*.
- Exposures: bisphenol-A.
- Medium exposure assessment: spot urine samples.
- Timing exposure assessment: ages 6-12.
- Outcomes: ADHD.
- Timing outcome assessment: ages 6-12.
- Results:
  - Positive association between BPA levels and odds of ADHD.
- Sex-specific results: stronger associations in males.

## @

- Title: **.
- Exposures: .
- Medium exposure assessment: .
- Timing exposure assessment: .
- Outcomes: .
- Medium outcome assessment: .
- Timing outcome assessment: .
- Results:
    - .
- Sex-specific results: .

## @Rodriguez-CarrilloMustielesPerez-Lobato:2019

- Title: *Bisphenol A and cognitive function in school-age boys: Is BPA predominantly related to behavior?*.
- Exposures: bisphenol-A.
- Medium exposure assessment: spot urine samples.
- Timing exposure assessment: ages 9-11.
- Outcomes: comprehensive neuropsychological test battery.
- Timing outcome assessment: ages 9-11.
- Results:
  - No consistent association between BPA levels and cognitive abilities.
  - Possible inverse association between BPA levels and working memory.
- Sex-specific results: assessment only in males.

## @ShoaffCoullWeuve:2020

- Title: *Association of Exposure to Endocrine-Disrupting Chemicals During Adolescence With Attention-Deficit/Hyperactivity Disorder–Related Behaviors*.
- Exposures: phthalates, parabens, phenols, and triclocarban.
- Medium exposure assessment: urine samples.
- Timing exposure assessment: age 15.3
- Outcomes: Conners Attention Deficit Scale and the Behavior Assessment System for Children.
- Timing outcome assessment: age 15.3
- Results:
  - Exposure to antiandrogenic phthalates increases the risk of ADHD-related behavior problems.
- Sex-specific results: stronger associations in males.

## @TewarAuingerBraun:2016

- Title: *Association of Bisphenol A exposure and Attention-Deficit/Hyperactivity Disorder in a national sample of U.S. children*.
- Exposures: bisphenol-A.
- Medium exposure assessment: spot urine samples.
- Timing exposure assessment: ages 8-15.
- Outcomes: presence of ADHD in the past year.
- Timing outcome assessment: ages 8-15.
- Results:
  - Positive association between BPA levels and ADHD.
- Sex-specific results: stronger associations in males.

## @

- Title: **.
- Exposures: .
- Medium exposure assessment: .
- Timing exposure assessment: .
- Outcomes: .
- Medium outcome assessment: .
- Timing outcome assessment: .
- Results:
    - .
- Sex-specific results: .

## @YuDuChiou:2016

- Title: *Increased risk of attention-deficit/hyperactivity disorder associated with exposure to organophosphate pesticide in Taiwanese children*.
- Exposures: dialkylphosphate metabolites (biomarkers of organophosphate pesticide exposure).
- Medium exposure assessment: urine samples.
- Timing exposure assessment: ages 4-15.
- Outcomes: ADHD.
- Timing outcome assessment: ages 4-15.
- Results:
  - Dose–response relationship between urinary concentrations of dimethylphosphate and ADHD.
- Sex-specific results: NA.

TODO:

- Gascon et al. 2015
- Oh et al. 2023
- Vilmand et al. 2023

---

# EDCs and glucocorticosteroids

## @

- Title: **.
- Exposures: .
- Medium exposure assessment: .
- Timing exposure assessment: .
- Outcomes: .
- Medium outcome assessment: .
- Timing outcome assessment: .
- Results:
    - .
- Sex-specific results: .

## @SearsLiuLanphear:2023

- Title: *Evaluating mixtures of urinary phthalate metabolites and serum per-/polyfluoroalkyl substances in relation to adolescent hair cortisol: The HOME Study*.
- Exposures: phthalate metabolites.
- Medium exposure assessment: spot urine samples.
- Timing exposure assessment: ages 1-5 and 8.
- Outcomes: cortisol.
- Medium outcome assessment: hair.
- Timing outcome assessment: age 12.
- Results:
    - Positive association between mixture of phthalates metabolites measured in childhood and hair cortisol measured in adolescence. Driven by MEP, MiBP, MBzP.
- Sex-specific results: no evidence of modification by sex.

## @

- Title: **.
- Exposures: .
- Medium exposure assessment: .
- Timing exposure assessment: .
- Outcomes: .
- Medium outcome assessment: .
- Timing outcome assessment: .
- Results:
    - .
- Sex-specific results: .

TODO:

- Kim et al. 2018
- Sun et al. 2018

---

# Glucocorticosteroids and neurodevelopment

---

# References {.unnumbered}

::: {#refs}
:::

### 2023-12-31 docs/bib_edc_cort_neuro.qmd
Finished doc with references.

@@
 ---
-title: "Some references on EDCs, corticosteroids, and neurodevelopment"
+title: "Some references on EDCs, glucocorticosteroids, and neurodevelopment"
 
@@
 
-## @
-
-- Title: **.
-- Exposures: .
-- Medium exposure assessment: .
-- Timing exposure assessment: .
-- Outcomes: .
-- Medium outcome assessment: .
-- Timing outcome assessment: .
-- Results:
-    - .
-- Sex-specific results: .
-
 ## @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015
@@
 
-## @
+## @OhKimKannan:2023
 
-- Title: **.
-- Exposures: .
-- Medium exposure assessment: .
-- Timing exposure assessment: .
-- Outcomes: .
-- Medium outcome assessment: .
-- Timing outcome assessment: .
+- Title: *Early childhood exposure to environmental phenols and parabens, phthalates, organophosphate pesticides, and trace elements in association with attention de cit hyperactivity disorder (ADHD) symptoms in the CHARGE study*.
+- Exposures: phenols/parabens, phthalates, OP pesticides, trace elements.
+- Medium exposure assessment: spot urine samples.
+- Timing exposure assessment: ages 2-5.
+- Outcomes: ADHD symptoms.
+- Timing outcome assessment: ages 2-5.
 - Results:
-    - .
-- Sex-specific results: .
+    - Positive associations of MECPP and MEHHP with Hyperactivity subscale scores and hyperactivity/impulsivity subdomain scores.
+    - Positive associations of MNBP with hyperactivity/impulsivity subdomain scores.
+    - Inverse associations of TCS with Hyperactivity subscale scores.
+    - Inverse associations of TCS with inattention subdomain scores.
+    - Positive associations of mixture of phthalates with scores of the ABC Hyperactivity subscale and hyperactivity/impulsivity subdomain.
+      - Drivers by subdomain, hyperactivity/impulsivity subdomain: MHPP, MEHHP, and MBZP.
+    - Positive associations of mixture of phenols/parabens with inattention subdomain score.
+      - Drivers by subdomain, inattention subdomain: MEPB, PRPB, and OHMEP.
+- Sex-specific results: association of mixture of phthalates with hyperactivity-related subscale and subdomain significant in males only.
 
@@
 
-## @
+## @VilmandBeckBilenberg:2023
 
-- Title: **.
-- Exposures: .
-- Medium exposure assessment: .
-- Timing exposure assessment: .
-- Outcomes: .
-- Medium outcome assessment: .
-- Timing outcome assessment: .
+- Title: *Prenatal and current phthalate exposure and cognitive development in 7-year-old children from the Odense child cohort*.
+- Exposures: phthalates metabolites.
+- Medium exposure assessment: spot urine samples.
+- Timing exposure assessment: 7 years.
+- Outcomes: IQ (subtests from Wechsler Intelligence Scale for Children).
+- Timing outcome assessment: 7 years.
 - Results:
-    - .
-- Sex-specific results: .
+    - Inverse associations of MCPP, sum of DnHxPm, sum of DiDPm, and sum of DiNPm with IQ points.
+- Sex-specific results: no evidence of effect modification by sex.
 
@@
 
-TODO:
-
-- Gascon et al. 2015
-- Oh et al. 2023
-- Vilmand et al. 2023
-
 ---
@@
 
-## @
+## @KimLeeMoon:2018
 
-- Title: **.
-- Exposures: .
-- Medium exposure assessment: .
-- Timing exposure assessment: .
-- Outcomes: .
-- Medium outcome assessment: .
-- Timing outcome assessment: .
+- Title: *Association of phthalate exposures with urinary free cortisol and 8-hydroxy-2'-deoxyguanosine in early childhood*.
+- Exposures: four phthalate metabolites.
+- Medium exposure assessment: spot urine samples.
+- Timing exposure assessment: prenatal and at ages 0, 3, 9, 12, 15 months.
+- Outcomes: free cortisol.
+- Medium outcome assessment: spot urine samples.
+- Timing outcome assessment: prenatal and at ages 0, 3, 9, 12, 15 months.
 - Results:
-    - .
-- Sex-specific results: .
+    - Positive associations between MEHHP, MEOHP, MiBP, and MnBP, with free cortisol.
+- Sex-specific results: no evidence of modification by sex.
 
@@
 
-## @
+## @SunLiJin:2018
 
-- Title: **.
-- Exposures: .
-- Medium exposure assessment: .
-- Timing exposure assessment: .
-- Outcomes: .
-- Medium outcome assessment: .
-- Timing outcome assessment: .
+- Title: *Associations between repeated measures of maternal urinary phthalate metabolites during pregnancy and cord blood glucocorticoids*.
+- Exposures: phthalate metabolites.
+- Medium exposure assessment: spot urine samples.
+- Timing exposure assessment: 14, 24, and 36 weeks of gestation.
+- Outcomes: cortisol and cortisone.
+- Medium outcome assessment: cord blood.
+- Timing outcome assessment: 14, 24, and 36 weeks of gestation.
 - Results:
-    - .
-- Sex-specific results: .
-
-TODO:
-
-- Kim et al. 2018
-- Sun et al. 2018
+    - Positive association of MEHP 3rd trimester with cortisol.
+    - Inverse association of MECPP and MEOHP 3rd trimester with cortisone.
+    - Positive association of MBzP 1st trimester with cortisol/cortisone ration.
+- Sex-specific results: evidence of modification by sex, but not significant.
+    - Females: positive associations with glucocorticoids (MEHP and MBzP 1st trimester, and DEHP 3rd trimester).
+    - Males: inverse associations with glucocorticoids (MECPP, MEHHP, and MEOHP 3rd trimester).
 
@@
 # Glucocorticosteroids and neurodevelopment
+
+## @LupienMcEwenGunnar:2009
+
+- Title: *Effects of stress throughout the lifespan on the brain, behaviour and cognition*.
+
+The HPA axis, which can be activated by stress, is responsible for the production of glucocorticoids.
+The brain, and its proper functioning, is a potential target, due to the presence of receptors for these hormones.
+Glucocorticoids are necessary for brain maturation, although their under- or over-production might interfere with its normal development and ultimately lead to long-term impaired functioning.
 

### 2024-01-01 output/paper/_figures.qmd
Added YML file for paper-specific authors and added some authors. Modyfing tables so that match style of AJE. Added abbreviations EDCs and steroids in text paper. Extending labels tabels.

@@
 
-## Marginal comparisons
+## Marginal contrasts
 
@@
 #| label: fig-marginal-1
-#| fig-cap: Marginal comparison results for $hitrtse \sim edc$
+#| fig-cap: Marginal contrast results for $hitrtse \sim edc$
 #| fig-width: 9
@@
 #| label: fig-marginal-2
-#| fig-cap: Marginal comparison results for $production \sim edc$
+#| fig-cap: Marginal contrast results for $production \sim edc$
 #| fig-width: 9
@@
 #| label: fig-marginal-3
-#| fig-cap: Marginal comparison results for $hitrtse \sim metabolite$
+#| fig-cap: Marginal contrast results for $hitrtse \sim metabolite$
 #| fig-width: 9

### 2024-01-01 output/paper/_si.qmd
Added YML file for paper-specific authors and added some authors. Modyfing tables so that match style of AJE. Added abbreviations EDCs and steroids in text paper. Extending labels tabels.

@@
 ```
-Lower limits of quantification for the corticosteroids
+Lower limits of quantification for the glucocorticosteroids
 :::
@@
 ```
-Classification of the measured corticosteroids, by cohort.
+Classification of the measured glucocorticosteroids, by cohort.
 :::
@@
 
-### Marginal comparisons for effect modification
+### Marginal contrasts for effect modification
 
@@
 ```
-Marginal comparison results for $hitrtse \sim edc \times sex$
+Marginal contrast results for $hitrtse \sim edc \times sex$
 :::
@@
 ```
-Marginal comparison results for $metabolite \sim edc \times sex$
+Marginal contrast results for $metabolite \sim edc \times sex$
 :::
@@
 ```
-Marginal comparison results for $hitrtse \sim metabolite \times sex$
+Marginal contrast results for $hitrtse \sim metabolite \times sex$
 :::

### 2024-01-01 output/paper/_tables.qmd
Added YML file for paper-specific authors and added some authors. Modyfing tables so that match style of AJE. Added abbreviations EDCs and steroids in text paper. Extending labels tabels.

@@
 #| label: tbl-pop-desc
-#| tbl-cap: Characteristics of the children included in the analyses, by cohort and overall
+#| tbl-cap: Participant characteristics, by cohort and overall (HELIX Study; 2013-2016).
 
@@
 #| label: tbl-edc-desc
-#| tbl-cap: Levels of unprocessed chemicals, by cohort and overall
+#| tbl-cap: Participants endocrine disruptors concentrations, by cohort and overall (HELIX Study; 2013-2016).
 
@@
 
-## Corticosteroids
+## Glucocorticosteroids
 
@@
 #| label: tbl-met-desc
-#| tbl-cap: Levels of unprocessed corticosteroids, by cohort and overall
+#| tbl-cap: Participants glucocorticosteroids concentrations, by cohort and overall (HELIX Study; 2013-2016).
 
@@
 #| label: tbl-hypothesis-1
-#| tbl-cap: Hypothesis testing results of effect modification by sex for $hitrtse \sim edc$
+#| tbl-cap: Hypothesis testing results of effect modification by sex for $hitrtse \sim edc$.
 
@@
 #| label: tbl-hypothesis-2
-#| tbl-cap: Hypothesis testing results of effect modification by sex for $metabolite \sim edc$
+#| tbl-cap: Hypothesis testing results of effect modification by sex for $metabolite \sim edc$.
 
@@
 #| label: tbl-hypothesis-3
-#| tbl-cap: Hypothesis testing results of effect modification by sex for $hitrtse \sim metabolite$
+#| tbl-cap: Hypothesis testing results of effect modification by sex for $hitrtse \sim metabolite$.
 

### 2024-01-01 output/paper/manuscript.qmd
Added YML file for paper-specific authors and added some authors. Modyfing tables so that match style of AJE. Added abbreviations EDCs and steroids in text paper. Extending labels tabels.

@@
 ---
-title: "Estimating associations between childhood exposure to non-persistent endocrine disruptors, corticosteroids, and neurodevelopment: A study based on the parametric g-formula"
+title: "Childhood exposure to non-persistent endocrine disruptors, glucocorticosteroids, and neurodevelopment: A study based on the parametric g-formula"
 
 # metadata-files:
-#   - ../../../../all_phd/authors.yml
+#   - ../../docs/authors.yml
 
@@
 The prevalence of several neurodevelopmental disorders has increased in the pediatric population [@GrandjeanLandrigan:2014], and multiple environmental pollutants play a role in the increased rates of these disorders [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
-\acr{edcs}, chemicals that are capable of interfering with the endocrine system, have been shown to have a role in childhood neurodevelopment [@Braun:2017].
+\acr{edcs}, chemicals that are capable of interfering with the endocrine system, have been shown to have a role in childhood neurodevelopment and behavior [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @HuangChenSu:2015; @CartierWarembourgLeManer-Idrissi:2016; @TewarAuingerBraun:2016; @YuDuChiou:2016; @HuangTsaiChen:2017; @KimHongShin:2017; @FurlongHerringBuckley:2017; @Braun:2017; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019; @ShoaffCoullWeuve:2020; @OhKimKannan:2023; @VilmandBeckBilenberg:2023].
 Although both pregnancy and early childhood are crucial stages of (neuro)development, most of the available literature is focused on the effects of prenatal exposure to \acr{edcs} on child neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
@@
 
-There is furthermore evidence that exposure to certain \acr{edcs}, specifically phthalates, might interfere with the \acr{hpa} axis and might interact with the glucocorticoid receptor [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023].
+There is also evidence that exposure to certain \acr{edcs}, specifically phthalates, might interfere with the \acr{hpa} axis and might interact with the glucocorticoid receptor [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023].
 
-Finally, the development of the fetal brain is regulated by, among others, glucocorticoids [@SunLiJin:2018], with regulation of their homeostasis being crucial for neurodevelopment in childhood [@SearsLiuLanphear:2023].
+The HPA axis, which can be activated by stress, is responsible for the production of glucocorticosteroids.
+The brain, and its proper functioning, is a potential target, due to the presence of receptors for these hormones [@LupienMcEwenGunnar:2009; @SunLiJin:2018].
+Glucocorticosteroids are necessary for brain maturation, although their under- or over-production might interfere with its normal development and ultimately lead to long-term impaired functioning [@LupienMcEwenGunnar:2009; @SearsLiuLanphear:2023].
 
 Taken together, these results suggest that the negative influence of exposure to certain \acr{edcs} on neurodevelopmental outcomes, might be mediated, at least partially, by disruption of the \acr{hpa} axis' homeostasis.
-In the present study, we thus estimated associations between 1) non-persistent \acr{edcs} and attention, 2) non-persistent \acr{edcs} and corticosteroids, and 3) corticosteroids and attention, using the parametric g-formula and average contrasts, in children of a large prospective birth cohort in Europe.
+In the present study, we thus estimated associations between 1) non-persistent \acr{edcs} and attention, 2) non-persistent \acr{edcs} and glucocorticosteroids, and 3) glucocorticosteroids and attention, using the parametric g-formula and marginal contrasts, in children of a large prospective birth cohort in Europe.
 
@@
 
+```{r}
+info_edcs <- myphd::edcs_information()
+print_tidy_abbreviations_edcs <- function(df, class_) {
+  tmp <- df |>
+    dplyr::filter(class == class_) |>
+    dplyr::select(chem_id, full_name) |>
+    dplyr::arrange(full_name) |>
+    dplyr::mutate(
+      ret = glue::glue("{full_name} ({chem_id})")
+    )
+  tmp$ret |>
+    paste0(collapse = ", ")
+}
+```
+
 Children were assessed between December 2013 and February 2016, and included neurological testing and urine collection.
@@
 A list of the environmental chemicals determined in urine samples and used for the present study is given in @supptbl-info-chems.
-Briefly, we analyzed a total of 7 phenols, 4 non-specific organophosphate pesticide metabolites, and 10 phthalate metabolites originating from 6 distinct phthalate parent compounds.
+Briefly, we analyzed a total of 7 phenols (`r print_tidy_abbreviations_edcs(info_edcs, "Phenols")`), 6 non-specific organophosphate pesticide metabolites (`r print_tidy_abbreviations_edcs(info_edcs, "OP pesticide metabolites")`), and 10 phthalate metabolites (`r print_tidy_abbreviations_edcs(info_edcs, "Phthalate metabolites")`) originating from 6 distinct phthalate parent compounds.
 The laboratory protocols for the analysis are described elsewhere [@HaugSakhiCequier:2018].
 
-### Corticosteroids {#sec-steroids}
+### Glucocorticosteroids {#sec-steroids}
 
-Urine samples of the night before the day of the visit were used to measure levels of the corticosteroids.
+```{r}
+info_steroids <- readODS::read_ods("docs/steroids.ods")
+print_tidy_abbreviations_steroids <- function(df, score) {
+  cortisol_prod <- c(
+    "20aDHF", "20bDHF", "5bDHF", "5aTHF", "5bTHF",
+    "6OHF",
+    "5a20acortol", "5a20bcortol", "5b20acortol", "5b20bcortol"
+  )
+  cortisone_prod <- c(
+    "20aDHE", "20bDHE", "5aTHE", "5bTHE",
+    "6OHE",
+    "5b20acortolone", "5b20bcortolone"
+  )
+  corticosterone_prod <- c(
+    "5aTHB", "5bTHB", "A", "17-DO-cortolone"
+  )
+  tmp <- df |>
+    dplyr::filter(dplyr::case_when(
+      score == "cortisol_prod" ~ acronym %in% cortisol_prod,
+      score == "cortisone_prod" ~ acronym %in% cortisone_prod,
+      score == "corticosterone_prod" ~ acronym %in% corticosterone_prod,
+    )) |>
+    dplyr::select(analyte, acronym) |>
+    dplyr::arrange(analyte) |>
+    dplyr::mutate(
+      ret = glue::glue("{analyte} ({acronym})")
+    )
+  tmp$ret |>
+    paste0(collapse = ", ")
+}
+```
+
+Urine samples of the night before the day of the visit were used to measure levels of the glucocorticosteroids.
 These included glucocorticosteroids, glucocorticosteroid metabolites, glucocorticosteroid precursors, glucocorticosteroid precursor metabolites, androgens, and androgen metabolites.
-A list of the corticosteroids determined in urine samples and used for the present study is given in @supptbl-info-mets.
+A list of the glucocorticosteroids determined in urine samples and used for the present study is given in @supptbl-info-mets.
 
-To assess the levels of corticosteroids and their metabolites, LC-MS/MS analysis was applied at the Applied Metabolomics Research Group, IMIM (Hospital del Mar Medical Research Institute).
+To assess the levels of glucocorticosteroids and their metabolites, LC-MS/MS analysis was applied at the Applied Metabolomics Research Group, IMIM (Hospital del Mar Medical Research Institute).
 The laboratory protocols for the analysis are described elsewhere [@MarcosRenauCasals:2014; @Gomez-GomezPozo:2020].
@@
 
-Three additional markers, cortisol production, cortisol metabolism, cortisone production, and 11bHSD activity, were computed based on the following: cortisol production as the sum of cortisol and its metabolites (20aDHF, 20bDHF, 5bDHF, 5aTHF, 5bTHF, 6OHF, 5a20acortol, 5a20bcortol, 5b20acortol, 5b20bcortol), cortisol metabolism as the inverse of the ratio between cortisol and its metabolites, cortisone production as the sum of cortisone and its metabolites (20aDHE, 20bDHE, 5aTHE, 5bTHE, 6OHE, 5b20acortolone, b20bcortolone), and 11bHSD activity as the ratio between cortisone production and cortisol production.
-11bHSD activity gives a measure of conversion of cortisone to cortisol.
+Three additional markers, total cortisol production, total cortisone production, and total corticosterone production, were computed based on the following: cortisol production as the sum of cortisol and its metabolites (`r print_tidy_abbreviations_steroids(info_steroids, "cortisol_prod")`), cortisone production as the sum of cortisone and its metabolites (`r print_tidy_abbreviations_steroids(info_steroids, "cortisone_prod")`), and corticosterone production as the sum of `r print_tidy_abbreviations_steroids(info_steroids, "corticosterone_prod")`.
 
@@
 
-Concentrations of the corticosteroids were classified as quantifiable, below the \acr{loq}, possible interference or out of range, and not detected.
+Concentrations of the glucocorticosteroids were classified as quantifiable, below the \acr{loq}, possible interference or out of range, and not detected.
 For each metabolite, we computed the fraction of values below the \acr{loq} and not detected, both within each cohort and overall.
 We proceeded to impute these values using half the value of the corresponding \acr{lloq}, for those metabolites that had less than `r threshold_within_lod`% of missings within each cohort and `r threshold_overall_lod`% of missings overall.
-Information about the \acr{lloq} for the corticosteroids is provided in @supptbl-lloq-mets.
+Information about the \acr{lloq} for the glucocorticosteroids is provided in @supptbl-lloq-mets.
 The remaining missing values were imputed using kNN from the `VIM` R package [@KowarikTempl:2016], for those metabolites that had less than `r threshold_within_nan`% of remaining missings within each cohort and `r threshold_overall_nan`% of remaining missings overall.
@@
 
-We estimated average contrasts with the parametric g-formula, a method of standardization.
+We estimated marginal contrasts with the parametric g-formula, a method of standardization.
 The parametric g-formula involves the following steps: 1) fit a outcome model including both covariates and balancing weights; 2) create two new datasets identical to the original one but with the exposure shifted according to a user-specified intervention set by a deterministic function of the observed exposure levels; 3) use the outcome model to compute adjusted predictions in the two counterfactual datasets; 4) compute the difference between the means of the adjusted predictions in the counterfactual datasets.
@@
 
-To estimate the average contrasts, we used the `avg_comparisons` function from the `marginaleffects` R package [@Arel-Bundock:2023].
+To estimate the marginal contrasts, we used the `avg_comparisons` function from the `marginaleffects` R package [@Arel-Bundock:2023].
 The two counterfactual datasets were obtained by setting the exposures levels to 90th percentile ($d_1$) and the 10th percentile ($d_2$), for each cohort separately.
-The average contrasts were computed using the estimated balancing weights above.
+The marginal contrasts were computed using the estimated balancing weights above.
 Robust standard errors were computed with the `sandwich` R package, using cohort as variable indicating clustering of observations [@Zeileis:2004; @ZeileisKollGraham:2020].
-For each outcome, we report the results as differences between average contrasts.
+For each outcome, we report the results as differences between marginal contrasts.
 
 We further estimated the \acr{adrf} using the `avg_predictions` function from the `marginaleffects` R package, examining 50 exposure values from the 10th to the 90th percentiles of the exposure.
-As done for the average contrasts, we included the estimated balancing weights and used cohort as a clustering variable when computing robust standard errors.
+As done for the marginal contrasts, we included the estimated balancing weights and used cohort as a clustering variable when computing robust standard errors.
 
@@
 To do so, balancing weights were estimated separately for each level of the sex variable, and an interaction term between the exposure and sex was included in the outcome model.
-Similarly, the average contrasts were aggregated separately for each level of sex.
-We further tested for significance of the difference between the average contrasts of males and females.
+Similarly, the marginal contrasts were aggregated separately for each level of sex.
+We further tested for significance of the difference between the marginal contrasts of males and females.
 
@@
 In total, 1,297 children were included in the dataset.
-Of these, 976 had information also about the corticosteroids.
+Of these, 976 had information also about the glucocorticosteroids.
 The sample consisted of 55% girls.
@@
 
-Levels of unprocessed non-persistent \acr{edcs} and corticosteroids, by cohort and overall, are presented in @tbl-edc-desc and @tbl-met-desc, respectively.
+Levels of unprocessed non-persistent \acr{edcs} and glucocorticosteroids, by cohort and overall, are presented in @tbl-edc-desc and @tbl-met-desc, respectively.
 
@@
 
-@Fig-marginal-1 presents the forest plot for the marginal effects of the non-persistent \acr{edcs} on \acr{hitrtse}.
+@Fig-marginal-1 presents the forest plot for the marginal contrasts of the non-persistent \acr{edcs} on \acr{hitrtse}.
 Overall, a cohort-specific increase in the levels of the exposures from the 10th to the 90th percentiles was associated with a positive marginal contrast, indicating an increase in the values of \acr{hitrtse}.
 Most of the \acr{ci}s included the null effect, though.
-Statistically significant associations were observed for the phenol MEPA (`r format_res(meffects1, "mepa", "hitrtse")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "hitrtse")`), oh-MiNP (`r format_res(meffects1, "ohminp", "hitrtse")`), and MEHP (`r format_res(meffects1, "mehp", "hitrtse")`).
+Statistically significant effects were observed for the phenol MEPA (`r format_res(meffects1, "mepa", "hitrtse")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "hitrtse")`), oh-MiNP (`r format_res(meffects1, "ohminp", "hitrtse")`), and MEHP (`r format_res(meffects1, "mehp", "hitrtse")`).
 The \acr{opp} DETP was strongly negatively associated with \acr{hitrtse} (`r format_res(meffects1, "detp", "hitrtse")`).
 
-@Fig-marginal-2 presents the forest plot for the marginal effects of the non-persistent \acr{edcs} on cortisone production, cortisol production, and corticosterone production.
+@Fig-marginal-2 presents the forest plot for the marginal contrasts of the non-persistent \acr{edcs} on cortisone production, cortisol production, and corticosterone production.
 Overall, a cohort-specific increase in the levels of the exposures from the 10th to the 90th percentiles was associated with a positive marginal contrast, indicating an increase in the total production of these metabolites, with the exception of BUPA, which was associated with negative marginal contrasts for all three outcomes, and MiBP, which was associated with a negative marginal contrast for total cortisone production only.
-The majority of the associations for the phenols and phthalate metabolites included the null.
+The majority of the effects for the phenols and phthalate metabolites included the null.
 The phenol BPA showed the largest marginal contrasts across all three outcomes (cortisone production, `r format_res(meffects2, "bpa", "cortisone production")`; cortisol production, `r format_res(meffects1, "bpa", "cortisol production")`; corticosterone production, `r format_res(meffects1, "bpa", "corticosterone production")`).
 
-@Fig-marginal-3 presents the forest plot for the marginal effects of the corticosteroids on \acr{hitrtse}.
+@Fig-marginal-3 presents the forest plot for the marginal contrasts of the glucocorticosteroids on \acr{hitrtse}.
 All marginal contrasts included the null, with no clear indication of directionality of the effect.
@@
 In this study, consisting of almost 1,000 children from 6 European birth cohorts, we observed that short-term childhood exposure to non-persistent \acr{edcs} had negative effects on \acr{hitrtse} and total production of cortisol, cortisone, and corticosterone, with the majority of the \acr{ci}s including the null effect.
-Increased production of these glucocorticoids did not seem to affect \acr{hitrtse}.
+Increased production of these glucocorticosteroids did not seem to affect \acr{hitrtse}.
 Some of these effects differed for females and males, including significant differences for the effects of increased production of cortisone and corticosterone on \acr{hitrtse}.
-Specifically, an increased production of these glucocorticoids was associated with lower values of \acr{hitrtse} for females, and higher values for males.
-Taken together, these results suggest that these non-persistent \acr{edcs} might be responsible for perturbations of the \acr{hpa} axis' homeostasis, and that higher levels of these glucocorticoids might interfere with different functions of attention in a sex-specific manner.
+Specifically, an increased production of these glucocorticosteroids was associated with lower values of \acr{hitrtse} for females, and higher values for males.
+Taken together, these results suggest that these non-persistent \acr{edcs} might be responsible for perturbations of the \acr{hpa} axis' homeostasis, and that higher levels of these glucocorticosteroids might interfere with different functions of attention in a sex-specific manner.
 
@@
 Contrary to these prior studies, we did find effect modification by sex.
-We are not aware of other epidemiological studies investigating phthalates metabolites, phenols, and \acr{opps}, in relation to urinary glucocorticoids in childhood.
-We are also not aware of prior studies specifically investigating the effects of exposure to \acr{edcs} and elevated levels of glucocorticoids in relation to \acr{hitrtse}.
+We are not aware of other epidemiological studies investigating phthalates metabolites, phenols, and \acr{opps}, in relation to urinary glucocorticosteroids in childhood.
+We are also not aware of prior studies specifically investigating the effects of exposure to \acr{edcs} and elevated levels of glucocorticosteroids in relation to \acr{hitrtse}.
 
@@
 Limitations include the cross-sectional design of the present study.
-Importantly, the non-persistent \acr{edcs} were measured in a pool of night and morning urine samples before the clinical visit, whereas the corticosteroids were measured in the night urine sample only.
+Importantly, the non-persistent \acr{edcs} were measured in a pool of night and morning urine samples before the clinical visit, whereas the glucocorticosteroids were measured in the night urine sample only.
 Although we included a wide range of confounders there is the possibility, as with other observational studies, of residual confounding, which might lead to a bias away from the null.
@@
 We decided to model both the *treatment* mechanisms, for the estimation of balancing weights, and the outcomes, with traditional covariates adjustment, to try to obtain *doubly robust* effect estimates.
-Finally, we decided not to interpret our results by focusing on the estimated coefficients of the regression models, but by making use of the g-computation procedure and estimate average comparisons.
+Finally, we decided not to interpret our results by focusing on the estimated coefficients of the regression models, but by making use of the g-computation procedure and estimate marginal contrasts.
 
@@
 
-{{< include _figures.qmd >}}
+<!-- {{< include _figures.qmd >}} -->
 

### 2024-01-02 output/paper/_figures.qmd
Adapting tables and figures to style AJE.

--- 2023-12-19 output/paper/_figures.qmd
@@
 
-- Figure titles and legends should be submitted as single paragraph.
-
 # Figures for main results
 
-## Marginal comparisons
+## Marginal contrasts
 
@@
 
+::: {#fig-marginal-1}
 ```{r marginal-1}
-#| label: fig-marginal-1
-#| fig-cap: Marginal comparison results for $hitrtse \sim edc$
 #| fig-width: 9
@@
 ```
+**Marginal contrasts (exposures: \acr{edcs}; outcome: \acr{hitrtse}) (HELIX Study; 2013-2016).**
+Circles indicate effect estimates. Solid lines indicate the $95\%$ \acr{ci}.
+:::
 
+::: {#fig-marginal-2}
 ```{r marginal-2}
-#| label: fig-marginal-2
-#| fig-cap: Marginal comparison results for $production \sim edc$
 #| fig-width: 9
@@
 ```
+**Marginal contrasts (exposures: \acr{edcs}; outcomes: glucocorticosteroids) (HELIX Study; 2013-2016).**
+Circles, triangles, and squares indicate effect estimates. Solid lines indicate the $95\%$ \acr{ci}.
+:::
 
+::: {#fig-marginal-3}
 ```{r marginal-3}
-#| label: fig-marginal-3
-#| fig-cap: Marginal comparison results for $hitrtse \sim metabolite$
 #| fig-width: 9
@@
 ```
+**Marginal contrasts (exposures: glucocorticosteroids; outcome: \acr{hitrtse}) (HELIX Study; 2013-2016).**
+Circles indicate effect estimates. Solid lines indicate the $95\%$ \acr{ci}.
+:::

### 2024-01-02 output/paper/_si.qmd
Adapting tables and figures to style AJE.

@@
 ```
-Codebook for covariates used in $hitrtse \sim edc$
+Codebook for the covariates used in the estimation of the marginal comparisons of \acr{edcs} on \acr{hitrtse}.
 :::
@@
 ```
-Codebook for covariates used in $metabolite \sim edc$
+Codebook for the covariates used in the estimation of the marginal comparisons of \acr{edcs} on the glucocorticosteroids.
 :::
@@
 ```
-Codebook for covariates used in $hitrtse \sim metabolite$
+Codebook for the covariates used in the estimation of the marginal comparisons of the glucocorticosteroids on \acr{hitrtse}.
 :::
@@
 ```
-Information about non-persistent EDCs
+Information about non-persistent EDCs, including the full compound name, the standard symbol, the used variable name, the identifier from PubChem, and the parental compound.
 :::
@@
   gt::tab_footnote(
-    footnote = "Human Metabolome Database",
-    locations = gt::cells_column_labels(
-      columns = hmdb_id
-    )
+    footnote = "Abbreviations: Human Metabolome Database (HMDB); Chemical Abstracts Service (CAS).",
   ) |>
-  gt::tab_footnote(
-    footnote = "Chemical Abstracts Service",
-    locations = gt::cells_column_labels(
-      columns = cas_number
-    )
+  gt::opt_footnote_marks(
+    marks = "letters"
   )
 ```
-Information about metabolites
+Information about the glucocorticosteroids, including the full metabolite name, the standard symbol, the identifier from the HMDB, and the CAS number.
 :::
@@
   tidylog::mutate(
-    metabolite = stringr::str_replace(
+    Metabolite = stringr::str_replace(
       metabolite,
@@
   tidylog::rename(
-    lloq = "loq"
-  )
-gt::gt(lloq_mets |> dplyr::arrange(metabolite)) |>
+    LLOQ = "loq"
+  ) |>
+  tidylog::select(-metabolite)
+gt::gt(lloq_mets |> dplyr::arrange(Metabolite)) |>
   gt::tab_footnote(
-    footnote = "Lower limit of quantification expressed in nanograms per millilitre.",
-    locations = gt::cells_column_labels(
-      columns = lloq
-    )
+    footnote = "Abbreviations: lower limit of quantification (LLOQ), expressed in nanograms per millilitre."
+  ) |>
+  gt::opt_footnote_marks(
+    marks = "letters"
   )
 ```
-Lower limits of quantification for the glucocorticosteroids
+Lower limits of quantification for the glucocorticosteroids.
 :::
@@
   gt::tab_footnote(
-    footnote = "N.",
-    locations = gt::cells_column_labels(
-      columns = c(Unadjusted, Adjusted)
-    )
-  ) |>
-  gt::tab_footnote(
     footnote = "Truncated weights.",
     locations = gt::cells_column_labels(columns = Adjusted)
+  ) |>
+  gt::opt_footnote_marks(
+    marks = "letters"
   )
 ```
-Effective sample size before and after weights estimation for $hitrtse \sim edc$
+Effective sample size before and after balancing weights estimation (exposures: \acr{edcs}; outcome: \acr{hitrtse}).
 :::
@@
   gt::tab_footnote(
-    footnote = "N.",
-    locations = gt::cells_column_labels(
-      columns = c(Unadjusted, Adjusted)
-    )
-  ) |>
-  gt::tab_footnote(
     footnote = "Truncated weights.",
     locations = gt::cells_column_labels(columns = Adjusted)
+  ) |>
+  gt::opt_footnote_marks(
+    marks = "letters"
   )
 ```
-Effective sample size before and after weights estimation for $metabolite \sim edc$
+Effective sample size before and after balancing weights estimation (exposures: \acr{edcs}; outcomes: glucocorticosteroids).
 :::
@@
   gt::tab_footnote(
-    footnote = "N.",
-    locations = gt::cells_column_labels(
-      columns = c(Unadjusted, Adjusted)
-    )
-  ) |>
-  gt::tab_footnote(
     footnote = "Truncated weights.",
     locations = gt::cells_column_labels(columns = Adjusted)
+  ) |>
+  gt::opt_footnote_marks(
+    marks = "letters"
   )
 ```
-Effective sample size before and after weights estimation for $hitrtse \sim metabolite$
+Effective sample size before and after balancing weights estimation (exposures: glucocorticosteroids; outcome: \acr{hitrtse}).
 :::
@@
 ```
-Summary statistics of the estimated balancing weights for $hitrtse \sim edc$
+Summary statistics of the estimated balancing weights (exposures: \acr{edcs}; outcome: \acr{hitrtse}).
 :::
@@
 ```
-Summary statistics of the estimated balancing weights for $metabolite \sim edc$
+Summary statistics of the estimated balancing weights (exposures: \acr{edcs}; outcomes: glucocorticosteroids).
 :::
@@
 ```
-Summary statistics of the estimated balancing weights for $hitrtse \sim metabolite$
+Summary statistics of the estimated balancing weights (exposures: glucocorticosteroids; outcome: \acr{hitrtse}).
 :::
@@
 ```
-Summary statistics of the estimated balancing weights for $hitrtse \sim edc \times sex$
+Summary statistics of the estimated balancing weights for effect modification (exposures: \acr{edcs}; outcome: \acr{hitrtse}; modifier: sex).
 :::
@@
 ```
-Summary statistics of the estimated balancing weights for $metabolite \sim edc \times sex$
+Summary statistics of the estimated balancing weights for effect modification (exposures: \acr{edcs}; outcomes: glucocorticosteroids; modifier: sex).
 :::
@@
 ```
-Summary statistics of the estimated balancing weights for $hitrtse \sim metabolite \times sex$
+Summary statistics of the estimated balancing weights for effect modification (exposures: glucocorticosteroids; outcome: \acr{hitrtse}; modifier: sex).
 :::
@@
 ```
-Classification of the measured non-persistent EDCs, by cohort.
+**Measurement classification of \acr{edcs}, by cohort.**
+Coding: 1, quantifiable; 2, <LOD; 3, interference or out of range; 4. not analysed.
 :::
@@
 ```
-Classification of the measured glucocorticosteroids, by cohort.
+**Measurement classification of the glucocorticosteroids, by cohort.**
+Coding: 1, quantifiable; 2, <LOQ; 3, interference or out of range; 4, not detected.
 :::
@@
 ```
-Marginal contrast results for $hitrtse \sim edc \times sex$
+**Marginal contrasts for effect modification (exposures: \acr{edcs}; outcome: \acr{hitrtse}; modifier: sex) (HELIX Study; 2013-2016).**
+Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ \acr{ci}.
 :::
@@
 ```
-Marginal contrast results for $metabolite \sim edc \times sex$
+**Marginal contrasts for effect modification (exposures: \acr{edcs}; outcomes: glucocorticosteroids; modifier: sex) (HELIX Study; 2013-2016).**
+Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ \acr{ci}.
 :::
@@
 ```
-Marginal contrast results for $hitrtse \sim metabolite \times sex$
+**Marginal contrasts for effect modification (exposures: glucocorticosteroids; outcome: \acr{hitrtse}; modifier: sex) (HELIX Study; 2013-2016).**
+Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ \acr{ci}.
 :::

### 2024-01-02 output/paper/_tables.qmd
Adapting tables and figures to style AJE.

@@
 ```
-
-**Table titles should give details on the place of the study, the time of the study, and the study population (if applicable).
-Example, “Baseline Characteristics of Infants With Initial and Follow-up Screening, London, United Kingdom, 2001–2003.” In the table body, leave blank spaces for no entry; avoid using dashes.
-Order of footnotes: 1) Abbreviations: (no footnote symbol, listed alphabetically, separated by semicolons); 2) other footnotes as necessary, each preceded by a superscript lowercase letter.
-Footnotes must be cited in order when reading from left to right and top to bottom.**
 
@@
 #| label: tbl-hypothesis-1
-#| tbl-cap: Hypothesis testing results of effect modification by sex for $hitrtse \sim edc$.
+#| tbl-cap: Pairwise differences between sex-specific marginal contrasts for the effect of \acr{edcs} on \acr{hitrtse} (HELIX Study; 2013-2016).
 
@@
 #| label: tbl-hypothesis-2
-#| tbl-cap: Hypothesis testing results of effect modification by sex for $metabolite \sim edc$.
+#| tbl-cap: Pairwise differences between sex-specific marginal contrasts for the effect of \acr{edcs} on the glucocorticosteroids (HELIX Study; 2013-2016).
 
@@
 #| label: tbl-hypothesis-3
-#| tbl-cap: Hypothesis testing results of effect modification by sex for $hitrtse \sim metabolite$.
+#| tbl-cap: Pairwise differences between sex-specific marginal contrasts for the effect of the glucocorticosteroids on \acr{hitrtse} (HELIX Study; 2013-2016).
 

### 2024-01-02 R/proc_res.R

@@ -80,6 +80,9 @@ tidy_codebooks <- function(rq) {
         "Percentage of confounders included in the models: {perc_included}%."
       ),
       locations = gt::cells_column_labels(columns = included)
+    ) |>
+    gt::opt_footnote_marks(
+      marks = "letters"
     )
   
   return(ret)
@@ -241,13 +244,61 @@ tbl_desc_vars <- function() {
   #   filename = paste0(path, name)
   # )
   
+  edcs <- myphd::edcs_information()
   desc_chems <- rq02_desc_data_exp$step3 |>
     gtsummary::as_gt() |>
+    gt::tab_row_group(
+      label = "OP pesticide metabolites",
+      rows = variable %in% as.character(edcs[edcs$class == "OP pesticide metabolites", ]$chem_id)
+    ) |>
+    gt::tab_row_group(
+      label = "Phenols",
+      rows = variable %in% as.character(edcs[edcs$class == "Phenols", ]$chem_id)
+    ) |>
+    gt::tab_row_group(
+      label = "Phthalate metabolites",
+      rows = variable %in% as.character(edcs[edcs$class == "Phthalate metabolites", ]$chem_id)
+    ) |>
+    gt::row_group_order(
+      groups = c("OP pesticide metabolites",
+                 "Phenols",
+                 "Phthalate metabolites")
+    ) |>
+    gt::tab_style(
+      locations = gt::cells_row_groups(group = dplyr::everything()),
+      style = list(gt::cell_text(weight = "bold"))
+    ) |>
     gt::opt_footnote_marks(
       marks = "letters"
     )
+  
+  steroids <- readODS::read_ods("docs/steroids.ods")
   desc_mets <- rq02_desc_data_out$step3 |>
-    gtsummary::as_gt() |>
+    gtsummary::as_gt()
+  for (x in unique(steroids$type)) {
+    desc_mets <- desc_mets |>
+      gt::tab_row_group(
+        label = x,
+        rows = variable %in% as.character(
+          steroids[steroids$type == x, ]$acronym
+        )
+      )
+  }
+  desc_mets <- desc_mets |>
+    gt::tab_style(
+      locations = gt::cells_row_groups(group = dplyr::everything()),
+      style = list(gt::cell_text(weight = "bold"))
+    ) |>
+    gt::row_group_order(
+      groups = c(
+        "Glucocorticosteroid",
+        "Glucocorticosteroid metabolite",
[…]

### 2024-01-02 output/paper/_si.qmd
First round of modifications for tables and figures done.

@@
 ```
-Codebook for the covariates used in the estimation of the marginal comparisons of \acr{edcs} on \acr{hitrtse}.
+**Codebook for the covariates used in the estimation of the marginal comparisons of \acr{edcs} on \acr{hitrtse}.**
 :::
@@
 ```
-Codebook for the covariates used in the estimation of the marginal comparisons of \acr{edcs} on the glucocorticosteroids.
+**Codebook for the covariates used in the estimation of the marginal comparisons of \acr{edcs} on the glucocorticosteroids.**
 :::
@@
 ```
-Codebook for the covariates used in the estimation of the marginal comparisons of the glucocorticosteroids on \acr{hitrtse}.
+**Codebook for the covariates used in the estimation of the marginal comparisons of the glucocorticosteroids on \acr{hitrtse}.**
 :::
@@
 ```
-Information about non-persistent EDCs, including the full compound name, the standard symbol, the used variable name, the identifier from PubChem, and the parental compound.
+**Information about non-persistent EDCs, including the full compound name, the standard symbol, the used variable name, the identifier from PubChem, and the parental compound.**
 :::
@@
 ```
-Information about the glucocorticosteroids, including the full metabolite name, the standard symbol, the identifier from the HMDB, and the CAS number.
+**Information about the glucocorticosteroids, including the full metabolite name, the standard symbol, the identifier from the HMDB, and the CAS number.**
 :::
@@
   ) |>
-  tidylog::select(-metabolite)
+  tidylog::select(-metabolite, Metabolite, LLOQ)
 gt::gt(lloq_mets |> dplyr::arrange(Metabolite)) |>
@@
 ```
-Lower limits of quantification for the glucocorticosteroids.
+**Lower limits of quantification for the glucocorticosteroids (HELIX Study; 2013-2016).**
 :::
@@
 balance_1 |>
+  tidylog::rename(Exposure = "exposure") |>
   tidylog::group_by(class) |>
@@
     marks = "letters"
+  ) |>
+  gt::tab_style(
+    style = gt::cell_text(
+      weight = "bold"
+    ),
+    locations = gt::cells_row_groups()
   )
 ```
-Effective sample size before and after balancing weights estimation (exposures: \acr{edcs}; outcome: \acr{hitrtse}).
+**Effective sample size before and after balancing weights estimation (exposures: \acr{edcs}; outcome: \acr{hitrtse}) (HELIX Study; 2013-2016).**
 :::
@@
     marks = "letters"
+  ) |>
+  gt::tab_style(
+    style = gt::cell_text(
+      weight = "bold"
+    ),
+    locations = gt::cells_row_groups()
   )
 ```
-Effective sample size before and after balancing weights estimation (exposures: \acr{edcs}; outcomes: glucocorticosteroids).
+**Effective sample size before and after balancing weights estimation (exposures: \acr{edcs}; outcomes: glucocorticosteroids) (HELIX Study; 2013-2016).**
 :::
@@
     marks = "letters"
+  ) |>
+  gt::tab_style(
+    style = gt::cell_text(
+      weight = "bold"
+    ),
+    locations = gt::cells_row_groups()
   )
 ```
-Effective sample size before and after balancing weights estimation (exposures: glucocorticosteroids; outcome: \acr{hitrtse}).
+**Effective sample size before and after balancing weights estimation (exposures: glucocorticosteroids; outcome: \acr{hitrtse}) (HELIX Study; 2013-2016).**
 :::
@@
 ```
-Summary statistics of the estimated balancing weights (exposures: \acr{edcs}; outcome: \acr{hitrtse}).
+**Summary statistics of the estimated balancing weights (exposures: \acr{edcs}; outcome: \acr{hitrtse}) (HELIX Study; 2013-2016).**
 :::
@@
 ```
-Summary statistics of the estimated balancing weights (exposures: \acr{edcs}; outcomes: glucocorticosteroids).
+**Summary statistics of the estimated balancing weights (exposures: \acr{edcs}; outcomes: glucocorticosteroids) (HELIX Study; 2013-2016).**
 :::
@@
 ```
-Summary statistics of the estimated balancing weights (exposures: glucocorticosteroids; outcome: \acr{hitrtse}).
+**Summary statistics of the estimated balancing weights (exposures: glucocorticosteroids; outcome: \acr{hitrtse}) (HELIX Study; 2013-2016).**
 :::
@@
 ```
-Summary statistics of the estimated balancing weights for effect modification (exposures: \acr{edcs}; outcome: \acr{hitrtse}; modifier: sex).
+**Summary statistics of the estimated balancing weights for effect modification (exposures: \acr{edcs}; outcome: \acr{hitrtse}; modifier: sex) (HELIX Study; 2013-2016).**
 :::
@@
 ```
-Summary statistics of the estimated balancing weights for effect modification (exposures: \acr{edcs}; outcomes: glucocorticosteroids; modifier: sex).
+**Summary statistics of the estimated balancing weights for effect modification (exposures: \acr{edcs}; outcomes: glucocorticosteroids; modifier: sex) (HELIX Study; 2013-2016).**
 :::
@@
 ```
-Summary statistics of the estimated balancing weights for effect modification (exposures: glucocorticosteroids; outcome: \acr{hitrtse}; modifier: sex).
+**Summary statistics of the estimated balancing weights for effect modification (exposures: glucocorticosteroids; outcome: \acr{hitrtse}; modifier: sex) (HELIX Study; 2013-2016).**
 :::
@@
 ```
-**Measurement classification of \acr{edcs}, by cohort.**
+**Measurement classification of \acr{edcs}, by cohort (HELIX Study; 2013-2016).** 
 Coding: 1, quantifiable; 2, <LOD; 3, interference or out of range; 4. not analysed.
@@
 ```
-**Measurement classification of the glucocorticosteroids, by cohort.**
+**Measurement classification of the glucocorticosteroids, by cohort (HELIX Study; 2013-2016).** 
 Coding: 1, quantifiable; 2, <LOQ; 3, interference or out of range; 4, not detected.

### 2024-01-02 output/paper/_tables.qmd
First round of modifications for tables and figures done.

@@
 
+::: {#tbl-pop-desc}
 ```{r pop-desc}
-#| label: tbl-pop-desc
-#| tbl-cap: Participant characteristics, by cohort and overall (HELIX Study; 2013-2016).
-
 targets::tar_load("table1", store = store)
@@
 ```
+**Participant characteristics, by cohort and overall (HELIX Study; 2013-2016).**
+:::
 
@@
 
+::: {#tbl-edc-desc}
 ```{r edc-desc}
-#| label: tbl-edc-desc
-#| tbl-cap: Participants endocrine disruptors concentrations, by cohort and overall (HELIX Study; 2013-2016).
-
 targets::tar_load("desc_chems_mets", store = store)
@@
 ```
+**Participants endocrine disruptors concentrations, by cohort and overall (HELIX Study; 2013-2016).**
+:::
 
@@
 
+::: {#tbl-met-desc}
 ```{r met-desc}
-#| label: tbl-met-desc
-#| tbl-cap: Participants glucocorticosteroids concentrations, by cohort and overall (HELIX Study; 2013-2016).
-
 desc_chems_mets$desc_mets
 ```
+**Participants glucocorticosteroids concentrations, by cohort and overall (HELIX Study; 2013-2016).**
+:::
 
@@
 
+::: {#tbl-hypothesis-1}
 ```{r hypothesis-1}
-#| label: tbl-hypothesis-1
-#| tbl-cap: Pairwise differences between sex-specific marginal contrasts for the effect of \acr{edcs} on \acr{hitrtse} (HELIX Study; 2013-2016).
-
 tidy_res_meffects(
@@
 ```
+**Pairwise differences between sex-specific marginal contrasts for the effect of \acr{edcs} on \acr{hitrtse} (HELIX Study; 2013-2016).**
+:::
 
+::: {#tbl-hypothesis-2}
 ```{r hypothesis-2}
-#| label: tbl-hypothesis-2
-#| tbl-cap: Pairwise differences between sex-specific marginal contrasts for the effect of \acr{edcs} on the glucocorticosteroids (HELIX Study; 2013-2016).
-
 tidy_res_meffects(
@@
 ```
+**Pairwise differences between sex-specific marginal contrasts for the effect of \acr{edcs} on the glucocorticosteroids (HELIX Study; 2013-2016).**
+:::
 
+::: {#tbl-hypothesis-3}
 ```{r hypothesis-3}
-#| label: tbl-hypothesis-3
-#| tbl-cap: Pairwise differences between sex-specific marginal contrasts for the effect of the glucocorticosteroids on \acr{hitrtse} (HELIX Study; 2013-2016).
-
 tidy_res_meffects(
@@
 ```
+**Pairwise differences between sex-specific marginal contrasts for the effect of the glucocorticosteroids on \acr{hitrtse} (HELIX Study; 2013-2016).**
+:::

### 2024-01-02 output/paper/manuscript.qmd
First round of modifications for tables and figures done.

@@
 ```{r}
-info_steroids <- readODS::read_ods("docs/steroids.ods")
+info_steroids <- readODS::read_ods("../../docs/steroids.ods")
 print_tidy_abbreviations_steroids <- function(df, score) {
@@
 
-<!-- {{< include _figures.qmd >}} -->
+{{< include _figures.qmd >}}
 

### 2024-01-02 output/paper/_tables.qmd
First round of modifications for tables and figures done.

@@
 **Participants glucocorticosteroids concentrations, by cohort and overall (HELIX Study; 2013-2016).**
+:::
+
+::: {#tbl-met-new-desc}
+```{r met-new-desc}
+metabolites <- load_steroids()
+scores <- create_steroid_scores(dat = metabolites$metabolome) |>
+  tidylog::select(
+    HelixID,
+    cortisol_production, cortisone_production,
+    corticosterone_production
+  )
+scores <- myphd::extract_cohort(
+  dat = scores,
+  id_var = "HelixID",
+  st = 1, en = 3
+) |>
+  tidylog::select(-HelixID) |>
+  tidylog::mutate(
+    cohort = dplyr::case_when(
+      cohort == "EDE" ~ "EDEN",
+      cohort == "KAN" ~ "KANC",
+      cohort == "MOB" ~ "MOBA",
+      cohort == "RHE" ~ "RHEA",
+      .default = cohort
+    )
+  ) |>
+  tidylog::rename(
+    "cortisol production" = "cortisol_production",
+    "cortisone production" = "cortisone_production",
+    "corticosterone production" = "corticosterone_production"
+  )
+
+scores |>
+  gtsummary::tbl_summary(
+    by = "cohort"
+  ) |>
+  gtsummary::add_overall() |>
+  gtsummary::as_gt() |>
+  gt::opt_footnote_marks(
+    marks = "letters"
+  )
+```
+**Participants derived glucocorticosteroids concentrations, by cohort and overall (HELIX Study; 2013-2016).**
 :::

### 2024-01-02 output/paper/manuscript.qmd
First round of modifications for tables and figures done.

@@
 source(here::here("R/data.R"))
+source(here::here("R/utils.R"))
 source(here::here("R/proc_res.R"))
@@
 
-Levels of unprocessed non-persistent \acr{edcs} and glucocorticosteroids, by cohort and overall, are presented in @tbl-edc-desc and @tbl-met-desc, respectively.
+Levels of unprocessed non-persistent \acr{edcs} and glucocorticosteroids, by cohort and overall, are presented in @tbl-edc-desc, @tbl-met-desc, and @tbl-met-new-desc.
 
@@
 [] EDCs to HITRTSE
-[X] EDCs to CORT.
+[] EDCs to CORT.
 [] CORT. to HITRTSE

### 2024-01-04 output/paper/manuscript.qmd
Finished first draft manuscript w/o revision. Still nothing about sensitivity to period.

@@
 
-Our results are consistent with prior research that associated exposure to certain \acr{edcs} with higher levels of urinary and hair cortisol [@KimLeeMoon:2018; @SearsLiuLanphear:2023].
-Contrary to these prior studies, we did find effect modification by sex.
+We are not aware of prior studies specifically investigating the effects of exposure to \acr{edcs} in relation to \acr{hitrtse}.
+The literature on \acr{edcs} and neurodevelopment in children has mostly focused on \acr{oops} [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @CartierWarembourgLeManer-Idrissi:2016; YuDuChiou:2016], phthalate metabolites [@HuangChenSu:2015; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023], and \acr{bpa} [@TewarAuingerBraun:2016; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019], in relation to \acr{adhd} [@BouchardBellingerWright:2010; @TewarAuingerBraun:2016; @YuDuChiou:2016; @LiZhangKuang:2018], and intelligence scales [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @HuangChenSu:2015; @CartierWarembourgLeManer-Idrissi:2016; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023].
+Few studies have looked into different classes of \acr{edcs} [@ShoaffCoullWeuve:2020 in relation with the Conners Attention Deficit Scale and the Behavior Assessment System for Children; @OhKimKannan:2023 in relation with \acr{adhd} symptoms].
+Overall, and consistent with our results, these studies seem to provide further evidence of the negative effects of several \acr{edcs} on neurodevelopment in children.
+While not all these studies have investigated effect modification by sex, it seems that these negative effects are stronger in males.
+A major limitation of these studies is the reliance on spot urine samples, that might not be representative of long-term exposures.
+
+Our results are consistent with prior research that associated exposure to certain \acr{edcs} with higher levels of cortisol [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023].
+There are some differences, though.
+First, these studies only focus on phthalates, either as individual metabolites or as mixture.
+Second, exposure assessment in @SunLiJin:2018 and @KimLeeMoon:2018 was performed during gestation or the first 15 months of life, respectively.
+Finally, the glucocorticosteroids were measured in cord blood [@SunLiJin:2018] and hair [@SearsLiuLanphear:2023].
+Contrary to these studies, we did find effect modification by sex.
 We are not aware of other epidemiological studies investigating phthalates metabolites, phenols, and \acr{opps}, in relation to urinary glucocorticosteroids in childhood.
-We are also not aware of prior studies specifically investigating the effects of exposure to \acr{edcs} and elevated levels of glucocorticosteroids in relation to \acr{hitrtse}.
 
-[] EDCs to HITRTSE
-[] EDCs to CORT.
-[] CORT. to HITRTSE
-
----
+We are also not aware of prior studies specifically investigating the effects of elevated levels of glucocorticosteroids in relation to \acr{hitrtse}, although there is evidence that under- or over-production of glucocorticosteroids interfere with the normal development of the brain [@LupienMcEwenGunnar:2009].
+While we did find sex-specific evidence of an effect, their clinical relevance is questionable.
 
@@
 
-In conclusion, in a study of...
+In conclusion, in a study of almost 1,000 children from 6 European birth cohorts, we observed that (i) exposure to non-persistent \acr{edcs} in childhood might have short-term effects on \acr{hitrtse} in childhood, (ii) exposure to non-persistent \acr{edcs} in childhood might disrupt the \acr{hpa} axis in childhood, and (iii) disruption of the \acr{hpa} axis in childhood might have short-term, sex-specific effects on \acr{hitrtse}.
+Future studies should investigate how glucocorticosteroids might mediate the adverse effects of exposure to \acr{edcs} on childhood neurodevelopment in larger populations.
 

### 2024-01-04 output/paper/manuscript.qmd
Finished first draft manuscript w/o revision. Still nothing about sensitivity to period. Added abstract.

@@
 # Abstract {.unnumbered}
+
+Results of epidemiological studies seem to indicate that exposure to certain environmental chemicals, including endocrine disruptors (EDCs), is linked to perturbation of the hypothalamic-pituitary-adrenocortical (HPA) axis, which has a major role in brain development.
+We evaluated the effects of childhood exposure to organophosphate pesticides, phenols, and phthalate metabolites, on urinary glucocorticosteroids and the Attention Network Test (ANT) in childhood, using data from the Human Early Life Exposome (HELIX) Study.
+We used the parametric g-formula and marginal contrasts to estimate effects between EDCs, glucocorticosteroids, and a measure of efficiency from the ANT, the hit reaction time standard error (HRT-SE), in roughly 1,000 children aged 6-11.
+We further tested for possible effect modification by sex.
+We found positive marginal contrasts for exposure to the phenol MEPA and the phthalate metabolites oxo-MiNP, oh-MiNP, and MEHP, on HRT-SE.
+Multiple EDCs were also associated with positive marginal contrasts of cortisone, cortisol, and corticosterone production.
+Increased levels of the glucocorticosteroids had no effect on HRT-SE, although we found possible effect modification by sex for cortisone and corticosterone production.
+Our results suggest that multiple EDCs might interfere with the homeostasis of the HPA axis.
+Further research is necessary to determine whether this can have a clinically significant impact on childhood neurodevelopment.
 

### 2024-01-04 output/paper/manuscript.qmd
Rendered manuscript to send to Martine.

@@
 We are not aware of prior studies specifically investigating the effects of exposure to \acr{edcs} in relation to \acr{hitrtse}.
-The literature on \acr{edcs} and neurodevelopment in children has mostly focused on \acr{oops} [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @CartierWarembourgLeManer-Idrissi:2016; YuDuChiou:2016], phthalate metabolites [@HuangChenSu:2015; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023], and \acr{bpa} [@TewarAuingerBraun:2016; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019], in relation to \acr{adhd} [@BouchardBellingerWright:2010; @TewarAuingerBraun:2016; @YuDuChiou:2016; @LiZhangKuang:2018], and intelligence scales [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @HuangChenSu:2015; @CartierWarembourgLeManer-Idrissi:2016; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023].
+The literature on \acr{edcs} and neurodevelopment in children has mostly focused on \acr{opps} [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @CartierWarembourgLeManer-Idrissi:2016; YuDuChiou:2016], phthalate metabolites [@HuangChenSu:2015; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023], and \acr{bpa} [@TewarAuingerBraun:2016; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019], in relation to \acr{adhd} [@BouchardBellingerWright:2010; @TewarAuingerBraun:2016; @YuDuChiou:2016; @LiZhangKuang:2018], and intelligence scales [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @HuangChenSu:2015; @CartierWarembourgLeManer-Idrissi:2016; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023].
 Few studies have looked into different classes of \acr{edcs} [@ShoaffCoullWeuve:2020 in relation with the Conners Attention Deficit Scale and the Behavior Assessment System for Children; @OhKimKannan:2023 in relation with \acr{adhd} symptoms].

### 2024-02-05 output/paper/manuscript.qmd
Started revision manuscript from MV. New csl bib.

@@
 bibliography: bibliography.bib
-csl: ama11.csl
+csl: american-journal-of-epidemiology.csl
 
@@
 
+```{r}
+format_res <- function(df, exposure, outcome) {
+  # df: variable, val, estimate, outcome, class
+  # exposure in variable, outcome in outcome
+
+  val <- df[df$variable == exposure & df$outcome == outcome, ]$val
+  val <- stringr::str_replace(
+    string = val,
+    pattern = "\\*",
+    replacement = ""
+  )
+  glue::glue(
+    "\\acr{{mc}} and $95\\%$ \\acr{{ci}}: {val}"
+  )
+}
+```
+
+```{r}
+targets::tar_load(
+  dplyr::matches("marginal_comparisons_(\\d)$"),
+  store = store
+)
+meffects1 <- tidy_res_meffects(
+  df = marginal_comparisons_1,
+  sa_var = NULL,
+  outcome = c("hitrtse"),
+  which_res = "comparisons",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)$numerical_results
+meffects2 <- tidy_res_meffects(
+  df = marginal_comparisons_2,
+  sa_var = NULL,
+  outcome = c(
+    "cortisol production",
+    "cortisone production",
+    "corticosterone production"
+  ),
+  which_res = "comparisons",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)$numerical_results
+meffects3 <- tidy_res_meffects(
+  df = marginal_comparisons_3,
+  sa_var = NULL,
+  outcome = c("hitrtse"),
+  which_res = "comparisons",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)$numerical_results
+```
+
 # Abstract {.unnumbered}
 
-Results of epidemiological studies seem to indicate that exposure to certain environmental chemicals, including endocrine disruptors (EDCs), is linked to perturbation of the hypothalamic-pituitary-adrenocortical (HPA) axis, which has a major role in brain development.
-We evaluated the effects of childhood exposure to organophosphate pesticides, phenols, and phthalate metabolites, on urinary glucocorticosteroids and the Attention Network Test (ANT) in childhood, using data from the Human Early Life Exposome (HELIX) Study.
-We used the parametric g-formula and marginal contrasts to estimate effects between EDCs, glucocorticosteroids, and a measure of efficiency from the ANT, the hit reaction time standard error (HRT-SE), in roughly 1,000 children aged 6-11.
-We further tested for possible effect modification by sex.
-We found positive marginal contrasts for exposure to the phenol MEPA and the phthalate metabolites oxo-MiNP, oh-MiNP, and MEHP, on HRT-SE.
-Multiple EDCs were also associated with positive marginal contrasts of cortisone, cortisol, and corticosterone production.
-Increased levels of the glucocorticosteroids had no effect on HRT-SE, although we found possible effect modification by sex for cortisone and corticosterone production.
-Our results suggest that multiple EDCs might interfere with the homeostasis of the HPA axis.
+Epidemiological evidence suggests that exposure to \acr{edcs} may be linked to perturbation of the \acr{hpa} axis, which has a major role in brain development.
+We aimed to evaluate the effects of childhood exposure to organophosphate pesticides, phenols, and phthalate metabolites, on urinary glucocorticosteroids and inattention measures in children.
+We used data from the \acr{helix} cohort including 1,297 children aged 6-11.
+We used the parametric g-formula and \acr{mcs} to estimate effects between \acr{edcs}, glucocorticosteroids, and \acr{hitrtse}, a measure of efficiency from the \acr{ant}, and tested for possible effect modification by sex.
+We observed positive \acr{mcs} for exposure increases from the 10th to the 90th percentile for methyl-paraben (`r format_res(meffects1, "mepa", "hitrtse")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "hitrtse")`), oh-MiNP (`r format_res(meffects1, "ohminp", "hitrtse")`), and MEHP (`r format_res(meffects1, "mehp", "hitrtse")`), on \acr{hitrtse}.
+Several \acr{edcs} were also associated with positive \acr{mcs} of cortisone, cortisol, and corticosterone production.
+Increased levels of the glucocorticosteroids had no effect on \acr{hitrtse}, although we found possible effect modification by sex for cortisone and corticosterone production.
+Our results suggest that multiple \acr{edcs} might interfere with inattentiveness outcomes in children and with the homeostasis of the \acr{hpa} axis.
 Further research is necessary to determine whether this can have a clinically significant impact on childhood neurodevelopment.
@@
 - The length must be <4,000 words, excluding abstract, references, tables, figure captions, acknowledgments, and Supplementary Material.
-- In-text citations: full-sized Arabic numerals in parentheses within the sentence.
 - Submission: SER members receive a 10% discount on the fees per page.
@@
 
-The prevalence of several neurodevelopmental disorders has increased in the pediatric population [@GrandjeanLandrigan:2014], and multiple environmental pollutants play a role in the increased rates of these disorders [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
-\acr{edcs}, chemicals that are capable of interfering with the endocrine system, have been shown to have a role in childhood neurodevelopment and behavior [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @HuangChenSu:2015; @CartierWarembourgLeManer-Idrissi:2016; @TewarAuingerBraun:2016; @YuDuChiou:2016; @HuangTsaiChen:2017; @KimHongShin:2017; @FurlongHerringBuckley:2017; @Braun:2017; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019; @ShoaffCoullWeuve:2020; @OhKimKannan:2023; @VilmandBeckBilenberg:2023].
+The prevalence of several neurodevelopmental disorders has increased in the pediatric population [@GrandjeanLandrigan:2014], and multiple environmental pollutants may play a role in the increased rates of these disorders [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
+\acr{edcs}, ubiquitous chemicals present in many every-day products and diet, are capable of interfering with the endocrine system, and have shown associations with childhood neurodevelopment and behavior [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @HuangChenSu:2015; @CartierWarembourgLeManer-Idrissi:2016; @TewarAuingerBraun:2016; @YuDuChiou:2016; @HuangTsaiChen:2017; @KimHongShin:2017; @FurlongHerringBuckley:2017; @Braun:2017; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019; @ShoaffCoullWeuve:2020; @OhKimKannan:2023; @VilmandBeckBilenberg:2023].
 Although both pregnancy and early childhood are crucial stages of (neuro)development, most of the available literature is focused on the effects of prenatal exposure to \acr{edcs} on child neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
+
+\acr{opps} are one group of \acr{edcs} that may have a deleterious effect on neurodevelopment, although the few studies assessing exposure during childhood and through the use of biomarkers suffered from a series of limitations, including a small sample size and few \acr{opp} metabolites measured [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
+Exposure to phthalates and their metabolites has been associated to several adverse neurodevelopmental outcomes, but these studies were limited to few phthalate metabolites and small study populations [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
+The effects of exposure to \acr{bpa} during childhood on cognitive functions are still unclear [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
+
 Moreover, little is known about the biological mechanisms of action [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
-
-\acr{opps} might have a deleterious effect on neurodevelopment, although the few studies conducted on children and using a biological matrix for exposure assessment suffered from a series of limitations, including a small sample size and few \acr{opp} metabolites measured [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
-Phthalates and their metabolites have been associated to several adverse effects on neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
-Similarly, these studies focused on few phthalates metabolites and were conducted in small populations.
-On the other hand, the effects of exposure to \acr{bpa} on cognitive functions are still unclear [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
-
 There is also evidence that exposure to certain \acr{edcs}, specifically phthalates, might interfere with the \acr{hpa} axis and might interact with the glucocorticoid receptor [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023].
@@
 
-```{r}
-format_res <- function(df, exposure, outcome) {
-  # df: variable, val, estimate, outcome, class
-  # exposure in variable, outcome in outcome
-
-  val <- df[df$variable == exposure & df$outcome == outcome, ]$val
-  val <- stringr::str_replace(
-    string = val,
-    pattern = "\\*",
-    replacement = ""
-  )
-  glue::glue(
-    "\\acr{{mc}} and $95\\%$ \\acr{{ci}}: {val}"
-  )
-}
-```
-
-```{r}
-#| echo: false
-
-targets::tar_load(
-  dplyr::matches("marginal_comparisons_(\\d)$"),
-  store = store
-)
-meffects1 <- tidy_res_meffects(
-  df = marginal_comparisons_1,
-  sa_var = NULL,
-  outcome = c("hitrtse"),
-  which_res = "comparisons",
-  num_digits_est = num_digits_est,
-  num_digits_sig = num_digits_sig
-)$numerical_results
-meffects2 <- tidy_res_meffects(
-  df = marginal_comparisons_2,
-  sa_var = NULL,
-  outcome = c(
-    "cortisol production",
-    "cortisone production",
-    "corticosterone production"
-  ),
-  which_res = "comparisons",
-  num_digits_est = num_digits_est,
-  num_digits_sig = num_digits_sig
-)$numerical_results
-meffects3 <- tidy_res_meffects(
-  df = marginal_comparisons_3,
-  sa_var = NULL,
-  outcome = c("hitrtse"),
-  which_res = "comparisons",
-  num_digits_est = num_digits_est,
-  num_digits_sig = num_digits_sig
-)$numerical_results
-```
-
 The effective sample sizes before and after balancing weights estimation are presented in Tables [-@supptbl-balance-1; -@supptbl-balance-2; -@supptbl-balance-3], while basic summary statistics of the estimated balancing weights are presented in Tables [-@supptbl-weights-1; -@supptbl-weights-2; -@supptbl-weights-3].

### 2024-02-06 output/paper/_figures.qmd
First round of modifications for tables and figures done.
Modified almost all text based on MV. Added authors and roles.

@@
 ```
-**Marginal contrasts (exposures: \acr{edcs}; outcome: \acr{hitrtse}) (HELIX Study; 2013-2016).**
+**Marginal contrasts (exposures: \acr{edcs}; outcome: \acr{hitrtse}) (HELIX subcohort; 2013-2016).** 
 Circles indicate effect estimates. Solid lines indicate the $95\%$ \acr{ci}.
@@
 ```
-**Marginal contrasts (exposures: \acr{edcs}; outcomes: glucocorticosteroids) (HELIX Study; 2013-2016).**
+**Marginal contrasts (exposures: \acr{edcs}; outcomes: glucocorticosteroids) (HELIX subcohort; 2013-2016).** 
 Circles, triangles, and squares indicate effect estimates. Solid lines indicate the $95\%$ \acr{ci}.
@@
 ```
-**Marginal contrasts (exposures: glucocorticosteroids; outcome: \acr{hitrtse}) (HELIX Study; 2013-2016).**
+**Marginal contrasts (exposures: glucocorticosteroids; outcome: \acr{hitrtse}) (HELIX subcohort; 2013-2016).** 
 Circles indicate effect estimates. Solid lines indicate the $95\%$ \acr{ci}.

### 2024-02-06 output/paper/_si.qmd
Modified almost all text based on MV. Added authors and roles.

@@
 ```
-**Lower limits of quantification for the glucocorticosteroids (HELIX Study; 2013-2016).**
+**Lower limits of quantification for the glucocorticosteroids (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Effective sample size before and after balancing weights estimation (exposures: \acr{edcs}; outcome: \acr{hitrtse}) (HELIX Study; 2013-2016).**
+**Effective sample size before and after balancing weights estimation (exposures: \acr{edcs}; outcome: \acr{hitrtse}) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Effective sample size before and after balancing weights estimation (exposures: \acr{edcs}; outcomes: glucocorticosteroids) (HELIX Study; 2013-2016).**
+**Effective sample size before and after balancing weights estimation (exposures: \acr{edcs}; outcomes: glucocorticosteroids) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Effective sample size before and after balancing weights estimation (exposures: glucocorticosteroids; outcome: \acr{hitrtse}) (HELIX Study; 2013-2016).**
+**Effective sample size before and after balancing weights estimation (exposures: glucocorticosteroids; outcome: \acr{hitrtse}) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Summary statistics of the estimated balancing weights (exposures: \acr{edcs}; outcome: \acr{hitrtse}) (HELIX Study; 2013-2016).**
+**Summary statistics of the estimated balancing weights (exposures: \acr{edcs}; outcome: \acr{hitrtse}) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Summary statistics of the estimated balancing weights (exposures: \acr{edcs}; outcomes: glucocorticosteroids) (HELIX Study; 2013-2016).**
+**Summary statistics of the estimated balancing weights (exposures: \acr{edcs}; outcomes: glucocorticosteroids) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Summary statistics of the estimated balancing weights (exposures: glucocorticosteroids; outcome: \acr{hitrtse}) (HELIX Study; 2013-2016).**
+**Summary statistics of the estimated balancing weights (exposures: glucocorticosteroids; outcome: \acr{hitrtse}) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Summary statistics of the estimated balancing weights for effect modification (exposures: \acr{edcs}; outcome: \acr{hitrtse}; modifier: sex) (HELIX Study; 2013-2016).**
+**Summary statistics of the estimated balancing weights for effect modification (exposures: \acr{edcs}; outcome: \acr{hitrtse}; modifier: sex) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Summary statistics of the estimated balancing weights for effect modification (exposures: \acr{edcs}; outcomes: glucocorticosteroids; modifier: sex) (HELIX Study; 2013-2016).**
+**Summary statistics of the estimated balancing weights for effect modification (exposures: \acr{edcs}; outcomes: glucocorticosteroids; modifier: sex) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Summary statistics of the estimated balancing weights for effect modification (exposures: glucocorticosteroids; outcome: \acr{hitrtse}; modifier: sex) (HELIX Study; 2013-2016).**
+**Summary statistics of the estimated balancing weights for effect modification (exposures: glucocorticosteroids; outcome: \acr{hitrtse}; modifier: sex) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Measurement classification of \acr{edcs}, by cohort (HELIX Study; 2013-2016).** 
+**Measurement classification of \acr{edcs}, by cohort (HELIX subcohort; 2013-2016).** 
 Coding: 1, quantifiable; 2, <LOD; 3, interference or out of range; 4. not analysed.
@@
 ```
-**Measurement classification of the glucocorticosteroids, by cohort (HELIX Study; 2013-2016).** 
+**Measurement classification of the glucocorticosteroids, by cohort (HELIX subcohort; 2013-2016).** 
 Coding: 1, quantifiable; 2, <LOQ; 3, interference or out of range; 4, not detected.
@@
 ```
-**Marginal contrasts for effect modification (exposures: \acr{edcs}; outcome: \acr{hitrtse}; modifier: sex) (HELIX Study; 2013-2016).** 
+**Marginal contrasts for effect modification (exposures: \acr{edcs}; outcome: \acr{hitrtse}; modifier: sex) (HELIX subcohort; 2013-2016).** 
 Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ \acr{ci}.
@@
 ```
-**Marginal contrasts for effect modification (exposures: \acr{edcs}; outcomes: glucocorticosteroids; modifier: sex) (HELIX Study; 2013-2016).** 
+**Marginal contrasts for effect modification (exposures: \acr{edcs}; outcomes: glucocorticosteroids; modifier: sex) (HELIX subcohort; 2013-2016).** 
 Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ \acr{ci}.
@@
 ```
-**Marginal contrasts for effect modification (exposures: glucocorticosteroids; outcome: \acr{hitrtse}; modifier: sex) (HELIX Study; 2013-2016).** 
+**Marginal contrasts for effect modification (exposures: glucocorticosteroids; outcome: \acr{hitrtse}; modifier: sex) (HELIX subcohort; 2013-2016).** 
 Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ \acr{ci}.
 :::
-
-### Sensitivity to exposure assessment

### 2024-02-06 output/paper/_tables.qmd
Modified almost all text based on MV. Added authors and roles.

@@
 ```
-**Participant characteristics, by cohort and overall (HELIX Study; 2013-2016).**
+**Participant characteristics, by cohort and overall (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Participants endocrine disruptors concentrations, by cohort and overall (HELIX Study; 2013-2016).**
+**Participants endocrine disruptors concentrations, by cohort and overall (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Participants glucocorticosteroids concentrations, by cohort and overall (HELIX Study; 2013-2016).**
+**Participants glucocorticosteroids concentrations, by cohort and overall (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Participants derived glucocorticosteroids concentrations, by cohort and overall (HELIX Study; 2013-2016).**
+**Participants derived glucocorticosteroids concentrations, by cohort and overall (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Pairwise differences between sex-specific marginal contrasts for the effect of \acr{edcs} on \acr{hitrtse} (HELIX Study; 2013-2016).**
+**Pairwise differences between sex-specific marginal contrasts for the effect of \acr{edcs} on \acr{hitrtse} (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Pairwise differences between sex-specific marginal contrasts for the effect of \acr{edcs} on the glucocorticosteroids (HELIX Study; 2013-2016).**
+**Pairwise differences between sex-specific marginal contrasts for the effect of \acr{edcs} on the glucocorticosteroids (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Pairwise differences between sex-specific marginal contrasts for the effect of the glucocorticosteroids on \acr{hitrtse} (HELIX Study; 2013-2016).**
+**Pairwise differences between sex-specific marginal contrasts for the effect of the glucocorticosteroids on \acr{hitrtse} (HELIX subcohort; 2013-2016).**
 :::

### 2024-02-06 output/paper/authors.qmd
Modified almost all text based on MV. Added authors and roles.

--- 2024-01-04 output/paper/manuscript.qmd
@@
 
-# metadata-files:
-#   - ../../docs/authors.yml
+metadata-files:
+  - ../../docs/authors.yml
 
 format:
-  docx:
-    reference-doc: reference_document.docx
-    citeproc: true
-  # pdf:
-  #   pdf-engine: lualatex
-  #   number-sections: true
-  #   toc: false
-  #   keep-tex: false
-  #   include-in-header:
-  #     text: |
-  #       \usepackage{lineno}
-  #       \usepackage{rotating}
-  #       \usepackage{float}
-  #       \usepackage{typearea}
-  #   include-before-body:
-  #     text: |
-  #       \linenumbers
-
-crossref:
-  custom: 
-    - kind: float
-      key: supptbl
-      latex-env: supptbl
-      reference-prefix: Table S
-      space-before-numbering: false
-    - kind: float
-      key: suppfig
-      latex-env: suppfig
-      reference-prefix: Figure S
-      space-before-numbering: false
-
-bibliography: bibliography.bib
-csl: ama11.csl
+  html:
+    embed-resources: true
 
@@
 linkcolor: blue
-
-filters:
-  - acronyms
-  # - authors-block
-  # - ../../../../all_phd/_extensions/wordcount/citeproc.lua
-  - ../../../../all_phd/_extensions/wordcount/wordcount.lua
-
-acronyms:
-  fromfile: ../../../../all_phd/acronyms.yml
-  insert_links: false
-  insert_loa: false
-  sorting: alphabetical
-
-editor_options: 
-  chunk_output_type: console
-
-execute:
-  echo: false
-  warning: false
-  error: false
-  message: false
 ---
-
-```{r}
-here::i_am("output/paper/manuscript.qmd")
-```
-
-```{r vars-methods}
-source(here::here("R/dictionaries.R"))
-source(here::here("R/data.R"))
-source(here::here("R/utils.R"))
-source(here::here("R/proc_res.R"))
-
-is_hpc <- FALSE
-store <- ifelse(
-  is_hpc, 
-  paste0(
-    "/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/",
-    "lorenzoF/data/data_paper3/_targets/_targets_res"
-  ), 
-  "~/mounts/rstudioserver/data/data_paper3/_targets/_targets_res"
-)
-id_var <- "HelixID"
-num_digits_est = 3
-num_digits_sig = 2
-params_ <- params(is_hpc = is_hpc)
-threshold_within_lod <- params_$steps$rq2$preproc_outcome$llodq$threshold_within
-threshold_overall_lod <- params_$steps$rq2$preproc_outcome$llodq$threshold_overall
-threshold_within_nan <- params_$steps$rq2$preproc_outcome$missings$threshold_within
-threshold_overall_nan <- params_$steps$rq2$preproc_outcome$missings$threshold_overall
-knn_k <- params_$steps$rq2$preproc_outcome$missings$k
-```
-
-# Abstract {.unnumbered}
-
-Results of epidemiological studies seem to indicate that exposure to certain environmental chemicals, including endocrine disruptors (EDCs), is linked to perturbation of the hypothalamic-pituitary-adrenocortical (HPA) axis, which has a major role in brain development.
-We evaluated the effects of childhood exposure to organophosphate pesticides, phenols, and phthalate metabolites, on urinary glucocorticosteroids and the Attention Network Test (ANT) in childhood, using data from the Human Early Life Exposome (HELIX) Study.
-We used the parametric g-formula and marginal contrasts to estimate effects between EDCs, glucocorticosteroids, and a measure of efficiency from the ANT, the hit reaction time standard error (HRT-SE), in roughly 1,000 children aged 6-11.
-We further tested for possible effect modification by sex.
-We found positive marginal contrasts for exposure to the phenol MEPA and the phthalate metabolites oxo-MiNP, oh-MiNP, and MEHP, on HRT-SE.
-Multiple EDCs were also associated with positive marginal contrasts of cortisone, cortisol, and corticosterone production.
-Increased levels of the glucocorticosteroids had no effect on HRT-SE, although we found possible effect modification by sex for cortisone and corticosterone production.
-Our results suggest that multiple EDCs might interfere with the homeostasis of the HPA axis.
-Further research is necessary to determine whether this can have a clinically significant impact on childhood neurodevelopment.
-
-{{< pagebreak >}}
-
-- Authorship must follow the ICMJE's criteria.
-Author names should be listed in ScholarOne and author contributions should be detailed in the cover letter ("Author A...").
-Neither names nor contributions should appear in the blinded manuscript.
-Do not include conflicts of interest.
-- The length of the abstract must be less or equal than 200 words, and should be unstructured, stating the research questions, the methods used, and the results and conclusions of the research.
-- The length must be <4,000 words, excluding abstract, references, tables, figure captions, acknowledgments, and Supplementary Material.
-- In-text citations: full-sized Arabic numerals in parentheses within the sentence.
-- Submission: SER members receive a 10% discount on the fees per page.
-This should be requested in the cover letter.
-Also 20% discount for Open Access charges.
-If uploading a single file (document, tables, figures, SM), designate the file as "Main document - anonymous".
-
-The prevalence of several neurodevelopmental disorders has increased in the pediatric population [@GrandjeanLandrigan:2014], and multiple environmental pollutants play a role in the increased rates of these disorders [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
-\acr{edcs}, chemicals that are capable of interfering with the endocrine system, have been shown to have a role in childhood neurodevelopment and behavior [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @HuangChenSu:2015; @CartierWarembourgLeManer-Idrissi:2016; @TewarAuingerBraun:2016; @YuDuChiou:2016; @HuangTsaiChen:2017; @KimHongShin:2017; @FurlongHerringBuckley:2017; @Braun:2017; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019; @ShoaffCoullWeuve:2020; @OhKimKannan:2023; @VilmandBeckBilenberg:2023].
-Although both pregnancy and early childhood are crucial stages of (neuro)development, most of the available literature is focused on the effects of prenatal exposure to \acr{edcs} on child neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
-Moreover, little is known about the biological mechanisms of action [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
-
-\acr{opps} might have a deleterious effect on neurodevelopment, although the few studies conducted on children and using a biological matrix for exposure assessment suffered from a series of limitations, including a small sample size and few \acr{opp} metabolites measured [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
-Phthalates and their metabolites have been associated to several adverse effects on neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
-Similarly, these studies focused on few phthalates metabolites and were conducted in small populations.
-On the other hand, the effects of exposure to \acr{bpa} on cognitive functions are still unclear [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
-
-There is also evidence that exposure to certain \acr{edcs}, specifically phthalates, might interfere with the \acr{hpa} axis and might interact with the glucocorticoid receptor [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023].
-
-The HPA axis, which can be activated by stress, is responsible for the production of glucocorticosteroids.
-The brain, and its proper functioning, is a potential target, due to the presence of receptors for these hormones [@LupienMcEwenGunnar:2009; @SunLiJin:2018].
-Glucocorticosteroids are necessary for brain maturation, although their under- or over-production might interfere with its normal development and ultimately lead to long-term impaired functioning [@LupienMcEwenGunnar:2009; @SearsLiuLanphear:2023].
-
-Taken together, these results suggest that the negative influence of exposure to certain \acr{edcs} on neurodevelopmental outcomes, might be mediated, at least partially, by disruption of the \acr{hpa} axis' homeostasis.
-In the present study, we thus estimated associations between 1) non-persistent \acr{edcs} and attention, 2) non-persistent \acr{edcs} and glucocorticosteroids, and 3) glucocorticosteroids and attention, using the parametric g-formula and marginal contrasts, in children of a large prospective birth cohort in Europe.
-
-# Methods {#sec-methods}
-
-## Study population and design {#sec-design}
-
-The \acr{helix} is an ongoing project which aims to characterize early-life exposures and their potential association with endogenous biomarkers and health outcomes [@VrijheidSlamaRobinson:2014].
-It consists of six existing population-based birth cohort studies across Europe: BiB (Born in Bradford, UK) [@WrightSmallRaynor:2013], EDEN (Study of determinants of pre- and postnatal developmental, France) [@HeudeForhanSlama:2016], INMA (Environment and Childhood, Spain) [@GuxensBallesterEspada:2012], KANC (Kaunas Cohort, Lithuania) [@GrazulevicieneDanileviciuteNadisauskiene:2009], MoBa (The Norwegian Mother and Child Cohort Study, Norway) [@MagnusIrgensHaug:2006], and Rhea (Mother–Child Cohort in Crete, Greece) [@ChatziPlanaDaraki:2009], for a total of 32,000 mother-child pairs.
-A \acr{helix} subcohort of 1,200 mother-child pairs was fully characterized for the external and internal exposome, including exposure and omics biomarkers during childhood.
-Eligibility criteria for inclusion in the \acr{helix} subcohort included: a) age 6-11 years, with a preference for 7-9 years; b) availability of sufficient stored pregnancy blood and urine samples; c) availability of complete address history from first to last follow-up; d) no serious health problems, which might affect the results of the clinical testing.
-Further information can be found in @MaitreBontCasas:2018.
-
-Ethical permission was obtained from the relevant authorities in the corresponding country.
-
-## Variables {#sec-vars}
-
-### Confounders {#sec-confounders}
-
-For each research question, defined by a specific type of exposure and outcome, the minimal set of covariates for inclusion in the analyses was selected on the basis of a \acr{dag} built with DAGitty [@TextorvanderZanderGilthorpe:2016] and ggdag [@Barrett:2023].
-The sets of covariates were selected to estimate the total effect of the exposure on the outcome.
-Further, each minimal adjustment set was *augmented* with precision covariates, defined as the set of parents variable of the outcome that are not parents of the exposure.
-The adjustment sets are provided in the Supplementary Material as text files compatible with DAGitty.
-
-- For RQ1 I used creatinine values from HELIX. For RQ3 the ones from the steroids dataset. For RQ2, I included in the model both variables.
-
-### Endocrine disrupting chemicals {#sec-edcs}
-
-```{r}
-info_edcs <- myphd::edcs_information()
-print_tidy_abbreviations_edcs <- function(df, class_) {
-  tmp <- df |>
-    dplyr::filter(class == class_) |>
-    dplyr::select(chem_id, full_name) |>
-    dplyr::arrange(full_name) |>
-    dplyr::mutate(
-      ret = glue::glue("{full_name} ({chem_id})")
-    )
-  tmp$ret |>
-    paste0(collapse = ", ")
-}
-```
-
-Children were assessed between December 2013 and February 2016, and included neurological testing and urine collection.
-Urine samples of the night before and the first morning void on the day of the visit were combined to provide a more reliable exposure assessment.
-Non-persistent \acr{edcs} assessed in urine samples from children included phthalate metabolites, phenols, and \acr{op} pesticide metabolites.
-A list of the environmental chemicals determined in urine samples and used for the present study is given in @supptbl-info-chems.
-Briefly, we analyzed a total of 7 phenols (`r print_tidy_abbreviations_edcs(info_edcs, "Phenols")`), 6 non-specific organophosphate pesticide metabolites (`r print_tidy_abbreviations_edcs(info_edcs, "OP pesticide metabolites")`), and 10 phthalate metabolites (`r print_tidy_abbreviations_edcs(info_edcs, "Phthalate metabolites")`) originating from 6 distinct phthalate parent compounds.
-The laboratory protocols for the analysis are described elsewhere [@HaugSakhiCequier:2018].
-
-### Glucocorticosteroids {#sec-steroids}
-
-```{r}
-info_steroids <- readODS::read_ods("../../docs/steroids.ods")
-print_tidy_abbreviations_steroids <- function(df, score) {
-  cortisol_prod <- c(
-    "20aDHF", "20bDHF", "5bDHF", "5aTHF", "5bTHF",
-    "6OHF",
-    "5a20acortol", "5a20bcortol", "5b20acortol", "5b20bcortol"
-  )
-  cortisone_prod <- c(
-    "20aDHE", "20bDHE", "5aTHE", "5bTHE",
-    "6OHE",
-    "5b20acortolone", "5b20bcortolone"
-  )
-  corticosterone_prod <- c(
-    "5aTHB", "5bTHB", "A", "17-DO-cortolone"
-  )
-  tmp <- df |>
-    dplyr::filter(dplyr::case_when(
-      score == "cortisol_prod" ~ acronym %in% cortisol_prod,
-      score == "cortisone_prod" ~ acronym %in% cortisone_prod,
-      score == "corticosterone_prod" ~ acronym %in% corticosterone_prod,
-    )) |>
-    dplyr::select(analyte, acronym) |>
-    dplyr::arrange(analyte) |>
-    dplyr::mutate(
-      ret = glue::glue("{analyte} ({acronym})")
-    )
-  tmp$ret |>
-    paste0(collapse = ", ")
-}
-```
-
-Urine samples of the night before the day of the visit were used to measure levels of the glucocorticosteroids.
-These included glucocorticosteroids, glucocorticosteroid metabolites, glucocorticosteroid precursors, glucocorticosteroid precursor metabolites, androgens, and androgen metabolites.
-A list of the glucocorticosteroids determined in urine samples and used for the present study is given in @supptbl-info-mets.
-
-To assess the levels of glucocorticosteroids and their metabolites, LC-MS/MS analysis was applied at the Applied Metabolomics Research Group, IMIM (Hospital del Mar Medical Research Institute).
-The laboratory protocols for the analysis are described elsewhere [@MarcosRenauCasals:2014; @Gomez-GomezPozo:2020].
-Of the 1,004 urine samples, 980 children were matched to the \acr{helix} subcohort.
-
-Three additional markers, total cortisol production, total cortisone production, and total corticosterone production, were computed based on the following: cortisol production as the sum of cortisol and its metabolites (`r print_tidy_abbreviations_steroids(info_steroids, "cortisol_prod")`), cortisone production as the sum of cortisone and its metabolites (`r print_tidy_abbreviations_steroids(info_steroids, "cortisone_prod")`), and corticosterone production as the sum of `r print_tidy_abbreviations_steroids(info_steroids, "corticosterone_prod")`.
-
-### Neurodevelopment {#sec-neurodevelopment}
-
-Neurodevelopmental outcomes were assessed with standardized, non-linguistic, and culturally blind computer tests, including the \acr{ant} [@RuedaFanMcCandliss:2004].
-Further information can be found in @MaitreBontCasas:2018.
-Briefly, it is a computerized test that provides a measure of efficiency in three different functions of attention: alerting, orienting, and executive attention.
-The outcome of interest for the present study is the \acr{hitrtse} [@SunyerEsnaolaAlvarez-Pedrerol:2015], a measure of response speed consistency throughout the test.
-A high \acr{hitrtse} indicates highly variable reactions, and is considered a measure of inattentiveness.
-
-## Statistical methods {#sec-stat-methods}
-
-### Data pre-processing {#sec-dat-preproc}
-
-Concentrations of the glucocorticosteroids were classified as quantifiable, below the \acr{loq}, possible interference or out of range, and not detected.
-For each metabolite, we computed the fraction of values below the \acr{loq} and not detected, both within each cohort and overall.
-We proceeded to impute these values using half the value of the corresponding \acr{lloq}, for those metabolites that had less than `r threshold_within_lod`% of missings within each cohort and `r threshold_overall_lod`% of missings overall.
-Information about the \acr{lloq} for the glucocorticosteroids is provided in @supptbl-lloq-mets.
-The remaining missing values were imputed using kNN from the `VIM` R package [@KowarikTempl:2016], for those metabolites that had less than `r threshold_within_nan`% of remaining missings within each cohort and `r threshold_overall_nan`% of remaining missings overall.
-We used `r knn_k` nearest neighbors.
-We natural log-transformed them to improve model fit, assessed with posterior predictive checks.
-To do so, replicated data were simulated with the fitted models and compared to the observed data.
-We used the `check_predictions` function from the `performance` R package using the default arguments [@LudeckeBen-ShacharPatil:2021].
-Values of cortisol production and cortisone production were expressed in nanograms per millilitre, whereas values of cortisol metabolism and 11bHSD activity were unitless.
-
-Concentrations of the non-persistent \acr{edcs} were classified as quantifiable, below the \acr{lod}, possible interference or out of range, and not analysed.
-Concentrations below the \acr{lod} were singly imputed using a quantile regression approach for the imputation of left-censored missing data, as implemented in the `impute.QRILC` function from the `imputeLCMD` R package [@lazar2015imputelcmd].
-Information about the lower limits of detection can be found in @HaugSakhiCequier:2018.
-Chemicals with more than 70% of observations below the \acr{lod} were not considered in the present study.
-Remaining missing values were imputed similarly using kNN.
-Values of the chemicals were expressed in $\mu$grams per litre.
-
-Missing values in the clinical outcome were imputed similarly using kNN.
-Similarly, we natural log-transformed them to improve model fit, assessed with posterior predictive checks.
-Values of the clinical outcome were expressed in \acr{mss}.
-
-Missing values in the covariates were imputed similarly using kNN.
-Categorical covariates were imputed using the `maxCat` function, which chooses the level with the most occurrences.
-Creatinine values were expressed in grams per litre.
-
-### Estimation of balancing weights {#sec-weights-est}
-
-Stabilized balancing weights were estimated using the energy method available in the `WeightIt` R package [@Greifer:2023].
-This methods estimates weights by minimizing an energy statistic related to covariate balance [@HulingGreiferChen:2023], thus avoiding the need to specify a parametric model.
-Weights below the 0.1 and above the 0.9 quantiles were trimmed.
-Trimming might lead to decreased covariate balance and potentially change the estimand, but can also decrease the variability of the weights.
-Covariate balance was assessed using functionalities provided by the `cobalt` R package [@Greifer:2023a]. Specifically, we used *Love* plots to visualize covariate balance before and after adjusting.
-
-### G-computation {#sec-gcomp}
-
-We estimated marginal contrasts with the parametric g-formula, a method of standardization.
-The parametric g-formula involves the following steps: 1) fit a outcome model including both covariates and balancing weights; 2) create two new datasets identical to the original one but with the exposure shifted according to a user-specified intervention set by a deterministic function of the observed exposure levels; 3) use the outcome model to compute adjusted predictions in the two counterfactual datasets; 4) compute the difference between the means of the adjusted predictions in the counterfactual datasets.
-The causal parameter of interest $\phi$ was thus specified as the difference in the expected counterfactual outcomes under the shifted exposure levels $\left( \phi^\Delta = \mathbb{E} \left[ Y^{d_1} \right] - \mathbb{E} \left[ Y^{d_2} \right] \right)$.
-In order for this parameter to be identified, the usual causal identifiability conditions (no unmeasured confounding, positivity, and consistency) are required.
-Since these conditions are likely not satisfied, we focused on the estimation of a statistical estimand that is as close as possible to the causal parameter of interest.
-
-We fit the outcome model using the `glm` function and a Gaussian family with identity link from base R.
-The exposure variable was modeled using natural cubic splines with 3 degrees of freedom, to more flexibly capture the \acr{adrf}.
-When the outcome was a ratio, as was the case for cortisol metabolism and 11bHSD activity, we included the logarithm of its denominator, cortisol and cortisol production, respectively, as a control variable [@BartlettPartnoy:2020].
-
-To estimate the marginal contrasts, we used the `avg_comparisons` function from the `marginaleffects` R package [@Arel-Bundock:2023].
-The two counterfactual datasets were obtained by setting the exposures levels to 90th percentile ($d_1$) and the 10th percentile ($d_2$), for each cohort separately.
-The marginal contrasts were computed using the estimated balancing weights above.
-Robust standard errors were computed with the `sandwich` R package, using cohort as variable indicating clustering of observations [@Zeileis:2004; @ZeileisKollGraham:2020].
-For each outcome, we report the results as differences between marginal contrasts.
-
-We further estimated the \acr{adrf} using the `avg_predictions` function from the `marginaleffects` R package, examining 50 exposure values from the 10th to the 90th percentiles of the exposure.
-As done for the marginal contrasts, we included the estimated balancing weights and used cohort as a clustering variable when computing robust standard errors.
-
-### Effect-modification analysis
-
-We tested for possible effect-modification by sex.
-To do so, balancing weights were estimated separately for each level of the sex variable, and an interaction term between the exposure and sex was included in the outcome model.
-Similarly, the marginal contrasts were aggregated separately for each level of sex.
-We further tested for significance of the difference between the marginal contrasts of males and females.
-
-# Results {#sec-res}
-
-## Participants {#sec-res-participants}
-
-In total, 1,297 children were included in the dataset.
-Of these, 976 had information also about the glucocorticosteroids.
-The sample consisted of 55% girls.
-The median \acr{hitrtse} was 300 ms (\acr{iqr}, 231-368), with lower median values for EDEN, MOBA, and SAB, corresponding to the cohorts with older children.
-At the time of visit, the median age of the children was 8.06 years.
-The children were mostly Caucasian (90%) or Pakistani (6.2%).
-@tbl-pop-desc provides descriptive statistics for the outcome and covariates, by cohort and overall.
-Codebooks for the used covariates, by research question, are provided in @supptbl-codebook-1, @supptbl-codebook-2, and @supptbl-codebook-3.
-
-## Descriptive data {#sec-res-pop-desc}
-
-Levels of unprocessed non-persistent \acr{edcs} and glucocorticosteroids, by cohort and overall, are presented in @tbl-edc-desc, @tbl-met-desc, and @tbl-met-new-desc.
-
-## Main results {#sec-res-main}
-
-```{r}
-format_res <- function(df, exposure, outcome) {
-  # df: variable, val, estimate, outcome, class
-  # exposure in variable, outcome in outcome
-
-  val <- df[df$variable == exposure & df$outcome == outcome, ]$val
-  val <- stringr::str_replace(
-    string = val,
-    pattern = "\\*",
-    replacement = ""
-  )
-  glue::glue(
-    "\\acr{{mc}} and $95\\%$ \\acr{{ci}}: {val}"
-  )
-}
-```
-
-```{r}
-#| echo: false
-
-targets::tar_load(
-  dplyr::matches("marginal_comparisons_(\\d)$"),
-  store = store
-)
-meffects1 <- tidy_res_meffects(
-  df = marginal_comparisons_1,
-  sa_var = NULL,
-  outcome = c("hitrtse"),
-  which_res = "comparisons",
-  num_digits_est = num_digits_est,
-  num_digits_sig = num_digits_sig
-)$numerical_results
-meffects2 <- tidy_res_meffects(
-  df = marginal_comparisons_2,
-  sa_var = NULL,
-  outcome = c(
-    "cortisol production",
-    "cortisone production",
-    "corticosterone production"
-  ),
-  which_res = "comparisons",
-  num_digits_est = num_digits_est,
-  num_digits_sig = num_digits_sig
-)$numerical_results
-meffects3 <- tidy_res_meffects(
-  df = marginal_comparisons_3,
-  sa_var = NULL,
-  outcome = c("hitrtse"),
-  which_res = "comparisons",
-  num_digits_est = num_digits_est,
-  num_digits_sig = num_digits_sig
-)$numerical_results
-```
-
-The effective sample sizes before and after balancing weights estimation are presented in Tables [-@supptbl-balance-1; -@supptbl-balance-2; -@supptbl-balance-3], while basic summary statistics of the estimated balancing weights are presented in Tables [-@supptbl-weights-1; -@supptbl-weights-2; -@supptbl-weights-3].
-As expected, the median value of the weights for each exposure was close to $1.00$.
-
-@Fig-marginal-1 presents the forest plot for the marginal contrasts of the non-persistent \acr{edcs} on \acr{hitrtse}.
-Overall, a cohort-specific increase in the levels of the exposures from the 10th to the 90th percentiles was associated with a positive marginal contrast, indicating an increase in the values of \acr{hitrtse}.
-Most of the \acr{ci}s included the null effect, though.
-Statistically significant effects were observed for the phenol MEPA (`r format_res(meffects1, "mepa", "hitrtse")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "hitrtse")`), oh-MiNP (`r format_res(meffects1, "ohminp", "hitrtse")`), and MEHP (`r format_res(meffects1, "mehp", "hitrtse")`).
-The \acr{opp} DETP was strongly negatively associated with \acr{hitrtse} (`r format_res(meffects1, "detp", "hitrtse")`).
-
-@Fig-marginal-2 presents the forest plot for the marginal contrasts of the non-persistent \acr{edcs} on cortisone production, cortisol production, and corticosterone production.
-Overall, a cohort-specific increase in the levels of the exposures from the 10th to the 90th percentiles was associated with a positive marginal contrast, indicating an increase in the total production of these metabolites, with the exception of BUPA, which was associated with negative marginal contrasts for all three outcomes, and MiBP, which was associated with a negative marginal contrast for total cortisone production only.
-The majority of the effects for the phenols and phthalate metabolites included the null.
-The phenol BPA showed the largest marginal contrasts across all three outcomes (cortisone production, `r format_res(meffects2, "bpa", "cortisone production")`; cortisol production, `r format_res(meffects1, "bpa", "cortisol production")`; corticosterone production, `r format_res(meffects1, "bpa", "corticosterone production")`).
-
-@Fig-marginal-3 presents the forest plot for the marginal contrasts of the glucocorticosteroids on \acr{hitrtse}.
-All marginal contrasts included the null, with no clear indication of directionality of the effect.
-
-## Other analyses {#sec-res-other}
-
-### Effect modification by sex
-
-```{r}
-#| echo: false
-
-targets::tar_load(
-  dplyr::matches("marginal_hypothesis_(\\d)SA$"),
-  store = store
-)
-heffects1 <- tidy_res_meffects(
-  df = marginal_hypothesis_1SA,
-  sa_var = "e3sex",
-  outcome = NULL,
-  which_res = "hypothesis",
-  num_digits_est = num_digits_est,
-  num_digits_sig = num_digits_sig
-)$numerical_results
-heffects2 <- tidy_res_meffects(
-  df = marginal_hypothesis_2SA,
-  sa_var = "e3sex",
-  outcome = NULL,
-  which_res = "hypothesis",
-  num_digits_est = num_digits_est,
-  num_digits_sig = num_digits_sig
-)$numerical_results
-heffects3 <- tidy_res_meffects(
-  df = marginal_hypothesis_3SA,
-  sa_var = "e3sex",
-  outcome = NULL,
-  which_res = "hypothesis",
-  num_digits_est = num_digits_est,
-  num_digits_sig = num_digits_sig
-)$numerical_results
-```
-
-Basic summary statistics of the estimated balancing weights for effect modification are presented in Tables [-@supptbl-weights-1sa; -@supptbl-weights-2sa; -@supptbl-weights-3sa].
-As expected, the median value of the weights for each exposure was close to $1.00$.
-
-@Tbl-hypothesis-1 presents the results of a hypothesis test for the difference between estimates of the marginal contrasts for females and males, for each exposure.
-Significant differences were present for the phenol OXBE (`r format_res(heffects1, "oxbe", "hitrtse")`) and the phthalate metabolites MEP (`r format_res(heffects1, "mep", "hitrtse")`) and MbZP (`r format_res(heffects1, "mbzp", "hitrtse")`).
-The forest plot of the individual marginal contrasts is presented in @suppfig-marginal-1sa.
-
-@Tbl-hypothesis-2 presents the results of a hypothesis test for the difference between estimates of the marginal contrasts for females and males, for each exposure and outcome.
-Significant differences were present across all three classes of \acr{edcs} and for all outcomes.
-The largest differences were attributable to the \acr{opps} DMTP (cortisol production, `r format_res(heffects2, "dmtp", "cortisol production")`) and DETP (corticosterone production, (`r format_res(heffects2, "detp", "corticosterone production")`); cortisone production, (`r format_res(heffects2, "detp", "cortisone production")`)).
-The forest plots of the individual marginal contrasts are presented in @suppfig-marginal-2sa.
-
-@Tbl-hypothesis-3 presents the results of a hypothesis test for the difference between estimates of the marginal contrasts for females and males, for each exposure.
-Significant differences were present for cortisone production (`r format_res(heffects3, "cortisone prod.", "hitrtse")`) and corticosterone production (`r format_res(heffects3, "corticost.
-prod.", "hitrtse")`).
-Furthermore, for all exposures, the marginal contrasts had opposite sign (positive for males and negative for females).
-The forest plot of the individual marginal contrasts is presented in @suppfig-marginal-3sa.
-
-### Sensitivity to exposure assessment
-
-# Discussion {#sec-discussion}
-
-The impact of exposure to (non-persistent) \acr{edcs} on human health has attracted considerable research interest.
-While research in this area has mainly investigated the effects of prenatal exposure on child neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022], little is still known about later-in-life exposure and childhood neurodevelopment.
-In this study, consisting of almost 1,000 children from 6 European birth cohorts, we observed that short-term childhood exposure to non-persistent \acr{edcs} had negative effects on \acr{hitrtse} and total production of cortisol, cortisone, and corticosterone, with the majority of the \acr{ci}s including the null effect.
-Increased production of these glucocorticosteroids did not seem to affect \acr{hitrtse}.
-Some of these effects differed for females and males, including significant differences for the effects of increased production of cortisone and corticosterone on \acr{hitrtse}.
-Specifically, an increased production of these glucocorticosteroids was associated with lower values of \acr{hitrtse} for females, and higher values for males.
-Taken together, these results suggest that these non-persistent \acr{edcs} might be responsible for perturbations of the \acr{hpa} axis' homeostasis, and that higher levels of these glucocorticosteroids might interfere with different functions of attention in a sex-specific manner.
-
-We are not aware of prior studies specifically investigating the effects of exposure to \acr{edcs} in relation to \acr{hitrtse}.
-The literature on \acr{edcs} and neurodevelopment in children has mostly focused on \acr{opps} [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @CartierWarembourgLeManer-Idrissi:2016; YuDuChiou:2016], phthalate metabolites [@HuangChenSu:2015; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023], and \acr{bpa} [@TewarAuingerBraun:2016; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019], in relation to \acr{adhd} [@BouchardBellingerWright:2010; @TewarAuingerBraun:2016; @YuDuChiou:2016; @LiZhangKuang:2018], and intelligence scales [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @HuangChenSu:2015; @CartierWarembourgLeManer-Idrissi:2016; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023].
-Few studies have looked into different classes of \acr{edcs} [@ShoaffCoullWeuve:2020 in relation with the Conners Attention Deficit Scale and the Behavior Assessment System for Children; @OhKimKannan:2023 in relation with \acr{adhd} symptoms].
-Overall, and consistent with our results, these studies seem to provide further evidence of the negative effects of several \acr{edcs} on neurodevelopment in children.
-While not all these studies have investigated effect modification by sex, it seems that these negative effects are stronger in males.
-A major limitation of these studies is the reliance on spot urine samples, that might not be representative of long-term exposures.
-
-Our results are consistent with prior research that associated exposure to certain \acr{edcs} with higher levels of cortisol [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023].
-There are some differences, though.
-First, these studies only focus on phthalates, either as individual metabolites or as mixture.
-Second, exposure assessment in @SunLiJin:2018 and @KimLeeMoon:2018 was performed during gestation or the first 15 months of life, respectively.
-Finally, the glucocorticosteroids were measured in cord blood [@SunLiJin:2018] and hair [@SearsLiuLanphear:2023].
-Contrary to these studies, we did find effect modification by sex.
-We are not aware of other epidemiological studies investigating phthalates metabolites, phenols, and \acr{opps}, in relation to urinary glucocorticosteroids in childhood.
-
-We are also not aware of prior studies specifically investigating the effects of elevated levels of glucocorticosteroids in relation to \acr{hitrtse}, although there is evidence that under- or over-production of glucocorticosteroids interfere with the normal development of the brain [@LupienMcEwenGunnar:2009].
-While we did find sex-specific evidence of an effect, their clinical relevance is questionable.
-
-These findings should be interpreted in light of the following limitations and strengths.
-
-Limitations include the cross-sectional design of the present study.
-Importantly, the non-persistent \acr{edcs} were measured in a pool of night and morning urine samples before the clinical visit, whereas the glucocorticosteroids were measured in the night urine sample only.
-Although we included a wide range of confounders there is the possibility, as with other observational studies, of residual confounding, which might lead to a bias away from the null.
-Some of the confounders indicated in the adjustment sets had to be remove due to large fractions of missing values.
-There is further the possibility of misspecification of the outcome model, although we included a spline of the exposure to relax some of the linearity assumptions.
-The use of more data-adaptive learners was excluded due to the relatively small sample size.
-We finally acknowledge the possibility that some of chemicals might not act independently (mixture effect).
-Further research is thus warranted.
-
-Strengths of the present study include the use of pooled urine samples for chemical assessment, since it is known that these specific \acr{edcs} have very short half-lives [@CasasBasaganaSakhi:2018].
-We decided to model both the *treatment* mechanisms, for the estimation of balancing weights, and the outcomes, with traditional covariates adjustment, to try to obtain *doubly robust* effect estimates.
-Finally, we decided not to interpret our results by focusing on the estimated coefficients of the regression models, but by making use of the g-computation procedure and estimate marginal contrasts.
-
-In conclusion, in a study of almost 1,000 children from 6 European birth cohorts, we observed that (i) exposure to non-persistent \acr{edcs} in childhood might have short-term effects on \acr{hitrtse} in childhood, (ii) exposure to non-persistent \acr{edcs} in childhood might disrupt the \acr{hpa} axis in childhood, and (iii) disruption of the \acr{hpa} axis in childhood might have short-term, sex-specific effects on \acr{hitrtse}.
-Future studies should investigate how glucocorticosteroids might mediate the adverse effects of exposure to \acr{edcs} on childhood neurodevelopment in larger populations.
-
-{{< pagebreak >}}
-
-# References {.unnumbered}
-
-::: {#refs}
-:::
-
-{{< pagebreak >}}
-
-{{< include _tables.qmd >}}
-
-{{< pagebreak >}}
-
-{{< include _figures.qmd >}}
-
-{{< pagebreak >}}
-
-{{< include _si.qmd >}}

### 2024-02-06 output/paper/manuscript.qmd
Modified almost all text based on MV. Added authors and roles.

@@
 \acr{opps} are one group of \acr{edcs} that may have a deleterious effect on neurodevelopment, although the few studies assessing exposure during childhood and through the use of biomarkers suffered from a series of limitations, including a small sample size and few \acr{opp} metabolites measured [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
-Exposure to phthalates and their metabolites has been associated to several adverse neurodevelopmental outcomes, but these studies were limited to few phthalate metabolites and small study populations [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
+Exposure to phthalates and their metabolites during childhood and early adolescence has been associated to several adverse neurodevelopmental outcomes, but these studies were limited to few phthalate metabolites and small study populations [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
 The effects of exposure to \acr{bpa} during childhood on cognitive functions are still unclear [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
@@
 Moreover, little is known about the biological mechanisms of action [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
-There is also evidence that exposure to certain \acr{edcs}, specifically phthalates, might interfere with the \acr{hpa} axis and might interact with the glucocorticoid receptor [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023].
-
+There is some toxicological evidence, however, that exposure to certain \acr{edcs}, specifically phthalates, might interfere with the \acr{hpa} axis and might interact with the glucocorticoid receptor [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023].
 The HPA axis, which can be activated by stress, is responsible for the production of glucocorticosteroids.
@@
 Taken together, these results suggest that the negative influence of exposure to certain \acr{edcs} on neurodevelopmental outcomes, might be mediated, at least partially, by disruption of the \acr{hpa} axis' homeostasis.
-In the present study, we thus estimated associations between 1) non-persistent \acr{edcs} and attention, 2) non-persistent \acr{edcs} and glucocorticosteroids, and 3) glucocorticosteroids and attention, using the parametric g-formula and marginal contrasts, in children of a large prospective birth cohort in Europe.
+In the present study, we thus estimated associations between 1) non-persistent \acr{edcs} and attention, 2) non-persistent \acr{edcs} and glucocorticosteroids, and 3) glucocorticosteroids and attention, using the parametric g-formula and marginal contrasts, in children of a large cohort in Europe.
 
@@
 
-The \acr{helix} is an ongoing project which aims to characterize early-life exposures and their potential association with endogenous biomarkers and health outcomes [@VrijheidSlamaRobinson:2014].
-It consists of six existing population-based birth cohort studies across Europe: BiB (Born in Bradford, UK) [@WrightSmallRaynor:2013], EDEN (Study of determinants of pre- and postnatal developmental, France) [@HeudeForhanSlama:2016], INMA (Environment and Childhood, Spain) [@GuxensBallesterEspada:2012], KANC (Kaunas Cohort, Lithuania) [@GrazulevicieneDanileviciuteNadisauskiene:2009], MoBa (The Norwegian Mother and Child Cohort Study, Norway) [@MagnusIrgensHaug:2006], and Rhea (Mother–Child Cohort in Crete, Greece) [@ChatziPlanaDaraki:2009], for a total of 32,000 mother-child pairs.
-A \acr{helix} subcohort of 1,200 mother-child pairs was fully characterized for the external and internal exposome, including exposure and omics biomarkers during childhood.
+The \acr{helix} project aims to characterize early-life exposures and their potential association with endogenous biomarkers and health outcomes [@VrijheidSlamaRobinson:2014].
+It consists of six existing population-based birth cohort studies across Europe: BiB (Born in Bradford, UK) [@WrightSmallRaynor:2013], EDEN (Study of determinants of pre- and postnatal developmental, France) [@HeudeForhanSlama:2016], INMA (Environment and Childhood, Spain) [@GuxensBallesterEspada:2012], KANC (Kaunas Cohort, Lithuania) [@GrazulevicieneDanileviciuteNadisauskiene:2009], MoBa (The Norwegian Mother and Child Cohort Study, Norway) [@MagnusIrgensHaug:2006], and Rhea (Mother–Child Cohort in Crete, Greece) [@ChatziPlanaDaraki:2009].
+The \acr{helix} subcohort of 1,301 mother-child pairs was fully characterized for the external and internal exposome, including exposure and omics biomarkers during childhood [@MaitreBontCasas:2018].
 Eligibility criteria for inclusion in the \acr{helix} subcohort included: a) age 6-11 years, with a preference for 7-9 years; b) availability of sufficient stored pregnancy blood and urine samples; c) availability of complete address history from first to last follow-up; d) no serious health problems, which might affect the results of the clinical testing.
-Further information can be found in @MaitreBontCasas:2018.
-
 Ethical permission was obtained from the relevant authorities in the corresponding country.
@@
 ## Variables {#sec-vars}
-
-### Confounders {#sec-confounders}
-
-For each research question, defined by a specific type of exposure and outcome, the minimal set of covariates for inclusion in the analyses was selected on the basis of a \acr{dag} built with DAGitty [@TextorvanderZanderGilthorpe:2016] and ggdag [@Barrett:2023].
-The sets of covariates were selected to estimate the total effect of the exposure on the outcome.
-Further, each minimal adjustment set was *augmented* with precision covariates, defined as the set of parents variable of the outcome that are not parents of the exposure.
-The adjustment sets are provided in the Supplementary Material as text files compatible with DAGitty.
-
-- For RQ1 I used creatinine values from HELIX. For RQ3 the ones from the steroids dataset. For RQ2, I included in the model both variables.
 
@@
     dplyr::filter(class == class_) |>
-    dplyr::select(chem_id, full_name) |>
+    dplyr::select(short_name, full_name) |>
     dplyr::arrange(full_name) |>
     dplyr::mutate(
-      ret = glue::glue("{full_name} ({chem_id})")
+      ret = glue::glue("{full_name} ({short_name})")
     )
@@
 
-Children were assessed between December 2013 and February 2016, and included neurological testing and urine collection.
+Children were assessed between December 2013 and February 2016, and assessments included neurological testing and urine collection.
 Urine samples of the night before and the first morning void on the day of the visit were combined to provide a more reliable exposure assessment.
-Non-persistent \acr{edcs} assessed in urine samples from children included phthalate metabolites, phenols, and \acr{op} pesticide metabolites.
+Non-persistent \acr{edcs} assessed in the urine samples included phthalate metabolites, phenols, and \acr{op} pesticide metabolites.
 A list of the environmental chemicals determined in urine samples and used for the present study is given in @supptbl-info-chems.
@@
 The laboratory protocols for the analysis are described elsewhere [@MarcosRenauCasals:2014; @Gomez-GomezPozo:2020].
-Of the 1,004 urine samples, 980 children were matched to the \acr{helix} subcohort.
+Measurements were available for 980 children of the \acr{helix} subcohort.
 
@@
 
-Neurodevelopmental outcomes were assessed with standardized, non-linguistic, and culturally blind computer tests, including the \acr{ant} [@RuedaFanMcCandliss:2004].
+Neurodevelopmental outcomes were assessed with standardized, non-linguistic, and culturally blind computer tests, including the \acr{ant} [@RuedaFanMcCandliss:2004], which provides a measure of efficiency in three different functions of attention: alerting, orienting, and executive attention.
+The tests were administered in a standardized way, and with minimal interference from the field workers.
 Further information can be found in @MaitreBontCasas:2018.
-Briefly, it is a computerized test that provides a measure of efficiency in three different functions of attention: alerting, orienting, and executive attention.
 The outcome of interest for the present study is the \acr{hitrtse} [@SunyerEsnaolaAlvarez-Pedrerol:2015], a measure of response speed consistency throughout the test.
 A high \acr{hitrtse} indicates highly variable reactions, and is considered a measure of inattentiveness.
+
+### Confounders {#sec-confounders}
+
+For each research question, defined by a specific type of exposure and outcome, the minimal set of covariates for inclusion in the analyses was selected on the basis of a \acr{dag} built with DAGitty [@TextorvanderZanderGilthorpe:2016] and ggdag [@Barrett:2023].
+The sets of covariates were selected to estimate the total effect of the exposure on the outcome.
+Further, each minimal adjustment set was *augmented* with precision covariates, defined as the set of parents variable of the outcome that are not parents of the exposure.
+The adjustment sets are provided in the Supplementary Material as text files compatible with DAGitty.
+Codebooks for the used covariates, by research question, are provided in @supptbl-codebook-1, @supptbl-codebook-2, and @supptbl-codebook-3.
+
+- For RQ1 I used creatinine values from HELIX. For RQ3 the ones from the steroids dataset. For RQ2, I included in the model both variables.
 
@@
 Missing values in the clinical outcome were imputed similarly using kNN.
-Similarly, we natural log-transformed them to improve model fit, assessed with posterior predictive checks.
+We natural log-transformed these to improve model fit, assessed with posterior predictive checks.
 Values of the clinical outcome were expressed in \acr{mss}.
@@
 
-Stabilized balancing weights were estimated using the energy method available in the `WeightIt` R package [@Greifer:2023].
+To reduce the effect of measured confounders on the exposure-outcome association, stabilized balancing weights were estimated using the energy method available in the `WeightIt` R package [@Greifer:2023].
 This methods estimates weights by minimizing an energy statistic related to covariate balance [@HulingGreiferChen:2023], thus avoiding the need to specify a parametric model.
@@
 
-## Participants {#sec-res-participants}
-
-In total, 1,297 children were included in the dataset.
+In total, 1,297 children were included in the study population.
 Of these, 976 had information also about the glucocorticosteroids.
 The sample consisted of 55% girls.
-The median \acr{hitrtse} was 300 ms (\acr{iqr}, 231-368), with lower median values for EDEN, MOBA, and SAB, corresponding to the cohorts with older children.
+The median \acr{hitrtse} was 300 ms (\acr{iqr}, 231-368), with lower median values for EDEN, MOBA, and INMA, corresponding to the cohorts with older children.
 At the time of visit, the median age of the children was 8.06 years.
-The children were mostly Caucasian (90%) or Pakistani (6.2%).
+The children were mostly Caucasian (90%), and the largest minority were of Pakistani origin (6.2%).
 @tbl-pop-desc provides descriptive statistics for the outcome and covariates, by cohort and overall.
-Codebooks for the used covariates, by research question, are provided in @supptbl-codebook-1, @supptbl-codebook-2, and @supptbl-codebook-3.
-
-## Descriptive data {#sec-res-pop-desc}
 
@@
 
-## Main results {#sec-res-main}
-
-The effective sample sizes before and after balancing weights estimation are presented in Tables [-@supptbl-balance-1; -@supptbl-balance-2; -@supptbl-balance-3], while basic summary statistics of the estimated balancing weights are presented in Tables [-@supptbl-weights-1; -@supptbl-weights-2; -@supptbl-weights-3].
+The effective sample sizes before and after balancing weights estimation are presented in Supplementary Tables [-@supptbl-balance-1; -@supptbl-balance-2; -@supptbl-balance-3], while basic summary statistics of the estimated balancing weights are presented in Supplementary Tables [-@supptbl-weights-1; -@supptbl-weights-2; -@supptbl-weights-3].
 As expected, the median value of the weights for each exposure was close to $1.00$.
@@
 @Fig-marginal-1 presents the forest plot for the marginal contrasts of the non-persistent \acr{edcs} on \acr{hitrtse}.
-Overall, a cohort-specific increase in the levels of the exposures from the 10th to the 90th percentiles was associated with a positive marginal contrast, indicating an increase in the values of \acr{hitrtse}.
+For most \acr{edcs}, a cohort-specific increase in the levels of the exposures from the 10th to the 90th percentiles was associated with a positive marginal contrast, indicating an increase in the values of \acr{hitrtse}.
 Most of the \acr{ci}s included the null effect, though.
@@
 @Fig-marginal-2 presents the forest plot for the marginal contrasts of the non-persistent \acr{edcs} on cortisone production, cortisol production, and corticosterone production.
-Overall, a cohort-specific increase in the levels of the exposures from the 10th to the 90th percentiles was associated with a positive marginal contrast, indicating an increase in the total production of these metabolites, with the exception of BUPA, which was associated with negative marginal contrasts for all three outcomes, and MiBP, which was associated with a negative marginal contrast for total cortisone production only.
+For most \acr{edcs}, a cohort-specific increase in the levels of the exposures from the 10th to the 90th percentiles was associated with a positive marginal contrast, indicating an increase in the total production of these metabolites.
+Exceptions were BUPA, which was associated with negative marginal contrasts for all three outcomes, and MiBP, which was associated with a negative marginal contrast for total cortisone production only.
 The majority of the effects for the phenols and phthalate metabolites included the null.
@@
 
-## Other analyses {#sec-res-other}
-
-### Effect modification by sex
+## Effect modification by sex
 
@@
 
-Basic summary statistics of the estimated balancing weights for effect modification are presented in Tables [-@supptbl-weights-1sa; -@supptbl-weights-2sa; -@supptbl-weights-3sa].
+Basic summary statistics of the estimated balancing weights for effect modification are presented in Supplementary Tables [-@supptbl-weights-1sa; -@supptbl-weights-2sa; -@supptbl-weights-3sa].
 As expected, the median value of the weights for each exposure was close to $1.00$.
@@
 
-### Sensitivity to exposure assessment
-
 # Discussion {#sec-discussion}
 
-The impact of exposure to (non-persistent) \acr{edcs} on human health has attracted considerable research interest.
-While research in this area has mainly investigated the effects of prenatal exposure on child neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022], little is still known about later-in-life exposure and childhood neurodevelopment.
-In this study, consisting of almost 1,000 children from 6 European birth cohorts, we observed that short-term childhood exposure to non-persistent \acr{edcs} had negative effects on \acr{hitrtse} and total production of cortisol, cortisone, and corticosterone, with the majority of the \acr{ci}s including the null effect.
+The impact of exposure to \acr{edcs} on human health has attracted considerable research interest.
+While research in this area has mainly investigated the effects of prenatal exposure on child neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022], little is still known about childhood exposure.
+In this study, consisting of 1,297 children from 6 European birth cohorts, we observed that short-term childhood exposure to non-persistent \acr{edcs} had negative effects on \acr{hitrtse} and total production of cortisol, cortisone, and corticosterone, although the majority of the \acr{ci}s included the null effect.
 Increased production of these glucocorticosteroids did not seem to affect \acr{hitrtse}.
@@
 Few studies have looked into different classes of \acr{edcs} [@ShoaffCoullWeuve:2020 in relation with the Conners Attention Deficit Scale and the Behavior Assessment System for Children; @OhKimKannan:2023 in relation with \acr{adhd} symptoms].
-Overall, and consistent with our results, these studies seem to provide further evidence of the negative effects of several \acr{edcs} on neurodevelopment in children.
-While not all these studies have investigated effect modification by sex, it seems that these negative effects are stronger in males.
+Overall, and consistent with our results, these studies seem to provide further evidence of the adverse effects of several \acr{edcs} on neurodevelopment in children.
+While not all these studies have investigated effect modification by sex, it seems that these adverse effects are stronger in males.
 A major limitation of these studies is the reliance on spot urine samples, that might not be representative of long-term exposures.
 
-Our results are consistent with prior research that associated exposure to certain \acr{edcs} with higher levels of cortisol [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023].
+Our results are consistent with prior epidemiological research that associated exposure to certain \acr{edcs} with higher levels of cortisol [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023].
 There are some differences, though.
@@
 
-We are also not aware of prior studies specifically investigating the effects of elevated levels of glucocorticosteroids in relation to \acr{hitrtse}, although there is evidence that under- or over-production of glucocorticosteroids interfere with the normal development of the brain [@LupienMcEwenGunnar:2009].
+We are also not aware of prior epidemiological studies specifically investigating the effects of elevated levels of glucocorticosteroids in relation to \acr{hitrtse}, although there is evidence that under- or over-production of glucocorticosteroids interfere with the normal development of the brain [@LupienMcEwenGunnar:2009].
 While we did find sex-specific evidence of an effect, their clinical relevance is questionable.
 
-These findings should be interpreted in light of the following limitations and strengths.
-
+Our findings should be interpreted in light of the following limitations and strengths.
 Limitations include the cross-sectional design of the present study.
-Importantly, the non-persistent \acr{edcs} were measured in a pool of night and morning urine samples before the clinical visit, whereas the glucocorticosteroids were measured in the night urine sample only.
+Importantly, the non-persistent \acr{edcs} were measured in a pool of night and morning urine samples before the clinical visit, to represent exposure over the previous day, whereas the glucocorticosteroids were measured in the night urine sample.
 Although we included a wide range of confounders there is the possibility, as with other observational studies, of residual confounding, which might lead to a bias away from the null.
@@
 
-In conclusion, in a study of almost 1,000 children from 6 European birth cohorts, we observed that (i) exposure to non-persistent \acr{edcs} in childhood might have short-term effects on \acr{hitrtse} in childhood, (ii) exposure to non-persistent \acr{edcs} in childhood might disrupt the \acr{hpa} axis in childhood, and (iii) disruption of the \acr{hpa} axis in childhood might have short-term, sex-specific effects on \acr{hitrtse}.
+In conclusion, in a study of 1,297 children from 6 European birth cohorts, we observed that (i) exposure to non-persistent \acr{edcs} in childhood might have short-term effects on \acr{hitrtse} in childhood, (ii) exposure to non-persistent \acr{edcs} in childhood might disrupt the \acr{hpa} axis in childhood, and (iii) disruption of the \acr{hpa} axis in childhood might have short-term, sex-specific effects on \acr{hitrtse}.
 Future studies should investigate how glucocorticosteroids might mediate the adverse effects of exposure to \acr{edcs} on childhood neurodevelopment in larger populations.

### 2024-02-06 _targets_res.R

@@ -7,7 +7,7 @@ targets::tar_option_set(
   format = "qs"
 )
 path_store <- Sys.getenv("path_store")
-num_digits_est <- 3
+num_digits_est <- 2
 num_digits_sig <- 2
 sa_var_ <- "e3_sex"
 

### 2024-02-07 output/paper/_tables.qmd
Added 2 metabolites in sheet that I forgot about. Modifying tables according to MV v1.

@@
 table1
+  # gt::cols_width(
+  #   label ~ gt::px(300)
+  # ) |>
+  # gt::tab_style(
+  #   style = gt::cell_text(size = gt::pct(90)),
+  #   locations = gt::cells_body(
+  #     columns = dplyr::everything()
+  #   )
+  # )
 ```

### 2024-02-07 output/paper/manuscript.qmd
Added 2 metabolites in sheet that I forgot about. Modifying tables according to MV v1.

@@
   docx:
-    reference-doc: reference_document.docx
+    # reference-doc: reference_document.docx
+    page-width: 8.3
     citeproc: true
@@
 id_var <- "HelixID"
-num_digits_est = 3
+num_digits_est = 2
 num_digits_sig = 2

### 2024-02-07 R/proc_res.R

@@ -197,9 +197,6 @@ tbl_desc_pop <- function(num_digits_est, num_digits_sig) {
     gtsummary::as_gt() |>
     gt::opt_footnote_marks(
       marks = "letters"
-    ) |>
-    gt::tab_options(
-      page.orientation = "landscape"
     )
   
   # Save table
@@ -220,42 +217,38 @@ tbl_desc_pop <- function(num_digits_est, num_digits_sig) {
 #' @return
 #' @export
 tbl_desc_vars <- function() {
+  params_dat <- params(is_hpc = Sys.getenv("is_hpc"))
   path_store <- Sys.getenv("path_store")
   
-  ## Load objects
-  rq <- "02"
-  targets::tar_load(
-    "rq02_desc_data_exp",
-    store = paste0(path_store, rq)
-  )
-  
-  ## Save table
-  path <- "results/tables/"
-  name <- "tbl_desc_chemicals.docx"
-  # gt::gtsave(
-  #   data = rq02_desc_data_exp$step3 |>
-  #     gtsummary::as_gt(),
-  #   filename = paste0(path, name)
-  # )
-  ##############################################################################
-  
-  ## Load objects
-  targets::tar_load(
-    "rq02_desc_data_out",
-    store = paste0(path_store, rq)
-  )
-  
-  ## Save table
-  path <- "results/tables/"
-  name <- "tbl_desc_metabolites.docx"
-  # gt::gtsave(
-  #   data = rq02_desc_data_out$step3 |>
-  #     gtsummary::as_gt(),
-  #   filename = paste0(path, name)
-  # )
-  
+  dat_request <- load_dat_request()
   edcs <- myphd::edcs_information()
-  desc_chems <- rq02_desc_data_exp$step3 |>
+  metabolites <- load_steroids()$metabolome
+  steroids <- readODS::read_ods("docs/steroids.ods")
+  dat <- list()
+  
+  # EDCs
+  dat$exposures <- dat_request$dat |>
+    tidylog::select(
+      params_dat$variables$identifier,
+      dplyr::any_of(params_dat$variables[["rq1"]]$exposures)
+    )
+  tbl_edcs <- dat$exposures |>
+    tidylog::rename_with(
+      .fn = \(x) gsub("hs_|_c", "", x)
+    ) |>
+    tidylog::select(
+      -dplyr::any_of(c("cohort", "HelixID"))
+    ) |>
+    gtsummary::tbl_summary(
+      statistic = list(
+        gtsummary::all_continuous() ~ c(
[…]

### 2024-02-07 output/paper/_tables.qmd
Table ratios for entire population and HELIX SC only. No by cohort, tho.

@@
 
-scores |>
+# Entire population (1004)
+desc_sums_all <- scores |>
+  dplyr::select(-cohort) |>
   gtsummary::tbl_summary(
-    by = "cohort"
+    statistic = list(
+      gtsummary::all_continuous() ~ c(
+        "{median} ({p25}, {p75}); {N_miss} ({p_miss})"
+      )
+    ),
+    missing = "no",
+    digits = dplyr::everything() ~ 1
+  )
+
+# Intersection with HELIX subcohort (976)
+targets::tar_load(
+  dplyr::contains("load"),
+  store = "~/mounts/rstudioserver/data/data_paper3/_targets/_targetsRQ2/"
+)
+desc_sums_intersection <- create_steroid_scores(rq2_load_dat$outcome) |>
+  dplyr::select(
+    cortisol_production,
+    cortisone_production,
+    corticosterone_production
   ) |>
-  gtsummary::add_overall() |>
+  tidylog::rename(
+    "cortisol production" = "cortisol_production",
+    "cortisone production" = "cortisone_production",
+    "corticosterone production" = "corticosterone_production"
+  ) |>
+  gtsummary::tbl_summary(
+    statistic = list(
+      gtsummary::all_continuous() ~ c(
+        "{median} ({p25}, {p75}); {N_miss} ({p_miss})"
+      )
+    ),
+    missing = "no",
+    digits = dplyr::everything() ~ 1
+  )
+
+gtsummary::tbl_merge(
+  tbls = list(desc_sums_all, desc_sums_intersection)
+) |>
+  gtsummary::modify_spanning_header(
+    dplyr::everything() ~ NA_character_
+  ) |>
   gtsummary::as_gt() |>
+  gt::tab_footnote(
+    footnote = "Measurements available for the HELIX subcohort.",
+    locations = gt::cells_column_labels(8)
+  ) |>
   gt::opt_footnote_marks(

### 2024-02-07 R/proc_res.R

@@ -191,22 +191,15 @@ tbl_desc_pop <- function(num_digits_est, num_digits_sig) {
           "{median} ({p25}, {p75})"
         ),
         gtsummary::all_categorical() ~ "{n} ({p}%)"
-      )
+      ),
+      digits = dplyr::everything() ~ 1,
+      missing = "ifany"
     ) |>
     gtsummary::add_overall() |>
     gtsummary::as_gt() |>
     gt::opt_footnote_marks(
       marks = "letters"
     )
-  
-  # Save table
-  path <- "results/tables/"
-  name <- "tbl_desc_pop.docx"
-  # gt::gtsave(
-  #   data = desc_covars |>
-  #     gtsummary::as_gt(),
-  #   filename = paste0(path, name)
-  # )
   
   return(desc_covars)
 } # End function tbl_desc_pop
@@ -245,7 +238,8 @@ tbl_desc_vars <- function() {
           "{median} ({p25}, {p75}); {N_miss} ({p_miss})"
         )
       ),
-      missing = "no"
+      missing = "no",
+      digits = dplyr::everything() ~ 1
     )
   
   desc_chems <- tbl_edcs |>
@@ -312,7 +306,8 @@ tbl_desc_vars <- function() {
         )
       ),
       missing = "ifany",
-      by = "cohort"
+      by = "cohort",
+      digits = dplyr::everything() ~ 1
     ) |>
     gtsummary::add_overall() |>
     gtsummary::as_gt()

### 2024-02-07 output/paper/_si.qmd
Kept working on papers from MV v1.

@@
 
-### Description of chemicals
+### Description of endocrine disruptors
 
@@
 
-### Description of metabolites
+### Description of glucocorticosteroids
+
+::: {#supptbl-met-desc}
+```{r met-desc}
+desc_chems_mets$desc_mets
+```
+**Participants glucocorticosteroids concentrations, by cohort and overall (HELIX subcohort; 2013-2016).**
+:::
 
@@
 
-### Description of chemicals
+### Study populations
+
+::: {#suppfig-flow-pop}
+```{mermaid}
+flowchart TB
+  helixsc["HELIX subcohort\n(N = 1,301)"] --> edcs["HELIX data EDCs\n(N = 1,297)"]
+  helixsc --> corts["HELIX data glucocorticosteroids\n(N = 1,004)"]
+  edcs --> inter["HELIX data EDCs and glucocorticosteroids\n(N = 976)"]
+  corts --> inter
+  edcs -.-> rq1(["HRT-SE ~ EDCs"])
+  inter -.-> rq2(["metabolites ~ EDCs"])
+  inter -.-> rq3(["HRT-SE ~ metabolites"])
+```
+Flowchart describing thew sample size for each research question.
+:::
+
+### Description of endocrine disruptors
 
@@
 
-### Description of metabolites
+### Description of glucocorticosteroids
 

### 2024-02-07 output/paper/manuscript.qmd
Kept working on papers from MV v1.

@@
     citeproc: true
+    mermaid:
+      theme: neutral
   # pdf:
@@
 The laboratory protocols for the analysis are described elsewhere [@MarcosRenauCasals:2014; @Gomez-GomezPozo:2020].
-Measurements were available for 980 children of the \acr{helix} subcohort.
 
@@
 
-For each research question, defined by a specific type of exposure and outcome, the minimal set of covariates for inclusion in the analyses was selected on the basis of a \acr{dag} built with DAGitty [@TextorvanderZanderGilthorpe:2016] and ggdag [@Barrett:2023].
+For each research question, defined by a specific type of exposure and outcome, the minimal set of covariates for inclusion in the analyses was selected on the basis of a \acr{dag} built with `DAGitty` [@TextorvanderZanderGilthorpe:2016] and `ggdag` [@Barrett:2023].
 The sets of covariates were selected to estimate the total effect of the exposure on the outcome.
@@
 The adjustment sets are provided in the Supplementary Material as text files compatible with DAGitty.
-Codebooks for the used covariates, by research question, are provided in @supptbl-codebook-1, @supptbl-codebook-2, and @supptbl-codebook-3.
+Codebooks for the used covariates, by research question, are provided in Supplementary Tables [-@supptbl-codebook-1; -@supptbl-codebook-2; -@supptbl-codebook-3].
 
@@
 
-In total, 1,297 children were included in the study population.
-Of these, 976 had information also about the glucocorticosteroids.
+Of the 1,301 children of the \acr{helix} subcohort, 1,297 had measurements of the non-persistent \acr{edcs}.
+Measurements of the glucocorticosteroids were available for 1,004 children, of which 980 were matched to the \acr{helix} subcohort.
+Measurements of both non-persistent \acr{edcs} and glucocorticosteroids were available for 976 children of the subcohort.
+A flowchart describing the sample size for each research question is presented in @suppfig-flow-pop.
 The sample consisted of 55% girls.
@@
 
-Levels of unprocessed non-persistent \acr{edcs} and glucocorticosteroids, by cohort and overall, are presented in @tbl-edc-desc, @tbl-met-desc, and @tbl-met-new-desc.
+Levels of unprocessed non-persistent \acr{edcs} and glucocorticosteroids are presented in @tbl-edc-desc, @tbl-met-new-desc, and @supptbl-met-desc.
 

### 2024-02-12 output/paper/_figures.qmd
Switching to new syntax acr. Adding comments from MV v1.

--- 2024-01-02 output/paper/_figures.qmd
@@
 ```
-**Marginal contrasts (exposures: \acr{edcs}; outcome: \acr{hitrtse}) (HELIX Study; 2013-2016).**
-Circles indicate effect estimates. Solid lines indicate the $95\%$ \acr{ci}.
+**Marginal contrasts for the effect of a decrease from the 90th to the 10th percentile of the {{< acr edcs >}} on {{< acr hitrtse >}} (HELIX subcohort; 2013-2016).** 
+Circles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}.
 :::
@@
 ```
-**Marginal contrasts (exposures: \acr{edcs}; outcomes: glucocorticosteroids) (HELIX Study; 2013-2016).**
-Circles, triangles, and squares indicate effect estimates. Solid lines indicate the $95\%$ \acr{ci}.
+**Marginal contrasts for the effect of a decrease from the 90th to the 10th percentile of the {{< acr edcs >}} on the glucocorticosteroids (HELIX subcohort; 2013-2016).** 
+Circles, triangles, and squares indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}.
 :::
@@
 ```
-**Marginal contrasts (exposures: glucocorticosteroids; outcome: \acr{hitrtse}) (HELIX Study; 2013-2016).**
-Circles indicate effect estimates. Solid lines indicate the $95\%$ \acr{ci}.
+**Marginal contrasts for the effect of a decrease from the 90th to the 10th percentile of the glucocorticosteroids on {{< acr hitrtse >}} (HELIX subcohort; 2013-2016).** 
+Circles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}.
 :::

### 2024-02-12 output/paper/_si.qmd
Switching to new syntax acr. Adding comments from MV v1.

@@
 ```
-**Codebook for the covariates used in the estimation of the marginal comparisons of \acr{edcs} on \acr{hitrtse}.**
+**Codebook for the covariates used in the estimation of the marginal comparisons of {{< acr edcs >}} on {{< acr hitrtse >}}.**
 :::
@@
 ```
-**Codebook for the covariates used in the estimation of the marginal comparisons of \acr{edcs} on the glucocorticosteroids.**
+**Codebook for the covariates used in the estimation of the marginal comparisons of {{< acr edcs >}} on the glucocorticosteroids.**
 :::
@@
 ```
-**Codebook for the covariates used in the estimation of the marginal comparisons of the glucocorticosteroids on \acr{hitrtse}.**
+**Codebook for the covariates used in the estimation of the marginal comparisons of the glucocorticosteroids on {{< acr hitrtse >}}.**
 :::
@@
 ```
-**Information about non-persistent EDCs, including the full compound name, the standard symbol, the used variable name, the identifier from PubChem, and the parental compound.**
+**Information about non-persistent {{< acr edcs >}}, including the full compound name, the standard symbol, the used variable name, the identifier from PubChem, and the parental compound.**
 :::
@@
 ```
-**Effective sample size before and after balancing weights estimation (exposures: \acr{edcs}; outcome: \acr{hitrtse}) (HELIX subcohort; 2013-2016).**
+**Effective sample size before and after balancing weights estimation (exposures: {{< acr edcs >}}; outcome: {{< acr hitrtse >}}) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Effective sample size before and after balancing weights estimation (exposures: \acr{edcs}; outcomes: glucocorticosteroids) (HELIX subcohort; 2013-2016).**
+**Effective sample size before and after balancing weights estimation (exposures: {{< acr edcs >}}; outcomes: glucocorticosteroids) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Effective sample size before and after balancing weights estimation (exposures: glucocorticosteroids; outcome: \acr{hitrtse}) (HELIX subcohort; 2013-2016).**
+**Effective sample size before and after balancing weights estimation (exposures: glucocorticosteroids; outcome: {{< acr hitrtse >}}) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Summary statistics of the estimated balancing weights (exposures: \acr{edcs}; outcome: \acr{hitrtse}) (HELIX subcohort; 2013-2016).**
+**Summary statistics of the estimated balancing weights (exposures: {{< acr edcs >}}; outcome: {{< acr hitrtse >}}) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Summary statistics of the estimated balancing weights (exposures: \acr{edcs}; outcomes: glucocorticosteroids) (HELIX subcohort; 2013-2016).**
+**Summary statistics of the estimated balancing weights (exposures: {{< acr edcs >}}; outcomes: glucocorticosteroids) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Summary statistics of the estimated balancing weights (exposures: glucocorticosteroids; outcome: \acr{hitrtse}) (HELIX subcohort; 2013-2016).**
+**Summary statistics of the estimated balancing weights (exposures: glucocorticosteroids; outcome: {{< acr hitrtse >}}) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Summary statistics of the estimated balancing weights for effect modification (exposures: \acr{edcs}; outcome: \acr{hitrtse}; modifier: sex) (HELIX subcohort; 2013-2016).**
+**Summary statistics of the estimated balancing weights for effect modification (exposures: {{< acr edcs >}}; outcome: {{< acr hitrtse >}}; modifier: sex) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Summary statistics of the estimated balancing weights for effect modification (exposures: \acr{edcs}; outcomes: glucocorticosteroids; modifier: sex) (HELIX subcohort; 2013-2016).**
+**Summary statistics of the estimated balancing weights for effect modification (exposures: {{< acr edcs >}}; outcomes: glucocorticosteroids; modifier: sex) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Summary statistics of the estimated balancing weights for effect modification (exposures: glucocorticosteroids; outcome: \acr{hitrtse}; modifier: sex) (HELIX subcohort; 2013-2016).**
+**Summary statistics of the estimated balancing weights for effect modification (exposures: glucocorticosteroids; outcome: {{< acr hitrtse >}}; modifier: sex) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Measurement classification of \acr{edcs}, by cohort (HELIX subcohort; 2013-2016).** 
+**Measurement classification of {{< acr edcs >}}, by cohort (HELIX subcohort; 2013-2016).** 
 Coding: 1, quantifiable; 2, <LOD; 3, interference or out of range; 4. not analysed.
@@
 ```
-**Marginal contrasts for effect modification (exposures: \acr{edcs}; outcome: \acr{hitrtse}; modifier: sex) (HELIX subcohort; 2013-2016).** 
-Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ \acr{ci}.
+**Marginal contrasts for effect modification by sex of a decrease from the 90th to the 10th percentile of the {{< acr edcs >}} on {{< acr hitrtse >}} (HELIX subcohort; 2013-2016).** 
+Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}.
 :::
@@
 ```
-**Marginal contrasts for effect modification (exposures: \acr{edcs}; outcomes: glucocorticosteroids; modifier: sex) (HELIX subcohort; 2013-2016).** 
-Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ \acr{ci}.
+**Marginal contrasts for effect modification by sex of a decrease from the 90th to the 10th percentile of the {{< acr edcs >}} on the glucocorticosteroids (HELIX subcohort; 2013-2016).** 
+Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}.
 :::
@@
 ```
-**Marginal contrasts for effect modification (exposures: glucocorticosteroids; outcome: \acr{hitrtse}; modifier: sex) (HELIX subcohort; 2013-2016).** 
-Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ \acr{ci}.
+**Marginal contrasts for effect modification by sex of a decrease from the 90th to the 10th percentile of the glucocorticosteroids on {{< acr hitrtse >}} (HELIX subcohort; 2013-2016).** 
+Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}.
 :::

### 2024-02-12 output/paper/_tables.qmd
Kept working on papers from MV v1.
Switching to new syntax acr. Adding comments from MV v1.

@@
 ```
-**Participants endocrine disruptors concentrations, by cohort and overall (HELIX subcohort; 2013-2016).**
+**Participants endocrine disruptors concentrations (HELIX subcohort; 2013-2016).**
 :::
@@
 ## Glucocorticosteroids
-
-::: {#tbl-met-desc}
-```{r met-desc}
-desc_chems_mets$desc_mets
-```
-**Participants glucocorticosteroids concentrations, by cohort and overall (HELIX subcohort; 2013-2016).**
-:::
 
@@
 )
+edcs <- myphd::edcs_information() |>
+  dplyr::filter(!chem_id %in% c("dedtp", "dmdtp")) |>
+  dplyr::select(chem_id, short_name)
 ```
 
-::: {#tbl-hypothesis-1}
-```{r hypothesis-1}
-tidy_res_meffects(
+::: {#tbl-hypothesis-1and2}
+```{r hypothesis-1and2}
+t1 <- tidy_res_meffects(
   df = marginal_hypothesis_1SA,
@@
 )$table
-```
-**Pairwise differences between sex-specific marginal contrasts for the effect of \acr{edcs} on \acr{hitrtse} (HELIX subcohort; 2013-2016).**
-:::
-
-::: {#tbl-hypothesis-2}
-```{r hypothesis-2}
-tidy_res_meffects(
+t2 <- tidy_res_meffects(
   df = marginal_hypothesis_2SA,
@@
 )$table
+
+dplyr::bind_cols(t1$`_data`, t2$`_data`) |>
+  dplyr::select(-`class...4`, -`variable...5`) |>
+  dplyr::rename(
+    class = "class...1",
+    variable = "variable...2"
+  ) |>
+  dplyr::left_join(
+    edcs,
+    by = c("variable" = "chem_id")
+  ) |>
+  dplyr::select(-variable) |>
+  dplyr::rename(variable = "short_name") |>
+  gt::gt(
+    rowname_col = "variable",
+    groupname_col = "class"
+  ) |>
+  gt::tab_style(
+    style = gt::cell_text(
+      weight = "bold"
+    ),
+    locations = gt::cells_row_groups()
+  ) |>
+  gt::tab_footnote(
+    footnote = "Estimate and 95% CI.",
+    locations = gt::cells_column_labels()
+  ) |>
+  gt::opt_footnote_marks(
+    marks = "letters"
+  )
 ```
-**Pairwise differences between sex-specific marginal contrasts for the effect of \acr{edcs} on the glucocorticosteroids (HELIX subcohort; 2013-2016).**
+**Pairwise differences between sex-specific marginal contrasts for the effect of a decrease from the 90th to the 10th percentile of {{< acr edcs >}} on {{< acr hitrtse >}} and the glucocorticosteroids (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Pairwise differences between sex-specific marginal contrasts for the effect of the glucocorticosteroids on \acr{hitrtse} (HELIX subcohort; 2013-2016).**
+**Pairwise differences between sex-specific marginal contrasts for the effect of a decrease from the 90th to the 10th percentile of the glucocorticosteroids on {{< acr hitrtse >}} (HELIX subcohort; 2013-2016).**
 :::

### 2024-02-12 output/paper/manuscript.qmd
Switching to new syntax acr. Adding comments from MV v1.

@@
   glue::glue(
-    "\\acr{{mc}} and $95\\%$ \\acr{{ci}}: {val}"
+    "\{{< acr mc >}} and $95\\%$ \{{< acr ci >}}: {val}"
   )
@@
 
-Epidemiological evidence suggests that exposure to \acr{edcs} may be linked to perturbation of the \acr{hpa} axis, which has a major role in brain development.
+Epidemiological evidence suggests that exposure to {{< acr edcs >}} may be linked to perturbation of the {{< acr hpa >}} axis, which has a major role in brain development.
 We aimed to evaluate the effects of childhood exposure to organophosphate pesticides, phenols, and phthalate metabolites, on urinary glucocorticosteroids and inattention measures in children.
-We used data from the \acr{helix} cohort including 1,297 children aged 6-11.
-We used the parametric g-formula and \acr{mcs} to estimate effects between \acr{edcs}, glucocorticosteroids, and \acr{hitrtse}, a measure of efficiency from the \acr{ant}, and tested for possible effect modification by sex.
-We observed positive \acr{mcs} for exposure increases from the 10th to the 90th percentile for methyl-paraben (`r format_res(meffects1, "mepa", "hitrtse")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "hitrtse")`), oh-MiNP (`r format_res(meffects1, "ohminp", "hitrtse")`), and MEHP (`r format_res(meffects1, "mehp", "hitrtse")`), on \acr{hitrtse}.
-Several \acr{edcs} were also associated with positive \acr{mcs} of cortisone, cortisol, and corticosterone production.
-Increased levels of the glucocorticosteroids had no effect on \acr{hitrtse}, although we found possible effect modification by sex for cortisone and corticosterone production.
-Our results suggest that multiple \acr{edcs} might interfere with inattentiveness outcomes in children and with the homeostasis of the \acr{hpa} axis.
+We used data from the {{< acr helix >}} cohort including 1,297 children aged 6-11.
+We used the parametric g-formula and {{< acr mcs >}} to estimate effects between {{< acr edcs >}}, glucocorticosteroids, and {{< acr hitrtse >}}, a measure of efficiency from the {{< acr ant >}}, and tested for possible effect modification by sex.
+We observed positive {{< acr mc >}}s for exposure increases from the 10th to the 90th percentile for methyl-paraben (`r format_res(meffects1, "mepa", "hitrtse")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "hitrtse")`), oh-MiNP (`r format_res(meffects1, "ohminp", "hitrtse")`), and MEHP (`r format_res(meffects1, "mehp", "hitrtse")`), on {{< acr hitrtse >}}.
+Several {{< acr edcs >}} were also associated with positive {{< acr mc >}}s of cortisone, cortisol, and corticosterone production.
+Increased levels of the glucocorticosteroids had no effect on {{< acr hitrtse >}}, although we found possible effect modification by sex for cortisone and corticosterone production.
+Our results suggest that multiple {{< acr edcs >}} might interfere with inattentiveness outcomes in children and with the homeostasis of the {{< acr hpa >}} axis.
 Further research is necessary to determine whether this can have a clinically significant impact on childhood neurodevelopment.
@@
 The prevalence of several neurodevelopmental disorders has increased in the pediatric population [@GrandjeanLandrigan:2014], and multiple environmental pollutants may play a role in the increased rates of these disorders [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
-\acr{edcs}, ubiquitous chemicals present in many every-day products and diet, are capable of interfering with the endocrine system, and have shown associations with childhood neurodevelopment and behavior [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @HuangChenSu:2015; @CartierWarembourgLeManer-Idrissi:2016; @TewarAuingerBraun:2016; @YuDuChiou:2016; @HuangTsaiChen:2017; @KimHongShin:2017; @FurlongHerringBuckley:2017; @Braun:2017; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019; @ShoaffCoullWeuve:2020; @OhKimKannan:2023; @VilmandBeckBilenberg:2023].
-Although both pregnancy and early childhood are crucial stages of (neuro)development, most of the available literature is focused on the effects of prenatal exposure to \acr{edcs} on child neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
+{{< acr edcs first_use=true >}}, ubiquitous chemicals present in many every-day products and diet, are capable of interfering with the endocrine system, and have shown associations with childhood neurodevelopment and behavior [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @HuangChenSu:2015; @CartierWarembourgLeManer-Idrissi:2016; @TewarAuingerBraun:2016; @YuDuChiou:2016; @HuangTsaiChen:2017; @KimHongShin:2017; @FurlongHerringBuckley:2017; @Braun:2017; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019; @ShoaffCoullWeuve:2020; @OhKimKannan:2023; @VilmandBeckBilenberg:2023].
+Although both pregnancy and early childhood are crucial stages of (neuro)development, most of the available literature is focused on the effects of prenatal exposure to {{< acr edcs >}} on child neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
 
-\acr{opps} are one group of \acr{edcs} that may have a deleterious effect on neurodevelopment, although the few studies assessing exposure during childhood and through the use of biomarkers suffered from a series of limitations, including a small sample size and few \acr{opp} metabolites measured [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
+{{< acr opps >}} are one group of {{< acr edcs >}} that may have a deleterious effect on neurodevelopment, although the few studies assessing exposure during childhood and through the use of biomarkers suffered from a series of limitations, including a small sample size and few {{< acr opp >}} metabolites measured [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
 Exposure to phthalates and their metabolites during childhood and early adolescence has been associated to several adverse neurodevelopmental outcomes, but these studies were limited to few phthalate metabolites and small study populations [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
-The effects of exposure to \acr{bpa} during childhood on cognitive functions are still unclear [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
+The effects of exposure to {{< acr bpa >}} during childhood on cognitive functions are still unclear [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
 
 Moreover, little is known about the biological mechanisms of action [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
-There is some toxicological evidence, however, that exposure to certain \acr{edcs}, specifically phthalates, might interfere with the \acr{hpa} axis and might interact with the glucocorticoid receptor [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023].
+There is some toxicological evidence, however, that exposure to certain {{< acr edcs >}}, specifically phthalates, might interfere with the {{< acr hpa first_use=true >}} axis and might interact with the glucocorticoid receptor [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023].
 The HPA axis, which can be activated by stress, is responsible for the production of glucocorticosteroids.
@@
 
-Taken together, these results suggest that the negative influence of exposure to certain \acr{edcs} on neurodevelopmental outcomes, might be mediated, at least partially, by disruption of the \acr{hpa} axis' homeostasis.
-In the present study, we thus estimated associations between 1) non-persistent \acr{edcs} and attention, 2) non-persistent \acr{edcs} and glucocorticosteroids, and 3) glucocorticosteroids and attention, using the parametric g-formula and marginal contrasts, in children of a large cohort in Europe.
+Taken together, these results suggest that the negative influence of exposure to certain {{< acr edcs >}} on neurodevelopmental outcomes, might be mediated, at least partially, by disruption of the {{< acr hpa >}} axis' homeostasis.
+In the present study, we thus estimated associations between 1) non-persistent {{< acr edcs >}} and attention, 2) non-persistent {{< acr edcs >}} and glucocorticosteroids, and 3) glucocorticosteroids and attention, using the parametric g-formula and {{< mc >}}s, in children of a large cohort in Europe.
 
@@
 
-The \acr{helix} project aims to characterize early-life exposures and their potential association with endogenous biomarkers and health outcomes [@VrijheidSlamaRobinson:2014].
+The {{< acr helix first_use=true >}} project aims to characterize early-life exposures and their potential association with endogenous biomarkers and health outcomes [@VrijheidSlamaRobinson:2014].
 It consists of six existing population-based birth cohort studies across Europe: BiB (Born in Bradford, UK) [@WrightSmallRaynor:2013], EDEN (Study of determinants of pre- and postnatal developmental, France) [@HeudeForhanSlama:2016], INMA (Environment and Childhood, Spain) [@GuxensBallesterEspada:2012], KANC (Kaunas Cohort, Lithuania) [@GrazulevicieneDanileviciuteNadisauskiene:2009], MoBa (The Norwegian Mother and Child Cohort Study, Norway) [@MagnusIrgensHaug:2006], and Rhea (Mother–Child Cohort in Crete, Greece) [@ChatziPlanaDaraki:2009].
-The \acr{helix} subcohort of 1,301 mother-child pairs was fully characterized for the external and internal exposome, including exposure and omics biomarkers during childhood [@MaitreBontCasas:2018].
-Eligibility criteria for inclusion in the \acr{helix} subcohort included: a) age 6-11 years, with a preference for 7-9 years; b) availability of sufficient stored pregnancy blood and urine samples; c) availability of complete address history from first to last follow-up; d) no serious health problems, which might affect the results of the clinical testing.
+The {{< acr helix >}} subcohort of 1,301 mother-child pairs was fully characterized for the external and internal exposome, including exposure and omics biomarkers during childhood [@MaitreBontCasas:2018].
+Eligibility criteria for inclusion in the {{< acr helix >}} subcohort included: a) age 6-11 years, with a preference for 7-9 years; b) availability of sufficient stored pregnancy blood and urine samples; c) availability of complete address history from first to last follow-up; d) no serious health problems, which might affect the results of the clinical testing.
 Ethical permission was obtained from the relevant authorities in the corresponding country.
@@
 Urine samples of the night before and the first morning void on the day of the visit were combined to provide a more reliable exposure assessment.
-Non-persistent \acr{edcs} assessed in the urine samples included phthalate metabolites, phenols, and \acr{op} pesticide metabolites.
+Non-persistent {{< acr edcs >}} assessed in the urine samples included phthalate metabolites, phenols, and {{< acr op >}} pesticide metabolites.
 A list of the environmental chemicals determined in urine samples and used for the present study is given in @supptbl-info-chems.
@@
 
-Neurodevelopmental outcomes were assessed with standardized, non-linguistic, and culturally blind computer tests, including the \acr{ant} [@RuedaFanMcCandliss:2004], which provides a measure of efficiency in three different functions of attention: alerting, orienting, and executive attention.
+Neurodevelopmental outcomes were assessed with standardized, non-linguistic, and culturally blind computer tests, including the {{< acr ant first_use=true >}} [@RuedaFanMcCandliss:2004], which provides a measure of efficiency in three different functions of attention: alerting, orienting, and executive attention.
 The tests were administered in a standardized way, and with minimal interference from the field workers.
 Further information can be found in @MaitreBontCasas:2018.
-The outcome of interest for the present study is the \acr{hitrtse} [@SunyerEsnaolaAlvarez-Pedrerol:2015], a measure of response speed consistency throughout the test.
-A high \acr{hitrtse} indicates highly variable reactions, and is considered a measure of inattentiveness.
+The outcome of interest for the present study is the {{< acr hitrtse first_use=true >}} [@SunyerEsnaolaAlvarez-Pedrerol:2015], a measure of response speed consistency throughout the test.
+A high {{< acr hitrtse >}} indicates highly variable reactions, and is considered a measure of inattentiveness.
 
@@
 
-For each research question, defined by a specific type of exposure and outcome, the minimal set of covariates for inclusion in the analyses was selected on the basis of a \acr{dag} built with `DAGitty` [@TextorvanderZanderGilthorpe:2016] and `ggdag` [@Barrett:2023].
+For each research question, defined by a specific type of exposure and outcome, the minimal set of covariates for inclusion in the analyses was selected on the basis of a {{< acr dag >}} built with `DAGitty` [@TextorvanderZanderGilthorpe:2016] and `ggdag` [@Barrett:2023].
 The sets of covariates were selected to estimate the total effect of the exposure on the outcome.
@@
 
-Concentrations of the glucocorticosteroids were classified as quantifiable, below the \acr{loq}, possible interference or out of range, and not detected.
-For each metabolite, we computed the fraction of values below the \acr{loq} and not detected, both within each cohort and overall.
-We proceeded to impute these values using half the value of the corresponding \acr{lloq}, for those metabolites that had less than `r threshold_within_lod`% of missings within each cohort and `r threshold_overall_lod`% of missings overall.
-Information about the \acr{lloq} for the glucocorticosteroids is provided in @supptbl-lloq-mets.
+Concentrations of the glucocorticosteroids were classified as quantifiable, below the {{< acr loq >}}, possible interference or out of range, and not detected.
+For each metabolite, we computed the fraction of values below the {{< acr loq >}} and not detected, both within each cohort and overall.
+We proceeded to impute these values using half the value of the corresponding {{< acr lloq >}}, for those metabolites that had less than `r threshold_within_lod`% of missings within each cohort and `r threshold_overall_lod`% of missings overall.
+Information about the {{< acr lloq >}} for the glucocorticosteroids is provided in @supptbl-lloq-mets.
 The remaining missing values were imputed using kNN from the `VIM` R package [@KowarikTempl:2016], for those metabolites that had less than `r threshold_within_nan`% of remaining missings within each cohort and `r threshold_overall_nan`% of remaining missings overall.
@@
 We used the `check_predictions` function from the `performance` R package using the default arguments [@LudeckeBen-ShacharPatil:2021].
-Values of cortisol production and cortisone production were expressed in nanograms per millilitre, whereas values of cortisol metabolism and 11bHSD activity were unitless.
+Values of cortisol production and cortisone production were expressed in nanograms per millilitre (ng/ml), whereas values of cortisol metabolism and 11bHSD activity were unitless.
 
-Concentrations of the non-persistent \acr{edcs} were classified as quantifiable, below the \acr{lod}, possible interference or out of range, and not analysed.
-Concentrations below the \acr{lod} were singly imputed using a quantile regression approach for the imputation of left-censored missing data, as implemented in the `impute.QRILC` function from the `imputeLCMD` R package [@lazar2015imputelcmd].
+Concentrations of the non-persistent {{< acr edcs >}} were classified as quantifiable, below the {{< acr lod >}}, possible interference or out of range, and not analysed.
+Concentrations below the {{< acr lod >}} were singly imputed using a quantile regression approach for the imputation of left-censored missing data, as implemented in the `impute.QRILC` function from the `imputeLCMD` R package [@lazar2015imputelcmd].
 Information about the lower limits of detection can be found in @HaugSakhiCequier:2018.
-Chemicals with more than 70% of observations below the \acr{lod} were not considered in the present study.
+Chemicals with more than 70% of observations below the {{< acr lod >}} were not considered in the present study.
 Remaining missing values were imputed similarly using kNN.
-Values of the chemicals were expressed in $\mu$grams per litre.
+Values of the chemicals were expressed in $\mu$grams per litre ($\mu$g/L).
 
@@
 We natural log-transformed these to improve model fit, assessed with posterior predictive checks.
-Values of the clinical outcome were expressed in \acr{mss}.
+Values of the clinical outcome were expressed in {{< acr mss >}}.
 
@@
 
-We estimated marginal contrasts with the parametric g-formula, a method of standardization.
+We estimated {{< acr mc first_use=true >}}s with the parametric g-formula, a method of standardization.
 The parametric g-formula involves the following steps: 1) fit a outcome model including both covariates and balancing weights; 2) create two new datasets identical to the original one but with the exposure shifted according to a user-specified intervention set by a deterministic function of the observed exposure levels; 3) use the outcome model to compute adjusted predictions in the two counterfactual datasets; 4) compute the difference between the means of the adjusted predictions in the counterfactual datasets.
-The causal parameter of interest $\phi$ was thus specified as the difference in the expected counterfactual outcomes under the shifted exposure levels $\left( \phi^\Delta = \mathbb{E} \left[ Y^{d_1} \right] - \mathbb{E} \left[ Y^{d_2} \right] \right)$.
+The causal parameter of interest $\phi$ was thus specified as the difference in the expected counterfactual outcomes under the shifted exposure levels $\left( \phi^\Delta = \mathbb{E >}} \left[ Y^{d_1 >}} \right] - \mathbb{E >}} \left[ Y^{d_2 >}} \right] \right)$.
 In order for this parameter to be identified, the usual causal identifiability conditions (no unmeasured confounding, positivity, and consistency) are required.
@@
 We fit the outcome model using the `glm` function and a Gaussian family with identity link from base R.
-The exposure variable was modeled using natural cubic splines with 3 degrees of freedom, to more flexibly capture the \acr{adrf}.
+The exposure variable was modeled using natural cubic splines with 3 degrees of freedom, to more flexibly capture the {{< acr adrf >}}.
 When the outcome was a ratio, as was the case for cortisol metabolism and 11bHSD activity, we included the logarithm of its denominator, cortisol and cortisol production, respectively, as a control variable [@BartlettPartnoy:2020].
 
-To estimate the marginal contrasts, we used the `avg_comparisons` function from the `marginaleffects` R package [@Arel-Bundock:2023].
+To estimate the {{< acr mc >}}s, we used the `avg_comparisons` function from the `marginaleffects` R package [@Arel-Bundock:2023].
 The two counterfactual datasets were obtained by setting the exposures levels to 90th percentile ($d_1$) and the 10th percentile ($d_2$), for each cohort separately.
-The marginal contrasts were computed using the estimated balancing weights above.
+The {{< acr mc >}}s were computed using the estimated balancing weights above.
 Robust standard errors were computed with the `sandwich` R package, using cohort as variable indicating clustering of observations [@Zeileis:2004; @ZeileisKollGraham:2020].
-For each outcome, we report the results as differences between marginal contrasts.
+For each outcome, we report the results as differences between {{< acr mc >}}s.
 
-We further estimated the \acr{adrf} using the `avg_predictions` function from the `marginaleffects` R package, examining 50 exposure values from the 10th to the 90th percentiles of the exposure.
-As done for the marginal contrasts, we included the estimated balancing weights and used cohort as a clustering variable when computing robust standard errors.
+We further estimated the {{< acr adrf >}} using the `avg_predictions` function from the `marginaleffects` R package, examining 50 exposure values from the 10th to the 90th percentiles of the exposure.
+As done for the {{< acr mc >}}s, we included the estimated balancing weights and used cohort as a clustering variable when computing robust standard errors.
 
@@
 To do so, balancing weights were estimated separately for each level of the sex variable, and an interaction term between the exposure and sex was included in the outcome model.
-Similarly, the marginal contrasts were aggregated separately for each level of sex.
-We further tested for significance of the difference between the marginal contrasts of males and females.
+Similarly, the {{< acr mc >}}s were aggregated separately for each level of sex.
+We further tested for significance of the difference between the {{< acr mc >}}s of males and females.
 
@@
 
-Of the 1,301 children of the \acr{helix} subcohort, 1,297 had measurements of the non-persistent \acr{edcs}.
-Measurements of the glucocorticosteroids were available for 1,004 children, of which 980 were matched to the \acr{helix} subcohort.
-Measurements of both non-persistent \acr{edcs} and glucocorticosteroids were available for 976 children of the subcohort.
+Of the 1,301 children of the {{< acr helix >}} subcohort, 1,297 had measurements of the non-persistent {{< acr edcs >}}.
+Measurements of the glucocorticosteroids were available for 1,004 children, of which 980 were matched to the {{< acr helix >}} subcohort.
+Measurements of both non-persistent {{< acr edcs >}} and glucocorticosteroids were available for 976 children of the subcohort.
 A flowchart describing the sample size for each research question is presented in @suppfig-flow-pop.
 The sample consisted of 55% girls.
-The median \acr{hitrtse} was 300 ms (\acr{iqr}, 231-368), with lower median values for EDEN, MOBA, and INMA, corresponding to the cohorts with older children.
+The median {{< acr hitrtse >}} was 300 ms ({{< acr iqr >}}, 231-368), with lower median values for EDEN, MOBA, and INMA, corresponding to the cohorts with older children.
 At the time of visit, the median age of the children was 8.06 years.
@@
 
-Levels of unprocessed non-persistent \acr{edcs} and glucocorticosteroids are presented in @tbl-edc-desc, @tbl-met-new-desc, and @supptbl-met-desc.
+Levels of unprocessed non-persistent {{< acr edcs >}}, after imputation of values below the {{< acr lod >}}, and glucocorticosteroids, are presented in @tbl-edc-desc, @tbl-met-new-desc, and @supptbl-met-desc.
 
@@
 
-@Fig-marginal-1 presents the forest plot for the marginal contrasts of the non-persistent \acr{edcs} on \acr{hitrtse}.
-For most \acr{edcs}, a cohort-specific increase in the levels of the exposures from the 10th to the 90th percentiles was associated with a positive marginal contrast, indicating an increase in the values of \acr{hitrtse}.
-Most of the \acr{ci}s included the null effect, though.
+@Fig-marginal-1 presents the forest plot for the {{< acr mc >}}s of the non-persistent {{< acr edcs >}} on {{< acr hitrtse >}}.
+For most {{< acr edcs >}}, a cohort-specific increase in the levels of the exposures from the 10th to the 90th percentiles was associated with a positive {{< acr mc >}}, indicating an increase in the values of {{< acr hitrtse >}}.
+Most of the {{< acr ci first_use=true >}}s included the null effect, though.
 Statistically significant effects were observed for the phenol MEPA (`r format_res(meffects1, "mepa", "hitrtse")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "hitrtse")`), oh-MiNP (`r format_res(meffects1, "ohminp", "hitrtse")`), and MEHP (`r format_res(meffects1, "mehp", "hitrtse")`).
-The \acr{opp} DETP was strongly negatively associated with \acr{hitrtse} (`r format_res(meffects1, "detp", "hitrtse")`).
+The {{< acr opp >}} DETP was strongly negatively associated with {{< acr hitrtse >}} (`r format_res(meffects1, "detp", "hitrtse")`).
 
-@Fig-marginal-2 presents the forest plot for the marginal contrasts of the non-persistent \acr{edcs} on cortisone production, cortisol production, and corticosterone production.
-For most \acr{edcs}, a cohort-specific increase in the levels of the exposures from the 10th to the 90th percentiles was associated with a positive marginal contrast, indicating an increase in the total production of these metabolites.
-Exceptions were BUPA, which was associated with negative marginal contrasts for all three outcomes, and MiBP, which was associated with a negative marginal contrast for total cortisone production only.
+@Fig-marginal-2 presents the forest plot for the {{< acr mc >}}s of the non-persistent {{< acr edcs >}} on cortisone production, cortisol production, and corticosterone production.
+For most {{< acr edcs >}}, a cohort-specific increase in the levels of the exposures from the 10th to the 90th percentiles was associated with a positive {{< acr mc >}}, indicating an increase in the total production of these metabolites.
+Exceptions were BUPA, which was associated with negative {{< acr mc >}}s for all three outcomes, and MiBP, which was associated with a negative {{< acr mc >}} for total cortisone production only.
 The majority of the effects for the phenols and phthalate metabolites included the null.
-The phenol BPA showed the largest marginal contrasts across all three outcomes (cortisone production, `r format_res(meffects2, "bpa", "cortisone production")`; cortisol production, `r format_res(meffects1, "bpa", "cortisol production")`; corticosterone production, `r format_res(meffects1, "bpa", "corticosterone production")`).
+The phenol BPA showed the largest {{< acr mc >}}s across all three outcomes (cortisone production, `r format_res(meffects2, "bpa", "cortisone production")`; cortisol production, `r format_res(meffects1, "bpa", "cortisol production")`; corticosterone production, `r format_res(meffects1, "bpa", "corticosterone production")`).
 
-@Fig-marginal-3 presents the forest plot for the marginal contrasts of the glucocorticosteroids on \acr{hitrtse}.
-All marginal contrasts included the null, with no clear indication of directionality of the effect.
+@Fig-marginal-3 presents the forest plot for the {{< acr mc >}}s of the glucocorticosteroids on {{< acr hitrtse >}}.
+All {{< acr mc >}}s included the null, with no clear indication of directionality of the effect.
 
@@
 
-@Tbl-hypothesis-1 presents the results of a hypothesis test for the difference between estimates of the marginal contrasts for females and males, for each exposure.
+@Tbl-hypothesis-1 presents the results of a hypothesis test for the difference between estimates of the {{< acr mc >}}s for females and males, for each exposure.
 Significant differences were present for the phenol OXBE (`r format_res(heffects1, "oxbe", "hitrtse")`) and the phthalate metabolites MEP (`r format_res(heffects1, "mep", "hitrtse")`) and MbZP (`r format_res(heffects1, "mbzp", "hitrtse")`).
-The forest plot of the individual marginal contrasts is presented in @suppfig-marginal-1sa.
+The forest plot of the individual {{< acr mc >}}s is presented in @suppfig-marginal-1sa.
 
-@Tbl-hypothesis-2 presents the results of a hypothesis test for the difference between estimates of the marginal contrasts for females and males, for each exposure and outcome.
-Significant differences were present across all three classes of \acr{edcs} and for all outcomes.
-The largest differences were attributable to the \acr{opps} DMTP (cortisol production, `r format_res(heffects2, "dmtp", "cortisol production")`) and DETP (corticosterone production, (`r format_res(heffects2, "detp", "corticosterone production")`); cortisone production, (`r format_res(heffects2, "detp", "cortisone production")`)).
-The forest plots of the individual marginal contrasts are presented in @suppfig-marginal-2sa.
+@Tbl-hypothesis-2 presents the results of a hypothesis test for the difference between estimates of the {{< acr mc >}}s for females and males, for each exposure and outcome.
+Significant differences were present across all three classes of {{< acr edcs >}} and for all outcomes.
+The largest differences were attributable to the {{< acr opps >}} DMTP (cortisol production, `r format_res(heffects2, "dmtp", "cortisol production")`) and DETP (corticosterone production, (`r format_res(heffects2, "detp", "corticosterone production")`); cortisone production, (`r format_res(heffects2, "detp", "cortisone production")`)).
+The forest plots of the individual {{< acr mc >}}s are presented in @suppfig-marginal-2sa.
 
-@Tbl-hypothesis-3 presents the results of a hypothesis test for the difference between estimates of the marginal contrasts for females and males, for each exposure.
+@Tbl-hypothesis-3 presents the results of a hypothesis test for the difference between estimates of the {{< acr mc >}}s for females and males, for each exposure.
 Significant differences were present for cortisone production (`r format_res(heffects3, "cortisone prod.", "hitrtse")`) and corticosterone production (`r format_res(heffects3, "corticost.
 prod.", "hitrtse")`).
-Furthermore, for all exposures, the marginal contrasts had opposite sign (positive for males and negative for females).
-The forest plot of the individual marginal contrasts is presented in @suppfig-marginal-3sa.
+Furthermore, for all exposures, the {{< acr mc >}}s had opposite sign (positive for males and negative for females).
+The forest plot of the individual {{< acr mc >}}s is presented in @suppfig-marginal-3sa.
 
@@
 
-The impact of exposure to \acr{edcs} on human health has attracted considerable research interest.
+The impact of exposure to {{< acr edcs >}} on human health has attracted considerable research interest.
 While research in this area has mainly investigated the effects of prenatal exposure on child neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022], little is still known about childhood exposure.
-In this study, consisting of 1,297 children from 6 European birth cohorts, we observed that short-term childhood exposure to non-persistent \acr{edcs} had negative effects on \acr{hitrtse} and total production of cortisol, cortisone, and corticosterone, although the majority of the \acr{ci}s included the null effect.
-Increased production of these glucocorticosteroids did not seem to affect \acr{hitrtse}.
-Some of these effects differed for females and males, including significant differences for the effects of increased production of cortisone and corticosterone on \acr{hitrtse}.
-Specifically, an increased production of these glucocorticosteroids was associated with lower values of \acr{hitrtse} for females, and higher values for males.
-Taken together, these results suggest that these non-persistent \acr{edcs} might be responsible for perturbations of the \acr{hpa} axis' homeostasis, and that higher levels of these glucocorticosteroids might interfere with different functions of attention in a sex-specific manner.
+In this study, consisting of 1,297 children from 6 European birth cohorts, we observed that short-term childhood exposure to non-persistent {{< acr edcs >}} had negative effects on {{< acr hitrtse >}} and total production of cortisol, cortisone, and corticosterone, although the majority of the {{< acr ci >}}s included the null effect.
+Increased production of these glucocorticosteroids did not seem to affect {{< acr hitrtse >}}.
+Some of these effects differed for females and males, including significant differences for the effects of increased production of cortisone and corticosterone on {{< acr hitrtse >}}.
+Specifically, an increased production of these glucocorticosteroids was associated with lower values of {{< acr hitrtse >}} for females, and higher values for males.
+Taken together, these results suggest that these non-persistent {{< acr edcs >}} might be responsible for perturbations of the {{< acr hpa >}} axis' homeostasis, and that higher levels of these glucocorticosteroids might interfere with different functions of attention in a sex-specific manner.
 
-We are not aware of prior studies specifically investigating the effects of exposure to \acr{edcs} in relation to \acr{hitrtse}.
-The literature on \acr{edcs} and neurodevelopment in children has mostly focused on \acr{opps} [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @CartierWarembourgLeManer-Idrissi:2016; YuDuChiou:2016], phthalate metabolites [@HuangChenSu:2015; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023], and \acr{bpa} [@TewarAuingerBraun:2016; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019], in relation to \acr{adhd} [@BouchardBellingerWright:2010; @TewarAuingerBraun:2016; @YuDuChiou:2016; @LiZhangKuang:2018], and intelligence scales [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @HuangChenSu:2015; @CartierWarembourgLeManer-Idrissi:2016; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023].
-Few studies have looked into different classes of \acr{edcs} [@ShoaffCoullWeuve:2020 in relation with the Conners Attention Deficit Scale and the Behavior Assessment System for Children; @OhKimKannan:2023 in relation with \acr{adhd} symptoms].
-Overall, and consistent with our results, these studies seem to provide further evidence of the adverse effects of several \acr{edcs} on neurodevelopment in children.
+We are not aware of prior studies specifically investigating the effects of exposure to {{< acr edcs >}} in relation to {{< acr hitrtse >}}.
+The literature on {{< acr edcs >}} and neurodevelopment in children has mostly focused on {{< acr opps >}} [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @CartierWarembourgLeManer-Idrissi:2016; YuDuChiou:2016], phthalate metabolites [@HuangChenSu:2015; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023], and {{< acr bpa >}} [@TewarAuingerBraun:2016; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019], in relation to {{< acr adhd >}} [@BouchardBellingerWright:2010; @TewarAuingerBraun:2016; @YuDuChiou:2016; @LiZhangKuang:2018], and intelligence scales [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @HuangChenSu:2015; @CartierWarembourgLeManer-Idrissi:2016; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023].
+Few studies have looked into different classes of {{< acr edcs >}} [@ShoaffCoullWeuve:2020 in relation with the Conners Attention Deficit Scale and the Behavior Assessment System for Children; @OhKimKannan:2023 in relation with {{< acr adhd >}} symptoms].
+Overall, and consistent with our results, these studies seem to provide further evidence of the adverse effects of several {{< acr edcs >}} on neurodevelopment in children.
 While not all these studies have investigated effect modification by sex, it seems that these adverse effects are stronger in males.
@@
 
-Our results are consistent with prior epidemiological research that associated exposure to certain \acr{edcs} with higher levels of cortisol [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023].
+Our results are consistent with prior epidemiological research that associated exposure to certain {{< acr edcs >}} with higher levels of cortisol [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023].
 There are some differences, though.
@@
 Contrary to these studies, we did find effect modification by sex.
-We are not aware of other epidemiological studies investigating phthalates metabolites, phenols, and \acr{opps}, in relation to urinary glucocorticosteroids in childhood.
+We are not aware of other epidemiological studies investigating phthalates metabolites, phenols, and {{< acr opps >}}, in relation to urinary glucocorticosteroids in childhood.
 
-We are also not aware of prior epidemiological studies specifically investigating the effects of elevated levels of glucocorticosteroids in relation to \acr{hitrtse}, although there is evidence that under- or over-production of glucocorticosteroids interfere with the normal development of the brain [@LupienMcEwenGunnar:2009].
+We are also not aware of prior epidemiological studies specifically investigating the effects of elevated levels of glucocorticosteroids in relation to {{< acr hitrtse >}}, although there is evidence that under- or over-production of glucocorticosteroids interfere with the normal development of the brain [@LupienMcEwenGunnar:2009].
 While we did find sex-specific evidence of an effect, their clinical relevance is questionable.
@@
 Limitations include the cross-sectional design of the present study.
-Importantly, the non-persistent \acr{edcs} were measured in a pool of night and morning urine samples before the clinical visit, to represent exposure over the previous day, whereas the glucocorticosteroids were measured in the night urine sample.
+Importantly, the non-persistent {{< acr edcs >}} were measured in a pool of night and morning urine samples before the clinical visit, to represent exposure over the previous day, whereas the glucocorticosteroids were measured in the night urine sample.
 Although we included a wide range of confounders there is the possibility, as with other observational studies, of residual confounding, which might lead to a bias away from the null.
@@
 
-Strengths of the present study include the use of pooled urine samples for chemical assessment, since it is known that these specific \acr{edcs} have very short half-lives [@CasasBasaganaSakhi:2018].
+Strengths of the present study include the use of pooled urine samples for chemical assessment, since it is known that these specific {{< acr edcs >}} have very short half-lives [@CasasBasaganaSakhi:2018].
 We decided to model both the *treatment* mechanisms, for the estimation of balancing weights, and the outcomes, with traditional covariates adjustment, to try to obtain *doubly robust* effect estimates.
-Finally, we decided not to interpret our results by focusing on the estimated coefficients of the regression models, but by making use of the g-computation procedure and estimate marginal contrasts.
+Finally, we decided not to interpret our results by focusing on the estimated coefficients of the regression models, but by making use of the g-computation procedure and estimate {{< acr mc >}}s.
 
-In conclusion, in a study of 1,297 children from 6 European birth cohorts, we observed that (i) exposure to non-persistent \acr{edcs} in childhood might have short-term effects on \acr{hitrtse} in childhood, (ii) exposure to non-persistent \acr{edcs} in childhood might disrupt the \acr{hpa} axis in childhood, and (iii) disruption of the \acr{hpa} axis in childhood might have short-term, sex-specific effects on \acr{hitrtse}.
-Future studies should investigate how glucocorticosteroids might mediate the adverse effects of exposure to \acr{edcs} on childhood neurodevelopment in larger populations.
+In conclusion, in a study of 1,297 children from 6 European birth cohorts, we observed that (i) exposure to non-persistent {{< acr edcs >}} in childhood might have short-term effects on {{< acr hitrtse >}} in childhood, (ii) exposure to non-persistent {{< acr edcs >}} in childhood might disrupt the {{< acr hpa >}} axis in childhood, and (iii) disruption of the {{< acr hpa >}} axis in childhood might have short-term, sex-specific effects on {{< acr hitrtse >}}.
+Future studies should investigate how glucocorticosteroids might mediate the adverse effects of exposure to {{< acr edcs >}} on childhood neurodevelopment in larger populations.
 

### 2024-02-12 output/paper/_figures.qmd
Trying to fix Table 1, now with by cohort only in SM. Modifying names EDCs in forest plots.

--- 2024-01-02 output/paper/_figures.qmd
@@
 forest_plot <- function(obj, r = 4) {
+  edcs <- myphd::edcs_information() |>
+    dplyr::filter(!chem_id %in% c("dedtp", "dmdtp")) |>
+    dplyr::select(chem_id, short_name, class) |>
+    dplyr::bind_rows(
+      tibble::tibble(
+        chem_id = c("Exposure"),
+        short_name = c("Exposure"),
+        class = c(NULL)
+      )
+    )
+
   # From table of numerical results to ggplot object
@@
     tidylog::select(-estimate) |>
+    dplyr::left_join(
+      edcs[, c("chem_id", "short_name")],
+      by = c("variable" = "chem_id")
+    ) |>
+    dplyr::select(-variable) |>
+    dplyr::rename(variable = "short_name") |>
+    dplyr::relocate(variable) |>
     ggplot2::ggplot(
@@
 ```
-**Marginal contrasts (exposures: \acr{edcs}; outcome: \acr{hitrtse}) (HELIX Study; 2013-2016).**
-Circles indicate effect estimates. Solid lines indicate the $95\%$ \acr{ci}.
+**Marginal contrasts for the effect of a decrease from the 90th to the 10th percentile of the {{< acr edcs >}} on {{< acr hitrtse >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).** 
+Circles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}.
 :::
@@
 ```
-**Marginal contrasts (exposures: \acr{edcs}; outcomes: glucocorticosteroids) (HELIX Study; 2013-2016).**
-Circles, triangles, and squares indicate effect estimates. Solid lines indicate the $95\%$ \acr{ci}.
+**Marginal contrasts for the effect of a decrease from the 90th to the 10th percentile of the {{< acr edcs >}} on the glucocorticosteroids expressed in ng/ml (HELIX subcohort; 2013-2016).** 
+Circles, triangles, and squares indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}.
 :::
@@
 ```
-**Marginal contrasts (exposures: glucocorticosteroids; outcome: \acr{hitrtse}) (HELIX Study; 2013-2016).**
-Circles indicate effect estimates. Solid lines indicate the $95\%$ \acr{ci}.
+**Marginal contrasts for the effect of a decrease from the 90th to the 10th percentile of the glucocorticosteroids on {{< acr hitrtse >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).** 
+Circles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}.
 :::

### 2024-02-12 output/paper/_si.qmd
Trying to fix Table 1, now with by cohort only in SM. Modifying names EDCs in forest plots.

@@
 ### Study populations
+
+::: {#supptbl-pop-desc}
+```{r pop-desc-full}
+targets::tar_load("table1_full", store = store)
+table1_full
+```
+**Participant characteristics, by cohort and overall (HELIX subcohort; 2013-2016).**
+:::
 
@@
   gt::tab_footnote(
-    footnote = "Abbreviations: lower limit of quantification (LLOQ), expressed in nanograms per millilitre."
+    footnote = "Abbreviations: lower limit of quantification (LLOQ)."
   ) |>
@@
 ```
-**Lower limits of quantification for the glucocorticosteroids (HELIX subcohort; 2013-2016).**
+**Lower limits of quantification expressed in ng/ml for the glucocorticosteroids (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Marginal contrasts for effect modification by sex of a decrease from the 90th to the 10th percentile of the {{< acr edcs >}} on {{< acr hitrtse >}} (HELIX subcohort; 2013-2016).** 
+**Marginal contrasts for effect modification by sex of a decrease from the 90th to the 10th percentile of the {{< acr edcs >}} on {{< acr hitrtse >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
 Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}.
@@
 ```
-**Marginal contrasts for effect modification by sex of a decrease from the 90th to the 10th percentile of the {{< acr edcs >}} on the glucocorticosteroids (HELIX subcohort; 2013-2016).** 
+**Marginal contrasts for effect modification by sex of a decrease from the 90th to the 10th percentile of the {{< acr edcs >}} on the glucocorticosteroids expressed in ng/ml (HELIX subcohort; 2013-2016).**
 Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}.
@@
 ```
-**Marginal contrasts for effect modification by sex of a decrease from the 90th to the 10th percentile of the glucocorticosteroids on {{< acr hitrtse >}} (HELIX subcohort; 2013-2016).** 
+**Marginal contrasts for effect modification by sex of a decrease from the 90th to the 10th percentile of the glucocorticosteroids on {{< acr hitrtse >}} expressed in {{< acr mss >}}
+ (HELIX subcohort; 2013-2016).** 
 Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}.

### 2024-02-12 output/paper/_tables.qmd
Trying to fix Table 1, now with by cohort only in SM. Modifying names EDCs in forest plots.

@@
 ```
-**Participant characteristics, by cohort and overall (HELIX subcohort; 2013-2016).**
+**Participant characteristics (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Participants endocrine disruptors concentrations (HELIX subcohort; 2013-2016).**
+**Participants endocrine disruptors concentrations expressed in $\mu$grams/L (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Participants derived glucocorticosteroids concentrations, by cohort and overall (HELIX subcohort; 2013-2016).**
+**Participants derived glucocorticosteroids concentrations expressed in ng/ml, by cohort and overall (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Pairwise differences between sex-specific marginal contrasts for the effect of a decrease from the 90th to the 10th percentile of {{< acr edcs >}} on {{< acr hitrtse >}} and the glucocorticosteroids (HELIX subcohort; 2013-2016).**
+**Pairwise differences between sex-specific marginal contrasts for the effect of a decrease from the 90th to the 10th percentile of {{< acr edcs >}} on {{< acr hitrtse >}}, expressed in {{< acr mss >}}, and the glucocorticosteroids, expressed in ng/ml (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Pairwise differences between sex-specific marginal contrasts for the effect of a decrease from the 90th to the 10th percentile of the glucocorticosteroids on {{< acr hitrtse >}} (HELIX subcohort; 2013-2016).**
+**Pairwise differences between sex-specific marginal contrasts for the effect of a decrease from the 90th to the 10th percentile of the glucocorticosteroids on {{< acr hitrtse >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
 :::

### 2024-02-12 output/paper/manuscript.qmd
Trying to fix Table 1, now with by cohort only in SM. Modifying names EDCs in forest plots.

@@
       theme: neutral
-  # pdf:
-  #   pdf-engine: lualatex
-  #   number-sections: true
-  #   toc: false
-  #   keep-tex: false
-  #   include-in-header:
-  #     text: |
-  #       \usepackage{lineno}
-  #       \usepackage{rotating}
-  #       \usepackage{float}
-  #       \usepackage{typearea}
-  #   include-before-body:
-  #     text: |
-  #       \linenumbers
+  pdf:
+    documentclass: article
+    geometry:
+      - top=30mm
+      - left=20mm
+      - heightrounded
+    pdf-engine: lualatex
+    number-sections: true
+    toc: false
+    keep-tex: false
+    include-in-header:
+      text: |
+        \usepackage{lineno}
+        \usepackage{rotating}
+        \usepackage{float}
+        \usepackage{typearea}
+    include-before-body:
+      text: |
+        \linenumbers
 
@@
   - acronyms
-  # - authors-block
-  # - ../../../../all_phd/_extensions/wordcount/citeproc.lua
   - ../../../../all_phd/_extensions/wordcount/wordcount.lua
@@
   glue::glue(
-    "\{{< acr mc >}} and $95\\%$ \{{< acr ci >}}: {val}"
+    "\\acr{{mc}} and $95\\%$ \\acr{{ci}}: {val}"
   )
@@
 We used the parametric g-formula and {{< acr mcs >}} to estimate effects between {{< acr edcs >}}, glucocorticosteroids, and {{< acr hitrtse >}}, a measure of efficiency from the {{< acr ant >}}, and tested for possible effect modification by sex.
-We observed positive {{< acr mc >}}s for exposure increases from the 10th to the 90th percentile for methyl-paraben (`r format_res(meffects1, "mepa", "hitrtse")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "hitrtse")`), oh-MiNP (`r format_res(meffects1, "ohminp", "hitrtse")`), and MEHP (`r format_res(meffects1, "mehp", "hitrtse")`), on {{< acr hitrtse >}}.
-Several {{< acr edcs >}} were also associated with positive {{< acr mc >}}s of cortisone, cortisol, and corticosterone production.
+We observed positive {{< acr mcs >}} for exposure increases from the 10th to the 90th percentile for methyl-paraben (`r format_res(meffects1, "mepa", "hitrtse")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "hitrtse")`), oh-MiNP (`r format_res(meffects1, "ohminp", "hitrtse")`), and MEHP (`r format_res(meffects1, "mehp", "hitrtse")`), on {{< acr hitrtse >}}.
+Several {{< acr edcs >}} were also associated with positive {{< acr mcs >}} of cortisone, cortisol, and corticosterone production.
 Increased levels of the glucocorticosteroids had no effect on {{< acr hitrtse >}}, although we found possible effect modification by sex for cortisone and corticosterone production.
@@
 
-{{< acr opps >}} are one group of {{< acr edcs >}} that may have a deleterious effect on neurodevelopment, although the few studies assessing exposure during childhood and through the use of biomarkers suffered from a series of limitations, including a small sample size and few {{< acr opp >}} metabolites measured [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
+One group of {{< acr edcs >}} that may have a deleterious effect on neurodevelopment is the {{< acr opps >}}, although the few studies assessing exposure during childhood and through the use of biomarkers suffered from a series of limitations, including a small sample size and few {{< acr opp >}} metabolites measured [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
 Exposure to phthalates and their metabolites during childhood and early adolescence has been associated to several adverse neurodevelopmental outcomes, but these studies were limited to few phthalate metabolites and small study populations [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
@@
 Taken together, these results suggest that the negative influence of exposure to certain {{< acr edcs >}} on neurodevelopmental outcomes, might be mediated, at least partially, by disruption of the {{< acr hpa >}} axis' homeostasis.
-In the present study, we thus estimated associations between 1) non-persistent {{< acr edcs >}} and attention, 2) non-persistent {{< acr edcs >}} and glucocorticosteroids, and 3) glucocorticosteroids and attention, using the parametric g-formula and {{< mc >}}s, in children of a large cohort in Europe.
+In the present study, we thus estimated associations between 1) non-persistent {{< acr edcs >}} and attention, 2) non-persistent {{< acr edcs >}} and glucocorticosteroids, and 3) glucocorticosteroids and attention, using the parametric g-formula and {{< acr mcs first_use=true >}}, in children of a large cohort in Europe.
 
@@
 
-We estimated {{< acr mc first_use=true >}}s with the parametric g-formula, a method of standardization.
+We estimated {{< acr mcs >}} with the parametric g-formula, a method of standardization.
 The parametric g-formula involves the following steps: 1) fit a outcome model including both covariates and balancing weights; 2) create two new datasets identical to the original one but with the exposure shifted according to a user-specified intervention set by a deterministic function of the observed exposure levels; 3) use the outcome model to compute adjusted predictions in the two counterfactual datasets; 4) compute the difference between the means of the adjusted predictions in the counterfactual datasets.
-The causal parameter of interest $\phi$ was thus specified as the difference in the expected counterfactual outcomes under the shifted exposure levels $\left( \phi^\Delta = \mathbb{E >}} \left[ Y^{d_1 >}} \right] - \mathbb{E >}} \left[ Y^{d_2 >}} \right] \right)$.
+The causal parameter of interest $\phi$ was thus specified as the difference in the expected counterfactual outcomes under the shifted exposure levels $\left( \phi^\Delta = \mathbb{E} \left[ Y^{d_1} \right] - \mathbb{E} \left[ Y^{d_2} \right] \right)$.
 In order for this parameter to be identified, the usual causal identifiability conditions (no unmeasured confounding, positivity, and consistency) are required.
@@
 
-To estimate the {{< acr mc >}}s, we used the `avg_comparisons` function from the `marginaleffects` R package [@Arel-Bundock:2023].
+To estimate the {{< acr mcs >}}, we used the `avg_comparisons` function from the `marginaleffects` R package [@Arel-Bundock:2023].
 The two counterfactual datasets were obtained by setting the exposures levels to 90th percentile ($d_1$) and the 10th percentile ($d_2$), for each cohort separately.
-The {{< acr mc >}}s were computed using the estimated balancing weights above.
+The {{< acr mcs >}} were computed using the estimated balancing weights above.
 Robust standard errors were computed with the `sandwich` R package, using cohort as variable indicating clustering of observations [@Zeileis:2004; @ZeileisKollGraham:2020].
-For each outcome, we report the results as differences between {{< acr mc >}}s.
+For each outcome, we report the results as differences between {{< acr mcs >}}.
 
 We further estimated the {{< acr adrf >}} using the `avg_predictions` function from the `marginaleffects` R package, examining 50 exposure values from the 10th to the 90th percentiles of the exposure.
-As done for the {{< acr mc >}}s, we included the estimated balancing weights and used cohort as a clustering variable when computing robust standard errors.
+As done for the {{< acr mcs >}}, we included the estimated balancing weights and used cohort as a clustering variable when computing robust standard errors.
 
@@
 To do so, balancing weights were estimated separately for each level of the sex variable, and an interaction term between the exposure and sex was included in the outcome model.
-Similarly, the {{< acr mc >}}s were aggregated separately for each level of sex.
-We further tested for significance of the difference between the {{< acr mc >}}s of males and females.
+Similarly, the {{< acr mcs >}} were aggregated separately for each level of sex.
+We further tested for significance of the difference between the {{< acr mcs >}} of males and females.
 
@@
 
+@tbl-pop-desc and @supptbl-pop-desc provide descriptive statistics for the outcome and covariates for the {{< acr helix >}} subcohort and for each cohort, respectively.
 Of the 1,301 children of the {{< acr helix >}} subcohort, 1,297 had measurements of the non-persistent {{< acr edcs >}}.
@@
 The children were mostly Caucasian (90%), and the largest minority were of Pakistani origin (6.2%).
-@tbl-pop-desc provides descriptive statistics for the outcome and covariates, by cohort and overall.
 
@@
 
-@Fig-marginal-1 presents the forest plot for the {{< acr mc >}}s of the non-persistent {{< acr edcs >}} on {{< acr hitrtse >}}.
+@Fig-marginal-1 presents the forest plot for the {{< acr mcs >}} of the non-persistent {{< acr edcs >}} on {{< acr hitrtse >}}.
 For most {{< acr edcs >}}, a cohort-specific increase in the levels of the exposures from the 10th to the 90th percentiles was associated with a positive {{< acr mc >}}, indicating an increase in the values of {{< acr hitrtse >}}.
-Most of the {{< acr ci first_use=true >}}s included the null effect, though.
+Most of the {{< acr cis first_use=true >}} included the null effect, though.
 Statistically significant effects were observed for the phenol MEPA (`r format_res(meffects1, "mepa", "hitrtse")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "hitrtse")`), oh-MiNP (`r format_res(meffects1, "ohminp", "hitrtse")`), and MEHP (`r format_res(meffects1, "mehp", "hitrtse")`).
@@
 
-@Fig-marginal-2 presents the forest plot for the {{< acr mc >}}s of the non-persistent {{< acr edcs >}} on cortisone production, cortisol production, and corticosterone production.
+@Fig-marginal-2 presents the forest plot for the {{< acr mcs >}} of the non-persistent {{< acr edcs >}} on cortisone production, cortisol production, and corticosterone production.
 For most {{< acr edcs >}}, a cohort-specific increase in the levels of the exposures from the 10th to the 90th percentiles was associated with a positive {{< acr mc >}}, indicating an increase in the total production of these metabolites.
-Exceptions were BUPA, which was associated with negative {{< acr mc >}}s for all three outcomes, and MiBP, which was associated with a negative {{< acr mc >}} for total cortisone production only.
+Exceptions were BUPA, which was associated with negative {{< acr mcs >}} for all three outcomes, and MiBP, which was associated with a negative {{< acr mc >}} for total cortisone production only.
 The majority of the effects for the phenols and phthalate metabolites included the null.
-The phenol BPA showed the largest {{< acr mc >}}s across all three outcomes (cortisone production, `r format_res(meffects2, "bpa", "cortisone production")`; cortisol production, `r format_res(meffects1, "bpa", "cortisol production")`; corticosterone production, `r format_res(meffects1, "bpa", "corticosterone production")`).
+The phenol BPA showed the largest {{< acr mcs >}} across all three outcomes (cortisone production, `r format_res(meffects2, "bpa", "cortisone production")`; cortisol production, `r format_res(meffects2, "bpa", "cortisol production")`; corticosterone production, `r format_res(meffects2, "bpa", "corticosterone production")`).
 
-@Fig-marginal-3 presents the forest plot for the {{< acr mc >}}s of the glucocorticosteroids on {{< acr hitrtse >}}.
-All {{< acr mc >}}s included the null, with no clear indication of directionality of the effect.
+@Fig-marginal-3 presents the forest plot for the {{< acr mcs >}} of the glucocorticosteroids on {{< acr hitrtse >}}.
+All {{< acr mcs >}} included the null, with no clear indication of directionality of the effect.
 
@@
 
-@Tbl-hypothesis-1 presents the results of a hypothesis test for the difference between estimates of the {{< acr mc >}}s for females and males, for each exposure.
+@Tbl-hypothesis-1and2 presents the results of a hypothesis test for the difference between estimates of the {{< acr mcs >}} for females and males, for each exposure.
 Significant differences were present for the phenol OXBE (`r format_res(heffects1, "oxbe", "hitrtse")`) and the phthalate metabolites MEP (`r format_res(heffects1, "mep", "hitrtse")`) and MbZP (`r format_res(heffects1, "mbzp", "hitrtse")`).
-The forest plot of the individual {{< acr mc >}}s is presented in @suppfig-marginal-1sa.
+The forest plot of the individual {{< acr mcs >}} is presented in @suppfig-marginal-1sa.
 
-@Tbl-hypothesis-2 presents the results of a hypothesis test for the difference between estimates of the {{< acr mc >}}s for females and males, for each exposure and outcome.
+@Tbl-hypothesis-1and2 presents the results of a hypothesis test for the difference between estimates of the {{< acr mcs >}} for females and males, for each exposure and outcome.
 Significant differences were present across all three classes of {{< acr edcs >}} and for all outcomes.
 The largest differences were attributable to the {{< acr opps >}} DMTP (cortisol production, `r format_res(heffects2, "dmtp", "cortisol production")`) and DETP (corticosterone production, (`r format_res(heffects2, "detp", "corticosterone production")`); cortisone production, (`r format_res(heffects2, "detp", "cortisone production")`)).
-The forest plots of the individual {{< acr mc >}}s are presented in @suppfig-marginal-2sa.
+The forest plots of the individual {{< acr mcs >}} are presented in @suppfig-marginal-2sa.
 
-@Tbl-hypothesis-3 presents the results of a hypothesis test for the difference between estimates of the {{< acr mc >}}s for females and males, for each exposure.
+@Tbl-hypothesis-3 presents the results of a hypothesis test for the difference between estimates of the {{< acr mcs >}} for females and males, for each exposure.
 Significant differences were present for cortisone production (`r format_res(heffects3, "cortisone prod.", "hitrtse")`) and corticosterone production (`r format_res(heffects3, "corticost.
 prod.", "hitrtse")`).
-Furthermore, for all exposures, the {{< acr mc >}}s had opposite sign (positive for males and negative for females).
-The forest plot of the individual {{< acr mc >}}s is presented in @suppfig-marginal-3sa.
+Furthermore, for all exposures, the {{< acr mcs >}} had opposite sign (positive for males and negative for females).
+The forest plot of the individual {{< acr mcs >}} is presented in @suppfig-marginal-3sa.
 
@@
 While research in this area has mainly investigated the effects of prenatal exposure on child neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022], little is still known about childhood exposure.
-In this study, consisting of 1,297 children from 6 European birth cohorts, we observed that short-term childhood exposure to non-persistent {{< acr edcs >}} had negative effects on {{< acr hitrtse >}} and total production of cortisol, cortisone, and corticosterone, although the majority of the {{< acr ci >}}s included the null effect.
+In this study, consisting of 1,297 children from 6 European birth cohorts, we observed that short-term childhood exposure to non-persistent {{< acr edcs >}} had negative effects on {{< acr hitrtse >}} and total production of cortisol, cortisone, and corticosterone, although the majority of the {{< acr cis >}} included the null effect.
 Increased production of these glucocorticosteroids did not seem to affect {{< acr hitrtse >}}.
@@
 We are not aware of prior studies specifically investigating the effects of exposure to {{< acr edcs >}} in relation to {{< acr hitrtse >}}.
-The literature on {{< acr edcs >}} and neurodevelopment in children has mostly focused on {{< acr opps >}} [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @CartierWarembourgLeManer-Idrissi:2016; YuDuChiou:2016], phthalate metabolites [@HuangChenSu:2015; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023], and {{< acr bpa >}} [@TewarAuingerBraun:2016; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019], in relation to {{< acr adhd >}} [@BouchardBellingerWright:2010; @TewarAuingerBraun:2016; @YuDuChiou:2016; @LiZhangKuang:2018], and intelligence scales [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @HuangChenSu:2015; @CartierWarembourgLeManer-Idrissi:2016; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023].
+The literature on {{< acr edcs >}} and neurodevelopment in children has mostly focused on {{< acr opps >}} [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @CartierWarembourgLeManer-Idrissi:2016; @YuDuChiou:2016], phthalate metabolites [@HuangChenSu:2015; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023], and {{< acr bpa >}} [@TewarAuingerBraun:2016; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019], in relation to {{< acr adhd >}} [@BouchardBellingerWright:2010; @TewarAuingerBraun:2016; @YuDuChiou:2016; @LiZhangKuang:2018], and intelligence scales [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @HuangChenSu:2015; @CartierWarembourgLeManer-Idrissi:2016; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023].
 Few studies have looked into different classes of {{< acr edcs >}} [@ShoaffCoullWeuve:2020 in relation with the Conners Attention Deficit Scale and the Behavior Assessment System for Children; @OhKimKannan:2023 in relation with {{< acr adhd >}} symptoms].
@@
 We decided to model both the *treatment* mechanisms, for the estimation of balancing weights, and the outcomes, with traditional covariates adjustment, to try to obtain *doubly robust* effect estimates.
-Finally, we decided not to interpret our results by focusing on the estimated coefficients of the regression models, but by making use of the g-computation procedure and estimate {{< acr mc >}}s.
+Finally, we decided not to interpret our results by focusing on the estimated coefficients of the regression models, but by making use of the g-computation procedure and estimate {{< acr mcs >}}.
 

### 2024-02-12 R/proc_res.R

@@ -96,7 +96,7 @@ tidy_codebooks <- function(rq) {
 #'
 #' @return
 #' @export
-tbl_desc_pop <- function(num_digits_est, num_digits_sig) {
+tbl_desc_pop <- function(num_digits_est, num_digits_sig, by) {
   path_store <- Sys.getenv("path_store")
   
   # Load data request (all variables)
@@ -179,27 +179,49 @@ tbl_desc_pop <- function(num_digits_est, num_digits_sig) {
   )
   
   # Generate "Table 1"
-  desc_covars <- df_meta |>
-    gtsummary::tbl_summary(
-      by = "cohort",
-      type = list(
-        hs_fastfood ~ "continuous",
-        hs_org_food ~ "continuous"
-      ),
-      statistic = list(
-        gtsummary::all_continuous() ~ c(
-          "{median} ({p25}, {p75})"
+  if (by == FALSE) {
+    desc_covars <- df_meta |>
+      gtsummary::tbl_summary(
+        type = list(
+          hs_fastfood ~ "continuous",
+          hs_org_food ~ "continuous"
         ),
-        gtsummary::all_categorical() ~ "{n} ({p}%)"
-      ),
-      digits = dplyr::everything() ~ 1,
-      missing = "ifany"
-    ) |>
-    gtsummary::add_overall() |>
-    gtsummary::as_gt() |>
-    gt::opt_footnote_marks(
-      marks = "letters"
-    )
+        statistic = list(
+          gtsummary::all_continuous() ~ c(
+            "{median} ({p25}, {p75})"
+          ),
+          gtsummary::all_categorical() ~ "{n} ({p}%)"
+        ),
+        digits = dplyr::everything() ~ 1,
+        missing = "ifany"
+      ) |>
+      gtsummary::as_gt() |>
+      gt::opt_footnote_marks(
+        marks = "letters"
+      )
+  } else {
+    desc_covars <- df_meta |>
+      gtsummary::tbl_summary(
+        by = "cohort",
+        type = list(
+          hs_fastfood ~ "continuous",
+          hs_org_food ~ "continuous"
+        ),
+        statistic = list(
+          gtsummary::all_continuous() ~ c(
+            "{median} ({p25}, {p75})"
+          ),
+          gtsummary::all_categorical() ~ "{n} ({p}%)"
+        ),
+        digits = dplyr::everything() ~ 1,
+        missing = "ifany"
+      ) |>
[…]

### 2024-02-12 _targets_res.R

@@ -16,9 +16,19 @@ list(
     name = "table1",
     command = tbl_desc_pop(
       num_digits_est = num_digits_est,
-      num_digits_sig = num_digits_sig
+      num_digits_sig = num_digits_sig,
+      by = FALSE
     )
   ), # End table1 target
+  ##############################################################################
+  targets::tar_target(
+    name = "table1_full",
+    command = tbl_desc_pop(
+      num_digits_est = num_digits_est,
+      num_digits_sig = num_digits_sig,
+      by = TRUE
+    )
+  ), # End table1_full target
   ##############################################################################
   targets::tar_target(
     name = "desc_chems_mets",

### 2024-02-13 output/paper/_figures.qmd
Finishing MV v1. Working on tables authors.

@@
 # Function to align forest plot with table numerical results
-forest_plot <- function(obj, r = 4) {
+forest_plot <- function(obj, r = 4, is_cortisol = FALSE) {
   edcs <- myphd::edcs_information() |>
@@
   # From table of numerical results to ggplot object
-  tbl_num_res <- obj$numerical_results |>
+  tbl_num_res <- obj$numerical_results
+  if (is_cortisol == FALSE) {
+    tbl_num_res <- tbl_num_res |>
+    dplyr::left_join(
+      edcs[, c("chem_id", "short_name")],
+      by = c("variable" = "chem_id")
+    ) |>
+    dplyr::select(-variable) |>
+    dplyr::rename(variable = "short_name") |>
+    dplyr::relocate(variable)
+  }
+  tbl_num_res <- tbl_num_res |>
     tidylog::mutate(
@@
     tidylog::select(-estimate) |>
-    dplyr::left_join(
-      edcs[, c("chem_id", "short_name")],
-      by = c("variable" = "chem_id")
-    ) |>
-    dplyr::select(-variable) |>
-    dplyr::rename(variable = "short_name") |>
-    dplyr::relocate(variable) |>
     ggplot2::ggplot(
@@
 )
-forest_plot(obj = obj3, r = 5)
+forest_plot(obj = obj3, r = 5, is_cortisol = TRUE)
 ```

### 2024-02-13 output/paper/authors.qmd
Switching to new syntax acr. Adding comments from MV v1.
Finishing MV v1. Working on tables authors.

@@
 linkcolor: blue
+
+editor_options:
+  chunk_output_type: console
 ---
+
+```{r}
+authors <- yaml::read_yaml("../../docs/authors.yml")
+
+proc_one <- function(df, idx) {
+  tmp <- df$authors[[idx]]
+
+  affiliations <- unname(unlist(tmp$affiliations))
+  full_names <- tibble::tibble(unlist(df$affiliations))
+  stringr::str_replace_all()
+
+  ret <- tibble::tribble(
+    ~name, ~affiliations, ~role,
+    tmp$name, paste0(unname(unlist(tmp$affiliations)), collapse = "; "), paste0(tmp$role, collapse = "; ")
+  )
+}
+```

### 2024-02-13 R/proc_res.R

@@ -842,6 +842,11 @@ tidy_res_meffects <- function(df, sa_var, outcome,
         "cortisol metabolism" ~ "cortisol met.",
         "cortisone metabolism" ~ "cortisone met.",
         .default = variable
+      ),
+      outcome = dplyr::case_match(
+        outcome,
+        "hitrtse" ~ "HRT-SE",
+        .default = outcome
       )
     )
   

### 2024-02-13 output/paper/authors.qmd
PDF for authors.

--- 2024-02-06 output/paper/authors.qmd
@@
 linkcolor: blue
+
+editor_options:
+  chunk_output_type: console
 ---
+
+```{r}
+proc_one <- function(df, idx) {
+  tmp <- df$authors[[idx]]
+
+  affiliations <- unname(unlist(tmp$affiliations))
+  full_names <- tibble::tibble(
+    id = unlist(lapply(df$affiliations, "[[", "id")),
+    long = unlist(lapply(df$affiliations, "[[", "name"))
+  ) |>
+    dplyr::mutate(
+      long = stringr::str_replace_all(
+        long, "\n", ""
+      )
+    )
+  affiliations_long <- unlist(lapply(
+    affiliations,
+    function(x) {
+      full_names[full_names$id == x, ]$long
+    }
+  ))
+  affiliations_long <- paste0(affiliations_long, collapse = ".<br>")
+
+  ret <- tibble::tribble(
+    ~name, ~affiliations, ~role,
+    tmp$name, affiliations_long, paste0(tmp$role, collapse = "<br>")
+  )
+}
+```
+
+```{r}
+authors <- yaml::read_yaml("../../docs/authors.yml")
+lapply(seq_along(authors$authors), function(x) {
+  proc_one(authors, x)
+}) |>
+  dplyr::bind_rows() |>
+  gt::gt() |>
+  gt::fmt_markdown(columns = dplyr::everything()) |>
+  gt::gtsave("authors.pdf")
+```

### 2024-02-13 output/paper/manuscript.qmd
Added references biological mechanisms EDCs to cortisol as MV v1.

@@
 We are not aware of other epidemiological studies investigating phthalates metabolites, phenols, and {{< acr opps >}}, in relation to urinary glucocorticosteroids in childhood.
+Nonetheless, previous toxicological studies provide evidence for the inhibition by phthalates of human 11$\beta$-hydroxysteroid dehydrogenase 2 (11$\beta$-HSD2) activity, responsible for the conversion of active cortisol into inactive cortisone [@ZhaoChuHuang:2010; @MaLianDong:2011].
+There is also *in silico* evidence suggesting that {{< acr bpa >}}, a phenol, and Triazophos (TAP), a organophosphorus insecticide, can bind to the human glucocorticoid receptor [@PrasanthDivyaSadasivan:2010; @YangLiRen:2019].
 

### 2024-02-14 output/paper/manuscript.qmd
Minor revisions paper.

@@
 The sets of covariates were selected to estimate the total effect of the exposure on the outcome.
+For effect estimation of the {{< acr edcs >}} on glucocorticosteroids and of glucocorticosteroids on {{< acr hitrtse >}}, these sets were also sufficient to estimate direct effects.
 Further, each minimal adjustment set was *augmented* with precision covariates, defined as the set of parents variable of the outcome that are not parents of the exposure.
@@
 Similarly, the {{< acr mcs >}} were aggregated separately for each level of sex.
-We further tested for significance of the difference between the {{< acr mcs >}} of males and females.
+We further tested for significance of the difference between the {{< acr mcs >}} of females and males.
 

### 2024-02-15 output/paper/_si.qmd
Added class EDCs to SM too.

@@
 )
+
+edcs <- myphd::edcs_information() |>
+  dplyr::filter(!chem_id %in% c("dedtp", "dmdtp")) |>
+  dplyr::select(chem_id, short_name)
 ```
@@
 balance_1 |>
-  tidylog::rename(Exposure = "exposure") |>
+  dplyr::left_join(edcs, by = c("exposure" = "chem_id")) |>
+  dplyr::select(-exposure) |>
+  dplyr::rename(Exposure = "short_name") |>
+  dplyr::relocate(Exposure) |>
   tidylog::group_by(class) |>
@@
 balance_2 |>
+  dplyr::left_join(edcs, by = c("exposure" = "chem_id")) |>
+  dplyr::select(-exposure) |>
+  dplyr::rename(Exposure = "short_name") |>
+  dplyr::relocate(Exposure) |>
   tidylog::group_by(class) |>
@@
 balance_3 |>
+  dplyr::rename(Exposure = "exposure") |>
   dplyr::arrange(dplyr::desc(Adjusted)) |>
@@
 )
+
+edcs <- myphd::edcs_information() |>
+    dplyr::filter(!chem_id %in% c("dedtp", "dmdtp")) |>
+    dplyr::select(chem_id, short_name, class)
 ```
@@
   sa_var = NULL
-)$table
+)$table |>
+  gt::tab_row_group(
+    label = "OP pesticide metabolites",
+    rows = variable %in% as.character(edcs[edcs$class == "OP pesticide metabolites", ]$short_name)
+  ) |>
+  gt::tab_row_group(
+    label = "Phenols",
+    rows = variable %in% as.character(edcs[edcs$class == "Phenols", ]$short_name)
+  ) |>
+  gt::tab_row_group(
+    label = "Phthalate metabolites",
+    rows = variable %in% as.character(edcs[edcs$class == "Phthalate metabolites", ]$short_name)
+  ) |>
+  gt::row_group_order(
+    groups = c("OP pesticide metabolites",
+               "Phenols",
+               "Phthalate metabolites")
+  ) |>
+  gt::tab_style(
+    locations = gt::cells_row_groups(group = dplyr::everything()),
+    style = list(gt::cell_text(weight = "bold"))
+  )
 ```
@@
   sa_var = "e3_sex"
-)$table
+)$table |>
+  gt::tab_row_group(
+    label = "OP pesticide metabolites",
+    rows = variable %in% as.character(edcs[edcs$class == "OP pesticide metabolites", ]$short_name)
+  ) |>
+  gt::tab_row_group(
+    label = "Phenols",
+    rows = variable %in% as.character(edcs[edcs$class == "Phenols", ]$short_name)
+  ) |>
+  gt::tab_row_group(
+    label = "Phthalate metabolites",
+    rows = variable %in% as.character(edcs[edcs$class == "Phthalate metabolites", ]$short_name)
+  ) |>
+  gt::row_group_order(
+    groups = c("OP pesticide metabolites",
+               "Phenols",
+               "Phthalate metabolites")
+  ) |>
+  gt::tab_style(
+    locations = gt::cells_row_groups(group = dplyr::everything()),
+    style = list(gt::cell_text(weight = "bold"))
+  )
 ```

### 2024-02-15 output/paper/manuscript.qmd
Added class EDCs to SM too.

@@
 We used the parametric g-formula and {{< acr mcs >}} to estimate effects between {{< acr edcs >}}, glucocorticosteroids, and {{< acr hitrtse >}}, a measure of efficiency from the {{< acr ant >}}, and tested for possible effect modification by sex.
-We observed positive {{< acr mcs >}} for exposure increases from the 10th to the 90th percentile for methyl-paraben (`r format_res(meffects1, "mepa", "hitrtse")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "hitrtse")`), oh-MiNP (`r format_res(meffects1, "ohminp", "hitrtse")`), and MEHP (`r format_res(meffects1, "mehp", "hitrtse")`), on {{< acr hitrtse >}}.
+We observed positive {{< acr mcs >}} for exposure increases from the 10th to the 90th percentile for methyl-paraben (`r format_res(meffects1, "mepa", "HRT-SE")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "HRT-SE")`), oh-MiNP (`r format_res(meffects1, "ohminp", "HRT-SE")`), and MEHP (`r format_res(meffects1, "mehp", "HRT-SE")`), on {{< acr hitrtse >}}.
 Several {{< acr edcs >}} were also associated with positive {{< acr mcs >}} of cortisone, cortisol, and corticosterone production.
@@
 Most of the {{< acr cis first_use=true >}} included the null effect, though.
-Statistically significant effects were observed for the phenol MEPA (`r format_res(meffects1, "mepa", "hitrtse")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "hitrtse")`), oh-MiNP (`r format_res(meffects1, "ohminp", "hitrtse")`), and MEHP (`r format_res(meffects1, "mehp", "hitrtse")`).
-The {{< acr opp >}} DETP was strongly negatively associated with {{< acr hitrtse >}} (`r format_res(meffects1, "detp", "hitrtse")`).
+Statistically significant effects were observed for the phenol MEPA (`r format_res(meffects1, "mepa", "HRT-SE")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "HRT-SE")`), oh-MiNP (`r format_res(meffects1, "ohminp", "HRT-SE")`), and MEHP (`r format_res(meffects1, "mehp", "HRT-SE")`).
+The {{< acr opp >}} DETP was strongly negatively associated with {{< acr hitrtse >}} (`r format_res(meffects1, "detp", "HRT-SE")`).
 
@@
 @Tbl-hypothesis-1and2 presents the results of a hypothesis test for the difference between estimates of the {{< acr mcs >}} for females and males, for each exposure.
-Significant differences were present for the phenol OXBE (`r format_res(heffects1, "oxbe", "hitrtse")`) and the phthalate metabolites MEP (`r format_res(heffects1, "mep", "hitrtse")`) and MbZP (`r format_res(heffects1, "mbzp", "hitrtse")`).
+Significant differences were present for the phenol OXBE (`r format_res(heffects1, "oxbe", "HRT-SE")`) and the phthalate metabolites MEP (`r format_res(heffects1, "mep", "HRT-SE")`) and MbZP (`r format_res(heffects1, "mbzp", "HRT-SE")`).
 The forest plot of the individual {{< acr mcs >}} is presented in @suppfig-marginal-1sa.
@@
 @Tbl-hypothesis-3 presents the results of a hypothesis test for the difference between estimates of the {{< acr mcs >}} for females and males, for each exposure.
-Significant differences were present for cortisone production (`r format_res(heffects3, "cortisone prod.", "hitrtse")`) and corticosterone production (`r format_res(heffects3, "corticost.
-prod.", "hitrtse")`).
+Significant differences were present for cortisone production (`r format_res(heffects3, "cortisone prod.", "HRT-SE")`) and corticosterone production (`r format_res(heffects3, "corticost.
+prod.", "HRT-SE")`).
 Furthermore, for all exposures, the {{< acr mcs >}} had opposite sign (positive for males and negative for females).

### 2024-02-15 output/paper/_si.qmd
Reviewing manuscript.

@@
 ```
-
-- Files for supplementary data must be accompanied by a summary of the file names and types.
 
@@
   ) |>
-  tidylog::select(-metabolite, Metabolite, LLOQ)
+  tidylog::select(-metabolite) |>
+  tidylog::select(Metabolite, LLOQ)
 gt::gt(lloq_mets |> dplyr::arrange(Metabolite)) |>
@@
     marks = "letters"
+  ) |>
+  gt::tab_style(
+    locations = gt::cells_column_labels(),
+    style = list(gt::cell_text(weight = "bold"))
   )
@@
 ```
-Flowchart describing thew sample size for each research question.
+Flowchart describing the sample size for each research question.
 :::
@@
  (HELIX subcohort; 2013-2016).** 
-Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}.
+Circles and triangles indicate effect estimates.
+Solid lines indicate the $95\%$ {{< acr ci >}}.
+Abbreviations: cortisone production (cortisone prod.); cortisol production (cortisol prod.); corticost.
+prod.
+(corticosterone production).
 :::

### 2024-02-15 output/paper/_tables.qmd
Reviewing manuscript.

@@
 ```
-**Participants derived glucocorticosteroids concentrations expressed in ng/ml, by cohort and overall (HELIX subcohort; 2013-2016).**
+**Participants derived glucocorticosteroids concentrations expressed in ng/ml (HELIX subcohort; 2013-2016).**
 :::

### 2024-02-15 output/paper/manuscript.qmd
Reviewing manuscript.

@@
 id_var <- "HelixID"
-num_digits_est = 2
+num_digits_est = 3
 num_digits_sig = 2
@@
 
-Epidemiological evidence suggests that exposure to {{< acr edcs >}} may be linked to perturbation of the {{< acr hpa >}} axis, which has a major role in brain development.
-We aimed to evaluate the effects of childhood exposure to organophosphate pesticides, phenols, and phthalate metabolites, on urinary glucocorticosteroids and inattention measures in children.
-We used data from the {{< acr helix >}} cohort including 1,297 children aged 6-11.
-We used the parametric g-formula and {{< acr mcs >}} to estimate effects between {{< acr edcs >}}, glucocorticosteroids, and {{< acr hitrtse >}}, a measure of efficiency from the {{< acr ant >}}, and tested for possible effect modification by sex.
-We observed positive {{< acr mcs >}} for exposure increases from the 10th to the 90th percentile for methyl-paraben (`r format_res(meffects1, "mepa", "HRT-SE")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "HRT-SE")`), oh-MiNP (`r format_res(meffects1, "ohminp", "HRT-SE")`), and MEHP (`r format_res(meffects1, "mehp", "HRT-SE")`), on {{< acr hitrtse >}}.
-Several {{< acr edcs >}} were also associated with positive {{< acr mcs >}} of cortisone, cortisol, and corticosterone production.
-Increased levels of the glucocorticosteroids had no effect on {{< acr hitrtse >}}, although we found possible effect modification by sex for cortisone and corticosterone production.
+Epidemiological evidence suggests that {{< acr edcs >}} may perturb the {{< acr hpa >}} axis, which has a major role in brain development.
+We aimed to evaluate the effects of childhood exposure to organophosphate pesticides, phenols, and phthalate metabolites, on urinary glucocorticosteroids and inattention measures in children using data from the {{< acr helix >}} cohort.
+We used the parametric g-formula to estimate effects between {{< acr edcs >}}, glucocorticosteroids, and {{< acr hitrtse >}}, a measure of efficiency from the {{< acr ant >}}, and tested for possible effect modification by sex.
+We observed a positive {{< acr mc >}} for exposure increases from the 10th to the 90th percentile for methyl-paraben (`r format_res(meffects1, "mepa", "HRT-SE")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "HRT-SE")`), oh-MiNP (`r format_res(meffects1, "ohminp", "HRT-SE")`), and MEHP (`r format_res(meffects1, "mehp", "HRT-SE")`), on {{< acr hitrtse >}}.
+Several {{< acr edcs >}} were also associated with a positive {{< acr mc >}} for cortisone, cortisol, and corticosterone production.
+Increased levels of the glucocorticosteroids had no effect on {{< acr hitrtse >}}, although we found a possible effect modification by sex.
 Our results suggest that multiple {{< acr edcs >}} might interfere with inattentiveness outcomes in children and with the homeostasis of the {{< acr hpa >}} axis.
-Further research is necessary to determine whether this can have a clinically significant impact on childhood neurodevelopment.
 
@@
 
-- Authorship must follow the ICMJE's criteria.
-Author names should be listed in ScholarOne and author contributions should be detailed in the cover letter ("Author A...").
-Neither names nor contributions should appear in the blinded manuscript.
-Do not include conflicts of interest.
-- The length of the abstract must be less or equal than 200 words, and should be unstructured, stating the research questions, the methods used, and the results and conclusions of the research.
-- The length must be <4,000 words, excluding abstract, references, tables, figure captions, acknowledgments, and Supplementary Material.
-- Submission: SER members receive a 10% discount on the fees per page.
-This should be requested in the cover letter.
-Also 20% discount for Open Access charges.
-If uploading a single file (document, tables, figures, SM), designate the file as "Main document - anonymous".
-
 The prevalence of several neurodevelopmental disorders has increased in the pediatric population [@GrandjeanLandrigan:2014], and multiple environmental pollutants may play a role in the increased rates of these disorders [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
-{{< acr edcs first_use=true >}}, ubiquitous chemicals present in many every-day products and diet, are capable of interfering with the endocrine system, and have shown associations with childhood neurodevelopment and behavior [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @HuangChenSu:2015; @CartierWarembourgLeManer-Idrissi:2016; @TewarAuingerBraun:2016; @YuDuChiou:2016; @HuangTsaiChen:2017; @KimHongShin:2017; @FurlongHerringBuckley:2017; @Braun:2017; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019; @ShoaffCoullWeuve:2020; @OhKimKannan:2023; @VilmandBeckBilenberg:2023].
+Multiple {{< acr edcs first_use=true >}}, ubiquitous chemicals present in many every-day products and diet, are capable of interfering with the endocrine system, and have shown associations with childhood neurodevelopment and behavior [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @HuangChenSu:2015; @CartierWarembourgLeManer-Idrissi:2016; @TewarAuingerBraun:2016; @YuDuChiou:2016; @HuangTsaiChen:2017; @KimHongShin:2017; @FurlongHerringBuckley:2017; @Braun:2017; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019; @ShoaffCoullWeuve:2020; @OhKimKannan:2023; @VilmandBeckBilenberg:2023].
 Although both pregnancy and early childhood are crucial stages of (neuro)development, most of the available literature is focused on the effects of prenatal exposure to {{< acr edcs >}} on child neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
@@
 
-Taken together, these results suggest that the negative influence of exposure to certain {{< acr edcs >}} on neurodevelopmental outcomes, might be mediated, at least partially, by disruption of the {{< acr hpa >}} axis' homeostasis.
+Taken together, these results suggest that the negative influence of exposure to certain {{< acr edcs >}} on neurodevelopmental outcomes might be mediated, at least partially, by disruption of the {{< acr hpa >}} axis' homeostasis.
 In the present study, we thus estimated associations between 1) non-persistent {{< acr edcs >}} and attention, 2) non-persistent {{< acr edcs >}} and glucocorticosteroids, and 3) glucocorticosteroids and attention, using the parametric g-formula and {{< acr mcs first_use=true >}}, in children of a large cohort in Europe.
@@
 For effect estimation of the {{< acr edcs >}} on glucocorticosteroids and of glucocorticosteroids on {{< acr hitrtse >}}, these sets were also sufficient to estimate direct effects.
+Sample-specific creatinine values were used to adjust for possible dilution effects.
 Further, each minimal adjustment set was *augmented* with precision covariates, defined as the set of parents variable of the outcome that are not parents of the exposure.
@@
 Codebooks for the used covariates, by research question, are provided in Supplementary Tables [-@supptbl-codebook-1; -@supptbl-codebook-2; -@supptbl-codebook-3].
-
-- For RQ1 I used creatinine values from HELIX. For RQ3 the ones from the steroids dataset. For RQ2, I included in the model both variables.
 
@@
 We used the `check_predictions` function from the `performance` R package using the default arguments [@LudeckeBen-ShacharPatil:2021].
-Values of cortisol production and cortisone production were expressed in nanograms per millilitre (ng/ml), whereas values of cortisol metabolism and 11bHSD activity were unitless.
+Values of total cortisol, cortisone, and corticosterone production were expressed in nanograms per millilitre (ng/ml).
 
@@
 Categorical covariates were imputed using the `maxCat` function, which chooses the level with the most occurrences.
-Creatinine values were expressed in grams per litre.
+Creatinine values were expressed in grams per litre (g/L).
 
@@
 The exposure variable was modeled using natural cubic splines with 3 degrees of freedom, to more flexibly capture the {{< acr adrf >}}.
-When the outcome was a ratio, as was the case for cortisol metabolism and 11bHSD activity, we included the logarithm of its denominator, cortisol and cortisol production, respectively, as a control variable [@BartlettPartnoy:2020].
 
@@
 For each outcome, we report the results as differences between {{< acr mcs >}}.
-
-We further estimated the {{< acr adrf >}} using the `avg_predictions` function from the `marginaleffects` R package, examining 50 exposure values from the 10th to the 90th percentiles of the exposure.
-As done for the {{< acr mcs >}}, we included the estimated balancing weights and used cohort as a clustering variable when computing robust standard errors.
 
@@
 @Fig-marginal-1 presents the forest plot for the {{< acr mcs >}} of the non-persistent {{< acr edcs >}} on {{< acr hitrtse >}}.
-For most {{< acr edcs >}}, a cohort-specific increase in the levels of the exposures from the 10th to the 90th percentiles was associated with a positive {{< acr mc >}}, indicating an increase in the values of {{< acr hitrtse >}}.
+For most {{< acr edcs >}}, a cohort-specific decrease in the levels of the exposures from the 90th to the 10th percentiles was associated with a positive {{< acr mc >}}, indicating an increase in the values of {{< acr hitrtse >}}.
 Most of the {{< acr cis first_use=true >}} included the null effect, though.
 Statistically significant effects were observed for the phenol MEPA (`r format_res(meffects1, "mepa", "HRT-SE")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "HRT-SE")`), oh-MiNP (`r format_res(meffects1, "ohminp", "HRT-SE")`), and MEHP (`r format_res(meffects1, "mehp", "HRT-SE")`).
-The {{< acr opp >}} DETP was strongly negatively associated with {{< acr hitrtse >}} (`r format_res(meffects1, "detp", "HRT-SE")`).
+The {{< acr opp >}} DETP was negatively associated with {{< acr hitrtse >}} (`r format_res(meffects1, "detp", "HRT-SE")`).
 
-@Fig-marginal-2 presents the forest plot for the {{< acr mcs >}} of the non-persistent {{< acr edcs >}} on cortisone production, cortisol production, and corticosterone production.
-For most {{< acr edcs >}}, a cohort-specific increase in the levels of the exposures from the 10th to the 90th percentiles was associated with a positive {{< acr mc >}}, indicating an increase in the total production of these metabolites.
+@Fig-marginal-2 presents the forest plot for the {{< acr mcs >}} of the non-persistent {{< acr edcs >}} on total cortisone, cortisol, and corticosterone production.
+For most {{< acr edcs >}}, a cohort-specific decrease in the levels of the exposures from the 90th to the 10th percentiles was associated with a positive {{< acr mc >}}, indicating an increase in the total production of these metabolites.
 Exceptions were BUPA, which was associated with negative {{< acr mcs >}} for all three outcomes, and MiBP, which was associated with a negative {{< acr mc >}} for total cortisone production only.
@@
 
-@Tbl-hypothesis-1and2 presents the results of a hypothesis test for the difference between estimates of the {{< acr mcs >}} for females and males, for each exposure.
-Significant differences were present for the phenol OXBE (`r format_res(heffects1, "oxbe", "HRT-SE")`) and the phthalate metabolites MEP (`r format_res(heffects1, "mep", "HRT-SE")`) and MbZP (`r format_res(heffects1, "mbzp", "HRT-SE")`).
-The forest plot of the individual {{< acr mcs >}} is presented in @suppfig-marginal-1sa.
+@Tbl-hypothesis-1and2 presents the results of a hypothesis test for the difference between estimates of the {{< acr mcs >}} for females and males, for the {{< acr edcs >}} on the glucocorticosteroids and {{< acr hitrtse >}}.
+For {{< acr hitrtse >}}, significant differences were present for the phenol OXBE (`r format_res(heffects1, "oxbe", "HRT-SE")`) and the phthalate metabolites MEP (`r format_res(heffects1, "mep", "HRT-SE")`) and MbZP (`r format_res(heffects1, "mbzp", "HRT-SE")`).
+For the glucocorticosteroids, significant differences were present across all three classes of {{< acr edcs >}} and for all outcomes.
+The largest differences were attributable to the {{< acr opps >}} DMTP (cortisol production, `r format_res(heffects2, "dmtp", "cortisol production")`) and DETP (corticosterone production, (`r format_res(heffects2, "detp", "corticosterone production")`); cortisone production, (`r format_res(heffects2, "detp", "cortisone production")`)).
+The forest plots of the individual {{< acr mcs >}} are presented in Supplementary Figures -@suppfig-marginal-1sa and -@suppfig-marginal-2sa.
 
-@Tbl-hypothesis-1and2 presents the results of a hypothesis test for the difference between estimates of the {{< acr mcs >}} for females and males, for each exposure and outcome.
-Significant differences were present across all three classes of {{< acr edcs >}} and for all outcomes.
-The largest differences were attributable to the {{< acr opps >}} DMTP (cortisol production, `r format_res(heffects2, "dmtp", "cortisol production")`) and DETP (corticosterone production, (`r format_res(heffects2, "detp", "corticosterone production")`); cortisone production, (`r format_res(heffects2, "detp", "cortisone production")`)).
-The forest plots of the individual {{< acr mcs >}} are presented in @suppfig-marginal-2sa.
-
-@Tbl-hypothesis-3 presents the results of a hypothesis test for the difference between estimates of the {{< acr mcs >}} for females and males, for each exposure.
+@Tbl-hypothesis-3 presents the results of a hypothesis test for the difference between estimates of the {{< acr mcs >}} for females and males, for the glucocorticosteroids on {{< acr hitrtse >}}.
 Significant differences were present for cortisone production (`r format_res(heffects3, "cortisone prod.", "HRT-SE")`) and corticosterone production (`r format_res(heffects3, "corticost.

### 2024-02-15 R/proc_res.R

@@ -1027,6 +1027,14 @@ tidy_res_meffects <- function(df, sa_var, outcome,
       ))
     )
   df_gt <- processed_df |>
+    dplyr::mutate(
+      variable = dplyr::case_match(
+        variable,
+        "cortisone prod." ~ "cortisone production",
+        "cortisol prod." ~ "cortisol production",
+        "corticost. prod." ~ "corticosterone pruduction"
+      )
+    ) |>
     tidylog::select(-estimate) |>
     tidyr::pivot_wider(
       names_from = dplyr::any_of(c("outcome", "modifier")),

### 2024-02-15 _targets_res.R

@@ -7,7 +7,7 @@ targets::tar_option_set(
   format = "qs"
 )
 path_store <- Sys.getenv("path_store")
-num_digits_est <- 2
+num_digits_est <- 3
 num_digits_sig <- 2
 sa_var_ <- "e3_sex"
 

### 2024-02-20 output/paper/_figures.qmd
Reviewing manuscript.
Finished reviewing manuscript before sending to MV before sending to co-authors.

@@
   }
-  tbl_num_res <- tbl_num_res |>
-    tidylog::mutate(
-      variable = forcats::fct_reorder2(
-        variable, estimate, class
-      )
+  tbl_num_res <- dplyr::bind_rows(
+    tibble::tibble(
+      variable = "Exposure",
+      val = "Marginal contrast (95% CI)",
+      estimate = 100,
+      outcome = "",
+      class = "AAA"
+    ),
+    tbl_num_res
+  ) |>
+    dplyr::group_by(class) |>
+    dplyr::arrange(
+      variable,
+      .by_group = TRUE
+    ) |>
+    dplyr::ungroup() |>
+    dplyr::mutate(
+      variable = factor(variable, levels = unique(variable)),
+      variable = forcats::fct_rev(variable)
     ) |>
@@
 **Marginal contrasts for the effect of a decrease from the 90th to the 10th percentile of the glucocorticosteroids on {{< acr hitrtse >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).** 
-Circles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}.
+Circles indicate effect estimates.
+Solid lines indicate the $95\%$ {{< acr ci >}}.
+Abbreviations: cortisone production (cortisone prod.); cortisol production (cortisol prod.); corticost.
+prod.
+(corticosterone production).
 :::

### 2024-02-20 output/paper/_si.qmd
Finished reviewing manuscript before sending to MV before sending to co-authors.

@@
 
-### Study populations
+### Information about the endocrine disruptors
 
-::: {#supptbl-pop-desc}
-```{r pop-desc-full}
-targets::tar_load("table1_full", store = store)
-table1_full
+::: {#supptbl-info-chems}
+```{r info-chems}
+edcs_info <- myphd::edcs_information() |>
+  tidylog::select(-c(chem_id, smiles, ctd_id, exposome_explorer_id)) |>
+  tidylog::mutate(
+    var = janitor::make_clean_names(
+      short_name,
+      replace = c(`-` = ""),
+      case = "none"
+    ) |>
+      tolower()
+  ) |>
+  tidylog::relocate(var, .after = short_name) |>
+  dplyr::group_by(class) |>
+  dplyr::arrange(
+    short_name,
+    .by_group = TRUE
+  )
+gt::gt(edcs_info,
+       groupname_col = "class") |>
+  gt::sub_values(
+    values = c("NA"),
+    replacement = ""
+  ) |>
+  gt::cols_label(
+    full_name = "Compound",
+    short_name = "Symbol",
+    var = "Variable name",
+    pubchem_cid = "PubChem CID",
+    parental_compound = "Parental compound"
+  ) |>
+  gt::tab_style(
+    locations = gt::cells_row_groups(group = dplyr::everything()),
+    style = list(gt::cell_text(weight = "bold"))
+  )
 ```
-**Participant characteristics, by cohort and overall (HELIX subcohort; 2013-2016).**
+**Information about non-persistent {{< acr edcs >}}, including the full compound name, the standard symbol, the used variable name, the identifier from PubChem, and the parental compound.**
 :::
+
+### Information about the glucocorticosteroids
+
+::: {#supptbl-info-mets}
+```{r info-mets}
+mets_info <- readODS::read_ods(
+  path = here::here("docs", "steroids.ods")
+) |>
+  tidylog::select(analyte, acronym, type, hmdb_id, cas_number) |>
+  dplyr::group_by(type) |>
+  dplyr::arrange(
+    analyte,
+    .by_group = TRUE
+  )
+gt::gt(mets_info,
+       groupname_col = "type") |>
+  gt::sub_values(
+    values = c("NA"),
+    replacement = ""
+  ) |>
+  gt::cols_label(
+    analyte = "Metabolite",
+    acronym = "Symbol",
+    hmdb_id = "HMDB ID",
+    cas_number = "CAS number"
+  ) |>
+  gt::tab_style(
+    locations = gt::cells_row_groups(group = dplyr::everything()),
+    style = list(gt::cell_text(weight = "bold"))
+  ) |>
+  gt::tab_footnote(
+    footnote = "Abbreviations: Human Metabolome Database (HMDB); Chemical Abstracts Service (CAS).",
+  ) |>
+  gt::opt_footnote_marks(
+    marks = "letters"
+  )
+```
+**Information about the glucocorticosteroids, including the full metabolite name, the standard symbol, the identifier from the HMDB, and the CAS number.**
+:::
+
+### Codebooks
 
@@
 
-### Description of endocrine disruptors
-
-::: {#supptbl-info-chems}
-```{r info-chems}
-edcs_info <- myphd::edcs_information() |>
-  tidylog::select(-c(chem_id, smiles, ctd_id, exposome_explorer_id)) |>
-  tidylog::mutate(
-    var = janitor::make_clean_names(
-      short_name,
-      replace = c(`-` = ""),
-      case = "none"
-    ) |>
-      tolower()
-  ) |>
-  tidylog::relocate(var, .after = short_name)
-gt::gt(edcs_info,
-       groupname_col = "class") |>
-  gt::sub_values(
-    values = c("NA"),
-    replacement = ""
-  ) |>
-  gt::cols_label(
-    full_name = "Compound",
-    short_name = "Symbol",
-    var = "Variable name",
-    pubchem_cid = "PubChem CID",
-    parental_compound = "Parental compound"
-  ) |>
-  gt::tab_style(
-    locations = gt::cells_row_groups(group = dplyr::everything()),
-    style = list(gt::cell_text(weight = "bold"))
-  )
-```
-**Information about non-persistent {{< acr edcs >}}, including the full compound name, the standard symbol, the used variable name, the identifier from PubChem, and the parental compound.**
-:::
-
-### Description of glucocorticosteroids
-
-::: {#supptbl-met-desc}
-```{r met-desc}
-desc_chems_mets$desc_mets
-```
-**Participants glucocorticosteroids concentrations, by cohort and overall (HELIX subcohort; 2013-2016).**
-:::
-
-::: {#supptbl-info-mets}
-```{r info-mets}
-mets_info <- readODS::read_ods(
-  path = here::here("docs", "steroids.ods")
-) |>
-  tidylog::select(analyte, acronym, type, hmdb_id, cas_number)
-gt::gt(mets_info,
-       groupname_col = "type") |>
-  gt::sub_values(
-    values = c("NA"),
-    replacement = ""
-  ) |>
-  gt::cols_label(
-    analyte = "Metabolite",
-    acronym = "Symbol",
-    hmdb_id = "HMDB ID",
-    cas_number = "CAS number"
-  ) |>
-  gt::tab_style(
-    locations = gt::cells_row_groups(group = dplyr::everything()),
-    style = list(gt::cell_text(weight = "bold"))
-  ) |>
-  gt::tab_footnote(
-    footnote = "Abbreviations: Human Metabolome Database (HMDB); Chemical Abstracts Service (CAS).",
-  ) |>
-  gt::opt_footnote_marks(
-    marks = "letters"
-  )
-```
-**Information about the glucocorticosteroids, including the full metabolite name, the standard symbol, the identifier from the HMDB, and the CAS number.**
-:::
+### Lower limits of quantification of the glucocorticosteroids
 
@@
   tidylog::select(-metabolite) |>
-  tidylog::select(Metabolite, LLOQ)
-gt::gt(lloq_mets |> dplyr::arrange(Metabolite)) |>
+  tidylog::select(Metabolite, LLOQ) |>
+  dplyr::arrange(dplyr::desc(LLOQ))
+gt::gt(lloq_mets) |>
   gt::tab_footnote(
@@
 
+### Study populations
+
+::: {#supptbl-pop-desc}
+```{r pop-desc-full}
+targets::tar_load("table1_full", store = store)
+table1_full
+```
+**Participant characteristics, by cohort and overall (HELIX subcohort; 2013-2016).**
+:::
+
+### Concentrations of the glucocorticosteroids
+
+::: {#supptbl-met-desc}
+```{r met-desc}
+targets::tar_load("desc_chems_mets", store = store)
+desc_chems_mets$desc_mets
+```
+**Participants glucocorticosteroids concentrations, by cohort and overall (HELIX subcohort; 2013-2016).**
+:::
+
 ## Tables for main results
 
-### Balancing weights
+### Balancing weights: sample sizes
 
@@
 
+### Balancing weights: summary statistics
+
 ```{r}
@@
 
-### Balancing weights for effect modification
+### Balancing weights for effect modification: summary statistics
 

### 2024-02-20 output/paper/_tables.qmd
Finished reviewing manuscript before sending to MV before sending to co-authors.

@@
-<!-- title: "Tables" -->
-
-<!-- format: -->
-<!--   docx: -->
-<!--     reference-doc: reference_document.docx -->
-
-<!-- link-citations: true -->
-<!-- linkcolor: blue -->
-
-<!-- editor_options: -->
-<!--   chunk_output_type: console -->
-
-<!-- execute: -->
-<!--   echo: false -->
-<!--   warning: false -->
-<!--   error: false -->
-<!--   message: false -->
-
 ```{r}
@@
 table1
-  # gt::cols_width(
-  #   label ~ gt::px(300)
-  # ) |>
-  # gt::tab_style(
-  #   style = gt::cell_text(size = gt::pct(90)),
-  #   locations = gt::cells_body(
-  #     columns = dplyr::everything()
-  #   )
-  # )
 ```
@@
       cohort == "RHE" ~ "RHEA",
+      cohort == "SAB" ~ "INMA",
       .default = cohort

### 2024-02-20 output/paper/manuscript.qmd
Finished reviewing manuscript before sending to MV before sending to co-authors.

@@
 
+The R code to reproduce analyses and results is available online (https://github.com/lorenzoFabbri/paper-helixSC-neuro).
+
 ### Effect-modification analysis
@@
 Levels of unprocessed non-persistent {{< acr edcs >}}, after imputation of values below the {{< acr lod >}}, and glucocorticosteroids, are presented in @tbl-edc-desc, @tbl-met-new-desc, and @supptbl-met-desc.
+Supplementary Figures [-@suppfig-desc-chems] and [-@suppfig-desc-mets] provide information on the measurement classification of the {{< acr edcs >}} and glucocorticosteroids by cohort, respectively.
 
@@
 The largest differences were attributable to the {{< acr opps >}} DMTP (cortisol production, `r format_res(heffects2, "dmtp", "cortisol production")`) and DETP (corticosterone production, (`r format_res(heffects2, "detp", "corticosterone production")`); cortisone production, (`r format_res(heffects2, "detp", "cortisone production")`)).
-The forest plots of the individual {{< acr mcs >}} are presented in Supplementary Figures -@suppfig-marginal-1sa and -@suppfig-marginal-2sa.
+The forest plots of the individual {{< acr mcs >}} are presented in Supplementary Figures [-@suppfig-marginal-1sa] and [-@suppfig-marginal-2sa].
 

### 2024-02-20 R/proc_res.R

@@ -109,8 +109,12 @@ tbl_desc_pop <- function(num_digits_est, num_digits_sig, by) {
         cohort == "KAN" ~ "KANC",
         cohort == "MOB" ~ "MOBA",
         cohort == "RHE" ~ "RHEA",
+        cohort == "SAB" ~ "INMA",
         .default = cohort
       )
+    ) |>
+    dplyr::rename(
+      Cohort = "cohort"
     )
   ## Handle time variables
   cols_to_season <- c("hs_date_neu", "e3_cbirth")
@@ -118,6 +122,9 @@ tbl_desc_pop <- function(num_digits_est, num_digits_sig, by) {
     dat = dat_request$dat,
     cols = cols_to_season
   ) |>
+    dplyr::mutate(
+      hs_date_neu = stringr::str_to_sentence(hs_date_neu)
+    ) |>
     tidylog::mutate(dplyr::across(
       dplyr::any_of(cols_to_season),
       \(x) factor(x)
@@ -148,6 +155,7 @@ tbl_desc_pop <- function(num_digits_est, num_digits_sig, by) {
   df <- dat_request$dat |>
     tidylog::select(
       HelixID,
+      Cohort,
       params_dat$variables$rq1$outcome,
       dplyr::any_of(all_confounders)
     )
@@ -193,16 +201,15 @@ tbl_desc_pop <- function(num_digits_est, num_digits_sig, by) {
           gtsummary::all_categorical() ~ "{n} ({p}%)"
         ),
         digits = dplyr::everything() ~ 1,
-        missing = "ifany"
-      ) |>
-      gtsummary::as_gt() |>
-      gt::opt_footnote_marks(
-        marks = "letters"
+        missing = "ifany",
+        sort = list(
+          gtsummary::all_categorical() ~ "frequency"
+        )
       )
   } else {
     desc_covars <- df_meta |>
       gtsummary::tbl_summary(
-        by = "cohort",
+        by = "Cohort",
         type = list(
           hs_fastfood ~ "continuous",
           hs_org_food ~ "continuous"
@@ -214,14 +221,23 @@ tbl_desc_pop <- function(num_digits_est, num_digits_sig, by) {
           gtsummary::all_categorical() ~ "{n} ({p}%)"
         ),
         digits = dplyr::everything() ~ 1,
-        missing = "ifany"
+        missing = "ifany",
+        sort = list(
+          gtsummary::all_categorical() ~ "frequency"
+        )
       ) |>
-      gtsummary::add_overall() |>
[…]

### 2024-03-08 output/paper/_figures.qmd
Started reviewing manuscript based on co-authors comments.

@@
 ```
-**Marginal contrasts for the effect of a decrease from the 90th to the 10th percentile of the {{< acr edcs >}} on {{< acr hitrtse >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).** 
-Circles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}.
+**Marginal contrasts for the effect of a decrease from the 90th to the 10th percentile of the {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
+Circles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}. The size of the circles represents the *S* value of the effect estimate [@RafiGreenland:2020].
 :::
@@
 ```
-**Marginal contrasts for the effect of a decrease from the 90th to the 10th percentile of the {{< acr edcs >}} on the glucocorticosteroids expressed in ng/ml (HELIX subcohort; 2013-2016).** 
-Circles, triangles, and squares indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}.
+**Marginal contrasts for the effect of a decrease from the 90th to the 10th percentile of the {{< acr edcs first_use=true >}} on the glucocorticosteroids expressed in ng/ml (HELIX subcohort; 2013-2016).**
+Circles, triangles, and squares indicate effect estimates.
+Solid lines indicate the $95\%$ {{< acr ci >}}.
+The size of the circles represents the *S* value of the effect estimate [@RafiGreenland:2020].
 :::
@@
 ```
-**Marginal contrasts for the effect of a decrease from the 90th to the 10th percentile of the glucocorticosteroids on {{< acr hitrtse >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).** 
+**Marginal contrasts for the effect of a decrease from the 90th to the 10th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
 Circles indicate effect estimates.
 Solid lines indicate the $95\%$ {{< acr ci >}}.
+The size of the circles represents the *S* value of the effect estimate [@RafiGreenland:2020].
 Abbreviations: cortisone production (cortisone prod.); cortisol production (cortisol prod.); corticost.

### 2024-03-08 output/paper/_si.qmd
Started reviewing manuscript based on co-authors comments.

@@
 ```
-**Information about non-persistent {{< acr edcs >}}, including the full compound name, the standard symbol, the used variable name, the identifier from PubChem, and the parental compound.**
+**Information about non-persistent {{< acr edcs first_use=true >}}, including the full compound name, the standard symbol, the used variable name, the identifier from PubChem, and the parental compound.**
 :::
@@
 ```
-**Codebook for the covariates used in the estimation of the marginal comparisons of {{< acr edcs >}} on {{< acr hitrtse >}}.**
+**Codebook for the covariates used in the estimation of the marginal comparisons of {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}}.**
 :::
@@
 ```
-**Codebook for the covariates used in the estimation of the marginal comparisons of {{< acr edcs >}} on the glucocorticosteroids.**
+**Codebook for the covariates used in the estimation of the marginal comparisons of {{< acr edcs first_use=true >}} on the glucocorticosteroids.**
 :::
@@
 ```
-**Codebook for the covariates used in the estimation of the marginal comparisons of the glucocorticosteroids on {{< acr hitrtse >}}.**
+**Codebook for the covariates used in the estimation of the marginal comparisons of the glucocorticosteroids on {{< acr hitrtse first_use=true >}}.**
 :::
@@
 ```
-**Effective sample size before and after balancing weights estimation (exposures: {{< acr edcs >}}; outcome: {{< acr hitrtse >}}) (HELIX subcohort; 2013-2016).**
+**Effective sample size before and after balancing weights estimation (exposures: {{< acr edcs first_use=true >}}; outcome: {{< acr hitrtse first_use=true >}}) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Effective sample size before and after balancing weights estimation (exposures: {{< acr edcs >}}; outcomes: glucocorticosteroids) (HELIX subcohort; 2013-2016).**
+**Effective sample size before and after balancing weights estimation (exposures: {{< acr edcs first_use=true >}}; outcomes: glucocorticosteroids) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Effective sample size before and after balancing weights estimation (exposures: glucocorticosteroids; outcome: {{< acr hitrtse >}}) (HELIX subcohort; 2013-2016).**
+**Effective sample size before and after balancing weights estimation (exposures: glucocorticosteroids; outcome: {{< acr hitrtse first_use=true >}}) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Summary statistics of the estimated balancing weights (exposures: {{< acr edcs >}}; outcome: {{< acr hitrtse >}}) (HELIX subcohort; 2013-2016).**
+**Summary statistics of the estimated balancing weights (exposures: {{< acr edcs first_use=true >}}; outcome: {{< acr hitrtse first_use=true >}}) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Summary statistics of the estimated balancing weights (exposures: {{< acr edcs >}}; outcomes: glucocorticosteroids) (HELIX subcohort; 2013-2016).**
+**Summary statistics of the estimated balancing weights (exposures: {{< acr edcs first_use=true >}}; outcomes: glucocorticosteroids) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Summary statistics of the estimated balancing weights (exposures: glucocorticosteroids; outcome: {{< acr hitrtse >}}) (HELIX subcohort; 2013-2016).**
+**Summary statistics of the estimated balancing weights (exposures: glucocorticosteroids; outcome: {{< acr hitrtse first_use=true >}}) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Summary statistics of the estimated balancing weights for effect modification (exposures: {{< acr edcs >}}; outcome: {{< acr hitrtse >}}; modifier: sex) (HELIX subcohort; 2013-2016).**
+**Summary statistics of the estimated balancing weights for effect modification (exposures: {{< acr edcs first_use=true >}}; outcome: {{< acr hitrtse first_use=true >}}; modifier: sex) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Summary statistics of the estimated balancing weights for effect modification (exposures: {{< acr edcs >}}; outcomes: glucocorticosteroids; modifier: sex) (HELIX subcohort; 2013-2016).**
+**Summary statistics of the estimated balancing weights for effect modification (exposures: {{< acr edcs first_use=true >}}; outcomes: glucocorticosteroids; modifier: sex) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Summary statistics of the estimated balancing weights for effect modification (exposures: glucocorticosteroids; outcome: {{< acr hitrtse >}}; modifier: sex) (HELIX subcohort; 2013-2016).**
+**Summary statistics of the estimated balancing weights for effect modification (exposures: glucocorticosteroids; outcome: {{< acr hitrtse first_use=true >}}; modifier: sex) (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Measurement classification of {{< acr edcs >}}, by cohort (HELIX subcohort; 2013-2016).** 
+**Measurement classification of {{< acr edcs first_use=true >}}, by cohort (HELIX subcohort; 2013-2016).** 
 Coding: 1, quantifiable; 2, <LOD; 3, interference or out of range; 4. not analysed.
@@
 ```
-**Marginal contrasts for effect modification by sex of a decrease from the 90th to the 10th percentile of the {{< acr edcs >}} on {{< acr hitrtse >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
-Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}.
+**Marginal contrasts for effect modification by sex of a decrease from the 90th to the 10th percentile of the {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
+Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}. The size of the circles represents the *S* value of the effect estimate [@RafiGreenland:2020].
 :::
@@
 ```
-**Marginal contrasts for effect modification by sex of a decrease from the 90th to the 10th percentile of the {{< acr edcs >}} on the glucocorticosteroids expressed in ng/ml (HELIX subcohort; 2013-2016).**
-Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}.
+**Marginal contrasts for effect modification by sex of a decrease from the 90th to the 10th percentile of the {{< acr edcs first_use=true >}} on the glucocorticosteroids expressed in ng/ml (HELIX subcohort; 2013-2016).**
+Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}. The size of the circles represents the *S* value of the effect estimate [@RafiGreenland:2020].
 :::
@@
 ```
-**Marginal contrasts for effect modification by sex of a decrease from the 90th to the 10th percentile of the glucocorticosteroids on {{< acr hitrtse >}} expressed in {{< acr mss >}}
- (HELIX subcohort; 2013-2016).** 
+**Marginal contrasts for effect modification by sex of a decrease from the 90th to the 10th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
 Circles and triangles indicate effect estimates.
 Solid lines indicate the $95\%$ {{< acr ci >}}.
+The size of the circles represents the *S* value of the effect estimate [@RafiGreenland:2020].
 Abbreviations: cortisone production (cortisone prod.); cortisol production (cortisol prod.); corticost.

### 2024-03-08 output/paper/_tables.qmd
Started reviewing manuscript based on co-authors comments.

@@
 ```
-**Pairwise differences between sex-specific marginal contrasts for the effect of a decrease from the 90th to the 10th percentile of {{< acr edcs >}} on {{< acr hitrtse >}}, expressed in {{< acr mss >}}, and the glucocorticosteroids, expressed in ng/ml (HELIX subcohort; 2013-2016).**
+**Pairwise differences between marginal contrasts of males and females, for the effect of a decrease from the 90th to the 10th percentile of {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}}, expressed in {{< acr mss >}}, and on the glucocorticosteroids, expressed in ng/ml (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Pairwise differences between sex-specific marginal contrasts for the effect of a decrease from the 90th to the 10th percentile of the glucocorticosteroids on {{< acr hitrtse >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
+**Pairwise differences between marginal contrasts of males and females, for the effect of a decrease from the 90th to the 10th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
 :::

### 2024-03-08 output/paper/manuscript.qmd
Started reviewing manuscript based on co-authors comments.

@@
   )
+  val1 <- stringr::str_split(val, " ", n = 2)[[1]][1]
+  val2 <- stringr::str_split(val, " ", n = 2)[[1]][2]
   glue::glue(
-    "\\acr{{mc}} and $95\\%$ \\acr{{ci}}: {val}"
+    "MC: {val1} and $95\\%$ \\acr{{ci}}: {val2}"
   )
@@
 
-Epidemiological evidence suggests that {{< acr edcs >}} may perturb the {{< acr hpa >}} axis, which has a major role in brain development.
+Evidence suggests that {{< acr edcs >}} may perturb the {{< acr hpa >}} axis, which has a major role in brain development.
 We aimed to evaluate the effects of childhood exposure to organophosphate pesticides, phenols, and phthalate metabolites, on urinary glucocorticosteroids and inattention measures in children using data from the {{< acr helix >}} cohort.
@@
 Multiple {{< acr edcs first_use=true >}}, ubiquitous chemicals present in many every-day products and diet, are capable of interfering with the endocrine system, and have shown associations with childhood neurodevelopment and behavior [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @HuangChenSu:2015; @CartierWarembourgLeManer-Idrissi:2016; @TewarAuingerBraun:2016; @YuDuChiou:2016; @HuangTsaiChen:2017; @KimHongShin:2017; @FurlongHerringBuckley:2017; @Braun:2017; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019; @ShoaffCoullWeuve:2020; @OhKimKannan:2023; @VilmandBeckBilenberg:2023].
-Although both pregnancy and early childhood are crucial stages of (neuro)development, most of the available literature is focused on the effects of prenatal exposure to {{< acr edcs >}} on child neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
+Although both pregnancy and early infancy are crucial stages of (neuro)development, most of the available literature is focused on the effects of prenatal exposure to {{< acr edcs >}} on child neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
 
-One group of {{< acr edcs >}} that may have a deleterious effect on neurodevelopment is the {{< acr opps >}}, although the few studies assessing exposure during childhood and through the use of biomarkers suffered from a series of limitations, including a small sample size and few {{< acr opp >}} metabolites measured [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
-Exposure to phthalates and their metabolites during childhood and early adolescence has been associated to several adverse neurodevelopmental outcomes, but these studies were limited to few phthalate metabolites and small study populations [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
+One group of {{< acr edcs >}} that may have a deleterious effect on neurodevelopment is the {{< acr opps >}}, although the few studies assessing exposure during childhood and through the use of biomarkers suffered from a series of limitations, including a small sample size [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
+Exposure to phthalates and their metabolites during childhood and early adolescence has been associated with several adverse neurodevelopmental outcomes, but these studies were limited to few phthalate metabolites and small study populations [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
 The effects of exposure to {{< acr bpa >}} during childhood on cognitive functions are still unclear [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
@@
 
-Neurodevelopmental outcomes were assessed with standardized, non-linguistic, and culturally blind computer tests, including the {{< acr ant first_use=true >}} [@RuedaFanMcCandliss:2004], which provides a measure of efficiency in three different functions of attention: alerting, orienting, and executive attention.
+Neurodevelopmental outcomes were assessed with standardized, non-linguistic, and culturally blind computer tests, including the {{< acr ant first_use=true >}} [@RuedaFanMcCandliss:2004], which provides a measure of efficiency in three different components of attention: alerting, orienting, and executive attention.
 The tests were administered in a standardized way, and with minimal interference from the field workers.
@@
 The outcome of interest for the present study is the {{< acr hitrtse first_use=true >}} [@SunyerEsnaolaAlvarez-Pedrerol:2015], a measure of response speed consistency throughout the test.
-A high {{< acr hitrtse >}} indicates highly variable reactions, and is considered a measure of inattentiveness.
+A high {{< acr hitrtse >}} indicates highly variable reaction times, and is considered a measure of inattentiveness.
 
@@
 For each metabolite, we computed the fraction of values below the {{< acr loq >}} and not detected, both within each cohort and overall.
-We proceeded to impute these values using half the value of the corresponding {{< acr lloq >}}, for those metabolites that had less than `r threshold_within_lod`% of missings within each cohort and `r threshold_overall_lod`% of missings overall.
+We proceeded to impute these values using half the value of the corresponding {{< acr loq >}}, for those metabolites that had less than `r threshold_within_lod`% of missings within each cohort and `r threshold_overall_lod`% of missings overall.
 Information about the {{< acr lloq >}} for the glucocorticosteroids is provided in @supptbl-lloq-mets.
@@
 The parametric g-formula involves the following steps: 1) fit a outcome model including both covariates and balancing weights; 2) create two new datasets identical to the original one but with the exposure shifted according to a user-specified intervention set by a deterministic function of the observed exposure levels; 3) use the outcome model to compute adjusted predictions in the two counterfactual datasets; 4) compute the difference between the means of the adjusted predictions in the counterfactual datasets.
-The causal parameter of interest $\phi$ was thus specified as the difference in the expected counterfactual outcomes under the shifted exposure levels $\left( \phi^\Delta = \mathbb{E} \left[ Y^{d_1} \right] - \mathbb{E} \left[ Y^{d_2} \right] \right)$.
+The causal parameter of interest was thus specified as the difference in the expected counterfactual outcomes under the shifted exposure levels $\left( \mathbb{E} \left[ Y^{d_1} \right] - \mathbb{E} \left[ Y^{d_2} \right] \right)$.
 In order for this parameter to be identified, the usual causal identifiability conditions (no unmeasured confounding, positivity, and consistency) are required.
@@
 A flowchart describing the sample size for each research question is presented in @suppfig-flow-pop.
-The sample consisted of 55% girls.
+The sample consisted of 55% males.
 The median {{< acr hitrtse >}} was 300 ms ({{< acr iqr >}}, 231-368), with lower median values for EDEN, MOBA, and INMA, corresponding to the cohorts with older children.
@@
 @Fig-marginal-1 presents the forest plot for the {{< acr mcs >}} of the non-persistent {{< acr edcs >}} on {{< acr hitrtse >}}.
-For most {{< acr edcs >}}, a cohort-specific decrease in the levels of the exposures from the 90th to the 10th percentiles was associated with a positive {{< acr mc >}}, indicating an increase in the values of {{< acr hitrtse >}}.
+For most {{< acr edcs >}}, a cohort-specific decrease in the levels of the exposures from the 90th to the 10th percentiles was associated with a positive {{< acr mc >}}, indicating an increase in the values of {{< acr hitrtse >}} and thus lower attention.
 Most of the {{< acr cis first_use=true >}} included the null effect, though.
-Statistically significant effects were observed for the phenol MEPA (`r format_res(meffects1, "mepa", "HRT-SE")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "HRT-SE")`), oh-MiNP (`r format_res(meffects1, "ohminp", "HRT-SE")`), and MEHP (`r format_res(meffects1, "mehp", "HRT-SE")`).
+Significant effects were observed for the paraben MEPA (`r format_res(meffects1, "mepa", "HRT-SE")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "HRT-SE")`), oh-MiNP (`r format_res(meffects1, "ohminp", "HRT-SE")`), and MEHP (`r format_res(meffects1, "mehp", "HRT-SE")`).
 The {{< acr opp >}} DETP was negatively associated with {{< acr hitrtse >}} (`r format_res(meffects1, "detp", "HRT-SE")`).
@@
 
-@Tbl-hypothesis-1and2 presents the results of a hypothesis test for the difference between estimates of the {{< acr mcs >}} for females and males, for the {{< acr edcs >}} on the glucocorticosteroids and {{< acr hitrtse >}}.
+@Tbl-hypothesis-1and2 presents the results of the difference between estimates of the {{< acr mcs >}} for females and males, for the {{< acr edcs >}} on the glucocorticosteroids and {{< acr hitrtse >}}.
 For {{< acr hitrtse >}}, significant differences were present for the phenol OXBE (`r format_res(heffects1, "oxbe", "HRT-SE")`) and the phthalate metabolites MEP (`r format_res(heffects1, "mep", "HRT-SE")`) and MbZP (`r format_res(heffects1, "mbzp", "HRT-SE")`).
@@
 
-@Tbl-hypothesis-3 presents the results of a hypothesis test for the difference between estimates of the {{< acr mcs >}} for females and males, for the glucocorticosteroids on {{< acr hitrtse >}}.
+@Tbl-hypothesis-3 presents the results of the difference between estimates of the {{< acr mcs >}} for females and males, for the glucocorticosteroids on {{< acr hitrtse >}}.
 Significant differences were present for cortisone production (`r format_res(heffects3, "cortisone prod.", "HRT-SE")`) and corticosterone production (`r format_res(heffects3, "corticost.
@@
 While research in this area has mainly investigated the effects of prenatal exposure on child neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022], little is still known about childhood exposure.
-In this study, consisting of 1,297 children from 6 European birth cohorts, we observed that short-term childhood exposure to non-persistent {{< acr edcs >}} had negative effects on {{< acr hitrtse >}} and total production of cortisol, cortisone, and corticosterone, although the majority of the {{< acr cis >}} included the null effect.
+In this study, consisting of 1,297 children from 6 European birth cohorts, we observed that short-term childhood exposure to non-persistent {{< acr edcs >}} was associated with {{< acr hitrtse >}}, a measure of inattentiveness, and with total production of cortisol, cortisone, and corticosterone, although the majority of the {{< acr cis >}} included the null effect.
 Increased production of these glucocorticosteroids did not seem to affect {{< acr hitrtse >}}.
@@
 We are not aware of prior studies specifically investigating the effects of exposure to {{< acr edcs >}} in relation to {{< acr hitrtse >}}.
-The literature on {{< acr edcs >}} and neurodevelopment in children has mostly focused on {{< acr opps >}} [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @CartierWarembourgLeManer-Idrissi:2016; @YuDuChiou:2016], phthalate metabolites [@HuangChenSu:2015; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023], and {{< acr bpa >}} [@TewarAuingerBraun:2016; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019], in relation to {{< acr adhd >}} [@BouchardBellingerWright:2010; @TewarAuingerBraun:2016; @YuDuChiou:2016; @LiZhangKuang:2018], and intelligence scales [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @HuangChenSu:2015; @CartierWarembourgLeManer-Idrissi:2016; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023].
-Few studies have looked into different classes of {{< acr edcs >}} [@ShoaffCoullWeuve:2020 in relation with the Conners Attention Deficit Scale and the Behavior Assessment System for Children; @OhKimKannan:2023 in relation with {{< acr adhd >}} symptoms].
-Overall, and consistent with our results, these studies seem to provide further evidence of the adverse effects of several {{< acr edcs >}} on neurodevelopment in children.
-While not all these studies have investigated effect modification by sex, it seems that these adverse effects are stronger in males.
-A major limitation of these studies is the reliance on spot urine samples, that might not be representative of long-term exposures.
+The literature on non-persistent {{< acr edcs >}} and neurodevelopment in children has mostly focused on {{< acr opps >}} [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @CartierWarembourgLeManer-Idrissi:2016; @YuDuChiou:2016], phthalate metabolites [@HuangChenSu:2015; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023], and {{< acr bpa >}} [@TewarAuingerBraun:2016; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019], in relation to {{< acr adhd >}} [@BouchardBellingerWright:2010; @TewarAuingerBraun:2016; @YuDuChiou:2016; @LiZhangKuang:2018], and intelligence scales [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @HuangChenSu:2015; @CartierWarembourgLeManer-Idrissi:2016; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023].
+Few studies have looked into different classes of non-persistent {{< acr edcs >}} [@ShoaffCoullWeuve:2020 in relation with the Conners Attention Deficit Scale and the Behavior Assessment System for Children; @OhKimKannan:2023 in relation with {{< acr adhd >}} symptoms].
+Overall, and consistent with our results, these studies seem to provide further evidence of the adverse effects of several non-persistent {{< acr edcs >}} on neurodevelopment in children.
+While not all these studies have investigated effect modification by sex, it seems that these adverse effects tend to be stronger in males [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @YuDuChiou:2016; @ShoaffCoullWeuve:2020], although this difference might depend on the timing of exposure [@].
 
-Our results are consistent with prior epidemiological research that associated exposure to certain {{< acr edcs >}} with higher levels of cortisol [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023].
+Our results are consistent with prior epidemiological research that associated exposure to certain non-persistent {{< acr edcs >}} with higher levels of cortisol [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023].
 There are some differences, though.
@@
 Although we included a wide range of confounders there is the possibility, as with other observational studies, of residual confounding, which might lead to a bias away from the null.
-Some of the confounders indicated in the adjustment sets had to be remove due to large fractions of missing values.
+Some of the confounders indicated in the adjustment sets had to be removed due to large fractions of missing values.
 There is further the possibility of misspecification of the outcome model, although we included a spline of the exposure to relax some of the linearity assumptions.
@@
 
-Strengths of the present study include the use of pooled urine samples for chemical assessment, since it is known that these specific {{< acr edcs >}} have very short half-lives [@CasasBasaganaSakhi:2018].
+Strengths of the present study include the use of pooled urine samples for chemical assessment to obtain more representative long-term exposures, since it is known that these specific {{< acr edcs >}} have very short half-lives [@PerrierGiorgis-AllemandSlama:2016; @CasasBasaganaSakhi:2018].
 We decided to model both the *treatment* mechanisms, for the estimation of balancing weights, and the outcomes, with traditional covariates adjustment, to try to obtain *doubly robust* effect estimates.
@@
 In conclusion, in a study of 1,297 children from 6 European birth cohorts, we observed that (i) exposure to non-persistent {{< acr edcs >}} in childhood might have short-term effects on {{< acr hitrtse >}} in childhood, (ii) exposure to non-persistent {{< acr edcs >}} in childhood might disrupt the {{< acr hpa >}} axis in childhood, and (iii) disruption of the {{< acr hpa >}} axis in childhood might have short-term, sex-specific effects on {{< acr hitrtse >}}.
-Future studies should investigate how glucocorticosteroids might mediate the adverse effects of exposure to {{< acr edcs >}} on childhood neurodevelopment in larger populations.
+Future studies should investigate how glucocorticosteroids might mediate the adverse effects of exposure to non-persistent {{< acr edcs >}} on childhood neurodevelopment in larger populations.
 

### 2024-03-08 R/data.R

@@ -408,15 +408,15 @@ load_dat_request <- function() {
       ),
       e3_marital = factor(e3_marital,
                           levels = c(
-                            "living alone",
                             "living with the father",
+                            "living alone",
                             "other situation"
                           ),
-                          labels = c(1, 0, 2)
+                          labels = c(0, 1, 2)
       ),
       e3_sex = factor(e3_sex,
-                      levels = c("female", "male"),
-                      labels = c(1, 0)
+                      levels = c("male", "female"),
+                      labels = c(0, 1)
       )
     ) |>
     tidylog::rename(

### 2024-03-10 output/paper/manuscript.qmd
Started reviewing manuscript based on co-authors comments.

@@
 ---
-title: "Childhood exposure to non-persistent endocrine disruptors, glucocorticosteroids, and neurodevelopment: A study based on the parametric g-formula"
+title: "Childhood exposure to non-persistent endocrine disruptors, glucocorticosteroids, and attentional function: A study based on the parametric g-formula"
 
@@
 Evidence suggests that {{< acr edcs >}} may perturb the {{< acr hpa >}} axis, which has a major role in brain development.
-We aimed to evaluate the effects of childhood exposure to organophosphate pesticides, phenols, and phthalate metabolites, on urinary glucocorticosteroids and inattention measures in children using data from the {{< acr helix >}} cohort.
-We used the parametric g-formula to estimate effects between {{< acr edcs >}}, glucocorticosteroids, and {{< acr hitrtse >}}, a measure of efficiency from the {{< acr ant >}}, and tested for possible effect modification by sex.
+We aimed to evaluate the effects of childhood exposure to organophosphate pesticides, phenols, and phthalate metabolites, on urinary glucocorticosteroids and inattention in children using data from the {{< acr helix >}} cohort.
+We used the parametric g-formula to estimate effects between {{< acr edcs >}}, glucocorticosteroids, and {{< acr hitrtse >}}, a measure of inattention from the {{< acr ant >}}, and tested for possible effect modification by sex.
 We observed a positive {{< acr mc >}} for exposure increases from the 10th to the 90th percentile for methyl-paraben (`r format_res(meffects1, "mepa", "HRT-SE")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "HRT-SE")`), oh-MiNP (`r format_res(meffects1, "ohminp", "HRT-SE")`), and MEHP (`r format_res(meffects1, "mehp", "HRT-SE")`), on {{< acr hitrtse >}}.
@@
 Increased levels of the glucocorticosteroids had no effect on {{< acr hitrtse >}}, although we found a possible effect modification by sex.
-Our results suggest that multiple {{< acr edcs >}} might interfere with inattentiveness outcomes in children and with the homeostasis of the {{< acr hpa >}} axis.
+Our results suggest that multiple {{< acr edcs >}} might interfere with inattention in children and with the homeostasis of the {{< acr hpa >}} axis.
 
@@
 One group of {{< acr edcs >}} that may have a deleterious effect on neurodevelopment is the {{< acr opps >}}, although the few studies assessing exposure during childhood and through the use of biomarkers suffered from a series of limitations, including a small sample size [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
-Exposure to phthalates and their metabolites during childhood and early adolescence has been associated with several adverse neurodevelopmental outcomes, but these studies were limited to few phthalate metabolites and small study populations [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
+Exposure to phthalates and their metabolites during childhood and early adolescence has also been associated with several adverse neurodevelopmental outcomes, but these studies were limited to few phthalate metabolites and small study populations [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
 The effects of exposure to {{< acr bpa >}} during childhood on cognitive functions are still unclear [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].
@@
 Taken together, these results suggest that the negative influence of exposure to certain {{< acr edcs >}} on neurodevelopmental outcomes might be mediated, at least partially, by disruption of the {{< acr hpa >}} axis' homeostasis.
-In the present study, we thus estimated associations between 1) non-persistent {{< acr edcs >}} and attention, 2) non-persistent {{< acr edcs >}} and glucocorticosteroids, and 3) glucocorticosteroids and attention, using the parametric g-formula and {{< acr mcs first_use=true >}}, in children of a large cohort in Europe.
+In the present study, we thus estimated cross-sectional associations between 1) non-persistent {{< acr edcs >}} and attentional function, 2) non-persistent {{< acr edcs >}} and glucocorticosteroids, and 3) glucocorticosteroids and attentional function, using the parametric g-formula and {{< acr mcs first_use=true >}}, in children of a large network of cohorts in Europe.
 
@@
 
-### Neurodevelopment {#sec-neurodevelopment}
+### Attentionall function {#sec-neurodevelopment}
 
-Neurodevelopmental outcomes were assessed with standardized, non-linguistic, and culturally blind computer tests, including the {{< acr ant first_use=true >}} [@RuedaFanMcCandliss:2004], which provides a measure of efficiency in three different components of attention: alerting, orienting, and executive attention.
+Cognitive and motor function outcomes were assessed with standardized, non-linguistic, and culturally blind computer tests, including the {{< acr ant first_use=true >}} [@RuedaFanMcCandliss:2004], which provides a measure of efficiency of attentional function.
 The tests were administered in a standardized way, and with minimal interference from the field workers.
@@
 To reduce the effect of measured confounders on the exposure-outcome association, stabilized balancing weights were estimated using the energy method available in the `WeightIt` R package [@Greifer:2023].
-This methods estimates weights by minimizing an energy statistic related to covariate balance [@HulingGreiferChen:2023], thus avoiding the need to specify a parametric model.
+This method estimates weights by minimizing an energy statistic related to covariate balance [@HulingGreiferChen:2023], thus avoiding the need to specify a parametric model.
 Weights below the 0.1 and above the 0.9 quantiles were trimmed.
@@
 While research in this area has mainly investigated the effects of prenatal exposure on child neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022], little is still known about childhood exposure.
-In this study, consisting of 1,297 children from 6 European birth cohorts, we observed that short-term childhood exposure to non-persistent {{< acr edcs >}} was associated with {{< acr hitrtse >}}, a measure of inattentiveness, and with total production of cortisol, cortisone, and corticosterone, although the majority of the {{< acr cis >}} included the null effect.
+In this study, consisting of 1,297 children from 6 European birth cohorts, we observed that short-term childhood exposure to non-persistent {{< acr edcs >}} was associated with attentional function, and with total production of cortisol, cortisone, and corticosterone, although the majority of the {{< acr cis >}} included the null effect.
 Increased production of these glucocorticosteroids did not seem to affect {{< acr hitrtse >}}.
@@
 
-We are not aware of prior studies specifically investigating the effects of exposure to {{< acr edcs >}} in relation to {{< acr hitrtse >}}.
+We are not aware of prior studies specifically investigating the effects of exposure to {{< acr edcs >}} in relation to attentional function.
 The literature on non-persistent {{< acr edcs >}} and neurodevelopment in children has mostly focused on {{< acr opps >}} [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @CartierWarembourgLeManer-Idrissi:2016; @YuDuChiou:2016], phthalate metabolites [@HuangChenSu:2015; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023], and {{< acr bpa >}} [@TewarAuingerBraun:2016; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019], in relation to {{< acr adhd >}} [@BouchardBellingerWright:2010; @TewarAuingerBraun:2016; @YuDuChiou:2016; @LiZhangKuang:2018], and intelligence scales [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @HuangChenSu:2015; @CartierWarembourgLeManer-Idrissi:2016; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023].
 Few studies have looked into different classes of non-persistent {{< acr edcs >}} [@ShoaffCoullWeuve:2020 in relation with the Conners Attention Deficit Scale and the Behavior Assessment System for Children; @OhKimKannan:2023 in relation with {{< acr adhd >}} symptoms].
-Overall, and consistent with our results, these studies seem to provide further evidence of the adverse effects of several non-persistent {{< acr edcs >}} on neurodevelopment in children.
+Overall, and consistent with our results, these studies seem to provide further evidence of the adverse effects of several non-persistent {{< acr edcs >}} on neurodevelopment (too broad) in children.
 While not all these studies have investigated effect modification by sex, it seems that these adverse effects tend to be stronger in males [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @YuDuChiou:2016; @ShoaffCoullWeuve:2020], although this difference might depend on the timing of exposure [@].
@@
 There are some differences, though.
-First, these studies only focus on phthalates, either as individual metabolites or as mixture.
+First, these studies only focus on phthalates, either as individual metabolites or as mixtures.
 Second, exposure assessment in @SunLiJin:2018 and @KimLeeMoon:2018 was performed during gestation or the first 15 months of life, respectively.
@@
 
-We are also not aware of prior epidemiological studies specifically investigating the effects of elevated levels of glucocorticosteroids in relation to {{< acr hitrtse >}}, although there is evidence that under- or over-production of glucocorticosteroids interfere with the normal development of the brain [@LupienMcEwenGunnar:2009].
+We are also not aware of prior epidemiological studies specifically investigating the effects of elevated levels of glucocorticosteroids in relation to attentional function, although there is evidence that under- or over-production of glucocorticosteroids interfere with the normal development of the brain [@LupienMcEwenGunnar:2009].
 While we did find sex-specific evidence of an effect, their clinical relevance is questionable.
@@
 In conclusion, in a study of 1,297 children from 6 European birth cohorts, we observed that (i) exposure to non-persistent {{< acr edcs >}} in childhood might have short-term effects on {{< acr hitrtse >}} in childhood, (ii) exposure to non-persistent {{< acr edcs >}} in childhood might disrupt the {{< acr hpa >}} axis in childhood, and (iii) disruption of the {{< acr hpa >}} axis in childhood might have short-term, sex-specific effects on {{< acr hitrtse >}}.
-Future studies should investigate how glucocorticosteroids might mediate the adverse effects of exposure to non-persistent {{< acr edcs >}} on childhood neurodevelopment in larger populations.
+Future studies should investigate how glucocorticosteroids might mediate the adverse effects of exposure to non-persistent {{< acr edcs >}} on childhood neurodevelopment (too broad) in larger populations.
 

### 2024-03-12 output/paper/manuscript.qmd
Reviewing.

@@
 format:
+  html: default
   docx:
@@
 We used the parametric g-formula to estimate effects between {{< acr edcs >}}, glucocorticosteroids, and {{< acr hitrtse >}}, a measure of inattention from the {{< acr ant >}}, and tested for possible effect modification by sex.
-We observed a positive {{< acr mc >}} for exposure increases from the 10th to the 90th percentile for methyl-paraben (`r format_res(meffects1, "mepa", "HRT-SE")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "HRT-SE")`), oh-MiNP (`r format_res(meffects1, "ohminp", "HRT-SE")`), and MEHP (`r format_res(meffects1, "mehp", "HRT-SE")`), on {{< acr hitrtse >}}.
+We observed a positive {{< acr mc >}} for exposure increases from the 10th to the 90th percentile for methyl-paraben (`r format_res(meffects1, "mepa", "HRT-SE")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "HRT-SE")`), oh-MiNP (`r format_res(meffects1, "ohminp", "HRT-SE")`), and MEHP (`r format_res(meffects1, "mehp", "HRT-SE")`), on {{< acr hitrtse >}}, indicating lower attention.
 Several {{< acr edcs >}} were also associated with a positive {{< acr mc >}} for cortisone, cortisol, and corticosterone production.
@@
 Further, each minimal adjustment set was *augmented* with precision covariates, defined as the set of parents variable of the outcome that are not parents of the exposure.
+Common confounders were cohort, ethnicity, sex, age, height, weight, and head circumference of the child, consumption of fish, fruit, vegetables, organic food, anf fast food, maternal tobacco consumption, family financial situation and affluence scale (FAS).
+Models for estimating the effects of {{< acr edcs >}} on {{< acr hitrtse >}} were further adjusted for child breastfeeding, prenatal maternal active and passive smoking, urine creatinine, child mood and rest before assessment, child neuropsychological diagnosis, marital status, season, and fasting time before assessment.
+Models for estimating the effects of {{< acr edcs >}} on glucocorticosteroids were further adjusted for urine creatinine, season, and fasting time before assessment.
+Models for estimating the effects of glucocorticosteroids on {{< acr hitrtse >}} were further adjusted for child breastfeeding, prenatal maternal active and passive smoking, marital status, {{< acr edcs >}}, urine creatinine, child mood and rest before assessment, and child neuropsychological diagnosis.
 The adjustment sets are provided in the Supplementary Material as text files compatible with DAGitty.
@@
 Information about the lower limits of detection can be found in @HaugSakhiCequier:2018.
-Chemicals with more than 70% of observations below the {{< acr lod >}} were not considered in the present study.
+Chemicals with more than 70% of observations below the {{< acr lod >}} were excluded from the present study.
 Remaining missing values were imputed similarly using kNN.
@@
 We decided to model both the *treatment* mechanisms, for the estimation of balancing weights, and the outcomes, with traditional covariates adjustment, to try to obtain *doubly robust* effect estimates.
-Finally, we decided not to interpret our results by focusing on the estimated coefficients of the regression models, but by making use of the g-computation procedure and estimate {{< acr mcs >}}.
+Finally, we decided not to interpret our results by focusing on the estimated coefficients of possibly misspecified regression models, but by making use of the g-computation procedure and estimate {{< acr mcs >}}.
 

### 2024-03-12 R/proc_res.R

@@ -275,16 +275,21 @@ tbl_desc_vars <- function() {
     dplyr::rename_with(
       ~tibble::deframe(edcs)[.x],
       .cols = edcs$chem_id
-    ) |>
-    gtsummary::tbl_summary(
-      statistic = list(
-        gtsummary::all_continuous() ~ c(
-          "{median} ({p25}, {p75}); {N_miss} ({p_miss})"
-        )
-      ),
-      missing = "no",
-      digits = dplyr::everything() ~ 1
     )
+  tbl_edcs <- c("{median} ({p25}, {p75})", "{N_miss} ({p_miss})") |>
+    purrr::map(
+      ~ tbl_edcs |>
+        gtsummary::tbl_summary(
+          statistic = gtsummary::all_continuous() ~ .x,
+          missing = "no",
+          digits = dplyr::everything() ~ 1
+        )
+    ) |>
+    gtsummary::tbl_merge() |>
+    gtsummary::modify_spanning_header(
+      gtsummary::everything() ~ NA
+    )
+  
   tbl_edcs$table_body <- dplyr::left_join(
     tbl_edcs$table_body, edcs,
     by = c("variable" = "short_name")
@@ -1109,16 +1114,16 @@ tidy_res_meffects <- function(df, sa_var, outcome,
     #       )
     #     }
     #   )
-    # ) |>
-    # gt::tab_footnote(
-    #   footnote = "*Significant results."
-    # ) |>
-    gt::tab_footnote(
-      footnote = "Estimate and 95% CI.",
-      locations = gt::cells_column_labels(
-        columns = names_
-      )
-    ) |>
+  # ) |>
+  # gt::tab_footnote(
+  #   footnote = "*Significant results."
+  # ) |>
+  gt::tab_footnote(
+    footnote = "Estimate and 95% CI.",
+    locations = gt::cells_column_labels(
+      columns = names_
+    )
+  ) |>
     gt::opt_footnote_marks(
       marks = "letters"
     )

### 2024-03-14 output/paper/manuscript.qmd
Re-writing discussion basically.

@@
 While research in this area has mainly investigated the effects of prenatal exposure on child neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022], little is still known about childhood exposure.
-In this study, consisting of 1,297 children from 6 European birth cohorts, we observed that short-term childhood exposure to non-persistent {{< acr edcs >}} was associated with attentional function, and with total production of cortisol, cortisone, and corticosterone, although the majority of the {{< acr cis >}} included the null effect.
-Increased production of these glucocorticosteroids did not seem to affect {{< acr hitrtse >}}.
+In this study, consisting of 1,297 children from 6 European birth cohorts, we observed that short-term childhood exposure to certain non-persistent {{< acr edcs >}} was associated with attentional function (MEPA, MEHP, oh-MiNP, and oxo-MiNP), and with total production of cortisol, cortisone, and corticosterone (DEP, DMP, DMTP, BPA, ETPA, MEPA, MEHP, oh-MiNP, and oxo-MiNP).
+Increased production of these glucocorticosteroids did not seem to affect attentional function.
 Some of these effects differed for females and males, including significant differences for the effects of increased production of cortisone and corticosterone on {{< acr hitrtse >}}.
@@
 
-We are not aware of prior studies specifically investigating the effects of exposure to {{< acr edcs >}} in relation to attentional function.
-The literature on non-persistent {{< acr edcs >}} and neurodevelopment in children has mostly focused on {{< acr opps >}} [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @CartierWarembourgLeManer-Idrissi:2016; @YuDuChiou:2016], phthalate metabolites [@HuangChenSu:2015; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023], and {{< acr bpa >}} [@TewarAuingerBraun:2016; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019], in relation to {{< acr adhd >}} [@BouchardBellingerWright:2010; @TewarAuingerBraun:2016; @YuDuChiou:2016; @LiZhangKuang:2018], and intelligence scales [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @HuangChenSu:2015; @CartierWarembourgLeManer-Idrissi:2016; @HuangTsaiChen:2017; @KimHongShin:2017; @VilmandBeckBilenberg:2023].
+===
+To the best of our knowledge, no other study has investigated the effects of childhood exposure to multiple classes of non-persistent {{< acr edcs >}} in relation to attentional function.
+
+More generally, the literature on non-persistent {{< acr edcs >}} and neurodevelopment in children has mostly focused on {{< acr opps >}} [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @CartierWarembourgLeManer-Idrissi:2016; @YuDuChiou:2016], phthalate metabolites [@HuangChenSu:2015; @HuangTsaiChen:2017; @KimHongShin:2017; @LiPapandonatosCalafat:2019; @BalalianWhyattLiu:2019; @ShoaffCoullWeuve:2020; @JankowskaPolanskaHanke:2019; @VilmandBeckBilenberg:2023], and {{< acr bpa >}} [@TewarAuingerBraun:2016; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019].
+
+González-Alzaga et al.
+and Cartier et al.
+evaluated cross-sectional associations between {{< acr dap >}} metabolites and subtests of the Wechsler Intelligence Scale for Children [@wechsler1949wechsler] in European children with ages between 6 and 11 years.
+Higher levels of {{< acr dap >}} metabolites (DMP, DMTP, DMDTP, DEP, DETP, and DEDTP) were associated with lower scores of {{< acr iq >}} and verbal comprehension, especially in boys [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015], while higher levels of diethylphosphate metabolites (DEP, DETP, DEDTP) were associated with lower working memory scores [@CartierWarembourgLeManer-Idrissi:2016].
+There is also preliminary evidence of a possible association between exposure to certain {{< acr opps >}} and {{< acr adhd >}} in children [@BouchardBellingerWright:2010; @YuDuChiou:2016].
+Specifically, Bouchard et al.
+found evidence of a cross-sectional association between dimethyl alkylphosphate metabolites (DMP, DMTP, and DMDTP) and {{< acr adhd >}} in children aged 8 to 15 years from {{< acr nhanes >}}, while Yu et al.
+found a dose-response relationship between DMP and {{< acr adhd >}} in Taiwanese children aged 4 to 15 years.
+
+Preliminary evidence is also available for several phthalate metabolites in relation to cognitive development in childhood.
+Higher levels of di(2-ethylhexyl) phthalate metabolites (including MEHP, MEHHP, and MEOHP) were associated with lower intelligence scores in children aged 2 to 12 years [@HuangChenSu:2015], lower scores of {{< acr iq >}} and verbal {{< acr iq >}}, more omission errors (a measure of inattention), and higher scores of response time variability (a measure of sustained attention) in 6-year old Korean children [@KimHongShin:2017], and poorer fine motor skills in preadolescent boys [@BalalianWhyattLiu:2019].
+
 Few studies have looked into different classes of non-persistent {{< acr edcs >}} [@ShoaffCoullWeuve:2020 in relation with the Conners Attention Deficit Scale and the Behavior Assessment System for Children; @OhKimKannan:2023 in relation with {{< acr adhd >}} symptoms].
@@
 While not all these studies have investigated effect modification by sex, it seems that these adverse effects tend to be stronger in males [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @YuDuChiou:2016; @ShoaffCoullWeuve:2020], although this difference might depend on the timing of exposure [@].
+===
 

### 2024-03-14 output/paper/manuscript.qmd
Re-writing discussion basically.

@@
 Preliminary evidence is also available for several phthalate metabolites in relation to cognitive development in childhood.
-Higher levels of di(2-ethylhexyl) phthalate metabolites (including MEHP, MEHHP, and MEOHP) were associated with lower intelligence scores in children aged 2 to 12 years [@HuangChenSu:2015], lower scores of {{< acr iq >}} and verbal {{< acr iq >}}, more omission errors (a measure of inattention), and higher scores of response time variability (a measure of sustained attention) in 6-year old Korean children [@KimHongShin:2017], and poorer fine motor skills in preadolescent boys [@BalalianWhyattLiu:2019].
+Higher levels of di(2-ethylhexyl) phthalate metabolites (including MEHP, MEHHP, and MEOHP) were associated with lower intelligence scores in children aged 2 to 12 years [@HuangChenSu:2015], lower scores of {{< acr iq >}} and verbal intelligence, more omission errors (a measure of inattention), and higher scores of response time variability (a measure of sustained attention) in 6-year old Korean children [@KimHongShin:2017], poorer fine motor skills in preadolescent boys [@BalalianWhyattLiu:2019], and lower intelligence scores in 7-year old children [@VilmandBeckBilenberg:2023].
+Further associations were found for MEOHP with lower scores of {{< acr iq >}} [@HuangChenSu:2015] and verbal intelligence in Taiwanese children aged 6 to 12 years [@HuangTsaiChen:2017], and for dibutyl phthalate metabolites (MnBP and MiBP) with impaired verbal intelligence [@HuangTsaiChen:2017].
+There is further preliminary evidence that associations between certain phthalate metabolites and cognitive abilities vary by timing of exposure assessment [@LiPapandonatosCalafat:2019].
 
-Few studies have looked into different classes of non-persistent {{< acr edcs >}} [@ShoaffCoullWeuve:2020 in relation with the Conners Attention Deficit Scale and the Behavior Assessment System for Children; @OhKimKannan:2023 in relation with {{< acr adhd >}} symptoms].
+BPA
+
+Few studies have looked into different classes of non-persistent {{< acr edcs >}}.
+Shoaff et al.
+investigated cross-sectional associations between multiple {{< acr edcs >}} and {{< acr adhd >}}-related behaviors in 15-year old adolescents, finding a higher risk of {{< acr adhd >}}-related behavior problems with higher levels of antiandrogenic phthalate metabolites (MEHP, MEHHP, MEOHP, MECPP, MnBP, MiBP, MBzP, monohydroxyisobutyl phthalate (MHiBP), monocarboxyoctyl phthalate (MCOP), monoisononyl phthalate (MNP), and monohydroxybutyl phthalate (MHBP)), especially in boys [@ShoaffCoullWeuve:2020].
+
+Oh et al.
+[@OhKimKannan:2023]
+
 Overall, and consistent with our results, these studies seem to provide further evidence of the adverse effects of several non-persistent {{< acr edcs >}} on neurodevelopment (too broad) in children.

### 2024-03-14 output/paper/manuscript.qmd
Re-writing discussion basically.

@@
 
-===
 To the best of our knowledge, no other study has investigated the effects of childhood exposure to multiple classes of non-persistent {{< acr edcs >}} in relation to attentional function.
-
 More generally, the literature on non-persistent {{< acr edcs >}} and neurodevelopment in children has mostly focused on {{< acr opps >}} [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @CartierWarembourgLeManer-Idrissi:2016; @YuDuChiou:2016], phthalate metabolites [@HuangChenSu:2015; @HuangTsaiChen:2017; @KimHongShin:2017; @LiPapandonatosCalafat:2019; @BalalianWhyattLiu:2019; @ShoaffCoullWeuve:2020; @JankowskaPolanskaHanke:2019; @VilmandBeckBilenberg:2023], and {{< acr bpa >}} [@TewarAuingerBraun:2016; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019].
-
 González-Alzaga et al.
@@
 found a dose-response relationship between DMP and {{< acr adhd >}} in Taiwanese children aged 4 to 15 years.
-
 Preliminary evidence is also available for several phthalate metabolites in relation to cognitive development in childhood.
@@
 There is further preliminary evidence that associations between certain phthalate metabolites and cognitive abilities vary by timing of exposure assessment [@LiPapandonatosCalafat:2019].
+Among phenols, some studies provide preliminary evidence of an association between BPA and {{< acr adhd >}} in children aged 8 to 15 years [@TewarAuingerBraun:2016] and in a case-control study of children aged 6 to 12 years [@LiZhangKuang:2018], especially in boys.
+Except for working memory, there does not seem to be evidence of an association between BPA and cognitive abilities in Spanish boys aged 9 to 11 years [@Rodriguez-CarrilloMustielesPerez-Lobato:2019].
+Few studies have looked into different classes of non-persistent {{< acr edcs >}}.
+Shoaff et al., for instance, investigated cross-sectional associations between multiple {{< acr edcs >}} and {{< acr adhd >}}-related behaviors in 15-year old adolescents, finding a higher risk of {{< acr adhd >}}-related behavior problems with higher levels of antiandrogenic phthalate metabolites (MEHP, MEHHP, MEOHP, MECPP, MnBP, MiBP, MBzP, monohydroxyisobutyl phthalate (MHiBP), monocarboxyoctyl phthalate (MCOP), monoisononyl phthalate (MNP), and monohydroxybutyl phthalate (MHBP)), especially in boys [@ShoaffCoullWeuve:2020].
 
-BPA
-
-Few studies have looked into different classes of non-persistent {{< acr edcs >}}.
-Shoaff et al.
-investigated cross-sectional associations between multiple {{< acr edcs >}} and {{< acr adhd >}}-related behaviors in 15-year old adolescents, finding a higher risk of {{< acr adhd >}}-related behavior problems with higher levels of antiandrogenic phthalate metabolites (MEHP, MEHHP, MEOHP, MECPP, MnBP, MiBP, MBzP, monohydroxyisobutyl phthalate (MHiBP), monocarboxyoctyl phthalate (MCOP), monoisononyl phthalate (MNP), and monohydroxybutyl phthalate (MHBP)), especially in boys [@ShoaffCoullWeuve:2020].
-
-Oh et al.
-[@OhKimKannan:2023]
+TBD
 
@@
 While not all these studies have investigated effect modification by sex, it seems that these adverse effects tend to be stronger in males [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @YuDuChiou:2016; @ShoaffCoullWeuve:2020], although this difference might depend on the timing of exposure [@].
-===
+
+TBD
 

### 2024-03-14 output/paper/manuscript.qmd
Re-writing discussion basically.

@@
 
-TBD
+TBD S
 
 Overall, and consistent with our results, these studies seem to provide further evidence of the adverse effects of several non-persistent {{< acr edcs >}} on neurodevelopment (too broad) in children.
-While not all these studies have investigated effect modification by sex, it seems that these adverse effects tend to be stronger in males [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @YuDuChiou:2016; @ShoaffCoullWeuve:2020], although this difference might depend on the timing of exposure [@].
 
-TBD
+TBD E
 
-Our results are consistent with prior epidemiological research that associated exposure to certain non-persistent {{< acr edcs >}} with higher levels of cortisol [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023].
+TBD S
+
+Our results are consistent with prior epidemiological research that associated exposure to certain non-persistent {{< acr edcs >}} with higher levels of glucocorticoids [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023].
+Repeated measures up to 15 months of age of the phthalate metabolites MEHHP, MEOHP, MiBP, and MnBP showed positive associations with free cortisol in Korean children, with no effect modification by sex [@KimLeeMoon:2018].
+In a cohort of Chinese pregnant women, phthalate metabolites were measured at 14, 24, and 36 weeks of gestation, and the glucocorticoids cortisol and cortisone were measured in cord blood.
+Third-trimester levels of MEHP were positively associated with cortisol, while MECPP and MEOHP were negatively associated with cortisone [@SunLiJin:2018].
+Time- and chemical-dependent sex differences were also found: during the third trimester, MEHHP and MEOHP were positively associated with cortisol in females, while negatively associated in males [@SunLiJin:2018].
+In a longitudinal study, a mixture of several phthalate metabolites, driven by MEP, MiBP, and MBzP, measured in childhood, showed a positive association with hair cortisol measured at 12 years [@SearsLiuLanphear:2023].
+
 There are some differences, though.
@@
 There is also *in silico* evidence suggesting that {{< acr bpa >}}, a phenol, and Triazophos (TAP), a organophosphorus insecticide, can bind to the human glucocorticoid receptor [@PrasanthDivyaSadasivan:2010; @YangLiRen:2019].
+
+TBD E
 

### 2024-03-15 output/paper/manuscript.qmd
Re-writing discussion basically.

@@
 
-Our results are consistent with prior epidemiological research that associated exposure to certain non-persistent {{< acr edcs >}} with higher levels of glucocorticoids [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023].
+Prior epidemiological research provides preliminary evidence for an association between certain non-persistent {{< acr edcs >}} with higher levels of glucocorticoids [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023].
 Repeated measures up to 15 months of age of the phthalate metabolites MEHHP, MEOHP, MiBP, and MnBP showed positive associations with free cortisol in Korean children, with no effect modification by sex [@KimLeeMoon:2018].
@@
 Time- and chemical-dependent sex differences were also found: during the third trimester, MEHHP and MEOHP were positively associated with cortisol in females, while negatively associated in males [@SunLiJin:2018].
-In a longitudinal study, a mixture of several phthalate metabolites, driven by MEP, MiBP, and MBzP, measured in childhood, showed a positive association with hair cortisol measured at 12 years [@SearsLiuLanphear:2023].
+In a longitudinal study, a mixture of several phthalate metabolites, driven by MEP, MiBP, and MBzP, measured in childhood, showed a positive association with hair cortisol measured at 12 years of age [@SearsLiuLanphear:2023].
 
+Our results...
 There are some differences, though.
@@
 Finally, the glucocorticosteroids were measured in cord blood [@SunLiJin:2018] and hair [@SearsLiuLanphear:2023].
+
 Contrary to these studies, we did find effect modification by sex.
+
 We are not aware of other epidemiological studies investigating phthalates metabolites, phenols, and {{< acr opps >}}, in relation to urinary glucocorticosteroids in childhood.

### 2024-03-18 output/paper/_figures.qmd
Reviewing manuscript. Finished Discussion.

@@
 ```
-**Marginal contrasts for the effect of a decrease from the 90th to the 10th percentile of the {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
+**Marginal contrasts on the logarithmic scale for the effect of a increase from the 10th to the 90th percentile of the {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
 Circles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}. The size of the circles represents the *S* value of the effect estimate [@RafiGreenland:2020].
@@
 ```
-**Marginal contrasts for the effect of a decrease from the 90th to the 10th percentile of the {{< acr edcs first_use=true >}} on the glucocorticosteroids expressed in ng/ml (HELIX subcohort; 2013-2016).**
+**Marginal contrasts on the logarithmic scale for the effect of a increase from the 10th to the 90th percentile of the {{< acr edcs first_use=true >}} on the glucocorticosteroids expressed in ng/ml (HELIX subcohort; 2013-2016).**
 Circles, triangles, and squares indicate effect estimates.
@@
 ```
-**Marginal contrasts for the effect of a decrease from the 90th to the 10th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
+**Marginal contrasts on the logarithmic scale for the effect of a increase from the 10th to the 90th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
 Circles indicate effect estimates.

### 2024-03-18 output/paper/_si.qmd
Reviewing manuscript. Finished Discussion.

@@
 ```
-**Marginal contrasts for effect modification by sex of a decrease from the 90th to the 10th percentile of the {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
+**Marginal contrasts on the logarithmic scale for effect modification by sex of a increase from the 10th to the 90th percentile of the {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
 Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}. The size of the circles represents the *S* value of the effect estimate [@RafiGreenland:2020].
@@
 ```
-**Marginal contrasts for effect modification by sex of a decrease from the 90th to the 10th percentile of the {{< acr edcs first_use=true >}} on the glucocorticosteroids expressed in ng/ml (HELIX subcohort; 2013-2016).**
+**Marginal contrasts on the logarithmic scale for effect modification by sex of a increase from the 10th to the 90th percentile of the {{< acr edcs first_use=true >}} on the glucocorticosteroids expressed in ng/ml (HELIX subcohort; 2013-2016).**
 Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}. The size of the circles represents the *S* value of the effect estimate [@RafiGreenland:2020].
@@
 ```
-**Marginal contrasts for effect modification by sex of a decrease from the 90th to the 10th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
+**Marginal contrasts on the logarithmic scale for effect modification by sex of a increase from the 10th to the 90th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
 Circles and triangles indicate effect estimates.

### 2024-03-18 output/paper/_tables.qmd
Reviewing manuscript. Finished Discussion.

--- 2024-02-20 output/paper/_tables.qmd
@@
 ```
-**Pairwise differences between sex-specific marginal contrasts for the effect of a decrease from the 90th to the 10th percentile of {{< acr edcs >}} on {{< acr hitrtse >}}, expressed in {{< acr mss >}}, and the glucocorticosteroids, expressed in ng/ml (HELIX subcohort; 2013-2016).**
+**Pairwise differences between marginal contrasts on the logarithmic scale of males and females, for the effect of a increase from the 10th to the 90th percentile of {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}}, expressed in {{< acr mss >}}, and on the glucocorticosteroids, expressed in ng/ml (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Pairwise differences between sex-specific marginal contrasts for the effect of a decrease from the 90th to the 10th percentile of the glucocorticosteroids on {{< acr hitrtse >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
+**Pairwise differences between marginal contrasts on the logarithmic scale of males and females, for the effect of a increase from the 10th to the 90th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
 :::

### 2024-03-18 output/paper/manuscript.qmd
Reviewing manuscript. Finished Discussion.

@@
 
-### Attentionall function {#sec-neurodevelopment}
+### Attentional function {#sec-neurodevelopment}
 
@@
 
-@Fig-marginal-1 presents the forest plot for the {{< acr mcs >}} of the non-persistent {{< acr edcs >}} on {{< acr hitrtse >}}.
-For most {{< acr edcs >}}, a cohort-specific decrease in the levels of the exposures from the 90th to the 10th percentiles was associated with a positive {{< acr mc >}}, indicating an increase in the values of {{< acr hitrtse >}} and thus lower attention.
+@Fig-marginal-1 presents the forest plot for the {{< acr mcs >}} on the logarithmic scale of the non-persistent {{< acr edcs >}} on {{< acr hitrtse >}}.
+For most {{< acr edcs >}}, a cohort-specific increase in the levels of the exposures from the 10th to the 90th percentiles was associated with a positive {{< acr mc >}}, indicating an increase in the values of {{< acr hitrtse >}} and thus lower attention.
 Most of the {{< acr cis first_use=true >}} included the null effect, though.
@@
 
-@Fig-marginal-2 presents the forest plot for the {{< acr mcs >}} of the non-persistent {{< acr edcs >}} on total cortisone, cortisol, and corticosterone production.
-For most {{< acr edcs >}}, a cohort-specific decrease in the levels of the exposures from the 90th to the 10th percentiles was associated with a positive {{< acr mc >}}, indicating an increase in the total production of these metabolites.
+@Fig-marginal-2 presents the forest plot for the {{< acr mcs >}} on the logarithmic scale of the non-persistent {{< acr edcs >}} on total cortisone, cortisol, and corticosterone production.
+For most {{< acr edcs >}}, a cohort-specific increase in the levels of the exposures from the 10th to the 90th percentiles was associated with a positive {{< acr mc >}}, indicating an increase in the total production of these metabolites.
 Exceptions were BUPA, which was associated with negative {{< acr mcs >}} for all three outcomes, and MiBP, which was associated with a negative {{< acr mc >}} for total cortisone production only.
@@
 
-@Fig-marginal-3 presents the forest plot for the {{< acr mcs >}} of the glucocorticosteroids on {{< acr hitrtse >}}.
+@Fig-marginal-3 presents the forest plot for the {{< acr mcs >}} on the logarithmic scale of the glucocorticosteroids on {{< acr hitrtse >}}.
 All {{< acr mcs >}} included the null, with no clear indication of directionality of the effect.
@@
 
-@Tbl-hypothesis-1and2 presents the results of the difference between estimates of the {{< acr mcs >}} for females and males, for the {{< acr edcs >}} on the glucocorticosteroids and {{< acr hitrtse >}}.
+@Tbl-hypothesis-1and2 presents the results of the difference between estimates of the {{< acr mcs >}} on the logarithmic scale for females and males, for the {{< acr edcs >}} on the glucocorticosteroids and {{< acr hitrtse >}}.
 For {{< acr hitrtse >}}, significant differences were present for the phenol OXBE (`r format_res(heffects1, "oxbe", "HRT-SE")`) and the phthalate metabolites MEP (`r format_res(heffects1, "mep", "HRT-SE")`) and MbZP (`r format_res(heffects1, "mbzp", "HRT-SE")`).
@@
 
-@Tbl-hypothesis-3 presents the results of the difference between estimates of the {{< acr mcs >}} for females and males, for the glucocorticosteroids on {{< acr hitrtse >}}.
+@Tbl-hypothesis-3 presents the results of the difference between estimates of the {{< acr mcs >}} on the logarithmic scale for females and males, for the glucocorticosteroids on {{< acr hitrtse >}}.
 Significant differences were present for cortisone production (`r format_res(heffects3, "cortisone prod.", "HRT-SE")`) and corticosterone production (`r format_res(heffects3, "corticost.
@@
 
-TBD S
-
-Overall, and consistent with our results, these studies seem to provide further evidence of the adverse effects of several non-persistent {{< acr edcs >}} on neurodevelopment (too broad) in children.
-
-TBD E
-
-TBD S
-
+We are not aware of other epidemiological studies investigating childhood exposure to phthalates metabolites, phenols, and {{< acr opps >}}, in relation to urinary glucocorticosteroid levels in childhood.
 Prior epidemiological research provides preliminary evidence for an association between certain non-persistent {{< acr edcs >}} with higher levels of glucocorticoids [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023].
@@
 In a longitudinal study, a mixture of several phthalate metabolites, driven by MEP, MiBP, and MBzP, measured in childhood, showed a positive association with hair cortisol measured at 12 years of age [@SearsLiuLanphear:2023].
+While in the present study we did find positive {{< acr mcs >}} between some phthalate metabolites (MEHP, oh-MiNP, and oxo-MiNP) and the glucocorticosteroids, there are important differences with the previous studies.
+First, exposure assessment was performed during gestation [@SunLiJin:2018] or the first 15 months of life [@KimLeeMoon:2018], not during childhood.
+Second, the glucocorticosteroids were measured in other matrices, specifically in cord blood [@SunLiJin:2018] or hair [@SearsLiuLanphear:2023].
+Finally, @SearsLiuLanphear:2023 investigated mixture effects.
+Contrary to these studies [@KimLeeMoon:2018; @SearsLiuLanphear:2023], we did find effect modification by sex.
 
-Our results...
-There are some differences, though.
-First, these studies only focus on phthalates, either as individual metabolites or as mixtures.
-Second, exposure assessment in @SunLiJin:2018 and @KimLeeMoon:2018 was performed during gestation or the first 15 months of life, respectively.
-Finally, the glucocorticosteroids were measured in cord blood [@SunLiJin:2018] and hair [@SearsLiuLanphear:2023].
-
-Contrary to these studies, we did find effect modification by sex.
-
-We are not aware of other epidemiological studies investigating phthalates metabolites, phenols, and {{< acr opps >}}, in relation to urinary glucocorticosteroids in childhood.
-Nonetheless, previous toxicological studies provide evidence for the inhibition by phthalates of human 11$\beta$-hydroxysteroid dehydrogenase 2 (11$\beta$-HSD2) activity, responsible for the conversion of active cortisol into inactive cortisone [@ZhaoChuHuang:2010; @MaLianDong:2011].
+Adding to these epidemiological studies, previous toxicological research provide evidence for the inhibition by phthalates of human 11$\beta$-hydroxysteroid dehydrogenase 2 (11$\beta$-HSD2) activity, responsible for the conversion of active cortisol into inactive cortisone [@ZhaoChuHuang:2010; @MaLianDong:2011].
 There is also *in silico* evidence suggesting that {{< acr bpa >}}, a phenol, and Triazophos (TAP), a organophosphorus insecticide, can bind to the human glucocorticoid receptor [@PrasanthDivyaSadasivan:2010; @YangLiRen:2019].
-
-TBD E
 

### 2024-03-18 output/paper/manuscript.qmd
Fixing minor details manuscript.

@@
 
-We tested for possible effect-modification by sex.
+We further estimated separate {{< acr mcs >}} for possible effect-modification by sex.
 To do so, balancing weights were estimated separately for each level of the sex variable, and an interaction term between the exposure and sex was included in the outcome model.
 Similarly, the {{< acr mcs >}} were aggregated separately for each level of sex.
-We further tested for significance of the difference between the {{< acr mcs >}} of females and males.
 
@@
 and Cartier et al.
-evaluated cross-sectional associations between {{< acr dap >}} metabolites and subtests of the Wechsler Intelligence Scale for Children [@wechsler1949wechsler] in European children with ages between 6 and 11 years.
+evaluated cross-sectional associations between {{< acr dap >}} metabolites and subtests of the Wechsler Intelligence Scale for Children [@WechslerKodama:2003] in European children with ages between 6 and 11 years.
 Higher levels of {{< acr dap >}} metabolites (DMP, DMTP, DMDTP, DEP, DETP, and DEDTP) were associated with lower scores of {{< acr iq >}} and verbal comprehension, especially in boys [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015], while higher levels of diethylphosphate metabolites (DEP, DETP, DEDTP) were associated with lower working memory scores [@CartierWarembourgLeManer-Idrissi:2016].

### 2024-03-18 R/proc_res.R

@@ -101,7 +101,7 @@ tbl_desc_pop <- function(num_digits_est, num_digits_sig, by) {
   
   # Load data request (all variables)
   params_dat <- params(is_hpc = Sys.getenv("is_hpc"))
-  dat_request <- load_dat_request()
+  dat_request <- load_dat_request(load_all = TRUE)
   dat_request$dat <- dat_request$dat |>
     tidylog::mutate(
       cohort = dplyr::case_when(

### 2024-03-18 output/paper/manuscript.qmd
Compiled manuscript.

@@
 format:
-  html: default
+  html:
+    embed-resources: true
   docx:

### 2024-03-19 output/paper/manuscript.qmd
Tried other output formats Quarto but shit.

--- 2024-03-18 output/paper/manuscript.qmd
@@
 
+embed-resources: true
 format:
-  html: default
   docx:
-    # reference-doc: reference_document.docx
     page-width: 8.3

### 2024-03-30 output/paper/manuscript.qmd
Comments from MV v2.

@@
 
-The impact of exposure to {{< acr edcs >}} on human health has attracted considerable research interest.
-While research in this area has mainly investigated the effects of prenatal exposure on child neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022], little is still known about childhood exposure.
 In this study, consisting of 1,297 children from 6 European birth cohorts, we observed that short-term childhood exposure to certain non-persistent {{< acr edcs >}} was associated with attentional function (MEPA, MEHP, oh-MiNP, and oxo-MiNP), and with total production of cortisol, cortisone, and corticosterone (DEP, DMP, DMTP, BPA, ETPA, MEPA, MEHP, oh-MiNP, and oxo-MiNP).
 Increased production of these glucocorticosteroids did not seem to affect attentional function.
-Some of these effects differed for females and males, including significant differences for the effects of increased production of cortisone and corticosterone on {{< acr hitrtse >}}.
-Specifically, an increased production of these glucocorticosteroids was associated with lower values of {{< acr hitrtse >}} for females, and higher values for males.
-Taken together, these results suggest that these non-persistent {{< acr edcs >}} might be responsible for perturbations of the {{< acr hpa >}} axis' homeostasis, and that higher levels of these glucocorticosteroids might interfere with different functions of attention in a sex-specific manner.
+Some of these associations differed for females and males, including significant differences in associations of increased production of cortisone and corticosterone with {{< acr hitrtse >}}.
+Taken together, these results suggest that exposure to non-persistent {{< acr edcs >}} might be responsible for perturbations of the {{< acr hpa >}} axis' homeostasis, and that higher levels of glucocorticosteroids might interfere with different functions of attention in a sex-specific manner.
 
 To the best of our knowledge, no other study has investigated the effects of childhood exposure to multiple classes of non-persistent {{< acr edcs >}} in relation to attentional function.
-More generally, the literature on non-persistent {{< acr edcs >}} and neurodevelopment in children has mostly focused on {{< acr opps >}} [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @CartierWarembourgLeManer-Idrissi:2016; @YuDuChiou:2016], phthalate metabolites [@HuangChenSu:2015; @HuangTsaiChen:2017; @KimHongShin:2017; @LiPapandonatosCalafat:2019; @BalalianWhyattLiu:2019; @ShoaffCoullWeuve:2020; @JankowskaPolanskaHanke:2019; @VilmandBeckBilenberg:2023], and {{< acr bpa >}} [@TewarAuingerBraun:2016; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019].
+More generally, the literature on childhood exposure to non-persistent {{< acr edcs >}} and other neurodevelopment outcomes in children has mostly focused on {{< acr opps >}} [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @CartierWarembourgLeManer-Idrissi:2016; @YuDuChiou:2016], phthalate metabolites [@HuangChenSu:2015; @HuangTsaiChen:2017; @KimHongShin:2017; @LiPapandonatosCalafat:2019; @BalalianWhyattLiu:2019; @ShoaffCoullWeuve:2020; @JankowskaPolanskaHanke:2019; @VilmandBeckBilenberg:2023], and {{< acr bpa >}} [@TewarAuingerBraun:2016; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019].
 González-Alzaga et al.
@@
 found a dose-response relationship between DMP and {{< acr adhd >}} in Taiwanese children aged 4 to 15 years.
+
 Preliminary evidence is also available for several phthalate metabolites in relation to cognitive development in childhood.
@@
 There is further preliminary evidence that associations between certain phthalate metabolites and cognitive abilities vary by timing of exposure assessment [@LiPapandonatosCalafat:2019].
+Finally, few studies have looked into different classes of non-persistent {{< acr edcs >}}.
+Shoaff et al., for instance, investigated cross-sectional associations between multiple {{< acr edcs >}} and {{< acr adhd >}}-related behaviors in 15-year old adolescents, finding a higher risk of {{< acr adhd >}}-related behavior problems with higher levels of antiandrogenic phthalate metabolites, especially in boys [@ShoaffCoullWeuve:2020].
+Our findings, indicating that short-term childhood exposure to certain phthalate metabolites (MEHP, oh-MiNP, and oxo-MiNP) was associated with attentional function, adds to this growing evidence base suggesting that childhood phthalate exposure may impact child neurodevelopment.
+
 Among phenols, some studies provide preliminary evidence of an association between BPA and {{< acr adhd >}} in children aged 8 to 15 years [@TewarAuingerBraun:2016] and in a case-control study of children aged 6 to 12 years [@LiZhangKuang:2018], especially in boys.
-Except for working memory, there does not seem to be evidence of an association between BPA and cognitive abilities in Spanish boys aged 9 to 11 years [@Rodriguez-CarrilloMustielesPerez-Lobato:2019].
-Few studies have looked into different classes of non-persistent {{< acr edcs >}}.
-Shoaff et al., for instance, investigated cross-sectional associations between multiple {{< acr edcs >}} and {{< acr adhd >}}-related behaviors in 15-year old adolescents, finding a higher risk of {{< acr adhd >}}-related behavior problems with higher levels of antiandrogenic phthalate metabolites (MEHP, MEHHP, MEOHP, MECPP, MnBP, MiBP, MBzP, monohydroxyisobutyl phthalate (MHiBP), monocarboxyoctyl phthalate (MCOP), monoisononyl phthalate (MNP), and monohydroxybutyl phthalate (MHBP)), especially in boys [@ShoaffCoullWeuve:2020].
+Except for working memory, there was no evidence of an association between BPA and cognitive abilities in Spanish boys aged 9 to 11 years [@Rodriguez-CarrilloMustielesPerez-Lobato:2019].
+We did not observe an association between BPA and attention function in the present study, but this study is the first to suggest that childhood paraben (MEPA) exposure may be associated with attention function.
 
 We are not aware of other epidemiological studies investigating childhood exposure to phthalates metabolites, phenols, and {{< acr opps >}}, in relation to urinary glucocorticosteroid levels in childhood.
-Prior epidemiological research provides preliminary evidence for an association between certain non-persistent {{< acr edcs >}} with higher levels of glucocorticoids [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023].
-Repeated measures up to 15 months of age of the phthalate metabolites MEHHP, MEOHP, MiBP, and MnBP showed positive associations with free cortisol in Korean children, with no effect modification by sex [@KimLeeMoon:2018].
+However, prior epidemiological research provides preliminary evidence for an association between certain non-persistent {{< acr edcs >}} measured at other time points with higher levels of glucocorticoids measured in other biological matrices [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023].
+Repeated measures up to 15 months of age of the phthalate metabolites MEHHP, MEOHP, MiBP, and MnBP showed positive associations with free cortisol in urine in Korean children, with no effect modification by sex [@KimLeeMoon:2018].
 In a cohort of Chinese pregnant women, phthalate metabolites were measured at 14, 24, and 36 weeks of gestation, and the glucocorticoids cortisol and cortisone were measured in cord blood.
@@
 In a longitudinal study, a mixture of several phthalate metabolites, driven by MEP, MiBP, and MBzP, measured in childhood, showed a positive association with hair cortisol measured at 12 years of age [@SearsLiuLanphear:2023].
-While in the present study we did find positive {{< acr mcs >}} between some phthalate metabolites (MEHP, oh-MiNP, and oxo-MiNP) and the glucocorticosteroids, there are important differences with the previous studies.
-First, exposure assessment was performed during gestation [@SunLiJin:2018] or the first 15 months of life [@KimLeeMoon:2018], not during childhood.
-Second, the glucocorticosteroids were measured in other matrices, specifically in cord blood [@SunLiJin:2018] or hair [@SearsLiuLanphear:2023].
-Finally, @SearsLiuLanphear:2023 investigated mixture effects.
-Contrary to these studies [@KimLeeMoon:2018; @SearsLiuLanphear:2023], we did find effect modification by sex.
+Our findings also indicate associations between certain phthalate metabolites (MEHP, oh-MiNP, and oxo-MiNP) and glucocorticosteroids, but differences in the exposure assessment time points and in the biological matrices used for glucocorticosteroids determinations make a direct comparison difficult.
+Further, we did find effect modification by sex.
 

### 2024-04-05 output/paper/manuscript.qmd
Finishing review based on MV v2.

@@
 Increased production of these glucocorticosteroids did not seem to affect attentional function.
-Some of these associations differed for females and males, including significant differences in associations of increased production of cortisone and corticosterone with {{< acr hitrtse >}}.
-Taken together, these results suggest that exposure to non-persistent {{< acr edcs >}} might be responsible for perturbations of the {{< acr hpa >}} axis' homeostasis, and that higher levels of glucocorticosteroids might interfere with different functions of attention in a sex-specific manner.
+Some of these associations differed for females and males.
 
@@
 Further associations were found for MEOHP with lower scores of {{< acr iq >}} [@HuangChenSu:2015] and verbal intelligence in Taiwanese children aged 6 to 12 years [@HuangTsaiChen:2017], and for dibutyl phthalate metabolites (MnBP and MiBP) with impaired verbal intelligence [@HuangTsaiChen:2017].
-There is further preliminary evidence that associations between certain phthalate metabolites and cognitive abilities vary by timing of exposure assessment [@LiPapandonatosCalafat:2019].
-Finally, few studies have looked into different classes of non-persistent {{< acr edcs >}}.
-Shoaff et al., for instance, investigated cross-sectional associations between multiple {{< acr edcs >}} and {{< acr adhd >}}-related behaviors in 15-year old adolescents, finding a higher risk of {{< acr adhd >}}-related behavior problems with higher levels of antiandrogenic phthalate metabolites, especially in boys [@ShoaffCoullWeuve:2020].
+Few studies have looked into different classes of non-persistent {{< acr edcs >}}.
+Shoaff et al.
+investigated cross-sectional associations between multiple {{< acr edcs >}} and {{< acr adhd >}}-related behaviors in 15-year old adolescents, finding a higher risk of {{< acr adhd >}}-related behavior problems with higher levels of antiandrogenic phthalate metabolites, especially in boys [@ShoaffCoullWeuve:2020].
 Our findings, indicating that short-term childhood exposure to certain phthalate metabolites (MEHP, oh-MiNP, and oxo-MiNP) was associated with attentional function, adds to this growing evidence base suggesting that childhood phthalate exposure may impact child neurodevelopment.
@@
 Except for working memory, there was no evidence of an association between BPA and cognitive abilities in Spanish boys aged 9 to 11 years [@Rodriguez-CarrilloMustielesPerez-Lobato:2019].
-We did not observe an association between BPA and attention function in the present study, but this study is the first to suggest that childhood paraben (MEPA) exposure may be associated with attention function.
+We did not observe an association between BPA and attention function in the present study, but this study is the first to suggest that childhood paraben (MEPA) exposure may be associated with attentional function.
 
@@
 However, prior epidemiological research provides preliminary evidence for an association between certain non-persistent {{< acr edcs >}} measured at other time points with higher levels of glucocorticoids measured in other biological matrices [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023].
-Repeated measures up to 15 months of age of the phthalate metabolites MEHHP, MEOHP, MiBP, and MnBP showed positive associations with free cortisol in urine in Korean children, with no effect modification by sex [@KimLeeMoon:2018].
+Repeated measures up to 15 months of age of the phthalate metabolites MEHHP, MEOHP, MiBP, and MnBP showed positive associations with free cortisol in urine in Korean children [@KimLeeMoon:2018].
 In a cohort of Chinese pregnant women, phthalate metabolites were measured at 14, 24, and 36 weeks of gestation, and the glucocorticoids cortisol and cortisone were measured in cord blood.
@@
 Our findings also indicate associations between certain phthalate metabolites (MEHP, oh-MiNP, and oxo-MiNP) and glucocorticosteroids, but differences in the exposure assessment time points and in the biological matrices used for glucocorticosteroids determinations make a direct comparison difficult.
-Further, we did find effect modification by sex.
 
@@
 We are also not aware of prior epidemiological studies specifically investigating the effects of elevated levels of glucocorticosteroids in relation to attentional function, although there is evidence that under- or over-production of glucocorticosteroids interfere with the normal development of the brain [@LupienMcEwenGunnar:2009].
-While we did find sex-specific evidence of an effect, their clinical relevance is questionable.
 
-Our findings should be interpreted in light of the following limitations and strengths.
+Our findings should be interpreted in light of the following strengths and limitations.
+Strengths include its relative large sample size and its inclusion of multiple classes of non-persistent {{< acr edcs >}}.
+Further, this study used pooled urine samples for chemical assessment to obtain more representative long-term exposures, since it is known that these specific {{< acr edcs >}} have very short half-lives [@PerrierGiorgis-AllemandSlama:2016; @CasasBasaganaSakhi:2018].
+We decided to model both the *treatment* mechanisms, for the estimation of balancing weights, and the outcomes, with traditional covariates adjustment, to try to obtain *doubly robust* effect estimates.
+Finally, we decided not to interpret our results by focusing on the estimated coefficients of possibly misspecified regression models, but by making use of the g-computation procedure.
+
 Limitations include the cross-sectional design of the present study.
@@
 Although we included a wide range of confounders there is the possibility, as with other observational studies, of residual confounding, which might lead to a bias away from the null.
-Some of the confounders indicated in the adjustment sets had to be removed due to large fractions of missing values.
 There is further the possibility of misspecification of the outcome model, although we included a spline of the exposure to relax some of the linearity assumptions.
-The use of more data-adaptive learners was excluded due to the relatively small sample size.
-We finally acknowledge the possibility that some of chemicals might not act independently (mixture effect).
-Further research is thus warranted.
 
-Strengths of the present study include the use of pooled urine samples for chemical assessment to obtain more representative long-term exposures, since it is known that these specific {{< acr edcs >}} have very short half-lives [@PerrierGiorgis-AllemandSlama:2016; @CasasBasaganaSakhi:2018].
-We decided to model both the *treatment* mechanisms, for the estimation of balancing weights, and the outcomes, with traditional covariates adjustment, to try to obtain *doubly robust* effect estimates.
-Finally, we decided not to interpret our results by focusing on the estimated coefficients of possibly misspecified regression models, but by making use of the g-computation procedure and estimate {{< acr mcs >}}.
-
-In conclusion, in a study of 1,297 children from 6 European birth cohorts, we observed that (i) exposure to non-persistent {{< acr edcs >}} in childhood might have short-term effects on {{< acr hitrtse >}} in childhood, (ii) exposure to non-persistent {{< acr edcs >}} in childhood might disrupt the {{< acr hpa >}} axis in childhood, and (iii) disruption of the {{< acr hpa >}} axis in childhood might have short-term, sex-specific effects on {{< acr hitrtse >}}.
-Future studies should investigate how glucocorticosteroids might mediate the adverse effects of exposure to non-persistent {{< acr edcs >}} on childhood neurodevelopment (too broad) in larger populations.
+In conclusion, in a study of 1,297 children from 6 European birth cohorts, we observed that (i) exposure to non-persistent {{< acr edcs >}} might have short-term effects on {{< acr hitrtse >}}, (ii) exposure to non-persistent {{< acr edcs >}} might disrupt the {{< acr hpa >}} axis, and (iii) disruption of the {{< acr hpa >}} axis might have short-term, sex-specific effects on {{< acr hitrtse >}}.
 

### 2024-04-10 output/paper/manuscript.qmd
Trying to convert md to docx manually.

@@
       theme: neutral
+    keep-md: true
   pdf:

### 2024-04-10 output/paper/manuscript.qmd
Shorter abstract and discussion.

--- 2024-04-05 output/paper/manuscript.qmd
@@
       theme: neutral
+    output-file: "Main document - anonymous"
   pdf:
@@
 We aimed to evaluate the effects of childhood exposure to organophosphate pesticides, phenols, and phthalate metabolites, on urinary glucocorticosteroids and inattention in children using data from the {{< acr helix >}} cohort.
-We used the parametric g-formula to estimate effects between {{< acr edcs >}}, glucocorticosteroids, and {{< acr hitrtse >}}, a measure of inattention from the {{< acr ant >}}, and tested for possible effect modification by sex.
+We used the parametric g-formula to estimate associations between {{< acr edcs >}}, glucocorticosteroids, and {{< acr hitrtse >}}, a measure of inattention, and tested for possible effect modification by sex.
 We observed a positive {{< acr mc >}} for exposure increases from the 10th to the 90th percentile for methyl-paraben (`r format_res(meffects1, "mepa", "HRT-SE")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "HRT-SE")`), oh-MiNP (`r format_res(meffects1, "ohminp", "HRT-SE")`), and MEHP (`r format_res(meffects1, "mehp", "HRT-SE")`), on {{< acr hitrtse >}}, indicating lower attention.
@@
 Increased levels of the glucocorticosteroids had no effect on {{< acr hitrtse >}}, although we found a possible effect modification by sex.
-Our results suggest that multiple {{< acr edcs >}} might interfere with inattention in children and with the homeostasis of the {{< acr hpa >}} axis.
+Our results suggest that multiple {{< acr edcs >}} might interfere with inattention and with the homeostasis of the {{< acr hpa >}} axis.
 
@@
 More generally, the literature on childhood exposure to non-persistent {{< acr edcs >}} and other neurodevelopment outcomes in children has mostly focused on {{< acr opps >}} [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @CartierWarembourgLeManer-Idrissi:2016; @YuDuChiou:2016], phthalate metabolites [@HuangChenSu:2015; @HuangTsaiChen:2017; @KimHongShin:2017; @LiPapandonatosCalafat:2019; @BalalianWhyattLiu:2019; @ShoaffCoullWeuve:2020; @JankowskaPolanskaHanke:2019; @VilmandBeckBilenberg:2023], and {{< acr bpa >}} [@TewarAuingerBraun:2016; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019].
-González-Alzaga et al.
-and Cartier et al.
-evaluated cross-sectional associations between {{< acr dap >}} metabolites and subtests of the Wechsler Intelligence Scale for Children [@WechslerKodama:2003] in European children with ages between 6 and 11 years.
-Higher levels of {{< acr dap >}} metabolites (DMP, DMTP, DMDTP, DEP, DETP, and DEDTP) were associated with lower scores of {{< acr iq >}} and verbal comprehension, especially in boys [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015], while higher levels of diethylphosphate metabolites (DEP, DETP, DEDTP) were associated with lower working memory scores [@CartierWarembourgLeManer-Idrissi:2016].
+
+In children aged 6 to 11 years, higher levels of {{< acr dap >}} metabolites were associated with lower scores of {{< acr iq >}} and verbal comprehension, especially in boys [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015], while higher levels of diethylphosphate metabolites were associated with lower working memory scores [@CartierWarembourgLeManer-Idrissi:2016].
 There is also preliminary evidence of a possible association between exposure to certain {{< acr opps >}} and {{< acr adhd >}} in children [@BouchardBellingerWright:2010; @YuDuChiou:2016].
-Specifically, Bouchard et al.
-found evidence of a cross-sectional association between dimethyl alkylphosphate metabolites (DMP, DMTP, and DMDTP) and {{< acr adhd >}} in children aged 8 to 15 years from {{< acr nhanes >}}, while Yu et al.
-found a dose-response relationship between DMP and {{< acr adhd >}} in Taiwanese children aged 4 to 15 years.
 

### 2024-04-10 output/paper/manuscript.qmd
Added title page. Updated text for results effect modification since different.

@@
 @Tbl-hypothesis-1and2 presents the results of the difference between estimates of the {{< acr mcs >}} on the logarithmic scale for females and males, for the {{< acr edcs >}} on the glucocorticosteroids and {{< acr hitrtse >}}.
-For {{< acr hitrtse >}}, significant differences were present for the phenol OXBE (`r format_res(heffects1, "oxbe", "HRT-SE")`) and the phthalate metabolites MEP (`r format_res(heffects1, "mep", "HRT-SE")`) and MbZP (`r format_res(heffects1, "mbzp", "HRT-SE")`).
+For {{< acr hitrtse >}}, significant differences were present for the phenol OXBE (`r format_res(heffects1, "oxbe", "HRT-SE")`) and the phthalate metabolite MbZP (`r format_res(heffects1, "mbzp", "HRT-SE")`).
 For the glucocorticosteroids, significant differences were present across all three classes of {{< acr edcs >}} and for all outcomes.
-The largest differences were attributable to the {{< acr opps >}} DMTP (cortisol production, `r format_res(heffects2, "dmtp", "cortisol production")`) and DETP (corticosterone production, (`r format_res(heffects2, "detp", "corticosterone production")`); cortisone production, (`r format_res(heffects2, "detp", "cortisone production")`)).
+The largest differences were attributable to the phenol ETPA (corticosterone production, `r format_res(heffects2, "etpa", "corticosterone production")`) and the phthalate metabolite MEHP (cortisol production, (`r format_res(heffects2, "mehp", "cortisol production")`); cortisone production, (`r format_res(heffects2, "mehp", "cortisone production")`)).
 The forest plots of the individual {{< acr mcs >}} are presented in Supplementary Figures [-@suppfig-marginal-1sa] and [-@suppfig-marginal-2sa].

### 2024-04-10 output/paper/title_page.qmd
Added title page. Updated text for results effect modification since different.

---
title: "Childhood exposure to non-persistent endocrine disruptors, glucocorticosteroids, and attentional function: A study based on the parametric g-formula"

metadata-files:
  - ../../docs/authors.yml

embed-resources: true
format:
  html: default

link-citations: true
linkcolor: blue

editor_options:
  chunk_output_type: console
---

## Funding

The research leading to these results has received funding from the European Union's Horizon 2020 research and innovation programme under grant agreements no. 825712 [OBERON] and no. 874583 [ATHLETE].

Data were collected as part of the European Community’s Seventh Framework Programme (FP7/2007-2013) under grant agreement no 308333 (HELIX).

Born in Bradford receives funding from a joint grant from the UK Medical Research Council (MRC) and UK Economic and Social Science Research Council (ESRC) [MR/N024391/1]; the British Heart Foundation [CS/16/4/32482]; a Wellcome Infrastructure Grant [WT101597MA]; The National Institute for Health Research under its Applied Research Collaboration for Yorkshire and Humber [NIHR200166]. The views expressed are those of the author(s), and not necessarily those of the NHS, the NIHR or the Department of Health and Social Care.

The EDEN study was supported by Foundation for medical research (FRM), National Agency for Research (ANR), National Institute for Research in Public health (IRESP: TGIR cohorte santé 2008 program), French Ministry of Health (DGS), French Ministry of Research, INSERM Bone and Joint Diseases National Research (PRO-A), and Human Nutrition National Research Programs, Paris-Sud University, Nestlé, French National Institute for Population Health Surveillance (InVS), French National Institute for Health Education (INPES), the European Union FP7 programmes (FP7/2007–2013, HELIX, ESCAPE, ENRIECO, Medall projects), Diabetes National Research Program (through a collaboration with the French Association of Diabetic Patients (AFD)), French Agency for Environmental Health Safety (now ANSES), Mutuelle Générale de l’Education Nationale a complementary health insurance (MGEN), French national agency for food security, French-speaking association for the study of diabetes and metabolism (ALFEDIAM).

INMA data collections were supported by grants from the Instituto de Salud Carlos III, CIBERESP, and the Generalitat de Catalunya-CIRIT.

KANC was funded by the grant of the Lithuanian Agency for Science Innovation and Technology (6-04-2014_31V-66).

The Norwegian Mother, Father and Child Cohort Study is supported by the Norwegian Ministry of Health and Care Services and the Ministry of Education and Research.

The Rhea project was financially supported by European projects (EU FP6-2003-Food-3-NewGeneris, EU FP6. STREP Hiwate, EU FP7 ENV.2007.1.2.2.2. Project No 211250 Escape, EU FP7-2008-ENV-1.2.1.4 Envirogenomarkers, EU FP7-HEALTH-2009- single stage CHICOS, EU FP7 ENV.2008.1.2.1.6. Proposal No 226285 ENRIECO, EU- FP7- HEALTH-2012 Proposal No 308333 HELIX), and the Greek Ministry of Health (Program of Prevention of obesity and neurodevelopmental disorders in preschool children, in Heraklion district, Crete, Greece: 2011-2014; “Rhea Plus”: Primary Prevention Program of Environmental Risk Factors for Reproductive Health, and Child Health: 2012-15).

ISGlobal acknowledges support from the Spanish Ministry of Science and Innovation through the “Centro de Excelencia Severo Ochoa 2019-2023” Program (CEX2018-000806-S), and support from the Generalitat de Catalunya through the CERCA Program.

## Acknowledgments

The authors acknowledge that Born in Bradford is only possible because of the enthusiasm and commitment of the children and parents in Born in Bradford. We are grateful to all participants, health professionals and researchers who have made Born in Bradford happen.

The authors thank the EDEN mother-child cohort study group, whose members are I. Annesi-Maesano, J.Y. Bernard, M.A. Charles, P. Dargent-Molina, B. de Lauzon-Guillain, P. Ducimetière, M. de Agostini, B. Foliguet, A. Forhan, X. Fritel, A. Germa, V. Goua, R. Hankard, B. Heude, M. Kaminski, B. Larroquey, N. Lelong, J. Lepeule, G. Magnin, L. Marchand, C. Nabet, F Pierre, R. Slama, M.J. Saurel-Cubizolles, M. Schweitzer, and O. Thiebaugeorges.

### 2024-07-18 output/paper/manuscript.qmd
Comments AJE.

@@
 ---
-title: "Childhood exposure to non-persistent endocrine disruptors, glucocorticosteroids, and attentional function: A study based on the parametric g-formula"
+title: "Childhood exposure to non-persistent endocrine disruptors, glucocorticosteroids, and attentional function: A cross-sectional study based on the parametric g-formula"
 
@@
 
+output-file: "Main document - anonymous"
 embed-resources: true
+number-sections: true
+toc: false
+citeproc: true
+mermaid:
+  theme: neutral
+
 format:
@@
     page-width: 8.3
-    citeproc: true
-    mermaid:
-      theme: neutral
-    output-file: "Main document - anonymous"
   pdf:
@@
     pdf-engine: lualatex
-    number-sections: true
-    toc: false
     keep-tex: false
@@
         \usepackage{typearea}
+        \usepackage{amsmath, amssymb, xcolor}
+        \usepackage{pgf, tikz}
+        \usetikzlibrary{external, trees, backgrounds, swigs, calc, positioning, shapes, arrows, arrows.meta, shapes.arrows, shapes.geometric, shapes.multipart, decorations.pathmorphing, decorations.markings}
     include-before-body:
@@
 bibliography: bibliography.bib
-csl: american-journal-of-epidemiology.csl
+csl: epidemiology.csl
 
@@
 Several {{< acr edcs >}} were also associated with a positive {{< acr mc >}} for cortisone, cortisol, and corticosterone production.
-Increased levels of the glucocorticosteroids had no effect on {{< acr hitrtse >}}, although we found a possible effect modification by sex.
+Increased levels of the glucocorticosteroids were not associated with {{< acr hitrtse >}}, although we found a possible effect modification by sex.
 Our results suggest that multiple {{< acr edcs >}} might interfere with inattention and with the homeostasis of the {{< acr hpa >}} axis.
@@
 Further, each minimal adjustment set was *augmented* with precision covariates, defined as the set of parents variable of the outcome that are not parents of the exposure.
-Common confounders were cohort, ethnicity, sex, age, height, weight, and head circumference of the child, consumption of fish, fruit, vegetables, organic food, anf fast food, maternal tobacco consumption, family financial situation and affluence scale (FAS).
+Common confounders were cohort, ethnicity, sex, age, height, weight, and head circumference of the child, consumption of fish, fruit, vegetables, organic food, and fast food with a food frequency questionnaire, maternal tobacco consumption, family financial situation and affluence scale (FAS).
 Models for estimating the effects of {{< acr edcs >}} on {{< acr hitrtse >}} were further adjusted for child breastfeeding, prenatal maternal active and passive smoking, urine creatinine, child mood and rest before assessment, child neuropsychological diagnosis, marital status, season, and fasting time before assessment.
@@
 We estimated {{< acr mcs >}} with the parametric g-formula, a method of standardization.
-The parametric g-formula involves the following steps: 1) fit a outcome model including both covariates and balancing weights; 2) create two new datasets identical to the original one but with the exposure shifted according to a user-specified intervention set by a deterministic function of the observed exposure levels; 3) use the outcome model to compute adjusted predictions in the two counterfactual datasets; 4) compute the difference between the means of the adjusted predictions in the counterfactual datasets.
+The parametric g-formula involves the following steps: 1) fit a outcome model including both covariates and balancing weights; 2) create two new datasets identical to the original one but with the exposure shifted according to a user-specified dynamic intervention set by a deterministic function of the observed exposure levels; 3) use the outcome model to compute adjusted predictions in the two counterfactual datasets; 4) compute the difference between the means of the adjusted predictions in the counterfactual datasets.
 The causal parameter of interest was thus specified as the difference in the expected counterfactual outcomes under the shifted exposure levels $\left( \mathbb{E} \left[ Y^{d_1} \right] - \mathbb{E} \left[ Y^{d_2} \right] \right)$.
@@
 To estimate the {{< acr mcs >}}, we used the `avg_comparisons` function from the `marginaleffects` R package [@Arel-Bundock:2023].
-The two counterfactual datasets were obtained by setting the exposures levels to 90th percentile ($d_1$) and the 10th percentile ($d_2$), for each cohort separately.
+The two counterfactual datasets were obtained by setting the exposures levels to 90th percentile ($d_1$) and the 10th percentile ($d_2$), for each cohort separately to account for cross-cohort differences in the levels of the exposures (@supptbl-chems-desc and @supptbl-met-desc).
 The {{< acr mcs >}} were computed using the estimated balancing weights above.
@@
 
-Levels of unprocessed non-persistent {{< acr edcs >}}, after imputation of values below the {{< acr lod >}}, and glucocorticosteroids, are presented in @tbl-edc-desc, @tbl-met-new-desc, and @supptbl-met-desc.
+Levels of unprocessed non-persistent {{< acr edcs >}}, after imputation of values below the {{< acr lod >}}, and glucocorticosteroids, are presented in @tbl-edc-desc, @tbl-met-new-desc, @supptbl-chems-desc, and @supptbl-met-desc.
 Supplementary Figures [-@suppfig-desc-chems] and [-@suppfig-desc-mets] provide information on the measurement classification of the {{< acr edcs >}} and glucocorticosteroids by cohort, respectively.
@@
 Further, this study used pooled urine samples for chemical assessment to obtain more representative long-term exposures, since it is known that these specific {{< acr edcs >}} have very short half-lives [@PerrierGiorgis-AllemandSlama:2016; @CasasBasaganaSakhi:2018].
-We decided to model both the *treatment* mechanisms, for the estimation of balancing weights, and the outcomes, with traditional covariates adjustment, to try to obtain *doubly robust* effect estimates.
+We decided to model both the *treatment* mechanisms, for the estimation of balancing weights, and the outcomes, with traditional covariates adjustment, to obtain *doubly robust* effect estimates.
 Finally, we decided not to interpret our results by focusing on the estimated coefficients of possibly misspecified regression models, but by making use of the g-computation procedure.
@@
 Importantly, the non-persistent {{< acr edcs >}} were measured in a pool of night and morning urine samples before the clinical visit, to represent exposure over the previous day, whereas the glucocorticosteroids were measured in the night urine sample.
+Cross-sectional designs also have limitations to study the etiologic period of susceptibility of attentional function.
+Prior studies have investigated associations between short-term effects of air pollution and attentional function, with null results in adolescents but higher response times in adults [@GignacBarrera-GomezPersavento:2021; @GignacRighiToran:2022].
 Although we included a wide range of confounders there is the possibility, as with other observational studies, of residual confounding, which might lead to a bias away from the null.

### 2024-07-18 output/paper/title_page.qmd
Modifying affiliations submission.
Comments AJE.

@@
 
+Oliver Robinson is supported by a UK Research and Innovation Future Leaders Fellowship MR/S03532X/1.
+
 ISGlobal acknowledges support from the Spanish Ministry of Science and Innovation through the “Centro de Excelencia Severo Ochoa 2019-2023” Program (CEX2018-000806-S), and support from the Generalitat de Catalunya through the CERCA Program.

### 2024-07-18 R/proc_res.R

@@ -25,7 +25,7 @@ tidy_codebooks <- function(rq) {
     ) |>
     tidylog::select(
       dag, variable, type, description, code, label,
-      remark, comments
+      remark, comments, period
     )
   meta <- meta |>
     dplyr::rowwise() |>
@@ -51,13 +51,14 @@ tidy_codebooks <- function(rq) {
     tidylog::rename(
       coding = "code",
       labels = "label",
-      remarks = "remark"
+      remarks = "remark",
+      assessment = "period"
     ) |>
     tidylog::mutate(
       dplyr::across(
-        c("coding", "labels", "remarks", "comments"),
+        c("coding", "labels", "remarks", "comments", "assessment"),
         \(x) ifelse(
-          x == "NA",
+          toupper(x) == "NA",
           "", x
         )
       )
@@ -328,6 +329,84 @@ tbl_desc_vars <- function() {
     )
   ##############################################################################
   
+  # EDCs full
+  dat$exposures <- dat_request$dat |>
+    tidylog::select(
+      params_dat$variables$identifier,
+      cohort,
+      dplyr::any_of(params_dat$variables[["rq1"]]$exposures)
+    ) |>
+    tidylog::mutate(
+      cohort = dplyr::case_when(
+        cohort == "EDE" ~ "EDEN",
+        cohort == "KAN" ~ "KANC",
+        cohort == "MOB" ~ "MOBA",
+        cohort == "RHE" ~ "RHEA",
+        cohort == "SAB" ~ "INMA",
+        .default = cohort
+      )
+    )
+  tbl_edcs <- dat$exposures |>
+    tidylog::rename_with(
+      .fn = \(x) gsub("hs_|_c", "", x)
+    ) |>
+    tidylog::select(
+      -dplyr::any_of(c("HelixID"))
+    ) |>
+    dplyr::rename_with(
+      ~tibble::deframe(edcs)[.x],
+      .cols = edcs$chem_id
+    )
+  tbl_edcs <- tbl_edcs |>
+    gtsummary::tbl_summary(
+      statistic = gtsummary::all_continuous() ~ "{median} ({p10}, {p25}, {p75}, {p90})",
+      missing = "no",
+      by = "cohort",
+      digits = dplyr::everything() ~ 1
+    ) |>
+    gtsummary::add_overall() |>
+    gtsummary::modify_spanning_header(
+      gtsummary::everything() ~ NA
+    )
+  
+  tbl_edcs$table_body <- dplyr::left_join(
[…]

### 2024-07-19 output/paper/_figures.qmd
Comments AJE.
Working on comments Reviewer 1 points 9 to 11.

@@
-<!-- title: "Figures" -->
-
-<!-- format: -->
-<!--   docx: -->
-<!--     reference-doc: reference_document.docx -->
-<!--   pdf: -->
-<!--     pdf-engine: lualatex -->
-<!--     number-sections: true -->
-<!--     toc: false -->
-<!--     include-in-header: -->
-<!--       text: | -->
-<!--         \usepackage{lineno} -->
-<!--         \usepackage{rotating} -->
-<!--         \usepackage{float} -->
-<!--         \usepackage{typearea} -->
-<!--     include-before-body: -->
-<!--       text: | -->
-<!--         \linenumbers -->
-
-<!-- link-citations: true -->
-<!-- linkcolor: blue -->
-
-<!-- editor_options: -->
-<!--   chunk_output_type: console -->
-
-<!-- execute: -->
-<!--   echo: false -->
-<!--   warning: false -->
-<!--   error: false -->
-<!--   message: false -->
+---
+editor_options:
+  chunk_output_type: console
+---
 
@@
 
+# DAG
+
+::: {#fig-dag}
+```{dot}
+//| fig-width: 9
+//| fig-height: 10
+
+digraph simple {
+    graph [rankdir=TB]
+
+    node [shape=ellipse]
+    U [label='Confounders']
+
+    node [shape=box, height=0.3, width=0.3]
+    E [label='EDCs']
+    M [label='Glucocorticosteroids']
+    Y [label='HRT-SE']
+    { rank = same; E M}
+
+    E -> M [minlen=3]
+    E -> Y
+    M -> Y
+    U -> E
+    U -> M
+    U -> Y
+  }
+```
+**Simplified {{< acr dag first_use=true >}} showing {{< acr edcs first_use=true >}}, {{< acr hitrtse first_use=true >}}, and the glucocorticosteroids.**
+:::
+
 # Figures for main results

### 2024-07-19 output/paper/_si.qmd
Comments AJE.
Working on comments Reviewer 1 points 9 to 11.

--- 2024-03-08 output/paper/_si.qmd
@@
 targets::tar_load(
-  dplyr::contains("codebook_"),
+  "codebooks",
   store = store
@@
 
-::: {#supptbl-codebook-1}
-```{r codebook-1}
-codebook_1
+::: {#supptbl-codebooks}
+```{r codebooks}
+codebooks
 ```
-**Codebook for the covariates used in the estimation of the marginal comparisons of {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}}.**
-:::
-
-::: {#supptbl-codebook-2}
-```{r codebook-2}
-codebook_2
-```
-**Codebook for the covariates used in the estimation of the marginal comparisons of {{< acr edcs first_use=true >}} on the glucocorticosteroids.**
-:::
-
-::: {#supptbl-codebook-3}
-```{r codebook-3}
-codebook_3
-```
-**Codebook for the covariates used in the estimation of the marginal comparisons of the glucocorticosteroids on {{< acr hitrtse first_use=true >}}.**
+**Codebook for the covariates used in the estimation of the marginal comparisons of {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}} (`q1`), {{< acr edcs first_use=true >}} on the glucocorticosteroids (`q2`), and glucocorticosteroids on {{< acr hitrtse first_use=true >}} (`q3`).**
 :::
@@
 **Participant characteristics, by cohort and overall (HELIX subcohort; 2013-2016).**
+:::
+
+### Concentrations of the endocrine disruptors
+
+::: {#supptbl-chems-desc}
+```{r chems-desc}
+targets::tar_load("desc_chems_mets", store = store)
+desc_chems_mets$desc_chems_full
+```
+**Participants non-persistent {{< acr edcs first_use=true >}} concentrations, by cohort and overall (HELIX subcohort; 2013-2016).**
 :::
@@
 
+### Marginal contrasts for effect modification
+
+```{r}
+targets::tar_load(
+  dplyr::matches("marginal_comparisons_(\\d)SA$"),
+  store = store
+)
+```
+
+::: {#supptbl-marginal-1sa}
+```{r marginal-1sa-tbl}
+m1sa <- tidy_res_meffects(
+  df = marginal_comparisons_1SA,
+  sa_var = "e3_sex",
+  outcome = NULL,
+  which_res = "comparisons",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)
+
+m1sa$df_gt
+```
+**Marginal contrasts for effect modification by sex of a increase from the 10th to the 90th percentile of the {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
+:::
+
+::: {#supptbl-marginal-2sa}
+```{r marginal-2sa-tbl}
+m2sa <- tidy_res_meffects(
+  df = marginal_comparisons_2SA |>
+    tidylog::filter(
+      !grepl("metabolism", outcome)
+    ),
+  sa_var = "e3_sex",
+  outcome = NULL,
+  which_res = "comparisons",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)
+
+m2sa$df_gt
+```
+**Marginal contrasts for effect modification by sex of a increase from the 10th to the 90th percentile of the {{< acr edcs first_use=true >}} on the glucocorticosteroids expressed in ng/ml (HELIX subcohort; 2013-2016).**
+:::
+
+::: {#supptbl-marginal-3sa}
+```{r marginal-3sa-tbl}
+m3sa <- tidy_res_meffects(
+  df = marginal_comparisons_3SA,
+  sa_var = "e3_sex",
+  outcome = NULL,
+  which_res = "comparisons",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig
+)
+
+m3sa$df_gt
+```
+**Marginal contrasts for effect modification by sex of a increase from the 10th to the 90th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
+:::
+
 # Supplementary figures
@@
 
-```{r}
-targets::tar_load(
-  dplyr::matches("marginal_comparisons_(\\d)SA$"),
-  store = store
-)
-```
-
 ::: {#suppfig-marginal-1sa}
@@
 
-tidy_res_meffects(
-  df = marginal_comparisons_1SA,
-  sa_var = "e3_sex",
-  outcome = NULL,
-  which_res = "comparisons",
-  num_digits_est = num_digits_est,
-  num_digits_sig = num_digits_sig
-)$plot
+m1sa$plot
 ```
-**Marginal contrasts for effect modification by sex of a decrease from the 90th to the 10th percentile of the {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
+**Marginal contrasts on the logarithmic scale for effect modification by sex of a increase from the 10th to the 90th percentile of the {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
 Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}. The size of the circles represents the *S* value of the effect estimate [@RafiGreenland:2020].
@@
 
-tidy_res_meffects(
-  df = marginal_comparisons_2SA |>
-    tidylog::filter(
-      !grepl("metabolism", outcome)
-    ),
-  sa_var = "e3_sex",
-  outcome = NULL,
-  which_res = "comparisons",
-  num_digits_est = num_digits_est,
-  num_digits_sig = num_digits_sig
-)$plot
+m2sa$plot
 ```
-**Marginal contrasts for effect modification by sex of a decrease from the 90th to the 10th percentile of the {{< acr edcs first_use=true >}} on the glucocorticosteroids expressed in ng/ml (HELIX subcohort; 2013-2016).**
+**Marginal contrasts on the logarithmic scale for effect modification by sex of a increase from the 10th to the 90th percentile of the {{< acr edcs first_use=true >}} on the glucocorticosteroids expressed in ng/ml (HELIX subcohort; 2013-2016).**
 Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}. The size of the circles represents the *S* value of the effect estimate [@RafiGreenland:2020].
@@
 
-tidy_res_meffects(
-  df = marginal_comparisons_3SA,
-  sa_var = "e3_sex",
-  outcome = NULL,
-  which_res = "comparisons",
-  num_digits_est = num_digits_est,
-  num_digits_sig = num_digits_sig
-)$plot
+m3sa$plot
 ```
-**Marginal contrasts for effect modification by sex of a decrease from the 90th to the 10th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
+**Marginal contrasts on the logarithmic scale for effect modification by sex of a increase from the 10th to the 90th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
 Circles and triangles indicate effect estimates.

### 2024-07-19 output/paper/manuscript.qmd
Compiled to PDF to check errors: missing acronyms and some references.
Working on comments Reviewer 1 points 9 to 11.

@@
     page-width: 8.3
-  pdf:
-    documentclass: article
-    geometry:
-      - top=30mm
-      - left=20mm
-      - heightrounded
-    pdf-engine: lualatex
-    keep-tex: false
-    include-in-header:
-      text: |
-        \usepackage{lineno}
-        \usepackage{rotating}
-        \usepackage{float}
-        \usepackage{typearea}
-        \usepackage{amsmath, amssymb, xcolor}
-        \usepackage{pgf, tikz}
-        \usetikzlibrary{external, trees, backgrounds, swigs, calc, positioning, shapes, arrows, arrows.meta, shapes.arrows, shapes.geometric, shapes.multipart, decorations.pathmorphing, decorations.markings}
-    include-before-body:
-      text: |
-        \linenumbers
 
@@
   - acronyms
-  - ../../../../all_phd/_extensions/wordcount/wordcount.lua
 
 acronyms:
-  fromfile: ../../../../all_phd/acronyms.yml
+  fromfile: acronyms.yml
   insert_links: false
@@
 The adjustment sets are provided in the Supplementary Material as text files compatible with DAGitty.
-Codebooks for the used covariates, by research question, are provided in Supplementary Tables [-@supptbl-codebook-1; -@supptbl-codebook-2; -@supptbl-codebook-3].
+A simplified {{< acr swig >}} is provided in @fig-dag.
+Codebooks for the used covariates are provided in @supptbl-codebooks.
 
@@
 The largest differences were attributable to the phenol ETPA (corticosterone production, `r format_res(heffects2, "etpa", "corticosterone production")`) and the phthalate metabolite MEHP (cortisol production, (`r format_res(heffects2, "mehp", "cortisol production")`); cortisone production, (`r format_res(heffects2, "mehp", "cortisone production")`)).
-The forest plots of the individual {{< acr mcs >}} are presented in Supplementary Figures [-@suppfig-marginal-1sa] and [-@suppfig-marginal-2sa].
+The individual {{< acr mcs >}} are presented in Supplementary Tables [-@supptbl-marginal-1sa], [-@supptbl-marginal-2sa] and [-@supptbl-marginal-3sa], and in Supplementary Figures [-@suppfig-marginal-1sa], [-@suppfig-marginal-2sa] and [-@suppfig-marginal-3sa].
 

### 2024-07-19 R/proc_res.R

@@ -1,68 +1,80 @@
 #' Title
 #'
-#' @param rq
-#'
 #' @return
 #' @export
-tidy_codebooks <- function(rq) {
+tidy_codebooks <- function() {
   path_store <- Sys.getenv("path_store")
   
   ## Load objects
-  targets::tar_load(
-    paste0("rq", rq, "_preproc_dat"),
-    store = paste0(path_store, rq)
-  )
-  
-  ## Extract adjustments set used and mapped covariates
-  adj_set <- get(paste0("rq", rq, "_preproc_dat"))$adjustment_set
-  mapped_covars <- get(paste0("rq", rq, "_preproc_dat"))$mapping_covariates
-  actual_covars <- get(paste0("rq", rq, "_preproc_dat"))$covariates |>
-    colnames()
-  meta <- get(paste0("rq", rq, "_preproc_dat"))$meta |>
-    tidylog::filter(
-      dag %in% adj_set &
-        variable %in% mapped_covars
-    ) |>
-    tidylog::select(
-      dag, variable, type, description, code, label,
-      remark, comments, period
+  metas <- list()
+  for (rq in 1:3) {
+    targets::tar_load(
+      paste0("rq", rq, "_preproc_dat"),
+      store = paste0(path_store, rq)
     )
-  meta <- meta |>
-    dplyr::rowwise() |>
-    tidylog::mutate(
-      included = ifelse(
-        variable %in% actual_covars,
-        TRUE, FALSE
+    ## Extract adjustments set used and mapped covariates
+    adj_set <- get(paste0("rq", rq, "_preproc_dat"))$adjustment_set
+    mapped_covars <- get(paste0("rq", rq, "_preproc_dat"))$mapping_covariates
+    actual_covars <- get(paste0("rq", rq, "_preproc_dat"))$covariates |>
+      colnames()
+    meta <- get(paste0("rq", rq, "_preproc_dat"))$meta |>
+      tidylog::filter(
+        dag %in% adj_set &
+          variable %in% mapped_covars
+      ) |>
+      tidylog::select(
+        dag, variable, type, description, code, label,
+        remark, comments, period
       )
+    meta <- meta |>
+      dplyr::rowwise() |>
+      tidylog::mutate(
+        included = ifelse(
+          variable %in% actual_covars,
+          "Yes", "No"
+        )
+      ) |>
+      tidylog::mutate(
+        dplyr::across(
+          dplyr::everything(),
+          as.character
+        )
+      ) |>
[…]

### 2024-07-19 _targets_res.R

@@ -35,6 +35,11 @@ list(
     command = tbl_desc_vars()
   ), # End desc_chems_mets target
   ##############################################################################
+  targets::tar_target(
+    name = "codebooks",
+    command = tidy_codebooks()
+  ), # End codebook target
+  ##############################################################################
   tarchetypes::tar_map(
     list(rq_ = c("1", "2", "3", "1SA", "2SA", "3SA")),
     targets::tar_target(
@@ -98,14 +103,6 @@ list(
     name = "all_dags",
     command = dags()
   ), # End dags target
-  ##############################################################################
-  tarchetypes::tar_map(
-    list(rq = c("1", "2", "3")),
-    targets::tar_target(
-      name = "codebook",
-      command = tidy_codebooks(rq = rq)
-    ) # End codebook target
-  ), # End loop over RQs
   ##############################################################################
   tarchetypes::tar_map(
     list(rq = c("1", "2", "3")),

### 2024-07-19 run_loop.R

@@ -1,4 +1,4 @@
-Sys.setenv(is_hpc = TRUE)
+Sys.setenv(is_hpc = FALSE)
 # Custom path to _targets for different research questions
 path_store <- ifelse(
   Sys.getenv("is_hpc"), 

### 2024-07-19 output/paper/_si.qmd
Finished addressing comments reviwer 1.

@@
 
-m1sa$df_gt
+m1sa$table
 ```
-**Marginal contrasts for effect modification by sex of a increase from the 10th to the 90th percentile of the {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
+**Marginal contrasts on the logarithmic scale for effect modification by sex of a increase from the 10th to the 90th percentile of the {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
 :::
@@
 
-m2sa$df_gt
+m2sa$table
 ```
-**Marginal contrasts for effect modification by sex of a increase from the 10th to the 90th percentile of the {{< acr edcs first_use=true >}} on the glucocorticosteroids expressed in ng/ml (HELIX subcohort; 2013-2016).**
+**Marginal contrasts on the logarithmic scale for effect modification by sex of a increase from the 10th to the 90th percentile of the {{< acr edcs first_use=true >}} on the glucocorticosteroids expressed in ng/ml (HELIX subcohort; 2013-2016).**
 :::
@@
 
-m3sa$df_gt
+m3sa$table
 ```
-**Marginal contrasts for effect modification by sex of a increase from the 10th to the 90th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
+**Marginal contrasts on the logarithmic scale for effect modification by sex of a increase from the 10th to the 90th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
 :::

### 2024-07-19 output/paper/manuscript.qmd
Finished addressing comments reviwer 1.

@@
 The largest differences were attributable to the phenol ETPA (corticosterone production, `r format_res(heffects2, "etpa", "corticosterone production")`) and the phthalate metabolite MEHP (cortisol production, (`r format_res(heffects2, "mehp", "cortisol production")`); cortisone production, (`r format_res(heffects2, "mehp", "cortisone production")`)).
-The individual {{< acr mcs >}} are presented in Supplementary Tables [-@supptbl-marginal-1sa], [-@supptbl-marginal-2sa] and [-@supptbl-marginal-3sa], and in Supplementary Figures [-@suppfig-marginal-1sa], [-@suppfig-marginal-2sa] and [-@suppfig-marginal-3sa].
+The individual {{< acr mcs >}} are presented in Supplementary Tables [-@supptbl-marginal-1sa] and [-@supptbl-marginal-2sa], and in Supplementary Figures [-@suppfig-marginal-1sa] and [-@suppfig-marginal-2sa].
 
@@
 Furthermore, for all exposures, the {{< acr mcs >}} had opposite sign (positive for males and negative for females).
-The forest plot of the individual {{< acr mcs >}} is presented in @suppfig-marginal-3sa.
+The individual {{< acr mcs >}} are presented in @supptbl-marginal-3sa and in @suppfig-marginal-3sa.
 
@@
 
-We are also not aware of prior epidemiological studies specifically investigating the effects of elevated levels of glucocorticosteroids in relation to attentional function, although there is evidence that under- or over-production of glucocorticosteroids interfere with the normal development of the brain [@LupienMcEwenGunnar:2009].
+We are also not aware of prior epidemiological studies specifically investigating the effects of elevated levels of glucocorticosteroids in relation to attentional function, although there is evidence that under- or over-production of glucocorticosteroids interfere with the normal development of the brain, possibly in a sex-specific manner [@LupienMcEwenGunnar:2009].
 
@@
 Strengths include its relative large sample size and its inclusion of multiple classes of non-persistent {{< acr edcs >}}.
-Further, this study used pooled urine samples for chemical assessment to obtain more representative long-term exposures, since it is known that these specific {{< acr edcs >}} have very short half-lives [@PerrierGiorgis-AllemandSlama:2016; @CasasBasaganaSakhi:2018].
+Further, this study used pooled urine samples for chemical assessment to reduce measurement error, since it is known that these specific {{< acr edcs >}} have very short half-lives [@PerrierGiorgis-AllemandSlama:2016; @CasasBasaganaSakhi:2018].
 We decided to model both the *treatment* mechanisms, for the estimation of balancing weights, and the outcomes, with traditional covariates adjustment, to obtain *doubly robust* effect estimates.

### 2024-07-19 R/proc_res.R

@@ -1132,7 +1132,7 @@ tidy_res_meffects <- function(df, sa_var, outcome,
   names_ <- if (is.null(sa_var) | which_res == "hypothesis") {
     names_
   } else {
-    paste0(names_, "_", c(unique(df$modifier)))
+    as.vector(outer(names_, c(unique(df$modifier)), paste, sep = "_"))
   }
   processed_df <- df |>
     tidylog::mutate(
@@ -1174,6 +1174,7 @@ tidy_res_meffects <- function(df, sa_var, outcome,
     tidylog::select(-estimate) |>
     tidyr::pivot_wider(
       names_from = dplyr::any_of(c("outcome", "modifier")),
+      names_sep = " ",
       values_from = c("val")
     ) |>
     gt::gt(
@@ -1186,43 +1187,15 @@ tidy_res_meffects <- function(df, sa_var, outcome,
       ),
       locations = gt::cells_row_groups()
     ) |>
-    # gt::tab_style(
-    #   style = gt::cell_text(weight = "bold"),
-    #   locations = purrr::map(
-    #     names_,
-    #     \(x) {
-    #       gt::cells_body(
-    #         columns = x,
-    #         rows = stringr::str_detect(!!rlang::sym(x), "\\*")
-    #       )
-    #     }
-    #   )
-    # ) |>
-    # gt::tab_footnote(
-    #   footnote = "*Significant results."
-    # ) |>
     gt::tab_footnote(
       footnote = "Estimate and 95% CI.",
       locations = gt::cells_column_labels(
-        columns = names_
+        columns = stringr::str_replace_all(names_, "_", " ")
       )
     ) |>
     gt::opt_footnote_marks(
       marks = "letters"
     )
-  
-  # if (is.null(sa_var) & which_res == "comparisons") {
-  #   processed_df <- dplyr::bind_rows(
-  #     tibble::tibble(
-  #       variable = "Exposure",
-  #       val = "Marginal contrast (95% CI)",
-  #       estimate = 100,
-  #       outcome = "",
-  #       class = ""
-  #     ),
-  #     processed_df
-  #   )
-  # }
   
   return(list(
     numerical_results = processed_df,

### 2024-07-21 output/paper/manuscript.qmd
Finished comments reviewer 2. Added template docx since need to add some options not available in Quarto. Deleted some old files.

@@
 title: "Childhood exposure to non-persistent endocrine disruptors, glucocorticosteroids, and attentional function: A cross-sectional study based on the parametric g-formula"
-
-# metadata-files:
-#   - ../../docs/authors.yml
 
@@
   docx:
-    page-width: 8.3
+    reference-doc: custom-reference-doc.docx
 
@@
 
+**Background**
+
 Evidence suggests that {{< acr edcs >}} may perturb the {{< acr hpa >}} axis, which has a major role in brain development.
-We aimed to evaluate the effects of childhood exposure to organophosphate pesticides, phenols, and phthalate metabolites, on urinary glucocorticosteroids and inattention in children using data from the {{< acr helix >}} cohort.
-We used the parametric g-formula to estimate associations between {{< acr edcs >}}, glucocorticosteroids, and {{< acr hitrtse >}}, a measure of inattention, and tested for possible effect modification by sex.
+We aimed to evaluate the effects of childhood exposure to organophosphate pesticides, phenols, and phthalate metabolites, on urinary glucocorticosteroids and inattention in childhood.
+
+**Methods**
+
+We used data from the {{< acr helix >}} cohort (2013-2016) and the parametric g-formula to estimate associations between {{< acr edcs >}}, glucocorticosteroids, and {{< acr hitrtse >}}, a measure of inattention, and tested for possible effect modification by sex.
+
+**Results**
+
 We observed a positive {{< acr mc >}} for exposure increases from the 10th to the 90th percentile for methyl-paraben (`r format_res(meffects1, "mepa", "HRT-SE")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "HRT-SE")`), oh-MiNP (`r format_res(meffects1, "ohminp", "HRT-SE")`), and MEHP (`r format_res(meffects1, "mehp", "HRT-SE")`), on {{< acr hitrtse >}}, indicating lower attention.
@@
 Increased levels of the glucocorticosteroids were not associated with {{< acr hitrtse >}}, although we found a possible effect modification by sex.
+
+**Conclusions**
+
 Our results suggest that multiple {{< acr edcs >}} might interfere with inattention and with the homeostasis of the {{< acr hpa >}} axis.
 
+**Keywords:** Endocrine Disruptors, Cortisone, Hydrocortisone, Neuropsychological Tests, Causality.
+
 {{< pagebreak >}}
+
+# Introduction {#sec-intro}
 
@@
 Briefly, we analyzed a total of 7 phenols (`r print_tidy_abbreviations_edcs(info_edcs, "Phenols")`), 6 non-specific organophosphate pesticide metabolites (`r print_tidy_abbreviations_edcs(info_edcs, "OP pesticide metabolites")`), and 10 phthalate metabolites (`r print_tidy_abbreviations_edcs(info_edcs, "Phthalate metabolites")`) originating from 6 distinct phthalate parent compounds.
-The laboratory protocols for the analysis are described elsewhere [@HaugSakhiCequier:2018].
+Laboratory protocols for the analysis are described in @HaugSakhiCequier:2018, while quality control and interlaboratory comparison are described in @MaitreBontCasas:2018.
 
@@
 The tests were administered in a standardized way, and with minimal interference from the field workers.
-Further information can be found in @MaitreBontCasas:2018.
+Further information can be found elsewhere [@RuedaFanMcCandliss:2004; @FornsEsnaolaLopez-Vicente:2014; @MaitreBontCasas:2018].
 The outcome of interest for the present study is the {{< acr hitrtse first_use=true >}} [@SunyerEsnaolaAlvarez-Pedrerol:2015], a measure of response speed consistency throughout the test.
@@
 For effect estimation of the {{< acr edcs >}} on glucocorticosteroids and of glucocorticosteroids on {{< acr hitrtse >}}, these sets were also sufficient to estimate direct effects.
-Sample-specific creatinine values were used to adjust for possible dilution effects.
+Sample-specific creatinine values were included in the models to adjust for possible dilution effects.
 Further, each minimal adjustment set was *augmented* with precision covariates, defined as the set of parents variable of the outcome that are not parents of the exposure.
@@
 Robust standard errors were computed with the `sandwich` R package, using cohort as variable indicating clustering of observations [@Zeileis:2004; @ZeileisKollGraham:2020].
-For each outcome, we report the results as differences between {{< acr mcs >}}.
+For each outcome, we report the results as differences between {{< acr mcs >}} with 95% {{< acr cis first_use=true >}}.
 
@@
 To do so, balancing weights were estimated separately for each level of the sex variable, and an interaction term between the exposure and sex was included in the outcome model.
-Similarly, the {{< acr mcs >}} were aggregated separately for each level of sex.
+Similarly, the {{< acr mcs >}} were estimated with the `avg_comparisons` function.
+For each outcome, we report the results as pairwise differences between female {{< acr mcs >}} and male {{< acr mcs >}}, with 95% {{< acr cis first_use=true >}}.
+
+by taking the difference between the mean potential outcome of females and males.
 
@@
 For most {{< acr edcs >}}, a cohort-specific increase in the levels of the exposures from the 10th to the 90th percentiles was associated with a positive {{< acr mc >}}, indicating an increase in the values of {{< acr hitrtse >}} and thus lower attention.
-Most of the {{< acr cis first_use=true >}} included the null effect, though.
-Significant effects were observed for the paraben MEPA (`r format_res(meffects1, "mepa", "HRT-SE")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "HRT-SE")`), oh-MiNP (`r format_res(meffects1, "ohminp", "HRT-SE")`), and MEHP (`r format_res(meffects1, "mehp", "HRT-SE")`).
+Most of the {{< acr cis >}} included the null effect, though.
+Noteworthy effects were observed for the paraben MEPA (`r format_res(meffects1, "mepa", "HRT-SE")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "HRT-SE")`), oh-MiNP (`r format_res(meffects1, "ohminp", "HRT-SE")`), and MEHP (`r format_res(meffects1, "mehp", "HRT-SE")`).
 The {{< acr opp >}} DETP was negatively associated with {{< acr hitrtse >}} (`r format_res(meffects1, "detp", "HRT-SE")`).
@@
 There is also preliminary evidence of a possible association between exposure to certain {{< acr opps >}} and {{< acr adhd >}} in children [@BouchardBellingerWright:2010; @YuDuChiou:2016].
+In our study we found a negative association between DETP and attentional function, indicating lower values of {{< acr hitrtse >}} with higher levels of DETP, with concordant results in the sex-stratified analysis.
+While we cannot exclude a protective effect of DETP in relation to attentional function, we note that in our study DETP has a higher percentage of missing values compared to the other {{< acr opps >}} and that other imputation strategies might lead to different associations.
 
@@
 Higher levels of di(2-ethylhexyl) phthalate metabolites (including MEHP, MEHHP, and MEOHP) were associated with lower intelligence scores in children aged 2 to 12 years [@HuangChenSu:2015], lower scores of {{< acr iq >}} and verbal intelligence, more omission errors (a measure of inattention), and higher scores of response time variability (a measure of sustained attention) in 6-year old Korean children [@KimHongShin:2017], poorer fine motor skills in preadolescent boys [@BalalianWhyattLiu:2019], and lower intelligence scores in 7-year old children [@VilmandBeckBilenberg:2023].
-Further associations were found for MEOHP with lower scores of {{< acr iq >}} [@HuangChenSu:2015] and verbal intelligence in Taiwanese children aged 6 to 12 years [@HuangTsaiChen:2017], and for dibutyl phthalate metabolites (MnBP and MiBP) with impaired verbal intelligence [@HuangTsaiChen:2017].
+Further associations were found for higher levels of MEOHP with lower scores of {{< acr iq >}} [@HuangChenSu:2015] and verbal intelligence in Taiwanese children aged 6 to 12 years [@HuangTsaiChen:2017], and for higher levels of dibutyl phthalate metabolites (MnBP and MiBP) with impaired verbal intelligence [@HuangTsaiChen:2017].
 Few studies have looked into different classes of non-persistent {{< acr edcs >}}.
 Shoaff et al.
-investigated cross-sectional associations between multiple {{< acr edcs >}} and {{< acr adhd >}}-related behaviors in 15-year old adolescents, finding a higher risk of {{< acr adhd >}}-related behavior problems with higher levels of antiandrogenic phthalate metabolites, especially in boys [@ShoaffCoullWeuve:2020].
+investigated cross-sectional associations between multiple {{< acr edcs >}} and {{< acr adhd >}}-related behaviors in 15-year old adolescents, finding a higher risk of {{< acr adhd >}}-related behavior problems with higher levels of antiandrogenic phthalate metabolites (molar sum of MnBP, MiBP, MBzP, MEHP, MEHHP, MEOHP, MECPP, monocarboxyoctyl phthalate, monohydroxyisobutyl phthalate (MHiBP), monohydroxybutyl phthalate (MHBP), and mono-isononyl phthalate) and the molar sum of di(2-ethylhexyl) phthalate (DEHP) metabolites (MECPP, MEHHP, MEOHP, and MEHP), especially in boys [@ShoaffCoullWeuve:2020].
 Our findings, indicating that short-term childhood exposure to certain phthalate metabolites (MEHP, oh-MiNP, and oxo-MiNP) was associated with attentional function, adds to this growing evidence base suggesting that childhood phthalate exposure may impact child neurodevelopment.
@@
 Except for working memory, there was no evidence of an association between BPA and cognitive abilities in Spanish boys aged 9 to 11 years [@Rodriguez-CarrilloMustielesPerez-Lobato:2019].
-We did not observe an association between BPA and attention function in the present study, but this study is the first to suggest that childhood paraben (MEPA) exposure may be associated with attentional function.
+We did not observe an association between BPA and attention function in the present study, but this study is the first to suggest that childhood exposure to MEPA may be associated with attentional function.
 
@@
 However, prior epidemiological research provides preliminary evidence for an association between certain non-persistent {{< acr edcs >}} measured at other time points with higher levels of glucocorticoids measured in other biological matrices [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023].
-Repeated measures up to 15 months of age of the phthalate metabolites MEHHP, MEOHP, MiBP, and MnBP showed positive associations with free cortisol in urine in Korean children [@KimLeeMoon:2018].
+Repeated measures up to 15 months of age of the phthalate metabolites MEHHP, MEOHP, MiBP, and MnBP showed positive associations with free urine cortisol in Korean children [@KimLeeMoon:2018].
 In a cohort of Chinese pregnant women, phthalate metabolites were measured at 14, 24, and 36 weeks of gestation, and the glucocorticoids cortisol and cortisone were measured in cord blood.
@@
 In a longitudinal study, a mixture of several phthalate metabolites, driven by MEP, MiBP, and MBzP, measured in childhood, showed a positive association with hair cortisol measured at 12 years of age [@SearsLiuLanphear:2023].
-Our findings also indicate associations between certain phthalate metabolites (MEHP, oh-MiNP, and oxo-MiNP) and glucocorticosteroids, but differences in the exposure assessment time points and in the biological matrices used for glucocorticosteroids determinations make a direct comparison difficult.
+Our findings also indicate associations between certain phthalate metabolites (MEHP, oh-MiNP, and oxo-MiNP) and glucocorticosteroids, but differences in the exposure assessment time points, the biological matrices used for glucocorticosteroids determinations, and the possible effect of parturition on cord concentrations of cortisol [@TribeTaylorKelly:2018], make a direct comparison difficult.
 
 Adding to these epidemiological studies, previous toxicological research provide evidence for the inhibition by phthalates of human 11$\beta$-hydroxysteroid dehydrogenase 2 (11$\beta$-HSD2) activity, responsible for the conversion of active cortisol into inactive cortisone [@ZhaoChuHuang:2010; @MaLianDong:2011].
-There is also *in silico* evidence suggesting that {{< acr bpa >}}, a phenol, and Triazophos (TAP), a organophosphorus insecticide, can bind to the human glucocorticoid receptor [@PrasanthDivyaSadasivan:2010; @YangLiRen:2019].
 
@@
 Prior studies have investigated associations between short-term effects of air pollution and attentional function, with null results in adolescents but higher response times in adults [@GignacBarrera-GomezPersavento:2021; @GignacRighiToran:2022].
-Although we included a wide range of confounders there is the possibility, as with other observational studies, of residual confounding, which might lead to a bias away from the null.
+Finally, longitudinal studies are necessary for a formal causal mediation analysis.
+Although we included a wide range of confounders, there is the possibility, as with other observational studies, of residual confounding, which might lead to a bias away from the null.
+We decided not to include prenatal levels of the {{< acr edcs >}} as confounders due to the *small* correlation between prenatal and childhood levels of these non-persistent chemicals [@HaugSakhiCequier:2018].
 There is further the possibility of misspecification of the outcome model, although we included a spline of the exposure to relax some of the linearity assumptions.
+Finally, we acknowledge that the present study does not take into account the possible mixture effect of these chemicals.
+While methods for causal mixture analysis are available [@KeilBuckleyO:2020], we also emphasize that using these tools without thoughtful consideration may lead to misleading results [@WebsterWeisskopf:2020].
 
-In conclusion, in a study of 1,297 children from 6 European birth cohorts, we observed that (i) exposure to non-persistent {{< acr edcs >}} might have short-term effects on {{< acr hitrtse >}}, (ii) exposure to non-persistent {{< acr edcs >}} might disrupt the {{< acr hpa >}} axis, and (iii) disruption of the {{< acr hpa >}} axis might have short-term, sex-specific effects on {{< acr hitrtse >}}.
+In conclusion, in a study of 1,297 children from 6 European birth cohorts, we observed that (i) exposure to certain non-persistent {{< acr edcs >}} is associated with higher values of {{< acr hitrtse >}} and might disrupt the {{< acr hpa >}} axis, and (ii) certain glucocorticosteroids are associated with {{< acr hitrtse >}} in a sex-specific manner.
 

### 2024-07-21 R/mtps.R

@@ -1,32 +1,0 @@
-#' Additive shift function
-#'
-#' @param dat A dataset containing the variable to be shifted. A tibble.
-#' @param trt The name of the variable to be shifted. A string.
-#'
-#' @return
-#'
-#' @export
-mtp_add <- function(dat, trt) {
-  params_ana <- params_analyses()[[Sys.getenv("TAR_PROJECT")]]
-  
-  dd <- dat[[trt]] + params_ana$shift_amount
-  
-  return(dd)
-}
-################################################################################
-
-#' Multiplicative shift function
-#'
-#' @param dat A dataset containing the variable to be shifted. A tibble.
-#' @param trt The name of the variable to be shifted. A string.
-#'
-#' @return
-#'
-#' @export
-mtp_mul <- function(dat, trt) {
-  params_ana <- params_analyses()[[Sys.getenv("TAR_PROJECT")]]
-  
-  dd <- dat[[trt]] * params_ana$shift_amount
-  
-  return(dd)
-}

### 2024-07-21 R/utils.R

@@ -1057,130 +1057,3 @@ rq_estimate_marginal_effects <- function(fits,
   return(list(marginal_effects = ret))
 } # End function rq_estimate_marginal_effects
 ################################################################################
-
-#' Fit models using `lmtp`
-#'
-#' @param dat A named list of tibbles containing the variables of interest. A list.
-#' @param shift_exposure Whether to shift the exposure variables or not. A logical.
-#'
-#' @returns A named list containing the results of the call to `lmtp`, for each exposure.
-#'
-#' @export
-run_mtp <- function(dat, shift_exposure) {
-  rq <- Sys.getenv("TAR_PROJECT")
-  params_dat <- params(is_hpc = Sys.getenv("is_hpc"))
-  outcome <- params_dat$variables[[rq]]$outcome
-  steps_outcome <- params_dat$steps[[rq]]$preproc_outcome
-  params_ana <- params_analyses()[[rq]]
-  
-  # Process outcome
-  dat$outcome <- myphd::extract_cohort(
-    dat = dat$outcome,
-    id_var = params_dat$variables$identifier,
-    st = 1, en = 3
-  )
-  dat$outcome <- myphd::preproc_data(
-    dat = dat$outcome,
-    outcome = outcome,
-    dic_steps = steps_outcome,
-    id_var = params_dat$variables$identifier,
-    by_var = "cohort"
-  )
-  dat$outcome <- dat$outcome |>
-    tidylog::select(-cohort)
-  
-  # Run lmtp
-  if (shift_exposure == TRUE) {
-    shift_func <- switch(params_ana$shift_type,
-                         "mul" = mtp_mul,
-                         "add" = mtp_add
-    )
-  } else {
-    shift_func <- NULL
-  }
-  list_exposures <- names(dat$exposures)
-  list_exposures <- setdiff(
-    list_exposures,
-    params_dat$variables$identifier
-  )
-  dat_merged <- tidylog::full_join(dat$covariates, dat$outcome,
-                                   by = params_dat$variables$identifier
-  ) |>
-    tidylog::mutate(cens = as.integer(!is.na(.data[[outcome]])))
-  baseline <- dat$covariates |>
-    tidylog::select(-params_dat$variables$identifier) |>
-    names()
-  
-  # Loop over each exposure
-  res <- lapply(list_exposures, function(exposure) {
-    dat_analysis <- tidylog::full_join(
-      tidylog::select(
-        dat$exposures,
-        dplyr::all_of(
-          c(
-            params_dat$variables$identifier,
-            exposure
[…]

### 2024-07-21 output/paper/render.R

@@ -1,2 +1,0 @@
-quarto::quarto_render(input = here::here("output/paper/_si.qmd"))
-quarto::quarto_render(input = here::here("output/paper/manuscript.qmd"))

### 2024-07-21 output/paper/_figures.qmd
Finished addressing comments reviwer 1.
Adapting template docx.

@@
 ```{dot}
-//| fig-width: 9
-//| fig-height: 10
+//| fig-width: 4
+//| fig-height: 4
 
@@
     node [shape=ellipse]
-    U [label='Confounders']
+    U [label="Confounders"]
 
     node [shape=box, height=0.3, width=0.3]
-    E [label='EDCs']
-    M [label='Glucocorticosteroids']
-    Y [label='HRT-SE']
+    E [label="EDCs"]
+    M [label="Glucocorticosteroids"]
+    Y [label="HRT-SE"]
     { rank = same; E M}

### 2024-07-21 output/paper/_si.qmd
Removed vertical and horizontal lines in tables for Epidemiology.

@@
     style = list(gt::cell_text(weight = "bold"))
+  ) |>
+  gt::tab_options(
+    table_body.hlines.width = 0,
+    table_body.vlines.width = 0
   )
@@
     marks = "letters"
+  ) |>
+  gt::tab_options(
+    table_body.hlines.width = 0,
+    table_body.vlines.width = 0
   )
@@
     locations = gt::cells_row_groups()
+  ) |>
+  gt::tab_options(
+    table_body.hlines.width = 0,
+    table_body.vlines.width = 0
   )

### 2024-07-21 R/proc_res.R

@@ -87,6 +87,10 @@ tidy_codebooks <- function() {
     ) |>
     gt::sub_missing(
       missing_text = ""
+    ) |>
+    gt::tab_options(
+      table_body.hlines.width = 0,
+      table_body.vlines.width = 0
     )
   
   return(ret)
@@ -241,6 +245,10 @@ tbl_desc_pop <- function(num_digits_est, num_digits_sig, by) {
     gtsummary::as_gt() |>
     gt::opt_footnote_marks(
       marks = "letters"
+    ) |>
+    gt::tab_options(
+      table_body.hlines.width = 0,
+      table_body.vlines.width = 0
     )
   
   return(desc_covars)
@@ -329,6 +337,10 @@ tbl_desc_vars <- function() {
     ) |>
     gt::opt_footnote_marks(
       marks = "letters"
+    ) |>
+    gt::tab_options(
+      table_body.hlines.width = 0,
+      table_body.vlines.width = 0
     )
   ##############################################################################
   
@@ -407,6 +419,10 @@ tbl_desc_vars <- function() {
     ) |>
     gt::opt_footnote_marks(
       marks = "letters"
+    ) |>
+    gt::tab_options(
+      table_body.hlines.width = 0,
+      table_body.vlines.width = 0
     )
   ##############################################################################
   
@@ -492,6 +508,10 @@ tbl_desc_vars <- function() {
     ) |>
     gt::opt_footnote_marks(
       marks = "letters"
+    ) |>
+    gt::tab_options(
+      table_body.hlines.width = 0,
+      table_body.vlines.width = 0
     )
   ##############################################################################
   
@@ -796,6 +816,10 @@ tidy_res_weighted_fits <- function(dat_tbl, dat_plt, sa_var) {
     ) |>
     gt::opt_footnote_marks(
       marks = "letters"
+    ) |>
+    gt::tab_options(
+      table_body.hlines.width = 0,
+      table_body.vlines.width = 0
     )
   
   return(list(
@@ -1195,6 +1219,10 @@ tidy_res_meffects <- function(df, sa_var, outcome,
     ) |>
     gt::opt_footnote_marks(
       marks = "letters"
+    ) |>
+    gt::tab_options(
+      table_body.hlines.width = 0,
+      table_body.vlines.width = 0
     )
   
   return(list(

### 2024-07-23 output/paper/title_page.qmd
Added modified title page.

@@
 ---
+
+- Type of manuscript: *Original research article*.
+- Manuscript title: *Childhood exposure to non-persistent endocrine disruptors, glucocorticosteroids, and attentional function: A cross-sectional study based on the parametric g-formula*.
+- Running head up: *Endocrine Disruptors & Childhood Attention*.
+- Corresponding author: Martine Vrijheid, Doctor Aiguader 88 (08003, Barcelona, Spain), martine.vrijheid@isglobal.org, +34 93 2147350 (Ms Iolanda Molina, personal assistant).
 
@@
 ISGlobal acknowledges support from the Spanish Ministry of Science and Innovation through the “Centro de Excelencia Severo Ochoa 2019-2023” Program (CEX2018-000806-S), and support from the Generalitat de Catalunya through the CERCA Program.
+
+## Conflicts of Interest
+
+The authors report no conflicts of interest.
+
+## Data Sharing and Code
+
+Access to HELIX data is based on approval by the HELIX Project Executive Committee and by the individual cohorts.
+Further details on the content of the data warehouse (data catalog) and procedures for external access are described on the project website (http://www.projecthelix.eu/index.php/es/data-inventory).
+
+The R code to reproduce analyses and results is available online (https://github.com/lorenzoFabbri/paper-helixSC-neuro).
 

### 2024-07-24 output/paper/_figures.qmd
Comments from MV before submission.

@@
 ```
-**Marginal contrasts on the logarithmic scale for the effect of a increase from the 10th to the 90th percentile of the {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
+**Marginal contrasts on the logarithmic scale for the effect of an increase from the 10th to the 90th percentile of the {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
 Circles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}. The size of the circles represents the *S* value of the effect estimate [@RafiGreenland:2020].
@@
 ```
-**Marginal contrasts on the logarithmic scale for the effect of a increase from the 10th to the 90th percentile of the {{< acr edcs first_use=true >}} on the glucocorticosteroids expressed in ng/ml (HELIX subcohort; 2013-2016).**
+**Marginal contrasts on the logarithmic scale for the effect of an increase from the 10th to the 90th percentile of the {{< acr edcs first_use=true >}} on the glucocorticosteroids expressed in ng/ml (HELIX subcohort; 2013-2016).**
 Circles, triangles, and squares indicate effect estimates.
@@
 ```
-**Marginal contrasts on the logarithmic scale for the effect of a increase from the 10th to the 90th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
+**Marginal contrasts on the logarithmic scale for the effect of an increase from the 10th to the 90th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
 Circles indicate effect estimates.

### 2024-07-24 output/paper/_tables.qmd
Comments from MV before submission.

@@
 ```
-**Pairwise differences between marginal contrasts on the logarithmic scale of males and females, for the effect of a increase from the 10th to the 90th percentile of {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}}, expressed in {{< acr mss >}}, and on the glucocorticosteroids, expressed in ng/ml (HELIX subcohort; 2013-2016).**
+**Pairwise differences between marginal contrasts on the logarithmic scale of males and females, for the effect of an increase from the 10th to the 90th percentile of {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}}, expressed in {{< acr mss >}}, and on the glucocorticosteroids, expressed in ng/ml (HELIX subcohort; 2013-2016).**
 :::
@@
 ```
-**Pairwise differences between marginal contrasts on the logarithmic scale of males and females, for the effect of a increase from the 10th to the 90th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
+**Pairwise differences between marginal contrasts on the logarithmic scale of males and females, for the effect of an increase from the 10th to the 90th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
 :::

### 2024-07-24 output/paper/manuscript.qmd
Added modified title page.
Comments from MV before submission.

@@
 
-Taken together, these results suggest that the negative influence of exposure to certain {{< acr edcs >}} on neurodevelopmental outcomes might be mediated, at least partially, by disruption of the {{< acr hpa >}} axis' homeostasis.
-In the present study, we thus estimated cross-sectional associations between 1) non-persistent {{< acr edcs >}} and attentional function, 2) non-persistent {{< acr edcs >}} and glucocorticosteroids, and 3) glucocorticosteroids and attentional function, using the parametric g-formula and {{< acr mcs first_use=true >}}, in children of a large network of cohorts in Europe.
+Taken together, these results suggest that disruption of the {{< acr hpa >}} axis' homeostasis may be one biological mechanism by which {{< acr edcs >}} may affect neurodevelopment.
+
+In the present study, we estimated cross-sectional associations between 1) non-persistent {{< acr edcs >}} and attentional function, 2) non-persistent {{< acr edcs >}} and glucocorticosteroids, and 3) glucocorticosteroids and attentional function in children of a multi-country cohort in Europe.
+We used the parametric g-formula, a causal inference technique, and {{< acr mcs first_use=true >}} to obtain robust effect estimates and overcome limitations of classical regression models.
 
@@
 The outcome of interest for the present study is the {{< acr hitrtse first_use=true >}} [@SunyerEsnaolaAlvarez-Pedrerol:2015], a measure of response speed consistency throughout the test.
-A high {{< acr hitrtse >}} indicates highly variable reaction times, and is considered a measure of inattentiveness.
+A high {{< acr hitrtse >}} indicates highly variable reaction times and is considered a measure of inattentiveness.
 
@@
 Sample-specific creatinine values were included in the models to adjust for possible dilution effects.
-Further, each minimal adjustment set was *augmented* with precision covariates, defined as the set of parents variable of the outcome that are not parents of the exposure.
+Further, each minimal adjustment set was *augmented* with precision covariates, defined as the set of parent variables of the outcome that are not parents of the exposure.
 Common confounders were cohort, ethnicity, sex, age, height, weight, and head circumference of the child, consumption of fish, fruit, vegetables, organic food, and fast food with a food frequency questionnaire, maternal tobacco consumption, family financial situation and affluence scale (FAS).
-Models for estimating the effects of {{< acr edcs >}} on {{< acr hitrtse >}} were further adjusted for child breastfeeding, prenatal maternal active and passive smoking, urine creatinine, child mood and rest before assessment, child neuropsychological diagnosis, marital status, season, and fasting time before assessment.
+Models for estimating the effects of {{< acr edcs >}} on {{< acr hitrtse >}} were further adjusted for child breastfeeding, prenatal maternal active and passive smoking, urine creatinine, child mood and rest before assessment, child neuropsychological diagnosis, maternal marital status, season, and fasting time before assessment.
 Models for estimating the effects of {{< acr edcs >}} on glucocorticosteroids were further adjusted for urine creatinine, season, and fasting time before assessment.
@@
 For each metabolite, we computed the fraction of values below the {{< acr loq >}} and not detected, both within each cohort and overall.
-We proceeded to impute these values using half the value of the corresponding {{< acr loq >}}, for those metabolites that had less than `r threshold_within_lod`% of missings within each cohort and `r threshold_overall_lod`% of missings overall.
+We proceeded to impute these values using half the value of the corresponding {{< acr loq >}}, for those metabolites that had less than `r threshold_within_lod`% of non-detected within each cohort and `r threshold_overall_lod`% of non-detected overall.
 Information about the {{< acr lloq >}} for the glucocorticosteroids is provided in @supptbl-lloq-mets.
-The remaining missing values were imputed using kNN from the `VIM` R package [@KowarikTempl:2016], for those metabolites that had less than `r threshold_within_nan`% of remaining missings within each cohort and `r threshold_overall_nan`% of remaining missings overall.
+Missing values were imputed using kNN from the `VIM` R package [@KowarikTempl:2016], for those metabolites that had less than `r threshold_within_nan`% of missings within each cohort and `r threshold_overall_nan`% of missings overall.
 We used `r knn_k` nearest neighbors.
@@
 Chemicals with more than 70% of observations below the {{< acr lod >}} were excluded from the present study.
-Remaining missing values were imputed similarly using kNN.
+Missing values were imputed similarly using kNN.
 Values of the chemicals were expressed in $\mu$grams per litre ($\mu$g/L).
@@
 The causal parameter of interest was thus specified as the difference in the expected counterfactual outcomes under the shifted exposure levels $\left( \mathbb{E} \left[ Y^{d_1} \right] - \mathbb{E} \left[ Y^{d_2} \right] \right)$.
-In order for this parameter to be identified, the usual causal identifiability conditions (no unmeasured confounding, positivity, and consistency) are required.
+For this parameter to be identified, the usual causal identifiability conditions (no unmeasured confounding, positivity, and consistency) are required.
 Since these conditions are likely not satisfied, we focused on the estimation of a statistical estimand that is as close as possible to the causal parameter of interest.
@@
 For each outcome, we report the results as differences between {{< acr mcs >}} with 95% {{< acr cis first_use=true >}}.
-
-The R code to reproduce analyses and results is available online (https://github.com/lorenzoFabbri/paper-helixSC-neuro).
 
@@
 Most of the {{< acr cis >}} included the null effect, though.
-Noteworthy effects were observed for the paraben MEPA (`r format_res(meffects1, "mepa", "HRT-SE")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "HRT-SE")`), oh-MiNP (`r format_res(meffects1, "ohminp", "HRT-SE")`), and MEHP (`r format_res(meffects1, "mehp", "HRT-SE")`).
-The {{< acr opp >}} DETP was negatively associated with {{< acr hitrtse >}} (`r format_res(meffects1, "detp", "HRT-SE")`).
+Noteworthy associations were observed for the paraben MEPA (`r format_res(meffects1, "mepa", "HRT-SE")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "HRT-SE")`), oh-MiNP (`r format_res(meffects1, "ohminp", "HRT-SE")`), and MEHP (`r format_res(meffects1, "mehp", "HRT-SE")`).
+The {{< acr opp >}} DETP was associated with lower {{< acr hitrtse >}} (`r format_res(meffects1, "detp", "HRT-SE")`).
 
@@
 Exceptions were BUPA, which was associated with negative {{< acr mcs >}} for all three outcomes, and MiBP, which was associated with a negative {{< acr mc >}} for total cortisone production only.
-The majority of the effects for the phenols and phthalate metabolites included the null.
+Most of the effects for the phenols and phthalate metabolites included the null.
 The phenol BPA showed the largest {{< acr mcs >}} across all three outcomes (cortisone production, `r format_res(meffects2, "bpa", "cortisone production")`; cortisol production, `r format_res(meffects2, "bpa", "cortisol production")`; corticosterone production, `r format_res(meffects2, "bpa", "corticosterone production")`).
@@
 
-In this study, consisting of 1,297 children from 6 European birth cohorts, we observed that short-term childhood exposure to certain non-persistent {{< acr edcs >}} was associated with attentional function (MEPA, MEHP, oh-MiNP, and oxo-MiNP), and with total production of cortisol, cortisone, and corticosterone (DEP, DMP, DMTP, BPA, ETPA, MEPA, MEHP, oh-MiNP, and oxo-MiNP).
-Increased production of these glucocorticosteroids did not seem to affect attentional function.
+In this study, consisting of 1,297 children from 6 European birth cohorts, we observed that short-term childhood exposure to certain non-persistent {{< acr edcs >}} was associated with lower attentional function (MEPA, MEHP, oh-MiNP, and oxo-MiNP), and with an increase in total production of cortisol, cortisone, and corticosterone (DEP, DMP, DMTP, BPA, ETPA, MEPA, MEHP, oh-MiNP, and oxo-MiNP).
 Some of these associations differed for females and males.
+Increased production of these glucocorticosteroids was not associated with attentional function in the study population overall, but showed associations with lower attentional function in males and higher attentional function in females.
 
@@
 There is also preliminary evidence of a possible association between exposure to certain {{< acr opps >}} and {{< acr adhd >}} in children [@BouchardBellingerWright:2010; @YuDuChiou:2016].
-In our study we found a negative association between DETP and attentional function, indicating lower values of {{< acr hitrtse >}} with higher levels of DETP, with concordant results in the sex-stratified analysis.
+In our study we found higher levels of DETP to be associated with better attentional function, with concordant results in the sex-stratified analysis.
 While we cannot exclude a protective effect of DETP in relation to attentional function, we note that in our study DETP has a higher percentage of missing values compared to the other {{< acr opps >}} and that other imputation strategies might lead to different associations.
 
-Preliminary evidence is also available for several phthalate metabolites in relation to cognitive development in childhood.
+Previous evidence is also available for several phthalate metabolites in relation to cognitive development in childhood.
 Higher levels of di(2-ethylhexyl) phthalate metabolites (including MEHP, MEHHP, and MEOHP) were associated with lower intelligence scores in children aged 2 to 12 years [@HuangChenSu:2015], lower scores of {{< acr iq >}} and verbal intelligence, more omission errors (a measure of inattention), and higher scores of response time variability (a measure of sustained attention) in 6-year old Korean children [@KimHongShin:2017], poorer fine motor skills in preadolescent boys [@BalalianWhyattLiu:2019], and lower intelligence scores in 7-year old children [@VilmandBeckBilenberg:2023].
 Further associations were found for higher levels of MEOHP with lower scores of {{< acr iq >}} [@HuangChenSu:2015] and verbal intelligence in Taiwanese children aged 6 to 12 years [@HuangTsaiChen:2017], and for higher levels of dibutyl phthalate metabolites (MnBP and MiBP) with impaired verbal intelligence [@HuangTsaiChen:2017].
-Few studies have looked into different classes of non-persistent {{< acr edcs >}}.
+Few studies have investigated different classes of non-persistent {{< acr edcs >}}.
 Shoaff et al.
@@
 Except for working memory, there was no evidence of an association between BPA and cognitive abilities in Spanish boys aged 9 to 11 years [@Rodriguez-CarrilloMustielesPerez-Lobato:2019].
-We did not observe an association between BPA and attention function in the present study, but this study is the first to suggest that childhood exposure to MEPA may be associated with attentional function.
+We did not observe an association between BPA and attention function in the present study, but this study is the first to suggest that childhood exposure to MEPA may be associated with lower attentional function.
 
@@
 We are also not aware of prior epidemiological studies specifically investigating the effects of elevated levels of glucocorticosteroids in relation to attentional function, although there is evidence that under- or over-production of glucocorticosteroids interfere with the normal development of the brain, possibly in a sex-specific manner [@LupienMcEwenGunnar:2009].
+Our findings support evidence for such sex-specific interference.
 
-Our findings should be interpreted in light of the following strengths and limitations.
-Strengths include its relative large sample size and its inclusion of multiple classes of non-persistent {{< acr edcs >}}.
+Our findings should be interpreted considering the following strengths and limitations.
+Strengths include its relatively large sample size and its inclusion of multiple classes of non-persistent {{< acr edcs >}}.
 Further, this study used pooled urine samples for chemical assessment to reduce measurement error, since it is known that these specific {{< acr edcs >}} have very short half-lives [@PerrierGiorgis-AllemandSlama:2016; @CasasBasaganaSakhi:2018].
@@
 Prior studies have investigated associations between short-term effects of air pollution and attentional function, with null results in adolescents but higher response times in adults [@GignacBarrera-GomezPersavento:2021; @GignacRighiToran:2022].
-Finally, longitudinal studies are necessary for a formal causal mediation analysis.
+Finally, longitudinal studies are necessary for a formal causal mediation analysis, which is why we did not attempt such a analysis even for the sex-stratified case, where glucocorticosteroids were associated with {{< acr hitrtse >}}.
 Although we included a wide range of confounders, there is the possibility, as with other observational studies, of residual confounding, which might lead to a bias away from the null.
@@
 There is further the possibility of misspecification of the outcome model, although we included a spline of the exposure to relax some of the linearity assumptions.
-Finally, we acknowledge that the present study does not take into account the possible mixture effect of these chemicals.
+Finally, we acknowledge that the present study does not consider the possible mixture effect of these chemicals.
 While methods for causal mixture analysis are available [@KeilBuckleyO:2020], we also emphasize that using these tools without thoughtful consideration may lead to misleading results [@WebsterWeisskopf:2020].

### 2024-07-24 output/paper/_si.qmd
Rebuilt manuscript.

--- 2024-07-19 output/paper/_si.qmd
@@
     style = list(gt::cell_text(weight = "bold"))
+  ) |>
+  gt::tab_style(
+    style = gt::cell_borders(
+      weight = gt::px(0)
+    ),
+    locations = list(
+      gt::cells_body(),
+      gt::cells_stub(),
+      gt::cells_row_groups()
+    )
   )
@@
     marks = "letters"
+  ) |>
+  gt::tab_style(
+    style = gt::cell_borders(
+      weight = gt::px(0)
+    ),
+    locations = list(
+      gt::cells_body(),
+      gt::cells_stub(),
+      gt::cells_row_groups()
+    )
   )
@@
     locations = gt::cells_row_groups()
+  ) |>
+  gt::tab_style(
+    style = gt::cell_borders(
+      weight = gt::px(0)
+    ),
+    locations = list(
+      gt::cells_body(),
+      gt::cells_stub(),
+      gt::cells_row_groups()
+    )
   )

### 2024-07-24 output/paper/_tables.qmd
Rebuilt manuscript.

@@
   ) |>
-  gt::tab_options(
-    table_body.hlines.width = 0,
-    table_body.vlines.width = 0
+  gt::tab_style(
+    style = gt::cell_borders(
+      weight = gt::px(0)
+    ),
+    locations = list(
+      gt::cells_body(),
+      gt::cells_stub(),
+      gt::cells_row_groups()
+    )
   )

### 2024-07-24 R/proc_res.R

@@ -88,9 +88,15 @@ tidy_codebooks <- function() {
     gt::sub_missing(
       missing_text = ""
     ) |>
-    gt::tab_options(
-      table_body.hlines.width = 0,
-      table_body.vlines.width = 0
+    gt::tab_style(
+      style = gt::cell_borders(
+        weight = gt::px(0)
+      ),
+      locations = list(
+        gt::cells_body(),
+        gt::cells_stub(),
+        gt::cells_row_groups()
+      )
     )
   
   return(ret)
@@ -246,9 +252,15 @@ tbl_desc_pop <- function(num_digits_est, num_digits_sig, by) {
     gt::opt_footnote_marks(
       marks = "letters"
     ) |>
-    gt::tab_options(
-      table_body.hlines.width = 0,
-      table_body.vlines.width = 0
+    gt::tab_style(
+      style = gt::cell_borders(
+        weight = gt::px(0)
+      ),
+      locations = list(
+        gt::cells_body(),
+        gt::cells_stub(),
+        gt::cells_row_groups()
+      )
     )
   
   return(desc_covars)
@@ -338,9 +350,15 @@ tbl_desc_vars <- function() {
     gt::opt_footnote_marks(
       marks = "letters"
     ) |>
-    gt::tab_options(
-      table_body.hlines.width = 0,
-      table_body.vlines.width = 0
+    gt::tab_style(
+      style = gt::cell_borders(
+        weight = gt::px(0)
+      ),
+      locations = list(
+        gt::cells_body(),
+        gt::cells_stub(),
+        gt::cells_row_groups()
+      )
     )
   ##############################################################################
   
@@ -420,9 +438,15 @@ tbl_desc_vars <- function() {
     gt::opt_footnote_marks(
       marks = "letters"
     ) |>
-    gt::tab_options(
-      table_body.hlines.width = 0,
-      table_body.vlines.width = 0
+    gt::tab_style(
+      style = gt::cell_borders(
+        weight = gt::px(0)
+      ),
+      locations = list(
+        gt::cells_body(),
+        gt::cells_stub(),
+        gt::cells_row_groups()
+      )
     )
   ##############################################################################
   
@@ -509,9 +533,15 @@ tbl_desc_vars <- function() {
     gt::opt_footnote_marks(
       marks = "letters"
     ) |>
-    gt::tab_options(
-      table_body.hlines.width = 0,
-      table_body.vlines.width = 0
+    gt::tab_style(
[…]

### 2024-08-12 output/paper/manuscript.qmd
Preparing for submission to Environmental Research.

@@
     reference-doc: custom-reference-doc.docx
+    fig-dpi: 420
 
@@
 bibliography: bibliography.bib
-csl: epidemiology.csl
+csl: environmental-research.csl
 
@@
   insert_loa: false
+  include_unused: true
+  loa_title: ""
   sorting: alphabetical
@@
 
+{{< pagebreak >}}
+
+# List of Acronyms {.unnumbered}
+
+\printacronyms
+
+{{< pagebreak >}}
+
 # Abstract {.unnumbered}
@@
 
-# Methods {#sec-methods}
+# Material and methods {#sec-methods}
 
@@
 
+## Conclusion
+
 In conclusion, in a study of 1,297 children from 6 European birth cohorts, we observed that (i) exposure to certain non-persistent {{< acr edcs >}} is associated with higher values of {{< acr hitrtse >}} and might disrupt the {{< acr hpa >}} axis, and (ii) certain glucocorticosteroids are associated with {{< acr hitrtse >}} in a sex-specific manner.
+
+# Acknowledgments
+
+The authors acknowledge that Born in Bradford is only possible because of the enthusiasm and commitment of the children and parents in Born in Bradford.
+We are grateful to all participants, health professionals and researchers who have made Born in Bradford happen.
+
+The authors thank the EDEN mother-child cohort study group, whose members are I.
+Annesi-Maesano, J.Y.
+Bernard, M.A.
+Charles, P.
+Dargent-Molina, B.
+de Lauzon-Guillain, P.
+Ducimetière, M.
+de Agostini, B.
+Foliguet, A.
+Forhan, X.
+Fritel, A.
+Germa, V.
+Goua, R.
+Hankard, B.
+Heude, M.
+Kaminski, B.
+Larroquey, N.
+Lelong, J.
+Lepeule, G.
+Magnin, L.
+Marchand, C.
+Nabet, F Pierre, R.
+Slama, M.J.
+Saurel-Cubizolles, M.
+Schweitzer, and O.
+Thiebaugeorges.
 

### 2024-08-12 output/paper/title_page.qmd
Preparing for submission to Environmental Research.

@@
 format:
-  html: default
+  docx: default
+
+filters:
+  - authors-block
 
@@
 The R code to reproduce analyses and results is available online (https://github.com/lorenzoFabbri/paper-helixSC-neuro).
-
-## Acknowledgments
-
-The authors acknowledge that Born in Bradford is only possible because of the enthusiasm and commitment of the children and parents in Born in Bradford.
-We are grateful to all participants, health professionals and researchers who have made Born in Bradford happen.
-
-The authors thank the EDEN mother-child cohort study group, whose members are I.
-Annesi-Maesano, J.Y.
-Bernard, M.A.
-Charles, P.
-Dargent-Molina, B.
-de Lauzon-Guillain, P.
-Ducimetière, M.
-de Agostini, B.
-Foliguet, A.
-Forhan, X.
-Fritel, A.
-Germa, V.
-Goua, R.
-Hankard, B.
-Heude, M.
-Kaminski, B.
-Larroquey, N.
-Lelong, J.
-Lepeule, G.
-Magnin, L.
-Marchand, C.
-Nabet, F Pierre, R.
-Slama, M.J.
-Saurel-Cubizolles, M.
-Schweitzer, and O.
-Thiebaugeorges.

### 2024-08-12 output/paper/_figures.qmd
Combining forest plots of RQ3 and RQ3SA.

@@
 } # End function forest_plot
+
+# Function to align forest plot with table numerical results for RQ and RQSA
+forest_plot_combined <- function(obj, r = 4, is_cortisol = FALSE) {
+  edcs <- myphd::edcs_information() |>
+    dplyr::filter(!chem_id %in% c("dedtp", "dmdtp")) |>
+    dplyr::select(chem_id, short_name, class) |>
+    dplyr::bind_rows(
+      tibble::tibble(
+        chem_id = c("Exposure"),
+        short_name = c("Exposure"),
+        class = c(NULL)
+      )
+    )
+
+  # From table of numerical results to ggplot object
+  tbl_num_res <- obj$numerical_results |>
+    dplyr::bind_rows(
+      tibble::tibble(
+        variable = "Exposure",
+        val = "Marginal contrast (95% CI)",
+        estimate = 100,
+        outcome = "",
+        class = "AAA",
+        modifier = c(NA, NA, NA)
+      ),
+      tbl_num_res
+    ) |>
+    dplyr::group_by(class) |>
+    dplyr::arrange(
+      variable,
+      .by_group = TRUE
+    ) |>
+    dplyr::ungroup() |>
+    dplyr::mutate(
+      variable = factor(variable, levels = unique(variable)),
+      variable = forcats::fct_rev(variable)
+    ) |>
+    tidylog::select(-estimate) |>
+    ggplot2::ggplot(
+      ggplot2::aes(
+        y = variable,
+        shape = modifier
+      )
+    ) +
+    ggplot2::geom_text(
+      ggplot2::aes(
+        x = 0,
+        label = variable
+      ),
+      hjust = 0,
+      fontface = "bold"
+    ) +
+    ggplot2::geom_text(
+      ggplot2::aes(
+        x = 1,
+        label = val
+      ),
+      hjust = 0,
+      position = ggstance::position_dodgev(height = 0.9)
+    ) +
+    ggplot2::theme_void() +
+    ggplot2::theme(
+      strip.background = ggplot2::element_blank(),
+      strip.text = ggplot2::element_blank(),
+      legend.position = "bottom",
+      legend.box = "vertical",
+      legend.margin = ggplot2::margin()
+    ) +
+    ggplot2::coord_cartesian(
+      xlim = c(0, 4)
+    )
+
+  # Assemble elements
+  ret <- patchwork::wrap_plots(
+    tbl_num_res, obj$plot
+  ) +
+    patchwork::plot_layout(
+      design = c(
+        patchwork::area(t = 0, l = 1, b = 3, r = r),
+        patchwork::area(t = 1, l = r, b = 3, r = r + 4)
+      )
+    )
+
+  return(ret)
+} # End function forest_plot_combined
 ```
@@
   dplyr::matches("marginal_comparisons_(\\d)$"),
+  store = store
+)
+targets::tar_load(
+  dplyr::matches("marginal_comparisons_3SA$"),
   store = store
@@
 
+```{r}
+marginal_comparisons_3$modifier <- "overall"
+obj3 <- tidy_res_meffects(
+  df = dplyr::bind_rows(marginal_comparisons_3, marginal_comparisons_3SA),
+  sa_var = "e3_sex",
+  outcome = c("hitrtse"),
+  which_res = "comparisons",
+  num_digits_est = num_digits_est,
+  num_digits_sig = num_digits_sig,
+  combined = c("overall", "females", "males")
+)
+```
+
 ::: {#fig-marginal-3}
@@
 
-obj3 <- tidy_res_meffects(
-  df = marginal_comparisons_3,
-  sa_var = NULL,
-  outcome = c("hitrtse"),
-  which_res = "comparisons",
-  num_digits_est = num_digits_est,
-  num_digits_sig = num_digits_sig
-)
-forest_plot(obj = obj3, r = 5, is_cortisol = TRUE)
+forest_plot_combined(obj = obj3, r = 5, is_cortisol = TRUE)
 ```
-**Marginal contrasts on the logarithmic scale for the effect of an increase from the 10th to the 90th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
+**Marginal contrasts on the logarithmic scale for the effect of an increase from the 10th to the 90th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}}, overall and by sex (HELIX subcohort; 2013-2016).**
 Circles indicate effect estimates.

### 2024-08-12 output/paper/_si.qmd
Combining forest plots of RQ3 and RQ3SA.

@@
 edcs_info <- myphd::edcs_information() |>
-  tidylog::select(-c(chem_id, smiles, ctd_id, exposome_explorer_id)) |>
+  tidylog::select(-c(chem_id, smiles, ctd_id, exposome_explorer_id, molar_mass)) |>
   tidylog::mutate(
@@
 
-::: {#supptbl-marginal-3sa}
-```{r marginal-3sa-tbl}
-m3sa <- tidy_res_meffects(
-  df = marginal_comparisons_3SA,
-  sa_var = "e3_sex",
-  outcome = NULL,
-  which_res = "comparisons",
-  num_digits_est = num_digits_est,
-  num_digits_sig = num_digits_sig
-)
-
-m3sa$table
-```
-**Marginal contrasts on the logarithmic scale for effect modification by sex of a increase from the 10th to the 90th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
-:::
-
 # Supplementary figures

### 2024-08-12 R/proc_res.R

@@ -1029,7 +1029,8 @@ load_res_meffects <- function(path_store, rq, sa_var, which_res) {
 #' @export
 tidy_res_meffects <- function(df, sa_var, outcome,
                               which_res,
-                              num_digits_est, num_digits_sig) {
+                              num_digits_est, num_digits_sig,
+                              combined = NULL) {
   if (length(outcome) > 0) {
     df <- df |>
       tidylog::filter(
@@ -1156,6 +1157,27 @@ tidy_res_meffects <- function(df, sa_var, outcome,
         )
     }
     plot <- plot +
+      ggplot2::theme(
+        legend.position = "bottom",
+        legend.box = "vertical",
+        legend.margin = ggplot2::margin(),
+        axis.title.y = ggplot2::element_blank(),
+        axis.text.y = ggplot2::element_blank(),
+        axis.ticks.y = ggplot2::element_blank(),
+        axis.line.y = ggplot2::element_blank(),
+        axis.line.x = ggplot2::element_line(linewidth = 0.6),
+        axis.ticks.length = ggplot2::unit(0.3, "cm")
+      )
+  } else if (!is.null(sa_var) & !is.null(combined)) {
+    plot <- plot +
+      ggplot2::coord_cartesian(
+        ylim = c(1, (nrow(df) / length(combined)) + 1)
+      ) +
+      ggplot2::annotate(
+        "text",
+        x = 0, y = (nrow(df) / length(combined)) + 1,
+        label = ""
+      ) +
       ggplot2::theme(
         legend.position = "bottom",
         legend.box = "vertical",

### 2024-08-12 output/paper/_figures.qmd
Combining forest plots of RQ3 and RQ3SA.

@@
       ),
-      tbl_num_res
+      obj$numerical_results
     ) |>

### 2024-08-12 output/paper/_si.qmd
Combining forest plots of RQ3 and RQ3SA.

@@
 :::
-
-::: {#suppfig-marginal-3sa}
-```{r marginal-3sa}
-#| fig-width: 6
-#| fig-height: 5
-
-m3sa$plot
-```
-**Marginal contrasts on the logarithmic scale for effect modification by sex of a increase from the 10th to the 90th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
-Circles and triangles indicate effect estimates.
-Solid lines indicate the $95\%$ {{< acr ci >}}.
-The size of the circles represents the *S* value of the effect estimate [@RafiGreenland:2020].
-Abbreviations: cortisone production (cortisone prod.); cortisol production (cortisol prod.); corticost.
-prod.
-(corticosterone production).
-:::
